### 2022-08-14 Run_SS3_Krill_481.Rmd
Add files via upload

---
title: "Run_SS3_Krill_481"
author: "Mauricio Mardones I."
date: '2022-06-15'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# 
# install.packages("caTools")
# library("caTools")
# 
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
```

```{r message=FALSE, warning=FALSE}
dir1<-here("~/DOCAS/SA_Krill/s1")
dir2<-here("~/DOCAS/SA_Krill/s2")
dir3<-here("~/DOCAS/SA_Krill/s3_longer")
dir4<-here("~/DOCAS/SA_Krill/s3")
```

```{r}
#shell(cmd="ss") # run SS para windows
# aqui debo cambiar el setwd() para correr el modelo

r4ss::run_SS_models(dirvec = dir4, model = './ss_osx', 
                    skipfinished = FALSE) 
```

```{r}
dir4<-here("s3")
base.model <- SS_output(dir=dir4,covar=F,forecast=F)
```

Saco los outputs en html

```{r eval=FALSE}
SS_plots(base.model, uncertainty=T,datplot = F, png=T, aalresids = F,btarg=0.75, minbthresh=0.25, forecast=F)
```

Data disponible para este escenario. Espinel es la serie mas consistente del conjunto de datos.

```{r}
SSplotData(base.model, subplot = 2, 
           fleetnames = c("Fishery", "Survey"),
           fleetcol = c(4,6))
```

Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso

```{r}
SSplotBiology(base.model, subplots =1, labels = c("Length (cm)", "Age (yr)", "Maturity", "Mean weight (kg) in last year",
    "Spawning output", "Length (cm, beginning of the year)", "Natural mortality",
    "Female weight (kg)", "Female length (cm)", "Fecundity", "Default fecundity label",
    "Year", "Hermaphroditism transition rate", "Fraction females by age at equilibrium"),
 )
```

aporte de las cohortes por año para las capturas.
```{r}
SSplotCohortCatch(base.model, subplots = 1)
```

\quad

AJuste de tallas por flota
```{r}
SSplotComps(base.model, subplots = 1)
```

Otros plots
```{r}
SSplotDynamicB0(base.model, uncertainty = F)
#SSplotSPR(base.model3)
SSplotPars(base.model)

```

Salida de las biomasas con las dos flotas

```{r}
SSplotTimeseries(base.model, subplot = 1)
```

```{r}
#########################################################################
#COMPARACION DE MODELOS con distinto desembarque
#########################################################################

#PLOT labels, name of each model run
legend.labels <- c('SSdir1','SSdir2','SSdir3')

#read in all model runs
#note if cover=T you need a hessian; if covar=F you do not need a hessian
biglist <- SSgetoutput(keyvec = NULL, dirvec = c(SSdir1,SSdir2,SSdir3),	getcovar = F)

#create summary of model runs from list above
summaryoutput <- SSsummarize(biglist)

SSplotComparisons(summaryoutput, subplots = 1:20, plot = TRUE,
                  print = T, endyrvec = "default", indexfleets = NULL, indexUncertainty = FALSE,
                  indexQlabel = TRUE, indexQdigits = 4, indexSEvec = "default",
                  indexPlotEach = FALSE, labels = c("Year", "Spawning biomass (t)",
                                                    "Relative spawning biomass", "Age-0 recruits (1,000s)",
                                                    "Recruitment deviations", "Index", "Log index", "1 - SPR", "Density",
                                                    "Management target", "Minimum stock size threshold", "Spawning output",
                                                    "Harvest rate"), col = NULL, shadecol = NULL, pch = NULL,
                  lty = 1, lwd = 2, spacepoints = 10, staggerpoints = 1,
                  initpoint = 0, tickEndYr = TRUE, shadeForecast = TRUE,
                  xlim = "default", ylimAdj = 1, xaxs = "r", yaxs = "r",
                  type = "o", uncertainty = TRUE, shadealpha = 0.1, legend = TRUE,
                  legendlabels = "default", legendloc = "topright",
                  legendorder = "default", legendncol = 1, sprtarg = NULL,
                  btarg = NULL, minbthresh = NULL, pwidth = 6.5, pheight = 5,
                  punits = "in", res = 300, ptsize = 10, cex.main = 1,
                  plotdir = "C:\\Users\\mauricio.mardones\\Documents\\IFOP\\Loco_Assessment_AMERB\\SA_Loco", filenameprefix = "", densitynames = c("SSB_Virgin",
                                                                                                                     "R0"), densityxlabs = "default", densityscalex = 1,
                  densityscaley = 1, densityadjust = 1, densitysymbols = TRUE,
                  densitytails = TRUE, densitymiddle = FALSE, densitylwd = 1,
                  fix0 = TRUE, new = TRUE, add = FALSE, par = list(mar = c(5, 4, 1,
                                                                           1) + 0.1), verbose = TRUE, mcmcVec = FALSE,
                  show_equilibrium = TRUE)

SStableComparisons(summaryoutput,
                   likenames = c("TOTAL", "Survey", "Length_comp", "Age_comp", "priors",
                                 "Size_at_age"), names = c("Recr_Virgin", "R0", "steep", "NatM",
                                                           "L_at_Amax", "VonBert_K", "SSB_Virg", "Bratio_2017", "SPRratio_2016"),
                   digits = NULL, modelnames = "default", csv = FALSE,
                   csvdir = "C:\\Users\\mauricio.mardones\\Documents\\IFOP\\Loco_Assessment_AMERB\\SA_Loco",
                   csvfile = "parameter_comparison_table.csv", verbose = TRUE,
                   mcmc = FALSE)
```

# Analisis retrospectivo

```{r}

#do retrospective model runs
SS_doRetro('C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\5', '', newsubdir = "retrospectives",
           subdirstart = "retro", years = 0:-5, overwrite = TRUE, exefile = "ss",
           extras = "-nox -nohess", intern = FALSE, CallType = "system",
           RemoveBlocks = FALSE)
retroModels <- SSgetoutput(dirvec=file.path('C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\5', "retrospectives",paste("retro",0:-5,sep="")))
retroSummary <- SSsummarize(retroModels)
endyrvec <- retroSummary$endyrs + 0:-5
SSplotComparisons(retroSummary, endyrvec=endyrvec, legendlabels=paste("Data",0:-5,"years"),
                  plotdir='C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\',plot=TRUE,print=T)
TableCompare <- SStableComparisons(retroSummary,likenames=like, names=names, 
                                   modelnames=c('B','-1','-2','-3','-4','-5'), csv=TRUE, csvfile="RetroRuns.csv",verbose=FALSE)
##############################################################################################

```

### 2024-03-26 Run_SS3_Krill_481.Rmd
innit commit

@@
 ---
-title: "Run_SS3_Krill_481"
-author: "Mauricio Mardones I."
-date: '2022-06-15'
-output: pdf_document
+title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
+subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
+author:
+  - Mardones Mauricio^[Instituo de Fomento Pesquero, mardones.mauricio@gmail.com]
+  - César Cárdenas^[Instituto Antártico Chileno]
+date:  "`r format(Sys.time(), '%d %B, %Y')`"
+bibliography: SA_krill.bib
+csl: apa.csl
+link-citations: yes
+linkcolor: blue
+always_allow_html: yes
+output:
+  pdf_document
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{=tex}
+\fontsize{12}{16}
+\selectfont{}
 ```
+\newpage
+
+```{=latex}
+\setcounter{tocdepth}{3}
+\tableofcontents
+```
+\newpage
+
+```{r setup1, echo=FALSE}
+rm(list = ls())
+knitr::opts_chunk$set(echo = FALSE,
+                      eval=FALSE,
+                      message = FALSE,
+                      warning = FALSE,
+                      fig.align = 'center',
+                      fig.pos = "H",
+                      dev = 'jpeg',
+                      dpi = 300,
+                      tidy.opts=list(width.cutoff=80), 
+                      tidy=TRUE)
+#XQuartz is a mess, put this in your onload to default to cairo instead
+options(bitmapType = "cairo") 
+# (https://github.com/tidyverse/ggplot2/issues/2655)
+# Lo mapas se hacen mas rapido
+```
+
+# ABSTRACT
+
+The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystemic framework, highlights the importance of incorporating spatial heterogeneity.
+The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill.
+By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics.
+This nuanced approach allows for a more refined understanding of diferente zones shape krill abundance, fishing mortalyty and recruitment patterns across the Antarctic Peninsula.
+Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
+
+*Keywords: Krill populations, dynamic population, stock assessment, SS3, spatial implicit, management, CCAMLR.*
+
+# INTRODUCTION
+
+# METHODOLOGY
+
+## Conceptual Model
+
+## Phases in model construction
+
+The process of modeling and statistical analysis of a database can be structured according to the following guidelines:
+
+- Contextualization of the problem. Definition of objectives and variables.
+- Experiment design and data collection.
+- Recording and preliminary processing of available information.
+- Graphical inspection and trend identification.
+- Consideration of distributional and relational hypotheses. Proposal of modeling.
+- Model adjustment. Comparison and selection of the best model.
+- Diagnosis and validation of the adjusted model.
+- Assessment of the predictive capacity of the model and prediction.
+- Interpretation and conclusions.
+
+## Scenarios
+
+In Table 1 we have ten scenarios to test different option in modeling about main consideration in assessment of krill population.
+
+| Scenario | Description                                      |
+|:---------:|:-------------------------------------------------|
+|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
+|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
+|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata and Predator and Env data |
+|    s3     | Same "s2" Without S-R relation         |
+|    s4     | Same "s2" Ricker S-R relation estimated          |
+|    s5     | Same "s2" BH S-R relation weak (0.9 steepness)     |
+|    s6     | Same "s2" BH S-R relation strong (0.6 steepness)     |
+|    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+
+## Statistical model (SS3)
+
+The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the most updated version (V.3.30.21).
+SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
+SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
+The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
+The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].
+
+## Model setup
 
@@
 # devtools::install_github("r4ss/r4ss", ref="development")
-# 
 # install.packages("caTools")
 # library("caTools")
-# 
 # install.packages("r4ss")
@@
 ```{r message=FALSE, warning=FALSE}
-dir1<-here("~/DOCAS/SA_Krill/s1")
-dir2<-here("~/DOCAS/SA_Krill/s2")
-dir3<-here("~/DOCAS/SA_Krill/s3_longer")
-dir4<-here("~/DOCAS/SA_Krill/s3")
+dir01 <- here("s01") # agreggate data (no spatial diferences)
+dir1<-here("s1") # Data strata flishery
+dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
+dir3<-here("s3") # whitout S-R
+dir4<-here("s4") # 
+dir5<-here("s5") # whitout S-R
+dir6<-here("s6") # whitout S-R
+dir7<-here("s7") # whitout S-R
 ```
 
-```{r}
-#shell(cmd="ss") # run SS para windows
-# aqui debo cambiar el setwd() para correr el modelo
+```{r eval=FALSE, message=F, include=FALSE}
+# Lista de directorios para correro tordos juntos
+directorios <- c("s01", 
+                 "s1", 
+                 "s2",
+                 "s3",
+                 "s4",
+                 "s5",
+                 "s6",
+                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
 
-r4ss::run_SS_models(dirvec = dir4, model = './ss_osx', 
-                    skipfinished = FALSE) 
+# Bucle para ejecutar el código en cada directorio
+for (dir in directorios) {
+  r4ss::run(
+    dir = dir,
+    exe = "ss_osx",
+    skipfinished = FALSE,
+    show_in_console = TRUE
+  )
+}
+
+#OR with rby separate
+
+#shell(cmd="ss") # run SS to windows
+# or 
+#system('./ss_osx') # to mac
+
+# r4ss::run(
+#   dir = dir1,
+#   exe = "ss_osx",
+#   skipfinished = FALSE, # TRUE will skip running if Report.sso present
+#   show_in_console = TRUE # change to true to watch the output go past
+# )
 ```
 
-```{r}
-dir4<-here("s3")
-base.model <- SS_output(dir=dir4,covar=F,forecast=F)
+```{r message=F, include=FALSE}
+#s01
+base.model01 <- SS_output(dir=dir01,
+                        covar=T,
+                        forecast=T)
+#s1
+base.model1 <- SS_output(dir=dir1,
+                        covar=T,
+                        forecast=T)
+#s2
+base.model2 <- SS_output(dir=dir2,
+                        covar=T,
+                        forecast=T)
+#s3
+base.model3 <- SS_output(dir=dir3,
+                        covar=T,
+                        forecast=T)
+#s4
+base.model4 <- SS_output(dir=dir4,
+                        covar=T,
+                        forecast=T)
+
+#s7
+base.model7 <- SS_output(dir=dir7,
+                        covar=T,
+                        forecast=T)
+
 ```
@@
 
-```{r eval=FALSE}
-SS_plots(base.model, uncertainty=T,datplot = F, png=T, aalresids = F,btarg=0.75, minbthresh=0.25, forecast=F)
+```{r eval=FALSE,  message=F, include=FALSE}
+SS_plots(base.model7, 
+         uncertainty = TRUE,
+         datplot = T, 
+         png=T, 
+         aalresids = F,
+         btarg=0.75, 
+         minbthresh=0.2, 
+         forecast=T)
 ```
@@
 ```{r}
-SSplotData(base.model, subplot = 2, 
-           fleetnames = c("Fishery", "Survey"),
-           fleetcol = c(4,6))
+SSplotData(base.model, subplots = 2)
+
 ```
@@
 ```{r}
-SSplotBiology(base.model, subplots =1, labels = c("Length (cm)", "Age (yr)", "Maturity", "Mean weight (kg) in last year",
-    "Spawning output", "Length (cm, beginning of the year)", "Natural mortality",
-    "Female weight (kg)", "Female length (cm)", "Fecundity", "Default fecundity label",
-    "Year", "Hermaphroditism transition rate", "Fraction females by age at equilibrium"),
- )
+SSplotSelex(base.model)
 ```
@@
 ```{r}
-SSplotTimeseries(base.model, subplot = 1)
+SSplotTimeseries(base.model7, subplot = 1)
 ```
@@
 ```{r}
-#########################################################################
-#COMPARACION DE MODELOS con distinto desembarque
-#########################################################################
+SSplotTimeseries(base.model,
+  subplot=1,
+  add = FALSE,
+  forecastplot = FALSE,
+  uncertainty = TRUE,
+  bioscale = 1,
+  minyr = -Inf,
+  maxyr = Inf,
+  plot = TRUE,
+  print = FALSE,
+  plotdir = "default",
+  verbose = TRUE,
+  btarg = "default",
+  minbthresh = "default",
+  xlab = "Year",
+  labels = NULL,
+  pwidth = 6.5,
+  pheight = 5,
+  punits = "in",
+  res = 300,
+  ptsize = 10,
+  cex.main = 1,
+  mainTitle = FALSE,
+  mar = NULL
+)
+```
 
+# Diagnosis Base Model (Jabba)
+
+step to do a good practice in diagnosis is;
+- 1. Convergence. Final convergence criteria is 1.0e-04
+- 2. Residual 
+- 3. Restrospective analysis
+- 4. Likelihood profile
+
+## Residual
+
+```{r}
+SSplotRunstest(base.model4, 
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model4, 
+               subplots = "len",
+               add=T)
+```
+
+```{r}
+SSplotJABBAres(base.model,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model,
+               subplots = "len",
+               add=T)
+```
+
+### Predator fleet with RW
+
+Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.
+
+Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter.
+In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process.
+Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.
+
+Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time.
+However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.
+
+# Retrospectivo
+
+Los análisis retrospectivo, dan cuenta de diferencias de estimación (sub - sobreestimación) en los patrones entre modelos evaluados.
+
+```{r eval=FALSE}
+retro(dir=dir7, oldsubdir="", 
+      newsubdir="Retrospective", 
+      years= 0:-5,
+      exe="ss_osx",
+      extras = "-nox", 
+      skipfinished = F)
+```
+
+```{r}
+retroModels <- SSgetoutput(dirvec=file.path(dir01,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary <- SSsummarize(retroModels)
+```
+
+```{r}
+#save(retroSummary, retroModels, file="retro5.Rdata")
+
+retro01 <- SSplotRetro(retroSummary,
+            add=T,
+            forecast = F,
+            legend = F,
+            verbose=F)
+retro2 <- SSplotRetro(retroSummary,
+            add=T,
+            forecast = F,
+            legend = F,
+            verbose=F)
+retro1 <- SSplotRetro(retroSummary,
+            add=T,
+            forecast = F,
+            legend = F,
+            verbose=F)
+
+retro7 <- SSplotRetro(retroSummary,
+            add=T,
+            forecast = F,
+            legend = F,
+            verbose=F,
+            legendlabels="default")
+```
+## Hindcast Cross-Validation and prediction skill
+
+Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated by `r4ss:SS_doRetro()`.
+As a robust measure of prediction skill, we implemented the mean absolute scaled error (MASE).
+In brief, the MASE score scales the mean absolute.
+Regarding (A MASE score > 1 indicates that the average model forecasts are worse than a random walk.
+Conversely, a MASE score of 0.5 indicates that the model forecasts twice as
+accurately as a naïve baseline prediction; thus, the model has prediction skill.
+
+```{r}
+hci = SSplotHCxval(retroSummary, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+```
+# Kobe
+
+```{r}
+mvln = SSdeltaMVLN(base.model01,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+```
+
+```{r}
+retroModels <- SSgetoutput(dirvec=file.path(mydir, "Retrospective",paste("retro",0:-5, sep="")))
+retroSummary <- SSsummarize(retroModels)
+endyrvec <- retroSummary$endyrs + 0:-5
+save(retroSummary, retroModels, file="retro5.Rdata")
+# Step 1. Identify restrospective period
+# e.g., for end.yr.vec   <- c(2015,2014,2013,2012,2011,2010)
+start.retro <- 0    #end year of model e.g., 2015
+end.retro   <- 5    #number of years for retrospective e.g., 2014,2013,2012,2011,2010
+
+# Step 2. Identify the base directory
+dirname.base <- here("SA_Krill")
+dirname.base
+
+# Step 3. Identify the directory where a completed model run is located
+dirname.completed.model.run <- here("s9")
+dirname.completed.model.run
+
+# Step 4. Create a subdirectory for the Retrospectives
+dirname.Retrospective <- here("Ejercicios","Retrospectivo") 
+dir.create(path=dirname.Retrospective, showWarnings = TRUE, recursive = TRUE)
+setwd(dirname.Retrospective)
+getwd()
+
+# Step 5. Create a subdirectory for the Plots
+dirname.plots <- paste0(dirname.Retrospective,"/plots_retro")
+dir.create(dirname.plots)
+
+# 7. Copiar los archivos necesarios de "Reference_run" al directorio "Perfil_Verosimilitud" ----
+
+copy_SS_inputs(dir.old = dirname.completed.model.run, 
+               dir.new =  dirname.Retrospective,
+               copy_exe = TRUE,
+               verbose = FALSE)
+
+# #----------------- copy model run files ----------------------------------------
+# file.copy(paste(dirname.completed.model.run,       "starter.ss_new", sep="/"),
+#           paste(dirname.Retrospective, "starter.ss", sep="/"))
+# 
+# file.copy(paste(dirname.completed.model.run,       "control.ss_new", sep="/"),
+#           paste(dirname.Retrospective, "CONTROL.SS", sep="/"))
+# 
+# file.copy(paste(dirname.completed.model.run,       "data.ss", sep="/"),
+#           paste(dirname.Retrospective, "DATA.SS", sep="/"))	
+# 
+# file.copy(paste(dirname.completed.model.run,       "forecast.ss", sep="/"),
+#           paste(dirname.Retrospective, "forecast.ss", sep="/"))
+# 
+# file.copy(paste(dirname.completed.model.run,       "ss_win.exe", sep="/"),
+#           paste(dirname.Retrospective, "SS.exe", sep="/"))
+
+# # Required for assessments with wtatage.ss files
+# file.copy(paste(dirname.completed.model.run,       "wtatage.ss", sep="/"),
+#           paste(dirname.Retrospective, "wtatage.ss", sep="/"))
+
+#------------Make Changes to the Starter.ss file (DC Example) ------------------------------- 
+starter <- readLines(paste(dirname.Retrospective, "/starter.ss", sep=""))
+
+# 1) Starter File changes to speed up model runs
+# Run Display Detail
+#[8] "2 # run display detail (0,1,2)" 
+linen <- grep("# run display detail", starter)
+starter[linen] <- paste0( 1 , " # run display detail (0,1,2)" )
+write(starter, paste(dirname.Retrospective, "starter.ss", sep="/"))
+
+#------------ r4SS retrospective calculations-------------------------------
+
+# Step 6. Run the retrospective analyses with r4SS function "SS_doRetro"
+retro(masterdir=dirname.Retrospective,
+           oldsubdir="", 
+           newsubdir="retrospectives",
+           years=start.retro:-end.retro)
+
+# Step 7. Read "SS_doRetro" output
+retroModels <- SSgetoutput(dirvec=file.path(dirname.Retrospective, "retrospectives",paste("retro",start.retro:-end.retro,sep="")))
+
+# Step 8. Summarize "SS_doRetro" output
+retroSummary <- SSsummarize(retroModels)
+
+sspar(mfrow=c(1,2),plot.cex=0.8)
+rb = SSplotRetro(retroSummary,add=T,forecast = F,legend = F,verbose=F)
+rf = SSplotRetro(retroSummary,add=T,subplots="F", ylim=c(0,0.4),
+                 forecast = F,legendloc="topleft",legendcex = 0.8,verbose=F)
+
+sspar(mfrow = c(2, 1), plot.cex = 0.8)
+
+SSplotRetro(retroSummary, 
+            add = T, 
+            forecast = T, 
+            legend = F, 
+            verbose = F, 
+            xmin = 1970)
+SSplotRetro(retroSummary, 
+            add = T, 
+            subplots = "F", 
+            ylim = c(0, 0.06), 
+            forecast = T,
+            legendloc = "topleft", 
+            legendcex = 0.8, 
+            verbose = F, 
+            xmin = 1970)
+
+SShcbias(retroSummary,quant="SSB",verbose=F)
+SShcbias(retroSummary,quant="F",verbose=F)
+```
+
+# Verosimilitud
+
+```{r}
+# 1. Identificar el directorio donde se encuentra el modelo base ----
+dirname.model.run <- here("s10")
+
+# 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"
+
+dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
+dir.create(path=dirname.R0.profile, showWarnings = TRUE, recursive = TRUE)
+
+# 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
+
+plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
+dir.create(path=plotdir, showWarnings = TRUE, recursive = TRUE)
+
+# 4. Crear un subdirectorio llamado "simple" ----
+
+reference.dir <- paste0(dirname.R0.profile,'/simple') 
+dir.create(path=reference.dir, showWarnings = TRUE, recursive = TRUE)
+
+# 5. Copiar el resultado del modelo base completo en este directorio ----
+
+file.copy(Sys.glob(paste(dirname.model.run, "*.*", sep="/"),
+                   dirmark = FALSE),
+                    reference.dir)
+
+# 6. Leer la salida del modelo base ----
+
+Base <- SS_output(dir=reference.dir,covar=T)
+
+# 7. Copiar los archivos necesarios de "simple" al directorio "Perfil_Verosimilitud" ----
+
+copy_SS_inputs(dir.old = reference.dir, 
+               dir.new =  dirname.R0.profile,
+               copy_exe = TRUE,
+               verbose = FALSE)
+
+# 8. Leer los archivos del modelo ----
+
+inputs <- r4ss::SS_read(dir = dirname.R0.profile)
+
+# 9. Editar el archivo control la fase de estimación recdev ----
+
+inputs$ctl$recdev_phase <- 1
+
+# 10. Editar el archivo starter para leer los valores de inicio ----
+
+inputs$start$init_values_src <- 0
+
+# 11. Vector de valores para el perfil ----
+
+R0.vec <- seq(7,11,0.5)  
+Nprof.R0 <- length(R0.vec)
+
+# 12. Cambiar el nombre del archivo control en el archivo starter.ss ----
+
+inputs$start$ctlfile <- "control_modified.ss"
+
+# 13. Incluir prior_like para parámetros no estimados ----
+
+inputs$start$prior_like <- 1
+
+# 14. Escribir los modelos modificados ----
+
+r4ss::SS_write(inputs, dir = dirname.R0.profile, overwrite = TRUE)
+
+# 15. Ejecutar la función profile() ----
+
+?SS_profile()
+profile <- profile(dir="dir10", # directory
+                      model="ss_osx",
+                      oldctlfile ="control.ss",
+                      newctlfile="control_modified.ss",
+                      string="SR_LN(R0)",
+                      profilevec=R0.vec)
+
+# 16. Leer los archivos de salida ----
+# (con nombres como Report1.sso, Report2.sso, etc.)
+prof.R0.models <- SSgetoutput(dirvec=dirname.R0.profile, 
+                              keyvec=1:Nprof.R0, 
+                              getcovar = FALSE)
+
+# 17. Resumir las salidas con la función SSsummarize()  ----
+prof.R0.summary <- SSsummarize(prof.R0.models)
+
+# 18. Identificar los componentes de Verosimilitud ----
+mainlike_components         <- c('TOTAL',"Survey", 'Length_comp',"Age_comp",'Size_at_age','Recruitment') 
+mainlike_components_labels  <- c('Total likelihood','Index likelihood','Length likelihood',"Age likelihood",'Size_at_age likelihood','Recruitment likelihood')
+
+# 19. Funciones para generar plots de perfil de verosimilitud ----
+### SSplotProfile() ----
+png(file.path(plotdir,"R0_profile_plot.png"),width=7,height=4.5,res=300,units='in')
+par(mar=c(5,4,1,1))
+
+SSplotProfile(prof.R0.summary,           # summary object
+              profile.string = "R0",     # substring of profile parameter
+              profile.label=expression(log(italic(R)[0])), ymax=150,minfraction = 0.001,
+              pheight=4.5, 
+              print=FALSE, 
+              plotdir=plotdir, 
+              components = mainlike_components, 
+              component.labels = mainlike_components_labels,
+              add_cutoff = TRUE,
+              cutoff_prob = 0.95)
+
+Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
+Baselab <- paste(Baseval,sep="")
+axis(1,at=Baseval,label=Baselab)
+abline(v = Baseval, lty=2)
+dev.off()
+
+### SSplotComparisons() ----
+# Comparación de series de tiempo 
+labs <- paste("SR_Ln(R0) = ",R0.vec)
+labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",Baseval,"(Base model)")
+
+SSplotComparisons(prof.R0.summary,
+                  legendlabels=labs,
+                  pheight=4.5,png=TRUE,
+                  plotdir=plotdir,
+                  legendloc='bottomleft')
+
+### PinerPlot() ----
+#### R0_profile_plot_Length_like ----
+png(file.path(plotdir,"R0_profile_plot_Length_like.png"),width=7,height=4.5,res=300,units='in')
+par(mar=c(5,4,1,1))
+PinerPlot(prof.R0.summary, 
+          profile.string = "R0", 
+          component = "Length_like",
+          main = "Changes in length-composition likelihoods by fleet",
+          add_cutoff = TRUE,
+          cutoff_prob = 0.95)
+Baseval <- round(Base$parameters$Value[grep("SR_LN",
+                                      Base$parameters$Label)],2)
+Baselab <- paste(Baseval,sep="")
+axis(1,at=Baseval,
+     label=Baselab)
+abline(v = Baseval, lty=2)
+dev.off()
+#### R0_profile_plot_Survey_like ----
+png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),width=7,height=4.5,res=300,units='in')
+par(mar=c(5,4,1,1))
+PinerPlot(prof.R0.summary, profile.string = "R0", component = "Surv_like",main = "Changes in Index likelihoods by fleet",
+          add_cutoff = TRUE,
+          cutoff_prob = 0.95, legendloc="topleft")
+Baseval <- round(Base$parameters$Value[grep("SR_LN",Base$parameters$Label)],2)
+Baselab <- paste(Baseval,sep="")
+axis(1,at=Baseval,label=Baselab)
+abline(v = Baseval, lty=2)
+dev.off()
+```
+
+# COMPARACION DE MODELOS con distinto desembarque
+
+```{r}
 #PLOT labels, name of each model run
-legend.labels <- c('SSdir1','SSdir2','SSdir3')
+legend.labels <- c( 
+                  "s1",
+                 "s2",
+                 "s3",
+                 "s4",
+                 "s5",
+                 "s6",
+                 "s7")
 
@@
 #note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist <- SSgetoutput(keyvec = NULL, dirvec = c(SSdir1,SSdir2,SSdir3),	getcovar = F)
+biglist <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir1,
+                                  dir2,
+                                  dir3,
+                                  dir4,
+                                  dir5,
+                                  dir6,
+                                  dir7),
+                       getcovar = F)
 
@@
 
-SSplotComparisons(summaryoutput, subplots = 1:20, plot = TRUE,
-                  print = T, endyrvec = "default", indexfleets = NULL, indexUncertainty = FALSE,
-                  indexQlabel = TRUE, indexQdigits = 4, indexSEvec = "default",
-                  indexPlotEach = FALSE, labels = c("Year", "Spawning biomass (t)",
-                                                    "Relative spawning biomass", "Age-0 recruits (1,000s)",
-                                                    "Recruitment deviations", "Index", "Log index", "1 - SPR", "Density",
-                                                    "Management target", "Minimum stock size threshold", "Spawning output",
-                                                    "Harvest rate"), col = NULL, shadecol = NULL, pch = NULL,
-                  lty = 1, lwd = 2, spacepoints = 10, staggerpoints = 1,
-                  initpoint = 0, tickEndYr = TRUE, shadeForecast = TRUE,
-                  xlim = "default", ylimAdj = 1, xaxs = "r", yaxs = "r",
-                  type = "o", uncertainty = TRUE, shadealpha = 0.1, legend = TRUE,
-                  legendlabels = "default", legendloc = "topright",
-                  legendorder = "default", legendncol = 1, sprtarg = NULL,
-                  btarg = NULL, minbthresh = NULL, pwidth = 6.5, pheight = 5,
-                  punits = "in", res = 300, ptsize = 10, cex.main = 1,
-                  plotdir = "C:\\Users\\mauricio.mardones\\Documents\\IFOP\\Loco_Assessment_AMERB\\SA_Loco", filenameprefix = "", densitynames = c("SSB_Virgin",
-                                                                                                                     "R0"), densityxlabs = "default", densityscalex = 1,
-                  densityscaley = 1, densityadjust = 1, densitysymbols = TRUE,
-                  densitytails = TRUE, densitymiddle = FALSE, densitylwd = 1,
-                  fix0 = TRUE, new = TRUE, add = FALSE, par = list(mar = c(5, 4, 1,
-                                                                           1) + 0.1), verbose = TRUE, mcmcVec = FALSE,
-                  show_equilibrium = TRUE)
+SSplotComparisons(summaryoutput,
+                  legendlabels = c("s1","s2",
+                 "s3",
+                 "s4",
+                 "s5",
+                 "s6",
+                 "s7"),
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"))
+```
 
+```{r}
 SStableComparisons(summaryoutput,
-                   likenames = c("TOTAL", "Survey", "Length_comp", "Age_comp", "priors",
-                                 "Size_at_age"), names = c("Recr_Virgin", "R0", "steep", "NatM",
-                                                           "L_at_Amax", "VonBert_K", "SSB_Virg", "Bratio_2017", "SPRratio_2016"),
-                   digits = NULL, modelnames = "default", csv = FALSE,
-                   csvdir = "C:\\Users\\mauricio.mardones\\Documents\\IFOP\\Loco_Assessment_AMERB\\SA_Loco",
-                   csvfile = "parameter_comparison_table.csv", verbose = TRUE,
+                   likenames = c("TOTAL", 
+                                 "Survey", 
+                                 "Length_comp",
+                                 "Age_comp", 
+                                 "priors",
+                                 "Size_at_age"), 
+                   names = c("Recr_Virgin",
+                             "R0", 
+                             "steep",
+                             "NatM",
+                             "L_at_Amax", 
+                             "VonBert_K", 
+                             "SSB_Virg", 
+                             "Bratio_2017",
+                             "SPRratio_2016"),
+                   digits = NULL,
+                   modelnames = c("s1","s2",
+                 "s3",
+                 "s4",
+                 "s5",
+                 "s6",
+                 "s7"),
+                   csv = FALSE,
+                   csvdir =~"/DOCAS/SA_Krill",
+                   csvfile = "parameter_comparison_table.csv",
+                 verbose = TRUE,
                    mcmc = FALSE)
 ```
-
-# Analisis retrospectivo
-
-```{r}
-
-#do retrospective model runs
-SS_doRetro('C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\5', '', newsubdir = "retrospectives",
-           subdirstart = "retro", years = 0:-5, overwrite = TRUE, exefile = "ss",
-           extras = "-nox -nohess", intern = FALSE, CallType = "system",
-           RemoveBlocks = FALSE)
-retroModels <- SSgetoutput(dirvec=file.path('C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\5', "retrospectives",paste("retro",0:-5,sep="")))
-retroSummary <- SSsummarize(retroModels)
-endyrvec <- retroSummary$endyrs + 0:-5
-SSplotComparisons(retroSummary, endyrvec=endyrvec, legendlabels=paste("Data",0:-5,"years"),
-                  plotdir='C:\\SAFS\\SS\\Modelos_Seattle\\Anchoveta\\month\\',plot=TRUE,print=T)
-TableCompare <- SStableComparisons(retroSummary,likenames=like, names=names, 
-                                   modelnames=c('B','-1','-2','-3','-4','-5'), csv=TRUE, csvfile="RetroRuns.csv",verbose=FALSE)
-##############################################################################################
-
-```

### 2024-03-27 Run_SS3_Krill_481.Rmd
td
Mensaje de commit inicial
test
some changes in .bib

@@
 author:
-  - Mardones Mauricio^[Instituo de Fomento Pesquero, mardones.mauricio@gmail.com]
-  - César Cárdenas^[Instituto Antártico Chileno]
+  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
+  - César Cárdenas^[Instituto Antártico Chileno, Chile]
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
 ```
+
 \newpage
@@
 knitr::opts_chunk$set(echo = FALSE,
-                      eval=FALSE,
                       message = FALSE,
@@
                       dpi = 300,
-                      tidy.opts=list(width.cutoff=80), 
+                      tidy.opts=list(width.cutoff=50), 
                       tidy=TRUE)
@@
 
-The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystemic framework, highlights the importance of incorporating spatial heterogeneity.
+The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity.
 The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill.
 By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics.
-This nuanced approach allows for a more refined understanding of diferente zones shape krill abundance, fishing mortalyty and recruitment patterns across the Antarctic Peninsula.
+This nuanced approach allows for a more refined understanding of different zones shape krill abundance, fishing mortality and recruitment patterns across the Antarctic Peninsula.
 Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
@@
 
+**(add refs)**
+
+Antarctica, with its pristine and unique ecosystems, hosts one of the most crucial marine resources, the Antarctic krill (Euphausia superba).
+These small crustaceans play a pivotal role in the Southern Ocean's food web, serving as a primary food source for a wide range of marine species, including whales, seals, and seabirds.
+Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
+
+In this context, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) plays a central role in the conservation and sustainable management of marine resources in the Southern Ocean.
+As part of its mandate, CCAMLR regularly conducts stock assessments to evaluate the status of key species, including Antarctic krill, and inform decision-making processes regarding fisheries management and conservation measures.
+
+Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
+However, these approaches may overlook critical aspects of krill ecology, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
+Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.
+
+In response to this need, we propose a comprehensive stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach.
+Unlike previous assessments, our model will integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics.
+Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.
+
+Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions, sea ice dynamics, and predator foraging behavior.
+By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
+This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies.
+
+Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent.
+These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
+By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
+
+Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
+Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds.
+Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem.
+By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
+
+The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill.
+By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
+This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
+
+In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula.
+By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR.
+Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.
+
 # METHODOLOGY
+
+## Study Area in the Antarctic Peninsula
+
+The study area for the assessment of krill populations in the Antarctic Peninsula is divided into strata to account for spatial heterogeneity.
+The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
+Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.
+
+The strata in the Antarctic Peninsula may include different regions such as coastal areas, offshore waters, and areas influenced by major oceanographic features like currents and frontal systems.
+Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability.
+
+![Study Area 48.1 and strata](Figs/481.png)
+
+## Data Sources and Recolection
+
+Stock Synthesis relies on multiple sources of data to inform the population dynamics model. These data sources may include:
+
+1. **Fisheries Data:** Information on krill catches and fishing effort from commercial fisheries operating in the Antarctic Peninsula region.
+
+2.
+**Population Monitoring Data:** Data collected from scientific research cruises and monitoring programs specifically designed to assess krill populations.
+This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.
+
+3.
+**Environmental Data:** Environmental data such as sea surface temperature, chlorophyll concentration, and sea ice extent are also incorporated into the assessment.
+These data help to understand the influence of environmental factors on krill populations and their distribution.
+
+4.
+**Predator information:** MAPPPD.
+Predator “fleet” to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.
+
+In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment.
+These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
 
@@
 
+## Population Dynamics Model
+
+In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time.
+This model incorporates key biological parameters such as growth rates, mortality rates, recruitment, and spawning biomass.
+The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.
+
+A typical state-space model for krill population dynamics can be represented as:
+
+\[
+N_t = N_{t-1} \cdot e^{(r - M)} + R
+\]
+
+Where:
+- \(N_t\) is the abundance of krill at time \(t\).
+- \(N_{t-1}\) is the abundance of krill at the previous time step.
+- \(r\) is the intrinsic growth rate of the population.
+- \(M\) is the mortality rate.
+- \(R\) is the recruitment of new individuals into the population.
+
+This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
+
+## Statistical model (SS3)
+
+Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations, including krill (Euphausia superba) in the Antarctic Peninsula region.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP.
+
+The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the most updated version (V.3.30.21).
+SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
+SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
+The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
+The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].
+
+By integrating data from multiple sources and considering spatial heterogeneity, the assessment methodology using Stock Synthesis v.3.30.21 provides a robust framework for evaluating the status of krill populations in the Antarctic Peninsula region.
+This information is essential for supporting management decisions aimed at ensuring the sustainable use of krill resources in this ecologically sensitive area.
+
 ## Scenarios
@@
 
-## Statistical model (SS3)
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 
-The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the most updated version (V.3.30.21).
-SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
-SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
-The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].
-
-## Model setup
-
-```{r message=FALSE}
+```{r message=FALSE, eval=TRUE}
 # install.packages("devtools")
 # devtools::install_github("r4ss/r4ss", ref="development")
+
+# 
 # install.packages("caTools")
 # library("caTools")
+# 
 # install.packages("r4ss")
@@
 library(ss3diags)
+library(kableExtra)
 ```
@@
                         forecast=T)
+#s5
+base.model5 <- SS_output(dir=dir5,
+                        covar=T,
+                        forecast=T)
+#s6
+base.model6 <- SS_output(dir=dir6,
+                        covar=T,
+                        forecast=T)
 
@@
                         forecast=T)
-
 ```
-
-Saco los outputs en html
 
@@
 
-Data disponible para este escenario. Espinel es la serie mas consistente del conjunto de datos.
+```{r}
+SSplotData(base.model01, subplot = 1, 
+           fleetnames = c("Fishery", "Survey", "Predator"),
+           fleetcol = c(7,6, 4))
 
-```{r}
-SSplotData(base.model, subplots = 2)
+SSplotData(base.model2, 
+           subplot = 1,
+           pheight = 15)
 
@@
 ```{r}
-SSplotSelex(base.model)
+SSplotSelex(base.model2,
+            subplots = 1)
+```
+
+```{r eval=FALSE}
+SSplotBiology(base.model2, 
+              subplots =2, 
+              labels = c("Length (cm)",
+                         "Age (yr)",
+                         "Maturity", 
+                         "Mean weight (kg) in last year",
+                          "Spawning output",
+                         "Length (cm, beginning of the year)", 
+                         "Natural mortality",
+                          "Female weight (kg)", 
+                         "Female length (cm)", 
+                         "Fecundity", 
+                         "Default fecundity label",
+                          "Year", 
+                         "Hermaphroditism transition rate", 
+                         "Fraction females by age at equilibrium"),
+ )
+
 ```
@@
 aporte de las cohortes por año para las capturas.
-```{r}
-SSplotCohortCatch(base.model, subplots = 1)
+```{r eval=FALSE}
+SSplotCohortCatch(base.model2, subplots = 1)
 ```
@@
 AJuste de tallas por flota
-```{r}
-SSplotComps(base.model, subplots = 1)
+```{r eval=FALSE}
+SSplotComps(base.model2, subplots = 1)
 ```
@@
 ```{r}
-SSplotDynamicB0(base.model, uncertainty = F)
+SSplotDynamicB0(base.model2, uncertainty = F)
 #SSplotSPR(base.model3)
-SSplotPars(base.model)
+```
+
+```{r eval=FALSE}
+SSplotPars(base.model2)
 
@@
 ```{r}
-SSplotTimeseries(base.model7, subplot = 1)
+SSplotTimeseries(base.model01, subplot = 7)
+```
+Salida de las biomasas con todas las flotas
+
+```{r}
+SSplotTimeseries(base.model2, subplot = 7)
 ```
 
-```{r}
-SSplotTimeseries(base.model,
-  subplot=1,
-  add = FALSE,
-  forecastplot = FALSE,
-  uncertainty = TRUE,
-  bioscale = 1,
-  minyr = -Inf,
-  maxyr = Inf,
-  plot = TRUE,
-  print = FALSE,
-  plotdir = "default",
-  verbose = TRUE,
-  btarg = "default",
-  minbthresh = "default",
-  xlab = "Year",
-  labels = NULL,
-  pwidth = 6.5,
-  pheight = 5,
-  punits = "in",
-  res = 300,
-  ptsize = 10,
-  cex.main = 1,
-  mainTitle = FALSE,
-  mar = NULL
-)
-```
+# RESULTS
 
-# Diagnosis Base Model (Jabba)
+## Diagnosis Base Model
 
-step to do a good practice in diagnosis is;
+Step to do a good practice in model diagnosis is;
+
 - 1. Convergence. Final convergence criteria is 1.0e-04
-- 2. Residual 
-- 3. Restrospective analysis
+
+- 2. Residual (visual and metrics)
+
+- 3. Retrospective analysis (Mhon Parameter)
+
 - 4. Likelihood profile
@@
 
-```{r}
-SSplotRunstest(base.model4, 
+```{r eval=TRUE}
+SSplotRunstest(base.model2, 
                subplots = "cpue",
                add=T)
-SSplotRunstest(base.model4, 
+SSplotRunstest(base.model2, 
                subplots = "len",
@@
 ```{r}
-SSplotJABBAres(base.model,
+SSplotJABBAres(base.model2,
                subplots = "cpue",
                add=T)
-SSplotJABBAres(base.model,
+SSplotJABBAres(base.model2,
                subplots = "len",
@@
 
-### Predator fleet with RW
+## Predator fleet with RW
 
@@
 
-# Retrospectivo
+## Retrospective analysis
 
@@
 ```{r}
-retroModels <- SSgetoutput(dirvec=file.path(dir01,
+retroModels <- SSgetoutput(dirvec=file.path(dir2,
                                             "Retrospective",
@@
 ```
-# Kobe
+## Kobe (status)
 
@@
 ```{r}
-retroModels <- SSgetoutput(dirvec=file.path(mydir, "Retrospective",paste("retro",0:-5, sep="")))
-retroSummary <- SSsummarize(retroModels)
-endyrvec <- retroSummary$endyrs + 0:-5
-save(retroSummary, retroModels, file="retro5.Rdata")
-# Step 1. Identify restrospective period
-# e.g., for end.yr.vec   <- c(2015,2014,2013,2012,2011,2010)
-start.retro <- 0    #end year of model e.g., 2015
-end.retro   <- 5    #number of years for retrospective e.g., 2014,2013,2012,2011,2010
-
-# Step 2. Identify the base directory
-dirname.base <- here("SA_Krill")
-dirname.base
-
-# Step 3. Identify the directory where a completed model run is located
-dirname.completed.model.run <- here("s9")
-dirname.completed.model.run
-
-# Step 4. Create a subdirectory for the Retrospectives
-dirname.Retrospective <- here("Ejercicios","Retrospectivo") 
-dir.create(path=dirname.Retrospective, showWarnings = TRUE, recursive = TRUE)
-setwd(dirname.Retrospective)
-getwd()
-
-# Step 5. Create a subdirectory for the Plots
-dirname.plots <- paste0(dirname.Retrospective,"/plots_retro")
-dir.create(dirname.plots)
-
-# 7. Copiar los archivos necesarios de "Reference_run" al directorio "Perfil_Verosimilitud" ----
-
-copy_SS_inputs(dir.old = dirname.completed.model.run, 
-               dir.new =  dirname.Retrospective,
-               copy_exe = TRUE,
-               verbose = FALSE)
-
-# #----------------- copy model run files ----------------------------------------
-# file.copy(paste(dirname.completed.model.run,       "starter.ss_new", sep="/"),
-#           paste(dirname.Retrospective, "starter.ss", sep="/"))
-# 
-# file.copy(paste(dirname.completed.model.run,       "control.ss_new", sep="/"),
-#           paste(dirname.Retrospective, "CONTROL.SS", sep="/"))
-# 
-# file.copy(paste(dirname.completed.model.run,       "data.ss", sep="/"),
-#           paste(dirname.Retrospective, "DATA.SS", sep="/"))	
-# 
-# file.copy(paste(dirname.completed.model.run,       "forecast.ss", sep="/"),
-#           paste(dirname.Retrospective, "forecast.ss", sep="/"))
-# 
-# file.copy(paste(dirname.completed.model.run,       "ss_win.exe", sep="/"),
-#           paste(dirname.Retrospective, "SS.exe", sep="/"))
-
-# # Required for assessments with wtatage.ss files
-# file.copy(paste(dirname.completed.model.run,       "wtatage.ss", sep="/"),
-#           paste(dirname.Retrospective, "wtatage.ss", sep="/"))
-
-#------------Make Changes to the Starter.ss file (DC Example) ------------------------------- 
-starter <- readLines(paste(dirname.Retrospective, "/starter.ss", sep=""))
-
-# 1) Starter File changes to speed up model runs
-# Run Display Detail
-#[8] "2 # run display detail (0,1,2)" 
-linen <- grep("# run display detail", starter)
-starter[linen] <- paste0( 1 , " # run display detail (0,1,2)" )
-write(starter, paste(dirname.Retrospective, "starter.ss", sep="/"))
-
-#------------ r4SS retrospective calculations-------------------------------
-
-# Step 6. Run the retrospective analyses with r4SS function "SS_doRetro"
-retro(masterdir=dirname.Retrospective,
-           oldsubdir="", 
-           newsubdir="retrospectives",
-           years=start.retro:-end.retro)
-
-# Step 7. Read "SS_doRetro" output
-retroModels <- SSgetoutput(dirvec=file.path(dirname.Retrospective, "retrospectives",paste("retro",start.retro:-end.retro,sep="")))
-
-# Step 8. Summarize "SS_doRetro" output
-retroSummary <- SSsummarize(retroModels)
-
-sspar(mfrow=c(1,2),plot.cex=0.8)
-rb = SSplotRetro(retroSummary,add=T,forecast = F,legend = F,verbose=F)
-rf = SSplotRetro(retroSummary,add=T,subplots="F", ylim=c(0,0.4),
-                 forecast = F,legendloc="topleft",legendcex = 0.8,verbose=F)
-
-sspar(mfrow = c(2, 1), plot.cex = 0.8)
-
-SSplotRetro(retroSummary, 
-            add = T, 
-            forecast = T, 
-            legend = F, 
-            verbose = F, 
-            xmin = 1970)
-SSplotRetro(retroSummary, 
-            add = T, 
-            subplots = "F", 
-            ylim = c(0, 0.06), 
-            forecast = T,
-            legendloc = "topleft", 
-            legendcex = 0.8, 
-            verbose = F, 
-            xmin = 1970)
-
 SShcbias(retroSummary,quant="SSB",verbose=F)
@@
 
-# Verosimilitud
+## Verosimilitud
 
-```{r}
+```{r eval=FALSE, echo= FALSE}
 # 1. Identificar el directorio donde se encuentra el modelo base ----
@@
 
-# COMPARACION DE MODELOS con distinto desembarque
+## Comparision outputs betwwen scenarios
 
@@
 #PLOT labels, name of each model run
-legend.labels <- c( 
+legend.labels <- c( "s01",
                   "s1",
@@
 biglist <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(
+                       dirvec = c(dir01,
                                   dir1,
@@
 SSplotComparisons(summaryoutput,
-                  legendlabels = c("s1","s2",
+                  legendlabels = c("s01",
+                                   "s1",
+                                   "s2",
                  "s3",
@@
 ```{r}
-SStableComparisons(summaryoutput,
+comtable <- SStableComparisons(summaryoutput,
                    likenames = c("TOTAL", 
@@
                    digits = NULL,
-                   modelnames = c("s1","s2",
+                   modelnames = c("s01", "s1","s2",
                  "s3",
@@
                    mcmc = FALSE)
+kbl(comtable, booktabs = T,format = "latex",
+    caption = "Comparacion likelihood y parámetros s01, s1, s2, s3, s4, s5, s6 y s7")  %>% 
+    kable_styling(latex_options = "scale_down")
+
 ```
+
+\newpage
+# DISCUSION
+
+\newpage
+# CONCLUSION
+
+\newpage
+# REFERENCES

### 2024-03-27 Run_SS3_Krill_481.Rmd
commit pdf

@@
 library(kableExtra)
+library(doParallel) # facilita la ejecución paralela en R
+detectCores()
+registerDoParallel(8)
+library(ggpubr)
 ```
@@
            fleetnames = c("Fishery", "Survey", "Predator"),
-           fleetcol = c(7,6, 4))
+           fleetcol = c("blue","green", "red"))
 
@@
 ```{r eval=TRUE}
-SSplotRunstest(base.model2, 
+SSplotRunstest(base.model01, 
                subplots = "cpue",
@@
 
-```{r}
-retroModels <- SSgetoutput(dirvec=file.path(dir2,
+```{r echo=FALSE}
+#s01
+retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
                                             "Retrospective",
@@
 
-retroSummary <- SSsummarize(retroModels)
+retroSummary01 <- SSsummarize(retroModels01)
+# s1
+retroModels1 <- SSgetoutput(dirvec=file.path(dir1,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary1 <- SSsummarize(retroModels1)
+#s2
+retroModels2 <- SSgetoutput(dirvec=file.path(dir2,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary2 <- SSsummarize(retroModels2)
+#s3
+retroModels3 <- SSgetoutput(dirvec=file.path(dir3,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary3 <- SSsummarize(retroModels3)
+#s4
+retroModels4 <- SSgetoutput(dirvec=file.path(dir4,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary4 <- SSsummarize(retroModels4)
+#s5
+retroModels5 <- SSgetoutput(dirvec=file.path(dir5,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary5 <- SSsummarize(retroModels5)
+#s6
+retroModels6 <- SSgetoutput(dirvec=file.path(dir6,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary6 <- SSsummarize(retroModels6)
+#s7
+retroModels7 <- SSgetoutput(dirvec=file.path(dir7,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary7 <- SSsummarize(retroModels7)
+
 ```
 
-```{r}
+```{r echo FALSE}
 #save(retroSummary, retroModels, file="retro5.Rdata")
 
-retro01 <- SSplotRetro(retroSummary,
+retro01 <- SSplotRetro(retroSummary01,
             add=T,
             forecast = F,
-            legend = F,
+            legend = T,
             verbose=F)
-retro2 <- SSplotRetro(retroSummary,
+retro1 <- SSplotRetro(retroSummary1,
             add=T,
             forecast = F,
-            legend = F,
+            legend = T,
             verbose=F)
-retro1 <- SSplotRetro(retroSummary,
+retro2 <- SSplotRetro(retroSummary2,
             add=T,
             forecast = F,
-            legend = F,
+            legend = T,
+            verbose=F)
+retro3 <- SSplotRetro(retroSummary3,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F)
+retro4 <- SSplotRetro(retroSummary4,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F)
+retro5 <- SSplotRetro(retroSummary5,
+            add=T,
+            forecast = F,
+            legend = T,
             verbose=F)
 
-retro7 <- SSplotRetro(retroSummary,
+retro6 <- SSplotRetro(retroSummary6,
             add=T,
             forecast = F,
-            legend = F,
-            verbose=F,
-            legendlabels="default")
+            legend = T,
+            verbose=F)
+
+retro7 <- SSplotRetro(retroSummary7,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F)
 ```
+```{r echo=FALSE}
+ggarrange(retro01,
+          retro1,
+          retro2,
+          retro3,
+          retro4,
+          retro5,
+          retro6,
+          retro7, ncol = 4, nrow = 4)
+```
+
 ## Hindcast Cross-Validation and prediction skill
@@
 ```{r}
-hci = SSplotHCxval(retroSummary, 
+hci = SSplotHCxval(retroSummary2, 
                    add = T, 
@@
 
+mvln = SSdeltaMVLN(base.model2,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
 ```
 
-```{r}
-SShcbias(retroSummary,quant="SSB",verbose=F)
-SShcbias(retroSummary,quant="F",verbose=F)
+```{r echo=FALSE}
+# model01
+tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
+tablebias02 <- SShcbias(retroSummary01,quant="F",verbose=F)
+
+kbl(tablebias01, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s01")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias02, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s01")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model2
+tablebias1 <- SShcbias(retroSummary2,quant="SSB",verbose=F)
+tablebias2 <- SShcbias(retroSummary2,quant="F",verbose=F)
+
+kbl(tablebias1, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s2")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias2, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s2")  %>% 
+    kable_styling(latex_options = "HOLD_position")
 ```
@@
 
-```{r eval=FALSE, echo= FALSE}
+```{r eval=FALSE, echo = FALSE}
 # 1. Identificar el directorio donde se encuentra el modelo base ----
-dirname.model.run <- here("s10")
-
+dirname.model.run <- here("s2")
 # 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
-
 dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
 dir.create(path=dirname.R0.profile, showWarnings = TRUE, recursive = TRUE)
-
 # 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
-
 plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
 dir.create(path=plotdir, showWarnings = TRUE, recursive = TRUE)
-
 # 4. Crear un subdirectorio llamado "simple" ----
-
 reference.dir <- paste0(dirname.R0.profile,'/simple') 
 dir.create(path=reference.dir, showWarnings = TRUE, recursive = TRUE)
-
 # 5. Copiar el resultado del modelo base completo en este directorio ----
-
 file.copy(Sys.glob(paste(dirname.model.run, "*.*", sep="/"),
@@
                     reference.dir)
-
 # 6. Leer la salida del modelo base ----
-
 Base <- SS_output(dir=reference.dir,covar=T)
-
 # 7. Copiar los archivos necesarios de "simple" al directorio "Perfil_Verosimilitud" ----
-
 copy_SS_inputs(dir.old = reference.dir, 
@@
                verbose = FALSE)
-
 # 8. Leer los archivos del modelo ----
-
 inputs <- r4ss::SS_read(dir = dirname.R0.profile)
-
 # 9. Editar el archivo control la fase de estimación recdev ----
-
 inputs$ctl$recdev_phase <- 1
-
 # 10. Editar el archivo starter para leer los valores de inicio ----
-
 inputs$start$init_values_src <- 0
-
 # 11. Vector de valores para el perfil ----
-
-R0.vec <- seq(7,11,0.5)  
+R0.vec <- seq(15,20,0.5)  
 Nprof.R0 <- length(R0.vec)
-
 # 12. Cambiar el nombre del archivo control en el archivo starter.ss ----
-
 inputs$start$ctlfile <- "control_modified.ss" 
-
 # 13. Incluir prior_like para parámetros no estimados ----
-
 inputs$start$prior_like <- 1                                 
-
 # 14. Escribir los modelos modificados ----
-
 r4ss::SS_write(inputs, dir = dirname.R0.profile, overwrite = TRUE)
-
 # 15. Ejecutar la función profile() ----
-
 ?SS_profile()
-profile <- profile(dir="dir10", # directory
-                      model="ss_osx",
+profile <- profile(dir=dirname.R0.profile, # directory
+                      exe="ss_osx",
                       oldctlfile ="control.ss",
@@
                       profilevec=R0.vec)
-
 # 16. Leer los archivos de salida ----
@@
                               getcovar = FALSE) 
-
 # 17. Resumir las salidas con la función SSsummarize()  ----
 prof.R0.summary <- SSsummarize(prof.R0.models)
-
 # 18. Identificar los componentes de Verosimilitud ----
-mainlike_components         <- c('TOTAL',"Survey", 'Length_comp',"Age_comp",'Size_at_age','Recruitment') 
-mainlike_components_labels  <- c('Total likelihood','Index likelihood','Length likelihood',"Age likelihood",'Size_at_age likelihood','Recruitment likelihood')
-
+mainlike_components         <- c('TOTAL',
+                                 "Survey", 
+                                 'Length_comp',
+                                 "Age_comp",
+                                 'Size_at_age',
+                                 'Recruitment') 
+mainlike_components_labels  <- c('Total likelihood',
+                                 'Index likelihood',
+                                 'Length likelihood',
+                                 "Age likelihood",
+                                 'Size_at_age likelihood',
+                                 'Recruitment likelihood') 
 # 19. Funciones para generar plots de perfil de verosimilitud ----
 ### SSplotProfile() ----
-png(file.path(plotdir,"R0_profile_plot.png"),width=7,height=4.5,res=300,units='in')
+png(file.path(plotdir,"R0_profile_plot.png"),
+    width=7,
+    height=4.5,
+    res=300,
+    units='in')
 par(mar=c(5,4,1,1))
-
 SSplotProfile(prof.R0.summary,           # summary object
               profile.string = "R0",     # substring of profile parameter
-              profile.label=expression(log(italic(R)[0])), ymax=150,minfraction = 0.001,
+              profile.label=expression(log(italic(R)[0])), 
+              ymax=150,minfraction = 0.001,
               pheight=4.5, 
@@
 labs <- paste("SR_Ln(R0) = ",R0.vec)
-labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",Baseval,"(Base model)")
+labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
+                                               Baseval,"(Base model)")
 
@@
 #### R0_profile_plot_Length_like ----
-png(file.path(plotdir,"R0_profile_plot_Length_like.png"),width=7,height=4.5,res=300,units='in')
+png(file.path(plotdir,"R0_profile_plot_Length_like.png"),
+    width=7,
+    height=4.5,
+    res=300,
+    units='in')
 par(mar=c(5,4,1,1))
@@
 #### R0_profile_plot_Survey_like ----
-png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),width=7,height=4.5,res=300,units='in')
+png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),
+    width=7,
+    height=4.5,
+    res=300,
+    units='in')
 par(mar=c(5,4,1,1))
-PinerPlot(prof.R0.summary, profile.string = "R0", component = "Surv_like",main = "Changes in Index likelihoods by fleet",
+PinerPlot(prof.R0.summary, 
+          profile.string = "R0", 
+          component = "Surv_like",
+          main = "Changes in Index likelihoods by fleet",
           add_cutoff = TRUE,
           cutoff_prob = 0.95, legendloc="topleft")
-Baseval <- round(Base$parameters$Value[grep("SR_LN",Base$parameters$Label)],2)
+Baseval <- round(Base$parameters$Value[grep("SR_LN",
+                                            Base$parameters$Label)],2)
 Baselab <- paste(Baseval,sep="")
@@
 
-```{r}
+```{r echo=FALSE}
 #PLOT labels, name of each model run
@@
                             "Harvest rate"))
+
+comp <- SSplotComparisons(summaryoutput,
+                  legendlabels = c("s01",
+                                   "s1",
+                                   "s2",
+                 "s3",
+                 "s4",
+                 "s5",
+                 "s6",
+                 "s7"),
+                 subplots = c(2, 16))
+comp
 ```

### 2024-04-04 Run_SS3_Krill_481.Rmd
add refs

@@
 
-The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the most updated version (V.3.30.21).
+The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21).
 SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
@@
 # r4ss::run(
-#   dir = dir1,
+#   dir = dir2,
 #   exe = "ss_osx",
@@
 ```{r eval=FALSE, message=F, include=FALSE}
-SS_plots(base.model7, 
+SS_plots(base.model2, 
          uncertainty = TRUE,
@@
 # model2
-tablebias1 <- SShcbias(retroSummary2,quant="SSB",verbose=F)
-tablebias2 <- SShcbias(retroSummary2,quant="F",verbose=F)
+tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
+tablebias07 <- SShcbias(retroSummary7,quant="F",verbose=F)
 
-kbl(tablebias1, booktabs = T,format = "latex",
+kbl(tablebias7, booktabs = T,format = "latex",
     caption = "Rho parameter in SSB model s2")  %>% 
@@
 
-kbl(tablebias2, booktabs = T,format = "latex",
+kbl(tablebias7, booktabs = T,format = "latex",
     caption = "Rho parameter in F model s2")  %>% 

### 2024-05-21 index.Rmd
adding html run code

--- 2024-04-04 Run_SS3_Krill_481.Rmd
@@
 title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
-subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
-author:
-  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
-  - César Cárdenas^[Instituto Antártico Chileno, Chile]
+subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
+author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
 linkcolor: blue
-always_allow_html: yes
 output:
-  pdf_document
+  html_document:
+    keep_md: true
+    toc: true
+    toc_deep: 3
+    toc_float:
+      collapsed: false
+      smooth_scroll: false
+    theme: cosmo
+    fontsize: 0.9em
+    linestretch: 1.7
+    html-math-method: katex
+    self-contained: true
+    code-tools: true
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
-
-```{=tex}
-\fontsize{12}{16}
-\selectfont{}
-```
-\newpage
-
-```{=latex}
-\setcounter{tocdepth}{3}
-\tableofcontents
-```
-
-\newpage
 
@@
 
-# ABSTRACT
-
-The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity.
-The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill.
-By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics.
-This nuanced approach allows for a more refined understanding of different zones shape krill abundance, fishing mortality and recruitment patterns across the Antarctic Peninsula.
-Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
-
-*Keywords: Krill populations, dynamic population, stock assessment, SS3, spatial implicit, management, CCAMLR.*
-
-# INTRODUCTION
-
-**(add refs)**
-
-Antarctica, with its pristine and unique ecosystems, hosts one of the most crucial marine resources, the Antarctic krill (Euphausia superba).
-These small crustaceans play a pivotal role in the Southern Ocean's food web, serving as a primary food source for a wide range of marine species, including whales, seals, and seabirds.
-Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-
-In this context, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) plays a central role in the conservation and sustainable management of marine resources in the Southern Ocean.
-As part of its mandate, CCAMLR regularly conducts stock assessments to evaluate the status of key species, including Antarctic krill, and inform decision-making processes regarding fisheries management and conservation measures.
-
-Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
-However, these approaches may overlook critical aspects of krill ecology, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.
-
-In response to this need, we propose a comprehensive stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach.
-Unlike previous assessments, our model will integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics.
-Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.
-
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions, sea ice dynamics, and predator foraging behavior.
-By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies.
-
-Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
-
-Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
-Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds.
-Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem.
-By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
-
-The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill.
-By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
-This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
-
-In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula.
-By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR.
-Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.
-
-# METHODOLOGY
-
-## Study Area in the Antarctic Peninsula
-
-The study area for the assessment of krill populations in the Antarctic Peninsula is divided into strata to account for spatial heterogeneity.
-The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.
-
-The strata in the Antarctic Peninsula may include different regions such as coastal areas, offshore waters, and areas influenced by major oceanographic features like currents and frontal systems.
-Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability.
-
-![Study Area 48.1 and strata](Figs/481.png)
-
-## Data Sources and Recolection
-
-Stock Synthesis relies on multiple sources of data to inform the population dynamics model. These data sources may include:
-
-1. **Fisheries Data:** Information on krill catches and fishing effort from commercial fisheries operating in the Antarctic Peninsula region.
-
-2.
-**Population Monitoring Data:** Data collected from scientific research cruises and monitoring programs specifically designed to assess krill populations.
-This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.
-
-3.
-**Environmental Data:** Environmental data such as sea surface temperature, chlorophyll concentration, and sea ice extent are also incorporated into the assessment.
-These data help to understand the influence of environmental factors on krill populations and their distribution.
-
-4.
-**Predator information:** MAPPPD.
-Predator “fleet” to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.
-
-In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment.
-These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
-
-## Conceptual Model
-
-## Phases in model construction
-
-The process of modeling and statistical analysis of a database can be structured according to the following guidelines:
-
-- Contextualization of the problem. Definition of objectives and variables.
-- Experiment design and data collection.
-- Recording and preliminary processing of available information.
-- Graphical inspection and trend identification.
-- Consideration of distributional and relational hypotheses. Proposal of modeling.
-- Model adjustment. Comparison and selection of the best model.
-- Diagnosis and validation of the adjusted model.
-- Assessment of the predictive capacity of the model and prediction.
-- Interpretation and conclusions.
-
-## Population Dynamics Model
+# OVERVIEW
 
@@
 The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.
-
-A typical state-space model for krill population dynamics can be represented as:
-
-\[
-N_t = N_{t-1} \cdot e^{(r - M)} + R
-\]
-
-Where:
-- \(N_t\) is the abundance of krill at time \(t\).
-- \(N_{t-1}\) is the abundance of krill at the previous time step.
-- \(r\) is the intrinsic growth rate of the population.
-- \(M\) is the mortality rate.
-- \(R\) is the recruitment of new individuals into the population.
-
-This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
 
@@
 The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].
+The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* and *ss3diags* library [@Taylor2019; @Winker2023].
 
@@
 |    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+
+## Run Models
 
@@
 ```
+
+Read outputs
 
@@
 
-```{r eval=TRUE}
+```{r eval=FALSE}
 SSplotRunstest(base.model01, 
@@
             verbose=F)
+
 ```
+
 ```{r echo=FALSE}
@@
 \newpage
-# DISCUSION
-
-\newpage
-# CONCLUSION
-
-\newpage
 # REFERENCES

### 2024-05-21 Run_SS3_Krill_481.Rmd
adding fixs

@@
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
+output:
+  bookdown::pdf_document2:
+    number_sections: false
+fig_caption: yes
 bibliography: SA_krill.bib
@@
 link-citations: yes
+toc: false
 linkcolor: blue
-always_allow_html: yes
-output:
-  pdf_document
+linestretch: 1.3
+header-includes:
+- \fontsize{12}{16}
+- \selectfont
+- \usepackage{lscape}
 ---
-
-```{=tex}
-\fontsize{12}{16}
-\selectfont{}
-```
 \newpage
-
-```{=latex}
-\setcounter{tocdepth}{3}
 \tableofcontents
-```
-
 \newpage
@@
 rm(list = ls())
-knitr::opts_chunk$set(echo = FALSE,
+knitr::opts_chunk$set(eval=FALSE,
+                      echo = FALSE,
                       message = FALSE,
@@
 The strata in the Antarctic Peninsula may include different regions such as coastal areas, offshore waters, and areas influenced by major oceanographic features like currents and frontal systems.
-Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability.
-
-![Study Area 48.1 and strata](Figs/481.png)
+Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability (Figure \@ref(fig:mapa))
 
@@
 
+(Figure \@ref(fig:path))
+
 ## Conceptual Model
-
-## Phases in model construction
 
@@
 
-```{r message=FALSE, warning=FALSE}
-dir01 <- here("s01") # agreggate data (no spatial diferences)
-dir1<-here("s1") # Data strata flishery
-dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
-dir3<-here("s3") # whitout S-R
-dir4<-here("s4") # 
-dir5<-here("s5") # whitout S-R
-dir6<-here("s6") # whitout S-R
-dir7<-here("s7") # whitout S-R
-```
-
-```{r eval=FALSE, message=F, include=FALSE}
-# Lista de directorios para correro tordos juntos
-directorios <- c("s01", 
-                 "s1", 
-                 "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
-
-# Bucle para ejecutar el código en cada directorio
-for (dir in directorios) {
-  r4ss::run(
-    dir = dir,
-    exe = "ss_osx",
-    skipfinished = FALSE,
-    show_in_console = TRUE
-  )
-}
-
-#OR with rby separate
-
-#shell(cmd="ss") # run SS to windows
-# or 
-#system('./ss_osx') # to mac
-
-# r4ss::run(
-#   dir = dir2,
-#   exe = "ss_osx",
-#   skipfinished = FALSE, # TRUE will skip running if Report.sso present
-#   show_in_console = TRUE # change to true to watch the output go past
-# )
-```
-
-```{r message=F, include=FALSE}
-#s01
-base.model01 <- SS_output(dir=dir01,
-                        covar=T,
-                        forecast=T)
-#s1
-base.model1 <- SS_output(dir=dir1,
-                        covar=T,
-                        forecast=T)
-#s2
-base.model2 <- SS_output(dir=dir2,
-                        covar=T,
-                        forecast=T)
-#s3
-base.model3 <- SS_output(dir=dir3,
-                        covar=T,
-                        forecast=T)
-#s4
-base.model4 <- SS_output(dir=dir4,
-                        covar=T,
-                        forecast=T)
-#s5
-base.model5 <- SS_output(dir=dir5,
-                        covar=T,
-                        forecast=T)
-#s6
-base.model6 <- SS_output(dir=dir6,
-                        covar=T,
-                        forecast=T)
-
-#s7
-base.model7 <- SS_output(dir=dir7,
-                        covar=T,
-                        forecast=T)
-```
-
-```{r eval=FALSE, message=F, include=FALSE}
-SS_plots(base.model2, 
-         uncertainty = TRUE,
-         datplot = T, 
-         png=T, 
-         aalresids = F,
-         btarg=0.75, 
-         minbthresh=0.2, 
-         forecast=T)
-```
-
-```{r}
+```{r fig.height=7}
 SSplotData(base.model01, subplot = 1, 
@@
 \newpage
+
+# FIGURES AND TABLES
+
+```{r mapa, out.width='100%', fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
+knitr::include_graphics('Figs/481.png')
+```
+\begin{landscape}
+
+```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
+knitr::include_graphics('Figs/pathmod.png')
+```
+
+Plot data
+
+```{r datas2, out.width='60%', fig.cap="Data series using in "s2" scenario"}
+knitr::include_graphics('s2/plots/data_plot.png')
+```
+
+\end{landscape}
+
+\newpage
 # REFERENCES

### 2024-05-21 index.Rmd
adding fixs

@@
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
 ```{r message=FALSE, eval=TRUE}
@@
 }
-
 #OR with rby separate
-
 #shell(cmd="ss") # run SS to windows
@@
 #system('./ss_osx') # to mac
-
 # r4ss::run(
@@
 
-```{r}
+```{r fig.height=7}
 SSplotData(base.model01, subplot = 1, 

### 2024-05-28 SS3_Krill_481.Rmd
improve Intro

--- 2024-05-21 Run_SS3_Krill_481.Rmd
@@
 rm(list = ls())
-knitr::opts_chunk$set(eval=FALSE,
+knitr::opts_chunk$set(eval=TRUE,
                       echo = FALSE,
@@
 
+```{r message=FALSE, eval=TRUE}
+library(here)
+library(kableExtra)
+library(ggpubr)
+library(tibble)
+```
+
 # ABSTRACT
@@
 By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics.
-This nuanced approach allows for a more refined understanding of different zones shape krill abundance, fishing mortality and recruitment patterns across the Antarctic Peninsula.
+This nuanced approach allows for a more refined understanding of different areas features krill structure, fishing mortality and recruitment patterns across the Antarctic Peninsula.
 Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
@@
 
-**(add refs)**
+Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
+On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
+Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
+As part of its mandate, CCAMLR routinely conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
+This regional fisheries management organization has intensified its efforts to promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
 
-Antarctica, with its pristine and unique ecosystems, hosts one of the most crucial marine resources, the Antarctic krill (Euphausia superba).
-These small crustaceans play a pivotal role in the Southern Ocean's food web, serving as a primary food source for a wide range of marine species, including whales, seals, and seabirds.
-Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
+Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
+However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
+Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
+However, these approaches may overlook critical aspects of krill dynamics, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
+Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.
+Integrating the complexity of krill population dynamics presents several challenges.
+Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension.
+Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics.
+Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
+Despite these gaps, the existing data on krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
 
-In this context, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) plays a central role in the conservation and sustainable management of marine resources in the Southern Ocean.
-As part of its mandate, CCAMLR regularly conducts stock assessments to evaluate the status of key species, including Antarctic krill, and inform decision-making processes regarding fisheries management and conservation measures.
-
-Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
-However, these approaches may overlook critical aspects of krill ecology, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.
-
-In response to this need, we propose a comprehensive stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach.
-Unlike previous assessments, our model will integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics.
+In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach that incorporates the fishing and biological history of krill over the past 20 years, as well as ecosystem variables such as environmental conditions, primary productivity levels affecting krill recruitment, and predator abundance.
+Unlike previous assessments, this apporach integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics.
 Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.
 
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions, sea ice dynamics, and predator foraging behavior.
+Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
 By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies.
+This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013].
+Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.
 
-Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
-
-Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
-Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds.
-Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem.
-By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
-
-The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill.
-By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
-This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
-
-In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula.
-By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR.
-Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.
+Beyond integrating available information for assessing krill population dynamics, we tested several natural scenarios to identify uncertainties associated with estimating key population variables.
+Using a set of eight scenarios, we aimed to diagnose, evaluate, assess performance, and compare the tested models.
+This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.
 
@@
 
-## Study Area in the Antarctic Peninsula
-
-The study area for the assessment of krill populations in the Antarctic Peninsula is divided into strata to account for spatial heterogeneity.
+The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity.
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
@@
 These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
-
-(Figure \@ref(fig:path))
+all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
 
@@
 Where:
-- \(N_t\) is the abundance of krill at time \(t\).
-- \(N_{t-1}\) is the abundance of krill at the previous time step.
-- \(r\) is the intrinsic growth rate of the population.
-- \(M\) is the mortality rate.
-- \(R\) is the recruitment of new individuals into the population.
+
+\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population.
 
@@
 
-The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21).
+The stock assessment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21).
 SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
 SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
-The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* library [@Taylor2019; @Winker2023].
+The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013: @methot2020stock].
+The analysis of results and outputs uses R tools and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].
 
@@
 
+## Environmental factor in modelation
+
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics.
+SS3 addresses these approaches.
+In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
+
+\[
+E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
+\]
+
+\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
+Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
+However, the model does not automatically adjust for this.
+
+Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent.
+These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
+By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
+
+## Predator components in modelation
+
+Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
+Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds.
+Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem.
+By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
+
 ## Scenarios
 
-In Table 1 we have ten scenarios to test different option in modeling about main consideration in assessment of krill population.
+The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill.
+By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
+This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
+
+In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.
 
@@
 |    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
-|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata and Predator and Env data |
-|    s3     | Same "s2" Without S-R relation         |
-|    s4     | Same "s2" Ricker S-R relation estimated          |
-|    s5     | Same "s2" BH S-R relation weak (0.9 steepness)     |
-|    s6     | Same "s2" BH S-R relation strong (0.6 steepness)     |
-|    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
+|    s3     | "s2" Without S-R relation         |
+|    s4     | "s2" Ricker S-R relation estimated          |
+|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
+|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
+|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-```{r message=FALSE, eval=TRUE}
-# install.packages("devtools")
-# devtools::install_github("r4ss/r4ss", ref="development")
-
-# 
-# install.packages("caTools")
-# library("caTools")
-# 
-# install.packages("r4ss")
-library(r4ss)
-library(here)
-#remotes::install_github("PIFSCstockassessments/ss3diags")
-library(ss3diags)
-library(kableExtra)
-library(doParallel) # facilita la ejecución paralela en R
-detectCores()
-registerDoParallel(8)
-library(ggpubr)
-```
-
-```{r fig.height=7}
-SSplotData(base.model01, subplot = 1, 
-           fleetnames = c("Fishery", "Survey", "Predator"),
-           fleetcol = c("blue","green", "red"))
-
-SSplotData(base.model2, 
-           subplot = 1,
-           pheight = 15)
-
-```
-
-Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso
-
-```{r}
-SSplotSelex(base.model2,
-            subplots = 1)
-```
-
-```{r eval=FALSE}
-SSplotBiology(base.model2, 
-              subplots =2, 
-              labels = c("Length (cm)",
-                         "Age (yr)",
-                         "Maturity", 
-                         "Mean weight (kg) in last year",
-                          "Spawning output",
-                         "Length (cm, beginning of the year)", 
-                         "Natural mortality",
-                          "Female weight (kg)", 
-                         "Female length (cm)", 
-                         "Fecundity", 
-                         "Default fecundity label",
-                          "Year", 
-                         "Hermaphroditism transition rate", 
-                         "Fraction females by age at equilibrium"),
- )
-
-```
-
-aporte de las cohortes por año para las capturas.
-```{r eval=FALSE}
-SSplotCohortCatch(base.model2, subplots = 1)
-```
-
-\quad
-
-AJuste de tallas por flota
-```{r eval=FALSE}
-SSplotComps(base.model2, subplots = 1)
-```
-
-Otros plots
-```{r}
-SSplotDynamicB0(base.model2, uncertainty = F)
-#SSplotSPR(base.model3)
-```
-
-```{r eval=FALSE}
-SSplotPars(base.model2)
-
-```
-
-Salida de las biomasas con las dos flotas
-
-```{r}
-SSplotTimeseries(base.model01, subplot = 7)
-```
-Salida de las biomasas con todas las flotas
-
-```{r}
-SSplotTimeseries(base.model2, subplot = 7)
-```
+: Scenarios to modelling dynamics in krill in spatial implicit context
 
@@
 - 4. Likelihood profile
-
-## Residual
-
-```{r eval=FALSE}
-SSplotRunstest(base.model01, 
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model2, 
-               subplots = "len",
-               add=T)
-```
-
-```{r}
-SSplotJABBAres(base.model2,
-               subplots = "cpue",
-               add=T)
-SSplotJABBAres(base.model2,
-               subplots = "len",
-               add=T)
-```
 
@@
 
-```{r eval=FALSE}
-retro(dir=dir7, oldsubdir="", 
-      newsubdir="Retrospective", 
-      years= 0:-5,
-      exe="ss_osx",
-      extras = "-nox", 
-      skipfinished = F)
-```
-
-```{r echo=FALSE}
-#s01
-retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary01 <- SSsummarize(retroModels01)
-# s1
-retroModels1 <- SSgetoutput(dirvec=file.path(dir1,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary1 <- SSsummarize(retroModels1)
-#s2
-retroModels2 <- SSgetoutput(dirvec=file.path(dir2,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary2 <- SSsummarize(retroModels2)
-#s3
-retroModels3 <- SSgetoutput(dirvec=file.path(dir3,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary3 <- SSsummarize(retroModels3)
-#s4
-retroModels4 <- SSgetoutput(dirvec=file.path(dir4,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary4 <- SSsummarize(retroModels4)
-#s5
-retroModels5 <- SSgetoutput(dirvec=file.path(dir5,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary5 <- SSsummarize(retroModels5)
-#s6
-retroModels6 <- SSgetoutput(dirvec=file.path(dir6,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary6 <- SSsummarize(retroModels6)
-#s7
-retroModels7 <- SSgetoutput(dirvec=file.path(dir7,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary7 <- SSsummarize(retroModels7)
-
-```
-
-```{r echo FALSE}
-#save(retroSummary, retroModels, file="retro5.Rdata")
-
-retro01 <- SSplotRetro(retroSummary01,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro1 <- SSplotRetro(retroSummary1,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro2 <- SSplotRetro(retroSummary2,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro3 <- SSplotRetro(retroSummary3,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro4 <- SSplotRetro(retroSummary4,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro5 <- SSplotRetro(retroSummary5,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
-retro6 <- SSplotRetro(retroSummary6,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
-retro7 <- SSplotRetro(retroSummary7,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
-```
-
-```{r echo=FALSE}
-ggarrange(retro01,
-          retro1,
-          retro2,
-          retro3,
-          retro4,
-          retro5,
-          retro6,
-          retro7, ncol = 4, nrow = 4)
-```
-
 ## Hindcast Cross-Validation and prediction skill
@@
 
-```{r}
-hci = SSplotHCxval(retroSummary2, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-```
 ## Kobe (status)
-
-```{r}
-mvln = SSdeltaMVLN(base.model01,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-mvln = SSdeltaMVLN(base.model2,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-```
-
-```{r echo=FALSE}
-# model01
-tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
-tablebias02 <- SShcbias(retroSummary01,quant="F",verbose=F)
-
-kbl(tablebias01, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s01")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias02, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s01")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-# model2
-tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
-tablebias07 <- SShcbias(retroSummary7,quant="F",verbose=F)
-
-kbl(tablebias7, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s2")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias7, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s2")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-```
 
@@
 
-```{r eval=FALSE, echo = FALSE}
-# 1. Identificar el directorio donde se encuentra el modelo base ----
-dirname.model.run <- here("s2")
-# 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
-dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
-dir.create(path=dirname.R0.profile, showWarnings = TRUE, recursive = TRUE)
-# 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
-plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
-dir.create(path=plotdir, showWarnings = TRUE, recursive = TRUE)
-# 4. Crear un subdirectorio llamado "simple" ----
-reference.dir <- paste0(dirname.R0.profile,'/simple') 
-dir.create(path=reference.dir, showWarnings = TRUE, recursive = TRUE)
-# 5. Copiar el resultado del modelo base completo en este directorio ----
-file.copy(Sys.glob(paste(dirname.model.run, "*.*", sep="/"),
-                   dirmark = FALSE),
-                    reference.dir)
-# 6. Leer la salida del modelo base ----
-Base <- SS_output(dir=reference.dir,covar=T)
-# 7. Copiar los archivos necesarios de "simple" al directorio "Perfil_Verosimilitud" ----
-copy_SS_inputs(dir.old = reference.dir, 
-               dir.new =  dirname.R0.profile,
-               copy_exe = TRUE,
-               verbose = FALSE)
-# 8. Leer los archivos del modelo ----
-inputs <- r4ss::SS_read(dir = dirname.R0.profile)
-# 9. Editar el archivo control la fase de estimación recdev ----
-inputs$ctl$recdev_phase <- 1
-# 10. Editar el archivo starter para leer los valores de inicio ----
-inputs$start$init_values_src <- 0
-# 11. Vector de valores para el perfil ----
-R0.vec <- seq(15,20,0.5)  
-Nprof.R0 <- length(R0.vec)
-# 12. Cambiar el nombre del archivo control en el archivo starter.ss ----
-inputs$start$ctlfile <- "control_modified.ss" 
-# 13. Incluir prior_like para parámetros no estimados ----
-inputs$start$prior_like <- 1                                 
-# 14. Escribir los modelos modificados ----
-r4ss::SS_write(inputs, dir = dirname.R0.profile, overwrite = TRUE)
-# 15. Ejecutar la función profile() ----
-?SS_profile()
-profile <- profile(dir=dirname.R0.profile, # directory
-                      exe="ss_osx",
-                      oldctlfile ="control.ss",
-                      newctlfile="control_modified.ss",
-                      string="SR_LN(R0)",
-                      profilevec=R0.vec)
-# 16. Leer los archivos de salida ----
-# (con nombres como Report1.sso, Report2.sso, etc.)
-prof.R0.models <- SSgetoutput(dirvec=dirname.R0.profile, 
-                              keyvec=1:Nprof.R0, 
-                              getcovar = FALSE) 
-# 17. Resumir las salidas con la función SSsummarize()  ----
-prof.R0.summary <- SSsummarize(prof.R0.models)
-# 18. Identificar los componentes de Verosimilitud ----
-mainlike_components         <- c('TOTAL',
-                                 "Survey", 
-                                 'Length_comp',
-                                 "Age_comp",
-                                 'Size_at_age',
-                                 'Recruitment') 
-mainlike_components_labels  <- c('Total likelihood',
-                                 'Index likelihood',
-                                 'Length likelihood',
-                                 "Age likelihood",
-                                 'Size_at_age likelihood',
-                                 'Recruitment likelihood') 
-# 19. Funciones para generar plots de perfil de verosimilitud ----
-### SSplotProfile() ----
-png(file.path(plotdir,"R0_profile_plot.png"),
-    width=7,
-    height=4.5,
-    res=300,
-    units='in')
-par(mar=c(5,4,1,1))
-SSplotProfile(prof.R0.summary,           # summary object
-              profile.string = "R0",     # substring of profile parameter
-              profile.label=expression(log(italic(R)[0])), 
-              ymax=150,minfraction = 0.001,
-              pheight=4.5, 
-              print=FALSE, 
-              plotdir=plotdir, 
-              components = mainlike_components, 
-              component.labels = mainlike_components_labels,
-              add_cutoff = TRUE,
-              cutoff_prob = 0.95)
-
-Baseval <- round(Base$parameters$Value[grep("R0",Base$parameters$Label)],2)
-Baselab <- paste(Baseval,sep="")
-axis(1,at=Baseval,label=Baselab)
-abline(v = Baseval, lty=2)
-dev.off()
-
-### SSplotComparisons() ----
-# Comparación de series de tiempo 
-labs <- paste("SR_Ln(R0) = ",R0.vec)
-labs[which(round(R0.vec,2)==Baseval)] <- paste("SR_Ln(R0) = ",
-                                               Baseval,"(Base model)")
-
-SSplotComparisons(prof.R0.summary,
-                  legendlabels=labs,
-                  pheight=4.5,png=TRUE,
-                  plotdir=plotdir,
-                  legendloc='bottomleft')
-
-### PinerPlot() ----
-#### R0_profile_plot_Length_like ----
-png(file.path(plotdir,"R0_profile_plot_Length_like.png"),
-    width=7,
-    height=4.5,
-    res=300,
-    units='in')
-par(mar=c(5,4,1,1))
-PinerPlot(prof.R0.summary, 
-          profile.string = "R0", 
-          component = "Length_like",
-          main = "Changes in length-composition likelihoods by fleet",
-          add_cutoff = TRUE,
-          cutoff_prob = 0.95)
-Baseval <- round(Base$parameters$Value[grep("SR_LN",
-                                      Base$parameters$Label)],2)
-Baselab <- paste(Baseval,sep="")
-axis(1,at=Baseval,
-     label=Baselab)
-abline(v = Baseval, lty=2)
-dev.off()
-#### R0_profile_plot_Survey_like ----
-png(file.path(plotdir,"R0_profile_plot_Survey_like.png"),
-    width=7,
-    height=4.5,
-    res=300,
-    units='in')
-par(mar=c(5,4,1,1))
-PinerPlot(prof.R0.summary, 
-          profile.string = "R0", 
-          component = "Surv_like",
-          main = "Changes in Index likelihoods by fleet",
-          add_cutoff = TRUE,
-          cutoff_prob = 0.95, legendloc="topleft")
-Baseval <- round(Base$parameters$Value[grep("SR_LN",
-                                            Base$parameters$Label)],2)
-Baselab <- paste(Baseval,sep="")
-axis(1,at=Baseval,label=Baselab)
-abline(v = Baseval, lty=2)
-dev.off()
-```
-
 ## Comparision outputs betwwen scenarios
-
-```{r echo=FALSE}
-#PLOT labels, name of each model run
-legend.labels <- c( "s01",
-                  "s1",
-                 "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7")
-
-#read in all model runs
-#note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(dir01,
-                                  dir1,
-                                  dir2,
-                                  dir3,
-                                  dir4,
-                                  dir5,
-                                  dir6,
-                                  dir7),
-                       getcovar = F)
-
-#create summary of model runs from list above
-summaryoutput <- SSsummarize(biglist)
-
-SSplotComparisons(summaryoutput,
-                  legendlabels = c("s01",
-                                   "s1",
-                                   "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
-                 labels = c("Year", 
-                            "Spawning biomass (t)",
-                            "Relative spawning biomass", 
-                            "Age-0 recruits (1,000s)",
-                            "Recruitment deviations", 
-                            "Index", "Log index", 
-                            "1 - SPR", 
-                            "Density",
-                            "Management target", 
-                            "Minimum stock size threshold",
-                            "Spawning output",
-                            "Harvest rate"))
-
-comp <- SSplotComparisons(summaryoutput,
-                  legendlabels = c("s01",
-                                   "s1",
-                                   "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
-                 subplots = c(2, 16))
-comp
-```
-
-```{r}
-comtable <- SStableComparisons(summaryoutput,
-                   likenames = c("TOTAL", 
-                                 "Survey", 
-                                 "Length_comp",
-                                 "Age_comp", 
-                                 "priors",
-                                 "Size_at_age"), 
-                   names = c("Recr_Virgin",
-                             "R0", 
-                             "steep",
-                             "NatM",
-                             "L_at_Amax", 
-                             "VonBert_K", 
-                             "SSB_Virg", 
-                             "Bratio_2017",
-                             "SPRratio_2016"),
-                   digits = NULL,
-                   modelnames = c("s01", "s1","s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
-                   csv = FALSE,
-                   csvdir =~"/DOCAS/SA_Krill",
-                   csvfile = "parameter_comparison_table.csv",
-                 verbose = TRUE,
-                   mcmc = FALSE)
-kbl(comtable, booktabs = T,format = "latex",
-    caption = "Comparacion likelihood y parámetros s01, s1, s2, s3, s4, s5, s6 y s7")  %>% 
-    kable_styling(latex_options = "scale_down")
-
-```
 
@@
 # DISCUSION
+
+ Some authors have idicated spatial heterogeneity and identified different zones, such as areas with higher or lower concentrations of juveniles or regions more affected by the environment, which structure the population differently even inside WAP (REF*).
+
+In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula.
+By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR.
+Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.
 
@@
 ```
+\end{landscape}
 
-Plot data
+```{r concem, out.width='80%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
+knitr::include_graphics('Figs/conceptual.png')
+```
 
-```{r datas2, out.width='60%', fig.cap="Data series using in "s2" scenario"}
+```{r datas2, out.width='60%', fig.cap="Data series using in s2 scenario"}
 knitr::include_graphics('s2/plots/data_plot.png')
@@
 
+\begin{landscape}
+
+```{r comparision, fig.cap="Data series using in s2 scenario"}
+par(mfrow=c(2,2))
+knitr::include_graphics(c('compare8_Fvalue_uncertainty.png',
+                          'compare2_spawnbio_uncertainty.png',
+                          'compare4_Bratio_uncertainty.png',
+                          'compare8_Fvalue_uncertainty.png'))
+```
 \end{landscape}
@@
 \newpage
+
 # REFERENCES

### 2024-05-28 index.Rmd
improve Intro

@@
 In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time.
-This model incorporates key biological parameters such as growth rates, mortality rates, recruitment, and spawning biomass.
+This model incorporates key biological, environmental and predator data sources.
 The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.
@@
 # devtools::install_github("r4ss/r4ss", ref="development")
-
-# 
 # install.packages("caTools")
 # library("caTools")
-# 
 # install.packages("r4ss")
@@
 library(ggpubr)
+library(tibble)
 ```
@@
 - 4. Likelihood profile
+
+all this framework try to follow recommendations of @Carvalho2021b
 
@@
 ## Verosimilitud
+
+```{r}
+like01 <- base.model01$likelihoods_used
+like01$model <- rep("s01", nrow(like01))
+like01 <- rownames_to_column(like01, var = "Description")
+like1 <- base.model1$likelihoods_used
+like1$model <- rep("s1", nrow(like1))
+like1 <- rownames_to_column(like1, var = "Description")
+like2 <- base.model1$likelihoods_used
+like2$model <- rep("s2", nrow(like2))
+like2 <- rownames_to_column(like2, var = "Description")
+like3 <- base.model3$likelihoods_used
+like3$model <- rep("s3", nrow(like3))
+like3 <- rownames_to_column(like3, var = "Description")
+like4 <- base.model4$likelihoods_used
+like4$model <- rep("s4", nrow(like4))
+like4 <- rownames_to_column(like4, var = "Description")
+like5 <- base.model5$likelihoods_used
+like5$model <- rep("s5", nrow(like5))
+like5 <- rownames_to_column(like5, var = "Description")
+like6 <- base.model6$likelihoods_used
+like6$model <- rep("s6", nrow(like6))
+like6 <- rownames_to_column(like6, var = "Description")
+like7 <- base.model7$likelihoods_used
+like7$model <- rep("s7", nrow(like7))
+like7 <- rownames_to_column(like7, var = "Description")
+totalike <- rbind(like01,
+                  like1,
+                   like2,
+                   like3,
+                   like4,
+                   like5,
+                   like6,
+                   like7)
+
+```
+
+```{r}
+ggplot(totalike)+
+  geom_bar( aes(y = reorder(Description, values), x =values),
+            stat = "identity", position = "dodge") +
+  theme_few() +
+  facet_wrap(.~model, 
+             ncol=4)+
+  labs(y="",
+       x="Likelihood")+
+  xlim(0, 1000)+
+  theme(axis.text.x = element_text(angle = 90, 
+                                   hjust = 1)) 
+```
 

### 2024-05-29 SS3_Krill_481.Rmd
commit adding GW in CReDIT

@@
 library(tibble)
+library(readxl)
+library(openxlsx)
 ```
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity.
+The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.
+Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in environmental factors.
+Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in structure population.
 
-The strata in the Antarctic Peninsula may include different regions such as coastal areas, offshore waters, and areas influenced by major oceanographic features like currents and frontal systems.
-Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability (Figure \@ref(fig:mapa))
+## Data Sources
 
-## Data Sources and Recolection
-
-Stock Synthesis relies on multiple sources of data to inform the population dynamics model. These data sources may include:
-
-1. **Fisheries Data:** Information on krill catches and fishing effort from commercial fisheries operating in the Antarctic Peninsula region.
-
-2.
-**Population Monitoring Data:** Data collected from scientific research cruises and monitoring programs specifically designed to assess krill populations.
+Multiple sources of data to inform the population dynamics model are included in this analysis.
+These data sources include: 1) Data from the monitoring program of the krill fishery were used, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+This database has information about vessel, nationality, georeferenced, among others.
+Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations.
 This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.
-
-3.
-**Environmental Data:** Environmental data such as sea surface temperature, chlorophyll concentration, and sea ice extent are also incorporated into the assessment.
-These data help to understand the influence of environmental factors on krill populations and their distribution.
-
-4.
-**Predator information:** MAPPPD.
-Predator “fleet” to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.
-
-In addition to existing data sources, researchers may submit requests to the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) Secretariat for access to additional data relevant to krill population assessment.
-These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
-all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+All this data was proceses in @Dornam2021; 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
+Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
+Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
+This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.
+Data processing involved sophisticated techniques, including data cleaning, spatial manipulation, rasterization, and projection transformations.
+These processes were facilitated by the R libraries *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], ensuring high-quality and accurate environmental datasets.
+To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST).
+The model included spatial component (cell) as random effect (formula: ~1 | cell).
+Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020].
+All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator **"fleet"** to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.
 
@@
 
+These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
+all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+
 # RESULTS
@@
 ```
+
 \begin{landscape}
@@
 ```
+
 \end{landscape}
 
-```{r concem, out.width='80%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
-knitr::include_graphics('Figs/conceptual.png')
+```{r parainit, fig.cap="Initial Parameters to modelling in stock process in krill in WAP"}
+parInit <- read_excel("DataKrill.xlsx", 
+    sheet = "parameters")
+parInit %>%
+  kbl(booktabs = TRUE,
+      format = "latex",
+      position="ht!",
+    caption = "Input parameters for the initial SS3 model of krill.
+Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
+If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
+  kable_paper("hover", 
+              full_width = F)%>%
+  kable_styling(latex_options = c("striped"),
+                full_width = FALSE,
+                font_size=9)%>% 
+  pack_rows(index = c("Natural Mortality" = 1,
+                        "Growth"= 5,
+                        "Length-Weigth Relation" = 2,
+                        "Maturity"=2,
+                        "Stock-Recruit Relation"=5,
+                        "Catchability"=2,
+                        "Selectivity"=4))
 ```
 
-```{r datas2, out.width='60%', fig.cap="Data series using in s2 scenario"}
+```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
+knitr::include_graphics('Figs/conceptual.jpeg')
+```
+
+```{r datas2, out.width='70%', fig.cap="Data series using in s2 scenario"}
 knitr::include_graphics('s2/plots/data_plot.png')
 ```
+
+```{r outs2, fig.cap="Main variables outputs from stock asssessment krill in WAP"}
+out_s2 <- read_excel("DataKrill.xlsx", 
+    sheet = "variable_s2")
+out_s2 %>%
+  kbl(booktabs = TRUE,
+      format = "latex",
+    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9)#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+\newpage
 
@@
 ```
+
 \end{landscape}

### 2024-05-29 index.Rmd
commit adding GW in CReDIT

@@
 rm(list = ls())
+set.seed(999)
 knitr::opts_chunk$set(echo = FALSE,
@@
 # Lo mapas se hacen mas rapido
+```
+
+```{r message=FALSE, eval=TRUE}
+# install.packages("devtools")
+# devtools::install_github("r4ss/r4ss", ref="development")
+# install.packages("caTools")
+# library("caTools")
+# install.packages("r4ss")
+library(r4ss)
+library(here)
+#remotes::install_github("PIFSCstockassessments/ss3diags")
+library(ss3diags)
+library(kableExtra)
+library(doParallel) # facilita la ejecución paralela en R
+detectCores()
+registerDoParallel(8)
+library(ggpubr)
+library(tibble)
+library(openxlsx)
+```
+
+```{r message=FALSE, warning=FALSE}
+dir01 <- here("s01") # agreggate data (no spatial diferences)
+dir1<-here("s1") # Data strata flishery
+dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
+dir3<-here("s3") # whitout S-R
+dir4<-here("s4") # 
+dir5<-here("s5") # whitout S-R
+dir6<-here("s6") # whitout S-R
+dir7<-here("s7") # whitout S-R
 ```
@@
 
+## Parametres
+
+read files
+
+```{r}
+# leo archivos para plotear y hacer tablas
+start2 <- SS_readstarter(file = file.path(dir2,
+                                          "starter.ss"),
+                              verbose = FALSE)
+# note the data and control file names can vary, so are determined from the 
+# starter file.
+dat2 <- SS_readdat(file = file.path(dir2, start2$datfile),
+                        verbose = FALSE)
+# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
+# assumes the correct data structure
+ctl2 <-  r4ss::SS_readctl(file = file.path(dir2,
+                                    start2$ctlfil),
+                        verbose = FALSE,
+                        use_datlist = TRUE, 
+                   datlist = dat2)
+fore2 <- r4ss::SS_readforecast(file = file.path(dir2, 
+                                                "forecast.ss"),
+                              verbose = FALSE)
+# can also read in wtatage.ss for an empirical wt at age model using
+# r4ss::SS_readwtatage()
+```
+
+```{r}
+parbio<-ctl2$MG_parms[1:10,c(1:3,7)]
+ row.names( parbio)<-c("Nat M",
+                       "Lmin", 
+                       "Lmax",
+                       "VonBert K",
+                       "CV young",
+                       "CV old", 
+                       "Wt a", 
+                       "Wt b",
+                       "L50%", 
+                       "Mat slope")
+
+ SRpar<-ctl2$SR_parms[1:5,c(1:3,7)]
+ Qpar<-ctl2$Q_parms[1:2,c(1:3,7)]
+ Selpar<-ctl2$size_selex_parms[1:22,c(1:3,7)]
+ parInit<-rbind(parbio,SRpar,Qpar,Selpar)
+
+wb <- createWorkbook()
+
+addWorksheet(wb, "parameters")
+writeData(wb, "parameters", parInit)
+
+# Guardar el workbook
+saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+
+```
+
+```{r}
+parInit %>%
+  kbl(booktabs = T,
+      format = "latex",
+      position="ht!",
+    caption = "\\label{Tab1}Input parameters for the initial SS3 model of krill.
+Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
+If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
+  kable_paper("hover", 
+              full_width = F)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9)%>% 
+  pack_rows(index = c("Mortalidad natural" = 1,
+                        "Crecimiento"= 5,
+                        "Relación longitud-peso" = 2,
+                        "Ojiva de madurez"=2,
+                        "Relación stock-recluta"=5,
+                        "Capturabilidad"=2,
+                        "Selectividad"=4))
+
+```
 ## Scenarios
@@
 ## Run Models
-
-```{r message=FALSE, eval=TRUE}
-# install.packages("devtools")
-# devtools::install_github("r4ss/r4ss", ref="development")
-# install.packages("caTools")
-# library("caTools")
-# install.packages("r4ss")
-library(r4ss)
-library(here)
-#remotes::install_github("PIFSCstockassessments/ss3diags")
-library(ss3diags)
-library(kableExtra)
-library(doParallel) # facilita la ejecución paralela en R
-detectCores()
-registerDoParallel(8)
-library(ggpubr)
-library(tibble)
-```
-
-```{r message=FALSE, warning=FALSE}
-dir01 <- here("s01") # agreggate data (no spatial diferences)
-dir1<-here("s1") # Data strata flishery
-dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
-dir3<-here("s3") # whitout S-R
-dir4<-here("s4") # 
-dir5<-here("s5") # whitout S-R
-dir6<-here("s6") # whitout S-R
-dir7<-here("s7") # whitout S-R
-```
 
@@
 #system('./ss_osx') # to mac
-# r4ss::run(
-#   dir = dir2,
-#   exe = "ss_osx",
-#   skipfinished = FALSE, # TRUE will skip running if Report.sso present
-#   show_in_console = TRUE # change to true to watch the output go past
-# )
+r4ss::run(
+  dir = dir2,
+  exe = "ss_osx",
+  skipfinished = FALSE, # TRUE will skip running if Report.sso present
+  show_in_console = TRUE # change to true to watch the output go past
+)
 ```
@@
 
+## Outputs Model Base
+
+```{r}
+outps2 <- base.model2$timeseries[1:53, 2:8]
+
+addWorksheet(wb, "variable_s2")
+writeData(wb, "variable_s2", outps2)
+
+# Guardar el workbook
+saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+```
+
 ## Comparision outputs betwwen scenarios
@@
                             "Spawning output",
-                            "Harvest rate"))
+                            "Harvest rate"),
+                 plotdir = "~/Figs")
 
@@
                  "s7"),
-                 subplots = c(2, 16))
+                 subplots = c(2, 16),
+                 plotdir = "~/Figs")
 comp

### 2024-05-29 SS3_Krill_481.Rmd
adding text and refs

@@
 
-## Statistical model (SS3)
+## Statistical model Stock Synthesis (SS3)
 
-Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations, including krill (Euphausia superba) in the Antarctic Peninsula region.
+Like most invertebrates, krill is a species for which aging is complex.
+Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
+Despite this, integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
+
+Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations.
 The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP.
@@
 
+Development of an integrated stock assessment like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
+Integrated models are not new to krill in SO.
+For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
+In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modelling strategies currently in use to manage krill"* and he stated that the an integrated modelling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
+
+@Kinzey2015a, @Kinzey2011 and @Kinzey2019a used integrated models with fishery and survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
+Additionally, @Wang2021 made significant contributions to integrated statistical models for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
+Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) and El Niño.
+
+Wang..
+Lo que es nuevo aca es
+
+**Paragraph 3:**  
+From an ecological standpoint, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
+In terms of the krill fishery, this area includes a significant fishing ground, specifically subarea 48.1.
+Notably, recent years have seen a higher concentration of krill catches in this subarea, both temporally and spatially.
+
+**Paragraph 4:**  
+Model estimation concludes when a convergent Hessian matrix is achieved, and all data sources are fitted using maximum likelihood estimates from the assessment method.
+
+**Paragraph 5:**  
+The independent survey data, which include survey abundance and length composition data, were collected by the United States Antarctic Living Marine Resources (AMLR) Program of the National Marine Fisheries Service, Southwest Fisheries Science Center.
+These surveys, conducted during the austral summer from 1992 to 2011, utilized zooplankton nets to gather sample data.
+
+**Paragraph 6:**  
+The fishery-dependent data, encompassing length composition data and catch biomass, are derived from CCAMLR's public database.
+Catch biomass data are available from 1992 to 2011, while length composition data are available for the same period.
+
+**Paragraph 7:**  
+The estimated recruitment of krill exhibits cyclical variations with an average period of 3–4 years.
+This cyclical variation is influenced by the physical environment of the polar ecosystem (Kinzey et al., 2019) and the behavior of its predators (Fraser et al., 1992; Fraser and Trivelpiece, 1995; Reid et al., 1999, 2002).
+
+**Paragraph 8:**  
+When evaluating any fishery resource, the chosen evaluation method should ensure that the results meet the reference data requirements for production management and optimize the use of existing data (Punt et al., 2013).
+
  Some authors have idicated spatial heterogeneity and identified different zones, such as areas with higher or lower concentrations of juveniles or regions more affected by the environment, which structure the population differently even inside WAP (REF*).

### 2024-05-30 SS3_Krill_481.Rmd
commit 31 may

@@
 ---
-title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
+title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
 subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
@@
 This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.
-Data processing involved sophisticated techniques, including data cleaning, spatial manipulation, rasterization, and projection transformations.
-These processes were facilitated by the R libraries *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], ensuring high-quality and accurate environmental datasets.
 To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST).
@@
 \(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population.
+This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
 
-This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
+Life history arameters taken as priors taken from @Smith2023a and @Maschette2020
 
@@
 
-Like most invertebrates, krill is a species for which aging is complex.
+Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
 Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
@@
 Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
-Antarctic krill are a vital prey species for a wide range of predators, including whales, seals, and seabirds.
-Changes in predator populations and foraging behavior can have profound effects on krill abundance and distribution, with cascading impacts throughout the Antarctic ecosystem.
 By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
+In this analysis we incorporated penguin population index as predator *"fleets"*.
+This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock].
 
@@
 
-The methodology involves comparing the statistical performance of the model concerning spatial heterogeneity of krill.
-By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
-This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
-
+To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
+The methodology involves comparing the statistical performance of the model  regardng this configurations.
 In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.
@@
 |:---------:|:-------------------------------------------------|
-|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
+|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
 |    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
 |    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
-|    s3     | "s2" Without S-R relation         |
+|    s3     | "s2" without S-R relation         |
 |    s4     | "s2" Ricker S-R relation estimated          |
@@
 |    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |
 
@@
 
+models S01, S1, S2, and S5 exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
+Models S3 and S4 show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
+Model S6 shows moderate autocorrelations, indicating some predictability based on recent past values.
+These findings help understand the temporal dynamics of krill recruitment and inform future modeling efforts.
+(fig autocorrelation)
 ## Kobe (status)
@@
 
-@Kinzey2015a, @Kinzey2011 and @Kinzey2019a used integrated models with fishery and survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
-Additionally, @Wang2021 made significant contributions to integrated statistical models for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
-Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) and El Niño.
+@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
+@Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula.
+Although these models had convergence issues, they represented significant advances in this type of framework.
+Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
+Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO).
 
-Wang..
-Lo que es nuevo aca es
+In that sense, the work carried out here differs substantially from the previously written integrated models since it incorporates aspects that have to do with ecosystem indicators such as environmental variables and predators.
+On the other hand, it considers the spatial component within the modeling scheme implicitly and finally evaluates the impact of the different sets of life history parameters that have been used and those calculated by Mardones et al (in press)
 
-**Paragraph 3:**  
-From an ecological standpoint, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
-In terms of the krill fishery, this area includes a significant fishing ground, specifically subarea 48.1.
-Notably, recent years have seen a higher concentration of krill catches in this subarea, both temporally and spatially.
+From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
+The intention to incorporate elements of predation in a stock is necessary to improve the modelling, because did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
+There is many studies that confirm the influence of predation on grill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
+For this, we consider penguin columns as one of the most powerful elements of natural predation on krill because have direct correlations between this type of taxonomic group and krill [@Kruger2021].
 
@@
  Some authors have idicated spatial heterogeneity and identified different zones, such as areas with higher or lower concentrations of juveniles or regions more affected by the environment, which structure the population differently even inside WAP (REF*).
+
+ By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
+This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
 

### 2024-05-30 index.Rmd
commit 31 may

@@
 ---
-title: "Implicit spatial model to krill Dynamic Population in Antarctic Peninsula, 48-1 SubArea. SS3 aplications"
+title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
 subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
@@
 library(openxlsx)
+library(ggthemes)
+library(forecast)
 ```
@@
 dir7<-here("s7") # whitout S-R
+dir8<-here("s8") # S2 set parametres EMM-2024/23 (Mardones)
 ```
@@
 |:---------:|:-------------------------------------------------|
-|    s01     | Fishery, Predator, Survey, Environmental data agreggate (Whole 48.1) |
+|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
 |    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
-|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata and Predator and Env data |
-|    s3     | Same "s2" Without S-R relation         |
-|    s4     | Same "s2" Ricker S-R relation estimated          |
-|    s5     | Same "s2" BH S-R relation weak (0.9 steepness)     |
-|    s6     | Same "s2" BH S-R relation strong (0.6 steepness)     |
-|    s7     | Same "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
+|    s3     | "s2" without S-R relation         |
+|    s4     | "s2" Ricker S-R relation estimated          |
+|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
+|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
+|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
+|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |
 
@@
                  "s6",
-                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
+                 "s7",
+                 "s8")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
 
@@
 }
+```
+
+```{r eval=FALSE, message=F, include=FALSE}
 #OR with rby separate
@@
 r4ss::run(
-  dir = dir2,
+  dir = dir01,
   exe = "ss_osx",
@@
                         forecast=T)
+
+#s7
+base.model8 <- SS_output(dir=dir8,
+                        covar=T,
+                        forecast=T)
 ```
@@
 ```{r eval=FALSE, message=F, include=FALSE}
-SS_plots(base.model2, 
+SS_plots(base.model3, 
          uncertainty = TRUE,
@@
 ```{r}
-SSplotTimeseries(base.model01, subplot = 7)
+SSplotTimeseries(base.model2, subplot = 7)
 ```
@@
 ```{r eval=FALSE}
-SSplotRunstest(base.model01, 
+SSplotRunstest(base.model8, 
                subplots = "cpue",
@@
 ```{r}
-SSplotJABBAres(base.model2,
-               subplots = "cpue",
+SSplotJABBAres(base.model01,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model1,
+               subplots = "len",
                add=T)
@@
                add=T)
+SSplotJABBAres(base.model3,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model4,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model5,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model6,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model7,
+               subplots = "len",
+               add=T)
+SSplotJABBAres(base.model7,
+               subplots = "len",
+               add=T)
+
+SSplotMnwt(base.model01, subplots = 2)
 ```
@@
 ```{r}
-mvln = SSdeltaMVLN(base.model01,
-                   Fref = "MSY", 
+mvln = SSdeltaMVLN(base.model8,
+                   Fref = "Btgt", 
                    plot = TRUE,
@@
                    virgin = TRUE,
-                   mc=100)
+                   mc=1000)
 
 mvln = SSdeltaMVLN(base.model2,
-                   Fref = "MSY", 
+                   Fref = "Btgt", 
                    plot = TRUE,
@@
                    virgin = TRUE,
-                   mc=100)
+                   mc=1000)
 ```
@@
 like7 <- rownames_to_column(like7, var = "Description")
+like8 <- base.model8$likelihoods_used
+like8$model <- rep("s8", nrow(like8))
+like8 <- rownames_to_column(like8, var = "Description")
 totalike <- rbind(like01,
@@
                    like6,
-                   like7)
+                   like7,
+                  like8)
 
@@
   facet_wrap(.~model, 
-             ncol=4)+
+             ncol=3)+
   labs(y="",
        x="Likelihood")+
-  xlim(0, 1000)+
   theme(axis.text.x = element_text(angle = 90, 
@@
                  "s6",
-                 "s7")
+                 "s7",
+                 "s8")
 
@@
                                   dir6,
-                                  dir7),
+                                  dir7,
+                                  dir8),
                        getcovar = F)
@@
 SSplotComparisons(summaryoutput,
-                  legendlabels = c("s01",
+                  legendlabels = c("No Spatial",
                                    "s1",
@@
                  "s6",
-                 "s7"),
+                 "s7",
+                 "s8"),
                  labels = c("Year", 
@@
 
-comp <- SSplotComparisons(summaryoutput,
-                  legendlabels = c("s01",
-                                   "s1",
-                                   "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
-                 subplots = c(2, 16),
-                 plotdir = "~/Figs")
-comp
 ```
@@
 ```
+comparision between select models
+
+```{r echo=FALSE}
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir8,
+
+                                  dir2,
+                                  dir3),
+                       getcovar = F)
+
+#create summary of model runs from list above
+summaryoutput <- SSsummarize(biglist)
+
+SSplotComparisons(summaryoutput,
+                  legendlabels = c(
+                                   "s8",
+
+                 "s2",
+                 "s3"),
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"),
+                 plotdir = "~/Figs")
+
+comp2_8 <- SSplotComparisons(summaryoutput,
+                  legendlabels = c(
+                                   "s8",
+
+                 "s2",
+                 "s3"),
+                 subplots = c(2, 16),
+                 plotdir = "~/Figs")
+comp2_8
+```
+## Comparsion in sd long term time series
+
+```{r}
+sblt01 <- base.model01$timeseries$Bio_all 
+sblt1 <- base.model1$timeseries$Bio_all 
+sblt2 <- base.model2$timeseries$Bio_all 
+sblt3 <- base.model3$timeseries$Bio_all 
+sblt4 <- base.model4$timeseries$Bio_all 
+sblt5 <- base.model5$timeseries$Bio_all 
+sblt6 <- base.model6$timeseries$Bio_all 
+sblt7 <- base.model7$timeseries$Bio_all 
+sblt8 <- base.model8$timeseries$Bio_all
+errt <- as.data.frame(cbind(sblt01,
+              sblt1,
+              sblt2,
+              sblt3,
+              sblt4,
+              sblt5,
+              sblt6,
+              sblt7,
+              sblt8))
+err_long <- errt %>%
+  pivot_longer(cols = everything(),
+               names_to = "variable",
+               values_to = "value")
+
+ggplot(err_long, aes(x = variable, y = value)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.2, alpha = 0.3) +
+  labs(title = "",
+       x = "",
+       y = "Spawning B.(mt)") +
+  theme_few() +
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1))
+base.model2$
+
+```
+
+## Autocorrelation in Recruit and Biomas
+
+```{r}
+recs01 <- base.model01$recruit$pred_recr
+recs1 <- base.model1$recruit$pred_recr
+recs2 <- base.model2$recruit$pred_recr
+recs3 <- base.model3$recruit$pred_recr
+recs4 <- base.model4$recruit$pred_recr
+recs5 <- base.model5$recruit$pred_recr
+recs6 <- base.model6$recruit$pred_recr
+recs7 <- base.model7$recruit$pred_recr
+recs8 <- base.model8$recruit$pred_recr
+
+p01 <- ggAcf(base.model01$recruit$pred_recr) +
+  ggtitle("S01") +
+  theme_few()
+p1 <- ggAcf(base.model1$recruit$pred_recr) +
+  ggtitle("S1") +
+  theme_few()
+p2 <- ggAcf(base.model2$recruit$pred_recr) +
+  ggtitle("S2") +
+  theme_few()
+p3 <- ggAcf(base.model3$recruit$pred_recr) +
+  ggtitle("S3") +
+  theme_few()
+p4 <- ggAcf(base.model4$recruit$pred_recr) +
+  ggtitle("S4") +
+  theme_few()
+p5 <- ggAcf(base.model5$recruit$pred_recr) +
+  ggtitle("S5") +
+  theme_few()
+p6 <- ggAcf(base.model6$recruit$pred_recr) +
+  ggtitle("S6") +
+  theme_few()
+p7 <- ggAcf(base.model7$recruit$pred_recr) +
+  ggtitle("S6") +
+  theme_few()
+p8 <- ggAcf(base.model8$recruit$pred_recr) +
+  ggtitle("S6") +
+  theme_few()
+
+ggarrange(p01,p1,
+          p2,
+          p3,
+          p4,
+          p5,
+          p6,
+          p7,
+          p8, ncol=3, nrow = 3,
+          common.legend = T)
+```
 

### 2024-05-31 index.Rmd
adding html but not in page

@@
 ```{r setup1, echo=FALSE}
-rm(list = ls())
-set.seed(999)
+#rm(list = ls())
+#set.seed(999)
 knitr::opts_chunk$set(echo = FALSE,
@@
 library(forecast)
+library(tidyr)
+library(mixR)
 ```
@@
 dir7<-here("s7") # whitout S-R
-dir8<-here("s8") # S2 set parametres EMM-2024/23 (Mardones)
+dir8<-here("s8")
+Figs <- here("Figs")# S2 set parametres EMM-2024/23 (Mardones)
 ```
@@
 
-In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time.
-This model incorporates key biological, environmental and predator data sources.
-The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.
+In a simple way, the core of Stock Synthesis is its population dynamics
+model, which represents the dynamics of krill populations over time.
+This model incorporates key biological, environmental and predator data
+sources. The model is typically formulated using mathematical equations
+that describe how these parameters interact to determine the abundance
+and distribution of krill in the study area.
 
@@
 
-Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations, including krill (Euphausia superba) in the Antarctic Peninsula region.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP.
-
-The stock assesment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21).
-SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
-SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
-The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-The analysis of results and outputs uses R tools and the graphical interface of the *r4ss* and *ss3diags* library [@Taylor2019; @Winker2023].
-
-By integrating data from multiple sources and considering spatial heterogeneity, the assessment methodology using Stock Synthesis v.3.30.21 provides a robust framework for evaluating the status of krill populations in the Antarctic Peninsula region.
-This information is essential for supporting management decisions aimed at ensuring the sustainable use of krill resources in this ecologically sensitive area.
+Stock Synthesis v.3.30.21 is a widely used software tool for assessing
+fish and invertebrate populations, including krill (Euphausia superba)
+in the Antarctic Peninsula region. The methodology employed by Stock
+Synthesis involves a comprehensive and integrated approach, utilizing
+various data sources and modeling techniques to estimate the main
+population variables of krill in WAP.
+
+The stock assesment model was configured using Stock Synthesis (SS3
+hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013;
+@methot2020stock] with the most updated version (V.3.30.21). SS3 is a
+structured age and size stock evaluation model, in the class of models
+called *"Integrated stock evaluation analysis model"*. SS3 has a stock
+population sub-model that simulates growth, maturity, fecundity,
+recruitment, movement, and mortality processes, and observation
+sub-models and expected values for different types of data. The model is
+coded in `C++` with estimation parameters enabled by automatic
+differentiation (ADMB) [@Fournier2012; @Methot2013]. The analysis of
+results and outputs uses R tools and the graphical interface of the
+*r4ss* and *ss3diags* library [@Taylor2019; @Winker2023].
+
+By integrating data from multiple sources and considering spatial
+heterogeneity, the assessment methodology using Stock Synthesis
+v.3.30.21 provides a robust framework for evaluating the status of krill
+populations in the Antarctic Peninsula region. This information is
+essential for supporting management decisions aimed at ensuring the
+sustainable use of krill resources in this ecologically sensitive area.
 
@@
 ```
+
 ## Scenarios
 
-In Table 1 we have ten scenarios to test different option in modeling about main consideration in assessment of krill population.
+In Table 1 we have ten scenarios to test different option in modeling
+about main consideration in assessment of krill population.
 
 | Scenario | Description                                                                          |
-|:---------:|:-------------------------------------------------|
-|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
+|:------------:|:---------------------------------------------------------|
+|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
 |    s1    | Fishery data (Length, Index, Catch) by strata and Predator and Env data              |
@@
 
-Respecto a los valores y parametros biologicos modelados, los siguientes graficos identifican los estimadores puntuales del recurso
+Respecto a los valores y parametros biologicos modelados, los siguientes
+graficos identifican los estimadores puntuales del recurso
 
@@
 ```
+```{r}
+SSplotComps(base.model01)
+SSplotComps(base.model1)
+SSplotComps(base.model2)
+SSplotComps(base.model3)
+SSplotComps(base.model4)
+SSplotComps(base.model5)
+SSplotComps(base.model6)
+SSplotComps(base.model7)
+SSplotComps(base.model8)
+
+```
+
+# RESULTS
+
+## Diagnosis Base Model
+
+Step to do a good practice in model diagnosis is;
+
+-     1.  Convergence. Final convergence criteria is 1.0e-04
+
+-     2.  Residual (visual and metrics)
+
+-     3.  Retrospective analysis (Mhon Parameter)
+
+-    4.  Likelihood profile
+
+all this framework try to follow recommendations of @Carvalho2021b
+
+## Residual consistency in `s2`
 
 ```{r eval=FALSE}
-SSplotBiology(base.model2, 
-              subplots =2, 
-              labels = c("Length (cm)",
-                         "Age (yr)",
-                         "Maturity", 
-                         "Mean weight (kg) in last year",
-                          "Spawning output",
-                         "Length (cm, beginning of the year)", 
-                         "Natural mortality",
-                          "Female weight (kg)", 
-                         "Female length (cm)", 
-                         "Fecundity", 
-                         "Default fecundity label",
-                          "Year", 
-                         "Hermaphroditism transition rate", 
-                         "Fraction females by age at equilibrium"),
- )
-
-```
-
-aporte de las cohortes por año para las capturas.
-```{r eval=FALSE}
-SSplotCohortCatch(base.model2, subplots = 1)
-```
-
-\quad
-
-AJuste de tallas por flota
-```{r eval=FALSE}
-SSplotComps(base.model2, subplots = 1)
-```
-
-Otros plots
-```{r}
-SSplotDynamicB0(base.model2, uncertainty = F)
-#SSplotSPR(base.model3)
-```
-
-```{r eval=FALSE}
-SSplotPars(base.model2)
-
-```
-
-Salida de las biomasas con las dos flotas
-
-```{r}
-SSplotTimeseries(base.model2, subplot = 7)
-```
-Salida de las biomasas con todas las flotas
-
-```{r}
-SSplotTimeseries(base.model2, subplot = 7)
-```
-
-# RESULTS
-
-## Diagnosis Base Model
-
-Step to do a good practice in model diagnosis is;
-
-- 1. Convergence. Final convergence criteria is 1.0e-04
-
-- 2. Residual (visual and metrics)
-
-- 3. Retrospective analysis (Mhon Parameter)
-
-- 4. Likelihood profile
-
-all this framework try to follow recommendations of @Carvalho2021b
-
-## Residual
-
-```{r eval=FALSE}
-SSplotRunstest(base.model8, 
+SSplotRunstest(base.model2, 
                subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model2, 
-               subplots = "len",
                add=T)
@@
 
-## Predator fleet with RW
-
-Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.
-
-Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter.
-In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process.
-Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.
-
-Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time.
-However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.
+## Relationship Stock-Recruit
+
+```{r eval=FALSE}
+l <- base.model4$recruit$SpawnBio
+m <- base.model4$recruit$pred_recr
+ggplot()+
+  aes(l,m)+
+  geom_point()+
+  geom_smooth(method = "lm", 
+              se =F)+
+  ylim(0, max(base.model2$recruit$pred_recr))
+```
 
@@
 
-Los análisis retrospectivo, dan cuenta de diferencias de estimación (sub - sobreestimación) en los patrones entre modelos evaluados.
+Los análisis retrospectivo, dan cuenta de diferencias de estimación
+(sub - sobreestimación) en los patrones entre modelos evaluados.
 
 ```{r eval=FALSE}
-retro(dir=dir7, oldsubdir="", 
-      newsubdir="Retrospective", 
-      years= 0:-5,
-      exe="ss_osx",
+for (dir in directorios) {
+  retro(
+    dir = dir,
+    oldsubdir = "",
+    newsubdir = "Retrospective",
+    years = 0:-5,
+    exe = "ss_osx",
     extras = "-nox",
-      skipfinished = F)
-```
-
-```{r echo=FALSE}
+    skipfinished = FALSE
+  )
+}
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 #s01
@@
 retroSummary7 <- SSsummarize(retroModels7)
+
+#s8
+retroModels8 <- SSgetoutput(dirvec=file.path(dir8,
+                                            "Retrospective",
+                                            paste("retro",0:-5,
+                                                  sep="")))
+
+retroSummary8 <- SSsummarize(retroModels8)
 
@@
             verbose=F)
-
-```
-
-```{r echo=FALSE}
-ggarrange(retro01,
-          retro1,
-          retro2,
-          retro3,
-          retro4,
-          retro5,
-          retro6,
-          retro7, ncol = 4, nrow = 4)
+retro8 <- SSplotRetro(retroSummary8,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F)
+
 ```
@@
 
-Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated by `r4ss:SS_doRetro()`.
-As a robust measure of prediction skill, we implemented the mean absolute scaled error (MASE).
-In brief, the MASE score scales the mean absolute.
-Regarding (A MASE score > 1 indicates that the average model forecasts are worse than a random walk.
-Conversely, a MASE score of 0.5 indicates that the model forecasts twice as
-accurately as a naïve baseline prediction; thus, the model has prediction skill.
+Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock
+Synthesis requires the same model outputs generated by
+`r4ss:SS_doRetro()`. As a robust measure of prediction skill, we
+implemented the mean absolute scaled error (MASE). In brief, the MASE
+score scales the mean absolute. Regarding (A MASE score \> 1 indicates
+that the average model forecasts are worse than a random walk.
+Conversely, a MASE score of 0.5 indicates that the model forecasts twice
+as accurately as a naïve baseline prediction; thus, the model has
+prediction skill.
 
@@
 ```
+
 ## Kobe (status)
@@
 ```{r}
+mvln = SSdeltaMVLN(base.model01,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model1,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model2,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model3,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model4,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model5,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model6,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+mvln = SSdeltaMVLN(base.model7,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
 mvln = SSdeltaMVLN(base.model8,
@@
                    virgin = TRUE,
-                   mc=1000)
-
-mvln = SSdeltaMVLN(base.model2,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=1000)
+                   mc=100)
+
 ```
@@
 # 11. Vector de valores para el perfil ----
-R0.vec <- seq(15,20,0.5)  
+R0.vec <- seq(15,30,1)  
 Nprof.R0 <- length(R0.vec)
@@
 dev.off()
-
-### SSplotComparisons() ----
+```
+
+## SSplotComparisons()
+
+```{r eval=FALSE, echo = FALSE}
+
 # Comparación de series de tiempo 
@@
                             "Harvest rate"),
-                 plotdir = "~/Figs")
+                 png=TRUE,plotdir=Figs)
 
@@
 ```
+
 comparision between select models
@@
                  "s3"),
-                 subplots = c(2, 16),
-                 plotdir = "~/Figs")
+                 subplots = c(2, 16))
 comp2_8
 ```
+
 ## Comparsion in sd long term time series
@@
                                    hjust = 1))
-base.model2$
-
 ```
@@
           common.legend = T)
+
+```
+
+## Recruit deviation
+
+```{r}
+dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S01")
+dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
+  mutate(Serie = "S1")
+dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S2")
+dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S3")
+dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S4")
+dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S5")
+dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S6")
+dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S7")
+dev8 <- base.model8$recruit[1:43,c(1,7)] %>% 
+  mutate(Serie = "S8")
+
+data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7, dev8)
+titles <- c("S01", "S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8")
+# Lista para almacenar los gráficos
+plots <- list()
+
+# Crear los gráficos en un ciclo for
+for (i in 1:length(data_list)) {
+  p <- ggplot(data_list[[i]], aes(x = Yr, y = dev)) +
+    geom_point() +
+    stat_smooth(method = "loess", 
+                span = 0.15, 
+                se = T) +
+    geom_hline(yintercept = 0,
+               col="red",
+               linetype = "dotdash")+
+    ggtitle(titles[i]) +
+    theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1))+
+    theme_few()
+  plots[[i]] <- p
+}
+
+# Organizar los gráficos con ggarrange
+combined_plot <- ggarrange(plotlist = plots, ncol = 3, nrow = 3)
+
+combined_plot
+```
+
+```{r}
+combined_data <- rbind(dev01, 
+                           dev1, 
+                           dev2, 
+                           dev3, 
+                           dev4,
+                           dev5,
+                           dev6,
+                           dev7,
+                           dev8)
+
+combined_plot_join <- ggplot(combined_data, 
+                             aes(x = Yr, 
+                                 y = dev, 
+                                 color = Serie)) +
+  geom_point() +
+  stat_smooth(method = "loess", 
+              span = 0.15, 
+              se = FALSE) +
+  ggtitle("Combined Series Plot") +
+  geom_hline(yintercept = 0,
+               col="red",
+               linetype = "dotdash")+
+  scale_color_viridis_d(option="turbo",
+                        name="Scenario")+
+  theme_few() +
+  theme(legend.position = "right")
+
+combined_plot_join
+
 ```

### 2024-06-05 SS3_Krill_481.Rmd
adding html but not in page
test model to example

@@
 library(openxlsx)
+library(r4ss)
+dir1<-here("s1")
 ```
@@
 
-The stock assessment model for *Euphausia superba* (hereafter krill) in the Antarctic Peninsula, operating within a spatial-temporal and ecosystem framework, highlights the importance of incorporating spatial heterogeneity.
-The spatial heterogeneity, such as variations in oceanographic conditions and habitat suitability across different regions, is likely to outperform a population dynamics of krill.
-By accounting for spatial heterogeneity, a integrated stock assessment model can better capture localized dynamics and interactions, leading to more accurate predictions of krill population dynamics.
-This nuanced approach allows for a more refined understanding of different areas features krill structure, fishing mortality and recruitment patterns across the Antarctic Peninsula.
-Consequently, by acknowledging and integrating spatial heterogeneity, the stock assessment model can provide more robust insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
+The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a stock assessment model.
+Being able to identify index data relative, systematize, integrate and estimate krill population variables in a context of change and fisheries management is vital for decision making.
+On the other hand, spatial consideration and ecosystem aspect are another vital component for the application of this kind of approach.
+To do this, we use Stock Synthesis (SS3) as a statistical platform, a modeling tool widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
+This approach has three main objectives.
+Firstly, integrate fishing, environmental and ecological variables in an integrated model.
+On the other hand, consider the spatial heterogeneity of the krill population structure.
+Finally, evaluate the impact of biological and population assumptions on the performance of this type of integrated stock assessment models.
+The work approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions, which in turn, allows the adequate configuration of control rules that may be useful in the fisheries management applied by CCAMLR of this important resource in the Southern Ocean.
+Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
 
-*Keywords: Krill populations, dynamic population, stock assessment, SS3, spatial implicit, management, CCAMLR.*
+*Keywords: Krill populations, dynamic population, integrated model, SS3, spatial implicit, ecosystem consideration, management, CCAMLR.*
+
+\newpage
 
@@
 Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-As part of its mandate, CCAMLR routinely conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
-This regional fisheries management organization has intensified its efforts to promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
-
+As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
+This regional organization has tried to conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
 Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
@@
 Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
-However, these approaches may overlook critical aspects of krill dynamics, such as spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-Recognizing these limitations, there is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables.
+However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
+There is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables in krill.
 Integrating the complexity of krill population dynamics presents several challenges.
@@
 Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
-Despite these gaps, the existing data on krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
+Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
+In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery biological and ecosystem history on krill population over the past 20 years, as well as environmental variables, and predator abundance.
 
-In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, leveraging an integrated modeling approach that incorporates the fishing and biological history of krill over the past 20 years, as well as ecosystem variables such as environmental conditions, primary productivity levels affecting krill recruitment, and predator abundance.
-Unlike previous assessments, this apporach integrate data from both fisheries and population monitoring efforts, providing a more holistic understanding of krill dynamics.
-Moreover, our approach will go beyond conventional methods by explicitly considering spatial heterogeneity in krill abundance and distribution across the Antarctic Peninsula.
-
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
-By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013].
-Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.
-
-Beyond integrating available information for assessing krill population dynamics, we tested several natural scenarios to identify uncertainties associated with estimating key population variables.
-Using a set of eight scenarios, we aimed to diagnose, evaluate, assess performance, and compare the tested models.
-This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.
+\newpage
 
@@
 The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.
-Each stratum may have distinct characteristics that affect krill biology, such as sea ice cover, water temperature, and nutrient availability (Figure \@ref(fig:mapa)).
+Each stratum may have distinct characteristics that affect krill biology (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows for a more detailed and accurate assessment of krill populations by considering local variations in structure population.
+Therefore, dividing the study area into strata allows you to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
 
@@
 Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include: 1) Data from the monitoring program of the krill fishery were used, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
 This database has information about vessel, nationality, georeferenced, among others.
@@
 This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.
-All this data was proceses in @Dornam2021; 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
+All this data was proceses in @Dornam2021 and length compositions (Suppl.
+Mat.
+3); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
 Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
 Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
-This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics in the Western Antarctic Peninsula (WAP) area.
+This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics on krill in the WAP area.
 To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST).
-The model included spatial component (cell) as random effect (formula: ~1 | cell).
+The model included spatial component (cell) as random effect (formula: ~1 | cell) (Suppl.
+Mat 3).
 Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020].
-All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator **"fleet"** to add like M2 in an integrated stock assessment, in this case, we use a index of penguins abundance to consider this like another kind of natural mortality.
+All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD".
+Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series (Figure \@ref(fig:peng)), we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.
+Al this data and temporal scale is showed in Figure \@ref(fig:datas2).
+
+### Environmental variable in modelling
+
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics.
+SS3 addresses these approaches.
+In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
+
+\[
+E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
+\]
+
+\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
+Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
+However, the model does not automatically adjust for this.
+
+Furthermore, this integrated model will take into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
+These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
+By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
+
+### Predator components in modelling
+
+This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics.
+By incorporating information on predator  abundance and feeding ecology, our model will provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
+In this analysis we incorporated penguin population index as predator *"fleets"* how  we describe previously.
+This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.
 
@@
 
-The process of modeling and statistical analysis of a database can be structured according to the following guidelines:
+A conceptual model for krill in the Antarctic Peninsula includes five spatial strata.
+Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)).
+Ecological and environmental variables include krill biomass, predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
+The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.
 
-- Contextualization of the problem. Definition of objectives and variables.
-- Experiment design and data collection.
-- Recording and preliminary processing of available information.
-- Graphical inspection and trend identification.
-- Consideration of distributional and relational hypotheses. Proposal of modeling.
-- Model adjustment. Comparison and selection of the best model.
-- Diagnosis and validation of the adjusted model.
-- Assessment of the predictive capacity of the model and prediction.
-- Interpretation and conclusions.
+## Statistical model Stock Synthesis (SS3)
 
-## Population Dynamics Model
+Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
+Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
+Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
+The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the updated version 3.30.21.
+SS3 is a widely used software tool for assessing fish and invertebrate populations.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
+SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
+The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+The analysis of results and outputs uses R tools (@R-base) and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].
+
+## Population dynamics Model
 
@@
 This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
+More detail in main equation and model configuration can be found in @methot2020stock.
+Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
 
-Life history arameters taken as priors taken from @Smith2023a and @Maschette2020
+\newpage
 
-## Statistical model Stock Synthesis (SS3)
-
-Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
-Despite this, integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
-
-Stock Synthesis v.3.30.21 is a widely used software tool for assessing fish and invertebrate populations.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP.
-
-The stock assessment model was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013; @methot2020stock] with the most updated version (V.3.30.21).
-SS3 is a structured age and size stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"*.
-SS3 has a stock population sub-model that simulates growth, maturity, fecundity, recruitment, movement, and mortality processes, and observation sub-models and expected values for different types of data.
-The model is coded in `C++` with estimation parameters enabled by automatic differentiation (ADMB) [@Fournier2012; @Methot2013: @methot2020stock].
-The analysis of results and outputs uses R tools and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].
-
-By integrating data from multiple sources and considering spatial heterogeneity, the assessment methodology using Stock Synthesis v.3.30.21 provides a robust framework for evaluating the status of krill populations in the Antarctic Peninsula region.
-This information is essential for supporting management decisions aimed at ensuring the sustainable use of krill resources in this ecologically sensitive area.
-
-## Environmental factor in modelation
-
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics.
-SS3 addresses these approaches.
-In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
-
-\[
-E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-\]
-
-\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
-However, the model does not automatically adjust for this.
-
-Furthermore, our integrated model will take into account environmental variables known to influence krill abundance and distribution, such as sea surface temperature, chlorophyll concentration, and sea ice extent.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
-
-## Predator components in modelation
-
-Additionally, our model will consider predator-prey interactions as a key driver of krill population dynamics.
-By incorporating information on predator distribution, abundance, and feeding ecology, our model will provide a more comprehensive understanding of the trophic interactions shaping krill dynamics.
-In this analysis we incorporated penguin population index as predator *"fleets"*.
-This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock].
-
-## Scenarios
+## Stock Assessment Scenarios
 
 To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
-The methodology involves comparing the statistical performance of the model  regardng this configurations.
+The methodology involves comparing the statistical performance of the model regarding this configurations.
 In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.
@@
 | Scenario | Description                                                                          |
-|:---------:|:-------------------------------------------------|
-|    s01     |Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1 |
-|    s1     | Fishery data (Length, Index, Catch) by strata  and Predator and Env data      |
-|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
-|    s3     | "s2" without S-R relation         |
-|    s4     | "s2" Ricker S-R relation estimated          |
-|    s5     | "s2" BH S-R relation weak (0.9 steepness)     |
-|    s6     | "s2" BH S-R relation strong (0.6 steepness)     |
-|    s7     | "s2" BH S-R relation mid-strong (0.65 steepness) estimated     |
-|    s8     | "s2"  w/ set of parameters estimated in @EMM-204/32 |
+|:------------:|:---------------------------------------------------------|
+|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate 48.1                |
+|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
+|    s2    | "s1" without S-R relation                                                            |
+|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
+|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
+|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
+|    s6    | "s1" Ricker S-R relation estimated                                                   |
+|    s7    | "s1" w/ set of parameters estimated in @SAM-2024/32                                   |
 
-: Scenarios to modelling dynamics in krill in spatial implicit context
+: Scenarios to modelling dynamics in krill
+
+For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @WP2024/34
 
 These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
-all this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+All this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+
+\newpage
 
@@
 
-Step to do a good practice in model diagnosis is;
+All models tested was the final gradient of the  model was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive definite which shows that the performance is within what was expected for this type of analysis.
 
-- 1. Convergence. Final convergence criteria is 1.0e-04
+Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years.
+In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
+The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`.
+For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000.
+However, there is variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
+For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
+The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall).
+
+Overall, the model's performance is robust, but there is room for improvement in capturing the observed variability in certain fleets and years.
+In residuals components for `s1`, `s01` and `s7`, the model collects the modes and proportions for each source of information that contains lengths, however, it is the fleet of predators that fails to replicate the data of juvenile individuals (Figure \@ref(fig:fitcom)).
 
@@
 
-## Predator fleet with RW
-
-Random Walk (RW) refers to a mathematical model that describes a stochastic process in which a variable changes randomly over time, without a clear trend or pattern.
-
-Specifically, a random walk can be used as a Bayesian estimation technique to infer the posterior distribution of an unknown parameter.
-In this approach, it is assumed that the prior distribution of the parameter is a normal distribution with mean zero and a known variance, and that the parameter value at each time step follows a random walk process.
-Based on the observed data and the prior distribution, the posterior distribution of the parameter can be calculated using Bayesian inference.
-
-Random walk is a useful tool for parameter estimation in dynamic models, as it allows for modeling uncertainty and variability in the parameter's evolution over time.
-However, it is important to note that the random walk assumes that the changes in the parameter are random and without a clear trend, which may not be appropriate in all cases.
+Main variables and forecasting can be found in Table \@ref(tab:mainvar)
 
@@
 
-## Verosimilitud
+@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern.
+
+## Likelihood profile in spatial Model `s1`
+
+The R0 likelihood component profiles for the SMA and HAKE models (Fig.
+6) were developed using the function r4ss:SS_profile() (see https://github.com/jabbamodel/ss3diags for example R code).
+For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources.
+The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro).
+The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.
+
+## Cross correlation
 
@@
 
+Regarding to autocorrelation in krill recruitment in time series analized by scenarios, **S01** most ACF values fall within the significance limits, suggesting no significant autocorrelation, which indicates the series might be random or lack temporal dependency.
+In **S1**, several ACF values at the initial lags are significant, indicating the presence of autocorrelation in the early periods and suggesting temporal dependency.
+Similarly, **S2** shows significant autocorrelation in the initial lags, pointing to early temporal dependency as well.
+**S3** mostly stays within the significance bounds, with some exceptions at the initial lags, hinting at a slight autocorrelation in the early periods, but it is not very pronounced.
+**S4** mirrors S01, with ACF values within the significance limits, indicating no significant autocorrelation and suggesting randomness in the series.
+**S5** displays some significant autocorrelation in the initial lags, although less pronounced than in S1 and S2, suggesting a mild early temporal dependency.
+**S6** follows the pattern of S01 and S4, with no significant autocorrelation, implying a potentially random series.
+Lastly, **S7** shows ACF values mostly within the significance bounds, with a few exceptions at the initial lags, indicating a slight autocorrelation in the early periods, though not strongly pronounced.In summary, S01, S4, and S6 exhibit no significant autocorrelation, suggesting randomness.
+
 \newpage
+
 # DISCUSION
+
+Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
+By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
+This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013].
+Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.
+
+Beyond integrating available information for assessing krill population dynamics, we tested several scenarios to identify uncertainties associated with estimating key population variables that aimed to diagnose, evaluate, assess performance, and compare the tested models.
+This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.
 
@@
 
+- The main challenge has been to identify sources of information about krill population dynamic.
+It was possible to integrate the various pieces of information (fishery, survey, environmental, predator, LH
+parameters) that are available for krill in 48.1 into an integrated model (SS3) that takes care of the fishing and biological history of this species.
+- Conducting an stock assessment and get “best fit” is not the ultimate goal; rather, it is to identify sources of uncertainty associated with krill population dynamics and test their impact on main variables and transfer this type of information into fisheries management in CCAMLR.
+- Different scenarios about spatial heterogeneity and the krill stock-recruit relationship (steepness) were tested.
+- There are various ways to check the performance of the models. More data sources, more parameters!
+- The retrospective pattern of the aggregate model s1 is greater than s01, but the RMSE indicates that model
+s01 has less predictive power
+
+\newpage
+
+# SUPPLEMENTARY MATERIAL
+
+- Supplementary Material 3:  Length composition templates preparation to SS3 from AMLR data [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
+
+- Supplementary Material 4:  Correlation analysis between.
+biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
+
 \newpage
@@
 
-```{r mapa, out.width='100%', fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
+```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
 knitr::include_graphics('Figs/481.png')
@@
 
+```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
+knitr::include_graphics('Figs/conceptual.jpeg')
+```
 \begin{landscape}
-
 ```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
@@
 
-```{r parainit, fig.cap="Initial Parameters to modelling in stock process in krill in WAP"}
-parInit <- read_excel("DataKrill.xlsx", 
-    sheet = "parameters")
+```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
+
+# leo archivos para plotear y hacer tablas
+start1 <- SS_readstarter(file = file.path(dir1,
+                                          "starter.ss"),
+                              verbose = FALSE)
+# note the data and control file names can vary, so are determined from the 
+# starter file.
+dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
+                        verbose = FALSE)
+# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
+# assumes the correct data structure
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
+                                    start1$ctlfil),
+                        verbose = FALSE,
+                        use_datlist = TRUE, 
+                   datlist = dat1)
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
+                                                "forecast.ss"),
+                              verbose = FALSE)
+# can also read in wtatage.ss for an empirical wt at age model using
+# r4ss::SS_readwtatage()
+
+parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
+ row.names( parbio)<-c("Nat M",
+                       "Lmin", 
+                       "Lmax",
+                       "VonBert K",
+                       "CV young",
+                       "CV old", 
+                       "Wt a", 
+                       "Wt b",
+                       "L50%", 
+                       "Mat slope")
+
+ SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
+ Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
+ Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
+ parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
+
 parInit %>%
@@
 
-```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
-knitr::include_graphics('Figs/conceptual.jpeg')
+\begin{landscape}
+
+```{r peng,  out.width='100%', fig.cap="Primary penguin species time series abundant populations in colonies located at 48.1"}
+knitr::include_graphics('Figs/Cunt_Penguins.png')
+```
+\end{landscape}
+
+```{r datas2, out.width='70%', fig.cap="Data presence by year for each fleet and data type used in s1 scenario"}
+knitr::include_graphics('Figs/data_plot_s2.png')
 ```
 
-```{r datas2, out.width='70%', fig.cap="Data series using in s2 scenario"}
-knitr::include_graphics('s2/plots/data_plot.png')
+```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
+knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
 ```
 
-```{r outs2, fig.cap="Main variables outputs from stock asssessment krill in WAP"}
+```{r res1a, fig.cap = "Pearson residuals, comparing across fleets.
+Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
+knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
+```
+
+```{r fitcom, out.width='50%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
+par(mfrow=c(3,1))
+knitr::include_graphics(c('s01/plots/comp_lenfit__aggregated_across_time.png',
+                          's1/plots/comp_lenfit__aggregated_across_time.png',
+                          's7/plots/comp_lenfit__aggregated_across_time.png'))
+
+```
+
+\begin{landscape}
+
+```{r resall, out.width='20%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes."}
+par(mfrow=c(2,5))
+knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))
+
+```
+\end{landscape}
+
+```{r likepro, out.width='60%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
+knitr::include_graphics('Figs/Profile_s1.png')
+```
+
+\begin{landscape}
+```{r acf, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-31-1.jpeg')
+```
+\end{landscape}
+
+\begin{landscape}
+```{r recdev, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-30-1.jpeg')
+```
+
+\end{landscape}
+
+```{r longterm, out.width='90%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
+knitr::include_graphics('Figs/longtermsb.png')
+```
+
+```{r mainvar, fig.cap="Main variables outputs from estimation in s1 model krill in WAP"}
 out_s2 <- read_excel("DataKrill.xlsx", 
@@
 
-```{r comparision, fig.cap="Data series using in s2 scenario"}
+```{r comparision, out.width='40%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
 par(mfrow=c(2,2))
-knitr::include_graphics(c('compare8_Fvalue_uncertainty.png',
-                          'compare2_spawnbio_uncertainty.png',
-                          'compare4_Bratio_uncertainty.png',
-                          'compare8_Fvalue_uncertainty.png'))
+knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
+                          'Figs/compare2_spawnbio_uncertainty.png',
+                          'Figs/compare4_Bratio_uncertainty.png',
+                          'Figs/compare17_densities_SSB_Virgin.png'))
 ```

### 2024-06-05 index.Rmd
test model to example

@@
 dir7<-here("s7") # whitout S-R
-dir8<-here("s8")
+dirtest<-here("test")
 Figs <- here("Figs")# S2 set parametres EMM-2024/23 (Mardones)
@@
 # leo archivos para plotear y hacer tablas
-start2 <- SS_readstarter(file = file.path(dir2,
+start1 <- SS_readstarter(file = file.path(dir1,
                                           "starter.ss"),
@@
 # starter file.
-dat2 <- SS_readdat(file = file.path(dir2, start2$datfile),
+dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
                         verbose = FALSE)
@@
 # assumes the correct data structure
-ctl2 <-  r4ss::SS_readctl(file = file.path(dir2,
-                                    start2$ctlfil),
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
+                                    start1$ctlfil),
                         verbose = FALSE,
                         use_datlist = TRUE, 
-                   datlist = dat2)
-fore2 <- r4ss::SS_readforecast(file = file.path(dir2, 
+                   datlist = dat1)
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
                                                 "forecast.ss"),
@@
 ```{r}
-parbio<-ctl2$MG_parms[1:10,c(1:3,7)]
+parbio<-ctl1$MG_parms[1:10,c(1:3,7)]
  row.names( parbio)<-c("Nat M",
@@
 
- SRpar<-ctl2$SR_parms[1:5,c(1:3,7)]
- Qpar<-ctl2$Q_parms[1:2,c(1:3,7)]
- Selpar<-ctl2$size_selex_parms[1:22,c(1:3,7)]
- parInit<-rbind(parbio,SRpar,Qpar,Selpar)
+ SRpar<-ctl1$SR_parms[1:5,c(1:3,7)]
+ Qpar<-ctl1$Q_parms[1:2,c(1:3,7)]
+ Selpar<-ctl1$size_selex_parms[1:22,c(1:3,7)]
+ parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
 
@@
 |   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
-|    s1    | Fishery data (Length, Index, Catch) by strata and Predator and Env data              |
-|    s2    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
-|    s3    | "s2" without S-R relation                                                            |
-|    s4    | "s2" Ricker S-R relation estimated                                                   |
-|    s5    | "s2" BH S-R relation weak (0.9 steepness)                                            |
-|    s6    | "s2" BH S-R relation strong (0.6 steepness)                                          |
-|    s7    | "s2" BH S-R relation mid-strong (0.65 steepness) estimated                           |
-|    s8    | "s2" w/ set of parameters estimated in @EMM-204/32                                   |
+|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
+|    s2    | "s1" without S-R relation                                                            |
+|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
+|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
+|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
+|    s6    | "s1" Ricker S-R relation estimated                                                   |
+|    s7    | "s1" w/ set of parameters estimated in @EMM-204/32                                   |
 
@@
                  "s6",
-                 "s7",
-                 "s8")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
+                 "s7")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
 
@@
 # or 
-#system('./ss_osx') # to mac
 r4ss::run(
-  dir = dir01,
+  dir = dir7,
   exe = "ss_osx",
@@
 
-#s7
-base.model8 <- SS_output(dir=dir8,
-                        covar=T,
-                        forecast=T)
+```
+# RESULTS
+
+### Main Variables poulation
+
+Total biomass
+
+```{r}
+total01 <- as.data.frame(base.model01$timeseries[1:50,c(2,5)]) %>%
+  mutate(Serie = "No Spatial")
+total1 <- as.data.frame(base.model1$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial implicit")
+total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ without S-R relation")
+total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
+total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
+total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
+total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
+total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")
+
+totalbio <- rbind(total01,
+              total1,
+              total2,
+              total3,
+              total4,
+              total5,
+              total6,
+              total7)
+
+ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
+  geom_point(size=0.7,
+             col="black") +
+  geom_smooth(method = "loess",
+              span=0.75, 
+              se=TRUE)+
+  labs(title = "Compare variables between scenarios",
+       x = "",
+       y = "Total biomass (mt)") +
+  theme_few() +
+  scale_color_viridis_d(option="H",
+                        name="Scenarios")+
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1))
 ```
@@
 ```{r eval=FALSE, message=F, include=FALSE}
-SS_plots(base.model3, 
+SS_plots(base.model7, 
          uncertainty = TRUE,
@@
 
+Data used en both (spatial and No spatial models)
+
 ```{r fig.height=7}
@@
 
-SSplotData(base.model2, 
+SSplotData(base.model1, 
            subplot = 1,
            pheight = 15)
-
 ```
@@
 ```{r}
+SSplotSelex(base.model01,
+            subplots = 1)
+SSplotSelex(base.model1,
+            subplots = 1)
 SSplotSelex(base.model2,
             subplots = 1)
-```
-```{r}
-SSplotComps(base.model01)
-SSplotComps(base.model1)
-SSplotComps(base.model2)
-SSplotComps(base.model3)
-SSplotComps(base.model4)
-SSplotComps(base.model5)
-SSplotComps(base.model6)
-SSplotComps(base.model7)
-SSplotComps(base.model8)
-
-```
-
-# RESULTS
+SSplotSelex(base.model3,
+            subplots = 1)
+SSplotSelex(base.model4,
+            subplots = 1)
+SSplotSelex(base.model5,
+            subplots = 1)
+SSplotSelex(base.model6,
+            subplots = 1)
+SSplotSelex(base.model7,
+            subplots = 1)
+```
+
+```{r eval=FALSE}
+SSplotComps(base.model01,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+
+SSplotComps(base.model1,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model2,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model3,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model4,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model5,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model6,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+SSplotComps(base.model7,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+```
 
@@
 
-## Residual consistency in `s2`
+## Residual consistency 
+```{r}
+SSplotRunstest(base.model01,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model1,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model2,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model3,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model4,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model5,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model6,
+               subplots = "len",
+               add=T)
+SSplotRunstest(base.model7,
+               subplots = "len",
+               add=T)
+```
+
+```{r}
+SSplotRunstest(base.model01,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model1,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model2,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model3,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model4,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model5,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model6,
+               subplots = "cpue",
+               add=T)
+SSplotRunstest(base.model7,
+               subplots = "cpue",
+               add=T)
+```
 
 ```{r eval=FALSE}
-SSplotRunstest(base.model2, 
-               subplots = "cpue",
-               add=T)
-```
-
-```{r}
 SSplotJABBAres(base.model01,
@@
                add=T)
+```
+
+```{r}
+SSplotJABBAres(base.model01,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model1,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model2,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model3,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model4,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model5,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model6,
+               subplots = "cpue",
+               add=T)
 SSplotJABBAres(base.model7,
-               subplots = "len",
-               add=T)
-
-SSplotMnwt(base.model01, subplots = 2)
+               subplots = "cpue",
+               add=T)
 ```
@@
 ```{r eval=FALSE}
-l <- base.model4$recruit$SpawnBio
-m <- base.model4$recruit$pred_recr
+l <- base.model1$recruit$SpawnBio
+m <- base.model1$recruit$pred_recr
 ggplot()+
@@
   geom_point()+
-  geom_smooth(method = "lm", 
+  geom_smooth(method = "loess", 
               se =F)+
-  ylim(0, max(base.model2$recruit$pred_recr))
+  theme_few()+
+  ylim(0, max(base.model1$recruit$pred_recr))
 ```
@@
 (sub - sobreestimación) en los patrones entre modelos evaluados.
+
+```{r eval=FALSE}
+#one by one
+retro(
+    dir = dir1,
+    oldsubdir = "",
+    newsubdir = "Retrospective",
+    years = 0:-5,
+    exe = "ss_osx",
+    extras = "-nox",
+    skipfinished = FALSE)
+```
 
@@
 retroSummary7 <- SSsummarize(retroModels7)
-
-#s8
-retroModels8 <- SSgetoutput(dirvec=file.path(dir8,
+```
+
+```{r eval=FALSE, message=FALSE, warning=FALSE}
+#stest
+retroModelstest <- SSgetoutput(dirvec=file.path(dirtest,
                                             "Retrospective",
-                                            paste("retro",0:-5,
+                                            paste("retro",0:-3,
                                                   sep="")))
 
-retroSummary8 <- SSsummarize(retroModels8)
-
-```
-
-```{r echo FALSE}
+retroSummary <- SSsummarize(retroModelstest)
+
+```
+
+```{r echo=FALSE}
 #save(retroSummary, retroModels, file="retro5.Rdata")
@@
             verbose=F)
-retro8 <- SSplotRetro(retroSummary8,
+
+```
+
+```{r eval=FALSE}
+retrotest <- SSplotRetro(retroSummarytest,
             add=T,
@@
             verbose=F)
-
 ```
@@
 ```{r}
+hci = SSplotHCxval(retroSummary01, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+hci = SSplotHCxval(retroSummary1, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
 hci = SSplotHCxval(retroSummary2, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+hci = SSplotHCxval(retroSummary3, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7) 
+hci = SSplotHCxval(retroSummary4, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+hci = SSplotHCxval(retroSummary5, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+hci = SSplotHCxval(retroSummary6, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+hci = SSplotHCxval(retroSummary7, 
                    add = T, 
@@
 
-mvln = SSdeltaMVLN(base.model8,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
+```
+
+another
+
+```{r eval=FALSE}
+
+nanios<-base.model1$Kobe$Yr
+SSBDep<-base.model1$Kobe$B.Bmsy
+y_BDPR<-base.model1$Kobe$F.Fmsy
+
+# Integrated Model
+
+kobe1 <- ggplot(base.model1$Kobe)+
+  #geom_line(aes(kobebro$BBMSY, kobebro$FFMSY)) +
+  geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0, ymax = 5),
+            fill = "#E43338", alpha = 0.5) +
+  geom_rect(aes(xmin = 0.5, xmax = 0.75, ymin = 0, ymax = 5),
+            fill = "#F2ED23", alpha = 0.5) +
+  geom_rect(aes(xmin = 0.75, xmax = 1.25, ymin = 0, ymax = 5),
+            fill = "#ACC39A", alpha = 0.5) +
+  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 0, ymax = 1),
+            fill = "#608D68", alpha = 0.5) +
+  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 1, ymax = 5),
+            fill = "#808080", alpha = 0.5) +
+  geom_path(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr))+
+  geom_point(aes(base.model1$Kobe$B.Bmsy, base.model1$Kobe$F.Fmsy),
+             lwd=2) +
+  geom_hline(yintercept = 1) +
+  geom_vline(xintercept = c(0.5, 0.75, 1.75, 1, 1.25), linetype=2)+
+  theme_few()+
+  labs(x = expression("BD/BD"[RMS]), y = expression("F/F"[RMS]))+
+  # geom_text(data = texto_coords, aes(x = x, y = y, label = etiqueta),
+  #            vjust = -0.5)+
+  geom_text(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr),
+            nudge_y = 0.1,size = 3,
+            check_overlap = TRUE)
+kobe1
 
@@
 tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
-tablebias02 <- SShcbias(retroSummary01,quant="F",verbose=F)
+tablebias01a <- SShcbias(retroSummary01,quant="F",verbose=F)
 
@@
 
-kbl(tablebias02, booktabs = T,format = "latex",
+kbl(tablebias01a, booktabs = T,format = "latex",
     caption = "Rho parameter in F model s01")  %>% 
@@
 
+# model1
+tablebias1 <- SShcbias(retroSummary1,quant="SSB",verbose=F)
+tablebias1a <- SShcbias(retroSummary1,quant="F",verbose=F)
+
+kbl(tablebias1, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s1")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias1a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s1")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
 # model2
+tablebias2 <- SShcbias(retroSummary2,quant="SSB",verbose=F)
+tablebias2a <- SShcbias(retroSummary2,quant="F",verbose=F)
+
+kbl(tablebias2, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s2")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias2a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s2")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model3
+tablebias3 <- SShcbias(retroSummary3,quant="SSB",verbose=F)
+tablebias3a <- SShcbias(retroSummary3,quant="F",verbose=F)
+
+kbl(tablebias3, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s3")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias3a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s3")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model4
+tablebias4 <- SShcbias(retroSummary4,quant="SSB",verbose=F)
+tablebias4a <- SShcbias(retroSummary4,quant="F",verbose=F)
+
+kbl(tablebias4, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s4")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias4a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s4")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model5
+tablebias5 <- SShcbias(retroSummary5,quant="SSB",verbose=F)
+tablebias5a <- SShcbias(retroSummary5,quant="F",verbose=F)
+
+kbl(tablebias5, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s5")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias5a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s5")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model6
+tablebias6 <- SShcbias(retroSummary6,quant="SSB",verbose=F)
+tablebias6a <- SShcbias(retroSummary6,quant="F",verbose=F)
+
+kbl(tablebias6, booktabs = T,format = "latex",
+    caption = "Rho parameter in SSB model s6")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias6a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s6")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+# model7
 tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
-tablebias07 <- SShcbias(retroSummary7,quant="F",verbose=F)
+tablebias7a <- SShcbias(retroSummary7,quant="F",verbose=F)
 
 kbl(tablebias7, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s2")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias7, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s2")  %>% 
-    kable_styling(latex_options = "HOLD_position")
+    caption = "Rho parameter in SSB model s7")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
+kbl(tablebias7a, booktabs = T,format = "latex",
+    caption = "Rho parameter in F model s7")  %>% 
+    kable_styling(latex_options = "HOLD_position")
+
 ```
@@
 like7 <- rownames_to_column(like7, var = "Description")
-like8 <- base.model8$likelihoods_used
-like8$model <- rep("s8", nrow(like8))
-like8 <- rownames_to_column(like8, var = "Description")
+
 totalike <- rbind(like01,
@@
                    like6,
-                   like7,
-                  like8)
+                   like7)
 
@@
 ggplot(totalike)+
-  geom_bar( aes(y = reorder(Description, values), x =values),
+  geom_bar( aes(y = reorder(Description, values), x =log(values)),
             stat = "identity", position = "dodge") +
@@
   facet_wrap(.~model, 
-             ncol=3)+
+             ncol=4)+
   labs(y="",
-       x="Likelihood")+
+       x="Log Likelihood")+
   theme(axis.text.x = element_text(angle = 90, 
-                                   hjust = 1)) 
+                                   hjust = 1))+
+  xlim(0,20)
 ```
@@
 # 1. Identificar el directorio donde se encuentra el modelo base ----
-dirname.model.run <- here("s2")
+dirname.model.run <- here("s1")
 # 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
@@
 # 15. Ejecutar la función profile() ----
-?SS_profile()
+#?SS_profile()
 profile <- profile(dir=dirname.R0.profile, # directory
@@
                                  "Age_comp",
+                                 "Catch",
                                  'Size_at_age',
@@
                                  "Age likelihood",
+                                 "Catch Likelihood",
                                  'Size_at_age likelihood',
@@
               profile.label=expression(log(italic(R)[0])), 
-              ymax=150,minfraction = 0.001,
+              ymax=2050,minfraction = 0.001,
               pheight=4.5, 
@@
 ```
+
+## Convergence criteria
+
+0.0001 final convergence criteria (e.g. 1.0e-04)
 
@@
 ```{r}
-outps2 <- base.model2$timeseries[1:53, 2:8]
+outps1 <- base.model1$timeseries[1:53, 2:8]
 
 addWorksheet(wb, "variable_s2")
-writeData(wb, "variable_s2", outps2)
+writeData(wb, "variable_s2", outps1)
 
@@
 ```{r echo=FALSE}
-#PLOT labels, name of each model run
-legend.labels <- c( "s01",
-                  "s1",
-                 "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7",
-                 "s8")
 
@@
                                   dir6,
-                                  dir7,
-                                  dir8),
+                                  dir7),
                        getcovar = F)
@@
 #create summary of model runs from list above
-summaryoutput <- SSsummarize(biglist)
-
-SSplotComparisons(summaryoutput,
+summaryoutputall <- SSsummarize(biglist)
+```
+
+```{r echo=FALSE}
+compall <- SSplotComparisons(summaryoutputall,
                   legendlabels = c("No Spatial",
@@
                  "s6",
-                 "s7",
-                 "s8"),
+                 "s7"),
                  labels = c("Year", 
@@
 ```{r}
-comtable <- SStableComparisons(summaryoutput,
+comtable <- SStableComparisons(summaryoutputall,
                    likenames = c("TOTAL", 
@@
                              "SSB_Virg", 
-                             "Bratio_2017",
-                             "SPRratio_2016"),
+                             "Bratio_2020",
+                             "SPRratio_2020"),
                    digits = NULL,
@@
                  "s7"),
-                   csv = FALSE,
+                   csv = F,
                    csvdir =~"/DOCAS/SA_Krill",
@@
 
-comparision between select models
+comparision between select models `No spatial`and `Spatial implicit` and `Spatial W/ new set parametres`
 
@@
 #note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist <- SSgetoutput(keyvec = NULL,
+biglist01_1_7 <- SSgetoutput(keyvec = NULL,
                        dirvec = c(
-                                  dir8,
-
-                                  dir2,
-                                  dir3),
+                                  dir01,
+                                  dir1,
+                                  dir7),
                        getcovar = F)
@@
 #create summary of model runs from list above
-summaryoutput <- SSsummarize(biglist)
-
-SSplotComparisons(summaryoutput,
+summaryoutput01_1_7 <- SSsummarize(biglist01_1_7)
+
+comp01_1_7 <- SSplotComparisons(summaryoutput01_1_7,
                   legendlabels = c(
-                                   "s8",
-
-                 "s2",
-                 "s3"),
+                                   "No Spatial",
+
+                 "Spatial",
+                 "Spatial w/ new set LH paramters"),
                  labels = c("Year", 
@@
                  png=TRUE,plotdir=Figs)
-
-comp2_8 <- SSplotComparisons(summaryoutput,
+```
+
+comparision between select models `Old Paramters` and `New Parameters WG SAM 2024/23`
+
+```{r echo=FALSE}
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist1_7 <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir1,
+                                  dir7),
+                       getcovar = F)
+
+#create summary of model runs from list above
+summaryoutput1_7 <- SSsummarize(biglist1_7)
+
+comp1_7 <- SSplotComparisons(summaryoutput1_7,
                   legendlabels = c(
-                                   "s8",
-
-                 "s2",
-                 "s3"),
-                 subplots = c(2, 16))
-comp2_8
+                                   "Old Parameters",
+
+                 "Parameters WG SAM 2024/23"),
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"),
+                 png=TRUE,plotdir=Figs)
 ```
@@
 ```{r}
-sblt01 <- base.model01$timeseries$Bio_all 
+
+sblt01 <- base.model01$timeseries$Bio_all[1:50]
 sblt1 <-  base.model1$timeseries$Bio_all
@@
 sblt7 <-  base.model7$timeseries$Bio_all
-sblt8 <- base.model8$timeseries$Bio_all
+
 errt <- as.data.frame(cbind(sblt01,
@@
               sblt6,
-              sblt7,
-              sblt8))
+              sblt7))
 err_long <- errt %>%
@@
                values_to = "value")  
+titles <- c("No Spatial",
+            "Spatial implicit", 
+            "Spatial w/ without S-R relation",
+            "Spatial w/ S-R relation weak (0.9 steepness)",
+            "Spatial w/ S-R relation strong (0.6 steepness)",
+            "Spatial w/ S-R relation mid-strong estimated",
+            "Spatial w/ Ricker S-R relation estimated", 
+            "Spatial w/ New set parameters Linf, k and M")
 
@@
   geom_jitter(width = 0.2, alpha = 0.3) +
-  labs(title = "",
+  labs(title = "Long Term Predictions",
        x = "",
-       y = "Spawning B.(mt)") +
+       y = "Spawning Biomass (t)") +
   theme_few() +
+  scale_x_discrete(labels= titles)+
   theme(axis.text.x = element_text(angle = 90,
@@
 recs7 <- base.model7$recruit$pred_recr
-recs8 <- base.model8$recruit$pred_recr
 
 p01 <- ggAcf(base.model01$recruit$pred_recr) +
-  ggtitle("S01") +
+  ggtitle("No Spatial") +
   theme_few()
 p1 <- ggAcf(base.model1$recruit$pred_recr) +
-  ggtitle("S1") +
+  ggtitle("Spatial implicit") +
   theme_few()
 p2 <- ggAcf(base.model2$recruit$pred_recr) +
-  ggtitle("S2") +
+  ggtitle("Spatial w/ without S-R relation") +
   theme_few()
 p3 <- ggAcf(base.model3$recruit$pred_recr) +
-  ggtitle("S3") +
+  ggtitle("Spatial w/ S-R relation weak (0.9 steepness)") +
   theme_few()
 p4 <- ggAcf(base.model4$recruit$pred_recr) +
-  ggtitle("S4") +
+  ggtitle("Spatial w/ S-R relation strong (0.6 steepness)") +
   theme_few()
 p5 <- ggAcf(base.model5$recruit$pred_recr) +
-  ggtitle("S5") +
+  ggtitle("Spatial w/ S-R relation mid-strong estimated") +
   theme_few()
 p6 <- ggAcf(base.model6$recruit$pred_recr) +
-  ggtitle("S6") +
+  ggtitle("Spatial w/ Ricker S-R relation estimated") +
   theme_few()
 p7 <- ggAcf(base.model7$recruit$pred_recr) +
-  ggtitle("S6") +
-  theme_few()
-p8 <- ggAcf(base.model8$recruit$pred_recr) +
-  ggtitle("S6") +
+  ggtitle("Spatial w/ New set parameters Linf, k and M") +
   theme_few()
@@
           p6,
-          p7,
-          p8, ncol=3, nrow = 3,
+          p7, ncol=3, nrow = 3,
           common.legend = T)
@@
   mutate(Serie = "S7")
-dev8 <- base.model8$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S8")
-
-data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7, dev8)
-titles <- c("S01", "S1", "S2", "S3", "S4", "S5", "S6", "S7", "S8")
+
+data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7)
+titles <- c("No Spatial",
+            "Spatial implicit", 
+            "Spatial w/ without S-R relation",
+            "Spatial w/ S-R relation weak (0.9 steepness)",
+            "Spatial w/ S-R relation strong (0.6 steepness)",
+            "Spatial w/ S-R relation mid-strong estimated",
+            "Spatial w/ Ricker S-R relation estimated", 
+            "Spatial w/ New set parameters Linf, k and M")
 # Lista para almacenar los gráficos
@@
 
-```{r}
-combined_data <- rbind(dev01, 
-                           dev1, 
-                           dev2, 
-                           dev3, 
-                           dev4,
-                           dev5,
-                           dev6,
-                           dev7,
-                           dev8)
-
-combined_plot_join <- ggplot(combined_data, 
-                             aes(x = Yr, 
-                                 y = dev, 
-                                 color = Serie)) +
-  geom_point() +
-  stat_smooth(method = "loess", 
-              span = 0.15, 
-              se = FALSE) +
-  ggtitle("Combined Series Plot") +
-  geom_hline(yintercept = 0,
-               col="red",
-               linetype = "dotdash")+
-  scale_color_viridis_d(option="turbo",
-                        name="Scenario")+
-  theme_few() +
-  theme(legend.position = "right")
-
-combined_plot_join
-
-```
-
 \newpage

### 2024-06-06 SS3_Krill_481.Rmd
add last changes in html

@@
 title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
-subtitle: "Working Paper to be submitted in CCAMLR WG-FSA 2024"
 author:
-  - Mardones Mauricio^[Universidad de Magallanes,Chile. mamardon@umag.cl]
-  - César Cárdenas^[Instituto Antártico Chileno, Chile]
+  - Mardones Mauricio^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
+  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
+  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
+  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
+  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
+
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a stock assessment model.
-Being able to identify index data relative, systematize, integrate and estimate krill population variables in a context of change and fisheries management is vital for decision making.
-On the other hand, spatial consideration and ecosystem aspect are another vital component for the application of this kind of approach.
-To do this, we use Stock Synthesis (SS3) as a statistical platform, a modeling tool widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
+The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a integrated stock assessment model.
+Being able to identify that sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management nd decision making.
+On the other hand, spatial consideration and ecosystem aspect are vital components to model dynamics population on krill.
 This approach has three main objectives.
-Firstly, integrate fishing, environmental and ecological variables in an integrated model.
+Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance.
 On the other hand, consider the spatial heterogeneity of the krill population structure.
-Finally, evaluate the impact of biological and population assumptions on the performance of this type of integrated stock assessment models.
-The work approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions, which in turn, allows the adequate configuration of control rules that may be useful in the fisheries management applied by CCAMLR of this important resource in the Southern Ocean.
+Finally, evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
 Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
@@
 As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
-This regional organization has tried to conducts stock assessments to assess the status of important species, such as Antarctic krill, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
-Various stock assessments research on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
-However, many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
-Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys.
+This regional organization has tried to conducts different modelling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
+Various stock assessments approach on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
+Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
+Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013].
 However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-There is a growing consensus on the need for integrated stock assessment models that incorporate multiple sources of information and consider the complex interplay of ecological variables in krill.
+There is a growing consensus on the need for use stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
 Integrating the complexity of krill population dynamics presents several challenges.
 Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension.
-Additionally, certain regions lack sufficient data or possess only limited information regarding krill dynamics.
+Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics.
 Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
 Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
-In response to this need, we propose a conceptual model to conducted a stock assessment of Antarctic krill in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery biological and ecosystem history on krill population over the past 20 years, as well as environmental variables, and predator abundance.
+In response to this need, we propose a conceptual model to conducted a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
 
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021.
-Each stratum may have distinct characteristics that affect krill biology (Figure \@ref(fig:mapa)).
+The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021 (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
@@
 These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others.
-Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019]; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations.
-This may include data on abundance, distribution, and biological characteristics of krill collected through net sampling, acoustic surveys, and other techniques.
-All this data was proceses in @Dornam2021 and length compositions (Suppl.
+This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations.
+This include data on abundance by strata, and biological characteristics of krill collected through net sampling.
+All this data was process in @Dornam2021; and length compositions from AMLR survey (provide by Ecosystem Science Division, NOAA) was handling according to the evaluation template requirements (Suppl.
 Mat.
-3); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
+2); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
 Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
@@
 All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD".
-Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series (Figure \@ref(fig:peng)), we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.
-Al this data and temporal scale is showed in Figure \@ref(fig:datas2).
+Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
 
-### Environmental variable in modelling
+## Conceptual Model
 
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics.
-SS3 addresses these approaches.
+A conceptual model for krill in the Antarctic Peninsula includes the five spatial strata described previously.
+Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)).
+Ecological and environmental variables that have impact in krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
+The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.
+
+## Statistical model Stock Synthesis (SS3)
+
+Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
+Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
+Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
+The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
+SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
+SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
+The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
+The analysis of results and outputs we use R tools (@R-base) and the graphical interface of the SS3 it was through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
+Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
+Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
+
+### Environmental variable modelling in `SS3`
+
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock].
 In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
@@
 
-### Predator components in modelling
+### Predator components modelling in `SS3`
 
@@
 
-## Conceptual Model
-
-A conceptual model for krill in the Antarctic Peninsula includes five spatial strata.
-Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)).
-Ecological and environmental variables include krill biomass, predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
-The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.
-
-## Statistical model Stock Synthesis (SS3)
-
-Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
-Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013] with the updated version 3.30.21.
-SS3 is a widely used software tool for assessing fish and invertebrate populations.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The analysis of results and outputs uses R tools (@R-base) and the graphical interface of the *r4ss*  and *ss3diags* libraries [@Taylor2019; @Winker2023].
-
-## Population dynamics Model
-
-In a simple way, the core of Stock Synthesis is its population dynamics model, which represents the dynamics of krill populations over time.
-This model incorporates key biological parameters such as growth rates, mortality rates, recruitment, and spawning biomass.
-The model is typically formulated using mathematical equations that describe how these parameters interact to determine the abundance and distribution of krill in the study area.
-
-A typical state-space model for krill population dynamics can be represented as:
-
-\[
-N_t = N_{t-1} \cdot e^{(r - M)} + R
-\]
-
-Where:
-
-\(N_t\) is the abundance of krill at time \(t\),  \(N_{t-1}\) is the abundance of krill at the previous time step, \(r\) is the intrinsic growth rate of the population, \(M\) is the mortality rate and \(R\) is the recruitment of new individuals into the population.
-This equation represents the basic dynamics of the krill population, with abundance changing over time due to growth, mortality, and recruitment.
-More detail in main equation and model configuration can be found in @methot2020stock.
-Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
+All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
 
@@
 
-To consider the uncertainty associated with the krill nature, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
+To consider the uncertainty associated with of krill nature dimension, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
 The methodology involves comparing the statistical performance of the model regarding this configurations.
-In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population.
+In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population, and for comparative propose, we consider `s1` as base model.
 
@@
 |    s6    | "s1" Ricker S-R relation estimated                                                   |
-|    s7    | "s1" w/ set of parameters estimated in @SAM-2024/32                                   |
+|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par                                  |
 
@@
 
-For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @WP2024/34
-
-These requests may include data from scientific research programs, environmental monitoring initiatives, and other sources that can contribute to improving the accuracy and reliability of the assessment.
-All this component and road to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @Mardones2024par (unpublished).
 
@@
 
-All models tested was the final gradient of the  model was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive definite which shows that the performance is within what was expected for this type of analysis.
-
-Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years.
+All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
+About goodness-of-fit, Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
 In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
@@
 For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
-The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall).
-
-Overall, the model's performance is robust, but there is room for improvement in capturing the observed variability in certain fleets and years.
-In residuals components for `s1`, `s01` and `s7`, the model collects the modes and proportions for each source of information that contains lengths, however, it is the fleet of predators that fails to replicate the data of juvenile individuals (Figure \@ref(fig:fitcom)).
-
-- 2. Residual (visual and metrics)
-
-- 3. Retrospective analysis (Mhon Parameter)
-
-- 4. Likelihood profile
-
-Main variables and forecasting can be found in Table \@ref(tab:mainvar)
+The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
 
@@
 
-Los análisis retrospectivo, dan cuenta de diferencias de estimación (sub - sobreestimación) en los patrones entre modelos evaluados.
-
-## Hindcast Cross-Validation and prediction skill
-
-Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis requires the same model outputs generated by `r4ss:SS_doRetro()`.
-As a robust measure of prediction skill, we implemented the mean absolute scaled error (MASE).
-In brief, the MASE score scales the mean absolute.
-Regarding (A MASE score > 1 indicates that the average model forecasts are worse than a random walk.
-Conversely, a MASE score of 0.5 indicates that the model forecasts twice as
-accurately as a naïve baseline prediction; thus, the model has prediction skill.
-
-models S01, S1, S2, and S5 exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
-Models S3 and S4 show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
-Model S6 shows moderate autocorrelations, indicating some predictability based on recent past values.
-These findings help understand the temporal dynamics of krill recruitment and inform future modeling efforts.
-(fig autocorrelation)
-## Kobe (status)
-
-@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern.
+Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
+@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
+To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we use the joint-index residual.
+Overall, `s1` have a good performance when evaluating the lengths compositions (14.2%), however  RMSE show a low predictive power with respect to the indices (71.3%).
+A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).
 
@@
 
-The R0 likelihood component profiles for the SMA and HAKE models (Fig.
-6) were developed using the function r4ss:SS_profile() (see https://github.com/jabbamodel/ss3diags for example R code).
 For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources.
@@
 
-## Cross correlation
+Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)
 
-## Comparision outputs betwwen scenarios
+## Comparision outputs between scenarios
 
-Regarding to autocorrelation in krill recruitment in time series analized by scenarios, **S01** most ACF values fall within the significance limits, suggesting no significant autocorrelation, which indicates the series might be random or lack temporal dependency.
-In **S1**, several ACF values at the initial lags are significant, indicating the presence of autocorrelation in the early periods and suggesting temporal dependency.
-Similarly, **S2** shows significant autocorrelation in the initial lags, pointing to early temporal dependency as well.
-**S3** mostly stays within the significance bounds, with some exceptions at the initial lags, hinting at a slight autocorrelation in the early periods, but it is not very pronounced.
-**S4** mirrors S01, with ACF values within the significance limits, indicating no significant autocorrelation and suggesting randomness in the series.
-**S5** displays some significant autocorrelation in the initial lags, although less pronounced than in S1 and S2, suggesting a mild early temporal dependency.
-**S6** follows the pattern of S01 and S4, with no significant autocorrelation, implying a potentially random series.
-Lastly, **S7** shows ACF values mostly within the significance bounds, with a few exceptions at the initial lags, indicating a slight autocorrelation in the early periods, though not strongly pronounced.In summary, S01, S4, and S6 exhibit no significant autocorrelation, suggesting randomness.
+In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions.
+We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years[@Perry2020] (Figure \@(fig:recdev)).
+
+Regarding to autocorrelation in krill recruitment in time series analized by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
+Models `s3` and `s4` show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
+Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary,
+`s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness. 
+(Figure \@(fig:acf)).
+
+To evaluate consistency of the source of information used in each of the tested scenarios, we verify its influence on the estimate of R0 through likelihood components and, like scenario 1, we verify that all scenarios show a high dependence on the composition lengths of lengths with respect to the estimate of the variable (as show the (Figure \@ref(fig:likeall)).
+
+With respect to population variables such as spawning biomass, we identify differences in the different scenarios with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios.
+This is relevant given that these scenarios consider low spatial dependence, density-dependence spawning biomass with respect to recruitment, which is plausible for resources such as krill where environmental dynamics shape population dynamics (Figure \@ref(fig:longterm)).
+
+We carried out a specific comparison between the most divergent scenarios that we tested in this case, `s1`,  `s01` and `s7` which according to the description are the aggregated spatial scenario, disaggregate spatial scenario and scenario with new set of parameters, respectively.
+All with ecosystem components such as predators and an environmental variable.In this sense, the impacts and difference on the estimates of recruitment,  R0, virginal biomass and relative biomass are high (Figure \@ref(fig:comparision)), demonstrating the impact of initial considerations and assumptions on population variables estimation.
 
@@
 
-# DISCUSION
+# DISCUSSION
 
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
-By incorporating spatially implicit information into our model, we aim to capture the complex spatial patterns of krill distribution and better understand the underlying drivers of variability in krill abundance.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013].
-Consequently, this approach aligns with current management frameworks, accurately reflecting krill population dynamics at a spatial scale and providing management recommendations at more localized scales than the current global ones.
-
-Beyond integrating available information for assessing krill population dynamics, we tested several scenarios to identify uncertainties associated with estimating key population variables that aimed to diagnose, evaluate, assess performance, and compare the tested models.
-This methodology represents a modern approach to assessing population dynamics and provides a foundation for making recommendations for the sustainable management of krill populations.
-
-Development of an integrated stock assessment like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
-Integrated models are not new to krill in SO.
+The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
+Development of an integrated stock assessment model like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
+However, integrated models are not new to krill in SO.
 For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
@@
 Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO).
+In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.
 
-In that sense, the work carried out here differs substantially from the previously written integrated models since it incorporates aspects that have to do with ecosystem indicators such as environmental variables and predators.
-On the other hand, it considers the spatial component within the modeling scheme implicitly and finally evaluates the impact of the different sets of life history parameters that have been used and those calculated by Mardones et al (in press)
+Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
+By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution.
+This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
 
 From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
+There is many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
+For this, we consider penguin colonies population as one of the most powerful elements of natural predation on krill because have direct correlations between this type of taxonomic group and krill [@Kruger2021].
 The intention to incorporate elements of predation in a stock is necessary to improve the modelling, because did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
-There is many studies that confirm the influence of predation on grill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
-For this, we consider penguin columns as one of the most powerful elements of natural predation on krill because have direct correlations between this type of taxonomic group and krill [@Kruger2021].
 
-**Paragraph 4:**  
-Model estimation concludes when a convergent Hessian matrix is achieved, and all data sources are fitted using maximum likelihood estimates from the assessment method.
+In a scientifically rigorous stock assessment, this analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics.
+It is important to note that the models presented here can be refined in terms of initial conditioning to improve fits on length compositions of predators and indices that were initially low.
+However, a primary objective was to evaluate the impact on key population variables, considering factors such as spatial heterogeneity and life history parameters.
+This impact was verified in this stock assessment approach, identifying aspects that must be considered when making recommendations for krill management by CCAMLR using models.
 
-**Paragraph 5:**  
-The independent survey data, which include survey abundance and length composition data, were collected by the United States Antarctic Living Marine Resources (AMLR) Program of the National Marine Fisheries Service, Southwest Fisheries Science Center.
-These surveys, conducted during the austral summer from 1992 to 2011, utilized zooplankton nets to gather sample data.
-
-**Paragraph 6:**  
-The fishery-dependent data, encompassing length composition data and catch biomass, are derived from CCAMLR's public database.
-Catch biomass data are available from 1992 to 2011, while length composition data are available for the same period.
-
-**Paragraph 7:**  
-The estimated recruitment of krill exhibits cyclical variations with an average period of 3–4 years.
-This cyclical variation is influenced by the physical environment of the polar ecosystem (Kinzey et al., 2019) and the behavior of its predators (Fraser et al., 1992; Fraser and Trivelpiece, 1995; Reid et al., 1999, 2002).
-
-**Paragraph 8:**  
-When evaluating any fishery resource, the chosen evaluation method should ensure that the results meet the reference data requirements for production management and optimize the use of existing data (Punt et al., 2013).
-
- Some authors have idicated spatial heterogeneity and identified different zones, such as areas with higher or lower concentrations of juveniles or regions more affected by the environment, which structure the population differently even inside WAP (REF*).
-
- By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, one can gauge the significance of incorporating spatial variation into the model.
+By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, the significance of incorporating spatial variation into the model can be assessed.
 This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
 
-In summary, our proposed stock assessment represents a novel and integrated approach to evaluating Antarctic krill populations in the Antarctic Peninsula.
-By leveraging multiple sources of information and explicitly considering spatial heterogeneity, environmental variables, and predator-prey interactions, our model aims to provide policymakers and stakeholders with actionable insights for informed decision-making within the framework of CCAMLR.
-Through collaborative efforts and interdisciplinary research, we are committed to advancing the science of krill ecology and supporting the conservation and sustainable management of Antarctic marine resources.
+Nowadays, one of the frontiers of stock assessment is the incorporation of ecosystem and spatial considerations into modeling.
+In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
+In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in subarea 48.1 of the Western Antarctic Peninsula (WAP).
 
 \newpage
-# CONCLUSION
 
-- The main challenge has been to identify sources of information about krill population dynamic.
-It was possible to integrate the various pieces of information (fishery, survey, environmental, predator, LH
-parameters) that are available for krill in 48.1 into an integrated model (SS3) that takes care of the fishing and biological history of this species.
-- Conducting an stock assessment and get “best fit” is not the ultimate goal; rather, it is to identify sources of uncertainty associated with krill population dynamics and test their impact on main variables and transfer this type of information into fisheries management in CCAMLR.
-- Different scenarios about spatial heterogeneity and the krill stock-recruit relationship (steepness) were tested.
-- There are various ways to check the performance of the models. More data sources, more parameters!
-- The retrospective pattern of the aggregate model s1 is greater than s01, but the RMSE indicates that model
-s01 has less predictive power
+# CONCLUSION AND FUTURE WORK
+
+- The main challenge in this analysis has been to identify sources of information about krill population dynamic (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
+This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions and evaluate the impact of this on population estimates.
+
+- Regarding to spatial heterogeneity, this can be considered in two ways in a modelling approach.
+First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer this considerations to the management for decision making.
+Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established in to population assessed.
+For this, it is necessary.
+have a spatially explicit model, which was not our case but can be considered a future approach to be analysed.
+
+- Ecosystem considerations for a model integrated with stock assessment are vital.Although SS3 do not modeling the ecosystem variables, allow these variables to be incorporated to model krill population dynamics.SS3 presents characteristics that allow the incorporation of this kind of variables.
+This not mean, however, this model is not modeling those components but rather they serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management that in the new strategy proposed by CCAMLR.
+
+- We consider that an integrated modeling of the krill adding the available information sources (fishery and surveys), as well as, the spatial heterogeneity and ecosystem variables is on the right path of the recommendation that has emerged in the last 10 years with respect to population dynamics and its link with the management procedure carried out by CCAMLR.
 
@@
 
-- Supplementary Material 3:  Length composition templates preparation to SS3 from AMLR data [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
+- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)
 
-- Supplementary Material 4:  Correlation analysis between.
-biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
+- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
+
+- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
 
@@
 ```
-\begin{landscape}
-```{r path, fig.cap="Framework path to stock assessment model in krill in WAP"}
-knitr::include_graphics('Figs/pathmod.png')
+
+```{r datas2, out.width='70%', fig.cap="Data presence by year for each fleet and data type used in s1 scenario"}
+knitr::include_graphics('Figs/data_plot_s2.png')
 ```
-
-\end{landscape}
 
@@
 
-```{r peng,  out.width='100%', fig.cap="Primary penguin species time series abundant populations in colonies located at 48.1"}
-knitr::include_graphics('Figs/Cunt_Penguins.png')
+```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
+knitr::include_graphics('Figs/pathmod.png')
 ```
+
 \end{landscape}
-
-```{r datas2, out.width='70%', fig.cap="Data presence by year for each fleet and data type used in s1 scenario"}
-knitr::include_graphics('Figs/data_plot_s2.png')
-```
 
@@
 
-```{r fitcom, out.width='50%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
-par(mfrow=c(3,1))
-knitr::include_graphics(c('s01/plots/comp_lenfit__aggregated_across_time.png',
-                          's1/plots/comp_lenfit__aggregated_across_time.png',
-                          's7/plots/comp_lenfit__aggregated_across_time.png'))
-
-```
-
 \begin{landscape}
 
-```{r resall, out.width='20%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes."}
+```{r resall, out.width='20%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
 par(mfrow=c(2,5))
@@
 
-```{r likepro, out.width='60%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
+```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
+knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')
+
+```
+
+```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
+knitr::include_graphics('Figs/retros1.png')
+```
+
+```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
+par(mfrow=c(2,1))
+knitr::include_graphics(c('Figs/rmse21.png',
+                          'Figs/rmselens1.png'))
+```
+
+```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
 knitr::include_graphics('Figs/Profile_s1.png')
@@
 
-\begin{landscape}
-```{r acf, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-31-1.jpeg')
-```
-\end{landscape}
-
-\begin{landscape}
-```{r recdev, out.width='70%', fig.cap="Cross correlation in autoregresive analysis"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-30-1.jpeg')
-```
-
-\end{landscape}
-
-```{r longterm, out.width='90%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
-knitr::include_graphics('Figs/longtermsb.png')
-```
-
-```{r mainvar, fig.cap="Main variables outputs from estimation in s1 model krill in WAP"}
+```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
 out_s2 <- read_excel("DataKrill.xlsx", 
@@
 
+\begin{landscape}
+```{r recdev, out.width='100%', fig.cap="Recruit deviation for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-42-1.jpeg')
+```
+\end{landscape}
+
+\begin{landscape}
+```{r acf, out.width='100%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-41-1.jpeg')
+```
+
+\end{landscape}
+
+```{r likeall, out.width='70%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
+knitr::include_graphics('Figs/Likelihoodtotal.png')
+```
+
+```{r longterm, out.width='90%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
+knitr::include_graphics('Figs/longtermsb.png')
+```
+
 \newpage
@@
 
-```{r comparision, out.width='40%', fig.show='hold', fig.cap="perfomance in spawning biomass in different scenarios"}
+```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s101, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
 par(mfrow=c(2,2))

### 2024-06-06 index.Rmd
add last changes in html

@@
                       warning = FALSE,
+                      fig.height = 3,
+                      fig.width = 5,
                       fig.align = 'center',
@@
 library(mixR)
-```
-
-```{r message=FALSE, warning=FALSE}
+library(readxl)
+```
+
+```{r message=FALSE, warning=FALSE, echo=TRUE}
 dir01 <- here("s01") # agreggate data (no spatial diferences)
@@
 dir4<-here("s4") # 
-dir5<-here("s5") # whitout S-R
-dir6<-here("s6") # whitout S-R
-dir7<-here("s7") # whitout S-R
+dir5<-here("s5") # 
+dir6<-here("s6") # 
+dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
 dirtest<-here("test")
-Figs <- here("Figs")# S2 set parametres EMM-2024/23 (Mardones)
+Figs <- here("Figs")# S
 ```
@@
  parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-
+```
+
+```{r eval=FALSE}
 wb <- createWorkbook()
@@
   kbl(booktabs = T,
-      format = "latex",
+      format = "html",
       position="ht!",
-    caption = "\\label{Tab1}Input parameters for the initial SS3 model of krill.
+    caption = "Input parameters for the initial SS3 model of krill.
 Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
@@
 
-```{r eval=FALSE, message=F, include=FALSE}
+```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
 # Lista de directorios para correro tordos juntos
@@
 
-```{r}
+```{r eval=FALSE}
 total01 <- as.data.frame(base.model01$timeseries[1:50,c(2,5)]) %>%
@@
 
-```{r fig.height=7}
+```{r }
 SSplotData(base.model01, subplot = 1, 
@@
 
-```{r}
+```{r fig.height=6}
 SSplotSelex(base.model01,
@@
 ```
+```{r}
+SSplotComps(base.model1,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+```
 
@@
             maxcols = 5)
-
 SSplotComps(base.model1,
@@
 ## Residual consistency
+
 ```{r}
+SSplotRunstest(base.model1,
+               subplots = "len",
+               add=T)
+```
+
+```{r eval=FALSE, fig.width=4, fig.height=5}
 SSplotRunstest(base.model01,
@@
 ```{r}
+SSplotRunstest(base.model1,
+               subplots = "cpue",
+               add=T)
+```
+
+```{r eval=FALSE}
 SSplotRunstest(base.model01,
@@
 ```{r}
+hci = SSplotHCxval(retroSummary1, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+```
+
+```{r eval=FALSE}
 hci = SSplotHCxval(retroSummary01, 
@@
 ```{r}
+mvln = SSdeltaMVLN(base.model1,
+                   Fref = "Btgt", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+```
+
+```{r eval=FALSE}
 mvln = SSdeltaMVLN(base.model01,
@@
 
-## Verosimilitud
+## Likelihood
 
@@
 
-## SSplotComparisons()
-
 ```{r eval=FALSE, echo = FALSE}
@@
 
-addWorksheet(wb, "variable_s2")
-writeData(wb, "variable_s2", outps1)
-
-# Guardar el workbook
-saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps1 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9)#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
 ```
@@
 
-comp01_1_7 <- SSplotComparisons(summaryoutput01_1_7,
+SSplotComparisons(summaryoutput01_1_7,
                   legendlabels = c(
@@
 
-```{r}
+```{r fig.height=6, fig.width=8}
 
@@
 
-```{r}
+```{r fig.width=8, fig.height=7}
 recs01 <- base.model01$recruit$pred_recr
@@
 
-```{r}
+```{r fig.width=8, fig.height=7}
 dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 

### 2024-06-09 SS3_Krill_481.Rmd
commit last submmit WG CCAMLR

@@
 ---
-title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
+title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
 author:
@@
 
-The integration of available source of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in a integrated stock assessment model.
-Being able to identify that sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management nd decision making.
-On the other hand, spatial consideration and ecosystem aspect are vital components to model dynamics population on krill.
+The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in an integrated stock assessment model.Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making.
 This approach has three main objectives.
@@
 On the other hand, consider the spatial heterogeneity of the krill population structure.
-Finally, evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
+Finally, to evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
 Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
 
-*Keywords: Krill populations, dynamic population, integrated model, SS3, spatial implicit, ecosystem consideration, management, CCAMLR.*
+*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem consideration, management, CCAMLR.*
 
@@
 Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
-On another hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
+On the other hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
 Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-As part of its mandate, CCAMLR promote sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
-This regional organization has tried to conducts different modelling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
-Various stock assessments approach on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
+As part of its mandate, CCAMLR promotes sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
+CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
+Various stock assessment approaches on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
 Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
@@
 However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-There is a growing consensus on the need for use stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
+There is a growing consensus on the need for using stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
 Integrating the complexity of krill population dynamics presents several challenges.
@@
 Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
-Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows build a integrated model to get populations signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
-In response to this need, we propose a conceptual model to conducted a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
+Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
+In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
 
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is block 48.1 subarea and strata subdivision inside to account for spatial heterogeneity into strata sensu @Dornam2021 (Figure \@ref(fig:mapa)).
+The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows you to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
+Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
 
@@
 These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations.
-This include data on abundance by strata, and biological characteristics of krill collected through net sampling.
-All this data was process in @Dornam2021; and length compositions from AMLR survey (provide by Ecosystem Science Division, NOAA) was handling according to the evaluation template requirements (Suppl.
+This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling.
+All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
 Mat.
-2); 3) For this analysis, chlorophyll (Chl) was assessed as a primary environmental driver.
+2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
 Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
 Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
-This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and dynamics on krill in the WAP area.
-To analyze the impact of chlorophyll on krill recruitment, We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + SIC + SST).
-The model included spatial component (cell) as random effect (formula: ~1 | cell) (Suppl.
-Mat 3).
-Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020].
-All this analysis could be found in this [repo](https://mauromardones.github.io/Krill_Length_Cor/); 4) Predator abundance data from "MAPPPD".
+This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP.
+We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)).
+The model included spatial component (cell) as random effect (formula: ~1 | cell).
+Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl.
+Mat 3); 4) Predator abundance data from "MAPPPD".
 Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
@@
 
-A conceptual model for krill in the Antarctic Peninsula includes the five spatial strata described previously.
-Despite differences among strata, krill population is closed to emigration and immigration, and their spatial distribution and interactions within this single, cohesive population (Figure \@ref(fig:concem)).
-Ecological and environmental variables that have impact in krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
-The model considers interanual fluctuations to dynamics of growth, maturity and reproduction, with a spawning pulse by year.
+A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
+Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
+The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.
 
@@
 SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
 The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-The analysis of results and outputs we use R tools (@R-base) and the graphical interface of the SS3 it was through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
+All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
 Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
@@
 One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock].
-In the context of krill, the estimated level of recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
+The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
 
@@
 \(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
+\(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable.
 Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
@@
 
-Furthermore, this integrated model will take into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
+Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
 These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
@@
 This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics.
-By incorporating information on predator  abundance and feeding ecology, our model will provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
-In this analysis we incorporated penguin population index as predator *"fleets"* how  we describe previously.
+By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
+In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously.
 This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.
@@
 
-To consider the uncertainty associated with of krill nature dimension, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
-The methodology involves comparing the statistical performance of the model regarding this configurations.
-In Table 1 we have eight scenarios to test different option in modeling about main consideration in assessment of krill population, and for comparative propose, we consider `s1` as base model.
+To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
+The methodology involves comparing the statistical performance of the model regarding these configurations.
+We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).
 
@@
 |:------------:|:---------------------------------------------------------|
-|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate 48.1                |
-|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
-|    s2    | "s1" without S-R relation                                                            |
-|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
+|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental data, aggregated at 48.1 level               |
+|    s1    | Fishery and Survey (AMLR) Length data, CPUE, Catch by strata. Predator and Env data |
+|    s2    | "s1" without Stock-Recruit relation                                      |
+|    s3    | "s1" Beverton & Holt S-R relation weak (0.9 steepness)                   |
 |    s4    | "s1" BH S-R relation strong (0.6 steepness)   |
@@
 
-For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of parameters life history respectively according to @Mardones2024par (unpublished).
+For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).
 
@@
 All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
-About goodness-of-fit, Figure \@ref(fig:res1)) and (Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
+About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
 In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
@@
 For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000.
-However, there is variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
+However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
 For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
-The residuals from the model fit to the sizes suggest that the model captures the inter annual and inter-fleet variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
+The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
 
@@
 Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
-@Carvalho2021b indicate that values of rho parametr that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
-To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we use the joint-index residual.
-Overall, `s1` have a good performance when evaluating the lengths compositions (14.2%), however  RMSE show a low predictive power with respect to the indices (71.3%).
+@Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
+To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual.
+Overall, `s1` had a good performance when evaluating the lengths compositions (14.2%), however  RMSE showed a low predictive power with respect to the indices (71.3%).
 A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).
@@
 
-In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions.
-We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years[@Perry2020] (Figure \@(fig:recdev)).
+In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions (Figure \@(fig:recdev)).
+We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @@McBride2021].
 
-Regarding to autocorrelation in krill recruitment in time series analized by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
+Autocorrelation in krill recruitment in time series analyzed by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
 Models `s3` and `s4` show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
-Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary,
-`s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness. 
-(Figure \@(fig:acf)).
+Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary, `s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness (Figure \@ref(fig:acf)).
 
@@
 
-With respect to population variables such as spawning biomass, we identify differences in the different scenarios with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios.
+We identify differences in the different scenarios related with population variables with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios.
 This is relevant given that these scenarios consider low spatial dependence, density-dependence spawning biomass with respect to recruitment, which is plausible for resources such as krill where environmental dynamics shape population dynamics (Figure \@ref(fig:longterm)).
@@
 The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
-Development of an integrated stock assessment model like this could be a support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
+The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
 However, integrated models are not new to krill in SO.
 For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
-In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modelling strategies currently in use to manage krill"* and he stated that the an integrated modelling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
+In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
 
@@
 Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
-Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Ossilation Southern (ENSO).
+Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Oscillation Southern (ENSO).
 In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.
@@
 From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
-There is many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
-For this, we consider penguin colonies population as one of the most powerful elements of natural predation on krill because have direct correlations between this type of taxonomic group and krill [@Kruger2021].
-The intention to incorporate elements of predation in a stock is necessary to improve the modelling, because did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
+There are many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
+For this, we consider the penguin  population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill  [@Kruger2021].
+The intention to incorporate elements of predation in a stock is necessary to improve the modeling, because it did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
 
@@
 In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
-In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in subarea 48.1 of the Western Antarctic Peninsula (WAP).
+In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
 
@@
 - The main challenge in this analysis has been to identify sources of information about krill population dynamic (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
-This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiples natural dimensions and evaluate the impact of this on population estimates.
+This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiple natural dimensions and evaluate the impact of this on population estimates.
 
-- Regarding to spatial heterogeneity, this can be considered in two ways in a modelling approach.
-First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer this considerations to the management for decision making.
-Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established in to population assessed.
+- Regarding to spatial heterogeneity, this can be considered in two ways in a modeling approach.
+First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer these considerations to the management for decision making.
+Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established into the population assessed.
 For this, it is necessary.
@@
 
-- Ecosystem considerations for a model integrated with stock assessment are vital.Although SS3 do not modeling the ecosystem variables, allow these variables to be incorporated to model krill population dynamics.SS3 presents characteristics that allow the incorporation of this kind of variables.
-This not mean, however, this model is not modeling those components but rather they serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management that in the new strategy proposed by CCAMLR.
+- Ecosystem considerations for a model integrated with stock assessment are vital.
+Although SS3 does not model the ecosystem variables, it allows for these variables to be incorporated to model krill population dynamics.
+SS3 presents characteristics that allow the incorporation of this kind of variable.
+This ecosystem information into stock assessment serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management in the new strategy proposed by CCAMLR.
 
@@
 
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, GERLASHE= Gerlashe strait, JOIN= Joinville Island, SSWI= South West)"}
+```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
 knitr::include_graphics('Figs/481.png')
@@
 
-```{r datas2, out.width='70%', fig.cap="Data presence by year for each fleet and data type used in s1 scenario"}
+```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
 knitr::include_graphics('Figs/data_plot_s2.png')
@@
 
-\begin{landscape}
-
-```{r resall, out.width='20%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
-par(mfrow=c(2,5))
+```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
+par(mfrow=c(5,2))
 knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
@@
 ```
-\end{landscape}
 
@@
                                   "condensed"),
-                full_width = FALSE,
-                font_size=9)#%>% 
+                full_width = TRUE,
+                font_size=6 )#%>% 
   #pack_rows(index = c("Estimation" = 1,
@@
 
-\begin{landscape}
-```{r recdev, out.width='100%', fig.cap="Recruit deviation for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-42-1.jpeg')
-```
-\end{landscape}
-
-\begin{landscape}
-```{r acf, out.width='100%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-41-1.jpeg')
+\newpage
+```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
 ```
 
-\end{landscape}
+```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
+```
 
-```{r likeall, out.width='70%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
+```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
 knitr::include_graphics('Figs/Likelihoodtotal.png')
@@
 
-```{r longterm, out.width='90%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
+```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
 knitr::include_graphics('Figs/longtermsb.png')
@@
 
-```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s101, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
+```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
 par(mfrow=c(2,2))

### 2024-06-18 index.Rmd
change name in WP 2024

@@
 
-                 "Parameters WG SAM 2024/23"),
+                 "Parameters WG SAM 2024/27"),
                  labels = c("Year", 

### 2024-08-26 SS3_Krill_481.Rmd
saving previus chgs

@@
 author:
-  - Mardones Mauricio^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
+  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
   - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]

### 2024-08-26 Corrections_SAM_SS3_Krill_481.Rmd
add WR to FSA

--- 2024-08-26 SS3_Krill_481.Rmd
@@
 ---
-title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+title: "FSA Working Report Stock Assessment Krill WAP"
 author:

### 2024-08-30 Corrections_SAM_SS3_Krill_481.Rmd
add 2nd file toFSA

@@
   - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
-
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
                       tidy=TRUE)
-#XQuartz is a mess, put this in your onload to default to cairo instead
-options(bitmapType = "cairo") 
-# (https://github.com/tidyverse/ggplot2/issues/2655)
-# Lo mapas se hacen mas rapido
 ```
@@
 
-The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in an integrated stock assessment model.Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making.
-This approach has three main objectives.
-Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance.
-On the other hand, consider the spatial heterogeneity of the krill population structure.
-Finally, to evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
-Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
+This study addresses key recommendations highlighted in the preliminary Report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024) regarding to analysis of integrating diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, especially in the face of environmental change and fisheries management challenges.
+This work systematically incorporates fishing, environmental, and ecological variables while accounting for the spatial heterogeneity of the krill population structure.
+Additionally, it evaluates the impact of biological and structural assumptions on model performance, calculates the harvest rate, and refines the proposed management procedures in a both conceptual and tactical approach.
+By acknowledging and integrating these diverse data sources, the model provides valuable insights into krill population dynamics, and continue contributing to the enhancement of CCAMLR fishery management and conservation strategies in the region.
 
@@
 \newpage
+
+# RECOMMENDATIONS SAM - 23
+
+2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016).
+It also noted that this would avoid the approximations required when converting length data to age data.
+The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).
+It encouraged the authors to provide standard model diagnostics, similar to those presented for toothfish assessments, to facilitate understanding of model performance (e.g., WG-SAM-2023, paragraphs 6.33 and 6.34).
+
+2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g.
+changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system.
+It further highlighted the need to discuss the development of a standardised data collection plan to support the ongoing revision of the krill fishery management approach.
+
+2.5 Dr S.
+Kasatkina (Russian Federation) noted that data on krill length and biological composition from catches of fishing vessels will be not suitable for such a modelling approach, and recalled that comparisons of krill length composition from catches of fishing trawls and catches of scientific trawls within the same fishing ground revealed significant differences (WG-ASAM-2021/03).
+Moreover, Dr Kasatkina noted that there were significant differences in the length composition of catches between fishing vessels and these differences are random in nature, which may be due to both the selective properties of commercial trawls, different fishing methods (continuous and traditional fishing) as well as the efficiency of krill sampling by observers at-sea (WG-ASAM-2021/03; WG-EMM-2024/37).
+Dr Kasatkina noted that this modelling effort required clarity regarding the interaction between the fishery and krill-dependent predators, which requires regular observations to study the spatial overlap of fishing zones and predator foraging zones and could be accompanied by krill distribution patterns (for example, such complex ecosystem observations were provided on RV Atlantida in 2020; SC-CAMLR-42/07).
+Dr Kasatkina recalled that the spatial and temporal dynamics of groups from the Bellingshausen and Weddell Seas (Fach et al, 2002; Murphy et al, 2004; WG-EMM-2024/43; WG-EMM-2024/39) and noted that without data on krill transport and standardised acoustic surveys of krill, it is impossible to assess the influence of factors such as spatial heterogeneity and life history parameters on key krill population variables.
+
+2.6 The Working Group encouraged the authors to provide a progression of model implementations from simple to more complex to facilitate understanding and evaluate the evidence for the model assumptions.
+It further noted the presence of patterns in the residuals shown in the paper (Figures 5 and 6) which warranted further investigation, as well as the need to assess the realism of some parameter values used in the model.
+The Working Group also indicated that this work would benefit from taking into consideration recent findings by SKEG regarding the krill stock hypothesis (e.g.
+WG-EMM-2024/39).
+
+- Harvest Rate
+
+- Diagram management procedure
 

### 2024-08-30 index.Rmd
saving previus chgs
add 2nd file toFSA

@@
                       warning = FALSE,
-                      fig.height = 3,
-                      fig.width = 5,
+                      fig.height = 5,
+                      fig.width = 7,
                       fig.align = 'center',

### 2024-09-04 Corrections_SAM_SS3_Krill_481.Rmd
adding layaout in correction file

@@
 ---
-title: "FSA Working Report Stock Assessment Krill WAP"
+title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+subtitle: "WG-FSA 2024. Working Paper"
 author:
@@
 
-# ABSTRACT
+# BACKGROUND
 
-This study addresses key recommendations highlighted in the preliminary Report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024) regarding to analysis of integrating diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, especially in the face of environmental change and fisheries management challenges.
-This work systematically incorporates fishing, environmental, and ecological variables while accounting for the spatial heterogeneity of the krill population structure.
-Additionally, it evaluates the impact of biological and structural assumptions on model performance, calculates the harvest rate, and refines the proposed management procedures in a both conceptual and tactical approach.
-By acknowledging and integrating these diverse data sources, the model provides valuable insights into krill population dynamics, and continue contributing to the enhancement of CCAMLR fishery management and conservation strategies in the region.
-
-*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem consideration, management, CCAMLR.*
-
-\newpage
-
-# RECOMMENDATIONS SAM - 23
-
-2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016).
-It also noted that this would avoid the approximations required when converting length data to age data.
-The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).
-It encouraged the authors to provide standard model diagnostics, similar to those presented for toothfish assessments, to facilitate understanding of model performance (e.g., WG-SAM-2023, paragraphs 6.33 and 6.34).
-
-2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g.
-changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system.
-It further highlighted the need to discuss the development of a standardised data collection plan to support the ongoing revision of the krill fishery management approach.
-
-2.5 Dr S.
-Kasatkina (Russian Federation) noted that data on krill length and biological composition from catches of fishing vessels will be not suitable for such a modelling approach, and recalled that comparisons of krill length composition from catches of fishing trawls and catches of scientific trawls within the same fishing ground revealed significant differences (WG-ASAM-2021/03).
-Moreover, Dr Kasatkina noted that there were significant differences in the length composition of catches between fishing vessels and these differences are random in nature, which may be due to both the selective properties of commercial trawls, different fishing methods (continuous and traditional fishing) as well as the efficiency of krill sampling by observers at-sea (WG-ASAM-2021/03; WG-EMM-2024/37).
-Dr Kasatkina noted that this modelling effort required clarity regarding the interaction between the fishery and krill-dependent predators, which requires regular observations to study the spatial overlap of fishing zones and predator foraging zones and could be accompanied by krill distribution patterns (for example, such complex ecosystem observations were provided on RV Atlantida in 2020; SC-CAMLR-42/07).
-Dr Kasatkina recalled that the spatial and temporal dynamics of groups from the Bellingshausen and Weddell Seas (Fach et al, 2002; Murphy et al, 2004; WG-EMM-2024/43; WG-EMM-2024/39) and noted that without data on krill transport and standardised acoustic surveys of krill, it is impossible to assess the influence of factors such as spatial heterogeneity and life history parameters on key krill population variables.
-
-2.6 The Working Group encouraged the authors to provide a progression of model implementations from simple to more complex to facilitate understanding and evaluate the evidence for the model assumptions.
-It further noted the presence of patterns in the residuals shown in the paper (Figures 5 and 6) which warranted further investigation, as well as the need to assess the realism of some parameter values used in the model.
-The Working Group also indicated that this work would benefit from taking into consideration recent findings by SKEG regarding the krill stock hypothesis (e.g.
-WG-EMM-2024/39).
-
-- Harvest Rate
-
-- Diagram management procedure
-
-# INTRODUCTION
-
-Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
-On the other hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
-Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-As part of its mandate, CCAMLR promotes sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
-CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
-Various stock assessment approaches on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
-Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
-Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013].
-However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-There is a growing consensus on the need for using stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
-Integrating the complexity of krill population dynamics presents several challenges.
-Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension.
-Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics.
-Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
-Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
-In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
+This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) [@Mardones2023] regarding the integration of diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, particularly in the context of environmental change and fisheries management challenges.
+The main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.Each observation will be explained in theoretical terms and, where possible, analyzed in detail.
+This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups.
+These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings.
+It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.
 
@@
 
-## Comparision outputs between scenarios
+\newpage
 
-In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions (Figure \@(fig:recdev)).
-We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @@McBride2021].
+# RECOMMENDATIONS SAM - 23
 
-Autocorrelation in krill recruitment in time series analyzed by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
-Models `s3` and `s4` show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
-Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary, `s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness (Figure \@ref(fig:acf)).
+The main recommendations were extracted from @CCAMLRWGSAM2024
 
-To evaluate consistency of the source of information used in each of the tested scenarios, we verify its influence on the estimate of R0 through likelihood components and, like scenario 1, we verify that all scenarios show a high dependence on the composition lengths of lengths with respect to the estimate of the variable (as show the (Figure \@ref(fig:likeall)).
+*2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016).
+It also noted that this would avoid the approximations required when converting length data to age data.
+The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).
+*
 
-We identify differences in the different scenarios related with population variables with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios.
-This is relevant given that these scenarios consider low spatial dependence, density-dependence spawning biomass with respect to recruitment, which is plausible for resources such as krill where environmental dynamics shape population dynamics (Figure \@ref(fig:longterm)).
+- **Response**:
 
-We carried out a specific comparison between the most divergent scenarios that we tested in this case, `s1`,  `s01` and `s7` which according to the description are the aggregated spatial scenario, disaggregate spatial scenario and scenario with new set of parameters, respectively.
-All with ecosystem components such as predators and an environmental variable.In this sense, the impacts and difference on the estimates of recruitment,  R0, virginal biomass and relative biomass are high (Figure \@ref(fig:comparision)), demonstrating the impact of initial considerations and assumptions on population variables estimation.
+We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data.
+We recognize that using a length-based model could reduce the need for approximations when converting length data to age data.
+We plan to explore this alternative by developing a length-based model to compare its performance and results with the current age-based model.
+Additionally, we will incorporate standard model diagnostics, as suggested, similar to those used in toothfish assessments, to ensure transparency and facilitate understanding of model performance.
+
+Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
+In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap.
+Length offers insights into the age structure of the population, as there is a correlation between age and length.
+However, it's still necessary to convert catch-at-length data into catch-at-age data.
+Several methods are available to achieve this [@ICCA2003].
+
+#### Age-Length Keys (ALKs)
+
+Put simple, age-length keys (ALKs) are generated through the ageing of a sub-sample of the population, and used to convert larger length samples from a population into ages.
+ALKs describe distributions of size for each age, and the relative number of individuals at each age, i.e.
+they represent a matrix detailing the probability that a fish of a given length is of a particular age.
+Once such a key is available, samples of fish that were only measured for length can be distributed over age groups according to the key.
+The use of ALKs assumes that the sample of aged fish and the sample of fish measured for length are simple random samples from the same population.
+Then, the probability that a fish is of a particular age, given its length is the same for both samples.
+The ALK should generally be applied to length data from the same time period, since variability in recruitment and survivorship at age will change the age-length composition over time, and hence the number of survivors at age used to weight the size at age compositions will vary.
+The ALK may need to be seasonal if growth is temporally distinctive or seasonal migration occurs.
+A single ALK should only be applied to size data from a number of years if growth is reasonably stationary, and the approach of Kimura and Chikuni (1987) is used.
+Suitable justification should be presented for multi-seasonal or multi-annual application of ALKs.
+It must also be noted that the application of an ALK derived from a single time of year can cause serious bias if used to compute catch at age for the entire year.
+Age-at-length data from hard parts should be combined into suitable length groups for required gears, time periods and locations.
+The size of these length groups will depend on the spread of lengths found within the catch length frequencies, the growth rate of the species, and the variability in length-at-age.
+Length-stratified sampling is required to ensure that the required number of fish (determined by the variability of length-at-age) is available over all length groups.
+Developing ALKs represent is laborious, and hence the optimum collection of information is desirable.
+Formulae exist to estimate the number of age determinations and length measurements necessary to guarantee a given level of accuracy.
+Oeberst (2000) developed a universal cost function for ALKs, for example [@ICCA2003].
+
+The proportion at age is calculated as:
+$$
+\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
+$$
+
+The ALK for the time period is raised to the length distribution for that time period:
+
+$$
+\text{Raised numbers at age by length group} = \text{Numbers at length} \times \text{Proportion at age for that length}
+$$
+
+If the ALK does not contain data for all the length groups in the length distribution then data in the length distribution may be assigned to adjacent length groups where data are present in the ALK.
+An appropriate age- length distribution by gear group and time period is then produced.
+Care must be taken since considerable biases can result, particularly at large lengths where individuals may be distributed over a wide range of ages.
+
+The numbers at age by length group are summed over the length range to give numbers at age.
+The variances are also summed over the length groups and the two components labelled variance due to ageing and variance due to length sampling.
+This gives the age composition for the required time period.
+Numbers at age for all gears can be calculated as:
+
+$$
+\sum N_a \times \left(\frac{W_{ct}}{W_{cs}}\right)
+$$
+where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
+sampled commercial catch weight.
+Variance due to ageing of numbers at age for all gears can be calculated as:
+
+$$
+\sum \text{Var}_a
+$$
+where $\sum Var_a$  is the sum of variances due to ageing.
+Variance due to length sampling of numbers at age can be calculated as:
+
+$$
+\sum \text{Var}_l
+$$
+where $\sum Var_a$  is the sum of variances due to length sampling.
+The variances should be raised by:
+
+$$
+\frac{W_{ct}}{W_{cs}}
+$$
+where W~ct~ is the total commercial catch weight, and W~cs~ is the sampled commercial catch weight.
+
+The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
+The method can work well where the distributions of size for each age are close to those in the length frequency, but convergence can be slow.
+The reader is referred to the paper for more information.
+
+Then Numbers-at-age in the model were converted to numbers-at-length (mm) using a von Bertalanffy growth relationship withestimated parameters L~inf~, k, and v(Eq.
+(A.6)).
+This length-at-age relationship, mediated by survey catchability and selectivity (Eqs.(A.2) and (A.3)), was used to compare survey biomasses with model estimates of vulnerable biomass (Eq.
+(A.4)).
+Weight-at-length was assumed known and calculated using Eq.
+3 from Hewitt et al.
+(2004)(W(L) = 2.236 × 10-6L3.314).
+(REF Kindzey)
+
+Another way that we implemented in `SS3` is regarding *platoons*.
+This approach allows for the exploration of survivorship that varies with size.
+A value of 1 will not create additional groups.
+Odd-numbered values (e.g., 3, 5) will divide the overall morph into that many groups, resulting in smaller, larger, and average growth groups.
+Increasing the number of groups slows down the model's performance, so it's recommended not to exceed a value of 5.
+The fraction of each morph assigned to each group can be either user-defined or approximated using a normal distribution.
+When multiple groups are designated, an additional input is required to specify the ratio of variability in size-at-age between groups versus within groups.
+This ratio is used to distribute the total growth variability.
+The size-at-age for each group is then calculated as a factor (based on the between-group and within-group variability) times the size-at-age of the central morph, which is determined from the growth parameters specific to that Growth Pattern x Sex combination [@methot2020stock].
+
+The age specific rates of fishing mortality can be linked to length specific selectivity.
+If growth platoons are invoked, then each platoon can have different age specific F due to interaction of length selectivity with their different size at age (Methot, *com pers*)
+
+In `SS3` colud be conditioned in this way;
+
+```{r eval=FALSE, echo=TRUE}
+#_N_Growth_Patterns (Growth Patterns, Morphs, Bio Patterns, GP are terms used interchangeably in SS)
+1  
+#_N_platoons_Within_GrowthPattern
+3 
+ #_Platoon_between/within_stdev_ratio (no read if N_platoons=1)
+0.4
+#vector_platoon_dist_(-1_in_first_val_gives_normal_approx)
+c(0.237, 0.464, 0.237)
+```
+
+```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
+knitr::include_graphics('Figs/Plattoons.png')
+```
+
+Regarding to observation in SAM -2023 (paragraph 4.3) that say;
+
+*4.3 its usefulness in providing an additional approach to assessing krill stock status, it recalled that the development of an integrated stock assessment for krill was considered desirable within three to five years"* but considering this;
+
+*(i) the impact of using an age-based or a length-based implementation could be assessed, noting that a preservation of the source data would be preferred (i.e.
+the conversion between length composition and age composition should be considered, with conversions from length to age to length avoided)*
+
+*It encouraged the authors to provide standard model diagnostics, similar to those presented for toothfish assessments, to facilitate understanding of model performance (e.g., WG-SAM-2023, paragraphs 6.33 and 6.34).*
+
+*2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g.
+changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system.
+It further highlighted the need to discuss the development of a standardized data collection plan to support the ongoing revision of the krill fishery management approach.*
+
+  - **Response**:
+
+We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system.
+We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs.
+Moreover, we acknowledge the need to reconsider the assumption of the Peninsula being a closed system, and we will explore alternative approaches to address this.
+Moving forward, we will engage in discussions to develop a standardized data collection plan that aligns with the revision of the krill fishery management approach, ensuring that future data inputs are consistent and reliable.
+
+*2.5 Dr S.
+Kasatkina (Russian Federation) noted that data on krill length and biological composition from catches of fishing vessels will be not suitable for such a modelling approach, and recalled that comparisons of krill length composition from catches of fishing trawls and catches of scientific trawls within the same fishing ground revealed significant differences (WG-ASAM-2021/03).
+Moreover, Dr Kasatkina noted that there were significant differences in the length composition of catches between fishing vessels and these differences are random in nature, which may be due to both the selective properties of commercial trawls, different fishing methods (continuous and traditional fishing) as well as the efficiency of krill sampling by observers at-sea (WG-ASAM-2021/03; WG-EMM-2024/37).
+Dr Kasatkina noted that this modelling effort required clarity regarding the interaction between the fishery and krill-dependent predators, which requires regular observations to study the spatial overlap of fishing zones and predator foraging zones and could be accompanied by krill distribution patterns (for example, such complex ecosystem observations were provided on RV Atlantida in 2020; SC-CAMLR-42/07).
+Dr Kasatkina recalled that the spatial and temporal dynamics of groups from the Bellingshausen and Weddell Seas (Fach et al, 2002; Murphy et al, 2004; WG-EMM-2024/43; WG-EMM-2024/39) and noted that without data on krill transport and standardised acoustic surveys of krill, it is impossible to assess the influence of factors such as spatial heterogeneity and life history parameters on key krill population variables.*
+
+- **Response**: We understand the concerns raised regarding the suitability of krill length and biological composition data from fishing vessels for this modeling approach.
+We agree that significant differences in length composition between commercial and scientific trawls, as well as among different vessels, pose challenges.
+To address this, we will review the selection criteria for the data used in our model and consider incorporating data exclusively from standardized scientific trawls to improve the reliability of the model outputs.
+We also acknowledge the importance of understanding the interaction between fisheries and krill-dependent predators.
+We will explore opportunities to integrate regular observations and data on the spatial overlap of fishing zones and predator foraging areas into the model, possibly leveraging existing ecosystem observations like those from RV Atlantida.
+
+*2.6 The Working Group encouraged the authors to provide a progression of model implementations from simple to more complex to facilitate understanding and evaluate the evidence for the model assumptions.
+It further noted the presence of patterns in the residuals shown in the paper (Figures 5 and 6) which warranted further investigation, as well as the need to assess the realism of some parameter values used in the model.
+The Working Group also indicated that this work would benefit from taking into consideration recent findings by SKEG regarding the krill stock hypothesis (e.g.
+WG-EMM-2024/39).*
+
+ - **Response**: We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions.
+We will outline a plan to present this progression in future work, beginning with a simpler model and gradually incorporating complexity.
+Furthermore, we acknowledge the presence of patterns in the residuals and the need to investigate these further.
+We will conduct a thorough analysis of the residuals to identify any underlying issues and reassess the realism of the parameter values used in the model.
+In addition, we will incorporate recent findings from SKEG regarding the krill stock hypothesis to ensure that our model assumptions are up-to-date and reflect the latest scientific understanding.
+
+*show  Harvest Rate*
+
+*Diagram management procedure*
 
@@
 In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
-
-\newpage
-
-# CONCLUSION AND FUTURE WORK
-
-- The main challenge in this analysis has been to identify sources of information about krill population dynamic (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
-This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiple natural dimensions and evaluate the impact of this on population estimates.
-
-- Regarding to spatial heterogeneity, this can be considered in two ways in a modeling approach.
-First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer these considerations to the management for decision making.
-Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established into the population assessed.
-For this, it is necessary.
-have a spatially explicit model, which was not our case but can be considered a future approach to be analysed.
-
-- Ecosystem considerations for a model integrated with stock assessment are vital.
-Although SS3 does not model the ecosystem variables, it allows for these variables to be incorporated to model krill population dynamics.
-SS3 presents characteristics that allow the incorporation of this kind of variable.
-This ecosystem information into stock assessment serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management in the new strategy proposed by CCAMLR.
-
-- We consider that an integrated modeling of the krill adding the available information sources (fishery and surveys), as well as, the spatial heterogeneity and ecosystem variables is on the right path of the recommendation that has emerged in the last 10 years with respect to population dynamics and its link with the management procedure carried out by CCAMLR.
 

### 2024-09-04 index.Rmd
adding layaout in correction file

@@
 dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
+dir8<-here("s8") # s1 test data weigthing
 dirtest<-here("test")
@@
 |    s7    | "s1" w/ set of parameters estimated in @EMM-204/32                                   |
+|    s8    | "s1" test data weighting                                                             |
 
@@
 r4ss::run(
-  dir = dir7,
+  dir = dir8,
   exe = "ss_osx",
@@
 
+#s8
+base.model8 <- SS_output(dir=dir8,
+                        covar=T,
+                        forecast=T)
+
+# html
+SS_plots(base.model8,
+         uncertainty=T,
+         datplot = T, 
+         png=T, 
+         aalresids = F,
+         btarg=0.75,
+         minbthresh=0.20, 
+         forecast=T)
+
 ```
@@
 ```{r}
-SSplotComps(base.model1,
+SSplotComps(base.model8,
             subplots = 1,
@@
 ```
+## Platoons analisis
+
+```{r}
+# Definir los valores del eje x
+x_values <- seq(0, 7, by=0.1)
+
+# Definir las medias y desviaciones estándar para los valores específicos
+means <- c( 2.9,  3.5, 4.1)
+std_devs <- c(0.65, 0.8, 0.65)
+
+# Crear un dataframe que contenga los valores de x, la media y las desviaciones estándar
+data <- data.frame(
+  x = rep(x_values, each = length(means)),
+  mean = rep(means, times = length(x_values)),
+  std_dev = rep(std_devs, times = length(x_values))
+)
+
+# Calcular las curvas de las desviaciones estándar
+data <- data %>%
+  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))
+
+# Graficar con ggplot
+ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
+  geom_line(size = 1) +
+  scale_linetype_manual(values = means,
+                        name ="Platoon") +
+  labs(x = "", y = "growth increment") +
+  theme_minimal()+
+  xlim(0,7)
+
+```
 

### 2024-09-08 Corrections_SAM_SS3_Krill_481.Rmd
wp2

@@
 ---
-title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
-subtitle: "WG-FSA 2024. Working Paper"
-author:
-  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
-  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
-  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
-  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
+title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+subtitle: "Appendix WP- SAM 2024/26"
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
                       dpi = 300,
-                      tidy.opts=list(width.cutoff=50), 
+                      tidy.opts=list(width.cutoff=80), 
                       tidy=TRUE)
@@
 library(r4ss)
+library(flextable)
 dir1<-here("s1")
@@
 
-# BACKGROUND
-
-This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) [@Mardones2023] regarding the integration of diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, particularly in the context of environmental change and fisheries management challenges.
-The main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.Each observation will be explained in theoretical terms and, where possible, analyzed in detail.
-This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups.
-These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings.
-It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.
-
 \newpage
 
-# METHODOLOGY
+# OBSERVATION SAM - 23
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
-The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
+The main recommendations were extracted from @CCAMLRWGSAM2024-
 
-## Data Sources
-
-Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling.
-All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
-Mat.
-2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
-Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
-Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
-This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP.
-We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)).
-The model included spatial component (cell) as random effect (formula: ~1 | cell).
-Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl.
-Mat 3); 4) Predator abundance data from "MAPPPD".
-Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
-
-## Conceptual Model
-
-A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
-Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
-The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.
-
-## Statistical model Stock Synthesis (SS3)
-
-Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
-Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
-SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
-SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
-Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
-Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
-
-### Environmental variable modelling in `SS3`
-
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock].
-The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
-
-\[
-E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-\]
-
-\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
-\(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
-However, the model does not automatically adjust for this.
-
-Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
-
-### Predator components modelling in `SS3`
-
-This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics.
-By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
-In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously.
-This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.
-
-All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
-
-\newpage
-
-## Stock Assessment Scenarios
-
-To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
-The methodology involves comparing the statistical performance of the model regarding these configurations.
-We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).
-
-| Scenario | Description                                                                          |
-|:------------:|:---------------------------------------------------------|
-|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental data, aggregated at 48.1 level               |
-|    s1    | Fishery and Survey (AMLR) Length data, CPUE, Catch by strata. Predator and Env data |
-|    s2    | "s1" without Stock-Recruit relation                                      |
-|    s3    | "s1" Beverton & Holt S-R relation weak (0.9 steepness)                   |
-|    s4    | "s1" BH S-R relation strong (0.6 steepness)   |
-|    s5    | "s1" BH S-R relation mid-strong estimated    |
-|    s6    | "s1" Ricker S-R relation estimated    |
-|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par   |
-
-: Scenarios to modelling dynamics in krill
-
-For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).
-
-\newpage
-
-# RESULTS
-
-## Diagnosis Base Model
-
-All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
-About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
-In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
-The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`.
-For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000.
-However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
-For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
-The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
-
-## Retrospective analysis
-
-Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
-@Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
-To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual.
-Overall, `s1` had a good performance when evaluating the lengths compositions (14.2%), however  RMSE showed a low predictive power with respect to the indices (71.3%).
-A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).
-
-## Likelihood profile in spatial Model `s1`
-
-For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources.
-The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro).
-The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.
-
-Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)
-
-\newpage
-
-# RECOMMENDATIONS SAM - 23
-
-The main recommendations were extracted from @CCAMLRWGSAM2024
+## Observation 1
 
@@
 It also noted that this would avoid the approximations required when converting length data to age data.
-The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).
-*
+The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).*
 
@@
 
-We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data.
-We recognize that using a length-based model could reduce the need for approximations when converting length data to age data.
-We plan to explore this alternative by developing a length-based model to compare its performance and results with the current age-based model.
-Additionally, we will incorporate standard model diagnostics, as suggested, similar to those used in toothfish assessments, to ensure transparency and facilitate understanding of model performance.
-
+We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
 Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
@@
 Length offers insights into the age structure of the population, as there is a correlation between age and length.
-However, it's still necessary to convert catch-at-length data into catch-at-age data.
-Several methods are available to achieve this [@ICCA2003].
+So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021].
+This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK).
+To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].
 
-#### Age-Length Keys (ALKs)
+In summary, ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples.
+ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
+Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias.
+Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003].
+Formulae exist to estimate the number of age determinations and length measurements is calculated as;
 
-Put simple, age-length keys (ALKs) are generated through the ageing of a sub-sample of the population, and used to convert larger length samples from a population into ages.
-ALKs describe distributions of size for each age, and the relative number of individuals at each age, i.e.
-they represent a matrix detailing the probability that a fish of a given length is of a particular age.
-Once such a key is available, samples of fish that were only measured for length can be distributed over age groups according to the key.
-The use of ALKs assumes that the sample of aged fish and the sample of fish measured for length are simple random samples from the same population.
-Then, the probability that a fish is of a particular age, given its length is the same for both samples.
-The ALK should generally be applied to length data from the same time period, since variability in recruitment and survivorship at age will change the age-length composition over time, and hence the number of survivors at age used to weight the size at age compositions will vary.
-The ALK may need to be seasonal if growth is temporally distinctive or seasonal migration occurs.
-A single ALK should only be applied to size data from a number of years if growth is reasonably stationary, and the approach of Kimura and Chikuni (1987) is used.
-Suitable justification should be presented for multi-seasonal or multi-annual application of ALKs.
-It must also be noted that the application of an ALK derived from a single time of year can cause serious bias if used to compute catch at age for the entire year.
-Age-at-length data from hard parts should be combined into suitable length groups for required gears, time periods and locations.
-The size of these length groups will depend on the spread of lengths found within the catch length frequencies, the growth rate of the species, and the variability in length-at-age.
-Length-stratified sampling is required to ensure that the required number of fish (determined by the variability of length-at-age) is available over all length groups.
-Developing ALKs represent is laborious, and hence the optimum collection of information is desirable.
-Formulae exist to estimate the number of age determinations and length measurements necessary to guarantee a given level of accuracy.
-Oeberst (2000) developed a universal cost function for ALKs, for example [@ICCA2003].
-
-The proportion at age is calculated as:
 $$
@@
 
-If the ALK does not contain data for all the length groups in the length distribution then data in the length distribution may be assigned to adjacent length groups where data are present in the ALK.
-An appropriate age- length distribution by gear group and time period is then produced.
-Care must be taken since considerable biases can result, particularly at large lengths where individuals may be distributed over a wide range of ages.
-
-The numbers at age by length group are summed over the length range to give numbers at age.
-The variances are also summed over the length groups and the two components labelled variance due to ageing and variance due to length sampling.
-This gives the age composition for the required time period.
+If the ALK lacks data for certain length groups, data from those groups may be assigned to adjacent ones in the ALK.
+This process generates an age-length distribution by gear and time period, but caution is needed to avoid significant biases, especially for larger lengths with a broad age range.
+The numbers at age are summed across length groups, combining variances from ageing and length sampling to produce the age composition for the specified period.
+Numbers at age for all gears can then be calculated.
 Numbers at age for all gears can be calculated as:
@@
 where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
-sampled commercial catch weight.
-Variance due to ageing of numbers at age for all gears can be calculated as:
+sampled commercial catch weight. Variance due to ageing of numbers at age for all gears can be calculated as:
 
@@
 $$
-where $\sum Var_a$  is the sum of variances due to ageing.
-Variance due to length sampling of numbers at age can be calculated as:
+where $\sum Var_a$  is the sum of variances due to ageing. Variance due to length sampling of numbers at age can be calculated as:
 
@@
 $$
-where $\sum Var_a$  is the sum of variances due to length sampling.
-The variances should be raised by:
+where $\sum Var_a$  is the sum of variances due to length sampling. The variances should be raised by:
 
@@
 The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
-The method can work well where the distributions of size for each age are close to those in the length frequency, but convergence can be slow.
-The reader is referred to the paper for more information.
+The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
+In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
+In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
+This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass (Eq.
+(A.4)).
+Weight-at-length was assumed known and calculated using @Maschette2020 parametres.
 
-Then Numbers-at-age in the model were converted to numbers-at-length (mm) using a von Bertalanffy growth relationship withestimated parameters L~inf~, k, and v(Eq.
-(A.6)).
-This length-at-age relationship, mediated by survey catchability and selectivity (Eqs.(A.2) and (A.3)), was used to compare survey biomasses with model estimates of vulnerable biomass (Eq.
-(A.4)).
-Weight-at-length was assumed known and calculated using Eq.
-3 from Hewitt et al.
-(2004)(W(L) = 2.236 × 10-6L3.314).
-(REF Kindzey)
+```{r AKL}
+alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz
 
-Another way that we implemented in `SS3` is regarding *platoons*.
-This approach allows for the exploration of survivorship that varies with size.
+# Convertir la matriz en un data.frame
+alk_df <- as.data.frame(as.table(alk_matrix))
+
+# Renombrar columnas para mayor claridad
+colnames(alk_df) <- c("Length", "TrueAge", "Value")
+
+# Asegúrate de que las columnas Length y TrueAge sean numéricas
+alk_df$Length <- as.numeric(as.character(alk_df$Length))
+alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
+
+# Asegúrate de que la columna Value sea numérica
+alk_df$Value <- as.numeric(alk_df$Value)
+
+# Crear el plot usando ggplot
+ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "blue") +
+  labs(x = "True Age",
+       y = "Length",
+       fill = "Value") +
+  theme_bw()+
+  theme(legend.position = "none")
+```
+
+One way to avoid dependence on converting variables such as size selectivity to age implemented in `SS3` is through the use of *platoons*.
+This approach allows for the exploration of survivorship that varies with size against the age.
 A value of 1 will not create additional groups.
@@
 
-In `SS3` colud be conditioned in this way;
+In `SS3`, this approach was conditioned as follows;
 
@@
 
-Regarding to observation in SAM -2023 (paragraph 4.3) that say;
+in this way, regarding to observation in SAM -2023 (paragraph 4.3);
 
@@
 
+Since we don't have age data monitoring program, we plan to explore *platoon* alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.
+
+## Observation 2
+
 *It encouraged the authors to provide standard model diagnostics, similar to those presented for toothfish assessments, to facilitate understanding of model performance (e.g., WG-SAM-2023, paragraphs 6.33 and 6.34).*
 
+Regarding @WGSAM2114, they provide diagnostic plots for the 2021 assessment model for Antarctic toothfish (*Dissostichus mawsoni*) in the Ross Sea region, following the recommendations of WG-SAM-2015 (SC-CAMLR-XXXIV 2015 Annex 5), and for which it is recommended to follow the diagnostic structure for this analysis.
+@WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Mote Carlo priors and posteriors distributions and fits to ages compositions.
+
+Some of the analyses presented in @WGSAM2114 are also implemented in @Mardones2023.
+
+- 1. Likelihood profile
+
+```{r likepro1, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
+knitr::include_graphics('Figs/Profile_s1.png')
+```
+
+- 2. Pearson residuals to fishery and surveys
+
+```{r resfleet, out.width='50%', fig.cap="Pearson residuals to fishery.
+Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected).
+in Antarctic krill"}
+knitr::include_graphics('s1/plots/comp_lenfit_residsflt6mkt0.png')
+```
+
+```{r ressurv, out.width='50%', fig.cap="Pearson residuals to survey.
+Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected).
+in Antarctic krill"}
+knitr::include_graphics('s1/plots/comp_lenfit_residsflt7mkt0.png')
+```
+- 3. Numbers by length
+
+```{r numsex, out.width='50%', fig.cap="Beginning of year expected numbers at length in (max ~ 2890.5 billion) in Antarctic krill"}
+knitr::include_graphics('s1/plots/numbers6_len_sex1.png')
+```
+- 4. Monte Carlo Prior distributions
+
+```{r marcov, out.width='90%', fig.cap="Each parameter distribution plots (plot 1 of 8).
+Deviation parameters are not included.
+Plotting range is equal to input limits on parameters in Antarctic krill"}
+knitr::include_graphics(c('s1/plots/parameter_distributions_page1.png',
+                          's1/plots/parameter_distributions_page2.png',
+                          's1/plots/parameter_distributions_page3.png',
+                          's1/plots/parameter_distributions_page4.png',
+                          's1/plots/parameter_distributions_page5.png',
+                          's1/plots/parameter_distributions_page6.png'))
+```
+
+- 5. Autocorrelation recruitment
+
+```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
+```
+
+- 6. Likelihood values
+
+```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
+knitr::include_graphics('Figs/Likelihoodtotal.png')
+```
+
+It is important to note that there is a degree of consistency in how the performance of stock assessment models for toothfish and krill is evaluated.
+For krill, the primary diagnostics conducted adhere to the "best practices" outlined in @Carvalho2021b work.
+This recipe of steps for testing model performance is illustrated in Figure \@ref(fig:recipediags).
+
+```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
+knitr::include_graphics('Figs/recipediags.png')
+```
+
+While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.
+
+\newpage
+
+## Observation 3.1
+
 *2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g.
-changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system.
+changes in mesh sizes and the presence of fine-mesh codend inserts)...*
+
+This observation raise implications for fishery selectivity.
+To develop this we configured  a selectivity block in an alternative scenario (`s9`) and compared its performance with the base model (`s1`).
+This is configured for all strata in 48.1.
+Tn this case, selectivity blocks are configured in `SS3` as show this template lines in `control` file:
+
+```{r eval=FALSE, echo=TRUE}
+1 #_Nblock_Patterns
+2 #_blocks_per_pattern
+#_begin and end years of blocks
+c(1980,2006,2006,2020)
+```
+
+The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of biomass (Figure \@ref(fig:comparebio)) and recruitment (Figure \@ref(fig:comparerec)).
+It can be observed that there are no significant differences, at least in the state variables.
+
+```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
+knitr::include_graphics('Figs/compare4_Bratio_uncertainty.png')
+```
+
+```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
+knitr::include_graphics('Figs/compare10_recruits_uncertainty.png')
+```
+## Observation 3.2
+
+*...and the likely invalid assumption of the Peninsula as being a closed system.
 It further highlighted the need to discuss the development of a standardized data collection plan to support the ongoing revision of the krill fishery management approach.*
 
-  - **Response**:
+Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an
+important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the
+northern Antarctic Peninsula, and into Subarea 48.2. The Weddell Front and Antarctic Slope
+Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG
+should be taken into account when allocation of catch limits, particularly the importance
+of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the
+directions of the important fluxes [WGEMM2439] (Figure \@ref(fig:ksh)).
 
-We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system.
-We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs.
-Moreover, we acknowledge the need to reconsider the assumption of the Peninsula being a closed system, and we will explore alternative approaches to address this.
-Moving forward, we will engage in discussions to develop a standardized data collection plan that aligns with the revision of the krill fishery management approach, ensuring that future data inputs are consistent and reliable.
+There are two important issues related to the current stock assessment approach presented in WG-SAM 24/27.
+The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult.
+However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean.
+Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the SAM.
+
+```{r ksh, out.width='80%', fig.cap="Graphic representation of Krill flux and retention in management strata proposed in the revised krill fishery management strategy within Subarea 48.1, as indicated by the SKEG WG"}
+knitr::include_graphics('Figs/KSH.png')
+```
+\newpage
+
+## Observation 4
 
@@
 
-- **Response**: We understand the concerns raised regarding the suitability of krill length and biological composition data from fishing vessels for this modeling approach.
-We agree that significant differences in length composition between commercial and scientific trawls, as well as among different vessels, pose challenges.
-To address this, we will review the selection criteria for the data used in our model and consider incorporating data exclusively from standardized scientific trawls to improve the reliability of the model outputs.
-We also acknowledge the importance of understanding the interaction between fisheries and krill-dependent predators.
-We will explore opportunities to integrate regular observations and data on the spatial overlap of fishing zones and predator foraging areas into the model, possibly leveraging existing ecosystem observations like those from RV Atlantida.
+We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system.
+We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs.
+According to the `C1` form from the database provided by the secretary through the SISO program, it is possible to identify (Figure \@ref(fig:trawl)) that there was a change in technique starting in 2006.
+
+```{r trawl, out.width='70%', fig.cap="Display of leght compositios with different trawl technique use in 48.1 subarea."}
+knitr::include_graphics('Figs/trawl.png')
+```
+
+This consideration was implemented as a new scenario (`s9`) as previously described.
+
+\newpage
+
+## Observation 5
 
@@
 
- - **Response**: We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions.
-We will outline a plan to present this progression in future work, beginning with a simpler model and gradually incorporating complexity.
-Furthermore, we acknowledge the presence of patterns in the residuals and the need to investigate these further.
-We will conduct a thorough analysis of the residuals to identify any underlying issues and reassess the realism of the parameter values used in the model.
-In addition, we will incorporate recent findings from SKEG regarding the krill stock hypothesis to ensure that our model assumptions are up-to-date and reflect the latest scientific understanding.
+We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions.
+However, and regarding this observation, a visual inspection of Pearson residuals can lead to a mistaken conclusion about the model *goodness-of-fit*.
+We have integrated a quantitative element to evaluate this type of test.
+The plotting options are kept mainly to those provided by `r4ss`.
+With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model.
+Green shading indicates
+no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively.
+The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series.
+We plot this in Figure \@ref(fig:runtest).
 
-*show  Harvest Rate*
+```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
+sspar(mfrow = c(5, 2), plot.cex = 0.8)
+knitr::include_graphics(c('Figs/run1.png',
+                          'Figs/run2.png',
+                          'Figs/run3.png',
+                          'Figs/run4.png',
+                          'Figs/run5.png',
+                          'Figs/run6.png',
+                          'Figs/run7.png',
+                          'Figs/run8.png',
+                          'Figs/run9.png'))
+```
 
-*Diagram management procedure*
+To facilitate automated processing, results from several diagnostic tests can also be called as table. Table \@ref(tab:runtable)
+
+```{r eval=TRUE, runtable, echo=FALSE}
+data <- data.frame(
+  Index = c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW", 
+            "SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "PREDATOR"),
+  runs.p = c(0.500, 0.145, 0.338, NA, 0.406, 0.189, 0.148, 0.334, 0.500, 0.599),
+  test = c("Passed", "Passed", "Passed", "Excluded", "Passed", 
+           "Passed", "Passed", "Passed", "Passed", "Passed"),
+  sigma3.lo = c(-0.1816665, -0.2319052, -0.1813395, NA, -0.1476043, 
+                -0.2452391, -0.2482065, -0.3723597, -0.5749614, -0.3154527),
+  sigma3.hi = c(0.1816665, 0.2319052, 0.1813395, NA, 0.1476043, 
+                0.2452391, 0.2482065, 0.3723597, 0.5749614, 0.3154527),
+  type = rep("len", 10)
+)
+
+kable(data, format = "markdown", 
+      col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
+      caption = "Table 1: Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
+  kable_styling(full_width = FALSE)
+```
+
+Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit).
+However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed.
+This calculation process is specifically designed to avoid dependence on the initial conditioning parameters of the model.
 
 \newpage
-
-# DISCUSSION
-
-The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
-The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
-However, integrated models are not new to krill in SO.
-For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
-In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
-
-@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
-@Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula.
-Although these models had convergence issues, they represented significant advances in this type of framework.
-Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
-Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Oscillation Southern (ENSO).
-In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.
-
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
-By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
-
-From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
-There are many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
-For this, we consider the penguin  population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill  [@Kruger2021].
-The intention to incorporate elements of predation in a stock is necessary to improve the modeling, because it did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
-
-In a scientifically rigorous stock assessment, this analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics.
-It is important to note that the models presented here can be refined in terms of initial conditioning to improve fits on length compositions of predators and indices that were initially low.
-However, a primary objective was to evaluate the impact on key population variables, considering factors such as spatial heterogeneity and life history parameters.
-This impact was verified in this stock assessment approach, identifying aspects that must be considered when making recommendations for krill management by CCAMLR using models.
-
-By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, the significance of incorporating spatial variation into the model can be assessed.
-This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
-
-Nowadays, one of the frontiers of stock assessment is the incorporation of ecosystem and spatial considerations into modeling.
-In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
-In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
-
-\newpage
-
-# SUPPLEMENTARY MATERIAL
-
-- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)
-
-- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
-
-- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
-
-\newpage
-
-# FIGURES AND TABLES
-
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
-knitr::include_graphics('Figs/481.png')
-```
-
-```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
-knitr::include_graphics('Figs/conceptual.jpeg')
-```
-
-```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
-knitr::include_graphics('Figs/data_plot_s2.png')
-```
-
-```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
+```{r eval=TRUE,, parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
 
@@
 
-\begin{landscape}
+\newpage
 
-```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
-knitr::include_graphics('Figs/pathmod.png')
-```
+## Observation 6
 
-\end{landscape}
+Graph the Harvest Rate
 
-```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
-```
+```{r}
+y <- base.model1$exploitation[,1]
+x <- base.model1$exploitation[,4]
 
-```{r res1a, fig.cap = "Pearson residuals, comparing across fleets.
-Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
-```
-
-```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
-par(mfrow=c(5,2))
-knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))
-
-```
-
-```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
-knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')
-
-```
-
-```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
-knitr::include_graphics('Figs/retros1.png')
-```
-
-```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
-par(mfrow=c(2,1))
-knitr::include_graphics(c('Figs/rmse21.png',
-                          'Figs/rmselens1.png'))
-```
-
-```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
-knitr::include_graphics('Figs/Profile_s1.png')
-```
-
-```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
-out_s2 <- read_excel("DataKrill.xlsx", 
-    sheet = "variable_s2")
-out_s2 %>%
-  kbl(booktabs = TRUE,
-      format = "latex",
-    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = TRUE,
-                font_size=6 )#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-
-\newpage
-```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
-```
-
-```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
-```
-
-```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
-knitr::include_graphics('Figs/Likelihoodtotal.png')
-```
-
-```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
-knitr::include_graphics('Figs/longtermsb.png')
+hr <- ggplot()+
+  geom_line(aes(y,x))+
+  geom_point(aes(y,x))+
+  theme_bw()+
+  labs(y="Harvest Rate",
+       x ="")+
+  ylim(0, 0.3)
+hr
 ```
@@
 
-\begin{landscape}
+# SUPPLEMENTARY MATERIAL
 
-```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
-par(mfrow=c(2,2))
-knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
-                          'Figs/compare2_spawnbio_uncertainty.png',
-                          'Figs/compare4_Bratio_uncertainty.png',
-                          'Figs/compare17_densities_SSB_Virgin.png'))
-```
+- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)
 
-\end{landscape}
+- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
+
+- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
 

### 2024-09-08 SS3_Krill_481_FSA.Rmd
wp2

--- 2024-08-26 SS3_Krill_481.Rmd
@@
 Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
+This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26), where main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.
+Each observation will be explained in theoretical terms and, where possible, analyzed in detail (Appendix).
+This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups.
+These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings.
+It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.
 
@@
 Multiple sources of data to inform the population dynamics model are included in this analysis.
+These data sources include those listed in Table \@ref(tab:datasources).
 These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
@@
 Mat.
-2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
+2; 3).
+For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
 Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
@@
 Mat 3); 4) Predator abundance data from "MAPPPD".
-Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
+Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.
+Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
+Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
 
@@
 
-Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
 Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
@@
 Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
-Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
+
+Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
+Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
+Regarding this, we consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
+Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
+In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap.
+Length offers insights into the age structure of the population, as there is a correlation between age and length.
+So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021].
+This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK).
+To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].
+ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples.
+ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
+Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias.
+Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003].
+Formulae exist to estimate the number of age determinations and length measurements is calculated as;
+
+$$
+\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
+$$
+
+The ALK for the time period is raised to the length distribution for that time period:
+
+$$
+\text{Raised numbers at age by length group} = \text{Numbers at length} \times \text{Proportion at age for that length}
+$$
+
+If the ALK lacks data for certain length groups, data from those groups may be assigned to adjacent ones in the ALK.
+This process generates an age-length distribution by gear and time period, but caution is needed to avoid significant biases, especially for larger lengths with a broad age range.
+The numbers at age are summed across length groups, combining variances from ageing and length sampling to produce the age composition for the specified period.
+Numbers at age for all gears can then be calculated.
+Numbers at age for all gears can be calculated as:
+
+$$
+\sum N_a \times \left(\frac{W_{ct}}{W_{cs}}\right)
+$$
+
+where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
+sampled commercial catch weight. Variance due to ageing of numbers at age for all gears can be calculated as:
+
+$$
+\sum \text{Var}_a
+$$
+where $\sum Var_a$  is the sum of variances due to ageing. Variance due to length sampling of numbers at age can be calculated as:
+
+$$
+\sum \text{Var}_l
+$$
+where $\sum Var_a$  is the sum of variances due to length sampling. The variances should be raised by:
+
+$$
+\frac{W_{ct}}{W_{cs}}
+$$
+where W~ct~ is the total commercial catch weight, and W~cs~ is the sampled commercial catch weight.
+
+The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
+The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
+In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
+In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
+This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass.
+Weight-at-length was assumed known and calculated using @Maschette2020 parametres.
+in krill, a AKL is exposed in Figure \@ref(fig:AKL)).
+
+```{r AKL, out.width='60%',  fig.cap="ALK Matrix Heatmap to krill in 48.1}
+alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz
+
+# Convertir la matriz en un data.frame
+alk_df <- as.data.frame(as.table(alk_matrix))
+
+# Renombrar columnas para mayor claridad
+colnames(alk_df) <- c("Length", "TrueAge", "Value")
+
+# Asegúrate de que las columnas Length y TrueAge sean numéricas
+alk_df$Length <- as.numeric(as.character(alk_df$Length))
+alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
+
+# Asegúrate de que la columna Value sea numérica
+alk_df$Value <- as.numeric(alk_df$Value)
+
+# Crear el plot usando ggplot
+ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "blue") +
+  labs(x = "True Age",
+       y = "Length",
+       fill = "Value") +
+  theme_bw()+
+  theme(legend.position = "none")
+```
+
+One way to avoid dependence on converting variables such as size selectivity to age implemented in `SS3` is through the use of *platoons*.
+This approach allows for the exploration of survivorship that varies with size against the age.
+A value of 1 will not create additional groups.
+Odd-numbered values (e.g., 3, 5) will divide the overall morph into that many groups, resulting in smaller, larger, and average growth groups.
+Increasing the number of groups slows down the model's performance, so it's recommended not to exceed a value of 5.
+The fraction of each morph assigned to each group can be either user-defined or approximated using a normal distribution.
+When multiple groups are designated, an additional input is required to specify the ratio of variability in size-at-age between groups versus within groups.
+This ratio is used to distribute the total growth variability.
+The size-at-age for each group is then calculated as a factor (based on the between-group and within-group variability) times the size-at-age of the central morph, which is determined from the growth parameters specific to that Growth Pattern x Sex combination [@methot2020stock].
+
+The age specific rates of fishing mortality can be linked to length specific selectivity.
+If growth platoons are invoked, then each platoon can have different age specific F due to interaction of length selectivity with their different size at age (Methot, *com pers*)
+
+In `SS3`, this approach was conditioned as follows;
+
+```{r eval=FALSE, echo=TRUE}
+#_N_Growth_Patterns (Growth Patterns, Morphs, Bio Patterns, GP are terms used interchangeably in SS)
+1  
+#_N_platoons_Within_GrowthPattern
+3 
+ #_Platoon_between/within_stdev_ratio (no read if N_platoons=1)
+0.4
+#vector_platoon_dist_(-1_in_first_val_gives_normal_approx)
+c(0.237, 0.464, 0.237)
+```
+
+Since we don't have age data monitoring program, we plan to explore *platoon* (Figure \@ref(fig:platoon)) alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.
 
@@
 |    s7    | "s1" w/ set of parameters estimated in @Mardones2024par   |
+|    s8    | "s1" w/ platoons conditions   |
+|    s9    | "s1" w/ Selectivity blocks in 2006   |
 
@@
 
+```{r datasources, echo=FALSE}
+data_sources <- data.frame(
+  `Data Source` = c("1. Krill Fishery Monitoring Program", 
+                    "2. AMLR Research Cruises & Monitoring Programs", 
+                    "3. Chlorophyll-a Concentration (Chl) Data", 
+                    "4. Predator Abundance Data from MAPPPD"),
+  Description = c("Data from CPUE and length compositions systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+This database includes information about vessel, nationality, and georeferenced data, among others.",
+                  "Data from scientific AMLR research cruises and monitoring programs designed to assess krill populations, including abundance by strata and biological characteristics of krill collected through net sampling.
+Length compositions from AMLR surveys (provided by Ecosystem Science Division, NOAA) were handled according to evaluation template requirements (Suppl.
+Mat.
+2).",
+                  "Chl was assessed as a primary environmental driver.
+Data was sourced from Bio-Geo-Chemical, L4 satellite observations (monthly and interpolated) obtained through the E.U.
+Copernicus Marine Service Information.
+This data helps evaluate ecosystem health and krill dynamics in the WAP.
+Fitted a GLMM to predict Length with Year and environmental factors.",
+                  "Data on penguin colonies within the study area, systematized as a vectorized indicator to consider as a fleet (M2) in an integrated stock assessment.
+Although penguin dynamics differ throughout the analyzed series, this data was included to account for predator abundance."),
+  `Temporal Scale` = c("2000 - Present", "Annual Surveys, 1990s - Present", "Monthly, 2002 - Present", "1979 - Present"),
+  `Reference/Notes` = c("Handled according to Dornam (2021).",
+                        "Processed in Dornam (2021).",
+                        "Processed with GLMM model.",
+                        "Handled with *tidyverse* library and dependencies (Wickham, 2019).")
+)
+kable(data_sources, format = "markdown", col.names = c("Data Source", "Description", "Temporal Scale", "Reference Notes"),
+      caption = "Summary of data sources used to inform the population dynamics model") %>%
+  kable_styling(full_width = FALSE) %>%
+  column_spec(1, width = "1in") %>%
+  column_spec(2, width = "2in") %>%
+  column_spec(3, width = "1in") %>%
+  column_spec(4, width = "1in")
+```
+
 ```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
@@
 
-```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
+```{r path, out.width='70%', fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
 knitr::include_graphics('Figs/pathmod.png')
@@
 \end{landscape}
+
+```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
+
+x_values <- seq(0, 7, by=0.1)
+means <- c( 2.9,  3.5, 4.1)
+std_devs <- c(0.65, 0.8, 0.65)
+data <- data.frame(
+  x = rep(x_values, each = length(means)),
+  mean = rep(means, times = length(x_values)),
+  std_dev = rep(std_devs, times = length(x_values))
+)
+
+data <- data %>%
+  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))
+
+ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
+  geom_line(size = 1) +
+  scale_linetype_manual(values = means,
+                        name ="Platoon") +
+  labs(x = "", y = "growth increment") +
+  theme_minimal()+
+  xlim(0,7)
+```
 

### 2024-09-08 index.Rmd
wp2

@@
 dir8<-here("s8") # s1 test data weigthing
+dir9<-here("s9") # s1 w/ blocks
 dirtest<-here("test")
@@
 |    s8    | "s1" test data weighting                                                             |
+|    s9    | "s1" w/ selectivity blocks                                                             |
 
@@
                  "s7",
-                 "s8")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
+                 "s8",
+                 "s9")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
 
@@
 r4ss::run(
-  dir = dir8,
+  dir = dir9,
   exe = "ss_osx",
@@
 
+#s9
+base.model9 <- SS_output(dir=dir9,
+                        covar=T,
+                        forecast=T)
+
 # html
-SS_plots(base.model8,
+SS_plots(base.model9,
          uncertainty=T,
@@
                subplots = "len",
-               add=T)
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
 ```
@@
 ```
-
+### Comparision between select models `s1` and `s9`
+
+```{r echo=FALSE}
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist1_9 <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir1,
+                                  dir9),
+                       getcovar = F)
+
+#create summary of model runs from list above
+summaryoutput1_9 <- SSsummarize(biglist1_9)
+
+comp1_9 <- SSplotComparisons(summaryoutput1_9,
+                  legendlabels = c(
+                                   "Base",
+                 "selectivity Block 2006"),
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"),
+                 png=TRUE,plotdir=Figs)
+```
 ## Comparsion in sd long term time series

### 2024-09-08 Corrections_SAM_SS3_Krill_481.Rmd
add html

@@
 title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
-subtitle: "Appendix WP- SAM 2024/26"
+subtitle: "Appendix 1 WP- SAM 2024/26"
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
@@
 
-The main recommendations were extracted from @CCAMLRWGSAM2024-
+The main recommendations were extracted from @CCAMLRWGSAM2024.
 
@@
 
+```{r include=FALSE, message=FALSE}
+base.model1 <- SS_output(dir=dir1,
+                        covar=T,
+                        forecast=T)
+```
+
 ```{r AKL}
@@
 Since we don't have age data monitoring program, we plan to explore *platoon* alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.
+
+\newpage
 
@@
 ```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
-knitr::include_graphics('Figs/compare4_Bratio_uncertainty.png')
+knitr::include_graphics('Figs/COM1_9compare4_Bratio_uncertainty.png')
 ```
@@
 ```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
-knitr::include_graphics('Figs/compare10_recruits_uncertainty.png')
+knitr::include_graphics('Figs/COM1_9compare10_recruits_uncertainty.png')
 ```
@@
 
-Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an
-important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the
-northern Antarctic Peninsula, and into Subarea 48.2. The Weddell Front and Antarctic Slope
-Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG
-should be taken into account when allocation of catch limits, particularly the importance
-of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the
-directions of the important fluxes [WGEMM2439] (Figure \@ref(fig:ksh)).
+Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the northern Antarctic Peninsula, and into Subarea 48.2.
+The Weddell Front and Antarctic Slope Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG should be taken into account when allocation of catch limits, particularly the importance of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the directions of the important fluxes [@WGEMM2439] (Figure \@ref(fig:ksh)).
 
@@
 
-This consideration was implemented as a new scenario (`s9`) as previously described.
+This consideration was implemented as a new scenario (`s9`) as previously described and implemented in stock assessment of krill.
 
@@
 ```
-
 To facilitate automated processing, results from several diagnostic tests can also be called as table. Table \@ref(tab:runtable)
 
-```{r eval=TRUE, runtable, echo=FALSE}
+```{r runtable, eval=TRUE, echo=FALSE}
 data <- data.frame(
@@
 
-Graph the Harvest Rate
+Although it was not recorded in the WGSAM-2024 report, one of the questions raised was related to the representation of population variables, such as the harvest rate.
+In this document, we present them in Figure \@ref(fig:hr).
+for the entire analyzed series and for each stratum.
 
-```{r}
-y <- base.model1$exploitation[,1]
-x <- base.model1$exploitation[,4]
+```{r hr, out.width='80%', fig.show='hold',  fig.cap="Harvest rate by strata in time series (1980-2020) in Antarctic krill fishey"}
+# Convertir la matriz en un formato largo (long format) para ggplot
+exploitation_long <- tidyr::pivot_longer(base.model1$exploitation, 
+                                         cols = starts_with("FISHERY"), 
+                                         names_to = "Fishery", 
+                                         values_to = "Exploitation")
 
-hr <- ggplot()+
-  geom_line(aes(y,x))+
-  geom_point(aes(y,x))+
+# Crear el gráfico
+ggplot(exploitation_long, 
+       aes(x = Yr, y = Exploitation, color = Fishery)) +
+  geom_point() +
+  geom_line() +
+   geom_line(aes(y = annual_F), color = "black", linetype = "dashed", size = 1) + 
+  scale_color_viridis_d()+
+  labs(title = "",
+       x = "Year",
+       y = "Harvest Rat",
+       color = "Strata") +
   theme_bw()+
-  labs(y="Harvest Rate",
-       x ="")+
-  ylim(0, 0.3)
-hr
+  xlim(1980,2020)
+
 ```
@@
 
-# SUPPLEMENTARY MATERIAL
-
-- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)
-
-- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
-
-- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
-
-\newpage
-
 # REFERENCES

### 2024-09-08 SS3_Krill_481_FSA.Rmd
add html

@@
   - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
-  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
+  - George Watters^[Ecosystem Antarctic Division, Southwest Fisheries Science Center, NOAA, USA]
 
@@
                       fig.pos = "H",
+                      fig.show='hold',
                       dev = 'jpeg',
@@
 
+```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
+knitr::include_graphics('Figs/481.png')
+```
+
 ## Data Sources
@@
 Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.
+
+```{r datasources, echo=FALSE}
+data_sources <- data.frame(
+  `Data Source` = c("1. Krill Fishery Monitoring Program", 
+                    "2. AMLR Research Cruises & Monitoring Programs", 
+                    "3. Chlorophyll-a Concentration (Chl) Data", 
+                    "4. Predator Abundance Data from MAPPPD"),
+  Description = c("Data from CPUE and length compositions systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+This database includes information about vessel, nationality, and georeferenced data, among others.",
+                  "Data from scientific AMLR research cruises and monitoring programs designed to assess krill populations, including abundance by strata and biological characteristics of krill collected through net sampling.
+Length compositions from AMLR surveys (provided by Ecosystem Science Division, NOAA) were handled according to evaluation template requirements (Suppl.
+Mat.
+2).",
+                  "Chl was assessed as a primary environmental driver.
+Data was sourced from Bio-Geo-Chemical, L4 satellite observations (monthly and interpolated) obtained through the E.U.
+Copernicus Marine Service Information.
+This data helps evaluate ecosystem health and krill dynamics in the WAP.
+Fitted a GLMM to predict Length with Year and environmental factors.",
+                  "Data on penguin colonies within the study area, systematized as a vectorized indicator to consider as a fleet (M2) in an integrated stock assessment.
+Although penguin dynamics differ throughout the analyzed series, this data was included to account for predator abundance."),
+  `Temporal Scale` = c("2000 - Present", "Annual Surveys, 1990s - Present", "Monthly, 2002 - Present", "1979 - Present"),
+  `Reference/Notes` = c("Handled according to Dornam (2021).",
+                        "Processed in Dornam (2021).",
+                        "Processed with GLMM model.",
+                        "Handled with *tidyverse* library and dependencies (Wickham, 2019).")
+)
+kable(data_sources, format = "markdown", col.names = c("Data Source", "Description", "Temporal Scale", "Reference Notes"),
+      caption = "Summary of data sources used to inform the population dynamics model") %>%
+  kable_styling(full_width = FALSE) %>%
+  column_spec(1, width = "1in") %>%
+  column_spec(2, width = "2in") %>%
+  column_spec(3, width = "1in") %>%
+  column_spec(4, width = "1in")
+```
+
 Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
 Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
+
+```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
+knitr::include_graphics('Figs/data_plot_s2.png')
+```
 
@@
 A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
-Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
+Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population.
+There are two important issues related to the current stock assessment approach.
+The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult.
+However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean.
+Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1 [@WGEMM2439], however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the @CCAMLRWGSAM2024.
+Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
 The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.
@@
 Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
+
+\newpage 
+```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
+
+# leo archivos para plotear y hacer tablas
+start1 <- SS_readstarter(file = file.path(dir1,
+                                          "starter.ss"),
+                              verbose = FALSE)
+# note the data and control file names can vary, so are determined from the 
+# starter file.
+dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
+                        verbose = FALSE)
+# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
+# assumes the correct data structure
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
+                                    start1$ctlfil),
+                        verbose = FALSE,
+                        use_datlist = TRUE, 
+                   datlist = dat1)
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
+                                                "forecast.ss"),
+                              verbose = FALSE)
+# can also read in wtatage.ss for an empirical wt at age model using
+# r4ss::SS_readwtatage()
+
+parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
+ row.names( parbio)<-c("Nat M",
+                       "Lmin", 
+                       "Lmax",
+                       "VonBert K",
+                       "CV young",
+                       "CV old", 
+                       "Wt a", 
+                       "Wt b",
+                       "L50%", 
+                       "Mat slope")
+
+ SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
+ Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
+ Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
+ parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
+
+parInit %>%
+  kbl(booktabs = TRUE,
+      format = "latex",
+      position="ht!",
+    caption = "Input parameters for the initial SS3 model of krill.
+Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
+If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
+  kable_paper("hover", 
+              full_width = F)%>%
+  kable_styling(latex_options = c("striped"),
+                full_width = FALSE,
+                font_size=9)%>% 
+  pack_rows(index = c("Natural Mortality" = 1,
+                        "Growth"= 5,
+                        "Length-Weigth Relation" = 2,
+                        "Maturity"=2,
+                        "Stock-Recruit Relation"=5,
+                        "Catchability"=2,
+                        "Selectivity"=4))
+```
+
+### Age-Length Key Matrix
 
@@
 
-```{r AKL, out.width='60%',  fig.cap="ALK Matrix Heatmap to krill in 48.1}
-alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz
-
-# Convertir la matriz en un data.frame
-alk_df <- as.data.frame(as.table(alk_matrix))
-
-# Renombrar columnas para mayor claridad
-colnames(alk_df) <- c("Length", "TrueAge", "Value")
-
-# Asegúrate de que las columnas Length y TrueAge sean numéricas
-alk_df$Length <- as.numeric(as.character(alk_df$Length))
-alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
-
-# Asegúrate de que la columna Value sea numérica
-alk_df$Value <- as.numeric(alk_df$Value)
-
-# Crear el plot usando ggplot
-ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
-  geom_tile() +
-  scale_fill_gradient(low = "white", high = "blue") +
-  labs(x = "True Age",
-       y = "Length",
-       fill = "Value") +
-  theme_bw()+
-  theme(legend.position = "none")
+```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
+knitr::include_graphics('Figs/AKL.png')
 ```
@@
 
+```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
+x_values <- seq(0, 7, by=0.1)
+means <- c( 2.9,  3.5, 4.1)
+std_devs <- c(0.65, 0.8, 0.65)
+data <- data.frame(
+  x = rep(x_values, each = length(means)),
+  mean = rep(means, times = length(x_values)),
+  std_dev = rep(std_devs, times = length(x_values))
+)
+
+data <- data %>%
+  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))
+
+ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
+  geom_line(size = 1) +
+  scale_linetype_manual(values = means,
+                        name ="Platoon") +
+  labs(x = "", y = "growth increment") +
+  theme_minimal()+
+  xlim(0,7)
+```
+
+### Selectivity Blocks
+
+According to the observations made at @CCAMLRWGSAM2024 regarding changes in net size and trawl type that could affect selectivity, we have incorporated a scenario (`s9`) with a selectivity block for krill.
+The inclusion of selectivity blocks aims to, on the one hand, reduce the effect of autocorrelation in the model fitting process, and on the other hand, explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms.
+According to the data identified, a shift from traditional to continuous trawling occurred in 2006.
+As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements.
+This is compared with the base scenario.
+
 ### Environmental variable modelling in `SS3`
@@
 
-\newpage
+\begin{landscape}
+
+```{r path, out.width='100%', fig.cap="Framework stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
+knitr::include_graphics('Figs/pathmod.png')
+```
+
+\end{landscape}
+
+It is important to note that there is a degree of consistency in how the performance of stock assessment models for toothfish and krill is evaluated.
+For krill, the primary diagnostics conducted adhere to the "best practices" outlined in @Carvalho2021b work.
+This recipe of steps for testing model performance is illustrated in Figure \@ref(fig:recipediags).
+
+```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
+knitr::include_graphics('Figs/recipediags.png')
+```
+
+While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.
 
@@
 
-For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).
+A comparative analysis of models `s01`, `s1`, and `s7` will be conducted, as they incorporate elements such as aggregate data, spatially structured data, and different sets of life history parameters according to @Mardones2024par.
+Additionally, to address the observations, scenarios `s8` and `s9` have been added and compared.
+The details of each observation made in the @CCAMLRWGSAM2024 report have been thoroughly addressed, and each response can be reviewed in **APPENDIX 1**.
 
@@
 For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
-The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
+
+```{r res1,  out.width='60%', fig.cap = "Pearson residuals, comparing across fleets.
+Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
+knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
+```
+
+```{r res1a, out.width='60%', fig.cap = "Pearson residuals, comparing across fleets.
+Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
+knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
+```
+
+Regarding this pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*.
+We have integrated a quantitative element to evaluate this type of test.
+The plotting options are kept mainly to those provided by `r4ss`.
+With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model.
+Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively.
+The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series.
+We plot this in Figure \@ref(fig:runtest) and Table \@ref(tab:runtable).
+
+```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
+sspar(mfrow = c(5, 2), plot.cex = 0.8)
+knitr::include_graphics(c('Figs/run1.png',
+                          'Figs/run2.png',
+                          'Figs/run3.png',
+                          'Figs/run4.png',
+                          'Figs/run5.png',
+                          'Figs/run6.png',
+                          'Figs/run7.png',
+                          'Figs/run8.png',
+                          'Figs/run9.png'))
+```
+
+```{r runtable, eval=TRUE,  echo=FALSE}
+data <- data.frame(
+  Index = c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW", 
+            "SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "PREDATOR"),
+  runs.p = c(0.500, 0.145, 0.338, NA, 0.406, 0.189, 0.148, 0.334, 0.500, 0.599),
+  test = c("Passed", "Passed", "Passed", "Excluded", "Passed", 
+           "Passed", "Passed", "Passed", "Passed", "Passed"),
+  sigma3.lo = c(-0.1816665, -0.2319052, -0.1813395, NA, -0.1476043, 
+                -0.2452391, -0.2482065, -0.3723597, -0.5749614, -0.3154527),
+  sigma3.hi = c(0.1816665, 0.2319052, 0.1813395, NA, 0.1476043, 
+                0.2452391, 0.2482065, 0.3723597, 0.5749614, 0.3154527),
+  type = rep("len", 10)
+)
+
+kable(data, format = "markdown", 
+      col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
+      caption = "Table 1: Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
+  kable_styling(full_width = FALSE)
+```
+
+Regarding the fitting of indices in general, the model attempts to interpret the trends but encounters some difficulties in effectively fitting all sources of information, as shown in (Figure \@ref(fig:resall)).
+
+```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
+par(mfrow=c(5,2))
+knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
+                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))
+
+```
+The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:fitcom)).
+
+```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
+knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')
+
+```
 
@@
 
+```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
+knitr::include_graphics('Figs/retros1.png')
+```
+
+```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
+par(mfrow=c(2,1))
+knitr::include_graphics(c('Figs/rmse21.png',
+                          'Figs/rmselens1.png'))
+```
+
 ## Likelihood profile in spatial Model `s1`
@@
 
+```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
+knitr::include_graphics('Figs/Profile_s1.png')
+```
+
 Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)
+
+```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
+out_s2 <- read_excel("DataKrill.xlsx", 
+    sheet = "variable_s2")
+out_s2 %>%
+  kbl(booktabs = TRUE,
+      format = "latex",
+    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = TRUE,
+                font_size=6 )#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
 
@@
 
+```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
+```
+
+```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
+knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
+```
+
+```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
+knitr::include_graphics('Figs/Likelihoodtotal.png')
+```
+
+```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
+knitr::include_graphics('Figs/longtermsb.png')
+```
+
 We carried out a specific comparison between the most divergent scenarios that we tested in this case, `s1`,  `s01` and `s7` which according to the description are the aggregated spatial scenario, disaggregate spatial scenario and scenario with new set of parameters, respectively.
 All with ecosystem components such as predators and an environmental variable.In this sense, the impacts and difference on the estimates of recruitment,  R0, virginal biomass and relative biomass are high (Figure \@ref(fig:comparision)), demonstrating the impact of initial considerations and assumptions on population variables estimation.
+
+\newpage
+
+\begin{landscape}
+
+```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
+par(mfrow=c(2,2))
+knitr::include_graphics(c('Figs/COM1compare10_recruits_uncertainty.png',
+                          'Figs/COM1compare2_spawnbio_uncertainty.png',
+                          'Figs/COM1compare4_Bratio_uncertainty.png',
+                          'Figs/COM1compare17_densities_SSB_Virgin.png'))
+```
+
+\end{landscape}
+
+\newpage
+
+A selectivity block in an alternative scenario (`s9`) was implemented and compared its performance with the base model (`s1`).
+The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)).
+It can be observed that there are no significant differences, at least in the state variables.
+
+\newpage
+
+\begin{landscape}
+
+```{r comparisionselec, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
+par(mfrow=c(2,2))
+knitr::include_graphics(c('Figs/COM1_9compare10_recruits_uncertainty.png',
+                          'Figs/COM1_9compare2_spawnbio_uncertainty.png',
+                          'Figs/COM1_9compare4_Bratio_uncertainty.png',
+                          'Figs/COM1_9compare17_densities_SSB_Virgin.png'))
+```
+
+\end{landscape}
 
@@
 In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
+Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit).
+However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed.
+This calculation process is specifically designed to avoid dependence on the initial conditioning parameters of the model.
 
@@
 
+- This work is considered a continuous inductive analytical exercise, which takes into account elements influencing krill dynamics, such as data, factors, opinions, and others, and integrates them through a mathematical process.
+This exercise is iteratively refined with observations and corrections as it is communicated to the Antarctic community.
+Therefore, we appreciate the feedback provided in the WG-SAM and WG-EMM 2024, which has contributed to its improvement.
+
 \newpage
@@
 
-# FIGURES AND TABLES
-
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
-knitr::include_graphics('Figs/481.png')
-```
-
-```{r datasources, echo=FALSE}
-data_sources <- data.frame(
-  `Data Source` = c("1. Krill Fishery Monitoring Program", 
-                    "2. AMLR Research Cruises & Monitoring Programs", 
-                    "3. Chlorophyll-a Concentration (Chl) Data", 
-                    "4. Predator Abundance Data from MAPPPD"),
-  Description = c("Data from CPUE and length compositions systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database includes information about vessel, nationality, and georeferenced data, among others.",
-                  "Data from scientific AMLR research cruises and monitoring programs designed to assess krill populations, including abundance by strata and biological characteristics of krill collected through net sampling.
-Length compositions from AMLR surveys (provided by Ecosystem Science Division, NOAA) were handled according to evaluation template requirements (Suppl.
-Mat.
-2).",
-                  "Chl was assessed as a primary environmental driver.
-Data was sourced from Bio-Geo-Chemical, L4 satellite observations (monthly and interpolated) obtained through the E.U.
-Copernicus Marine Service Information.
-This data helps evaluate ecosystem health and krill dynamics in the WAP.
-Fitted a GLMM to predict Length with Year and environmental factors.",
-                  "Data on penguin colonies within the study area, systematized as a vectorized indicator to consider as a fleet (M2) in an integrated stock assessment.
-Although penguin dynamics differ throughout the analyzed series, this data was included to account for predator abundance."),
-  `Temporal Scale` = c("2000 - Present", "Annual Surveys, 1990s - Present", "Monthly, 2002 - Present", "1979 - Present"),
-  `Reference/Notes` = c("Handled according to Dornam (2021).",
-                        "Processed in Dornam (2021).",
-                        "Processed with GLMM model.",
-                        "Handled with *tidyverse* library and dependencies (Wickham, 2019).")
-)
-kable(data_sources, format = "markdown", col.names = c("Data Source", "Description", "Temporal Scale", "Reference Notes"),
-      caption = "Summary of data sources used to inform the population dynamics model") %>%
-  kable_styling(full_width = FALSE) %>%
-  column_spec(1, width = "1in") %>%
-  column_spec(2, width = "2in") %>%
-  column_spec(3, width = "1in") %>%
-  column_spec(4, width = "1in")
-```
-
-```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
-knitr::include_graphics('Figs/conceptual.jpeg')
-```
-
-```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
-knitr::include_graphics('Figs/data_plot_s2.png')
-```
-
-```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
-
-# leo archivos para plotear y hacer tablas
-start1 <- SS_readstarter(file = file.path(dir1,
-                                          "starter.ss"),
-                              verbose = FALSE)
-# note the data and control file names can vary, so are determined from the 
-# starter file.
-dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
-                        verbose = FALSE)
-# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
-# assumes the correct data structure
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
-                                    start1$ctlfil),
-                        verbose = FALSE,
-                        use_datlist = TRUE, 
-                   datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
-                                                "forecast.ss"),
-                              verbose = FALSE)
-# can also read in wtatage.ss for an empirical wt at age model using
-# r4ss::SS_readwtatage()
-
-parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
- row.names( parbio)<-c("Nat M",
-                       "Lmin", 
-                       "Lmax",
-                       "VonBert K",
-                       "CV young",
-                       "CV old", 
-                       "Wt a", 
-                       "Wt b",
-                       "L50%", 
-                       "Mat slope")
-
- SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
- Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
- Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
- parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-
-parInit %>%
-  kbl(booktabs = TRUE,
-      format = "latex",
-      position="ht!",
-    caption = "Input parameters for the initial SS3 model of krill.
-Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
-If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-  kable_paper("hover", 
-              full_width = F)%>%
-  kable_styling(latex_options = c("striped"),
-                full_width = FALSE,
-                font_size=9)%>% 
-  pack_rows(index = c("Natural Mortality" = 1,
-                        "Growth"= 5,
-                        "Length-Weigth Relation" = 2,
-                        "Maturity"=2,
-                        "Stock-Recruit Relation"=5,
-                        "Catchability"=2,
-                        "Selectivity"=4))
-```
-
-\begin{landscape}
-
-```{r path, out.width='70%', fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
-knitr::include_graphics('Figs/pathmod.png')
-```
-
-\end{landscape}
-
-```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
-
-x_values <- seq(0, 7, by=0.1)
-means <- c( 2.9,  3.5, 4.1)
-std_devs <- c(0.65, 0.8, 0.65)
-data <- data.frame(
-  x = rep(x_values, each = length(means)),
-  mean = rep(means, times = length(x_values)),
-  std_dev = rep(std_devs, times = length(x_values))
-)
-
-data <- data %>%
-  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))
-
-ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
-  geom_line(size = 1) +
-  scale_linetype_manual(values = means,
-                        name ="Platoon") +
-  labs(x = "", y = "growth increment") +
-  theme_minimal()+
-  xlim(0,7)
-```
-
-```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
-```
-
-```{r res1a, fig.cap = "Pearson residuals, comparing across fleets.
-Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
-```
-
-```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
-par(mfrow=c(5,2))
-knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))
-
-```
-
-```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
-knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')
-
-```
-
-```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
-knitr::include_graphics('Figs/retros1.png')
-```
-
-```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
-par(mfrow=c(2,1))
-knitr::include_graphics(c('Figs/rmse21.png',
-                          'Figs/rmselens1.png'))
-```
-
-```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
-knitr::include_graphics('Figs/Profile_s1.png')
-```
-
-```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
-out_s2 <- read_excel("DataKrill.xlsx", 
-    sheet = "variable_s2")
-out_s2 %>%
-  kbl(booktabs = TRUE,
-      format = "latex",
-    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = TRUE,
-                font_size=6 )#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-
-\newpage
-```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
-```
-
-```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
-```
-
-```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
-knitr::include_graphics('Figs/Likelihoodtotal.png')
-```
-
-```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
-knitr::include_graphics('Figs/longtermsb.png')
-```
-
-\newpage
-
-\begin{landscape}
-
-```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
-par(mfrow=c(2,2))
-knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
-                          'Figs/compare2_spawnbio_uncertainty.png',
-                          'Figs/compare4_Bratio_uncertainty.png',
-                          'Figs/compare17_densities_SSB_Virgin.png'))
-```
-
-\end{landscape}
-
-\newpage
-
 # REFERENCES

### 2024-09-08 index.Rmd
add html

@@
                       warning = FALSE,
-                      fig.height = 5,
-                      fig.width = 7,
                       fig.align = 'center',
@@
                         forecast=T)
-
+```
+
+```{r eval= FALSE, message=F, include=FALSE}
 # html
@@
 
-comparision between select models `No spatial`and `Spatial implicit` and `Spatial W/ new set parametres`
+## Comparision between select models `No spatial`and `Spatial implicit` and `Spatial W/ new set parametres`
 
@@
                  "Spatial w/ new set LH paramters"),
+                 filenameprefix = "COM1",
                  labels = c("Year", 
@@
                  "selectivity Block 2006"),
+                 filenameprefix = "COM1_9",
                  labels = c("Year", 
@@
 ```
+## AKL
+
+```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
+alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz
+
+# Convertir la matriz en un data.frame
+alk_df <- as.data.frame(as.table(alk_matrix))
+
+# Renombrar columnas para mayor claridad
+colnames(alk_df) <- c("Length", "TrueAge", "Value")
+
+# Asegúrate de que las columnas Length y TrueAge sean numéricas
+alk_df$Length <- as.numeric(as.character(alk_df$Length))
+alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
+
+# Asegúrate de que la columna Value sea numérica
+alk_df$Value <- as.numeric(alk_df$Value)
+
+# Crear el plot usando ggplot
+ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "blue") +
+  labs(x = "True Age",
+       y = "Length",
+       fill = "Value") +
+  theme_bw()+
+  theme(legend.position = "none")
+```
 

### 2024-09-08 Observations_SAM_SS3_Krill_481.Rmd
commit add html 9 sept

--- 2024-09-08 Corrections_SAM_SS3_Krill_481.Rmd
@@
 ---
-title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
-subtitle: "Appendix 1 WP- SAM 2024/26"
+title: "APPENDIX 1: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+subtitle: "Observations WP- SAM 2024/26"
 date:  "`r format(Sys.time(), '%d %B, %Y')`"

### 2024-09-08 SS3_Krill_481_FSA.Rmd
commit add html 9 sept

@@
 bibliography: SA_krill.bib
+always_allow_html: true
 csl: apa.csl
@@
 Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
-@Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
+@Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@ref(fig:retros1)).
 To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual.
@@
 ```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
+knitr::include_graphics('Figs/recruitdev.png')
 ```

### 2024-09-08 index.Rmd
commit add html 9 sept

@@
 
-```{r fig.width=8, fig.height=7}
+```{r}
 dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 

### 2024-09-11 Observations_SAM_SS3_Krill_481.Rmd
test n
finish edits (i hope)

@@
 title: "APPENDIX 1: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
-subtitle: "Observations WP- SAM 2024/26"
-date:  "`r format(Sys.time(), '%d %B, %Y')`"
+subtitle: "Response to observations provided by the Working Group on WP- SAM 2024/26"
 output:
@@
 
-The main recommendations were extracted from @CCAMLRWGSAM2024.
+The main recommendations were extracted from the WG-SAM report @CCAMLRWGSAM2024.
 
@@
 
-We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
+We consider a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
 Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
@@
 (A.4)).
-Weight-at-length was assumed known and calculated using @Maschette2020 parametres.
+Weight-at-length was assumed known and calculated using @Maschette2020 parameters.
 
@@
 
-```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
+```{r platoon, out.width='70%', fig.cap="Representation of platoons in Krill length frequencies"}
 knitr::include_graphics('Figs/Plattoons.png')
@@
 
-Since we don't have age data monitoring program, we plan to explore *platoon* alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.
+Since we do not have an age data monitoring program, we explored a *platoon* alternative to compare its performance and results with the current length-based model currently in WG-SAM-2024/26 [@Mardones2023].
 
@@
 Regarding @WGSAM2114, they provide diagnostic plots for the 2021 assessment model for Antarctic toothfish (*Dissostichus mawsoni*) in the Ross Sea region, following the recommendations of WG-SAM-2015 (SC-CAMLR-XXXIV 2015 Annex 5), and for which it is recommended to follow the diagnostic structure for this analysis.
-@WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Mote Carlo priors and posteriors distributions and fits to ages compositions.
+@WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Monte Carlo priors and posteriors distributions and fits to ages compositions.
 
@@
 
-```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
+```{r acf, out.width='90%', fig.cap=" Cross correlation in autoregressive analysis for all scenarios in krill population"}
 knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
@@
 
-While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.
+While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a convergence towards these standard practices, regardless of the assessment platform and the structure of the integrated model.
 
@@
 
-This observation raise implications for fishery selectivity.
+This observation raises implications for fishery selectivity.
 To develop this we configured  a selectivity block in an alternative scenario (`s9`) and compared its performance with the base model (`s1`).
@@
 
-```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
+```{r comparebio, out.width='80%', fig.cap="Biomass estimates  in s1 and s9 scenarios."}
 knitr::include_graphics('Figs/COM1_9compare4_Bratio_uncertainty.png')
@@
 
-```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
+```{r comparerec, out.width='80%', fig.cap="Biomass estimates  in s1 and s9 scenarios."}
 knitr::include_graphics('Figs/COM1_9compare10_recruits_uncertainty.png')
@@
 There are two important issues related to the current stock assessment approach presented in WG-SAM 24/27.
-The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult.
+The first one, while we understand that there are krill flows to and from Subrea 48.1, the challenge of having time series data to implement a model that covers most of Area 48 becomes difficult.
 However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean.
-Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the SAM.
+Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Subrea 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at WG-SAM.
 
@@
 
-```{r trawl, out.width='70%', fig.cap="Display of leght compositios with different trawl technique use in 48.1 subarea."}
+```{r trawl, out.width='70%', fig.cap="Display of length compositions with different trawl technique use in 48.1 subarea."}
 knitr::include_graphics('Figs/trawl.png')
@@
 The plotting options are kept mainly to those provided by `r4ss`.
-With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model.
+With `SSplotRunstest()`, run tests results are illustrated for mean lengths of size composition data from the  `SS3`model.
 Green shading indicates
@@
 
-Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit).
+Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is shown in Table \@ref(tab:parainit).
 However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed.
@@
 Although it was not recorded in the WGSAM-2024 report, one of the questions raised was related to the representation of population variables, such as the harvest rate.
-In this document, we present them in Figure \@ref(fig:hr).
-for the entire analyzed series and for each stratum.
+The historical harvest rate of krill, as depicted in the graph, shows a significant decline from the 1970s through the 1990s, followed by a period of stabilization at much lower levels in the 2000s and 2010s.
+Initially, the exploitation rate was notably high, particularly in the Bransfield stratum, indicating intense harvesting activities during these early decades.
+However, after the early 1990s, the exploitation rate decreased sharply, suggesting the implementation of more effective management practices or a reduction in fishing pressure.
+In recent years, the exploitation rate across all strata has remained relatively low and stable, which may reflect sustainable fisheries management efforts.
+This trend highlights the importance of continued monitoring and adaptive management to ensure the long-term sustainability of krill populations in the face of environmental changes and evolving fishing practices (Figure \@ref(fig:hr)).
 

### 2024-09-11 SS3_Krill_481.Rmd
testc4
finish edits (i hope)

@@
 ---
-title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+title: "Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations"
 author:
-  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
-  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
-  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
+  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
+  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
+  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
   - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
-
-date:  "`r format(Sys.time(), '%d %B, %Y')`"
 output:

### 2024-09-11 SS3_Krill_481_FSA.Rmd
finish edits (i hope)

@@
 ---
-title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+title: "Integrated approach to modeling krill population dynamics in the Western Antarctic Peninsula: spatial and ecosystem considerations"
 author:
-  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
-  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
-  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
-  - George Watters^[Ecosystem Antarctic Division, Southwest Fisheries Science Center, NOAA, USA]
-
-date:  "`r format(Sys.time(), '%d %B, %Y')`"
+  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
+  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
+  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
 output:
@@
 
-The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in an integrated stock assessment model.Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making.
-This approach has three main objectives.
-Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance.
-On the other hand, consider the spatial heterogeneity of the krill population structure.
-Finally, to evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
-Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
-This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26), where main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.
-Each observation will be explained in theoretical terms and, where possible, analyzed in detail (Appendix).
-This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups.
-These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings.
-It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.
+The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill (*Euphausia superba*) in an integrated stock assessment model.
+Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making.
+The analysis represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
+We have three main objectives.
+i)integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance; ii) to consider the spatial heterogeneity of the krill population structure and, iii) to evaluate the impact of biological and  population structure assumptions on the performance of this type of integrated stock assessment models.
+Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management strategies in this region.
+This study addresses key recommendations provided by the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) .
+The base model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions.
+Retrospective analysis showed a slight underestimation bias in spawning biomass; however, this was within an acceptable range.
+The model suggests a higher dependence on length composition data compared to other sources for estimating recruitment.
+Spatial considerations and assumptions about stock-recruitment relationships significantly impact population variable estimates.
+The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the base model (`s1`).
+Overall, this study demonstrates the potential of using a spatially explicit integrated model for krill stock assessment.
+The model provides valuable insights into krill population dynamics and can be used to inform sustainable management practices.
 
-*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem consideration, management, CCAMLR.*
+*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem-based management, CCAMLR*
 
@@
 
-Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
-On the other hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
+Antarctic krill (*Euphausia superba*, thereafter krill) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
+WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
 Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-As part of its mandate, CCAMLR promotes sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
+As part of its mandate, CCAMLR promotes sustainable management of krill through an ecosystem-based approach  [@Watters2013].
 CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
@@
 However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-There is a growing consensus on the need for using stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
+There is a growing consensus on the need for using integrated stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
 Integrating the complexity of krill population dynamics presents several challenges.
@@
 Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
-Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
+Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich [@Kawaguchi2024], and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
 In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
+The study area for the assessment of krill populations in Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
@@
 
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
+```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Bransfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
 knitr::include_graphics('Figs/481.png')
@@
 
-Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include those listed in Table \@ref(tab:datasources).
-These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling.
+The analysis includes multiple sources of data to inform the population dynamics model listed in Table \@ref(tab:datasources).
+These data sources include: 1) Haul-by-haul data informed to CCAMLR in the C1 form, which include information about vessel, nationality, and georeferenced catch and effort, particularly useful for CPUE estimates.
+2) Data from the CCAMLR Scheme of International Scientific Observation (SISO) program, where records of krill length composition, is systematically collected on board by scientific observers.
+3) Data collected from U.S.
+Antarctic Marine Living Resources (AMLR) research cruises and monitoring programs specifically designed to assess krill abundance by strata, and biological characteristics of krill collected through net sampling.
 All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
@@
 2; 3).
+
 For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
@@
 Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl.
-Mat 3); 4) Predator abundance data from "MAPPPD".
+Mat 3; 4).
+Predator data was analyzed by using penguin abundances at different colonies along WAP available in the Mapping Application for Penguin Populations and Projected Dynamics "MAPPPD" portal.
 Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.
@@
 
-Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
-Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
+Source data and temporal scale used in this stock assessment is shown in Figure \@ref(fig:datas2).
+Data handling for modeling purposes was elaborated with *tidyverse* library and dependencies [@Wickham2019].
 
@@
 
-A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
+A conceptual model for krill in Subarea 48.1 includes  spatial strata described previously.
 Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population.
-There are two important issues related to the current stock assessment approach.
-The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult.
-However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean.
-Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1 [@WGEMM2439], however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the @CCAMLRWGSAM2024.
 Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
@@
 Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
+The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013; @methot2020stock] with the updated version 3.30.21.
 SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
 SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
+The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012] with estimation parameters.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in the WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that influence the analyzed population.
+All analyses were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
 Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
@@
 \newpage
+
 ```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
@@
                        "Mat slope")
-
  SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
@@
 Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
-Regarding this, we consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
+Regarding this, we consider a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring.
 Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
-In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap.
+In this sense, reading hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap.
 Length offers insights into the age structure of the population, as there is a correlation between age and length.
@@
 
-```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
+```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in SUbarea 48.1"}
 knitr::include_graphics('Figs/AKL.png')
@@
 
-Since we don't have age data monitoring program, we plan to explore *platoon* (Figure \@ref(fig:platoon)) alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.
+Since age data monitoring data was not available, we explored *platoon* (Figure \@ref(fig:platoon)) as an alternative to compare its performance and results with the current length-based model currently used in WG-SAM-2024/26 [@Mardones2023].
 
-```{r platoon, out.width='70%', fig.cap="Reprsentation of platoons in Krill lenght frencuecies"}
+```{r platoon, out.width='70%', fig.cap="Representation of platoons in krill length frequencies"}
 x_values <- seq(0, 7, by=0.1)
@@
 
-According to the observations made at @CCAMLRWGSAM2024 regarding changes in net size and trawl type that could affect selectivity, we have incorporated a scenario (`s9`) with a selectivity block for krill.
-The inclusion of selectivity blocks aims to, on the one hand, reduce the effect of autocorrelation in the model fitting process, and on the other hand, explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms.
+Here we have incorporated a scenario (`s9`) with a selectivity block for krill.
+The inclusion of selectivity blocks aims to reduce the effect of autocorrelation in the model fitting process, and also, to explicitly account for fleet behavior in exploring new extraction areas, whether in bathymetric or latitudinal terms.
 According to the data identified, a shift from traditional to continuous trawling occurred in 2006.
@@
 \(R\)  represents the variability of deviations, adding to the variance caused by environmental factors.
-\(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable.
+\(SpBio\) represent spawning biomass, \(envdata\) is a vector of time series of environmental variables.
 Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
@@
 
-Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
+Furthermore, this integrated model took into account environmental variables, such as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
 These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
@@
 
-This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics.
-By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
-In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously.
+This stock assessment approach considers predator-prey interactions as a key driver of krill population dynamics.
+By incorporating information on predator abundance and feeding ecology, our model provides a more comprehensive understanding of the trophic interactions shaping krill dynamics.
+In this analysis we incorporated a penguin population index as predator *"fleets"* as we described previously.
 This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.
@@
 Additionally, to address the observations, scenarios `s8` and `s9` have been added and compared.
-The details of each observation made in the @CCAMLRWGSAM2024 report have been thoroughly addressed, and each response can be reviewed in **APPENDIX 1**.
+To evaluate consistency of the source of information used in each of the tested scenarios, we estimate of R0 through likelihood components.
 
@@
 
-All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
-About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
-In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
-The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`.
-For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000.
+All models tested were the final gradient using maximum likelihood is achieved and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
+Regarding goodness-of-fit, Figure \@ref(fig:res1) display the residuals of krill length caught by fleets over the years in `s1`.
+ultiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS) are recorded Figure \@ref(fig:res1).
+SURVEYEI, SURVEYGS, SURVEYJOIN and PREDATOR have consistency after the year 2000.
 However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
-For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
+For instance, in FISHERYBS and FISHERYEI, the distribution widened and slightly shifted towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
 
-```{r res1,  out.width='60%', fig.cap = "Pearson residuals, comparing across fleets.
-Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
+```{r res1,  out.width='100%', fig.cap = "Pearson residuals, comparing across fleets. red are positive residuals (observed > expected) and blue are negative residuals (observed < expected)."}
+knitr::include_graphics('Figs/pearson_heatmap.png')
 ```
 
-```{r res1a, out.width='60%', fig.cap = "Pearson residuals, comparing across fleets.
-Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
-```
+Regarding this Pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*.
+We have integrated a quantitative element to evaluate this type of test.
+We run a statistical test to illustrate the residual of mean lengths of size composition data from the SS3 model.
+Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively.
+The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series (see Figure \@ref(fig:runtest) and Table \@ref(tab:runtable)).
 
-Regarding this pearson residual analysis, a visual inspection can lead to a mistaken conclusion about the model *goodness-of-fit*.
-We have integrated a quantitative element to evaluate this type of test.
-The plotting options are kept mainly to those provided by `r4ss`.
-With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model.
-Green shading indicates no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively.
-The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series.
-We plot this in Figure \@ref(fig:runtest) and Table \@ref(tab:runtable).
-
-```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
+```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in krill."}
 sspar(mfrow = c(5, 2), plot.cex = 0.8)
@@
 
-Regarding the fitting of indices in general, the model attempts to interpret the trends but encounters some difficulties in effectively fitting all sources of information, as shown in (Figure \@ref(fig:resall)).
+In general,  in the fitting of indices the model attempts to interpret the trends but with some difficulties in effectively fitting all sources of information, in particular, in predator fleet (Figure \@ref(fig:resall)).
 
@@
 
-Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
+This analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
 @Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@ref(fig:retros1)).
@@
 
-```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
+```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in krill"}
 knitr::include_graphics('Figs/retros1.png')
@@
 
-```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
+```{r rmses1, out.width='60%', fig.show='hold', fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in krill"}
 par(mfrow=c(2,1))
@@
 For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources.
-The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro).
+The second strongest gradient in the log-likelihood profile was observed for the length compositions (Figure \@ref(fig:likepro)).
 The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.
 
-```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
+```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in krill"}
 knitr::include_graphics('Figs/Profile_s1.png')
@@
 
-Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)
+Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms (See Table \@ref(tab:mainvar)).
 
@@
 
-## Comparision outputs between scenarios
+## Outputs comparision between scenarios
 
-In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions (Figure \@(fig:recdev)).
-We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @@McBride2021].
+We checked the consistency through recruitment deviations (Figure \@ref(fig:recdev)) by using spatial autocorrelation analysis, assessing the fluctuation of recruitment through the time series analyzed and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @McBride2021].
 
@@
 Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary, `s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness (Figure \@ref(fig:acf)).
-
-To evaluate consistency of the source of information used in each of the tested scenarios, we verify its influence on the estimate of R0 through likelihood components and, like scenario 1, we verify that all scenarios show a high dependence on the composition lengths of lengths with respect to the estimate of the variable (as show the (Figure \@ref(fig:likeall)).
+Through likelihood profile, we verify that all scenarios show a high dependence on length composition with respect to the estimate of the variable (as show in Figure \@ref(fig:likeall)).
 
@@
 
-A selectivity block in an alternative scenario (`s9`) was implemented and compared its performance with the base model (`s1`).
+A selectivity block in an alternative scenario (`s9`) related trawl thecnics and mesh changes was implemented and compared its performance with the base model (`s1`).
 The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)).
-It can be observed that there are no significant differences, at least in the state variables.
+No significant differences were detected in the state variables.
 
@@
 
-The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
-The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
-However, integrated models are not new to krill in SO.
+The development of an integrated stock assessment model can provide good support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021; @Kinzey2011; @Kinzey2015a].
+Integrated models are not new to krill in the SO.
 For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
-In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
-Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011], as is showed in Table \@ref(tab:parainit).
+In 2016, through an external review process, @Thomson2016  explicitly recommended “the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill” .
+The mentioned work highlights that an integrated modeling framework has the potential to substantially improve the scientific basis for future krill management.
+Regarding the reliability of the parameters, it is important to note that the model is conditioned with input parameters gathered from the literature [@Smith2023a; @Maschette2020; @Kinzey2011; @Thanassekos2014], as is showed in Table \@ref(tab:parainit).
 However, the model calculates the most essential parameters through an iterative process, ensuring that their correlation is carefully managed.
@@
 
-@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
+There are two important issues related to the current stock assessment approach.
+The first one, while we understand that there are krill flows to and from SubArea 48.1, the challenge of having time series data to implement a model that covers most of ASubarea 48 becomes difficult.
+However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean.
+Secondly, the current model implemented in SS3 considered the spatial heterogeneity of the krill stock structure within SubArea 48.1 (Meyer et al., 2024), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at WG-the CCAMLR Working Group SAM (2024).
+
+@Kinzey2015a and @Kinzey2019a used catch-at-length integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
 @Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula.
 Although these models had convergence issues, they represented significant advances in this type of framework.
-Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
+Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the WAP, focused on the Bransfield Strait, testing different data weightings.
 Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Oscillation Southern (ENSO).
@@
 Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
-By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
+By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aimed to capture the complex spatial patterns of krill distribution.
+This spatially implicit approach allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
+To do this, we have two options in a stock assessment modeling approach.
+First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment to have a reliable conceptual model and main estimation of population variables and transfer these considerations to the management for decision making.
+This option was our case.
+Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established into the population assessed.
+For this, it is necessary to have a spatially explicit model, which was not our case but could be considered a future approach to be analysed.
 
@@
 There are many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula  [@Silvestro2023; @Reisinger2022].
-For this, we consider the penguin  population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill  [@Kruger2021].
-The intention to incorporate elements of predation in a stock is necessary to improve the modeling, because it did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
+To incorporate predator components into the stock assessment process, we consider the penguin population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill [@Kruger2021].
+Considering predation in a stock is necessary to improve the modeling in order to  account for other sources for natural mortality and underestimated uncertainty in current stock biomass
 
-In a scientifically rigorous stock assessment, this analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics.
+Our analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics.
 It is important to note that the models presented here can be refined in terms of initial conditioning to improve fits on length compositions of predators and indices that were initially low.
@@
 Nowadays, one of the frontiers of stock assessment is the incorporation of ecosystem and spatial considerations into modeling.
-In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
-In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
+In this context, the features of the SS3 allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
+In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of krill in Subarea 48.1.
 
@@
 
-- The main challenge in this analysis has been to identify sources of information about krill population dynamic (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
+- The main challenge in this analysis has been to identify sources of information about krill population dynamics (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
 This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiple natural dimensions and evaluate the impact of this on population estimates.
 
-- Regarding to spatial heterogeneity, this can be considered in two ways in a modeling approach.
-First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer these considerations to the management for decision making.
-Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established into the population assessed.
-For this, it is necessary.
-have a spatially explicit model, which was not our case but can be considered a future approach to be analysed.
-
-- Ecosystem considerations for a model integrated with stock assessment are vital.
-Although SS3 does not model the ecosystem variables, it allows for these variables to be incorporated to model krill population dynamics.
-SS3 presents characteristics that allow the incorporation of this kind of variable.
-This ecosystem information into stock assessment serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management in the new strategy proposed by CCAMLR.
+- Ecosystem considerations for a model integrated with stock assessment are key.
+Although SS3 does not model the ecosystem variables, it allows the incorporation of this kind of information, which is generally not considered in stock assessment approaches.
+This ecosystem information into stock assessment serves as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management in the new strategy proposed by CCAMLR..
 
 - We consider that an integrated modeling of the krill adding the available information sources (fishery and surveys), as well as, the spatial heterogeneity and ecosystem variables is on the right path of the recommendation that has emerged in the last 10 years with respect to population dynamics and its link with the management procedure carried out by CCAMLR.
-
-- This work is considered a continuous inductive analytical exercise, which takes into account elements influencing krill dynamics, such as data, factors, opinions, and others, and integrates them through a mathematical process.
-This exercise is iteratively refined with observations and corrections as it is communicated to the Antarctic community.
-Therefore, we appreciate the feedback provided in the WG-SAM and WG-EMM 2024, which has contributed to its improvement.
 

### 2024-09-11 index.Rmd
finish edits (i hope)

@@
 ```
+Heatmap
+
+```{r}
+df <- base.model1$lendbase
+
+# Filtra el dataframe
+df1 <- df %>% 
+  dplyr::filter(Pearson<5) %>% 
+  dplyr::select(c(1, 6, 19, 16)) %>% 
+  dplyr::mutate(Fleet = dplyr::case_when(
+    Fleet == 1 ~ "FISHERYBS",
+    Fleet == 2 ~ "FISHERYEI",
+    Fleet == 3 ~ "FISHERYGS",
+    Fleet == 4 ~ "FISHERYJOIN",
+    Fleet == 5 ~ "FISHERYSSIW",
+    Fleet == 6 ~ "SURVEYBS",
+    Fleet == 7 ~ "SURVEYEI",
+    Fleet == 8 ~ "SURVEYGS",
+    Fleet == 9 ~ "SURVEYJOIN",
+    Fleet == 10 ~ "SURVEYSSIW",
+    Fleet == 11 ~ "PREDATOR",
+    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
+  ))
+
+# Crea el heatmap
+ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
+  geom_tile() +
+  scale_fill_gradient2(low = "blue", 
+                       mid = "white", 
+                       high = "red",
+                       midpoint = 0) +
+  facet_wrap(Fleet~., ncol=5) + 
+  labs(title = "",
+       x = "",
+       y = "Length",
+       fill = "Pearson") +
+  theme_bw()+
+  theme(axis.text.x = element_text(angle = 90,
+                                   hjust = 1))
+
+```
+hexagon
+
+```{r eval=FALSE}
+# Assuming your data is in a data frame called 'lendbase'
+# Filter and select relevant columns
+data <- base.model1$lendbase %>%
+  dplyr::filter(Pearson<5) %>% 
+  dplyr::select(c(1, 6, 19, 16)) %>% 
+  dplyr::mutate(Fleet = dplyr::case_when(
+    Fleet == 1 ~ "FISHERYBS",
+    Fleet == 2 ~ "FISHERYEI",
+    Fleet == 3 ~ "FISHERYGS",
+    Fleet == 4 ~ "FISHERYJOIN",
+    Fleet == 5 ~ "FISHERYSSIW",
+    Fleet == 6 ~ "SURVEYBS",
+    Fleet == 7 ~ "SURVEYEI",
+    Fleet == 8 ~ "SURVEYGS",
+    Fleet == 9 ~ "SURVEYJOIN",
+    Fleet == 10 ~ "SURVEYSSIW",
+    Fleet == 11 ~ "PREDATOR",
+    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
+  )) %>% 
+  dplyr::select(Yr, Bin, Pearson)
+
+# Create the hexagonal heatmap
+ggplot(data, aes(x = Yr, y = Bin)) +
+  geom_hex(bins=50) +  # Hexagonal bins
+  scale_fill_gradient2(low = "blue", 
+                       mid = "white", 
+                       high = "red", 
+                       midpoint = 0) +  # Color scale
+  theme_minimal() +  # Minimal theme for cleaner look
+  labs(title = "Hexagonal Heatmap of Pearson Residuals",
+       x = "Year",
+       y = "Bin",
+       fill = "Pearson Residual")  # Labels for plot and axes
+
+ggplot(data, aes(x = Yr, y = Bin)) +
+  geom_hex(bins = 50) +
+  scale_fill_gradient2(low = "blue", 
+                       mid = "white", 
+                       high = "red", midpoint = median(data$Bin)) +
+  theme_minimal() +
+  labs(title = "Hexagonal Heatmap of Pearson Residuals",
+       x = "Year", y = "Bin", fill = "Pearson Residual")
+```
 
@@
 ```
+
+## Likelihood Profile
 
@@
                                  'Recruitment likelihood')
-# 19. Funciones para generar plots de perfil de verosimilitud ----
+
+```
+
+Funciones para generar plots de perfil de verosimilitud ----
+
 ### SSplotProfile() ----
+
+```{r eval=FALSE, echo = FALSE}
 png(file.path(plotdir,"R0_profile_plot.png"),
@@
 ```{r eval=FALSE, echo = FALSE}
-
 # Comparación de series de tiempo 
@@
 
-### PinerPlot() ----
+```
+
+piner Plot
+
+```{r echo = FALSE}
 #### R0_profile_plot_Length_like ----
@@
 dev.off()
+```
+
+```{r echo = FALSE}
 #### R0_profile_plot_Survey_like ----

### 2024-10-02 Observations_SAM_SS3_Krill_481.Rmd
changed lot

@@
   scale_fill_gradient(low = "white", high = "blue") +
-  labs(x = "True Age",
+  labs(x = "Age",
        y = "Length",
@@
       col.names = c("Index", "runs.p", "test", "sigma3.lo", "sigma3.hi", "type"),
-      caption = "Table 1: Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
+      caption = "Runs tests results for fits differente fleet used in SS3 krill assessment") %>%
   kable_styling(full_width = FALSE)
@@
 \newpage
-```{r eval=TRUE,, parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
+```{r parainit, eval=TRUE}
 
@@
 
+And total Harvest rate could be noted in Table \@ref(tab:hrt) and Harvest Rate by Strata regarding outputs of `SS3` in Table \@ref(tab:hrt2)
+
+\newpage
+
+```{r hrt}
+# Calcular el acumulado anual de F_std por año
+data <- exploitation_long %>%
+  dplyr::filter(Yr<2021) %>% 
+  group_by(Yr) %>%
+  dplyr::summarize(F_std_accumulated = mean(F_std)) %>% 
+  dplyr::select(Yr,  F_std_accumulated) %>% 
+  dplyr::rename("Harvest Rate" ="F_std_accumulated" ,
+                "Year" = "Yr")
+
+data %>%
+  kbl(booktabs = TRUE,
+      format = "latex",
+      position="ht!",
+  caption = "Accumulate Harvest Rate by year in whole 48.1 Subarea") %>%
+  kable_paper("hover", 
+              full_width = F)%>%
+  kable_styling(latex_options = c("striped"),
+                full_width = FALSE,
+                font_size=9)
+
+```
+
+\newpage
+
+```{r hrt2}
+data2 <- base.model1$exploitation %>% 
+  dplyr::select(1, 7:11) %>% 
+  dplyr::rename("Year" = "Yr")
+data2 %>% 
+  kbl(booktabs = TRUE,
+      format = "latex",
+      position="ht!",
+      digits = 3,
+  caption = "Harvest Rate by strata and year") %>%
+  kable_paper("hover", 
+              full_width = F)%>%
+  kable_styling(latex_options = c("striped"),
+                full_width = FALSE,
+                font_size=9) %>% 
+  pack_rows("Forecast", 44, 44)
+```
+
 \newpage

### 2024-10-02 SS3_Krill_481_FSA.Rmd
changed lot

@@
 This study addresses key recommendations provided by the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) .
-The base model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions.
+The reference model (s1) performed well in capturing the inter-annual and inter-fleet variability in krill length compositions.
 Retrospective analysis showed a slight underestimation bias in spawning biomass; however, this was within an acceptable range.
@@
 Spatial considerations and assumptions about stock-recruitment relationships significantly impact population variable estimates.
-The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the base model (`s1`).
+The inclusion of a selectivity block into the model (`s9`) improved the fit to some length composition data compared to the reference model (`s1`).
 Overall, this study demonstrates the potential of using a spatially explicit integrated model for krill stock assessment.
@@
 As a result, two blocks have been established (1980-2006 and 2007-2020) as improvements.
-This is compared with the base scenario.
+This is compared with the reference scenario.
 
@@
 
-```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
+```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a reference model."}
 knitr::include_graphics('Figs/recipediags.png')
@@
 The methodology involves comparing the statistical performance of the model regarding these configurations.
-We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).
+We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the reference model (Table 1).
 
@@
 
-## Diagnosis Base Model
+## Diagnosis Model
 
@@
 
-A selectivity block in an alternative scenario (`s9`) related trawl thecnics and mesh changes was implemented and compared its performance with the base model (`s1`).
+A selectivity block in an alternative scenario (`s9`) related trawl thecnics and mesh changes was implemented and compared its performance with the reference model (`s1`).
 The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of recruit, spawning biomass, relative biomass and virgin biomass  (Figure \@ref(fig:comparisionselec)).

### 2024-10-02 index.Rmd
changed lot

@@
                 full_width = FALSE,
-                font_size=9)%>% 
+                font_size=10,
+                html_font = "arial")%>% 
   pack_rows(index = c("Mortalidad natural" = 1,
@@
 r4ss::run(
-  dir = dir9,
+  dir = dir1,
   exe = "ss_osx",
@@
 # html
-SS_plots(base.model9,
+SS_plots(base.model1,
          uncertainty=T,
@@
 ```{r}
-SSplotComps(base.model8,
+SSplotComps(base.model1,
             subplots = 1,
@@
                 full_width = FALSE,
-                font_size=9)#%>% 
+                font_size=9,
+                html_font = "arial")#%>% 
   #pack_rows(index = c("Estimation" = 1,
@@
 
+In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process
+
 ```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}

### 2024-10-04 index.Rmd
adding testint phases

@@
 r4ss::run(
-  dir = dir1,
+  dir = dirtest,
   exe = "ss_osx",
@@
                         forecast=T)
+
+#stest
+base.modeltest <- SS_output(dir=dirtest,
+                        covar=T,
+                        forecast=T)
 ```
@@
 # html
-SS_plots(base.model1,
+SS_plots(base.modeltest,
          uncertainty=T,
@@
 ```{r}
-SSplotComps(base.model1,
+SSplotComps(base.modeltest,
             subplots = 1,
@@
 ```{r}
-SSplotRunstest(base.model1,
+SSplotRunstest(base.modeltest,
                subplots = "len",
@@
                add=T)
-SSplotRunstest(base.model7,
+SSplotRunstest(base.modeltest,
                subplots = "len",
@@
 ```{r}
-SSplotRunstest(base.model1,
+SSplotRunstest(base.modeltest,
                subplots = "cpue",
@@
                add=T)
-SSplotRunstest(base.model7,
+SSplotRunstest(base.modeltest,
                subplots = "cpue",
@@
                add=T)
-SSplotJABBAres(base.model7,
+SSplotJABBAres(base.modeltest,
                subplots = "len",
@@
                add=T)
-SSplotJABBAres(base.model7,
+SSplotJABBAres(base.modeltest,
                subplots = "cpue",
@@
 retro(
-    dir = dir1,
+    dir = dirtest,
     oldsubdir = "",
     newsubdir = "Retrospective",
-    years = 0:-5,
+    years = 0:-4,
     exe = "ss_osx",
@@
 
-retroSummary <- SSsummarize(retroModelstest)
+retroSummarytest <- SSsummarize(retroModelstest)
 
@@
 retro7 <- SSplotRetro(retroSummary7,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F)
+
+retrotest <- SSplotRetro(retroSummarytest,
             add=T,

### 2024-10-08 index.Rmd
adding test new test files

@@
 
+Just some time series plots
+
+```{r}
+SSplotTimeseries(base.modeltest,
+                 subplot = 1)
+SSplotTimeseries(base.modeltest,
+                 subplot = 9)
+```
+and fits
+
+```{r}
+SSplotComps(base.modeltest,
+            subplots = 1,
+            maxrows = 5,
+            maxcols = 5)
+```
+
 ```{r eval= FALSE, message=F, include=FALSE}
@@
             subplots = 1)
-SSplotSelex(base.model7,
+SSplotSelex(base.modeltest,
             subplots = 1)
-```
-```{r}
-SSplotComps(base.modeltest,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
 ```
@@
 dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
-dir.create(path=dirname.R0.profile, showWarnings = TRUE, recursive = TRUE)
+dir.create(path=dirname.R0.profile, 
+           showWarnings = TRUE, 
+           recursive = TRUE)
 # 3. Crear un subdirectorio llamado "plots_Verosimilitud" ----
 plotdir=paste0(dirname.R0.profile, "/plots_Verosimilitud")
-dir.create(path=plotdir, showWarnings = TRUE, recursive = TRUE)
+dir.create(path=plotdir,
+           showWarnings = TRUE, 
+           recursive = TRUE)
 # 4. Crear un subdirectorio llamado "simple" ----

### 2024-10-09 index.Rmd
update index file

@@
 dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
-dir8<-here("s8") # s1 test data weigthing
+dir8<-here("s8") # s1 platoons
 dir9<-here("s9") # s1 w/ blocks
@@
 r4ss::run(
-  dir = dirtest,
+  dir = dir8,
   exe = "ss_osx",
@@
 ```{r}
-df <- base.model1$lendbase
+df <- base.modeltest$lendbase
 

### 2025-03-07 index.Rmd
test run

--- 2024-10-08 index.Rmd
@@
 dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
-dir3<-here("s3") # whitout S-R
+dir3<-here("s3") # without S-R
 dir4<-here("s4") # 
@@
                         forecast=T)
-
+```
+
+```{r message=F, include=FALSE}
 #stest
@@
 SSplotTimeseries(base.modeltest,
-                 subplot = 1)
+                 subplot = 7)
 SSplotTimeseries(base.modeltest,
@@
 ```{r }
-SSplotData(base.model01, subplot = 1, 
-           fleetnames = c("Fishery", "Survey", "Predator"),
-           fleetcol = c("blue","green", "red"))
+SSplotData(base.modeltest, subplot = 1)
 
@@
        x = "",
-       y = "Length",
+       y = "Length (cm)",
        fill = "Pearson") +
-  theme_bw()+
+  theme_few()+
   theme(axis.text.x = element_text(angle = 90,
-                                   hjust = 1))
+                                   hjust = 1, 
+                                   size = 7))
 
@@
                                             "Retrospective",
-                                            paste("retro",0:-3,
+                                            paste("retro",0:-4,
                                                   sep="")))
@@
 ```{r eval=FALSE}
-retrotest <- SSplotRetro(retroSummarytest,
+SSplotRetro(retroSummarytest,
             add=T,
@@
 ```{r}
-hci = SSplotHCxval(retroSummary1, 
+hci = SSplotHCxval(retroSummarytest, 
                    add = T, 
@@
 ```{r}
-mvln = SSdeltaMVLN(base.model1,
+mvln = SSdeltaMVLN(base.modeltest,
                    Fref = "Btgt", 
@@
 ```{r}
-outps1 <- base.model1$timeseries[1:53, 2:8]
+outps1 <- base.modeltest$timeseries[1:53, 2:8]
 
@@
 
-# Crear los gráficos en un ciclo for
 for (i in 1:length(data_list)) {

### 2025-03-19 index.Rmd
adding test 2

@@
 dir9<-here("s9") # s1 w/ blocks
-dirtest<-here("test")
+dirtest<-here("test") # con todo
+dirtest2 <- here("test2") # sin predador ni ambiental
 Figs <- here("Figs")# S
@@
 #stest
+r4ss::run(
+  dir = dirtest,
+  exe = "ss_osx",
+  skipfinished = FALSE, # TRUE will skip running if Report.sso present
+  show_in_console = TRUE # change to true to watch the output go past
+)
 base.modeltest <- SS_output(dir=dirtest,
@@
 
-Just some time series plots
-
 ```{r}
 SSplotTimeseries(base.modeltest,
-                 subplot = 7)
+                 subplot = 1)
 SSplotTimeseries(base.modeltest,
                  subplot = 9)
+SSplotTimeseries(base.modeltest,
+                 subplot = 7)
 ```

### 2025-03-20 index.Rmd
adding some files test2

@@
 r4ss::run(
-  dir = dirtest,
+  dir = dirtest2,
   exe = "ss_osx",

### 2025-03-21 index.Rmd
adding improves

--- 2025-03-07 index.Rmd
@@
 dir9<-here("s9") # s1 w/ blocks
-dirtest<-here("test")
+dir1.1<-here("s1.1") # con todo
+dir1.2 <- here("s1.2") # sin predador ni ambiental
+dir1.3 <- here("s1.3") # borrado los registros de prreddador )igual q el S2
 Figs <- here("Figs")# S
@@
 # leo archivos para plotear y hacer tablas
-start1 <- SS_readstarter(file = file.path(dir1,
+start1 <- SS_readstarter(file = file.path(dir1.1,
                                           "starter.ss"),
@@
 # starter file.
-dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
+dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
                         verbose = FALSE)
@@
 # assumes the correct data structure
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                     start1$ctlfil),
@@
                    datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                 "forecast.ss"),
@@
 parInit %>%
-  kbl(booktabs = T,
+  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
+  kbl(booktabs = TRUE,
       format = "html",
-      position="ht!",
+      position = "ht!",
       caption = "Input parameters for the initial SS3 model of krill.
@@
 If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-  kable_paper("hover", 
-              full_width = F)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = FALSE,
-                font_size=10,
-                html_font = "arial")%>% 
+  kable_paper("hover", full_width = TRUE) %>%  # Expande el ancho
+  kable_styling(latex_options = c("striped", "condensed"),
+                full_width = TRUE,  # Expande la tabla al ancho completo
+                font_size = 12,
+                html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
   pack_rows(index = c("Mortalidad natural" = 1,
-                        "Crecimiento"= 5,
+                      "Crecimiento" = 5,
                       "Relación longitud-peso" = 2,
-                        "Ojiva de madurez"=2,
-                        "Relación stock-recluta"=5,
-                        "Capturabilidad"=2,
-                        "Selectividad"=4))
+                      "Ojiva de madurez" = 2,
+                      "Relación stock-recluta" = 5,
+                      "Capturabilidad" = 2,
+                      "Selectividad" = 4))
 
@@
 |    s9    | "s1" w/ selectivity blocks                                                           |
+|    s1.1  | Spatial data without environmental and predator components                          |
+|    s1.2  | "s1.1" with environmental and predator components                                    |
+|    s1.3  | "s1.2" w/ selectivity blocks                                                         |
 
@@
 ```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
-# Lista de directorios para correro tordos juntos
-directorios <- c("s01", 
-                 "s1", 
-                 "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7",
-                 "s8",
-                 "s9")  # Agrega aquí todos los nombres de las carpetas que deseas procesar
-
-# Bucle para ejecutar el código en cada directorio
+directorios <- c("s1.1",
+                 "s1.2",
+                 "s1.3")
+
 for (dir in directorios) {
@@
 r4ss::run(
-  dir = dirtest,
+  dir = dir1.1,
   exe = "ss_osx",
@@
 )
-```
-
-Read outputs
-
-```{r message=F, include=FALSE}
-#s01
-base.model01 <- SS_output(dir=dir01,
+
+base.model1.1 <- SS_output(dir=dir1.1,
                         covar=T,
                         forecast=T)
-#s1
-base.model1 <- SS_output(dir=dir1,
+```
+
+Read outputs
+
+```{r message=F, include=FALSE}
+
+#s1.1
+base.model1.1 <- SS_output(dir=dir1.1,
                         covar=T,
                         forecast=T)
-#s2
-base.model2 <- SS_output(dir=dir2,
+#s1.2
+base.model1.2 <- SS_output(dir=dir1.2,
                         covar=T,
                         forecast=T)
-#s3
-base.model3 <- SS_output(dir=dir3,
+#s1.3
+base.model1.3 <- SS_output(dir=dir1.3,
                         covar=T,
                         forecast=T)
-#s4
-base.model4 <- SS_output(dir=dir4,
-                        covar=T,
-                        forecast=T)
-#s5
-base.model5 <- SS_output(dir=dir5,
-                        covar=T,
-                        forecast=T)
-#s6
-base.model6 <- SS_output(dir=dir6,
-                        covar=T,
-                        forecast=T)
-
-#s7
-base.model7 <- SS_output(dir=dir7,
-                        covar=T,
-                        forecast=T)
-
-#s8
-base.model8 <- SS_output(dir=dir8,
-                        covar=T,
-                        forecast=T)
-
-#s9
-base.model9 <- SS_output(dir=dir9,
-                        covar=T,
-                        forecast=T)
-```
-
-```{r message=F, include=FALSE}
-#stest
-base.modeltest <- SS_output(dir=dirtest,
-                        covar=T,
-                        forecast=T)
-```
-
-Just some time series plots
+```
 
 ```{r}
-SSplotTimeseries(base.modeltest,
+SSplotTimeseries(base.model1.1,
+                 subplot = 1)
+SSplotTimeseries(base.model1.1,
+                 subplot = 9)
+SSplotTimeseries(base.model1.1,
                  subplot = 7)
-SSplotTimeseries(base.modeltest,
-                 subplot = 9)
 ```
@@
 ```{r}
-SSplotComps(base.modeltest,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-```
-
-```{r eval= FALSE, message=F, include=FALSE}
-# html
-SS_plots(base.modeltest,
-         uncertainty=T,
+SSplotBiology(base.model1.1,
+            subplots = 1:32)
+```
+
+```{r}
+SSplotComps(base.model1.1,
+            subplots = 1)
+
+```
+
+# RESULTS
+```{r eval=FALSE, message=F, include=FALSE}
+# Definir la ruta de la carpeta
+folder_path <- "s1.1/plots"
+# Verificar si la carpeta existe y eliminarla
+if (dir.exists(folder_path)) {
+  unlink(folder_path, recursive = TRUE)
+  message("La carpeta 'plots' ha sido eliminada.")
+} else {
+  message("La carpeta 'plots' no existe.")
+}
+
+SS_plots(base.model1.1, 
+         uncertainty = TRUE,
          datplot = T, 
@@
          btarg=0.75, 
-         minbthresh=0.20, 
+         minbthresh=0.2, 
          forecast=T)
-
-```
-# RESULTS
+```
 
@@
 
-```{r eval=FALSE, message=F, include=FALSE}
-SS_plots(base.model7, 
-         uncertainty = TRUE,
-         datplot = T, 
-         png=T, 
-         aalresids = F,
-         btarg=0.75, 
-         minbthresh=0.2, 
-         forecast=T)
-```
-
 Data used en both (spatial and No spatial models)
@@
 ```{r }
-SSplotData(base.modeltest, subplot = 1)
-
-SSplotData(base.model1, 
+SSplotData(base.model1.1, 
            subplot = 1,
@@
             subplots = 1)
-SSplotSelex(base.modeltest,
+SSplotSelex(base.model1.1,
             subplots = 1)
@@
 ```{r}
-SSplotRunstest(base.modeltest,
+SSplotRunstest(base.model1.1,
                subplots = "len",
@@
 
-```{r eval=FALSE, fig.width=4, fig.height=5}
+```{r eval=FALSE}
 SSplotRunstest(base.model01,
@@
 ```{r}
-SSplotRunstest(base.modeltest,
+SSplotRunstest(base.model1.1,
                subplots = "cpue",
@@
                add=T)
-SSplotJABBAres(base.modeltest,
+SSplotJABBAres(base.model1.1,
                subplots = "len",
@@
                add=T)
-SSplotJABBAres(base.modeltest,
+SSplotJABBAres(base.model1.1,
                subplots = "cpue",
@@
 ```
-
 ## Relationship Stock-Recruit
@@
 ```{r eval=FALSE}
-l <- base.model1$recruit$SpawnBio
-m <- base.model1$recruit$pred_recr
-ggplot()+
-  aes(l,m)+
-  geom_point()+
+l <- base.modeltest$recruit$SpawnBio
+m <- base.modeltest$recruit$pred_recr
+ggplot() +
+  aes(l, m) +
+  geom_point() +
   geom_smooth(method = "loess",
-              se =F)+
-  theme_few()+
-  ylim(0, max(base.model1$recruit$pred_recr))
+              se = TRUE, 
+              color = "blue") +
+  theme_few() +
+  ylim(0, max(base.modeltest$recruit$pred_recr))
+
 ```
@@
 retro(
-    dir = dirtest,
+    dir = dir1.1,
     oldsubdir = "",
@@
 #stest
-retroModelstest <- SSgetoutput(dirvec=file.path(dirtest,
+retroModels1.1 <- SSgetoutput(dirvec=file.path(dir1.1,
                                             "Retrospective",
@@
 
-retroSummarytest <- SSsummarize(retroModelstest)
+retroSummary1.1 <- SSsummarize(retroModels1.1)
 
@@
 
-retrotest <- SSplotRetro(retroSummarytest,
+retro1.1 <- SSplotRetro(retroSummary1.1,
             add=T,
@@
             legend = T,
-            verbose=F)
-
-```
-
-```{r eval=FALSE}
-SSplotRetro(retroSummarytest,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
+            verbose=F,
+            subplots = c("SSB"))
 ```
@@
 
-mvln = SSdeltaMVLN(base.model7,
+mvln = SSdeltaMVLN(base.model1.1,
                    Fref = "Btgt", 

### 2025-03-27 MS2_Krill.Rmd
adding stest to MSE

--- 2024-08-26 SS3_Krill_481.Rmd
@@
 ---
-title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
+title: "Assessing environmental and predation effects on Antarctic Krill (*Euphausia superba*) stock productivity: An Integrated Catch-at-Length Model Perspective"
 author:
-  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
-  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
-  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
-  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
-
-date:  "`r format(Sys.time(), '%d %B, %Y')`"
+  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
+  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
+  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
+  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
 output:
-  bookdown::pdf_document2:
+  bookdown::word_document2:
+    reference_docx: MS2_Krill_template.docx.docx
     number_sections: false
@@
 bibliography: SA_krill.bib
-csl: apa.csl
+#csl: apa.csl
+csl: icesjournal.csl
 link-citations: yes
@@
 rm(list = ls())
-knitr::opts_chunk$set(eval=TRUE,
+knitr::opts_chunk$set(eval=FALSE,
                       echo = FALSE,
@@
 library(r4ss)
-dir1<-here("s1")
 ```
+
+# ABSTRACT
+
+Understanding the drivers of krill (*Euphausia superba*) population dynamics is crucial for sustainable management.
+This study evaluates the impact of environmental variables and predation components on krill productivity, specifically stock-recruitment relationships and other population parameters, in the Western Antarctic Peninsula.
+Using an integrated Catch-at-Length model, we assess how ecosystem factors influence recruitment, biomass trends, and stock status.
+Our results indicate that krill population variables, as well as productivity estimates, tend to be more optimistic in terms of stock status when environmental and predators variables are not considered.
+This highlights a critical challenge in developing management recommendations and harvest control rules, as ignoring ecosystem influences may lead to overly optimistic assessments of krill stock health.
+Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics.
+We consider that the use of integrated models, which leverage the extensive population, fisheries, and ecosystem data available through various scientific programs, can provide valuable insights for management aligned with CCAMLR conservation and harvesting objectives.
+
+**Keywords:** Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation, fisheries management.
+
+# INTRODUCTION
+
+Antarctic krill (*Euphausia superba*) is a keystone species in the Antarctic ecosystem and represents one of the largest living populations on the planet in terms of biomass (REF).
+This species is widely distributed across the Antarctic continental shelf; however, its populations are currently concentrated mainly in the Antarctic Peninsula region (REF).
+This spatial concentration has heightened scientific interest in understanding the population dynamics of krill, both from an ecological perspective—due to its crucial role as a prey species for whales, penguins, and other marine organisms (REF)—and from an economic standpoint, given the commercial fishing pressure exerted on this resource (REF).
+
+Recognizing the ecological and economic importance of krill, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has prioritized conservation and management strategies aimed at ensuring the sustainability of this resource.
+The primary objective has been to reconcile commercial exploitation with a precautionary approach that safeguards the ecological community dependent on krill (REF).
+CCAMLR’s management framework relies on sustainable exploitation recommendations based largely on survey data obtained from acoustic biomass estimation campaigns and fisheries monitoring programs (REF).
+Over time, these efforts have advanced toward integrating multiple sources of information into assessment models that guide decision-making.
+One such approach is the use of annual yield models, which recommend catch limits based on predefined exploitation thresholds (REF).
+
+In this study, we integrate multiple sources of biological and fisheries-related data, including estimates of biomass from scientific surveys, fisheries catch data, and biological parameters, to construct a modern stock assessment model that evaluates key population variables and krill productivity.
+However, contemporary stock assessment models require the incorporation of ecosystem variables—such as environmental conditions and predator-prey interactions—which are essential components of the Antarctic ecosystem and must be considered when modeling krill population dynamics (REF).
+
+To address these challenges, we integrate environmental information that influences krill population dynamics, as well as data on predator species that are affected by krill population variability.
+We employ an integrated catch-at-length stock synthesis model to compare different scenarios and assess the impact of ecosystem variability on key productivity indicators, such as recruitment and other critical population variables (REF).
+By incorporating these ecosystem drivers into our stock assessment approach, we aim to enhance our understanding of krill population fluctuations and improve the scientific basis for sustainable management recommendations.
+
+Antarctic krill (*Euphausia superba*) plays a fundamental role in the Southern Ocean ecosystem, linking primary production to higher trophic levels, including marine mammals, seabirds, and fish.
+Krill populations are subject to complex ecological pressures, including environmental variability (e.g., sea ice extent, temperature fluctuations, and primary productivity) and predation from key species such as penguins.
+In addition, commercial fisheries target krill, further influencing population dynamics.
+
+Traditional stock assessments primarily rely on fishery-dependent and survey-based data but often overlook ecosystem variables that may significantly impact population structure and productivity.
+Integrating environmental and biological drivers into stock assessments is essential to improving our understanding of krill dynamics and ensuring robust management strategies.
+
+This study aims to assess the impact of ecosystem variables, particularly environmental conditions and predation pressure from penguins, on krill productivity and population structure.
+Using an integrated Catch-to-Length model, we evaluate stock-recruitment relationships and other key population parameters to determine how ecosystem drivers shape krill dynamics.
+
+# M&M
+
+#### **2.1. Study Area and Data Sources**  
+This study focuses on the Antarctic Peninsula, a region where krill populations exhibit strong variability due to environmental fluctuations and intense predation pressure.
+Data were obtained from the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR), including fishery-dependent catch data, scientific surveys, and predator monitoring datasets.
+Environmental data, such as sea surface temperature (SST), chlorophyll-a concentration, and sea ice extent, were sourced from satellite observations and oceanographic monitoring programs.
+
+#### **2.2. Catch-to-Length Model Implementation**  
+We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends.
+The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables.
+Key model parameters include:
+
+- Growth rates based on length-frequency distributions  
+- Natural mortality estimates derived from predator consumption data  
+- Stock-recruitment functions incorporating environmental predictors  
+- Fishery removals as a function of effort and environmental conditions
+
+#### **2.3. Statistical Analysis and Model Validation**  
+Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation against independent survey datasets.
+We tested multiple scenarios to assess the relative influence of environmental and predation variables on recruitment and biomass dynamics.
+
+---
+
+### **3. Results**  
+Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
+
+1.
+**Stock-Recruitment Relationship**: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.
+2.
+**Predation Impact**: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.
+3. **Biomass Trends**: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.  
+4. **Fisheries Interaction**: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
+
+---
+
+### **4. Discussion**  
+Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
+The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
+
+- **Ecosystem-Based Management Implications**: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
+By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.
+- **Predator-Krill Interactions**: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
+Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.
+- **Climate Change Considerations**: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
+Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
+
+---
+
+### **5. Conclusion**  
+This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
+The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
+Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
+
+Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
+
+---
+
+### **Acknowledgments**  
+We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
+
+---
+
+This draft presents a formal scientific structure suitable for submission to a peer-reviewed journal. Let me know if you need modifications or additional details! 🚀
 
@@
 
-```{r mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
+```{r eval=F, mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
 out_s2 <- read_excel("DataKrill.xlsx", 

### 2025-03-27 index.Rmd
adding stest to MSE

@@
 ---
-title: "Integrated approach to modeling to krill population dynamics in Wester Antarctic Peninsula: spatial and ecosystem considerations"
-subtitle: "Working Paper to be submitted in a CCAMLR EMM-WG 2024"
-author: "Mardones, M; Cárdenas, C., Krüger, L., Santa Cruz, F."
+title: "Supplementary Information 1"
+subtitle: "Assessing ecosystem effects on the stock productivity of Antarctic Krill (*Euphausia superba*): An Integrated modeling perspective"
+author: "Mardones, M; Jarvis Mason, E.T.; Pinones, A.;  Santa Cruz, F.; Cárdenas, C.A"
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
 bibliography: SA_krill.bib
-csl: apa.csl
+#csl: apa.csl
+csl: icesjournal.csl
 link-citations: yes
@@
                       dev = 'jpeg',
+                      fig.path = "Figs/",
                       dpi = 300,
@@
 library(readxl)
+library(tidyverse)
+library(ggridges)
 ```
@@
 ```{r message=FALSE, warning=FALSE, echo=TRUE}
-dir01 <- here("s01") # agreggate data (no spatial diferences)
-dir1<-here("s1") # Data strata flishery
-dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
-dir3<-here("s3") # without S-R
-dir4<-here("s4") # 
-dir5<-here("s5") # 
-dir6<-here("s6") # 
-dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
-dir8<-here("s8") # s1 platoons
-dir9<-here("s9") # s1 w/ blocks
+# dir01 <- here("s01") # agreggate data (no spatial diferences)
+# dir1<-here("s1") # Data strata flishery
+# dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
+# dir3<-here("s3") # without S-R
+# dir4<-here("s4") # 
+# dir5<-here("s5") # 
+# dir6<-here("s6") # 
+# dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
+# dir8<-here("s8") # s1 platoons
+# dir9<-here("s9") # s1 w/ blocks
 dir1.1<-here("s1.1") # con todo
@@
 
-# OVERVIEW
-
-In a simple way, the core of Stock Synthesis is its population dynamics
-model, which represents the dynamics of krill populations over time.
-This model incorporates key biological, environmental and predator data
-sources. The model is typically formulated using mathematical equations
-that describe how these parameters interact to determine the abundance
-and distribution of krill in the study area.
-
-## Statistical model (SS3)
-
-Stock Synthesis v.3.30.21 is a widely used software tool for assessing
-fish and invertebrate populations, including krill (Euphausia superba)
-in the Antarctic Peninsula region. The methodology employed by Stock
-Synthesis involves a comprehensive and integrated approach, utilizing
-various data sources and modeling techniques to estimate the main
-population variables of krill in WAP.
-
-The stock assesment model was configured using Stock Synthesis (SS3
-hereafter)[SS3](https://vlab.noaa.gov/web/stock-synthesis) [@Methot2013;
-@methot2020stock] with the most updated version (V.3.30.21). SS3 is a
-structured age and size stock evaluation model, in the class of models
-called *"Integrated stock evaluation analysis model"*. SS3 has a stock
-population sub-model that simulates growth, maturity, fecundity,
-recruitment, movement, and mortality processes, and observation
-sub-models and expected values for different types of data. The model is
-coded in `C++` with estimation parameters enabled by automatic
-differentiation (ADMB) [@Fournier2012; @Methot2013]. The analysis of
-results and outputs uses R tools and the graphical interface of the
-*r4ss* and *ss3diags* library [@Taylor2019; @Winker2023].
-
-By integrating data from multiple sources and considering spatial
-heterogeneity, the assessment methodology using Stock Synthesis
-v.3.30.21 provides a robust framework for evaluating the status of krill
-populations in the Antarctic Peninsula region. This information is
-essential for supporting management decisions aimed at ensuring the
-sustainable use of krill resources in this ecologically sensitive area.
-
-## Parametres
-
-read files
-
-```{r}
-# leo archivos para plotear y hacer tablas
+## Overview
+
+This study aims to evaluate the impact of ecosystem components—such as environmental variables and predator-prey interactions—on the productivity and key population dynamics of *Euphausia superba* in Subarea 48.1.
+By incorporating these factors into the assessment, we analyze how krill population variables respond to ecological variability, providing insights into their resilience and potential management implications.
+
+Here, the **reference model** represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables.
+This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure.
+By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.
+
+## Statistical Model (SS3)
+
+Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill.
+SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
+In this exercise, SS3 is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers.
+The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends.
+The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023].
+By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions.
+These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.
+
+## Parameters
+
+The following table summarizes the key parameters to conditioning the reference model, including biological, growth, and population dynamics factors.
+
+```{r}
 start1 <- SS_readstarter(file = file.path(dir1.1,
@@
                               verbose = FALSE)
-# note the data and control file names can vary, so are determined from the 
-# starter file.
-dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
+dat1 <- SS_readdat(file = file.path(dir1.2, start1$datfile),
                         verbose = FALSE)
-# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
-# assumes the correct data structure
 ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
@@
  parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-```
-
-```{r eval=FALSE}
+
 wb <- createWorkbook()
@@
 # Guardar el workbook
-saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+#saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
 
@@
                 html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
-  pack_rows(index = c("Mortalidad natural" = 1,
-                      "Crecimiento" = 5,
-                      "Relación longitud-peso" = 2,
-                      "Ojiva de madurez" = 2,
-                      "Relación stock-recluta" = 5,
-                      "Capturabilidad" = 2,
-                      "Selectividad" = 4))
-
+  pack_rows(index = c("Natural Mortality" = 1,
+                      "Growth" = 5,
+                      "relationship Length-Weigth" = 2,
+                      "Maturity" = 2,
+                      "S-R relation" = 5,
+                      "Catchability" = 2,
+                      "Selectivity" = 4))
+
+```
+Source of data inpit
+
+```{r fig.width=9, fig.height=4}
+dat1_std <- dat1$CPUE %>%
+  group_by(index) %>%
+  mutate(index_std = obs / mean(obs, na.rm = TRUE)) %>%
+  ungroup() %>% 
+  mutate(index = recode(index,
+                        `1` = "FISHERYBS",
+                        `2` = "FISHERYEI",
+                        `3` = "FISHERYGS",
+                        `4` = "FISHERYJOIN",
+                        `5` = "FISHERYSSIW",
+                        `6` = "SURVEYBS",
+                        `7` = "SURVEYEI",
+                        `8` = "SURVEYGS",
+                        `9` = "SURVEYJOIN",
+                        `10` = "SURVEYSSIW",
+                        `11` = "PREDATOR"))
+ggplot(dat1_std, aes(x = year, y = index_std, group=index))+
+  geom_point() +
+  geom_smooth(method="lm",
+              se=F)+
+  facet_wrap(.~index,
+             scales = "free_y",
+             ncol=6)+
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, 
+                                   hjust = 1))+
+  labs(title = "",
+       x = "",
+       y = "Standar index",
+       color = "index")
+```
+
+```{r fig.width=10, fig.height=4}
+dat_long <- dat1$lencomp %>% 
+  pivot_longer(cols = starts_with("l"),  
+               names_to = "Talla",  
+               values_to = "Frecuencia") %>% 
+  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
+  mutate(FltSvy = recode(FltSvy,
+                         `1` = "FISHERYBS",
+                         `2` = "FISHERYEI",
+                         `3` = "FISHERYGS",
+                         `4` = "FISHERYJOIN",
+                         `5` = "FISHERYSSIW",
+                         `6` = "SURVEYBS",
+                         `7` = "SURVEYEI",
+                         `8` = "SURVEYGS",
+                         `9` = "SURVEYJOIN",
+                         `10` = "SURVEYSSIW",
+                         `11` = "PREDATOR")) %>%
+  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
+                                            "FISHERYEI", 
+                                            "FISHERYGS",
+                                            "FISHERYJOIN", 
+                                            "FISHERYSSIW",
+                                            "SURVEYBS",
+                                            "SURVEYEI", 
+                                            "SURVEYGS", 
+                                            "SURVEYJOIN",
+                                            "SURVEYSSIW",
+                                            "PREDATOR")))
+
+mean_talla <- dat_long %>%
+  group_by(FltSvy) %>%
+  summarise(mean_Talla = mean(Talla, na.rm = TRUE))
+
+# Expandir datos según la frecuencia
+dat_expanded <- dat_long %>%
+  uncount(weights = Frecuencia)
+
+# Crear el gráfico con líneas verticales en cada facet
+ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), 
+                         fill = stat(x))) +
+  geom_density_ridges_gradient(scale = 1.8,
+                               rel_min_height = 0.001,
+                               col = "black",
+                               fill = "grey") +
+  facet_wrap(~FltSvy, ncol = 10) +
+  geom_vline(data = mean_talla, aes(xintercept = mean_Talla), 
+             color = "red", 
+             size = 0.7,
+             alpha=0.5) +
+  labs(title = "",
+       x = "cm",
+       y = "",
+       fill = "Densidad") +
+  theme_few()
 ```
@@
 |:------------:|:---------------------------------------------------------|
-|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental aggregate data in 48.1       |
-|    s1    | Fishery and Survey (AMLR) data Length, Index, Catch by strata. Predator and Env data |
-|    s2    | "s1" without S-R relation                                                            |
-|    s3    | "s1" BH S-R relation weak (0.9 steepness)                                            |
-|    s4    | "s1" BH S-R relation strong (0.6 steepness)                                          |
-|    s5    | "s1" BH S-R relation mid-strong estimated                                            |
-|    s6    | "s1" Ricker S-R relation estimated                                                   |
-|    s7    | "s1" w/ set of parameters estimated in @EMM-204/32                                   |
-|    s8    | "s1" test data weighting                                                             |
-|    s9    | "s1" w/ selectivity blocks                                                           |
 |    s1.1  | Spatial data without environmental and predator components                           |
@@
 
-base.model1.1 <- SS_output(dir=dir1.1,
-                        covar=T,
-                        forecast=T)
-```
-
-Read outputs
+# s1.2
+r4ss::run(
+  dir = dir1.2,
+  exe = "ss_osx",
+  skipfinished = FALSE, # TRUE will skip running if Report.sso present
+  show_in_console = TRUE # change to true to watch the output go past
+)
+
+# s1.3
+r4ss::run(
+  dir = dir1.3,
+  exe = "ss_osx",
+  skipfinished = FALSE, # TRUE will skip running if Report.sso present
+  show_in_console = TRUE # change to true to watch the output go past
+)
+
+```
 
 ```{r message=F, include=FALSE}
-
 #s1.1
@@
 
-```{r}
-SSplotTimeseries(base.model1.1,
-                 subplot = 1)
-SSplotTimeseries(base.model1.1,
-                 subplot = 9)
-SSplotTimeseries(base.model1.1,
-                 subplot = 7)
-```
-and fits
-
-```{r}
-SSplotBiology(base.model1.1,
-            subplots = 1:32)
-```
-
-```{r}
-SSplotComps(base.model1.1,
-            subplots = 1)
-
-```
-
+Data used en both (spatial and No spatial models)
+
+```{r }
+SSplotData(base.model1.1, 
+           subplot = 1,
+           pheight = 15)
+
+SSplotData(base.model1.2, 
+           subplot = 1,
+           pheight = 15)
+```
 # RESULTS
+
 ```{r eval=FALSE, message=F, include=FALSE}
@@
          forecast=T)
-```
-
-### Main Variables poulation
+
+# S 1.2
+
+# Definir la ruta de la carpeta
+folder_path2 <- "s1.2/plots"
+# Verificar si la carpeta existe y eliminarla
+if (dir.exists(folder_path2)) {
+  unlink(folder_path2, 
+         recursive = TRUE)
+  message("La carpeta 'plots' ha sido eliminada.")
+} else {
+  message("La carpeta 'plots' no existe.")
+}
+
+SS_plots(base.model1.2, 
+         uncertainty = TRUE,
+         datplot = T, 
+         png=T, 
+         aalresids = F,
+         btarg=0.75, 
+         minbthresh=0.2, 
+         forecast=T)
+
+# S 1.3
+
+# Definir la ruta de la carpeta
+folder_path3 <- "s1.3/plots"
+# Verificar si la carpeta existe y eliminarla
+if (dir.exists(folder_path3)) {
+  unlink(folder_path3, 
+         recursive = TRUE)
+  message("La carpeta 'plots' ha sido eliminada.")
+} else {
+  message("La carpeta 'plots' no existe.")
+}
+
+SS_plots(base.model1.3, 
+         uncertainty = TRUE,
+         datplot = T, 
+         png=T, 
+         aalresids = F,
+         btarg=0.75, 
+         minbthresh=0.2, 
+         forecast=T)
+```
+
+### Main Variables poulation in `s1.1` scenario
+
+```{r}
+SSplotTimeseries(base.model1.1,
+                 subplot = 1)
+SSplotTimeseries(base.model1.1,
+                 subplot = 9)
+SSplotTimeseries(base.model1.1,
+                 subplot = 7)
+```
+
+### Main Variables poulation in `s1.2` scenario
+
+```{r}
+SSplotTimeseries(base.model1.2,
+                 subplot = 1)
+SSplotTimeseries(base.model1.2,
+                 subplot = 9)
+SSplotTimeseries(base.model1.2,
+                 subplot = 7)
+```
+
+### Main Variables poulation in `s1.3` scenario
+
+```{r}
+SSplotTimeseries(base.model1.3,
+                 subplot = 1)
+SSplotTimeseries(base.model1.3,
+                 subplot = 9)
+SSplotTimeseries(base.model1.3,
+                 subplot = 7)
+```
+
+Selectivity
+
+```{r fig.height=6}
+SSplotSelex(base.model1.1,
+            subplots = 1)
+
+SSplotSelex(base.model1.2,
+            subplots = 1)
+```
+
+```{r}
+SSplotBiology(base.model1.1,
+            subplots = 1:32)
+```
+
+```{r}
+SSplotComps(base.model1.2,
+            subplots = 1)
+
+```
 
@@
 ```{r eval=FALSE}
-total01 <- as.data.frame(base.model01$timeseries[1:50,c(2,5)]) %>%
+total01 <- as.data.frame(base.model1.1$timeseries[1:31,c(2,5)]) %>%
   mutate(Serie = "No Spatial")
-total1 <- as.data.frame(base.model1$timeseries[,c(2,5)]) %>% 
+total1 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
   mutate(Serie = "Spatial implicit")
@@
                                    hjust = 1))
-```
-
-Data used en both (spatial and No spatial models)
-
-```{r }
-SSplotData(base.model1.1, 
-           subplot = 1,
-           pheight = 15)
-```
-
-Respecto a los valores y parametros biologicos modelados, los siguientes
-graficos identifican los estimadores puntuales del recurso
-
-```{r fig.height=6}
-SSplotSelex(base.model01,
-            subplots = 1)
-SSplotSelex(base.model1,
-            subplots = 1)
-SSplotSelex(base.model2,
-            subplots = 1)
-SSplotSelex(base.model3,
-            subplots = 1)
-SSplotSelex(base.model4,
-            subplots = 1)
-SSplotSelex(base.model5,
-            subplots = 1)
-SSplotSelex(base.model6,
-            subplots = 1)
-SSplotSelex(base.model1.1,
-            subplots = 1)
 ```
@@
                plotdir = Figs)
+SSplotRunstest(base.model1.2,
+               subplots = "len",
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
 ```
@@
                add=T)
+SSplotRunstest(base.model1.2,
+               subplots = "cpue",
+               add=T)
 ```
@@
                add=T)
-SSplotJABBAres(base.model1.1,
+SSplotJABBAres(base.model1.2,
                subplots = "len",
@@
                add=T)
-SSplotJABBAres(base.model1.1,
+SSplotJABBAres(base.model1.2,
                subplots = "cpue",
@@
 
-## Likelihood
+## Likelihood tables
 
@@
 # 1. Identificar el directorio donde se encuentra el modelo base ----
-dirname.model.run <- here("s1")
+dirname.model.run <- here("s1.2")
 # 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
-dirname.R0.profile <- here("Ejercicios","Perfil_Verosimilitud")
+dirname.R0.profile <- here("s1.2",
+                           "Perfil_Verosimilitud")
 dir.create(path=dirname.R0.profile, 
@@
 # 11. Vector de valores para el perfil ----
-R0.vec <- seq(15,30,1)  
+R0.vec <- seq(18,30,1)  
 Nprof.R0 <- length(R0.vec)
@@
 #note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist01_1_7 <- SSgetoutput(keyvec = NULL,
+biglist1_2_3 <- SSgetoutput(keyvec = NULL,
                        dirvec = c(
-                                  dir01,
-                                  dir1,
-                                  dir7),
+                                  dir1.1,
+                                  dir1.2,
+                                  dir1.3),
                        getcovar = F)
@@
 #create summary of model runs from list above
-summaryoutput01_1_7 <- SSsummarize(biglist01_1_7)
-
-SSplotComparisons(summaryoutput01_1_7,
-                  legendlabels = c(
-                                   "No Spatial",
-
-                 "Spatial",
-                 "Spatial w/ new set LH paramters"),
+summaryoutput1_2_3 <- SSsummarize(biglist1_2_3)
+
+SSplotComparisons(summaryoutput1_2_3,
+                  legendlabels = c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Predator data",
+                 "S1.1 w/ Env and Predator data"),
                  filenameprefix = "COM1",
@@
                             "Harvest rate"),
-                 png=TRUE,plotdir=Figs)
+                 png=TRUE,
+                 plotdir=Figs)
 ```

### 2025-03-28 MS2_Krill.Rmd
save last chgs

@@
 linkcolor: blue
-linestretch: 1.3
+linestretch: 1.5
 header-includes:
@@
 
+\newpage
+
 # ABSTRACT
@@
 Understanding the drivers of krill (*Euphausia superba*) population dynamics is crucial for sustainable management.
-This study evaluates the impact of environmental variables and predation components on krill productivity, specifically stock-recruitment relationships and other population parameters, in the Western Antarctic Peninsula.
+This study evaluates the impact of environmental variables and predation components on krill productivity and other population parameters, in the Western Antarctic Peninsula (WAP).
 Using an integrated Catch-at-Length model, we assess how ecosystem factors influence recruitment, biomass trends, and stock status.
@@
 
-**Keywords:** Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation, fisheries management.
+*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation components, fisheries management.*
+
+\newpage
 
@@
 
-Antarctic krill (*Euphausia superba*) is a keystone species in the Antarctic ecosystem and represents one of the largest living populations on the planet in terms of biomass (REF).
-This species is widely distributed across the Antarctic continental shelf; however, its populations are currently concentrated mainly in the Antarctic Peninsula region (REF).
-This spatial concentration has heightened scientific interest in understanding the population dynamics of krill, both from an ecological perspective—due to its crucial role as a prey species for whales, penguins, and other marine organisms (REF)—and from an economic standpoint, given the commercial fishing pressure exerted on this resource (REF).
+Antarctic krill (*Euphausia superba*) is a keystone species in the Antarctic ecosystem and one of the largest populations on the planet in terms of biomass (REF).
+Currently, its population is predominantly concentrated in the Western Antarctic Peninsula (WAP), a critical feeding ground for predators and the primary area for krill harvesting over the past four decades (REF).
+At the same time, the WAP is one of the most vulnerable regions to climate change (CC) and has already undergone rapid environmental shifts [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
+These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective.
+Given its ecological significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
 
-Recognizing the ecological and economic importance of krill, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has prioritized conservation and management strategies aimed at ensuring the sustainability of this resource.
-The primary objective has been to reconcile commercial exploitation with a precautionary approach that safeguards the ecological community dependent on krill (REF).
-CCAMLR’s management framework relies on sustainable exploitation recommendations based largely on survey data obtained from acoustic biomass estimation campaigns and fisheries monitoring programs (REF).
-Over time, these efforts have advanced toward integrating multiple sources of information into assessment models that guide decision-making.
-One such approach is the use of annual yield models, which recommend catch limits based on predefined exploitation thresholds (REF).
+Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has adopted a new management strategy (CCAMLR 2019) that follows an ecosystem-based approach.
+This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species.
+It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks.
 
-In this study, we integrate multiple sources of biological and fisheries-related data, including estimates of biomass from scientific surveys, fisheries catch data, and biological parameters, to construct a modern stock assessment model that evaluates key population variables and krill productivity.
-However, contemporary stock assessment models require the incorporation of ecosystem variables—such as environmental conditions and predator-prey interactions—which are essential components of the Antarctic ecosystem and must be considered when modeling krill population dynamics (REF).
+As part of this strategy, CCAMLR has tested various modeling approaches to assess krill stock status and support fisheries management decisions [@Hill2016].
+While model-based stock assessments have been widely used (REF), many struggle to align with management objectives or accurately represent krill population dynamics due to their spatial variability and ecosystem complexity.
+One widely applied approach is the General Yield Model (GRYM) (REF), which provides global catch limits based on predefined harvest thresholds.
+Traditionally, stock assessments have relied on single-source data, either from fisheries catches [@Kindzey; @Wang] or population monitoring surveys [@Green2023; @Watters2013].
+However, these approaches often overlook critical factors such as spatial heterogeneity in krill abundance, environmental drivers, and predator-prey interactions.
+Given the influence of environmental variables on krill recruitment and the role of predators in adding an additional component of natural mortality, any stock assessment framework must integrate these ecosystem factors to better represent population dynamics.
 
-To address these challenges, we integrate environmental information that influences krill population dynamics, as well as data on predator species that are affected by krill population variability.
-We employ an integrated catch-at-length stock synthesis model to compare different scenarios and assess the impact of ecosystem variability on key productivity indicators, such as recruitment and other critical population variables (REF).
-By incorporating these ecosystem drivers into our stock assessment approach, we aim to enhance our understanding of krill population fluctuations and improve the scientific basis for sustainable management recommendations.
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture spatial variability, and account for ecosystem interactions.
+However, integrating these complexities presents several challenges.
+Long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies.
+Additionally, some areas have sparse or incomplete datasets on krill dynamics.
+Furthermore, indirect drivers, such as environmental variability and predator abundance, introduce another layer of complexity.
+Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations.
+This requires not only refining existing datasets but also adopting a fully ecosystem-based management approach.
 
-Antarctic krill (*Euphausia superba*) plays a fundamental role in the Southern Ocean ecosystem, linking primary production to higher trophic levels, including marine mammals, seabirds, and fish.
-Krill populations are subject to complex ecological pressures, including environmental variability (e.g., sea ice extent, temperature fluctuations, and primary productivity) and predation from key species such as penguins.
-In addition, commercial fisheries target krill, further influencing population dynamics.
+To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past four decades.
+This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations.
+Specifically, we employ an integrated catch-at-length stock synthesis model that allows us to evaluate the effects of environmental drivers on krill recruitment and productivity, as well as the impact of predator-prey interactions on overall population dynamics.
+Our approach involves comparing different scenarios, one incorporating ecosystem-related variables and another based solely on traditional stock assessment data, to determine how environmental fluctuations and predator abundance influence key population variables and parameters.
+By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends.
+This comparative approach enables us to assess the relative importance of environmental conditions and predator dynamics in shaping krill stocks, providing critical insights for both fisheries management and conservation efforts.
 
-Traditional stock assessments primarily rely on fishery-dependent and survey-based data but often overlook ecosystem variables that may significantly impact population structure and productivity.
-Integrating environmental and biological drivers into stock assessments is essential to improving our understanding of krill dynamics and ensuring robust management strategies.
-
-This study aims to assess the impact of ecosystem variables, particularly environmental conditions and predation pressure from penguins, on krill productivity and population structure.
-Using an integrated Catch-to-Length model, we evaluate stock-recruitment relationships and other key population parameters to determine how ecosystem drivers shape krill dynamics.
+By integrating ecosystem drivers into the stock assessment process, we enhance the accuracy of population estimates and improve our ability to predict future stock trends under changing environmental conditions.
+This broader ecological perspective strengthens the scientific foundation for sustainable management recommendations, ensuring that harvesting strategies align with an ecosystem-based management approach.
+Ultimately, the insights gained from this integrated modeling effort contribute to the ongoing efforts of CCAMLR's new strategy to balance krill fishery sustainability with the conservation of the broader Antarctic marine ecosystem.
 
@@
 
+Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
+Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
+Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
+
+### **Tablas a incluir**
+1. **Comparación de parámetros clave entre escenarios:**
+   - Biomasa desovante (SSB, Spawning Stock Biomass) en años clave.
+   - Tasa de explotación (F).
+   - Reclutamiento medio.
+   - Relación B/B0 (biomasa actual respecto a la virgen).
+   - Relación F/FMSY.
+
+   | Parámetro       | Base | Con variables ecosistémicas |
+   |----------------|------|----------------------------|
+   | SSB (2020)     | X    | Y                          |
+   | SSB (2025)     | X    | Y                          |
+   | F (2020)       | X    | Y                          |
+   | Reclutamiento medio | X | Y                        |
+   | B/B0 (2020)    | X    | Y                          |
+   | F/FMSY (2020)  | X    | Y                          |
+
+2. **Resumen del ajuste del modelo:**
+   - Log-verosimilitud total.
+   - Verosimilitud de los datos de captura, índices de abundancia y composición de tallas/edades.
+   - Número de parámetros estimados.
+   - Error estándar de las estimaciones clave.
+
+3. **Comparación de reclutamiento por año:**
+   - Media y desviación estándar del reclutamiento en periodos clave (por ejemplo, 1990-2000, 2000-2010, 2010-2020).
+   - Probabilidad de años de reclutamiento bajo.
+
+### **Gráficos recomendados**
+1. **Evolución de la biomasa desovante (SSB) en el tiempo**  
+   - Líneas para cada escenario mostrando tendencias divergentes.
+   - Inclusión de intervalos de confianza.
+
+2. **Reclutamiento anual estimado por escenario**  
+   - Barras para cada año con comparación entre escenarios.
+   - Promedio móvil para suavizar tendencias.
+
+3. **Relación stock-reclutamiento con diferentes ajustes**  
+   - Ajustes de la curva con y sin variables ecosistémicas.
+   - Comparación de la fuerza del reclutamiento en diferentes condiciones ambientales.
+
+4. **Proyecciones de biomasa y F bajo diferentes escenarios de explotación**  
+   - Predicciones de la biomasa y la explotación para los próximos 5-10 años.
+   - Comparación entre políticas de manejo con y sin ecosistema.
+
+5. **Distribución de longitudes o edades en el tiempo**  
+   - Para ver si el escenario con variables ecosistémicas predice cambios en la estructura de la población.
+
+6. **Curvas de Kobe plot (B/BMSY vs. F/FMSY)**  
+   - Comparar el estado del stock en cada escenario respecto a puntos de referencia de manejo.
+
+(Imagen Verosimilitud)
+
+The comparison of likelihood components across the three scenarios revealed notable differences in their fit to the data.
+Scenario S1.1 had the lowest total likelihood, suggesting a better overall fit compared to S1.2 and S1.3.
+The catch and equilibrium catch components showed minor variations among scenarios, indicating that these data had a limited influence on the overall likelihood differences.
+In terms of survey likelihood, S1.1 exhibited a slightly better fit to survey abundance data, while S1.2 and S1.3 showed higher likelihood values, suggesting a reduced agreement with observed biomass trends.
+The length composition component differed significantly, with S1.2 and S1.3 having higher likelihood values than S1.1, implying a poorer fit to observed length distributions.
+Recruitment likelihood was highest in S1.1, indicating stronger consistency with recruitment estimates compared to S1.2 and S1.3.
+Additionally, parameter deviations were more constrained in S1.1, whereas S1.2 and S1.3 exhibited greater variability.
+These results suggest that S1.1 provides a better overall fit, particularly in survey abundance and length composition components, while S1.2 and S1.3 show greater discrepancies in these aspects.
+
 ---
@@
 We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
-
----
-
-This draft presents a formal scientific structure suitable for submission to a peer-reviewed journal. Let me know if you need modifications or additional details! 🚀
-
-# ABSTRACT
-
-The integration of available sources of information is a challenge for any attempt to know dynamics populations for Antarctic Krill *Euphausia superba* in an integrated stock assessment model.Being able to identify sources of information, systematize and integrate it in a model and estimate krill population variables in a context of change and fisheries management is vital for management and decision making.
-This approach has three main objectives.
-Firstly, integrate fishing, environmental and ecological variables in an integrated stock assessment model and test his performance.
-On the other hand, consider the spatial heterogeneity of the krill population structure.
-Finally, to evaluate the impact of biological and structure population assumptions on the performance of this type of integrated stock assessment models.
-Consequently, by acknowledging and integrating different data sources, the stock assessment model can provide insights into the ecological dynamics of krill populations and improve CCAMLR fishery management and conservation strategies in this region.
-
-*Keywords: Krill populations, dynamic population, integrated model, Stock Synthesis (SS3), spatial implicit, ecosystem consideration, management, CCAMLR.*
-
-\newpage
-
-# INTRODUCTION
-
-Antarctic krill (*Euphausia superba* krill thereafter) population is predominantly concentrated in the Western Antarctic Peninsula (WAP), serving as a vital food source for predators and being targeted by increasing fishing activities.
-On the other hand, WAP is considered one of the most sensitive areas to Climate Change (CC) and has already experienced fast changes in various dimensions [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
-Given its ecological significance and economic value, effective management of krill stocks is paramount to ensure the sustainability of Antarctic ecosystems and support fisheries-dependent communities.
-As part of its mandate, CCAMLR promotes sustainable management of Antarctic krill through an ecosystem-based approach [@Watters2013].
-CCAMLR has tried to conduct different modeling approaches to assess the status of important species, and to provide information for decision-making regarding fisheries management and conservation measures [@Hill2016].
-Various stock assessment approaches on krill have been conducted to provide management advice to CCAMLR using model-based methodologies.
-Many of these models face challenges in aligning with management schemes or accurately representing the population dynamics of krill, due to its variability and heterogeneity in spatial distribution and ecosystem complexity.
-Traditionally, stock assessments have relied on single-source data, primarily from fisheries catch data or population monitoring surveys [@Green2023; @Watters2013].
-However, these approaches may overlook critical aspects of krill dynamics, such as integrated different data components, spatial heterogeneity in abundance and distribution, as well as the influence of environmental factors and predator-prey interactions.
-There is a growing consensus on the need for using stock assessment models that incorporate multiple sources of information, capture spatial heterogeneity and consider the complex interplay of ecosystem variables in krill.
-Integrating the complexity of krill population dynamics presents several challenges.
-Firstly, data consistency over the years has been problematic, as many monitoring programs and studies have not adhered to a uniform standard throughout their temporal and spatial extension.
-Additionally, certain regions or areas lack sufficient data or possess only limited information regarding krill dynamics.
-Furthermore, interpreting indirect indicators that may affect the krill population, such as environmental variables, predator abundance, and fishery data, adds another layer of complexity.
-Despite these gaps, the existing data about krill population dynamics is abundant, considered data rich, and this availability allows building an integrated model to get population signals and estimate main variables such as spawning biomass, recruitment, and fishing mortality, among others.
-In response to this need, we propose a conceptual model to conduct a statistical model in Antarctic krill population in the Antarctic Peninsula, which in turn allows the application of an integrated model that incorporates the fishery, surveys and ecosystem information on krill population over the past 40 years.
 
@@
 In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
+
+Predator indesx
+
+A simple predation pressure index for modeling changes in naturalmortality: Application to Gulf of Maine northern shrimp stockassessment (Richard et al 2016)
+
+Env index.
+
+Simulation analyses indicated that the integrated approach with additional process error is superior to the traditional method of correlating model estimates with environmental variables outside the estimation procedure.
+The results suggest that, for the snapper stock, recruitment is positively correlated with SST at the time of spawning.
+(Maunder et al 2003)
+
+The ecosystem variables used in stock assessment processes vary in both their definition and how they are applied.
+Marshall et al.
+(2019) classify ecosystem variables based on their impact on population dynamics, as well as the ways in which they are incorporated into analytical models.
 

### 2025-03-28 index.Rmd
save last chgs

@@
 # dir9<-here("s9") # s1 w/ blocks
-dir1.1<-here("s1.1") # con todo
-dir1.2 <- here("s1.2") # sin predador ni ambiental
-dir1.3 <- here("s1.3") # borrado los registros de prreddador )igual q el S2
+dir1.1<-here("s1.1") # sin predador ni ambiental
+dir1.2 <- here("s1.2") # s1.1 C/ predador
+dir1.3 <- here("s1.3") # s1.1 solo env
+dir1.4 <- here("s1.4") # s1.2 predator and env
 Figs <- here("Figs")# S
@@
 |    s1.1  | Spatial data without environmental and predator components                           |
-|    s1.2  | "s1.1" with environmental and predator components                                    |
-|    s1.3  | "s1.2" w/ selectivity blocks                                                         |
+|    s1.2  | "s1.1" with predator components                                                      |
+|    s1.3  | "s1.1" with environmental variable                                              |
+|    s1.4  | "s1.1" w/ both, predator fleet and environmental variable                        |
 
@@
 
+# s1.4
+r4ss::run(
+  dir = dir1.4,
+  exe = "ss_osx",
+  skipfinished = FALSE, # TRUE will skip running if Report.sso present
+  show_in_console = TRUE # change to true to watch the output go past
+)
 ```
@@
                         forecast=T)
+#s1.4
+base.model1.4 <- SS_output(dir=dir1.4,
+                        covar=T,
+                        forecast=T)
+
 ```
@@
 #note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist1_2_3 <- SSgetoutput(keyvec = NULL,
+biglist1_4 <- SSgetoutput(keyvec = NULL,
                        dirvec = c(
@@
                                   dir1.2,
-                                  dir1.3),
+                                  dir1.3,
+                                  dir1.4),
                        getcovar = F)
@@
 #create summary of model runs from list above
-summaryoutput1_2_3 <- SSsummarize(biglist1_2_3)
-
-SSplotComparisons(summaryoutput1_2_3,
+summaryoutput1_4 <- SSsummarize(biglist1_4)
+
+SSplotComparisons(summaryoutput1_4,
                   legendlabels = c("Ref Model: No Env-Predator",
                  "S1.1 w/ Predator data",
+                 "S1.1 w/ Env data",
                  "S1.1 w/ Env and Predator data"),
@@
 ```
+
+## Comparision RMSE
+
+```{r}
+# Crear un data frame con los valores de RMSE
+df <- data.frame(
+  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
+  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
+  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10])
+)
+
+# Comparación de estadísticas descriptivas
+summary(df)
+
+# Comparación con boxplot
+boxplot(df, main = "Comparación de RMSE entre modelos", col = c("red", "blue", "green"))
+
+# Prueba de diferencias estadísticas (ANOVA)
+anova_result <- aov(as.numeric(unlist(df)) ~ rep(1:3, each=nrow(df)))
+summary(anova_result)
+
+t.test(df$Model1.1, df$Model1.2)  # Comparar Model1.1 vs Model1.2
+t.test(df$Model1.1, df$Model1.3)  # Comparar Model1.1 vs Model1.3
+t.test(df$Model1.2, df$Model1.3)  # Comparar Model1.2 vs Model1.3
+
+# ANOVA para ver si hay diferencias globales
+anova_rmse <- aov(as.numeric(unlist(df)) ~ rep(1:3, each=nrow(df)))
+summary(anova_rmse)
+
+# Prueba post-hoc si ANOVA es significativa
+TukeyHSD(anova_rmse)
+
+df_long <- reshape2::melt(df)
+ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
+  geom_boxplot() +
+  theme_minimal() +
+  labs(title = "Comparación de RMSE entre modelos", x = "Modelo", y = "RMSE")
+
+```
+
+```{r}
+
+# Unir los modelos en un solo dataframe y asegurarse de que los nombres de filas se mantienen
+df1 <- base.model1.1$likelihoods_used %>%
+  select(values) %>%
+  mutate(Componente = rownames(base.model1.1$likelihoods_used), Modelo = "M1.1")
+
+df2 <- base.model1.2$likelihoods_used %>%
+  select(values) %>%
+  mutate(Componente = rownames(base.model1.2$likelihoods_used), Modelo = "M1.2")
+
+df3 <- base.model1.3$likelihoods_used %>%
+  select(values) %>%
+  mutate(Componente = rownames(base.model1.3$likelihoods_used), Modelo = "M1.3")
+
+# Combinar todos los modelos
+df <- bind_rows(df1, df2, df3)
+
+# Crear gráfico de barras con valores transformados
+ggplot(df, aes(x = Componente, y = values, fill = Modelo)) +
+  geom_col(position = position_dodge(), width = 0.7) +  # Barras separadas por modelo
+  theme_bw() +
+  coord_flip() +  # Para que sea horizontal
+  scale_fill_manual(values = c("S1.1" = "grey",
+                               "S1.2" = "red", 
+                               "S1.3" = "black")) +  # Colores personalizados
+  labs(x = "",
+       y = "Log-Normal Likelihood", 
+       title = "")
+
+```
+
 ## Platoons analisis

### 2025-03-28 MS2_Krill.Rmd
last changes in MS

@@
 
-# M&M
-
-#### **2.1. Study Area and Data Sources**  
-This study focuses on the Antarctic Peninsula, a region where krill populations exhibit strong variability due to environmental fluctuations and intense predation pressure.
-Data were obtained from the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR), including fishery-dependent catch data, scientific surveys, and predator monitoring datasets.
-Environmental data, such as sea surface temperature (SST), chlorophyll-a concentration, and sea ice extent, were sourced from satellite observations and oceanographic monitoring programs.
-
-#### **2.2. Catch-to-Length Model Implementation**  
-We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends.
-The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables.
-Key model parameters include:
-
-- Growth rates based on length-frequency distributions  
-- Natural mortality estimates derived from predator consumption data  
-- Stock-recruitment functions incorporating environmental predictors  
-- Fishery removals as a function of effort and environmental conditions
-
-#### **2.3. Statistical Analysis and Model Validation**  
-Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation against independent survey datasets.
-We tested multiple scenarios to assess the relative influence of environmental and predation variables on recruitment and biomass dynamics.
-
----
-
-### **3. Results**  
-Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
-
-1.
-**Stock-Recruitment Relationship**: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.
-2.
-**Predation Impact**: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.
-3. **Biomass Trends**: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.  
-4. **Fisheries Interaction**: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
-
-Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
-Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
-Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
-
-### **Tablas a incluir**
-1. **Comparación de parámetros clave entre escenarios:**
-   - Biomasa desovante (SSB, Spawning Stock Biomass) en años clave.
-   - Tasa de explotación (F).
-   - Reclutamiento medio.
-   - Relación B/B0 (biomasa actual respecto a la virgen).
-   - Relación F/FMSY.
-
-   | Parámetro       | Base | Con variables ecosistémicas |
-   |----------------|------|----------------------------|
-   | SSB (2020)     | X    | Y                          |
-   | SSB (2025)     | X    | Y                          |
-   | F (2020)       | X    | Y                          |
-   | Reclutamiento medio | X | Y                        |
-   | B/B0 (2020)    | X    | Y                          |
-   | F/FMSY (2020)  | X    | Y                          |
-
-2. **Resumen del ajuste del modelo:**
-   - Log-verosimilitud total.
-   - Verosimilitud de los datos de captura, índices de abundancia y composición de tallas/edades.
-   - Número de parámetros estimados.
-   - Error estándar de las estimaciones clave.
-
-3. **Comparación de reclutamiento por año:**
-   - Media y desviación estándar del reclutamiento en periodos clave (por ejemplo, 1990-2000, 2000-2010, 2010-2020).
-   - Probabilidad de años de reclutamiento bajo.
-
-### **Gráficos recomendados**
-1. **Evolución de la biomasa desovante (SSB) en el tiempo**  
-   - Líneas para cada escenario mostrando tendencias divergentes.
-   - Inclusión de intervalos de confianza.
-
-2. **Reclutamiento anual estimado por escenario**  
-   - Barras para cada año con comparación entre escenarios.
-   - Promedio móvil para suavizar tendencias.
-
-3. **Relación stock-reclutamiento con diferentes ajustes**  
-   - Ajustes de la curva con y sin variables ecosistémicas.
-   - Comparación de la fuerza del reclutamiento en diferentes condiciones ambientales.
-
-4. **Proyecciones de biomasa y F bajo diferentes escenarios de explotación**  
-   - Predicciones de la biomasa y la explotación para los próximos 5-10 años.
-   - Comparación entre políticas de manejo con y sin ecosistema.
-
-5. **Distribución de longitudes o edades en el tiempo**  
-   - Para ver si el escenario con variables ecosistémicas predice cambios en la estructura de la población.
-
-6. **Curvas de Kobe plot (B/BMSY vs. F/FMSY)**  
-   - Comparar el estado del stock en cada escenario respecto a puntos de referencia de manejo.
-
-(Imagen Verosimilitud)
-
-The comparison of likelihood components across the three scenarios revealed notable differences in their fit to the data.
-Scenario S1.1 had the lowest total likelihood, suggesting a better overall fit compared to S1.2 and S1.3.
-The catch and equilibrium catch components showed minor variations among scenarios, indicating that these data had a limited influence on the overall likelihood differences.
-In terms of survey likelihood, S1.1 exhibited a slightly better fit to survey abundance data, while S1.2 and S1.3 showed higher likelihood values, suggesting a reduced agreement with observed biomass trends.
-The length composition component differed significantly, with S1.2 and S1.3 having higher likelihood values than S1.1, implying a poorer fit to observed length distributions.
-Recruitment likelihood was highest in S1.1, indicating stronger consistency with recruitment estimates compared to S1.2 and S1.3.
-Additionally, parameter deviations were more constrained in S1.1, whereas S1.2 and S1.3 exhibited greater variability.
-These results suggest that S1.1 provides a better overall fit, particularly in survey abundance and length composition components, while S1.2 and S1.3 show greater discrepancies in these aspects.
-
----
-
-### **4. Discussion**  
-Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
-The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
-
-- **Ecosystem-Based Management Implications**: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
-By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.
-- **Predator-Krill Interactions**: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
-Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.
-- **Climate Change Considerations**: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
-Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
-
----
-
-### **5. Conclusion**  
-This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
-The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
-Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
-
-Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
-
----
-
-### **Acknowledgments**  
-We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
-
 \newpage
@@
 # METHODOLOGY
+
+## Study Area
 
@@
 
-## Statistical model Stock Synthesis (SS3)
+## Catch-to-Length Model Implementation
 
@@
 Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
+
+We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends.
+The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables.
+Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation.
+Key model parameters include:
+
+### Natural Mortality and Growth Parameterization
+
+The natural mortality rate (\(M\)) was specified using a single parameter per morphotype.
+Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization.
+The model assumed linear growth below a predefined settlement age of 1.5 years.
+The asymptotic length (\(L_{\infty}\)) was determined by the estimated length-at-age of 999 years, indicating that the model used \(L_{\infty}\) as an upper growth boundary.
+Exponential decay of growth beyond maximum age was disabled (, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
+
+Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
+
+### Maturity and Reproduction
+
+Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
+Fecundity was estimated as a power function of body weight, where fecundity was computed as \(Eggs = Wt \times (a + b \times Wt)\).
+The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
+
+The model used a direct parameterization approach for natural mortality (\(M\)), growth (\(G\)), and growth variability (\(CV_G\)), without additional offsets between sexes.
+Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
+
+### Selectivity Parameterization
+
+Selectivity functions were modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
+The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
+
+The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
+The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm.
+Parameter estimation phases were assigned, with negative values indicating fixed parameters.
+
+The survey selectivity was also modeled using a length-based logistic function.
+Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
+The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
+
+The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
+These parameters were estimated in later phases (positive phase values), allowing flexibility in their estimation.
+
+Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization.
+
+### Spawner-Recruitment Relationship Parameterization
+
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation (option 3).
+This model was selected due to its robustness in representing recruitment dynamics in response to spawning biomass.
+
+The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value ($SR\_LN(R_0)$) of 23, with a prior of 10.
+The parameter was freely estimated in phase 1.
+The steepness of the stock-recruitment curve ($SR\_BH\_steep$) was fixed at 0.85 in phase -4, ensuring consistency with previous studies.
+The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
+The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
+The $SR\_autocorr$ parameter was set to 0 and fixed, indicating that no temporal autocorrelation in recruitment deviations was assumed.
+
+Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
+The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
+The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
+
+The recruitment deviation phase was set to 2.
+Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
+Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
+The maximum bias adjustment factor was 0.8274.
+
+To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
+No cyclic patterns in recruitment were assumed, and advanced recruitment options were enabled for model flexibility.
 
@@
 
-## Stock Assessment Scenarios
+### Stock Assessment Scenarios
 
@@
 |:------------:|:---------------------------------------------------------|
-|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental data, aggregated at 48.1 level               |
-|    s1    | Fishery and Survey (AMLR) Length data, CPUE, Catch by strata. Predator and Env data |
-|    s2    | "s1" without Stock-Recruit relation                                      |
-|    s3    | "s1" Beverton & Holt S-R relation weak (0.9 steepness)                   |
-|    s4    | "s1" BH S-R relation strong (0.6 steepness)   |
-|    s5    | "s1" BH S-R relation mid-strong estimated    |
-|    s6    | "s1" Ricker S-R relation estimated    |
-|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par   |
+|    s1.1  | Spatial data without environmental and predator components                           |
+|    s1.2  | "s1.1" with predator components                                                      |
+|    s1.3  | "s1.1" with environmental variable                                              |
+|    s1.4  | "s1.1" w/ both, predator fleet and environmental variable                        |
 
@@
 
-For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).
+## Results
+
+Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
+
+1.
+Stock-Recruitment Relationship: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.
+2.
+Predation Impact: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.
+3. Biomass Trends: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.  
+4. Fisheries Interaction: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
+
+Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
+Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
+Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
+
+### Tablas a incluir
+1. Comparación de parámetros clave entre escenarios:
+   - Biomasa desovante (SSB, Spawning Stock Biomass) en años clave.
+   - Tasa de explotación (F).
+   - Reclutamiento medio.
+   - Relación B/B0 (biomasa actual respecto a la virgen).
+   - Relación F/FMSY.
+
+   | Parámetro       | Base | Con variables ecosistémicas |
+   |----------------|------|----------------------------|
+   | SSB (2020)     | X    | Y                          |
+   | SSB (2025)     | X    | Y                          |
+   | F (2020)       | X    | Y                          |
+   | Reclutamiento medio | X | Y                        |
+   | B/B0 (2020)    | X    | Y                          |
+   | F/FMSY (2020)  | X    | Y                          |
+
+2. Resumen del ajuste del modelo:
+   - Log-verosimilitud total.
+   - Verosimilitud de los datos de captura, índices de abundancia y composición de tallas/edades.
+   - Número de parámetros estimados.
+   - Error estándar de las estimaciones clave.
+
+3. Comparación de reclutamiento por año:
+   - Media y desviación estándar del reclutamiento en periodos clave (por ejemplo, 1990-2000, 2000-2010, 2010-2020).
+   - Probabilidad de años de reclutamiento bajo.
+
+### Gráficos recomendados
+1. Evolución de la biomasa desovante (SSB) en el tiempo  
+   - Líneas para cada escenario mostrando tendencias divergentes.
+   - Inclusión de intervalos de confianza.
+
+2. Reclutamiento anual estimado por escenario  
+   - Barras para cada año con comparación entre escenarios.
+   - Promedio móvil para suavizar tendencias.
+
+3. Relación stock-reclutamiento con diferentes ajustes  
+   - Ajustes de la curva con y sin variables ecosistémicas.
+   - Comparación de la fuerza del reclutamiento en diferentes condiciones ambientales.
+
+4. Proyecciones de biomasa y F bajo diferentes escenarios de explotación  
+   - Predicciones de la biomasa y la explotación para los próximos 5-10 años.
+   - Comparación entre políticas de manejo con y sin ecosistema.
+
+5. Distribución de longitudes o edades en el tiempo  
+   - Para ver si el escenario con variables ecosistémicas predice cambios en la estructura de la población.
+
+6. Curvas de Kobe plot (B/BMSY vs. F/FMSY)  
+   - Comparar el estado del stock en cada escenario respecto a puntos de referencia de manejo.
+
+(Imagen Verosimilitud)
+
+The comparison of likelihood components across the three scenarios revealed notable differences in their fit to the data.
+Scenario S1.1 had the lowest total likelihood, suggesting a better overall fit compared to S1.2 and S1.3.
+The catch and equilibrium catch components showed minor variations among scenarios, indicating that these data had a limited influence on the overall likelihood differences.
+In terms of survey likelihood, S1.1 exhibited a slightly better fit to survey abundance data, while S1.2 and S1.3 showed higher likelihood values, suggesting a reduced agreement with observed biomass trends.
+The length composition component differed significantly, with S1.2 and S1.3 having higher likelihood values than S1.1, implying a poorer fit to observed length distributions.
+Recruitment likelihood was highest in S1.1, indicating stronger consistency with recruitment estimates compared to S1.2 and S1.3.
+Additionally, parameter deviations were more constrained in S1.1, whereas S1.2 and S1.3 exhibited greater variability.
+These results suggest that S1.1 provides a better overall fit, particularly in survey abundance and length composition components, while S1.2 and S1.3 show greater discrepancies in these aspects.
+
+---
+
+### 4. Discussion  
+Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
+The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
+
+- Ecosystem-Based Management Implications: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
+By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.
+- Predator-Krill Interactions: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
+Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.
+- Climate Change Considerations: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
+Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
+
+---
+
+### 5. Conclusion  
+This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
+The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
+Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
+
+Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
+
+---
+
+### Acknowledgments  
+We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
 

### 2025-03-28 index.Rmd
last changes in MS

@@
 SSplotData(base.model1.1, 
-           subplot = 1,
+           subplots = 1,
            pheight = 15)
 
-SSplotData(base.model1.2, 
-           subplot = 1,
+SSplotData(base.model1.4, 
+           subplots = 1,
            pheight = 15)
@@
 SS_plots(base.model1.3, 
+         uncertainty = TRUE,
+         datplot = T, 
+         png=T, 
+         aalresids = F,
+         btarg=0.75, 
+         minbthresh=0.2, 
+         forecast=T)
+
+# S 1.4
+
+# Definir la ruta de la carpeta
+folder_path4 <- "s1.4/plots"
+# Verificar si la carpeta existe y eliminarla
+if (dir.exists(folder_path4)) {
+  unlink(folder_path4, 
+         recursive = TRUE)
+  message("La carpeta 'plots' ha sido eliminada.")
+} else {
+  message("La carpeta 'plots' no existe.")
+}
+
+SS_plots(base.model1.4, 
          uncertainty = TRUE,
@@
 ```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
-alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz
+alk_matrix <- base.model1.2$ALK[,,1] # Ajusta las dimensiones según tu matriz
 

### 2025-03-31 index.Rmd
adding mod

@@
 ```{r}
-SSplotTimeseries(base.model1.3,
+SSplotTimeseries(base.model1.4,
                  subplot = 1)
-SSplotTimeseries(base.model1.3,
+SSplotTimeseries(base.model1.4,
                  subplot = 9)
-SSplotTimeseries(base.model1.3,
+SSplotTimeseries(base.model1.4,
                  subplot = 7)
@@
             subplots = 1)
-```
-
-```{r}
-SSplotBiology(base.model1.1,
+SSplotSelex(base.model1.3,
+            subplots = 1)
+
+SSplotSelex(base.model1.4,
+            subplots = 1)
+```
+
+```{r}
+SSplotIndices(base.model1.1,
             subplots = 1:32)

### 2025-04-01 index.Rmd
adding las  chgs in html

--- 2025-03-28 index.Rmd
@@
                       warning = FALSE,
+                      fig.height = 6,
+                      fig.width = 7,
                       fig.align = 'center',
@@
 
-## Overview
+# OVERVIEW
 
@@
 ```{r}
-start1 <- SS_readstarter(file = file.path(dir1.1,
+start1 <- SS_readstarter(file = file.path(dir1.4,
                                           "starter.ss"),
                               verbose = FALSE)
-dat1 <- SS_readdat(file = file.path(dir1.2, start1$datfile),
+dat1 <- SS_readdat(file = file.path(dir1.4, start1$datfile),
                         verbose = FALSE)
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.4,
                                     start1$ctlfil),
@@
                    datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1.4, 
                                                 "forecast.ss"),
@@
 dat1_std <- dat1$CPUE %>%
-  group_by(index) %>%
-  mutate(index_std = obs / mean(obs, na.rm = TRUE)) %>%
-  ungroup() %>% 
-  mutate(index = recode(index,
+  mutate(
+    # Standardize 'obs' between 0 and 1
+    index_std = (obs - min(obs)) / (max(obs) - min(obs)),
+
+    # Recode 'index' values
+    index = recode(index,
                    `1` = "FISHERYBS",
@@
                    `10` = "SURVEYSSIW",
-                        `11` = "PREDATOR"))
-ggplot(dat1_std, aes(x = year, y = index_std, group=index))+
+                   `11` = "PREDATOR"),
+
+    # Convert 'index' to a factor with specific levels
+    index = factor(index, levels = c("FISHERYBS", 
+                                     "FISHERYEI", 
+                                     "FISHERYGS",
+                                     "FISHERYJOIN", 
+                                     "FISHERYSSIW", 
+                                     "SURVEYBS", 
+                                     "SURVEYEI", 
+                                     "SURVEYGS", 
+                                     "SURVEYJOIN", 
+                                     "SURVEYSSIW", 
+                                     "PREDATOR")),
+
+    # Assign categories based on 'index'
+    category = case_when(
+      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
+      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
+      index == "PREDATOR" ~ "Predator",
+      TRUE ~ NA_character_  # In case there are any unexpected values
+    )
+  )
+
+# Gráfico con 3 filas, dejando 'PREDATOR' en la última fila
+ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
   geom_point() +
-  geom_smooth(method="lm",
-              se=F)+
-  facet_wrap(.~index,
-             scales = "free_y",
-             ncol=6)+
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~index, scales = "free_y", ncol = 5) +  # Ajusta ncol para obtener 3 filas
+  scale_color_manual(values = c("Fishery" = '#1b9e77', 
+                                "Survey" = '#d95f02',
+                                "Predator" = '#7570b3')) + 
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, 
-                                   hjust = 1))+
-  labs(title = "",
-       x = "",
-       y = "Standar index",
-       color = "index")
-```
-
-```{r fig.width=10, fig.height=4}
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  labs(title = "", x = "", y = "Standard index", color = "Category")
+
+```
+
+```{r fig.width=10, fig.height=4, warning=FALSE}
 dat_long <- dat1$lencomp %>% 
@@
 
-# Expandir datos según la frecuencia
 dat_expanded <- dat_long %>%
@@
 
-# Crear el gráfico con líneas verticales en cada facet
-ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), 
-                         fill = stat(x))) +
+ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
   geom_density_ridges_gradient(scale = 1.8,
@@
                                col = "black",
-                               fill = "grey") +
+                               alpha=0.4) +
   facet_wrap(~FltSvy, ncol = 10) +
@@
              size = 0.7,
-             alpha=0.5) +
+             alpha = 0.5) +
+  scale_fill_manual(values = c("FISHERYBS" = '#1b9e77',  
+                               "FISHERYEI" = '#1b9e77',  
+                               "FISHERYGS" = '#1b9e77',
+                               "FISHERYJOIN" = '#1b9e77',  
+                               "FISHERYSSIW" = '#1b9e77',  
+                               "SURVEYBS" = '#d95f02',  
+                               "SURVEYEI" = '#d95f02',  
+                               "SURVEYGS" = '#d95f02',  
+                               "SURVEYJOIN" = '#d95f02',  
+                               "SURVEYSSIW" = '#d95f02',  
+                               "PREDATOR" = '#7570b3')) +  
   labs(title = "",
@@
        y = "",
-       fill = "Densidad") +
-  theme_few()
+       fill = "Category") +
+  theme_few()+
+  theme(legend.position = "none")
+
 ```
@@
 
-## Run Models
+### Run Models
 
@@
                  "s1.2",
-                 "s1.3")
+                 "s1.3",
+                 "s1.4")
 
@@
   exe = "ss_osx",
-  skipfinished = FALSE, # TRUE will skip running if Report.sso present
-  show_in_console = TRUE # change to true to watch the output go past
+  skipfinished = FALSE, 
+  show_in_console = TRUE 
 )
@@
 
-Data used en both (spatial and No spatial models)
+Data used en both (spatial and No spatial models) `s1.1`
 
@@
            pheight = 15)
-
+```
+
+and `s1.4`
+
+```{r }
 SSplotData(base.model1.4, 
@@
 ```
+
 # RESULTS
@@
 
-### Main Variables poulation in `s1.1` scenario
+Main Variables poulation in `s1.1` scenario
 
@@
 
-### Main Variables poulation in `s1.2` scenario
+Main Variables poulation in `s1.2` scenario
 
@@
 
-### Main Variables poulation in `s1.3` scenario
+Main Variables poulation in `s1.3` scenario
 
@@
 
+Main Variables poulation in `s1.4` scenario
+
+```{r}
+SSplotTimeseries(base.model1.4,
+                 subplot = 1)
+SSplotTimeseries(base.model1.4,
+                 subplot = 9)
+SSplotTimeseries(base.model1.4,
+                 subplot = 7)
+```
 Selectivity
 
-```{r fig.height=6}
+```{r}
 SSplotSelex(base.model1.1,
             subplots = 1)
-
 SSplotSelex(base.model1.2,
             subplots = 1)
-```
-
-```{r}
-SSplotBiology(base.model1.1,
-            subplots = 1:32)
-```
-
-```{r}
+SSplotSelex(base.model1.3,
+            subplots = 1)
+SSplotSelex(base.model1.4,
+            subplots = 1)
+```
+
+```{r}
+SSplotIndices(base.model1.1,
+            subplots = 9)
+SSplotIndices(base.model1.2,
+            subplots = 9)
+SSplotIndices(base.model1.3,
+            subplots = 9)
+SSplotIndices(base.model1.4,
+            subplots = 9)
+```
+
+```{r}
+SSplotComps(base.model1.1,
+            subplots = 21)
 SSplotComps(base.model1.2,
-            subplots = 1)
-
+            subplots = 21)
+SSplotComps(base.model1.3,
+            subplots = 21)
+SSplotComps(base.model1.4,
+            subplots = 21)
 ```
@@
 
-```{r eval=FALSE}
-total01 <- as.data.frame(base.model1.1$timeseries[1:31,c(2,5)]) %>%
-  mutate(Serie = "No Spatial")
-total1 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial implicit")
-total2 <- as.data.frame(base.model2$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ without S-R relation")
-total3 <- as.data.frame(base.model3$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ S-R relation weak (0.9 steepness)")
-total4 <- as.data.frame(base.model4$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ S-R relation strong (0.6 steepness)")
-total5 <- as.data.frame(base.model5$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ S-R relation mid-strong estimated ")
-total6 <- as.data.frame(base.model6$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ Ricker S-R relation estimated")
-total7 <- as.data.frame(base.model7$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "Spatial w/ New set parameters Linf, k and M")
-
-totalbio <- rbind(total01,
-              total1,
+```{r}
+total1 <- as.data.frame(base.model1.1$timeseries[,c(2,5)]) %>%
+  mutate(Serie = "Without Ecosystem Variables")
+total2 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "W/ Predator")
+total3 <- as.data.frame(base.model1.3$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "W/ Environment")
+total4 <- as.data.frame(base.model1.4$timeseries[,c(2,5)]) %>% 
+  mutate(Serie = "With Predator and Ecosystem")
+
+totalbio <- rbind(total1,
               total2,
               total3,
-              total4,
-              total5,
-              total6,
-              total7)
+              total4)
 
 ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
-  geom_point(size=0.7,
-             col="black") +
+  geom_point() +
   geom_smooth(method = "loess",
-              span=0.75, 
-              se=TRUE)+
-  labs(title = "Compare variables between scenarios",
+              span=0.5, 
+              se=F)+
+  labs(title = "",
        x = "",
@@
 
-```{r eval=FALSE}
-SSplotComps(base.model01,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model1,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model2,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model3,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model4,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model5,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model6,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-SSplotComps(base.model7,
-            subplots = 1,
-            maxrows = 5,
-            maxcols = 5)
-```
 Heatmap
@@
 ```{r}
-df <- base.modeltest$lendbase
-
-# Filtra el dataframe
+df <- base.model1.4$lendbase
+
 df1 <- df %>% 
@@
 # Filter and select relevant columns
-data <- base.model1$lendbase %>%
+data <- base.model1.4$lendbase %>%
   dplyr::filter(Pearson<5) %>% 
@@
 
-ggplot(data, aes(x = Yr, y = Bin)) +
-  geom_hex(bins = 50) +
-  scale_fill_gradient2(low = "blue", 
-                       mid = "white", 
-                       high = "red", midpoint = median(data$Bin)) +
-  theme_minimal() +
-  labs(title = "Hexagonal Heatmap of Pearson Residuals",
-       x = "Year", y = "Bin", fill = "Pearson Residual")
 ```
@@
 
-Step to do a good practice in model diagnosis is;
-
--     1.  Convergence. Final convergence criteria is 1.0e-04
-
--     2.  Residual (visual and metrics)
-
--     3.  Retrospective analysis (Mhon Parameter)
-
--    4.  Likelihood profile
-
-all this framework try to follow recommendations of @Carvalho2021b
-
-## Residual consistency
+A rigorous model diagnosis is essential to ensure the reliability and robustness of stock assessment models. The key steps for a good practice in model diagnosis include:
+
+1. Convergence Check: The model must reach a final convergence criterion of 1.0e-04 to ensure numerical stability and reliable parameter estimation.
+
+2. Residual Analysis: Both visual inspection and statistical metrics are used to evaluate model residuals, helping to detect patterns of bias or misfit.
+
+3.
+Retrospective Analysis: The Mohn’s rho parameter is used to assess the consistency of model estimates when sequentially removing recent years of data, identifying potential overestimation or underestimation trends.
+
+4. Likelihood Profile Analysis: This approach examines how the likelihood function behaves across a range of parameter values, providing insight into parameter uncertainty and model sensitivity.
+
+This framework follows the recommendations outlined by @Carvalho2021b, aiming to enhance transparency and reproducibility in model evaluation.
+
+### Residual consistency
+
+Residual analysis is a critical component of model diagnostics in stock assessments.
+It helps evaluate the fit of the model to observed data and detect potential biases or inconsistencies.
+This process is applied to both length composition data and abundance indices such as CPUE (Catch Per Unit Effort) and survey-derived estimates.
+
+For length composition data, residuals represent the difference between observed and model-predicted length distributions.
+The standardized residuals are calculated as the difference between observed and expected proportions at each length bin.
+These residuals are plotted by year to identify systematic trends, biases, or inconsistencies in the data.
+Ideally, they should be randomly distributed around zero, indicating no systematic over- or underestimation.
+
+For abundance indices such as CPUE and fishery-independent surveys, residuals are analyzed to assess model fit and potential sources of bias.
+Residuals are computed as the difference between observed index values and those predicted by the model, typically standardized by dividing by the standard error to facilitate comparison across years.
+These residuals are then plotted over time to evaluate trends.
+A shaded confidence region, like the green area in the provided plot, represents expected variability, with outliers highlighted in red or other distinct markers.
+Persistent positive or negative residuals may indicate systematic bias in the model or data collection process.
+
+Statistical diagnostics are also performed to check for autocorrelation in residuals, which can indicate potential model misspecifications.
+When mean residual values are close to zero, the model fit is considered unbiased.
+
+By integrating these residual analyses for both length and abundance indices, stock assessment models can be refined, improving their reliability and increasing confidence in the assessment results.
 
@@
                plotdir = Figs)
-```
-
-```{r eval=FALSE}
-SSplotRunstest(base.model01,
+SSplotRunstest(base.model1.3,
                subplots = "len",
-               add=T)
-SSplotRunstest(base.model1,
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
+SSplotRunstest(base.model1.4,
                subplots = "len",
-               add=T)
-SSplotRunstest(base.model2,
-               subplots = "len",
-               add=T)
-SSplotRunstest(base.model3,
-               subplots = "len",
-               add=T)
-SSplotRunstest(base.model4,
-               subplots = "len",
-               add=T)
-SSplotRunstest(base.model5,
-               subplots = "len",
-               add=T)
-SSplotRunstest(base.model6,
-               subplots = "len",
-               add=T)
-SSplotRunstest(base.modeltest,
-               subplots = "len",
-               add=T)
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
 ```
@@
                subplots = "cpue",
-               add=T)
+               add=T,
+               print = TRUE)
 SSplotRunstest(base.model1.2,
@@
                add=T)
-```
-
-```{r eval=FALSE}
-SSplotRunstest(base.model01,
+SSplotRunstest(base.model1.3,
                subplots = "cpue",
                add=T)
-SSplotRunstest(base.model1,
+SSplotRunstest(base.model1.4,
                subplots = "cpue",
                add=T)
-SSplotRunstest(base.model2,
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model3,
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model4,
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model5,
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.model6,
-               subplots = "cpue",
-               add=T)
-SSplotRunstest(base.modeltest,
-               subplots = "cpue",
-               add=T)
-```
-
-```{r eval=FALSE}
-SSplotJABBAres(base.model01,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model1,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model2,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model3,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model4,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model5,
-               subplots = "len",
-               add=T)
-SSplotJABBAres(base.model6,
+```
+
+### Residual Analysis and RMSE
+
+Residual analysis of mean length data is a fundamental diagnostic tool in stock assessments.
+It helps evaluate whether the model provides an unbiased fit to the observed data and detects potential biases over time.
+In this figure, mean length residuals are plotted across years, differentiated by data source, including fishery-dependent (FISHERY) and fishery-independent (SURVEY) datasets, as well as predator-related observations (PREDATOR).
+The residuals represent the deviation of observed mean length from model-predicted values, standardized to facilitate interpretation.
+
+The black line represents a locally estimated scatterplot smoothing (Loess) curve, which provides a trend line to visualize systematic deviations over time.
+The presence of persistent positive or negative trends in the residuals may indicate biases in the growth model, selectivity assumptions, or misrepresentation of recruitment variability.
+The gray bars highlight periods where residual variability is particularly high, suggesting potential inconsistencies between observed and predicted size structures.
+
+RMSE quantifies the overall deviation between observed and predicted values, providing an aggregate measure of model fit.
+Lower RMSE values indicate better agreement between observed and predicted data.
+In fisheries stock assessment [@HurtadoF2015], RMSE thresholds for acceptable model performance typically range between *10% and 30%*, depending on the data quality and complexity of the population dynamics being modeled.
+Values exceeding this range suggest potential biases, requiring further investigation into the model structure, parameter estimation, or data sources.
+
+By analyzing residual patterns and RMSE values, the model can be refined to improve the accuracy of mean length predictions, ultimately enhancing the reliability of stock assessment outcomes and management recommendations.
+
+```{r}
+SSplotJABBAres(base.model1.1,
                subplots = "len",
@@
                add=T)
-```
-
-```{r}
-SSplotJABBAres(base.model01,
-               subplots = "cpue",
+SSplotJABBAres(base.model1.3,
+               subplots = "len",
                add=T)
-SSplotJABBAres(base.model1,
-               subplots = "cpue",
+SSplotJABBAres(base.model1.4,
+               subplots = "len",
                add=T)
-SSplotJABBAres(base.model2,
-               subplots = "cpue",
-               add=T)
-SSplotJABBAres(base.model3,
-               subplots = "cpue",
-               add=T)
-SSplotJABBAres(base.model4,
-               subplots = "cpue",
-               add=T)
-SSplotJABBAres(base.model5,
-               subplots = "cpue",
-               add=T)
-SSplotJABBAres(base.model6,
+```
+
+```{r}
+SSplotJABBAres(base.model1.1,
                subplots = "cpue",
@@
                add=T)
-```
-## Relationship Stock-Recruit
-
-```{r eval=FALSE}
-l <- base.modeltest$recruit$SpawnBio
-m <- base.modeltest$recruit$pred_recr
+SSplotJABBAres(base.model1.3,
+               subplots = "cpue",
+               add=T)
+SSplotJABBAres(base.model1.4,
+               subplots = "cpue",
+               add=T)
+```
+
+### Comparision RMSE
+
+```{r}
+# Crear un data frame con los valores de RMSE
+dfrmse <- data.frame(
+  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
+  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
+  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
+  Model1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])
+
+)
+summary(dfrmse)
+
+# Comparación con boxplot
+boxplot(dfrmse, main = "RMSE comparision", col = c('#ca0020','#f4a582','#bababa','#404040'))
+
+# Prueba de diferencias estadísticas (ANOVA)
+anova_result <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
+summary(anova_result)
+
+t.test(dfrmse$Model1.1, dfrmse$Model1.2)  # Comparar Model1.1 vs Model1.2
+t.test(dfrmse$Model1.1, dfrmse$Model1.3)  # Comparar Model1.1 vs Model1.3
+t.test(dfrmse$Model1.2, dfrmse$Model1.3)
+t.test(dfrmse$Model1.3, dfrmse$Model1.4)
+# ANOVA para ver si hay diferencias globales
+anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
+summary(anova_rmse)
+
+# Prueba post-hoc si ANOVA es significativa
+#TukeyHSD(anova_rmse)
+
+dfrmse_long <- reshape2::melt(dfrmse)
+
+ggplot(dfrmse_long, aes(x = variable, y = value, fill = variable)) +
+  geom_boxplot(width=0.3) +
+  geom_jitter(width = 0.01, alpha = 0.3) +
+  theme_minimal() +
+  scale_fill_manual(name= "Scenario",
+                    values = c("Model1.1" = "#ca0020", 
+                               "Model1.2" = "#f4a582", 
+                               "Model1.3" = "#bababa", 
+                               "Model1.4" = "#404040")) +
+  labs(title = "", x = "", y = "RMSE")
+
+```
+This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill.
+RMSE serves as an indicator of model accuracy, with lower values representing better predictive performance.
+
+s1.1 exhibits the lowest median RMSE, suggesting it has the best overall fit among the four models.
+In contrast, Models 1.2, 1.3, and 1.4 show higher RMSE values, indicating comparatively lower predictive accuracy.
+The interquartile range (IQR) of these models is relatively similar, suggesting comparable variability in RMSE across models.
+Additionally, s1.1 has an outlier above 1.5 RMSE, which could indicate a case where the model's predictions deviated significantly from observed values.
+
+Overall, this analysis highlights that incorporating different environmental or predator-related variables in the models impacts their predictive ability.
+The differences in RMSE suggest that some models may overfit or underfit the recruitment patterns of krill, emphasizing the need to refine model selection based on ecological and statistical considerations.
+
+### Relationship Stock-Recruit
+
+\[
+R = \frac{aS}{1 + bS}
+\]
+
+Where:  
+- \( S \) is the spawning stock biomass.  
+- \( R \) is the predicted recruitment.  
+- \( a \) is the maximum recruitment capacity.  
+- \( b \) regulates density dependence (higher values of \( b \) result in a lower plateau).
+
+The blue line will show the asymptotic curve that describes the relationship between spawning stock biomass and recruitment.
+
+```{r}
+# First dataset
+l <- base.model1.1$recruit$SpawnBio  
+m <- base.model1.1$recruit$pred_recr 
+yr <- base.model1.1$timeseries$Yr[1:35]
+
+# Second dataset
+leco <- base.model1.4$recruit$SpawnBio  
+meco <- base.model1.4$recruit$pred_recr 
+yreco <- base.model1.4$timeseries$Yr[1:35]
+
+# Beverton-Holt models
+bev_holt_model1 <- nls(m ~ (a * l) / (1 + b * l), 
+                        start = list(a = max(m), b = 1 / max(l)), 
+                        data = data.frame(l, m))
+
+bev_holt_model2 <- nls(meco ~ (a * leco) / (1 + b * leco), 
+                        start = list(a = max(meco), b = 1 / max(leco)), 
+                        data = data.frame(leco, meco))
+
+# Predictions
+l_pred1 <- seq(min(l), max(l), length.out = 100)
+m_pred1 <- predict(bev_holt_model1, newdata = data.frame(l = l_pred1))
+
+l_pred2 <- seq(min(leco), max(leco), length.out = 100)
+m_pred2 <- predict(bev_holt_model2, newdata = data.frame(leco = l_pred2))
+
+# Create data frames for plotting
+data1 <- data.frame(l = l, m = m, yr = yr, Group = "s1.1")
+data2 <- data.frame(l = leco, m = meco, yr = yreco, Group = "s1.4")
+
+# Plot
 ggplot() +
-  aes(l, m) +
-  geom_point() +
-  geom_smooth(method = "loess",
-              se = TRUE, 
-              color = "blue") +
+  geom_point(data = data1, aes(x = l, y = m, shape = Group),
+             size = 1.5) +
+  geom_text(data = data1, aes(x = l, y = m, label = yr), 
+            hjust = 0.5, vjust = -1.5, size = 3) +  
+  geom_line(data = data.frame(l = l_pred1, m = m_pred1), 
+            aes(x = l, y = m), linewidth = 1,
+            col="blue") +
+  geom_point(data = data2, aes(x = l, y = m, shape = Group), 
+             size = 1.5) +
+  geom_text(data = data2, aes(x = l, y = m, label = yr), 
+            hjust = 0.5, vjust = 1.5, size = 3) +
+  geom_line(data = data.frame(l = l_pred2, m = m_pred2), 
+            aes(x = l, y = m), linewidth = 1,
+            col="blue") +
+  scale_shape_manual(values = c("s1.1" = 5, 
+                                "s1.4" = 4)) +
   theme_few() +
-  ylim(0, max(base.modeltest$recruit$pred_recr))
-
-```
-
-## Retrospective analysis
-
-Los análisis retrospectivo, dan cuenta de diferencias de estimación
-(sub - sobreestimación) en los patrones entre modelos evaluados.
+  ylim(0, max(m, meco)) +
+  labs(x = "SSB", y = "Recruit", 
+       title = "", shape = "Model") +
+  theme(legend.position = "bottom")
+
+```
+
+### Retrospective Analysis in Model Evaluation
+
+Retrospective analyses provide insights into the differences in estimation patterns (underestimation or overestimation) among the models evaluated.
+These analyses assess the consistency and reliability of stock assessment models by systematically removing the most recent years of data and comparing the resulting estimates with the full dataset.
+
+In this study, we conducted a retrospective analysis to examine the sensitivity of our recruitment and spawning stock biomass (SSB) estimates to the inclusion or exclusion of recent data.
+By applying this approach to multiple models, we identified potential biases and evaluated the stability of the recruitment estimates over time.
+
+The retrospective patterns were assessed by calculating the relative error between the predictions of truncated datasets and the full dataset.
+These differences allowed us to detect trends in model performance, such as systematic overestimation or underestimation of key population parameters.
+Understanding these deviations is crucial for improving the robustness of the stock assessment models and ensuring more reliable projections for fisheries management.
 
@@
 ```{r eval=FALSE}
+directorios <- c("s1.1",
+                 "s1.2",
+                 "s1.3",
+                 "s1.4")  
 for (dir in directorios) {
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-#s01
-retroModels01 <- SSgetoutput(dirvec=file.path(dir01,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary01 <- SSsummarize(retroModels01)
-# s1
-retroModels1 <- SSgetoutput(dirvec=file.path(dir1,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary1 <- SSsummarize(retroModels1)
-#s2
-retroModels2 <- SSgetoutput(dirvec=file.path(dir2,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary2 <- SSsummarize(retroModels2)
-#s3
-retroModels3 <- SSgetoutput(dirvec=file.path(dir3,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary3 <- SSsummarize(retroModels3)
-#s4
-retroModels4 <- SSgetoutput(dirvec=file.path(dir4,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary4 <- SSsummarize(retroModels4)
-#s5
-retroModels5 <- SSgetoutput(dirvec=file.path(dir5,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary5 <- SSsummarize(retroModels5)
-#s6
-retroModels6 <- SSgetoutput(dirvec=file.path(dir6,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary6 <- SSsummarize(retroModels6)
-#s7
-retroModels7 <- SSgetoutput(dirvec=file.path(dir7,
-                                            "Retrospective",
-                                            paste("retro",0:-5,
-                                                  sep="")))
-
-retroSummary7 <- SSsummarize(retroModels7)
-```
-
-```{r eval=FALSE, message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE}
 #stest
@@
 retroSummary1.1 <- SSsummarize(retroModels1.1)
-
-```
-
-```{r echo=FALSE}
+#stest
+retroModels1.2 <- SSgetoutput(dirvec=file.path(dir1.2,
+                                            "Retrospective",
+                                            paste("retro",0:-4,
+                                                  sep="")))
+
+retroSummary1.2 <- SSsummarize(retroModels1.2)
+#stest
+retroModels1.3 <- SSgetoutput(dirvec=file.path(dir1.3,
+                                            "Retrospective",
+                                            paste("retro",0:-4,
+                                                  sep="")))
+
+retroSummary1.3 <- SSsummarize(retroModels1.3)
+#stest
+retroModels1.4 <- SSgetoutput(dirvec=file.path(dir1.4,
+                                            "Retrospective",
+                                            paste("retro",0:-4,
+                                                  sep="")))
+
+retroSummary1.4 <- SSsummarize(retroModels1.4)
+
+```
+
+```{r}
 #save(retroSummary, retroModels, file="retro5.Rdata")
-
-retro01 <- SSplotRetro(retroSummary01,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro1 <- SSplotRetro(retroSummary1,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro2 <- SSplotRetro(retroSummary2,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro3 <- SSplotRetro(retroSummary3,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro4 <- SSplotRetro(retroSummary4,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-retro5 <- SSplotRetro(retroSummary5,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
-retro6 <- SSplotRetro(retroSummary6,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
-retro7 <- SSplotRetro(retroSummary7,
-            add=T,
-            forecast = F,
-            legend = T,
-            verbose=F)
-
 retro1.1 <- SSplotRetro(retroSummary1.1,
@@
             subplots = c("SSB"))
-```
-
-## Hindcast Cross-Validation and prediction skill
-
-Implementing the Hindcast Cross-Validation (HCxval) diagnostic in Stock
-Synthesis requires the same model outputs generated by
-`r4ss:SS_doRetro()`. As a robust measure of prediction skill, we
-implemented the mean absolute scaled error (MASE). In brief, the MASE
-score scales the mean absolute. Regarding (A MASE score \> 1 indicates
-that the average model forecasts are worse than a random walk.
-Conversely, a MASE score of 0.5 indicates that the model forecasts twice
-as accurately as a naïve baseline prediction; thus, the model has
-prediction skill.
-
-```{r}
-hci = SSplotHCxval(retroSummarytest, 
+retro1.2 <- SSplotRetro(retroSummary1.2,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("SSB"))
+retro1.3 <- SSplotRetro(retroSummary1.3,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("SSB"))
+retro1.4 <- SSplotRetro(retroSummary1.4,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("SSB"))
+```
+
+### Hindcast Cross-Validation and Prediction Skill
+
+The Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis is implemented using the model outputs generated by the `r4ss::SS_doRetro()` function.
+This diagnostic evaluates the predictive performance of the model by comparing hindcast predictions with observed data.
+
+To assess prediction skill, we employ the Mean Absolute Scaled Error (MASE) as a robust metric.
+MASE is calculated by scaling the mean absolute error of the model predictions relative to the mean absolute error of a naïve baseline prediction.
+Specifically, the MASE score is computed as follows:
+
+- A MASE score greater than 1 indicates that the model’s average forecasts are less accurate than a random walk model.
+- A MASE score equal to 1 suggests that the model’s accuracy is similar to that of a random walk.
+- A MASE score less than 1 indicates that the model performs better than a random walk.
+- A MASE score of 0.5, for example, indicates that the model’s forecasts are twice as accurate as the naïve baseline prediction, suggesting the model has predictive skill.
+
+This approach provides a rigorous evaluation of model forecasting capabilities and helps identify improvements for model calibration.
+
+```{r}
+hci = SSplotHCxval(retroSummary1.1, 
                    add = T, 
@@
                    legendcex = 0.7)
-```
-
-```{r eval=FALSE}
-hci = SSplotHCxval(retroSummary01, 
+hci = SSplotHCxval(retroSummary1.2, 
                    add = T, 
@@
                    legendcex = 0.7)
-hci = SSplotHCxval(retroSummary1, 
+hci = SSplotHCxval(retroSummary1.3, 
                    add = T, 
@@
                    legendcex = 0.7)
-hci = SSplotHCxval(retroSummary2, 
+hci = SSplotHCxval(retroSummary1.4, 
                    add = T, 
@@
                    legendcex = 0.7) 
-hci = SSplotHCxval(retroSummary3, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7) 
-hci = SSplotHCxval(retroSummary4, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-hci = SSplotHCxval(retroSummary5, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-hci = SSplotHCxval(retroSummary6, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-hci = SSplotHCxval(retroSummary7, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-```
-
-## Kobe (status)
-
-```{r}
-mvln = SSdeltaMVLN(base.modeltest,
-                   Fref = "Btgt", 
+```
+
+### Kobe (status)
+
+```{r eval=FALSE}
+mvln = SSdeltaMVLN(base.model1.1,
+                   Fref = "MSY", 
                    plot = TRUE,
@@
                    mc=100)
-
-```
-
-```{r eval=FALSE}
-mvln = SSdeltaMVLN(base.model01,
-                   Fref = "Btgt", 
+mvln = SSdeltaMVLN(base.model1.2,
+                   Fref = "MSY", 
                    plot = TRUE,
@@
                    mc=100)
-
-mvln = SSdeltaMVLN(base.model1,
-                   Fref = "Btgt", 
+mvln = SSdeltaMVLN(base.model1.3,
+                   Fref = "MSY", 
                    plot = TRUE,
@@
                    mc=100)
-
-mvln = SSdeltaMVLN(base.model2,
-                   Fref = "Btgt", 
+mvln = SSdeltaMVLN(base.model1.4,
+                   Fref = "MSY", 
                    plot = TRUE,
@@
 
-mvln = SSdeltaMVLN(base.model3,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-mvln = SSdeltaMVLN(base.model4,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-mvln = SSdeltaMVLN(base.model5,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-mvln = SSdeltaMVLN(base.model6,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-mvln = SSdeltaMVLN(base.model1.1,
-                   Fref = "Btgt", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-```
-
+```
 another
@@
 
-nanios<-base.model1$Kobe$Yr
-SSBDep<-base.model1$Kobe$B.Bmsy
-y_BDPR<-base.model1$Kobe$F.Fmsy
-
-# Integrated Model
-
-kobe1 <- ggplot(base.model1$Kobe)+
-  #geom_line(aes(kobebro$BBMSY, kobebro$FFMSY)) +
+nanios<-base.model1.1$Kobe$Yr
+SSBDep<-base.model1.1$Kobe$B.Bmsy
+y_BDPR<-base.model1.1$Kobe$F.Fmsy
+
+kobe1 <- ggplot(base.model1.1$Kobe)+
   geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0, ymax = 5),
@@
             fill = "#808080", alpha = 0.5) +
-  geom_path(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr))+
-  geom_point(aes(base.model1$Kobe$B.Bmsy, base.model1$Kobe$F.Fmsy),
+  geom_path(aes(x=base.model1.1$Kobe$B.Bmsy,y=base.model1.1$Kobe$F.Fmsy,
+                label=base.model1.1$Kobe$Yr))+
+  geom_point(aes(base.model1.1$Kobe$B.Bmsy, base.model1.1$Kobe$F.Fmsy),
              lwd=2) +
@@
   #            vjust = -0.5)+
-  geom_text(aes(x=base.model1$Kobe$B.Bmsy,y=base.model1$Kobe$F.Fmsy,label=base.model1$Kobe$Yr),
+  geom_text(aes(x=base.model1.1$Kobe$B.Bmsy,y=base.model1.1$Kobe$F.Fmsy,label=base.model1.1$Kobe$Yr),
             nudge_y = 0.1,size = 3,
@@
 
-```{r echo=FALSE}
+```{r}
 # model01
-tablebias01 <- SShcbias(retroSummary01,quant="SSB",verbose=F)
-tablebias01a <- SShcbias(retroSummary01,quant="F",verbose=F)
+tablebias01 <- SShcbias(retroSummary1.1,quant="SSB",verbose=F)
+tablebias01a <- SShcbias(retroSummary1.1,quant="F",verbose=F)
 
@@
 
-# model1
-tablebias1 <- SShcbias(retroSummary1,quant="SSB",verbose=F)
-tablebias1a <- SShcbias(retroSummary1,quant="F",verbose=F)
-
-kbl(tablebias1, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s1")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias1a, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s1")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
 # model2
-tablebias2 <- SShcbias(retroSummary2,quant="SSB",verbose=F)
-tablebias2a <- SShcbias(retroSummary2,quant="F",verbose=F)
+tablebias2 <- SShcbias(retroSummary1.2,quant="SSB",verbose=F)
+tablebias2a <- SShcbias(retroSummary1.2,quant="F",verbose=F)
 
@@
 # model3
-tablebias3 <- SShcbias(retroSummary3,quant="SSB",verbose=F)
-tablebias3a <- SShcbias(retroSummary3,quant="F",verbose=F)
+tablebias3 <- SShcbias(retroSummary1.3,quant="SSB",verbose=F)
+tablebias3a <- SShcbias(retroSummary1.3,quant="F",verbose=F)
 
@@
 # model4
-tablebias4 <- SShcbias(retroSummary4,quant="SSB",verbose=F)
-tablebias4a <- SShcbias(retroSummary4,quant="F",verbose=F)
+tablebias4 <- SShcbias(retroSummary1.4,quant="SSB",verbose=F)
+tablebias4a <- SShcbias(retroSummary1.4,quant="F",verbose=F)
 
@@
 
-# model5
-tablebias5 <- SShcbias(retroSummary5,quant="SSB",verbose=F)
-tablebias5a <- SShcbias(retroSummary5,quant="F",verbose=F)
-
-kbl(tablebias5, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s5")  %>% 
+# Combine the results from all models into one table
+tablebias_combined <- bind_rows(
+  data.frame(Model = "s1.1", 
+             Quant = "SSB", 
+             Rho = SShcbias(retroSummary1.1, 
+                            quant = "SSB", verbose = F)),
+  data.frame(Model = "s1.1", 
+             Quant = "F", 
+             Rho = SShcbias(retroSummary1.1, 
+                            quant = "F", verbose = F)),
+  data.frame(Model = "s1.2", 
+             Quant = "SSB", 
+             Rho = SShcbias(retroSummary1.2, 
+                            quant = "SSB", verbose = F)),
+  data.frame(Model = "s1.2", 
+             Quant = "F", 
+             Rho = SShcbias(retroSummary1.2, 
+                            quant = "F", verbose = F)),
+  data.frame(Model = "s1.3", 
+             Quant = "SSB", 
+             Rho = SShcbias(retroSummary1.3, 
+                            quant = "SSB", verbose = F)),
+  data.frame(Model = "s1.3", 
+             Quant = "F", 
+             Rho = SShcbias(retroSummary1.3, 
+                            quant = "F", verbose = F)),
+  data.frame(Model = "s1.4", 
+             Quant = "SSB", 
+             Rho = SShcbias(retroSummary1.4, 
+                            quant = "SSB", verbose = F)),
+  data.frame(Model = "s1.4",
+             Quant = "F", 
+             Rho = SShcbias(retroSummary1.4, 
+                            quant = "F", verbose = F))
+)
+
+kbl(tablebias_combined, booktabs = T, 
+    format = "html", 
+    caption = "Rho parameter by model and quantity (SSB and F)") %>% 
   kable_styling(latex_options = "HOLD_position")
 
-kbl(tablebias5a, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s5")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-# model6
-tablebias6 <- SShcbias(retroSummary6,quant="SSB",verbose=F)
-tablebias6a <- SShcbias(retroSummary6,quant="F",verbose=F)
-
-kbl(tablebias6, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s6")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias6a, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s6")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-# model7
-tablebias7 <- SShcbias(retroSummary7,quant="SSB",verbose=F)
-tablebias7a <- SShcbias(retroSummary7,quant="F",verbose=F)
-
-kbl(tablebias7, booktabs = T,format = "latex",
-    caption = "Rho parameter in SSB model s7")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-kbl(tablebias7a, booktabs = T,format = "latex",
-    caption = "Rho parameter in F model s7")  %>% 
-    kable_styling(latex_options = "HOLD_position")
-
-```
-
-## Likelihood tables
-
-```{r}
-like01 <- base.model01$likelihoods_used
-like01$model <- rep("s01", nrow(like01))
-like01 <- rownames_to_column(like01, var = "Description")
-like1 <- base.model1$likelihoods_used
-like1$model <- rep("s1", nrow(like1))
+```
+
+### Likelihood tables
+
+```{r}
+
+like1 <- base.model1.1$likelihoods_used
+like1$model <- rep("s1.1", nrow(like1))
 like1 <- rownames_to_column(like1, var = "Description")
-like2 <- base.model1$likelihoods_used
-like2$model <- rep("s2", nrow(like2))
+like2 <- base.model1.2$likelihoods_used
+like2$model <- rep("s1.2", nrow(like2))
 like2 <- rownames_to_column(like2, var = "Description")
-like3 <- base.model3$likelihoods_used
-like3$model <- rep("s3", nrow(like3))
+like3 <- base.model1.3$likelihoods_used
+like3$model <- rep("s1.3", nrow(like3))
 like3 <- rownames_to_column(like3, var = "Description")
-like4 <- base.model4$likelihoods_used
-like4$model <- rep("s4", nrow(like4))
+like4 <- base.model1.4$likelihoods_used
+like4$model <- rep("s1.4", nrow(like4))
 like4 <- rownames_to_column(like4, var = "Description")
-like5 <- base.model5$likelihoods_used
-like5$model <- rep("s5", nrow(like5))
-like5 <- rownames_to_column(like5, var = "Description")
-like6 <- base.model6$likelihoods_used
-like6$model <- rep("s6", nrow(like6))
-like6 <- rownames_to_column(like6, var = "Description")
-like7 <- base.model7$likelihoods_used
-like7$model <- rep("s7", nrow(like7))
-like7 <- rownames_to_column(like7, var = "Description")
-
-totalike <- rbind(like01,
-                  like1,
+
+totalike <- rbind(like1,
                    like2,
                    like3,
-                   like4,
-                   like5,
-                   like6,
-                   like7)
+                   like4)
 
@@
 ggplot(totalike)+
-  geom_bar( aes(y = reorder(Description, values), x =log(values)),
+  geom_bar( aes(y = reorder(Description, values), 
+                x =log(values)),
             stat = "identity", position = "dodge") +
@@
                                    hjust = 1))+
-  xlim(0,20)
-```
-
-## Likelihood Profile
+  xlim(0,30)
+```
+
+### Likelihood Profile
 
@@
 
-Funciones para generar plots de perfil de verosimilitud ----
-
-### SSplotProfile() ----
+### `SSplotProfile()`
 
@@
 
-## Convergence criteria
-
-0.0001 final convergence criteria (e.g. 1.0e-04)
+### Convergence Criteria
+
+The convergence criterion used for model calibration is set to a final threshold of **0.0001** (or equivalently **1.0e-04**).
+This criterion defines the minimum acceptable difference between successive model iterations.
+Convergence is considered achieved when the absolute change in the objective function value or key parameters falls below this threshold.
+A smaller convergence value ensures that the model achieves a high degree of accuracy and stability in its final estimates, indicating that further iterations are unlikely to result in significant changes to the parameter estimates.
 
@@
 
-```{r echo = FALSE}
+```{r eval = FALSE}
 #### R0_profile_plot_Length_like ----
@@
 
-```{r echo = FALSE}
+```{r eval=FALSE}
 #### R0_profile_plot_Survey_like ----
@@
 ```
-
-## Outputs Model Base
-
-```{r}
-outps1 <- base.modeltest$timeseries[1:53, 2:8]
+## Outputs
+
+### Outputs Model `s1.1`
+
+```{r}
+outps1 <- base.model1.1$timeseries[1:37, 2:8]
 
@@
       format = "html",
-    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
+    caption = "Main variables outputs from stock asssessment krill in WAP `s1.1`") %>%
   kable_paper("hover", 
@@
 ```
-
-## Comparision outputs betwwen scenarios
-
-```{r echo=FALSE}
+### Outputs Model `s1.2`
+
+```{r}
+outps2 <- base.model1.2$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps2 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.2`") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9,
+                html_font = "arial")#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Outputs Model `s1.3`
+
+```{r}
+outps3 <- base.model1.3$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps3 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.3`") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9,
+                html_font = "arial")#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Outputs Model `s1.4`
+
+```{r}
+outps4 <- base.model1.3$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps4 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.4`") %>%
+  kable_paper("hover", 
+              full_width = TRUE)%>%
+  kable_styling(latex_options = c("striped",
+                                  "condensed"),
+                full_width = FALSE,
+                font_size=9,
+                html_font = "arial")#%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+## Comparison
+
+### Comparison outputs betwwen scenarios
+
+```{r}
 
@@
 biglist <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(dir01,
-                                  dir1,
-                                  dir2,
-                                  dir3,
-                                  dir4,
-                                  dir5,
-                                  dir6,
-                                  dir7),
+                       dirvec = c(dir1.1,
+                                  dir1.2,
+                                  dir1.3,
+                                  dir1.4),
                        getcovar = F)
@@
 summaryoutputall <- SSsummarize(biglist)
-```
-
-```{r echo=FALSE}
-compall <- SSplotComparisons(summaryoutputall,
-                  legendlabels = c("No Spatial",
-                                   "s1",
-                                   "s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
-                 labels = c("Year", 
-                            "Spawning biomass (t)",
-                            "Relative spawning biomass", 
-                            "Age-0 recruits (1,000s)",
-                            "Recruitment deviations", 
-                            "Index", "Log index", 
-                            "1 - SPR", 
-                            "Density",
-                            "Management target", 
-                            "Minimum stock size threshold",
-                            "Spawning output",
-                            "Harvest rate"),
-                 png=TRUE,plotdir=Figs)
-
 ```
@@
                    digits = NULL,
-                   modelnames = c("s01", "s1","s2",
-                 "s3",
-                 "s4",
-                 "s5",
-                 "s6",
-                 "s7"),
+                   modelnames = c("s1.1", "s1.2",
+                                  "s1.3",
+                                  "s1.4"),
                    csv = F,
@@
                    mcmc = FALSE)
+
 kbl(comtable, booktabs = T,format = "latex",
-    caption = "Comparacion likelihood y parámetros s01, s1, s2, s3, s4, s5, s6 y s7")  %>% 
+    caption = "Comparision s1.1, s1.2, s1.3, s1.4")  %>% 
     kable_styling(latex_options = "scale_down")
@@
 
-## Comparision between select models `No spatial`and `Spatial implicit` and `Spatial W/ new set parametres`
-
-```{r echo=FALSE}
+```{r eval=FALSE}
 #read in all model runs
@@
 
-comparision between select models `Old Paramters` and `New Parameters WG SAM 2024/23`
-
-```{r echo=FALSE}
+Comparison between select models `Ref Model: No Env-Predator` and `S1.1 w/ Env and Predator data`
+
+```{r eval=FALSE}
 #read in all model runs
 #note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist1_7 <- SSgetoutput(keyvec = NULL,
+biglist14 <- SSgetoutput(keyvec = NULL,
                        dirvec = c(
-                                  dir1,
-                                  dir7),
+                                  dir1.1,
+                                  dir1.4),
                        getcovar = F)
@@
 #create summary of model runs from list above
-summaryoutput1_7 <- SSsummarize(biglist1_7)
-
-comp1_7 <- SSplotComparisons(summaryoutput1_7,
-                  legendlabels = c(
-                                   "Old Parameters",
-
-                 "Parameters WG SAM 2024/27"),
+summaryoutput14 <- SSsummarize(biglist14)
+
+comp14 <- SSplotComparisons(summaryoutput14,
+                  legendlabels = c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Env and Predator data"),
+                  filenameprefix = "COM2",
                  labels = c("Year", 
@@
 ```
-### Comparision between select models `s1` and `s9`
-
-```{r echo=FALSE}
-#read in all model runs
-#note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist1_9 <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(
-                                  dir1,
-                                  dir9),
-                       getcovar = F)
-
-#create summary of model runs from list above
-summaryoutput1_9 <- SSsummarize(biglist1_9)
-
-comp1_9 <- SSplotComparisons(summaryoutput1_9,
-                  legendlabels = c(
-                                   "Base",
-                 "selectivity Block 2006"),
-                 filenameprefix = "COM1_9",
-                 labels = c("Year", 
-                            "Spawning biomass (t)",
-                            "Relative spawning biomass", 
-                            "Age-0 recruits (1,000s)",
-                            "Recruitment deviations", 
-                            "Index", "Log index", 
-                            "1 - SPR", 
-                            "Density",
-                            "Management target", 
-                            "Minimum stock size threshold",
-                            "Spawning output",
-                            "Harvest rate"),
-                 png=TRUE,plotdir=Figs)
-```
-## Comparsion in sd long term time series
+
+### Comparsion in sd long term time series
 
@@
 
-sblt01 <- base.model01$timeseries$Bio_all[1:50]
-sblt1 <-  base.model1$timeseries$Bio_all
-sblt2 <-  base.model2$timeseries$Bio_all
-sblt3 <-  base.model3$timeseries$Bio_all
-sblt4 <-  base.model4$timeseries$Bio_all
-sblt5 <-  base.model5$timeseries$Bio_all
-sblt6 <-  base.model6$timeseries$Bio_all
-sblt7 <-  base.model7$timeseries$Bio_all
-
-errt <- as.data.frame(cbind(sblt01,
-              sblt1,
+sblt1 <-  base.model1.1$timeseries$Bio_all
+sblt2 <-  base.model1.2$timeseries$Bio_all
+sblt3 <-  base.model1.3$timeseries$Bio_all
+sblt4 <-  base.model1.4$timeseries$Bio_all
+
+errt <- as.data.frame(cbind(sblt1,
               sblt2,
               sblt3,
-              sblt4,
-              sblt5,
-              sblt6,
-              sblt7))
+              sblt4))
 err_long <- errt %>%
@@
                values_to = "value")  
-titles <- c("No Spatial",
-            "Spatial implicit", 
-            "Spatial w/ without S-R relation",
-            "Spatial w/ S-R relation weak (0.9 steepness)",
-            "Spatial w/ S-R relation strong (0.6 steepness)",
-            "Spatial w/ S-R relation mid-strong estimated",
-            "Spatial w/ Ricker S-R relation estimated", 
-            "Spatial w/ New set parameters Linf, k and M")
-
+titles <- c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Predator data",
+                 "S1.1 w/ Env data",
+                 "S1.1 w/ Env and Predator data")
 ggplot(err_long, aes(x = variable, y = value)) +
-  geom_boxplot() +
+  geom_boxplot(width=0.5) +
   geom_jitter(width = 0.2, alpha = 0.3) +
@@
 
-## Autocorrelation in Recruit and Biomas
-
-```{r fig.width=8, fig.height=7}
-recs01 <- base.model01$recruit$pred_recr
-recs1 <- base.model1$recruit$pred_recr
-recs2 <- base.model2$recruit$pred_recr
-recs3 <- base.model3$recruit$pred_recr
-recs4 <- base.model4$recruit$pred_recr
-recs5 <- base.model5$recruit$pred_recr
-recs6 <- base.model6$recruit$pred_recr
-recs7 <- base.model7$recruit$pred_recr
-
-p01 <- ggAcf(base.model01$recruit$pred_recr) +
-  ggtitle("No Spatial") +
+### Autocorrelation in Recruit
+
+To evaluate the temporal correlation structure of krill recruitment under different model configurations, we performed an Autocorrelation Function (ACF) analysis.
+The ACF measures the correlation between observations at different time lags, helping to assess whether recruitment estimates exhibit persistence or randomness across time.
+
+The analysis was conducted on recruitment estimates derived from four model scenarios:
+
+A reference model without environmental or predator influences (Ref Model: No Env-Predator).
+A model incorporating predator data (S1.1 w/ Predator data).
+A model incorporating environmental data (S1.1 w/ Env data).
+A model incorporating both environmental and predator data (S1.1 w/ Env and Predator data).
+Each model's residuals were extracted, and the autocorrelation function (ACF) was computed for a time lag range of up to 15 years.
+The dashed blue lines in the plots represent the 95% confidence intervals, indicating the threshold beyond which correlation values are statistically significant.
+If autocorrelation values remain within this range, it suggests that the recruitment estimates behave as a random process with no significant dependence on past values.
+Conversely, autocorrelation values exceeding these bounds indicate recruitment persistence or cyclic patterns.
+
+```{r fig.width=5, fig.height=7}
+recs1 <- base.model1.1$recruit$pred_recr
+recs2 <- base.model1.2$recruit$pred_recr
+recs3 <- base.model1.3$recruit$pred_recr
+recs4 <- base.model1.4$recruit$pred_recr
+
+p1 <- ggAcf(base.model1.1$recruit$pred_recr) +
+  ggtitle("Ref Model: No Env-Predator") +
   theme_few()
-p1 <- ggAcf(base.model1$recruit$pred_recr) +
-  ggtitle("Spatial implicit") +
+p2 <- ggAcf(base.model1.2$recruit$pred_recr) +
+  ggtitle("S1.1 w/ Predator data") +
   theme_few()
-p2 <- ggAcf(base.model2$recruit$pred_recr) +
-  ggtitle("Spatial w/ without S-R relation") +
+p3 <- ggAcf(base.model1.3$recruit$pred_recr) +
+  ggtitle("S1.1 w/ Env data") +
   theme_few()
-p3 <- ggAcf(base.model3$recruit$pred_recr) +
-  ggtitle("Spatial w/ S-R relation weak (0.9 steepness)") +
+p4 <- ggAcf(base.model1.4$recruit$pred_recr) +
+  ggtitle("S1.1 w/ Env and Predator data") +
   theme_few()
-p4 <- ggAcf(base.model4$recruit$pred_recr) +
-  ggtitle("Spatial w/ S-R relation strong (0.6 steepness)") +
-  theme_few()
-p5 <- ggAcf(base.model5$recruit$pred_recr) +
-  ggtitle("Spatial w/ S-R relation mid-strong estimated") +
-  theme_few()
-p6 <- ggAcf(base.model6$recruit$pred_recr) +
-  ggtitle("Spatial w/ Ricker S-R relation estimated") +
-  theme_few()
-p7 <- ggAcf(base.model7$recruit$pred_recr) +
-  ggtitle("Spatial w/ New set parameters Linf, k and M") +
-  theme_few()
-
-ggarrange(p01,p1,
+
+ggarrange(p1,
           p2,
           p3,
-          p4,
-          p5,
-          p6,
-          p7, ncol=3, nrow = 3,
+          p4, ncol=1, nrow = 4,
           common.legend = T)
@@
 
-## Recruit deviation
-
-```{r}
-dev01 <- base.model01$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S01")
-dev1 <- base.model1$recruit[1:43,c(1,7)]%>% 
-  mutate(Serie = "S1")
-dev2 <- base.model2$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S2")
-dev3 <- base.model3$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S3")
-dev4 <- base.model4$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S4")
-dev5 <- base.model5$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S5")
-dev6 <- base.model6$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S6")
-dev7 <- base.model7$recruit[1:43,c(1,7)] %>% 
-  mutate(Serie = "S7")
-
-data_list <- list(dev01, dev1, dev2, dev3, dev4, dev5, dev6, dev7)
-titles <- c("No Spatial",
-            "Spatial implicit", 
-            "Spatial w/ without S-R relation",
-            "Spatial w/ S-R relation weak (0.9 steepness)",
-            "Spatial w/ S-R relation strong (0.6 steepness)",
-            "Spatial w/ S-R relation mid-strong estimated",
-            "Spatial w/ Ricker S-R relation estimated", 
-            "Spatial w/ New set parameters Linf, k and M")
+The ACF plots indicate that the reference model (without environmental or predator data) exhibits weak but noticeable positive autocorrelation at certain lags, suggesting some degree of recruitment dependence over time.
+However, this autocorrelation does not appear strong or systematic.
+
+The model incorporating only predator data shows a slight reduction in autocorrelation magnitude, suggesting that predator-driven recruitment variability may have captured part of the temporal structure in the data.
+
+The model with only environmental data exhibits a further reduction in autocorrelation, implying that environmental variability explains a larger portion of recruitment trends than predator effects alone.
+
+Finally, the model that includes both environmental and predator data presents the lowest autocorrelation values, with nearly all bars remaining within the confidence bounds.
+This suggests that incorporating both factors provides the most effective explanation of recruitment fluctuations, reducing unexplained temporal structure.
+
+Overall, these results indicate that recruitment variability is at least partially driven by environmental and predator influences, and models integrating these factors provide more robust and independent recruitment estimates, minimizing systematic dependencies over time.
+
+### Recruit deviation
+
+```{r fig.width=5, fig.height=7}
+dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
+  mutate(Serie = "s1.1")
+dev2 <- base.model1.2$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.2")
+dev3 <- base.model1.3$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.3")
+dev4 <- base.model1.4$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.4")
+
+data_list <- list(dev1, 
+                  dev2,
+                  dev3,
+                  dev4)
+titles <- c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Predator data",
+                 "S1.1 w/ Env data",
+                 "S1.1 w/ Env and Predator data")
 # Lista para almacenar los gráficos
@@
                                    hjust = 1))+
+    ylim(-2,2)+
     theme_few()
@@
 # Organizar los gráficos con ggarrange
-combined_plot <- ggarrange(plotlist = plots, ncol = 3, nrow = 3)
+combined_plot <- ggarrange(plotlist = plots, ncol = 1, nrow = 4)
 
@@
 
-## Comparision RMSE
-
-```{r}
-# Crear un data frame con los valores de RMSE
-df <- data.frame(
-  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
-  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
-  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10])
-)
-
-# Comparación de estadísticas descriptivas
-summary(df)
-
-# Comparación con boxplot
-boxplot(df, main = "Comparación de RMSE entre modelos", col = c("red", "blue", "green"))
-
-# Prueba de diferencias estadísticas (ANOVA)
-anova_result <- aov(as.numeric(unlist(df)) ~ rep(1:3, each=nrow(df)))
-summary(anova_result)
-
-t.test(df$Model1.1, df$Model1.2)  # Comparar Model1.1 vs Model1.2
-t.test(df$Model1.1, df$Model1.3)  # Comparar Model1.1 vs Model1.3
-t.test(df$Model1.2, df$Model1.3)  # Comparar Model1.2 vs Model1.3
-
-# ANOVA para ver si hay diferencias globales
-anova_rmse <- aov(as.numeric(unlist(df)) ~ rep(1:3, each=nrow(df)))
-summary(anova_rmse)
-
-# Prueba post-hoc si ANOVA es significativa
-TukeyHSD(anova_rmse)
-
-df_long <- reshape2::melt(df)
-ggplot(df_long, aes(x = variable, y = value, fill = variable)) +
-  geom_boxplot() +
-  theme_minimal() +
-  labs(title = "Comparación de RMSE entre modelos", x = "Modelo", y = "RMSE")
-
-```
-
-```{r}
-
-# Unir los modelos en un solo dataframe y asegurarse de que los nombres de filas se mantienen
+```{r}
+
 df1 <- base.model1.1$likelihoods_used %>%
   select(values) %>%
-  mutate(Componente = rownames(base.model1.1$likelihoods_used), Modelo = "M1.1")
+  mutate(Componente = rownames(base.model1.1$likelihoods_used), Modelo = "s1.1")
 
@@
   select(values) %>%
-  mutate(Componente = rownames(base.model1.2$likelihoods_used), Modelo = "M1.2")
+  mutate(Componente = rownames(base.model1.2$likelihoods_used), Modelo = "s1.2")
 
@@
   select(values) %>%
-  mutate(Componente = rownames(base.model1.3$likelihoods_used), Modelo = "M1.3")
-
-# Combinar todos los modelos
-df <- bind_rows(df1, df2, df3)
-
-# Crear gráfico de barras con valores transformados
+  mutate(Componente = rownames(base.model1.3$likelihoods_used), Modelo = "s1.3")
+
+df4 <- base.model1.4$likelihoods_used %>%
+  select(values) %>%
+  mutate(Componente = rownames(base.model1.4$likelihoods_used), Modelo = "s1.4")
+
+df <- bind_rows(df1, 
+                df2,
+                df3,
+                df4)
+
 ggplot(df, aes(x = Componente, y = values, fill = Modelo)) +
-  geom_col(position = position_dodge(), width = 0.7) +  # Barras separadas por modelo
-  theme_bw() +
-  coord_flip() +  # Para que sea horizontal
-  scale_fill_manual(values = c("S1.1" = "grey",
-                               "S1.2" = "red", 
-                               "S1.3" = "black")) +  # Colores personalizados
+  geom_col(position = position_dodge(), width = 0.7) +  
+  theme_few() +
+  coord_flip() +  
+  scale_fill_manual(name= "Scenario",
+                    values = c("s1.1" = '#ca0020',
+                               "s1.2" = '#f4a582', 
+                               "s1.3" = '#bababa',
+                               "s1.4" = '#404040')) + 
   labs(x = "",
@@
        title = "") 
-
-```
-
-## Platoons analisis
+```
+
+### Platoons analisis
 
@@
 ```
-## AKL
+### AKL
 
@@
 ```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
-alk_matrix <- base.model1.2$ALK[,,1] # Ajusta las dimensiones según tu matriz
+alk_matrix <- base.model1.1$ALK[,,1] # Ajusta las dimensiones según tu matriz
 

### 2025-04-01 index.Rmd
ggarrange plots

@@
 ```{r}
+
+par(mfrow=c(1,3), mar=c(5,4,1,1)) 
 SSplotTimeseries(base.model1.1,
@@
 ```{r}
+par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.2,
@@
 ```{r}
+par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.3,
@@
 ```{r}
+par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.4,
@@
 ```{r}
+par(mfrow=c(2,2), mar=c(5,4,1,1))
 SSplotSelex(base.model1.1,
             subplots = 1)
+mtext("s1.1", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.2,
             subplots = 1)
+mtext("s1.2", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.3,
             subplots = 1)
+mtext("s1.3", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.4,
             subplots = 1)
-```
-
-```{r}
-SSplotIndices(base.model1.1,
-            subplots = 9)
-SSplotIndices(base.model1.2,
-            subplots = 9)
-SSplotIndices(base.model1.3,
-            subplots = 9)
-SSplotIndices(base.model1.4,
-            subplots = 9)
+mtext("s1.4", side=3, line=0.5, cex=1.2)
+```
+
+```{r}
+par(mfrow=c(2,2), mar=c(5,4,2,1))
+
+SSplotIndices(base.model1.1, subplots = 9)
+mtext("s1.1", side=3, line=0.5, cex=1.2)
+
+SSplotIndices(base.model1.2, subplots = 9)
+mtext("s1.2", side=3, line=0.5, cex=1.2)
+
+SSplotIndices(base.model1.3, subplots = 9)
+mtext("s1.3", side=3, line=0.5, cex=1.2)
+
+SSplotIndices(base.model1.4, subplots = 9)
+mtext("s1.4", side=3, line=0.5, cex=1.2)
+
 ```
@@
 ```{r}
+par(mfrow=c(2,4), mar=c(5,4,2,1)) 
 SSplotRunstest(base.model1.1,
@@
                plotdir = Figs)
+```
+
+```{r}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.2,
@@
                plotdir = Figs)
+```
+
+```{r}
+par(mfrow=c(2,4), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.3,
@@
                plotdir = Figs)
+```
+
+```{r}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.4,
@@
 
-```{r}
+```{r eval=FALSE}
 SSplotRunstest(base.model1.1,
@@
 ```{r}
+par(mfrow=c(2,2), mar=c(5,4,2,1))
 SSplotJABBAres(base.model1.1,
@@
 
-# Comparación con boxplot
-boxplot(dfrmse, main = "RMSE comparision", col = c('#ca0020','#f4a582','#bababa','#404040'))
-
-# Prueba de diferencias estadísticas (ANOVA)
-anova_result <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
-summary(anova_result)
-
 t.test(dfrmse$Model1.1, dfrmse$Model1.2)  # Comparar Model1.1 vs Model1.2
@@
 dfrmse_long <- reshape2::melt(dfrmse)
-
 ggplot(dfrmse_long, aes(x = variable, y = value, fill = variable)) +
@@
   labs(title = "", x = "", y = "RMSE")
-
 ```
@@
 ```{r}
-#save(retroSummary, retroModels, file="retro5.Rdata")
+
+par(mfrow=c(2,2), mar=c(5,4,1,1))
+
 retro1.1 <- SSplotRetro(retroSummary1.1,
@@
 ```
+```{r}
+
+par(mfrow=c(2,2), mar=c(5,4,1,1))
+
+retro1.1 <- SSplotRetro(retroSummary1.1,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("F"))
+retro1.2 <- SSplotRetro(retroSummary1.2,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("F"))
+retro1.3 <- SSplotRetro(retroSummary1.3,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("F"))
+retro1.4 <- SSplotRetro(retroSummary1.4,
+            add=T,
+            forecast = F,
+            legend = T,
+            verbose=F,
+            subplots = c("F"))
+```
 
@@
 ```{r}
+par(mfrow=c(2,4), mar=c(5,4,2,1))
 hci = SSplotHCxval(retroSummary1.1, 
@@
                    legendcex = 0.7)
+```
+
+```{r}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
 hci = SSplotHCxval(retroSummary1.2, 
@@
                    legendcex = 0.7)
+```
+
+```{r}
+par(mfrow=c(2,4), mar=c(5,4,2,1))
 hci = SSplotHCxval(retroSummary1.3, 
@@
                    legendcex = 0.7)
+```
+
+```{r}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
 hci = SSplotHCxval(retroSummary1.4, 
@@
 )
-
+```
+
+```{r}
 kbl(tablebias_combined, booktabs = T, 
@@
 
-```{r fig.width=5, fig.height=7}
+```{r}
 recs1 <- base.model1.1$recruit$pred_recr
@@
           p3,
-          p4, ncol=1, nrow = 4,
+          p4, ncol=2, nrow = 2,
           common.legend = T)
@@
 
-```{r fig.width=5, fig.height=7}
+```{r, warning=FALSE}
 dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
@@
 # Organizar los gráficos con ggarrange
-combined_plot <- ggarrange(plotlist = plots, ncol = 1, nrow = 4)
+combined_plot <- ggarrange(plotlist = plots, ncol = 2, nrow = 2)
 

### 2025-04-03 MS2_Krill.Rmd
adding text

@@
 ---
-title: "Assessing environmental and predation effects on Antarctic Krill (*Euphausia superba*) stock productivity: An Integrated Catch-at-Length Model Perspective"
+title: "Assessing environmental and predation impacts on Antarctic Krill (*Euphausia superba*) dynamic population from an integrated Catch-at-Length model perspective"
 author:
@@
 ---
-\newpage
-\tableofcontents
-\newpage
 
@@
 library(openxlsx)
+library(r4ss)
+library(readr)
+library(flextable)
+library(officer)
 library(r4ss)
@@
 Our results indicate that krill population variables, as well as productivity estimates, tend to be more optimistic in terms of stock status when environmental and predators variables are not considered.
+Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics.
 This highlights a critical challenge in developing management recommendations and harvest control rules, as ignoring ecosystem influences may lead to overly optimistic assessments of krill stock health.
-Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics.
 We consider that the use of integrated models, which leverage the extensive population, fisheries, and ecosystem data available through various scientific programs, can provide valuable insights for management aligned with CCAMLR conservation and harvesting objectives.
 
-*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation components, fisheries management.*
+*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, SS3, Catch-to-Length model, environmental factors, predation components, fisheries management.*
 
@@
 
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture spatial variability, and account for ecosystem interactions.
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
 However, integrating these complexities presents several challenges.
@@
 
-To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past four decades.
+To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past decades.
 This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations.
-Specifically, we employ an integrated catch-at-length stock synthesis model that allows us to evaluate the effects of environmental drivers on krill recruitment and productivity, as well as the impact of predator-prey interactions on overall population dynamics.
-Our approach involves comparing different scenarios, one incorporating ecosystem-related variables and another based solely on traditional stock assessment data, to determine how environmental fluctuations and predator abundance influence key population variables and parameters.
+Specifically, we employ an integrated catch-at-length implemented in Stock Synthesis model [@Methot2013] that allows us to evaluate the effects of environmental drivers on recruitment and productivity, as well as the impact of predator components into krill population dynamics.
+The primary objective of this study was to assess the statistical performance of different management metrics, population variables and parameters based on alternative approaches to incorporating both environmental and predator data into a typical krill stock assessment model.
 By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends.
@@
 
-By integrating ecosystem drivers into the stock assessment process, we enhance the accuracy of population estimates and improve our ability to predict future stock trends under changing environmental conditions.
-This broader ecological perspective strengthens the scientific foundation for sustainable management recommendations, ensuring that harvesting strategies align with an ecosystem-based management approach.
-Ultimately, the insights gained from this integrated modeling effort contribute to the ongoing efforts of CCAMLR's new strategy to balance krill fishery sustainability with the conservation of the broader Antarctic marine ecosystem.
+By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate conditions.
+This broader ecological perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
+Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
 
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
+The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021; REF] (Figure \@ref(fig:mapa)).
 The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
 Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
+
+## Conceptual Model
+
+In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets [REF].
+Differences among strata have been identified from various perspectives of krill population dynamics [REF], as well as from the influence of environmental variability within Subarea 48.1 [REF].
+This approach allows integrate spatial heterogeneity into the stock assessment model to ensure that spatially structured processes are adequately represented.
+Although strata are treated as separate fleets, recruitment and growth processes remain unified, ensuring that the overall krill stock is modeled as a single population while accounting for spatially structured impacts at the Subarea 48.1 level.
+Each stratum is assigned a unique catch and survey history, reflecting differences in harvesting intensity and spatial exploitation patterns.
+Selectivity is modeled separately for each strata to account for potential variations in size composition due to fishing gear, depth distribution, or environmental conditions.
+Survey indices and length composition data are fleet-specific, allowing for variations in data collection methods and regional abundance estimates.
+The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year ocurring in January [@REF].
+This method allows for spatial variability in exploitation patterns without explicitly modeling movement between areas.
+This approach enables a more realistic assessment of krill stock productivity under varying environmental conditions and harvesting scenarios, aligning with CCAMLR’s ecosystem-based management framework.
 
@@
 Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling.
-All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
+These data sources include data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+This database has information about vessel, nationality, georeferenced, among others.
+Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including biomass index by strata, and biological characteristics of krill collected through net sampling.
+All this data was processed in @Dornam2021.
+Krill length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
 Mat.
-2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
-Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
+2).
+For this analysis, Chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
+Chl data was downloaded from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
 Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
 This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and krill dynamics in the WAP.
-We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)).
-The model included spatial component (cell) as random effect (formula: ~1 | cell).
-Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl.
-Mat 3); 4) Predator abundance data from "MAPPPD".
-Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
+We applied a Generalized Linear Mixed Model (GLMM) to identify the environmental variable with main influence on krill population dynamics in the Antarctic Peninsula.
+Krill *Length* was used as a proxy for growth, incorporating both fixed and random effects to account for environmental and spatial variability.
+The icon model was formulated as:
 
-## Conceptual Model
+$$
+\text{Length}_i = \beta_0 + \beta_1 \text{ID}_i + \beta_2 \text{SIC}_i + \beta_3 \text{SST}_i + \beta_4 \text{Chla}_i + \beta_5 (\text{SST} \times \text{Chla})_i + (1 | \text{Year}_i) + \epsilon_i
+$$
 
-A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
-Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
-The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.
+where $\beta_0$ is the intercept, and the fixed effects include *ID* (management unit or stratum), sea ice concentration (SIC), sea surface temperature (SST), and chlorophyll-a concentration (Chl).
+An interaction term between SST and Chl was included to capture potential synergistic effects on krill growth.
+A random effect for *Year* was added to account for interannual variability, mitigating biases from year-specific influences and data scarcity across strata.
+The residual error, $\epsilon_i$, was assumed to follow a normal distribution.
+The estimated coefficients and their 95% confidence intervals (CIs) were evaluated to determine the significance of each predictor, where coefficients overlapping zero indicated no significant relationship with krill length.
+Variance components were used to assess the proportion of variability explained by the random factors.
+Model selection was based on performance, with six different models tested, including one using the 75th percentile of lengths as an alternative response variable.
+The GLMM analysis was conducted using the *lme4* and *easystats* packages [@REF], while spatial data processing and stratification were performed with the *sf* and *CCAMLRGIS* packages [@REF].
+Data handling for modeling purposes was carried out using the *tidyverse* library and its dependencies [@REF].
+A full description of the models used is provided in Supplementary Material 2.
+
+```{r eval=TRUE}
+data <- data.frame(
+  Source = c("Krill Fishery Monitoring Program",
+                 "Krill Fishery Monitoring Program, collected by scientific observers under CCAMLR SISO program (vessel, nationality, georeferenced)",
+                 "Krill Fishery Monitoring Program",
+                 "AMLR Research Cruises Program", 
+                  "Satellite data",
+                 "Predator Abundance",
+                 "AMLR Research Cruises Program"),
+  Description = c("Catches",
+                  "CPUE (standarized by vessel, area and deep)", 
+                  "Krill Length compositions (grouped by 0.2 cm)",
+                  "Biomass index from scientific AMLR cruises",
+                  "Satellite-based chlorophyll-a concentration (monthly, interpolated) from the EU Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).",
+                  "Data on penguin colonies, sourced from 'MAPPPD'",
+                  "Predator Length composition krill diet"),
+    TimeSeries = c("1997-2020",
+                 "1997-2020",
+                 "2000-2020",
+                 "1991-2011 (based on the research cruise periods).",
+                 "Monthly data from 1997 to present (based on satellite data availability).",
+                 "Data from 2000s-present.",
+                 "1977-2020 (from 2000 used in this analysis)"),
+   Scale = c("by Strata",
+                           "by Strata",
+                           "by Strata",
+                           "by Strata",
+                           "48.1",
+                           "by colony",
+                           "by colony"))
+ft <- flextable(data) %>%  
+  set_caption("Comparison of Model Parameters") %>%  
+  autofit() %>%  
+  theme_vanilla() %>%  
+  fontsize(size = 10, 
+           part = "all") %>%  
+  font(fontname = "Arial", 
+       part = "all") %>%  
+  set_table_properties(layout = "autofit", width = .75)
+ft
+```
 
@@
 Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
+Therefore, a Catch-to-Length model with age dynamics was used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
 Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
+Regarding this, we consider a length-based model as an alternative to the age-based approach, especially given the scarcity (or null) of direct age data in krill monitoring.
+Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017].
+In this sense, reading hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap.
+Length offers insights into the age structure of the population, as there is a correlation between age and length.
+So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021].
+This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK).
+To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].
+ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples.
+ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
+Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias.
+Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003].
+Formulae exist to estimate the number of age determinations and length measurements is calculated as;
+
+$$
+\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
+$$
+
+The ALK for the time period is raised to the length distribution for that time period:
+
+$$
+\text{Raised numbers at age by length group} = \text{Numbers at length} \times \text{Proportion at age for that length}
+$$
+
+If the ALK lacks data for certain length groups, data from those groups may be assigned to adjacent ones in the ALK.
+This process generates an age-length distribution by gear and time period, but caution is needed to avoid significant biases, especially for larger lengths with a broad age range.
+The numbers at age are summed across length groups, combining variances from ageing and length sampling to produce the age composition for the specified period.
+Numbers at age for all gears can then be calculated.
+Numbers at age for all gears can be calculated as:
+
+$$
+\sum N_a \times \left(\frac{W_{ct}}{W_{cs}}\right)
+$$
+
+where $\sum N_a$ is the sum of sampled numbers at age, W~ct~ is the total commercial catch weight, and W~cs~ is the
+sampled commercial catch weight. Variance due to ageing of numbers at age for all gears can be calculated as:
+
+$$
+\sum \text{Var}_a
+$$
+where $\sum Var_a$  is the sum of variances due to ageing. Variance due to length sampling of numbers at age can be calculated as:
+
+$$
+\sum \text{Var}_l
+$$
+where $\sum Var_a$  is the sum of variances due to length sampling. The variances should be raised by:
+
+$$
+\frac{W_{ct}}{W_{cs}}
+$$
+where W~ct~ is the total commercial catch weight, and W~cs~ is the sampled commercial catch weight.
+
+The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
+The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
+In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
+In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
+This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass.
+Weight-at-length was assumed known and calculated using @Maschette2020 parametres.
+in krill, a AKL is exposed in Figure \@ref(fig:AKL)).
+
+```{r AKL, eval=TRUE, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in SUbarea 48.1"}
+knitr::include_graphics('Figs/AKL.png')
+```
+
 The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
@@
 All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
-Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
+Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
 Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
 
-We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends.
-The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables.
-Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation.
-Key model parameters include:
+\newpage
 
-### Natural Mortality and Growth Parameterization
+```{r parainit1, eval=TRUE}
+dir1.1<-here("s1.1") 
+start1 <- SS_readstarter(file = file.path(dir1.1,
+                                          "starter.ss"),
+                              verbose = FALSE)
+dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
+                        verbose = FALSE)
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
+                                    start1$ctlfil),
+                        verbose = FALSE,
+                        use_datlist = TRUE, 
+                   datlist = dat1)
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
+                                                "forecast.ss"),
+                              verbose = FALSE)
+parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
+ row.names( parbio)<-c("Nat M",
+                       "Lmin", 
+                       "Lmax",
+                       "VonBert K",
+                       "CV young",
+                       "CV old", 
+                       "Wt a", 
+                       "Wt b",
+                       "L50%", 
+                       "Mat slope")
+ SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
+ Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
+ Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
 
-The natural mortality rate (\(M\)) was specified using a single parameter per morphotype.
+ parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
+parInit %>% 
+  kable(format = "latex",
+        booktabs = TRUE, 
+        caption = "Input parameters for the initial SS3 model of krill.
+Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
+If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
+  kable_styling(latex_options = c("striped"), 
+                full_width = FALSE, 
+                font_size = 9) %>%
+  pack_rows(index = c("Natural Mortality" = 1,
+                      "Growth" = 5,
+                      "Length-Weight Relation" = 2,
+                      "Maturity" = 2,
+                      "Stock-Recruit Relation" = 5,
+                      "Catchability" = 2,
+                      "Selectivity" = 4))
+```
+
+### Growth conditioning
+
 Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization.
 The model assumed linear growth below a predefined settlement age of 1.5 years.
-The asymptotic length (\(L_{\infty}\)) was determined by the estimated length-at-age of 999 years, indicating that the model used \(L_{\infty}\) as an upper growth boundary.
-Exponential decay of growth beyond maximum age was disabled (, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
-
+The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary.
+Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
 Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
@@
 Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
-Fecundity was estimated as a power function of body weight, where fecundity was computed as \(Eggs = Wt \times (a + b \times Wt)\).
+Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
 The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
-
-The model used a direct parameterization approach for natural mortality (\(M\)), growth (\(G\)), and growth variability (\(CV_G\)), without additional offsets between sexes.
+The model used a direct parameterization without additional offsets between sexes.
 Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
@@
 The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
-
 The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
@@
 Parameter estimation phases were assigned, with negative values indicating fixed parameters.
-
 The survey selectivity was also modeled using a length-based logistic function.
@@
 The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
-
+These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets.
 The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
+assuming that the predator can capture a wider range of krill sizes.
 These parameters were estimated in later phases (positive phase values), allowing flexibility in their estimation.
-
 Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization.
@@
 
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation (option 3).
-This model was selected due to its robustness in representing recruitment dynamics in response to spawning biomass.
-
-The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value ($SR\_LN(R_0)$) of 23, with a prior of 10.
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
+This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF].
+The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10.
 The parameter was freely estimated in phase 1.
-The steepness of the stock-recruitment curve ($SR\_BH\_steep$) was fixed at 0.85 in phase -4, ensuring consistency with previous studies.
 The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
 The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
-The $SR\_autocorr$ parameter was set to 0 and fixed, indicating that no temporal autocorrelation in recruitment deviations was assumed.
-
 Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
@@
 The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
-
-The recruitment deviation phase was set to 2.
 Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
@@
 The maximum bias adjustment factor was 0.8274.
-
 To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-No cyclic patterns in recruitment were assumed, and advanced recruitment options were enabled for model flexibility.
+Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.
 
@@
 One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock].
-The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
+The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment in this way;
 
-\[
+$$
 E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-\]
+$$
 
-\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
-\(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
+$R$ represents the variability of deviations, adding to the variance caused by environmental factors.
+$SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable.
+Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
 However, the model does not automatically adjust for this.
+Furthermore, this integrated model took into account environmental variables, as chlorophyll-a, known to influence krill abundance and reproductive outputs.
+These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
+Figure \@ref(fig:mapachl) presents a time series of chlorophyll-a (Chl-a) concentration (mg/m-3) in the waters around the Antarctic Peninsula from 2000 to 2020.
+Each panel represents a different year, showing spatial variations in Chl-a levels, which indicate phytoplankton biomass.
+The highest concentrations are observed along the coastal regions, while offshore areas exhibit lower Chl-a levels.
+This visualization helps assess interannual variability in primary productivity, which is a key driver of the Antarctic krill population dynamics.
 
-Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
+```{r mapachl, eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in 48.1 area"}
+knitr::include_graphics('Figs/chla_grilled.png')
+```
 
-### Predator components modelling in `SS3`
+### Predator components modelling
 
@@
 
-All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
+In SS3, natural mortality (*M*) can be estimated using different methods, depending on the data available and the model structure.
+Since a penguin population index is being incorporated as an ecosystem variable, natural mortality can be modeled to reflect predator-driven mortality effects on Antarctic krill.
+Generally, natural mortality can be fixed or estimated.
+If *M* is assumed constant, it is set as a fixed input parameter based on literature values or prior studies.
+In this case, in referencee model, we use *M* described by @Tanasekos and @Maschete.
+In this case, SS applies a penalized likelihood estimation, with constraints set through priors or data-informed boundaries.
+Since *M* is influenced by the **penguin population index**, it can be modeled as a function of this ecosystem variable.
+Stock Synthesis allows *M* to be linked to parameters covariates, meaning that predator abundance (penguins) can influence krill mortality as follows:
+
+$$
+M_t = M_0 + \beta \cdot X_t
+$$
+
+where $M_t$ is the time-varying natural mortality, $M_0$ is the baseline mortality, $X_t$ is the standardized penguin index, and $\beta$ is the estimated effect of the predator index on mortality.
+If predator abundance is highly variable, process error can be incorporated to allow *M* to fluctuate over time, rather than being purely deterministic.
+This introduces stochasticity, improving realism in the predator-prey interaction.
+The equation with process error is:
+
+$$
+M_t = M_0 + \beta \cdot X_t + \epsilon_t, \quad \epsilon_t \sim N(0, \sigma^2)
+$$
+
+where $\epsilon_t$ represents the process error term, normally distributed with mean zero and variance $\sigma^2$.
+
+Using a time-varying mortality function linked to penguin abundance improves realism in krill stock assessments by incorporating predator-driven mortality as a key ecosystem variable.
+Figure \@ref(fig:mapapre) presents a temporal and spatial overview of penguin prey length data from 2001 to 2020.
+The maps focus on the Antarctic Peninsula region, showing how spatial patterns of prey collection have changed over time.
+
+```{r mapapre, eval=TRUE, fig.cap="predator temporal and spatial variables.
+The top panel shows boxplots of prey lengths (in cm) across years, with different colors representing prey consumed by three penguin species: Adélie (red), chinstrap (orange), and gentoo (black).
+A red horizontal line marks the 3.6 cm as recruit size, likely for comparison purposes.
+The bottom panel consists of small maps for each year, displaying the geographic distribution of prey data.
+Colored circles indicate locations and intensity of sampling by species, with circle size suggesting sampling effort or frequency."}
+knitr::include_graphics('Figs/Predator_fleet.png')
+```
+
+For stock assessment purposes, we integrated all penguin colony abundances into a standardized index, which represents a comprehensive measure of population trends over time.
+This index accounts for variations in colony sizes and survey efforts, providing a robust metric that reflects the overall status of the penguin population.
+By using this standardized index, we ensure comparability across different time periods and regions, facilitating a more accurate assessment of the relationship between penguin population dynamics and ecosystem variability.
+Figure \@ref(fig:mapapre2) shows a time series of a standardized abundance index for penguin colonies, spanning from approximately 1990 to 2020.
+The index appears to exhibit fluctuations over time, with a general decline from the early 1990s to the mid-2000s, followed by a period of relative stabilization and another decline toward the end of the time series.
+The wide error bars suggest substantial variability in abundance estimates, likely reflecting natural population fluctuations, survey effort differences, or environmental influences.
+
+```{r mapapre2, eval=TRUE, out.width="60%", fig.cap="Standardized abundance index of penguin colonies (1990–2020). Open circles represent annual estimates, with error bars indicating uncertainty"}
+knitr::include_graphics('s1.4/plots/index1_cpuedata_PREDATOR.png')
+```
 
@@
 
-### Stock Assessment Scenarios
+## Scenarios
 
-To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
-The methodology involves comparing the statistical performance of the model regarding these configurations.
-We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).
+To account for the uncertainty associated with ecosystem influence in krill dynamic popupation, we evaluated a set of model configurations that incorporate variations in predator "fleet" and environmental effects, and the relationship between spawning biomass and recruitment.
+Spatial heterogeneity implicity is considering in all scenarios.
+The methodology involves assessing the statistical performance of the model under these configurations.
+Specifically, `s1.1` serves as the reference model, incorporating only fishery and survey data while accounting for spatial heterogeneity.
+The subsequent scenarios build upon this configuration to evaluate the effects of additional ecological components.
+`s1.2` extends `s1.1` by incorporating predator influences, allowing the model to assess the impact of predation on krill population dynamics.
+`s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers affecting krill recruitment and biomass.
+Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more comprehensive assessment of ecosystem interactions and their role in krill population dynamics (Table 1).
 
 | Scenario | Description                                                |
-|:------------:|:---------------------------------------------------------|
+|:--------:|:-----------------------------------------------------------|
 |   s1.1   | Spatial data without environmental and predator components |
@@
 
-## Results
+## Diagnosis
 
-Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
+Carvalho et al.
+(2021) propose a structured approach for diagnosing stock assessment models to ensure their reliability and robustness.
+The diagnostic process involves several key steps.First, convergence analysis is conducted to assess whether the estimation algorithm has reached a stable solution.
+This includes checking for model convergence by examining parameter trace plots, gradients, and the Hessian matrix to confirm that all parameters are well-estimated.
+Next, the fit to data is evaluated by comparing observed and predicted values for key data sources, such as catch, survey indices, and length composition.
+Residual plots are used to detect systematic patterns, which may indicate model misspecification.
+A retrospective analysis is then performed to assess the consistency by scenario estimates over time.
+This is done by sequentially removing recent years of data (peeling) and re-estimating the model.
+The retrospective bias (*Rho*) is calculated for key population metrics, such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
+The model’s forecasting ability is tested through one-step-ahead forecasts, comparing predicted and observed values for recent years.
+This step helps identify potential overfitting or biases in recruitment or abundance estimates.
+Sensitivity analysis is conducted by varying key parameters, in this case, ecosystem variables.
+Likelihood profiling is performed to explore parameter uncertainty and identify influential priors.
+Prediction skill testing involves using hindcasting methods to evaluate model performance in predicting past stock trends.
+Alternative models are compared to determine the most reliable configuration.
 
-1.
-Stock-Recruitment Relationship: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.
-2.
-Predation Impact: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.
-3. Biomass Trends: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.  
-4. Fisheries Interaction: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
+Four estimation models were developed depending on how the ecosystem components was integrated where it was included as an additional covariate within the stock-recruitment (S-R) relationship.
+To evaluate the differences in model performance, we analyzed the Root Mean Squared Error (RMSE) values across four scenario configurations (`s1.1`, `s1.2`, `s1.3`, and `s1.4`).
+The RMSE values were extracted from the variance tuning checks of each scenario and compiled into a data frame for statistical analysis.
+To test for overall differences among scenarios, a one-way ANOVA [@REF] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
+If the ANOVA indicated significant differences, a post-hoc Tukey test would be performed to determine which scenario comparisons were driving the differences.
+By systematically applying these diagnostic steps, stock assessment models can be rigorously evaluated, ensuring they provide robust and unbiased estimates for fisheries management.
 
-Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
-Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
-Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
+# RESULTS
 
 ### Tablas a incluir
+
 1.  Comparación de parámetros clave entre escenarios:
+
     -   Biomasa desovante (SSB, Spawning Stock Biomass) en años clave.
@@
     | Parámetro           | Base | Con variables ecosistémicas |
-   |----------------|------|----------------------------|
+    |---------------------|------|-----------------------------|
     | SSB (2020)          | X    | Y                           |
@@
 2.  Resumen del ajuste del modelo:
+
     -   Log-verosimilitud total.
@@
     -   Error estándar de las estimaciones clave.
+
+```{r eval=TRUE}
+parameter_comparison_table <- read_csv("parameter_comparison_table.csv")
+parameter_comparison_table %>%  
+  mutate(across(where(is.numeric), ~ format(., decimal.mark = "."))) %>%  
+  flextable() %>%  
+  set_caption("Comparison of Model Parameters") %>%  
+  autofit() %>%  
+  theme_vanilla() %>%  
+  fontsize(size = 10, part = "all") %>%  
+  font(fontname = "Arial", part = "all") %>%  
+  set_table_properties(layout = "autofit", width = .75)
+
+```
 
@@
 ### Gráficos recomendados
+
 1.  Evolución de la biomasa desovante (SSB) en el tiempo
@@
     -   Inclusión de intervalos de confianza.
-
 2.  Reclutamiento anual estimado por escenario
@@
     -   Promedio móvil para suavizar tendencias.
-
 3.  Relación stock-reclutamiento con diferentes ajustes
@@
     -   Comparación de la fuerza del reclutamiento en diferentes condiciones ambientales.
-
 4.  Proyecciones de biomasa y F bajo diferentes escenarios de explotación
@@
     -   Comparación entre políticas de manejo con y sin ecosistema.
-
 5.  Distribución de longitudes o edades en el tiempo
     -   Para ver si el escenario con variables ecosistémicas predice cambios en la estructura de la población.
-
 6.  Curvas de Kobe plot (B/BMSY vs. F/FMSY)
@@
 
----
-
-### 4. Discussion  
-Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
-The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
-
-- Ecosystem-Based Management Implications: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
-By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.
-- Predator-Krill Interactions: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
-Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.
-- Climate Change Considerations: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
-Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
-
----
-
-### 5. Conclusion  
-This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
-The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
-Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
-
-Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
-
----
-
-### Acknowledgments  
-We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
-
-\newpage
-
-# RESULTS
-
 ## Diagnosis Base Model
 
-All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
+All scenarios converged within the tolerance criteria and the final gradient using maximum likelihood is achieve and was relatively small (\< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
 About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
@@
 In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions (Figure \@(fig:recdev)).
-We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @@McBride2021].
+We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; \@ @McBride2021].
 
@@
 
+Statistical diferences.
+The *t-test* assumes normality and equal variance and resulted in a *statistically significant difference* between the two models (*t = 4.2682, df = 72, p \< 0.001*).
+The 95% confidence interval for the mean difference ranged from *1,387,011 to 3,818,074*, with *mean spawning biomass estimates of 4,600,304 t for Model 1.1 and 1,997,762 t for Model 1.4*.
+
+The *Wilcoxon rank sum test*, a non-parametric alternative that does not assume normality, also indicated a significant difference (*W = 1256, p \< 0.001*), confirming the robustness of the result.
+The warning regarding ties suggests some repeated values in the dataset, but this does not invalidate the conclusion.
+
+These findings indicate that incorporating environmental and predator effects into the model (Model 1.4) leads to significantly lower estimates of spawning biomass compared to the reference model (Model 1.1).
+This suggests that ecosystem variables play a crucial role in the assessment of krill population dynamics and should be considered in stock evaluation frameworks.
 \newpage
 
-# DISCUSSION
+# DISCUSION
 
@@
 
+Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
+
+1.
+Stock-Recruitment Relationship: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.\
+2.
+Predation Impact: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.\
+3.  Biomass Trends: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.\
+4.  Fisheries Interaction: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
+
+Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
+Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
+Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
+
+Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
+The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
+
+-   Ecosystem-Based Management Implications: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
+By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.\
+-   Predator-Krill Interactions: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
+Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.\
+-   Climate Change Considerations: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
+Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
+
 \newpage
@@
 
+This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
+The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
+Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
+
+Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
+
 \newpage
+
+# ACKNOWLEDGMENTS
+
+We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
 
@@
 
+````{=tex}
 \begin{landscape}
@@
 \end{landscape}
+````
 
@@
 \newpage
+
 ```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}

### 2025-04-03 index.Rmd
adding text

@@
   mutate(
-    # Standardize 'obs' between 0 and 1
     index_std = (obs - min(obs)) / (max(obs) - min(obs)),
-
-    # Recode 'index' values
     index = recode(index,
@@
                    `11` = "PREDATOR"),
-
-    # Convert 'index' to a factor with specific levels
     index = factor(index, levels = c("FISHERYBS", 
@@
                                      "PREDATOR")),
-
-    # Assign categories based on 'index'
     category = case_when(
@@
       index == "PREDATOR" ~ "Predator",
-      TRUE ~ NA_character_  # In case there are any unexpected values
+      TRUE ~ NA_character_  
     )
@@
 
-# Gráfico con 3 filas, dejando 'PREDATOR' en la última fila
-ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
+p1 <- ggplot(dat1_std, aes(x = year, y = index_std, group = index, color = category)) +
   geom_point() +
@@
   labs(title = "", x = "", y = "Standard index", color = "Category")
-
+p1
 ```
@@
 
-ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
+p2 <- ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
   geom_density_ridges_gradient(scale = 1.8,
@@
        fill = "Category") +
-  theme_few()+
+  theme_bw()+
   theme(legend.position = "none")
-
+p2
 ```
@@
 
-Data used en both (spatial and No spatial models) `s1.1`
+Data used en both (spatial and No spatial models) `s1.1` and `s1.4` to compare
 
 ```{r }
+par(mfrow=c(1,2), mar=c(5,4,2,1))
 SSplotData(base.model1.1, 
@@
            pheight = 15)
-```
-
-and `s1.4`
-
-```{r }
+
 SSplotData(base.model1.4, 
@@
 
-```{r}
-
+```{r fig.height=3, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1)) 
@@
 
-```{r}
+```{r fig.height=3, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
@@
             subplots = 1)
+mtext("s1.4", side=3, line=0.5, cex=1.2)
+```
+```{r}
+SSplotBiology(base.model1.1,
+            subplots = 24)
+```
+
+Natural Mortality
+
+```{r}
+par(mfrow=c(2,2), mar=c(5,4,1,1))
+SSplotBiology(base.model1.1,
+            subplots = 21)
+mtext("s1.1", side=3, line=0.5, cex=1.2)
+SSplotBiology(base.model1.2,
+            subplots = 21)
+mtext("s1.2", side=3, line=0.5, cex=1.2)
+SSplotBiology(base.model1.3,
+            subplots = 21)
+mtext("s1.3", side=3, line=0.5, cex=1.2)
+SSplotBiology(base.model1.4,
+            subplots = 21)
 mtext("s1.4", side=3, line=0.5, cex=1.2)
@@
 ```
+Viz
+
+```{r}
+tablebias_combined <- tablebias_combined %>%
+  mutate(Rho.peel = factor(Rho.peel, levels = unique(Rho.peel)))
+
+ggplot(tablebias_combined, aes(x = Rho.peel, y = Rho.Rho, group = Model, fill = Model)) +
+  geom_point(size = 3, shape = 21, color = "black") +  
+  geom_hline(yintercept = 0, color = "grey") +
+  facet_wrap(~Quant, scales = "free_y") + 
+  scale_fill_manual(name = "Scenario",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  scale_color_manual(name = "Scenario",
+                     values = c("s1.1" = "#ca0020", 
+                                "s1.2" = "#f4a582", 
+                                "s1.3" = "#bababa", 
+                                "s1.4" = "#404040")) +
+  theme_minimal() +
+  labs(title = "",
+       x = "",
+       y = "Rho Value") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        legend.position = "bottom")
+
+```
+
+The plot presents the retrospective bias (*Rho*) in Spawning Stock Biomass (SSB) and Fishing Mortality (F) across different peel years (2019–2016) and for the combined period under four model scenarios (s1.1 to s1.4).
+In the SSB panel, all models exhibit negative *Rho* values across the years, indicating a downward bias.
+The magnitude of the bias varies among models, with s1.3 and s1.4 showing the most pronounced deviations.
+In the F panel, *Rho* values are consistently positive, suggesting an upward bias.
+The combined *Rho* values for each metric confirm these trends, showing persistent differences among model scenarios.
+
+(Discussion)
+
+The observed retrospective bias in SSB and F aligns with findings from previous studies on krill population dynamics, where recruitment variability and model structure influence stock assessment estimates (Hill et al., 2016; Watters et al., 2020).
+The negative bias in SSB suggests that models may overestimate past biomass, potentially due to unaccounted environmental variability or underestimation of natural mortality (Meyer et al., 2021).
+The positive bias in F indicates a tendency to underestimate past fishing mortality, which could be linked to the inclusion of predator-based indices such as penguin foraging data (Trathan et al., 2018).
+These results highlight the importance of incorporating ecosystem interactions into stock assessments to improve predictive accuracy and support sustainable management of Antarctic krill.
 
@@
 ```
+
 ## Outputs
@@
 
+```{r}
+# Unir los data frames en uno solo
+outpsall <- rbind(outps1,  
+                  outps2,
+                  outps3,
+                outps4)
+```
+
+```{r}
+# Crear una columna de modelo en cada data frame
+outps1$Model <- "Ref Model: No Env-Predator"
+outps2$Model <- "s1.2 w/ Predator data"
+outps3$Model <- "s1.3 w/ Env data"
+outps4$Model <- "s1.4 w/ Env and Predator data"
+
+# Unir los data frames en uno solo
+outpsall <- rbind(outps1,  outps4)
+
+outpsall_long <- outpsall %>%
+  pivot_longer(cols = c(Bio_smry, SpawnBio, Recruit_0), 
+               names_to = "Variable", values_to = "Value")
+
+# Gráfico de barras agrupadas por variable y modelo
+ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  facet_wrap(~Variable, 
+             ncol=1,
+             scales = "free_y") + 
+  scale_fill_manual(name= "Scenario",
+                    values = c("Ref Model: No Env-Predator" = "#ca0020",
+
+                               "s1.4 w/ Env and Predator data" = "#404040")) +
+  labs(title = "",
+       x = "", y = "") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        legend.position = "bottom")
+```
+
 ## Comparison
@@
                        getcovar = F)
+summaryoutputall <- SSsummarize(biglist)
+```
+
+```{r}
+legend.labels <- c('s1.1','s1.2','s1.3', 's1.4')
+
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist <- SSgetoutput(keyvec = NULL, dirvec = c(dir1.1,
+                                                 dir1.2,
+                                                 dir1.3,
+                                                 dir1.4),	getcovar = F)
 
 #create summary of model runs from list above
-summaryoutputall <- SSsummarize(biglist)
-```
-
-```{r}
-comtable <- SStableComparisons(summaryoutputall,
+summaryoutput <- SSsummarize(biglist)
+
+tablelike <- SStableComparisons(summaryoutput,
                    likenames = c("TOTAL",
@@
                              "R0",
-                             "steep",
-                             "NatM",
-                             "L_at_Amax", 
-                             "VonBert_K", 
                              "SSB_Virg", 
@@
                    digits = NULL,
-                   modelnames = c("s1.1", "s1.2",
-                                  "s1.3",
-                                  "s1.4"),
-                   csv = F,
-                   csvdir =~"/DOCAS/SA_Krill",
-                   csvfile = "parameter_comparison_table.csv",
-                 verbose = TRUE,
+                   modelnames = legend.labels,
+                   csv=TRUE,
+                   csvdir = "/Users/mauriciomardones/DOCAS/SA_Krill",
+                   csvfile = "parameter_comparison_table.csv", verbose = TRUE,
                    mcmc = FALSE)
-
-kbl(comtable, booktabs = T,format = "latex",
-    caption = "Comparision s1.1, s1.2, s1.3, s1.4")  %>% 
+kbl(tablelike, booktabs = T,
+    format = "html",
+    caption = "")  %>% 
     kable_styling(latex_options = "scale_down")
-
 ```
@@
 
-### Comparsion in sd long term time series
-
-```{r fig.height=6, fig.width=8}
-
-sblt1 <-  base.model1.1$timeseries$Bio_all
-sblt2 <-  base.model1.2$timeseries$Bio_all
-sblt3 <-  base.model1.3$timeseries$Bio_all
-sblt4 <-  base.model1.4$timeseries$Bio_all
-
-errt <- as.data.frame(cbind(sblt1,
-              sblt2,
-              sblt3,
-              sblt4))
+### Comparsion in sd long term time series forecasting
+
+```{r fig.height=5, fig.width=8}
+
+errt <- data.frame(
+  Bio_all_1 = base.model1.1$timeseries$Bio_all,
+  Bio_all_2 = base.model1.2$timeseries$Bio_all,
+  Bio_all_3 = base.model1.3$timeseries$Bio_all,
+  Bio_all_4 = base.model1.4$timeseries$Bio_all,
+
+  Bio_smry_1 = base.model1.1$timeseries$Bio_smry,
+  Bio_smry_2 = base.model1.2$timeseries$Bio_smry,
+  Bio_smry_3 = base.model1.3$timeseries$Bio_smry,
+  Bio_smry_4 = base.model1.4$timeseries$Bio_smry,
+
+  SpawnBio_1 = base.model1.1$timeseries$SpawnBio,
+  SpawnBio_2 = base.model1.2$timeseries$SpawnBio,
+  SpawnBio_3 = base.model1.3$timeseries$SpawnBio,
+  SpawnBio_4 = base.model1.4$timeseries$SpawnBio,
+
+  Recruit_0_1 = base.model1.1$timeseries$Recruit_0,
+  Recruit_0_2 = base.model1.2$timeseries$Recruit_0,
+  Recruit_0_3 = base.model1.3$timeseries$Recruit_0,
+  Recruit_0_4 = base.model1.4$timeseries$Recruit_0
+)
+
 err_long <- errt %>%
   pivot_longer(cols = everything(), 
-               names_to = "variable",
-               values_to = "value")  
-titles <- c("Ref Model: No Env-Predator",
-                 "S1.1 w/ Predator data",
-                 "S1.1 w/ Env data",
-                 "S1.1 w/ Env and Predator data")
-ggplot(err_long, aes(x = variable, y = value)) +
-  geom_boxplot(width=0.5) +
-  geom_jitter(width = 0.2, alpha = 0.3) +
-  labs(title = "Long Term Predictions",
+               names_to = "Variable_Model",
+               values_to = "Value")
+
+err_long <- err_long %>%
+  extract(Variable_Model, into = c("Variable", "Model"), 
+          regex = "(.+?)_(\\d+)$")
+
+err_long <- err_long %>%
+  mutate(Model = case_when(
+    Model == "1" ~ "Ref Model: No Env-Predator",
+    Model == "2" ~ "S1.1 w/ Predator data",
+    Model == "3" ~ "S1.1 w/ Env data",
+    Model == "4" ~ "S1.1 w/ Env and Predator data",
+    TRUE ~ Model
+  ))
+
+ggplot(err_long, aes(x = Model, y = Value, fill = Model)) +
+  geom_boxplot(width = 0.5, outlier.shape = NA,
+               outliers = F) +
+  #geom_jitter(width = 0.2, alpha = 0.1) +
+  facet_wrap(~Variable, ncol=4, 
+             scales = "free_y") +
+  scale_fill_manual(name = "Scenario",
+                    values = c("#ca0020", "#f4a582", "#bababa", "#404040")) +
+  labs(title = "",
        x = "",
-       y = "Spawning Biomass (t)") +
+       y = "") +
   theme_few() +
-  scale_x_discrete(labels= titles)+
-  theme(axis.text.x = element_text(angle = 90,
-                                   hjust = 1))
-```
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        legend.position = "none")
+```
+
+Statistical diferences
+
+```{r}
+# Extraer los datos
+spawn_bio_1 <- base.model1.1$timeseries$SpawnBio
+spawn_bio_4 <- base.model1.4$timeseries$SpawnBio
+
+# Prueba de normalidad (Shapiro-Wilk)
+shapiro.test(spawn_bio_1)
+shapiro.test(spawn_bio_4)
+
+# Prueba de igualdad de varianzas (Levene)
+library(car)
+leveneTest(c(spawn_bio_1, spawn_bio_4), 
+           group = rep(c("Model 1.1", "Model 1.4"), each = length(spawn_bio_1)))
+
+# Prueba t de Student (si los datos son normales)
+t.test(spawn_bio_1, spawn_bio_4, var.equal = TRUE)
+
+# Prueba de Wilcoxon (si los datos no son normales)
+wilcox.test(spawn_bio_1, spawn_bio_4)
+
+```
+
+**Two-Sample t-Test**
+
+- **t-value:** 4.2682  
+- **Degrees of Freedom (df):** 72  
+- **p-value:** 5.915e-05 (very small)  
+- **95% Confidence Interval for the Mean Difference:** [1,387,011; 3,818,074]  
+- **Sample Means:**  
+  - Model 1.1: **4,600,304**  
+  - Model 1.4: **1,997,762**
+
+Since the **p-value (5.915e-05) is much smaller than 0.05**, we reject the null hypothesis.
+This means there is **a statistically significant difference** between the spawning biomass predicted by Model 1.1 and Model 1.4.
+The confidence interval suggests that Model 1.1 tends to estimate a higher spawning biomass than Model 1.4, with an estimated difference between **1.39 and 3.82 million tons**.
+
+**Wilcoxon Rank Sum Test (Mann-Whitney U Test)**
+
+- **W-statistic:** 1256  
+- **p-value:** 6.698e-10 (extremely small)  
+- **Warning:** "Cannot compute exact p-value with ties" (indicates that some values are repeated)
+
+The **Wilcoxon test also rejects the null hypothesis** with an extremely small **p-value (6.698e-10)**.
+This confirms that the distribution of spawning biomass values is significantly different between the two models.
+Since the Wilcoxon test does not assume normality, it supports the conclusion from the **t-test**, reinforcing that
+
+Both tests strongly indicate that **Model 1.1 and Model 1.4 produce significantly different estimates of spawning biomass**.
+Specifically, **Model 1.1 predicts higher values**.
+This suggests that the factors or assumptions included in Model 1.4 result in **lower spawning biomass estimates**, which could have important implications for stock assessment and fisheries management.
 
@@
 ```{r}
-
 df1 <- base.model1.1$likelihoods_used %>%
@@
   mutate(Componente = rownames(base.model1.1$likelihoods_used), Modelo = "s1.1")
-
 df2 <- base.model1.2$likelihoods_used %>%
@@
   mutate(Componente = rownames(base.model1.2$likelihoods_used), Modelo = "s1.2")
-
 df3 <- base.model1.3$likelihoods_used %>%
@@
   mutate(Componente = rownames(base.model1.3$likelihoods_used), Modelo = "s1.3")
-
 df4 <- base.model1.4$likelihoods_used %>%
@@
                 df4)
-
-ggplot(df, aes(x = Componente, y = values, fill = Modelo)) +
+ggplot(df %>% 
+         filter(values > 1), aes(x = Componente, y = values, fill = Modelo)) +
   geom_col(position = position_dodge(), width = 0.7) +  
@@
 ```
+This bar plot presents the log-normal likelihood contributions of different model components in an Antarctic krill stock assessment under four scenarios (s1.1, s1.2, s1.3, and s1.4).
+
+**TOTAL Likelihood:** The overall likelihood is highest for scenario s1.4 (black), followed by s1.2 (light orange).
+This suggests that s1.4 provides the best overall model fit, while s1.1 (dark red) has the lowest likelihood, indicating a poorer fit.
+**Survey Component:** The survey likelihood follows a similar pattern, with s1.4 and s1.2 showing better agreement with the data compared to s1.1.  
+**Recruitment, Parameter Priors, and Parameter Deviations:** These components contribute minimally to the total likelihood, indicating that they do not heavily influence model fit differences.  
+**Length Composition:** There is some variation in this component, with s1.2 showing a larger contribution than s1.1 and s1.3.
+
+### **Overall Conclusion:**  
+Scenario s1.4 seems to fit the data best, followed closely by s1.2.
+Scenario s1.1 has the poorest likelihood, meaning it is the least supported by the data.
+If s1.4 includes both environmental and predator effects, this suggests that incorporating ecosystem variables significantly improves the stock assessment model for Antarctic krill.
 

### 2025-04-11 MS2_Krill.Rmd
friday!

--- 2025-03-28 MS2_Krill.Rmd
@@
 ---
-title: "Assessing environmental and predation effects on Antarctic Krill (*Euphausia superba*) stock productivity: An Integrated Catch-at-Length Model Perspective"
+title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated catch-at-length model perspective"
 author:
-  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
+  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
   - Erica Jarvis Mason^[Antarctic División NOAA, USA]
-  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
+  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
   - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base]
+  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
+  - Richard Methot ^[NOAA, Fishery Science División, NorthWeast, Seattle, USA]
 output:
@@
 library(r4ss)
+library(readr)
+library(flextable)
+library(officer)
+library(r4ss)
 ```
@@
 
-Understanding the drivers of krill (*Euphausia superba*) population dynamics is crucial for sustainable management.
-This study evaluates the impact of environmental variables and predation components on krill productivity and other population parameters, in the Western Antarctic Peninsula (WAP).
-Using an integrated Catch-at-Length model, we assess how ecosystem factors influence recruitment, biomass trends, and stock status.
-Our results indicate that krill population variables, as well as productivity estimates, tend to be more optimistic in terms of stock status when environmental and predators variables are not considered.
-This highlights a critical challenge in developing management recommendations and harvest control rules, as ignoring ecosystem influences may lead to overly optimistic assessments of krill stock health.
-Our findings emphasize the necessity of incorporating ecosystem variables into population stock assessments process, as neglecting these factors may lead to biased estimations of krill population dynamics.
-We consider that the use of integrated models, which leverage the extensive population, fisheries, and ecosystem data available through various scientific programs, can provide valuable insights for management aligned with CCAMLR conservation and harvesting objectives.
+Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP).
+In this study, we implement an integrated Catch-at-Length model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
+We explored multiple scenarios that accounted for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species.
+Performance analyses reveal systematic differences across scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables.
+Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics.
+Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through post hoc analyses.
+Overlooking these drivers may result in less accurate conclusions and limit the effectiveness of management recommendations.
+Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s conservation and sustainable use objectives.
 
-*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Catch-to-Length model, environmental factors, predation components, fisheries management.*
+*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Stock Synthesis model, perfomance model, fisheries management, CCAMLR*
 
@@
 While model-based stock assessments have been widely used (REF), many struggle to align with management objectives or accurately represent krill population dynamics due to their spatial variability and ecosystem complexity.
-One widely applied approach is the General Yield Model (GRYM) (REF), which provides global catch limits based on predefined harvest thresholds.
-Traditionally, stock assessments have relied on single-source data, either from fisheries catches [@Kindzey; @Wang] or population monitoring surveys [@Green2023; @Watters2013].
-However, these approaches often overlook critical factors such as spatial heterogeneity in krill abundance, environmental drivers, and predator-prey interactions.
-Given the influence of environmental variables on krill recruitment and the role of predators in adding an additional component of natural mortality, any stock assessment framework must integrate these ecosystem factors to better represent population dynamics.
+Traditionally, stock assessments have relied on single-source data, either from fisheries catches [@Kinzey2015a; @Wang2021] or population monitoring surveys [@Green2023; @Watters2013].
+One widely applied approach is the General Yield Model (GRYM), which provides global catch limits based on predefined harvest thresholds [@Maschette2020].
+However, such approaches often neglect key ecological drivers.
+Considering the influence of habitat on population dynamics, the role of environmental variability in shaping krill productivity, and the contribution of predators as a source of natural mortality, stock assessment frameworks should explicitly incorporate these ecosystem components to more accurately represent population processes.
 
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture spatial variability, and account for ecosystem interactions.
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
 However, integrating these complexities presents several challenges.
@@
 Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations.
-This requires not only refining existing datasets but also adopting a fully ecosystem-based management approach.
 
-To address this need, we propose a conceptual framework for developing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past four decades.
-This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations.
-Specifically, we employ an integrated catch-at-length stock synthesis model that allows us to evaluate the effects of environmental drivers on krill recruitment and productivity, as well as the impact of predator-prey interactions on overall population dynamics.
-Our approach involves comparing different scenarios, one incorporating ecosystem-related variables and another based solely on traditional stock assessment data, to determine how environmental fluctuations and predator abundance influence key population variables and parameters.
+To address this need, we propose a conceptual framework for implementing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past decades.
+This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations and productivity.
+
+Specifically, we employ an integrated catch-at-length implemented in Stock Synthesis model [@Methot2013] that allows us to evaluate the effects of environmental drivers on recruitment and productivity, as well as the impact of predator components into an integrated stock assessment model to Antarctic krill.
+The primary objective of this study was to assess the statistical performance of different management metrics, population variables and parameters based on alternative approaches to incorporating both environmental and predator data into a typical krill stock assessment model.
 By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends.
@@
 
-By integrating ecosystem drivers into the stock assessment process, we enhance the accuracy of population estimates and improve our ability to predict future stock trends under changing environmental conditions.
-This broader ecological perspective strengthens the scientific foundation for sustainable management recommendations, ensuring that harvesting strategies align with an ecosystem-based management approach.
-Ultimately, the insights gained from this integrated modeling effort contribute to the ongoing efforts of CCAMLR's new strategy to balance krill fishery sustainability with the conservation of the broader Antarctic marine ecosystem.
+By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions.
+This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
+Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
 
@@
 
-The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)).
-The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill.
-Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.
+The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale.
+To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021] (Fig 1 S1).
+While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale.
+This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics.
+This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.
+
+## Conceptual Model
+
+In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets [REF].
+Differences among strata have been identified from various perspectives of krill population dynamics [REF], as well as from the influence of environmental variability within Subarea 48.1 [REF] (Fig 2 S1).
+This approach allows integrate spatial heterogeneity into the stock assessment model to ensure that spatially structured processes are adequately represented.
+Although strata are treated as separate fleets, recruitment and growth processes remain unified, ensuring that the overall krill stock is modeled as a single population while accounting for spatially structured impacts at the Subarea 48.1 level.
+Each stratum is assigned a unique fishery and survey history, reflecting differences in harvesting intensity and spatial exploitation patterns.
+Survey indices and length composition data are fleet-specific, allowing for variations in data collection methods and regional abundance estimates.
+Selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
+The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January regarding @REF.
+This approach represents a spatially implicit model, as it captures spatial variability in exploitation patterns without explicitly modeling movement or recruitment connectivity between areas.
+This allows for a more realistic assessment of krill stock productivity, in line with CCAMLR’s ecosystem-based management framework.
 
@@
 Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling.
-All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
+These data sources include data from the monitoring program of the krill fishery were used, as standar CPUE (kg/time hauls) and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
+This database has information about vessel, nationality, georeferenced, among others.
+Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including biomass index by strata, and biological characteristics of krill collected through net sampling.
+All this data was processed in @Dornam2021.
+Krill length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
 Mat.
-2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver.
-Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U.
-Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
-This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP.
-We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)).
-The model included spatial component (cell) as random effect (formula: ~1 | cell).
-Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl.
-Mat 3); 4) Predator abundance data from "MAPPPD".
-Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].
+2).
 
-## Conceptual Model
+```{r eval=TRUE}
+data <- data.frame(
+  Source = c("Krill Fishery Monitoring Program",
+                 "Krill Fishery Monitoring Program, collected by scientific observers under CCAMLR SISO program (vessel, nationality, georeferenced)",
+                 "Krill Fishery Monitoring Program",
+                 "AMLR Research Cruises Program", 
+                  "Satellite data",
+                 "Predator Abundance",
+                 "AMLR Research Cruises Program"),
+  Description = c("Catches",
+                  "CPUE (standarized by vessel, area and deep)", 
+                  "Krill Length compositions (grouped by 0.2 cm)",
+                  "Biomass index from scientific AMLR cruises",
+                  "Satellite-based chlorophyll-a concentration (monthly, interpolated) from the EU Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).",
+                  "Data on penguin colonies, sourced from 'MAPPPD'",
+                  "Predator Length composition krill diet"),
+    TimeSeries = c("1997-2020",
+                 "1997-2020",
+                 "2000-2020",
+                 "1991-2011 (based on the research cruise periods).",
+                 "Monthly data from 1997 to present (based on satellite data availability).",
+                 "Data from 2000s-present.",
+                 "1977-2020 (from 2000 used in this analysis)"),
+   Scale = c("by Strata",
+                           "by Strata",
+                           "by Strata",
+                           "by Strata",
+                           "48.1",
+                           "by colony",
+                           "by colony"))
+ft <- flextable(data) %>%  
+  set_caption("Comparison of Model Parameters") %>%  
+  autofit() %>%  
+  theme_vanilla() %>%  
+  fontsize(size = 9, 
+           part = "all") %>%  
+  font(fontname = "Arial", 
+       part = "all") %>%  
+  set_table_properties(layout = "autofit", width = .85)
+ft
+```
+Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
+Source data and temporal scale used in this stock assessment is showed in Fig 4, S1.
 
-A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously.
-Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels.
-The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.
-
-## Catch-to-Length Model Implementation
-
-Like most invertebrates, krill is a species for which aging is complex [@Punt2013].
-Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure.
-Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables.
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
-SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions.
-SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023].
-Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit).
-Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).
-
-We applied an integrated Catch-to-Length model to estimate krill stock-recruitment relationships and population trends.
-The model links catch-at-length data with growth, mortality, and recruitment processes while incorporating environmental and predation variables.
-Model performance was evaluated through likelihood-based comparisons, sensitivity analyses, and cross-validation.
-Key model parameters include:
-
-### Natural Mortality and Growth Parameterization
-
-The natural mortality rate (\(M\)) was specified using a single parameter per morphotype.
-Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization.
-The model assumed linear growth below a predefined settlement age of 1.5 years.
-The asymptotic length (\(L_{\infty}\)) was determined by the estimated length-at-age of 999 years, indicating that the model used \(L_{\infty}\) as an upper growth boundary.
-Exponential decay of growth beyond maximum age was disabled (, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
-
-Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
-
-### Maturity and Reproduction
-
-Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
-Fecundity was estimated as a power function of body weight, where fecundity was computed as \(Eggs = Wt \times (a + b \times Wt)\).
-The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
-
-The model used a direct parameterization approach for natural mortality (\(M\)), growth (\(G\)), and growth variability (\(CV_G\)), without additional offsets between sexes.
-Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
-
-### Selectivity Parameterization
-
-Selectivity functions were modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
-The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
-
-The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
-The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm.
-Parameter estimation phases were assigned, with negative values indicating fixed parameters.
-
-The survey selectivity was also modeled using a length-based logistic function.
-Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
-The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
-
-The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
-These parameters were estimated in later phases (positive phase values), allowing flexibility in their estimation.
-
-Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization.
-
-### Spawner-Recruitment Relationship Parameterization
-
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation (option 3).
-This model was selected due to its robustness in representing recruitment dynamics in response to spawning biomass.
-
-The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value ($SR\_LN(R_0)$) of 23, with a prior of 10.
-The parameter was freely estimated in phase 1.
-The steepness of the stock-recruitment curve ($SR\_BH\_steep$) was fixed at 0.85 in phase -4, ensuring consistency with previous studies.
-The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
-The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
-The $SR\_autocorr$ parameter was set to 0 and fixed, indicating that no temporal autocorrelation in recruitment deviations was assumed.
-
-Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
-The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
-The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
-
-The recruitment deviation phase was set to 2.
-Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
-Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
-The maximum bias adjustment factor was 0.8274.
-
-To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-No cyclic patterns in recruitment were assumed, and advanced recruitment options were enabled for model flexibility.
-
-### Environmental variable modelling in `SS3`
-
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock].
-The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.
-
-\[
-E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-\]
-
-\(R\) represents the variability of deviations, adding to the variance caused by environmental factors.
-\(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced.
-However, the model does not automatically adjust for this.
-
-Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs.
-These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics.
-By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.
-
-### Predator components modelling in `SS3`
-
-This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics.
-By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.
-In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously.
-This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.
-
-All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).
-
-\newpage
-
-### Stock Assessment Scenarios
-
-To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment.
-The methodology involves comparing the statistical performance of the model regarding these configurations.
-We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).
-
-| Scenario | Description                                                                          |
-|:------------:|:---------------------------------------------------------|
-|    s1.1  | Spatial data without environmental and predator components                           |
-|    s1.2  | "s1.1" with predator components                                                      |
-|    s1.3  | "s1.1" with environmental variable                                              |
-|    s1.4  | "s1.1" w/ both, predator fleet and environmental variable                        |
-
-: Scenarios to modelling dynamics in krill
-
-## Results
-
-Our model results indicate that ecosystem variables significantly influence krill productivity and population structure:
-
-1.
-Stock-Recruitment Relationship: Recruitment was positively correlated with sea ice extent and primary productivity, suggesting a strong link between krill reproduction and environmental conditions.
-2.
-Predation Impact: Penguin predation exerted a measurable effect on juvenile krill survival, with increased predation pressure leading to lower recruitment estimates in years with high predator abundance.
-3. Biomass Trends: Environmental drivers accounted for a substantial portion of biomass fluctuations, with warm years showing lower krill abundance, likely due to reduced primary productivity.  
-4. Fisheries Interaction: Years with high krill harvest coincided with lower biomass estimates, highlighting the cumulative impact of both natural and anthropogenic pressures on krill stocks.
-
-Para reportar los distintos escenarios obtenidos del stock assessment en SS3, puedes enfocarte en comparar las diferencias clave entre el escenario base y el que incorpora variables ecosistémicas.
-Como mencionaste, el escenario con variables ecosistémicas es más pesimista en términos de reclutamiento y otras variables poblacionales.
-Aquí tienes algunas ideas de tablas y gráficos para reportar los resultados:
-
-### Tablas a incluir
-1. Comparación de parámetros clave entre escenarios:
-   - Biomasa desovante (SSB, Spawning Stock Biomass) en años clave.
-   - Tasa de explotación (F).
-   - Reclutamiento medio.
-   - Relación B/B0 (biomasa actual respecto a la virgen).
-   - Relación F/FMSY.
-
-   | Parámetro       | Base | Con variables ecosistémicas |
-   |----------------|------|----------------------------|
-   | SSB (2020)     | X    | Y                          |
-   | SSB (2025)     | X    | Y                          |
-   | F (2020)       | X    | Y                          |
-   | Reclutamiento medio | X | Y                        |
-   | B/B0 (2020)    | X    | Y                          |
-   | F/FMSY (2020)  | X    | Y                          |
-
-2. Resumen del ajuste del modelo:
-   - Log-verosimilitud total.
-   - Verosimilitud de los datos de captura, índices de abundancia y composición de tallas/edades.
-   - Número de parámetros estimados.
-   - Error estándar de las estimaciones clave.
-
-3. Comparación de reclutamiento por año:
-   - Media y desviación estándar del reclutamiento en periodos clave (por ejemplo, 1990-2000, 2000-2010, 2010-2020).
-   - Probabilidad de años de reclutamiento bajo.
-
-### Gráficos recomendados
-1. Evolución de la biomasa desovante (SSB) en el tiempo  
-   - Líneas para cada escenario mostrando tendencias divergentes.
-   - Inclusión de intervalos de confianza.
-
-2. Reclutamiento anual estimado por escenario  
-   - Barras para cada año con comparación entre escenarios.
-   - Promedio móvil para suavizar tendencias.
-
-3. Relación stock-reclutamiento con diferentes ajustes  
-   - Ajustes de la curva con y sin variables ecosistémicas.
-   - Comparación de la fuerza del reclutamiento en diferentes condiciones ambientales.
-
-4. Proyecciones de biomasa y F bajo diferentes escenarios de explotación  
-   - Predicciones de la biomasa y la explotación para los próximos 5-10 años.
-   - Comparación entre políticas de manejo con y sin ecosistema.
-
-5. Distribución de longitudes o edades en el tiempo  
-   - Para ver si el escenario con variables ecosistémicas predice cambios en la estructura de la población.
-
-6. Curvas de Kobe plot (B/BMSY vs. F/FMSY)  
-   - Comparar el estado del stock en cada escenario respecto a puntos de referencia de manejo.
-
-(Imagen Verosimilitud)
-
-The comparison of likelihood components across the three scenarios revealed notable differences in their fit to the data.
-Scenario S1.1 had the lowest total likelihood, suggesting a better overall fit compared to S1.2 and S1.3.
-The catch and equilibrium catch components showed minor variations among scenarios, indicating that these data had a limited influence on the overall likelihood differences.
-In terms of survey likelihood, S1.1 exhibited a slightly better fit to survey abundance data, while S1.2 and S1.3 showed higher likelihood values, suggesting a reduced agreement with observed biomass trends.
-The length composition component differed significantly, with S1.2 and S1.3 having higher likelihood values than S1.1, implying a poorer fit to observed length distributions.
-Recruitment likelihood was highest in S1.1, indicating stronger consistency with recruitment estimates compared to S1.2 and S1.3.
-Additionally, parameter deviations were more constrained in S1.1, whereas S1.2 and S1.3 exhibited greater variability.
-These results suggest that S1.1 provides a better overall fit, particularly in survey abundance and length composition components, while S1.2 and S1.3 show greater discrepancies in these aspects.
-
----
-
-### 4. Discussion  
-Our findings emphasize the necessity of incorporating ecosystem variables into krill population assessments.
-The significant influence of environmental variability and predation highlights the limitations of traditional stock assessment models that assume static population processes.
-
-- Ecosystem-Based Management Implications: Our study supports the integration of ecosystem-based reference points in krill fishery regulations.
-By considering environmental and predation effects, management frameworks can better anticipate population fluctuations and minimize risks of overexploitation.
-- Predator-Krill Interactions: The observed impact of penguin predation aligns with previous studies that indicate krill availability is a primary driver of penguin reproductive success.
-Managing krill fisheries in a way that ensures sufficient prey availability for predators is critical for maintaining ecosystem balance.
-- Climate Change Considerations: Given the ongoing warming trends in the Antarctic Peninsula, krill populations may face increased environmental stress.
-Our results suggest that future management strategies should incorporate climate-driven changes in krill habitat suitability.
-
----
-
-### 5. Conclusion  
-This study demonstrates that incorporating environmental and predation variables into stock assessments significantly improves our understanding of Antarctic krill population dynamics.
-The integrated Catch-to-Length model provides a robust framework for assessing krill productivity under varying ecosystem conditions.
-Our results advocate for an ecosystem-based approach to krill fisheries management, ensuring that decision-making processes account for the broader ecological context.
-
-Future research should explore additional ecosystem interactions and refine predictive models to support adaptive management strategies under changing climatic conditions.
-
----
-
-### Acknowledgments  
-We thank CCAMLR for providing access to data and acknowledge the support of the scientific community contributing to krill population research.
-
-\newpage
-
-# RESULTS
-
-## Diagnosis Base Model
-
-All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis.
-About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`.
-In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS).
-The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`.
-For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000.
-However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity.
-For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices.
-The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).
-
-## Retrospective analysis
-
-Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better.
-@Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)).
-To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual.
-Overall, `s1` had a good performance when evaluating the lengths compositions (14.2%), however  RMSE showed a low predictive power with respect to the indices (71.3%).
-A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).
-
-## Likelihood profile in spatial Model `s1`
-
-For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources.
-The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro).
-The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.
-
-Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)
-
-## Comparision outputs between scenarios
-
-In comparative terms with all the tested scenarios, we check the consistency through recruitment diversions (Figure \@(fig:recdev)).
-We do this through spatial autocorrelation analysis, checking the fluctuation of recruitment through the time series analyzed, that is, between 1979 and 2020 and which has consistency as described by other authors regarding pulses between three and five years [@Perry2020; @@McBride2021].
-
-Autocorrelation in krill recruitment in time series analyzed by scenarios, models `s01`, `s1`, `s2`, and `s5` exhibit significant autocorrelations at various lags, indicating dependencies and potential cyclical patterns in recruitment.
-Models `s3` and `s4` show minimal to no significant autocorrelations, suggesting more random recruitment patterns.
-Model `s6` shows moderate autocorrelations, indicating some predictability based on recent past values, In summary, `s01`, `s4`, and `s6` exhibit no significant autocorrelation, suggesting randomness (Figure \@ref(fig:acf)).
-
-To evaluate consistency of the source of information used in each of the tested scenarios, we verify its influence on the estimate of R0 through likelihood components and, like scenario 1, we verify that all scenarios show a high dependence on the composition lengths of lengths with respect to the estimate of the variable (as show the (Figure \@ref(fig:likeall)).
-
-We identify differences in the different scenarios related with population variables with respect to the long-term estimates where the implicit model `s1` and the models without stock recruit relation or with a low density-dependence on the stock-recruit relation (`s2`and `s3`) show higher predictions than the rest of the tested scenarios.
-This is relevant given that these scenarios consider low spatial dependence, density-dependence spawning biomass with respect to recruitment, which is plausible for resources such as krill where environmental dynamics shape population dynamics (Figure \@ref(fig:longterm)).
-
-We carried out a specific comparison between the most divergent scenarios that we tested in this case, `s1`,  `s01` and `s7` which according to the description are the aggregated spatial scenario, disaggregate spatial scenario and scenario with new set of parameters, respectively.
-All with ecosystem components such as predators and an environmental variable.In this sense, the impacts and difference on the estimates of recruitment,  R0, virginal biomass and relative biomass are high (Figure \@ref(fig:comparision)), demonstrating the impact of initial considerations and assumptions on population variables estimation.
-
-\newpage
-
-# DISCUSSION
-
-The analysis presented here represents a modern approach for stock assessment of complex dynamics such as krill and the integration of ecosystem components as environmental variables and ecological aspects such as predators.
-The development of an integrated stock assessment model like could provide support to adaptive management, wherein management procedures resulting, like status or quota, are adjusted based on ecosystem monitoring indices [@Wang2021].
-However, integrated models are not new to krill in SO.
-For several years, CCAMLR has been developing and applying integrated assessment methodologies that incorporate diverse data sources, such as acoustic and trawl surveys, to improve the accuracy of their stock assessments.
-In 2016, through an external review process,@Thomson2016 explicitly recommended *"the use of an Integrated Analysis model for assessment of Antarctic krill in FAO Subarea 48.1 is appropriate and would constitute an improvement over the modeling strategies currently in use to manage krill"* and he stated that the an integrated modeling framework *"has the potential to substantially improve the scientific basis for future krill management"*.
-
-@Kinzey2015a and @Kinzey2019a used integrated models with survey data to test the impacts of selectivity and recruitment estimation in certain areas of Subarea 48.1.
-@Kinzey2011 implemented a model to test spawning biomass, recruitment and movement between areas associated with the northern Antarctic Peninsula.
-Although these models had convergence issues, they represented significant advances in this type of framework.
-Additionally, @Wang2021 made contributions to integrated statistical models with survey and fishery data for the northern sector of the Western Antarctic Peninsula, focusing on the Bransfield Strait, testing different data weightings.
-Furthermore, examined correlations between recruitment and climatic variables such as the Southern Annular Mode (SAM) El Niño Oscillation Southern (ENSO).
-In absolute values and with respect to biomass levels, although the estimates of our model were lower than the previously described models, there is a coincidence with respect to recruitment pulses as well as mortality levels.
-
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by a myriad of factors including oceanographic conditions and predator foraging behavior.
-By incorporating spatial heterogeneity into our model considering the strata as part of the analysis, we aim to capture the complex spatial patterns of krill distribution.
-This spatially implicit approach will allow us to identify krill hotspots, areas of high productivity, and potential ecological corridors, providing valuable insights for conservation planning and management strategies [@Watters2013] and transfer those signals of differences in krill population structure into the model to generate its estimates.
-
-From an ecological point of view, the study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
-There are many studies that confirm the influence of predation on krill populations within the Antarctic Peninsula [@Silvestro2023; @Reisinger2022].
-For this, we consider the penguin  population as one of the most powerful elements of natural predation on krill because they have direct correlations between this type of taxonomic group and krill  [@Kruger2021].
-The intention to incorporate elements of predation in a stock is necessary to improve the modeling, because it did not account for uncertainty in natural mortality underestimated uncertainty in current stock biomass by as much as 20% [@Hollowed2000].
-
-In a scientifically rigorous stock assessment, this analysis included comparisons between various natural scenarios and assumptions used to model krill population dynamics.
-It is important to note that the models presented here can be refined in terms of initial conditioning to improve fits on length compositions of predators and indices that were initially low.
-However, a primary objective was to evaluate the impact on key population variables, considering factors such as spatial heterogeneity and life history parameters.
-This impact was verified in this stock assessment approach, identifying aspects that must be considered when making recommendations for krill management by CCAMLR using models.
-
-By comparing model outputs under scenarios where spatial heterogeneity is assumed versus not assumed, the significance of incorporating spatial variation into the model can be assessed.
-This methodology enables researchers to make informed decisions about the necessity of accounting for spatial heterogeneity in understanding krill distribution and abundance, thereby enhancing the accuracy and reliability of ecological modeling efforts.
-
-Nowadays, one of the frontiers of stock assessment is the incorporation of ecosystem and spatial considerations into modeling.
-In this context, the features of the SS3 platform allow us to advance in this direction and contribute to the discussion, both in understanding the population dynamics of krill and providing advice for sustainable management.
-In summary, our proposed stock assessment scheme represents a pivotal approach to understanding the complex population dynamics of Antarctic krill in Subarea 48.1.
-
-Predator indesx
-
-A simple predation pressure index for modeling changes in naturalmortality: Application to Gulf of Maine northern shrimp stockassessment (Richard et al 2016)
-
-Env index.
-
-Simulation analyses indicated that the integrated approach with additional process error is superior to the traditional method of correlating model estimates with environmental variables outside the estimation procedure.
-The results suggest that, for the snapper stock, recruitment is positively correlated with SST at the time of spawning.
-(Maunder et al 2003)
-
-The ecosystem variables used in stock assessment processes vary in both their definition and how they are applied.
-Marshall et al.
-(2019) classify ecosystem variables based on their impact on population dynamics, as well as the ways in which they are incorporated into analytical models.
-
-\newpage
-
-# CONCLUSION AND FUTURE WORK
-
-- The main challenge in this analysis has been to identify sources of information about krill population dynamic (fishery, survey, environmental, predator and life history parameters) and model it into an integrated model, in this case, with Stock Synthesis platform.
-This approach is not related to the *"best model"*, but rather to the implementation of the different options of scenarios that krill faces in multiple natural dimensions and evaluate the impact of this on population estimates.
-
-- Regarding to spatial heterogeneity, this can be considered in two ways in a modeling approach.
-First one, the complexity and differences in distribution in abundance and other components of the krill population can be incorporated into an stock assessment in advance, as was carried out in this exercise, to have a more reliable estimate of the population variables that and transfer these considerations to the management for decision making.
-Second one, spatial heterogeneity can be considered for the decomposition of the management procedures that are established into the population assessed.
-For this, it is necessary.
-have a spatially explicit model, which was not our case but can be considered a future approach to be analysed.
-
-- Ecosystem considerations for a model integrated with stock assessment are vital.
-Although SS3 does not model the ecosystem variables, it allows for these variables to be incorporated to model krill population dynamics.
-SS3 presents characteristics that allow the incorporation of this kind of variable.
-This ecosystem information into stock assessment serve as an input to condition the moderate population dynamics, in any case we consider this an advance that can be perfected but that is inside discussion with the ecosystem management in the new strategy proposed by CCAMLR.
-
-- We consider that an integrated modeling of the krill adding the available information sources (fishery and surveys), as well as, the spatial heterogeneity and ecosystem variables is on the right path of the recommendation that has emerged in the last 10 years with respect to population dynamics and its link with the management procedure carried out by CCAMLR.
-
-\newpage
-
-# SUPPLEMENTARY MATERIAL
-
-- Supplementary Material 1:  Model template example and initial condition to `s1` scenario can be found in this [repo](https://github.com/MauroMardones/SA_Krill/tree/main/test) and the executable version (3.30.21) can be download from NOAA Virtual Lab [SS3](https://vlab.noaa.gov/web/stock-synthesis)
-
-- Supplementary Material 2:  Length composition templates preparation to SS3 from AMLR can be found [link](https://mauromardones.github.io/AMLR_Length_Data/) and routine code in this [repo](https://github.com/MauroMardones/AMLR_Length_Data)
-
-- Supplementary Material 3:  Correlation analysis between krill biology and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_Length_Cor/) and routine code in this [repo](https://github.com/MauroMardones/Krill_Length_Cor)
-
-\newpage
-
-# FIGURES AND TABLES
-
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
-knitr::include_graphics('Figs/481.png')
-```
-
-```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
-knitr::include_graphics('Figs/conceptual.jpeg')
-```
-
-```{r datas2, out.width='70%', fig.cap="Source information time series for each fleet used in s1 scenario"}
-knitr::include_graphics('Figs/data_plot_s2.png')
-```
-
-```{r parainit, fig.cap="Initial biology and fishery parameters to set a model s1 in krill in WAP"}
-
-# leo archivos para plotear y hacer tablas
-start1 <- SS_readstarter(file = file.path(dir1,
+```{r parainit1, eval=TRUE}
+dir1.1<-here("s1.1") 
+start1 <- SS_readstarter(file = file.path(dir1.1,
                                           "starter.ss"),
                               verbose = FALSE)
-# note the data and control file names can vary, so are determined from the 
-# starter file.
-dat1 <- SS_readdat(file = file.path(dir1, start1$datfile),
+dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
                         verbose = FALSE)
-# Read in ctl file. Note that the data fileR object is needed so that SS_readctl
-# assumes the correct data structure
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1,
+ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
                                     start1$ctlfil),
@@
                    datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1, 
+fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
                                                 "forecast.ss"),
                               verbose = FALSE)
-# can also read in wtatage.ss for an empirical wt at age model using
-# r4ss::SS_readwtatage()
-
 parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
@@
  Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
+
  parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-
 parInit %>% 
-  kbl(booktabs = TRUE,
-      format = "latex",
-      position="ht!",
+  kable(format = "latex",
+        booktabs = TRUE, 
         caption = "Input parameters for the initial SS3 model of krill.
@@
 If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-  kable_paper("hover", 
-              full_width = F)%>%
   kable_styling(latex_options = c("striped"), 
                 full_width = FALSE, 
-                font_size=9)%>% 
+                font_size = 9) %>%
   pack_rows(index = c("Natural Mortality" = 1,
-                        "Growth"= 5,
-                        "Length-Weigth Relation" = 2,
-                        "Maturity"=2,
-                        "Stock-Recruit Relation"=5,
-                        "Catchability"=2,
-                        "Selectivity"=4))
+                      "Growth" = 5,
+                      "Length-Weight Relation" = 2,
+                      "Maturity" = 2,
+                      "Stock-Recruit Relation" = 5,
+                      "Catchability" = 2,
+                      "Selectivity" = 4))
 ```
 
-\begin{landscape}
+## Catch-to-Length Model Implementation
 
-```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
-knitr::include_graphics('Figs/pathmod.png')
+The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
+SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
+The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources described previously and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
+All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
+
+Like many invertebrates, krill present challenges for age determination [@Punt2013].
+Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017].
+As a result, stock assessments have relied on catch-at-length models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length.
+Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics.
+Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
+Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure.
+To date, all integrated krill assessments have used catch-at-length frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021].
+Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003].
+An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes.
+ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
+This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks.
+The ALK matrix is typically defined as:
+
+$$
+P(A = a \mid L = l) = \frac{N_{a,l}}{\sum_{a} N_{a,l}}
+$$
+where \( P(A = a \mid L = l) \) is the probability that an individual of length \( l \) belongs to age \( a \),  \( N_{a,l} \) is the number of individuals of age \( a \) observed at length \( l \) and the denominator \( \sum_{a} N_{a,l} \) ensures that the probabilities sum to 1 for each length class.
+This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:
+
+$$
+N_a = \sum_{l} P(A = a \mid L = l) \cdot N_l
+$$
+
+where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age.
+In integrated stock assessment models, such as those using statistical catch-at-age (SCAA) or integrated length-based frameworks, the ALK provides a critical link between observable length data and underlying age-based population dynamics.
+
+The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
+The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
+In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
+In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
+This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass.
+Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in Figure \@ref(fig:AKL)).
+
+```{r AKL, eval=TRUE, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in Subarea 48.1"}
+knitr::include_graphics('Figs/AKL.png')
 ```
 
-\end{landscape}
+Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization.
+The model assumed linear growth below a predefined settlement age of 1.5 years.
+The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary.
+Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
+Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
+Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
+Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
+The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
+The model used a direct parameterization without additional offsets between sexes.
+Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
+Regarding selectivity functions, it was modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
+The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
+The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
+The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata.
+Parameter estimation phases were assigned, with negative values indicating fixed parameters.
+The survey selectivity was also modeled using a length-based logistic function.
+Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
+The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
+These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (Fig 6, S1).
+The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
+assuming that the predator can capture a wider range of krill sizes.
+These parameters were estimated in later phases, allowing flexibility in their estimation.
+Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
+This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF].
+The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10.
+The parameter was freely estimated in phase 1.
+The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
+The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
+Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
+The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
+The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
+Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
+Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
+To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
+Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.
+All of these configurations can be reviewed in the repository section of S1.
 
-```{r res1, fig.cap = "Pearson residuals, comparing across fleets. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page1_multi-fleet_comparison.png')
+## Environmental variable modelling
+
+To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables previously recognized as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
+SST data were obtained from ERA5 monthly means (1991–present), SIC data from Nimbus-7 (1978–present), and Chla data from the Bio-Geo-Chemical L4 satellite product, downloaded from the E.U.
+Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
+All environmental data were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations.
+We model the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power.
+The initial model considered only the effect of strata (spat), while subsequent models sequentially included Chla, SST, Sea Ice Index and interaction terms.
+This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process.
+All models and their results can be found in the Supplementary Material 2.
+
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [REF] and SS3 addresses these approaches.
+According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment.
+This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;
+
+$$
+E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
+$$
+$R$ represents the variability of deviations, adding to the variance caused by environmental factors.
+$SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable.
+Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
+However, the model does not automatically adjust for this.
+Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment.
+This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability.
+By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics esttimates.
+
+## Predator modelling
+
+As @methot2020stock establish in SS3 manual, natural mortality attributable to predation (M2) can be incorporated through different types of observational data, each informing distinct aspects of predator-prey dynamics.
+First, total mortality due to predation can be input as a "discard" component in the data structure.
+Since M2 represents a proportion of total mortality (Z), the quantity of individuals removed by predators can be estimated using the relative contribution of M2 to Z, i.e., `M2 / Z`.
+This allows the model to estimate the biomass consumed by predators, analogous to how catch is calculated for fishing fleets.
+Second, predator abundance can be input as an index of effort, treating M2 as a fishing mortality analog.
+This approach uses the same structure as a fishery-independent survey, where predator biomass serves as an index of predation pressure over time.
+In this framework, the predation rate is proportional to predator abundance scaled by a catchability coefficient \( Q \), following the equation:
+
+$$
+M2_t = Q * I_t
+$$
+where \( M2_t \) is the predation mortality in year \( t \), \( I_t \) is the predator abundance index in year \( t \), and \( Q \) is an estimated or fixed catchability parameter linking the predator index to predation pressure.
+Finally, information on the size or age composition of predated individuals, such as those derived from gut content analyses, can be used to inform the selectivity pattern of M2.
+These data are input analogously to length- or age-composition observations from fisheries, and allow the model to estimate the predation selectivity curve, capturing how predation probability varies with prey size or age.
+
+Regarding krill predators, we will incorporate both biomass and dietary data from the three main penguin colonies —Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*) —into the stock assessment model through two complementary pathways..
+First, the abundance of these predator colonies will be treated as an index of predation "effort" exerted on the krill population, allowing estimation of time-varying predation mortality rates.
+Second, empirical length-frequency distributions of krill derived from the diet composition of each penguin species will be used to parameterize the size-selective nature of M2, enabling the estimation of predator-specific selectivity functions.
+This dual approach facilitates a temporally and size-structured characterization of M2, consistent with ecosystem-based stock assessment principles and improving the ecological realism of krill population dynamics modeling.
+The total natural mortality (M) will thus explicitly include the predation component, expressed as \( M = M_1 + \sum M_2 \) [@methot2020stock], where \( M_1 \) represents background (non-predation) natural mortality, and \( \sum M_2 \) accounts for mortality attributed to each predator type.
+Consequently, total mortality is represented as \( Z = M_1 + M_2 + F \), incorporating both natural and fishing sources of mortality.
+
+\newpage
+
+## Scenarios
+
+To account for the uncertainty associated with ecosystem influence in krill dynamic popupation, we evaluated a set of model configurations that incorporate variations in predator "fleet" and environmental effects, and the relationship between spawning biomass and recruitment.
+Spatial heterogeneity implicity is considering in all scenarios.
+The methodology involves assessing the statistical performance of the model under these configurations.
+Specifically, `s1.1` serves as the reference model, incorporating only fishery and survey data while accounting for spatial heterogeneity.
+The subsequent scenarios build upon this configuration to evaluate the effects of additional ecological components.
+`s1.2` extends `s1.1` by incorporating predator influences, allowing the model to assess the impact of predation on krill population dynamics.
+`s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers affecting krill recruitment and biomass.
+Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more comprehensive assessment of ecosystem interactions and their role in krill population dynamics (Table 1).
+
+| Scenario | Description                                                |
+|:--------:|:-----------------------------------------------------------|
+|   s1.1   | Spatial data without environmental and predator components |
+|   s1.2   | "s1.1" with predator components                            |
+|   s1.3   | "s1.1" with environmental variable                         |
+|   s1.4   | "s1.1" w/ both, predator fleet and environmental variable  |
+
+: Scenarios to modelling dynamics in krill
+
+Regarding the performance and comparison between scenarios, @Carvalho2017 and @Carvalho2021b propose a structured approach for diagnosing stock assessment models to ensure their reliability and robustness.
+This diagnostic process was applied to the krill stock assessment and involved some key steps.
+First, convergence analysis was conducted to evaluate whether the estimation algorithm reached a stable solution.
+This included reviewing parameter trace plots, gradients, and the Hessian matrix to ensure that all parameters were well-estimated.
+Next, model fit to data was assessed by comparing observed and predicted values across key data sources, including catch, survey indices, and length composition.
+Residual plots were examined to detect potential patterns indicative of model misspecification.
+A retrospective analysis was performed by sequentially removing recent years of data (peeling) and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
+The model's forecasting capability was evaluated through one-step-ahead predictions and their comparison to observed values in recent years, helping identify potential bias or overfitting in recruitment or abundance estimates.
+Sensitivity analyses were conducted by varying ecosystem variables to assess their influence on model outputs.
+Likelihood profiling was also carried out to quantify parameter uncertainty and identify the influence of priors.
+Finally, prediction skill was assessed using hindcasting techniques to evaluate the model’s ability to reproduce historical stock dynamics.Four estimation models were developed depending on how the ecosystem components was integrated where it was included as an additional covariate within the stock-recruitment (S-R) relationship.
+To evaluate statistical differences in model performance, we analyzed the Root Mean Squared Error (RMSE) values across four scenario configurations (`s1.1`, `s1.2`, `s1.3`, and `s1.4`).
+The RMSE values were extracted from the variance tuning checks of each scenario and compiled into a data frame for statistical analysis.
+To test for overall differences among scenarios, a one-way ANOVA [@REF] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
+If the ANOVA indicated significant differences, a post-hoc Tukey test would be performed to determine which scenario comparisons were driving the differences.
+By systematically applying these diagnostic steps, stock assessment models can be rigorously evaluated, ensuring they provide robust and unbiased estimates.
+
+# RESULTS
+
+### Environmental data
+
+Based on linear models performances and ecological relevance, Chla emerged as the most influential predictor, therefore, this variable was selected for incorporation into the stock assessment model as an environmental covariate (Table 2, S2).
+This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
+Figure \@ref(fig:mapachl) presents a time series of chlorophyll-a (Chl-a) concentration (mg/m-3) in the waters around the Antarctic Peninsula from 2000 to 2020.
+Each panel represents a different year, showing spatial variations in Chl-a levels, which indicate phytoplankton biomass.
+The highest concentrations are observed along the coastal regions, while offshore areas exhibit lower Chl-a levels.
+This visualization helps assess interannual variability in primary productivity, which is a key driver of the Antarctic krill population dynamics.
+
+```{r mapachl, eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in 48.1 area"}
+knitr::include_graphics('Figs/chla_grilled.png')
 ```
 
-```{r res1a, fig.cap = "Pearson residuals, comparing across fleets.
-Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) (continued)."}
-knitr::include_graphics('s1/plots/comp_lenfit__page2_multi-fleet_comparison.png')
+In another hand, Figure \@ref(fig:chlindex) illustrates the interannual variability of a Chlorophyll Variability Index from the year 2000 to 2020.
+The figure shows notable fluctuations across years, highlighting periods of higher-than-average chlorophyll variability (e.g., 2003, 2016) as well as years with significantly lower values (e.g., 2011, 2020).
+These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
+
+```{r chlindex, eval=TRUE, fig.height=2, fig.width=9, fig.cap="Interannual variability of the Chlorophyll Variability Index from 2000 to 2020.
+Positive anomalies (in red) and negative anomalies (in black) indicate below-average conditions"}
+datos <- data.frame(
+  anio = 2001:2020,
+  variable = c(
+    0.684, 0.693, 0.737, 0.683, 0.716, 0.693, 0.670,
+    0.708, 0.653, 0.699, 0.642, 0.647, 0.672, 0.658, 0.694,
+    0.764, 0.675, 0.724, 0.662, 0.622
+  )
+)
+
+min_val <- min(datos$variable)
+max_val <- max(datos$variable)
+datos$indice_ambiental <- 2 * (datos$variable - min_val) / (max_val - min_val) - 1
+
+ggplot(datos, aes(x = anio, y = indice_ambiental)) +
+  geom_col(aes(fill = indice_ambiental > 0), width = 0.5) +
+  scale_fill_manual(values = c("black", "red")) +
+  #geom_line(color = "grey20", size = 1) +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  labs(title = "",
+       x = "", y = "Env Index") +
+  theme_bw() +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 90, hjust = 2))+
+  scale_x_continuous(breaks = seq(from = 2000, to = 2020, by = 1))
 ```
 
-```{r resall, out.width='30%', fig.show='hold', fig.cap="Mean length for each with 95% confidence intervals based on current sample sizes. blue line represent estimated"}
-par(mfrow=c(5,2))
-knitr::include_graphics(c('s1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYJOIN.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_FISHERYSSIW.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_PREDATOR.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYBS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYEI.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYGS.png',
-                          's1/plots/comp_lenfit_data_weighting_TA1.8_SURVEYJOIN.png'))
+### Predator comoponent as driver of krill dynamics
+
+The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lukcas!, REF Vicky].
+Figure \@ref(fig:mapapre) presents a temporal and spatial overview of penguin variables from 2001 to 2020.
+The top panel in Figure \@ref(fig:mapapre) shows annual boxplots of krill length distributions derived from penguin diet samples, stratified by predator species.
+Although median krill lengths remain relatively consistent over time, there is notable interannual variability in the spread and distribution of sizes, which may reflect fluctuations in recruitment strength and availability of krill age classes.
+Foraging activity is concentrated in the Bransfield Strait and surrounding areas, aligning with regions of persistent krill aggregation.
+This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics (middle panel in Figure \@ref(fig:mapapre)).
+For stock assessment purposes, we integrated all penguin colony abundances into a synoptic index of relative predator biomass, which represents a comprehensive measure of population trends over time (bottom panel in Figure \@ref(fig:mapapre).
+This index accounts for variations in colonies abundance providing a robust metric that reflects the overall status of the penguin population.
+The index appears to exhibit fluctuations over time, with a general decline from the early 1990s to the mid-2000s, followed by a period of relative stabilization and another decline toward the end of the time series.
+This index captures broad decadal declines in predator presence, particularly after 2015, potentially reflecting changes in prey availability, environmental conditions, or predator demography.
+
+```{r mapapre, eval=TRUE, fig.cap= "Ecosystem indicators derived from penguin predator data in the Western Antarctic Peninsula.
+Top panel: Annual krill length distributions from penguin diet samples by species: Adélie (red), chinstrap (orange), and gentoo (black).
+A red horizontal line marks the 3.6 cm as recruit size for reference purposes.
+Middle panel: Spatial foraging locations and relative intensity by species across years.
+Colored circles indicate locations and intensity of sampling by species, with circle size suggesting sampling effort or frequency.
+Bottom panel: Synoptic index of relative predator biomass, used as an ecosystem input in the stock assessment model."}
+knitr::include_graphics('Figs/Predator_fleet2.png')
+```
+
+## Assessment of model perfomance and complexity across ecological scenarios
+
+The krill population dynamics model was evaluated under four increasingly complex scenarios.
+Scenario s1.1 represented the baseline spatial implicit model, incorporating only survey and fishery data, and resulted in the estimation of 71 parameters.
+In scenario s1.2, predator-related components were added, including biomass and size-selective predation patterns from key penguin colonies, increasing the parameter count to 107.
+Scenario s1.3 introduced an environmental covariate as a recruitment driver into the baseline model, leading to the estimation of 96 parameters.
+Finally, scenario s1.4 integrated both predator and environmental components, culminating in the highest model complexity with 128 estimated parameters.
+All model scenarios successfully converged within acceptable tolerance thresholds, with final gradients below 1.00e-04, and positive-definite Hessian matrices confirming robust parameter estimation and local identifiability.
+
+Regarding model fit, Pearson residuals of predicted length distributions for krill were evaluated across four modeling scenarios, each incorporating different levels of ecosystem complexity.
+Pearson residuals are used to assess the goodness-of-fit by standardizing the differences between observed and predicted values, allowing for the detection of systematic patterns or biases in model predictions across size classes and sampling sources.
+
+The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population (Figure \@ref(fig:pearson)).
+In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
+Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
+The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals.
+However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
+In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
+This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
+Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure.
+The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
+
+```{r pearson, eval=TRUE, fig.cap= "Pearson residuals of predicted versus observed length-frequency distributions of Antarctic krill across four modeling scenarios (s1.1 to s1.4), each incorporating increasing levels of ecosystem complexity.
+Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components"}
+knitr::include_graphics('Figs/Pearson.png')
+```
+
+### Retrospective analysis
+
+Retrospective analyses were conducted to evaluate model performance across four scenarios incorporating different levels of ecosystem complexity (Figure \@ref(fig:rhoplot)).
+The retrospective pattern for fishing mortality (F) showed a marked increase in bias in scenario s1.4, which included both predator and environmental components, particularly in the terminal year.
+In this scenario, the combined rho value exceeded 3.5, indicating a strong retrospective pattern.
+Scenarios s1.2 (predator only) and s1.3 (environment only) also showed elevated rho values relative to s1.1 (baseline model without ecosystem variables), although the magnitude was lower than in s1.4.
+In contrast, the retrospective pattern for spawning stock biomass (SSB) was consistently negative across all scenarios and years.
+However, models that included environmental components—s1.3 and s1.4—tended to show larger negative biases in SSB, with scenario s1.4 again exhibiting the most pronounced pattern.
+The combined rho values for SSB in s1.3 and s1.4 were close to –0.5 and –0.6, respectively, whereas s1.1 showed a comparatively milder retrospective trend.
+
+```{r rhoplot, eval=TRUE, out.width="75%", fig.cap="Retrospective patterns for fishing mortality (F, left) and spawning stock biomass (SSB, right) across four modeling scenarios differing in ecosystem complexity.
+Each point represents the rho value for a given terminal year (2016–2019), with Combined indicating the average retrospective bias across years in different scenarios"}
+knitr::include_graphics('Figs/unnamed-chunk-46-1.jpeg')
+```
+
+### Likelihood components
+
+Log-likelihood profiles were examined for each model scenario (s1.1 to s1.4) to assess how different components contributed to the overall model fit (Figure \@ref(fig:liketotal).
+Across all scenarios, the TOTAL log-likelihood is primarily influenced by the Survey, Length composition, and Recruitment components, with minor contributions from parameter priors.
+The reference model (s1.1) displays the highest overall likelihood value, particularly due to a stronger fit to the Length composition and Survey data.
+However, this comes with an elevated penalty in the parm priors categories, suggesting potential overfitting or parameter instability.
+In contrast, the ecosystem-informed scenarios—s1.2 (with predator data), s1.3 (with environmental data), and s1.4 (with both)—show slightly lower total likelihoods.
+These models present more balanced likelihood contributions across data components and generally lower penalties for parameter priors and deviations, indicating a more stable parameterization despite reduced absolute fit.
+Notably, the Recruitment likelihood contribution decreases in the ecosystem scenarios compared to the reference, which aligns with previous results showing constrained recruitment variability when ecological drivers are included.
+These variations may reflect differences in model assumptions or input data between scenarios, which could affect how well each model fits particular data sources.
+
+```{r liketotal, eval=TRUE, fig.cap="Log-likelihood contributions for four model components across four krill population model scenarios.
+Bars show the contribution of each likelihood component to the total model fit"}
+knitr::include_graphics('Figs/likesel.png')
+```
+
+Table \@ref(tab:tablike) shows the log-likelihood component values for each scenario.
+It summarizes the contributions of different data sources—such as survey indices, length composition, recruitment, and parameter priors—to the total likelihood in each model configuration (s1.1 to s1.4).
+This comparison helps assess how each scenario fits the data and highlights potential differences in the influence of specific components across models.
+
+```{r tablike, eval=TRUE}
+parameter_comparison_table <- read_csv("parameter_comparison_table.csv") %>% 
+  dplyr::slice(1:6)
+
+parameter_comparison_table %>%
+  mutate(across(
+    where(is.numeric),
+    ~ formatC(., format = "f", digits = 3, decimal.mark = ".")
+  )) %>%
+  flextable() %>%
+  set_caption("Log-likelihood contributions of different components across model scenarios, including the total likelihood and individual contributions from survey indices, length composition data, recruitment, and parameter priors.") %>%
+  autofit() %>%
+  theme_vanilla() %>%
+  fontsize(size = 10, part = "all") %>%
+  #font(fontname = "Arial", part = "all") %>%
+  set_table_properties(layout = "autofit", width = .85)
 
@@
 
-```{r fitcom, out.width='80%', fig.show='hold', fig.cap="Fits to lengths compositions in different fleets in s1 scenario"}
-knitr::include_graphics('s1/plots/comp_lenfit__aggregated_across_time.png')
+### Root Mean Square Error (RMSE)
+
+Another step to measure the model performance in krill stock assessment was evaluated using the Root Mean Square Error (RMSE) across four scenarios of increasing ecological complexity (Figure \@ref(fig:rmseplot)).
+Scenario Model1.1, which excludes both environmental and predator variables, showed the lowest median RMSE and the narrowest interquartile range, suggesting higher consistency in model fits to observed data.
+The inclusion of ecosystem variables in Model1.2 (predator only), Model1.3 (environment only), and Model1.4 (both) resulted in increases in RMSE dispersion and median values, although the differences were relatively moderate.
+Among the ecosystem-informed scenarios, Model1.3 and Model1.4 had comparable RMSE distributions, slightly exceeding those of the predator-only model (Model1.2).
+The increased variability in RMSE under these scenarios suggests a reduction in overall fit quality when ecological complexity is introduced, likely reflecting increased model structural uncertainty or misalignment between ecological covariates and population dynamics.
+
+```{r rmseplot, eval=TRUE, fig.cap="Root Mean Square Error (RMSE) distributions for four krill population model scenarios"}
+knitr::include_graphics('Figs/unnamed-chunk-29-1.jpeg')
+```
+
+FINAL TABLE Suggested by MJ
+
+## Comparision outputs between scenarios
+
+Model outputs under different ecosystem scenarios revealed substantial variation in krill population indicators, including total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio) (Figure \@ref(fig:popvar)).
+The reference model (s1.1), which excluded environmental and predator components, consistently produced higher median estimates and wider uncertainty intervals across all population indicators.
+Inclusion of predator components (s1.2) or environmental variables (s1.3) led to reductions in central tendency and variability for all metrics, with s1.3 (environment-only) showing the most constrained estimates, particularly for recruitment and spawning biomass.
+Scenario s1.4, which incorporated both environmental and predator components, resulted in intermediate biomass and recruitment estimates, with moderate variance compared to the other scenarios.
+Recruitment (Recruit_0) was especially sensitive to the inclusion of ecosystem components.
+The highest median recruitment was observed under s1.2 (predator-only), while the lowest was seen in s1.3 (environment-only), highlighting the contrasting influences of these ecological factors.
+
+```{r popvar, eval=TRUE, fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
+Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)."}
+knitr::include_graphics('Figs/unnamed-chunk-65-1.jpeg')
+```
+
+Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under four modeling scenarios from 1990 to 2020.
+The reference model (blue), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015.
+In contrast, the models that incorporate ecosystem components—predator data (green), environmental data (yellow), or both (red)—estimate consistently lower spawning output throughout the time series.
+The red line, representing the model with both predator and environmental data, remains the most conservative, suggesting that ecosystem-informed models lead to more precautionary estimates of reproductive potential.
+The rigth panel displays the posterior density distributions of unfished spawning biomass (SSB~o~) for the same scenarios.
+The reference model estimates the highest SSBo values (mode ~ 3.2), while the ecosystem-informed models estimate considerably lower SSBo, with the most constrained distribution occurring under the full ecosystem model (red).
+The shift in the distributions indicates that including ecosystem drivers reduces estimates of unfished reproductive capacity, which has implications for reference point calculations and harvest strategies.
+
+```{r compara,eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
+knitr::include_graphics('Figs/COM1compare2_spawnbio_uncertainty.png')
 
@@
 
-```{r retros1, out.width='70%', fig.cap="Retrospective pattern in s1 scenario in Antarctic krill"}
-knitr::include_graphics('Figs/retros1.png')
+The different configurations to estimated posterior distribution of Natural log of virgin recruitment level (R0) suggesting that the choice of covariates has a substantial impact on the estimation of unfished recruitment (Figure \@ref(fig:compara2)).
+The model with both environmental and predator data yields the highest R0 values, while the model with only environmental data results in the lowest estimates.
+This highlights the sensitivity of the recruitment parameter to assumptions about ecological drivers in the stock-recruitment relationship.
+
+```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated spawning output for Antarctic krill under four assessment scenarios"}
+knitr::include_graphics('Figs/COM1compare16_densities_SR_LN(R0).png')
+
 ```
 
-```{r rmses1, out.width='60%', fig.show='hold',  fig.cap="Joint residual plots for index (upper panel) and lengths (bottom panel) fits from different fleets in Antarctic krill"}
-par(mfrow=c(2,1))
-knitr::include_graphics(c('Figs/rmse21.png',
-                          'Figs/rmselens1.png'))
+The stock–recruitment relationship was examined under two model scenarios: s1.1, a reference model excluding ecosystem components, and s1.4, which incorporates both environmental and predator data  (Figure \@ref(fig:proplot)).
+The plot illustrates the relationship between spawning stock biomass (SSB) and recruitment across years, overlaid with a fitted Beverton-Holt function derived from the model with ecosystem covariates.
+Scenario s1.1 (diamonds) displays a wider spread in recruitment estimates across the full range of SSB values, with several years (e.g., 2017, 2014) producing exceptionally high recruitment despite moderate to low SSB.
+In contrast, scenario s1.4 (crosses), which integrates ecosystem variables, generally shows lower recruitment estimates, with reduced variance and more clustered residuals at low SSB values.
+The fitted Beverton-Holt curve (in blue) from s1.4 captures the expected saturating recruitment response to increasing SSB.
+Notably, the inclusion of ecosystem variables appears to reduce extreme recruitment events and leads to a more stable, density-dependent recruitment pattern aligned with theoretical expectations.
+
+Estimating the productivity of Antarctic krill is critical for understanding the species capacity to replenish its population in response to varying levels of spawning biomass.
+Here, we defined the Productivity as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
+For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):
+
+$$
+\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
+$$
+This plot illustrates the relationship between recruitment per unit of spawning stock biomass (Recruitment/SSB) and relative spawning stock biomass (SSB/SSBo) for four scenarios (s1.1 to s1.4)  (Figure \@ref(fig:proplot)).
+As expected under compensatory recruitment dynamics, all scenarios show a declining trend in productivity as SSB increases, indicating higher per-capita recruitment at low stock levels and reduced productivity as the population approaches its unfished biomass.
+The scenarios differ markedly in the magnitude of predicted productivity.
+Scenario s1.1 and s1.3 show higher productivity levels throughout the SSB range when compared to s1.2 and, most notably, s1.4.
+The latter scenario presents the lowest productivity values, indicating either reduced recruitment compensation or differences in the stock-recruitment parameters.
+These contrasting trends highlight how scenario-specific assumptions or structural choices influence estimates of productivity.
+These patterns reflect a more precautionary and ecologically realistic recruitment dynamic, where environmental variability limits the ability of the stock to maintain high productivity under depleted conditions.
+Overall, the results emphasize the influence of ecosystem variables on shaping recruitment dynamics and demonstrate that incorporating environmental drivers into the assessment model leads to more conservative productivity estimates, particularly at low spawning biomass levels.
+
+```{r proplot, eval=TRUE, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios.
+Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
+Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
+knitr::include_graphics('Figs/produc_com.png')
 ```
 
-```{r likepro, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
-knitr::include_graphics('Figs/Profile_s1.png')
+Autocorrelation functions (ACF) of model residuals were used to assess the temporal structure and potential misspecification across the four configurations (`s1.1` to `s1.4`) (Figure \@ref(fig:acfplot)).
+The reference model (`s1.1`) displays significant autocorrelation at the first three lags, indicating serial correlation and potential misspecification.
+In `s1.2`, which includes predator dynamics, the autocorrelation is more dispersed, with significant values at early lags and at lag 13, suggesting the introduction of temporal structure not fully accounted for.
+In contrast, `s1.3`, which incorporates environmental variability, shows a marked reduction in autocorrelation beyond the first lag, with most values within confidence bounds—indicating improved model independence.
+The fully integrated model (`s1.4`), combining predator and environmental effects, exhibits an autocorrelation pattern comparable to `s1.3`, with low residual autocorrelation overall, though slight structure persists at early lags.
+
+```{r acfplot, eval=TRUE, fig.cap="Autocorrelation function (ACF) plots of model residuals for four krill assessment scenarios Blue dashed lines indicate 95% confidence intervals"}
+knitr::include_graphics('Figs/ACF2.png')
 ```
 
-```{r eval=F, mainvar, fig.cap="Main variables outputs (Total Biomass, Biomass at 75%, Spawining biomass and Recruit in t) from estimation in s1 model krill in WAP"}
-out_s2 <- read_excel("DataKrill.xlsx", 
-    sheet = "variable_s2")
-out_s2 %>%
-  kbl(booktabs = TRUE,
-      format = "latex",
-    caption = "Main variables outputs from stock asssessment krill in WAP") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = TRUE,
-                font_size=6 )#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
+A *t-test*, assuming normality and homoscedasticity, revealed a statistically significant difference in spawning biomass estimates between models `s1.1` and `s1.4` (*t = 4.2682, df = 72, p < 0.001*), with a 95% confidence interval for the mean difference ranging from *1,387,011 to 3,818,074*.
+Mean spawning biomass estimates were *4,600,304 t* for `s1.1` and *1,997,762 t* for `s1.4`.
+A non-parametric *Wilcoxon rank sum test* further confirmed this difference (*W = 1256, p < 0.001*), reinforcing the robustness of the result despite minor warnings due to ties in the dataset.
+These findings demonstrate that incorporating predator and environmental covariates into the population model (scenario `s1.4`) leads to significantly lower estimates of spawning biomass.
+This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill.
 
-\newpage
-```{r recdev, out.width='90%', fig.cap="Recruit deviation for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-47-1.jpeg')
-```
+# DISCUSION
 
-```{r acf, out.width='90%', fig.cap="Cross correlation in autoregresive analysis for all scenarios in krill population"}
-knitr::include_graphics('index_files/figure-html/unnamed-chunk-46-1.jpeg')
-```
+The use of integrated models for assessing krill populations in the Southern Ocean has posed significant challenges for CCAMLR, particularly in light of the inherent uncertainties and limitations associated with data-poor regions.
+One of the key challenges is the reliance on models like the General Yield Model (GRYM) [REF], which assumes stable environmental conditions—a premise that increasingly fails to hold true in the context of ongoing climate change [REF].
+GRYM, while useful for certain data-poor applications, faces difficulties due to its heavy dependence on life history parameters, such as growth rates, harvest strategies, and natural mortality.
+Moreover, the assumption of stable environmental conditions in such models is problematic [REF], as climate variability significantly influences krill productivity and survival.
+As a result, predictions based on GRYM and similar models become less reliable, particularly in the face of changing environmental conditions or ecological considerations.
+This highlights the need for more dynamic, ecosystem-informed approaches to stock assessment that consider the evolving nature of environmental and ecological drivers, such as changes in sea ice extent, primary productivity, and predator-prey interactions.
 
-```{r likeall, out.width='90%', fig.cap="Components of total likelihood for all sources and all tested scenarios"}
-knitr::include_graphics('Figs/Likelihoodtotal.png')
-```
+The transition from traditional, model-based approaches like GRYM to more integrated, ecosystem-based models presents a significant opportunity to improve the scientific basis of krill stock assessments.
+However, this shift also requires addressing the trade-offs between model complexity and data availability [REF].
+While integrated models offer a more holistic view of krill population dynamics by incorporating ecosystem variables, they also introduce new challenges related to the accuracy and reliability of model inputs.
+Thus, the development of integrated models requires continuous refinement to ensure that they remain relevant and robust in the face of environmental change.
+[REF]
 
-```{r longterm, out.width='100%', fig.cap="Long term estimation for different scenarios in krill spawning biomass for whole time series (1979-2020"}
-knitr::include_graphics('Figs/longtermsb.png')
-```
+Management strategies should incorporate ecosystem-based thresholds informed by environmental and predation dynamics, as these factors are critical in maintaining krill availability for predators, which is essential for overall ecosystem balance.
+Additionally, the ongoing warming in the Antarctic Peninsula highlights the urgent need to integrate climate-driven changes into habitat suitability models, ensuring that management approaches remain adaptive to shifting environmental conditions.
+
+Although integrated models are not new to krill in the Southern Ocean, CCAMLR has for years applied methodologies that incorporate diverse data sources.
+External reviews, such as @Thomson2016, have endorsed these approaches as improvements over traditional models.
+Previous studies like @Kinzey2011, @Kinzey2015a, @Kinzey2019a, and @Wang2021 have advanced integrated frameworks, though challenges like model convergence remain.
+
+This analysis represents a modern approach to stock assessment of complex dynamics such as krill, integrating ecosystem components like environmental variables and ecological aspects such as predators.
+Integrated stock assessment models can support adaptive management, wherein outputs like status or quotas are adjusted based on ecosystem monitoring indices [REF]
+
+Spatial heterogeneity is a fundamental aspect of krill ecology, driven by factors including oceanographic conditions and predator foraging behavior.
+Incorporating strata into our model captures complex spatial patterns of krill distribution, enabling identification of hotspots, productivity zones, and ecological corridors.
+[REF]
+
+These progressive increments in parameter numbers reflect the added ecological realism and mechanistic detail captured in each scenario, emphasizing the role of top-down and bottom-up processes in krill dynamics.
+
+### Predation and Natural Mortality
+
+The study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
+Evidence indicates that penguin predation exerts strong top-down control on krill populations.
+Penguins, as krill specialists, have reproductive success tied closely to krill availability.
+[REF]
+
+Incorporating predation into stock models addresses the underestimation of uncertainty in natural mortality, which can otherwise skew biomass and productivity estimates by as much as 20% [@Hollowed2000].
+
+Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the joint inclusion of predator and environmental variables yielding the most robust representation of observed length compositions.
+The inclusion of environmental covariates (s1.3 and s1.4) markedly improves the temporal independence of model residuals, as indicated by lower autocorrelation beyond lag 1.
+This implies that environmental variability contributes significantly to explaining observed trends in krill population dynamics.
+Conversely, the predator-only model (s1.2) introduces more persistent autocorrelation, indicating that its effects may require more nuanced representation or additional process structuring.
+Overall, these results highlight the benefit of including environmental signals to reduce residual autocorrelation and enhance model fit.
+This analysis coincides with...
+[REF]
+
+The analysis indicates that incorporating environmental and predator data into the stock assessment framework introduces greater variability in model fit, as measured by RMSE.
+While the baseline model (Model 1.1) demonstrates more stable performance, it potentially achieves this at the cost of biological realism.
+The modest trade-off in predictive accuracy associated with ecosystem-informed models is expected and highlights the need to balance goodness-of-fit with ecological plausibility when developing integrated assessment models.
+
+The inclusion of predator and environmental variables significantly alters model behavior in retrospective diagnostics, particularly increasing bias in terminal-year estimates of fishing mortality and deepening the underestimation of spawning stock biomass (SSB).
+These patterns suggest that while ecosystem-informed models improve the biological realism of the stock assessment, they also pose challenges in terms of stability and predictive consistency.
+Such outcomes underline the importance of further refining the structure and parameterization of ecosystem components when integrating them into assessment frameworks, to balance realism with diagnostic robustness.
+
+Likelihood decomposition reveals that models incorporating ecological data (s1.2–s1.4) yield slightly lower total likelihoods than the reference model (s1.1), reflecting trade-offs between model fit and biological realism.
+The reduced penalty terms and more evenly distributed likelihood contributions in the ecosystem scenarios suggest improved model stability and reduced overparameterization.
+These results reinforce the value of integrating ecosystem variables not only for enhancing ecological interpretation but also for promoting robust model structure and parameter behavior.
+
+### Stock-Recruitment and Biomass Estimates
+
+Integrating environmental and predator components into the population model (s1.4) substantially affects the shape and variability of the stock–recruitment relationship.
+The ecosystem-informed model reduces the occurrence of extreme recruitment outliers and yields a more consistent relationship between SSB and recruitment.
+[REF]
+
+Models that omit ecosystem components tend to produce higher and more uncertain projections, potentially overestimating stock productivity.
+The inclusion of these variables leads to more conservative and constrained outputs, which may better reflect the ecological dynamics governing krill populations.
+
+### Productivity
+
+The inclusion of environmental variables (s1.3 and s1.4) substantially lowers estimates of per-capita productivity, especially at low stock levels.
+These ecosystem-informed scenarios reflect more realistic ecological constraints and reduce the risk of overestimating krill productivity under unfavorable conditions.
+This highlights the importance of accounting for environmental drivers in stock assessment models to avoid overly optimistic projections of stock resilience.
+[REF]
+
+### Ecosystem Variables: Ecological Relevance
+
+Variations in krill size composition in penguin diets have been shown to reflect underlying changes in krill population dynamics, including recruitment pulses and cohort strength [Jeff REFs].
+Thus, penguin diet data serve as a proxy for detecting cyclical patterns in krill recruitment that may not be evident through direct surveys alone.
+[REF]
+
+The ecosystem variables used in assessment processes vary from their definition to how they are applied.
+Environmental variables influencing krill recruitment, as well as predators that introduce an additional component of natural mortality into the model, are among those considered in this study.
+[REF]
+
+Marshall et al.
+(2019) [REF] classify ecosystem variables based on their impact on population dynamics and how they are incorporated into stock assessment models.
+They suggest that integrating variables, especially recruitment, directly within the model results in a more dynamic and accurate representation of population dynamics, as it captures real-time interactions between ecological factors.
+In contrast, external incorporation may not fully capture ecosystem feedback loops, potentially leading to less reliable assessments.
+
+Our analysis, which integrates ecosystem variables directly within the model, aligns with the conclusions of Marshall et al.
+(2019) and provides further evidence that this approach offers superior accuracy in modeling krill population dynamics.
+By including both environmental drivers and predation pressures within the model, our study captures the complexity of krill dynamics more effectively, offering a more realistic representation of krill biomass, recruitment, and mortality processes.
+This integrated approach not only improves model precision but also enhances its predictive power, addressing key ecological factors that influence krill populations.
+
+This analysis represents a significant step forward in the incorporation of ecosystem-based considerations into krill stock assessments in the Southern Ocean.
+By adopting a more comprehensive, ecosystem-informed model, we contribute to the growing body of research supporting the integration of ecosystem variables in stock assessment frameworks [REF].
+This approach is consistent with the new and evolving management strategies of the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR), which increasingly emphasizes ecosystem-based management practices.
+As climate change and ecosystem variability continue to impact krill populations, this analysis provides a foundation for precautionary management strategies that are better aligned with the dynamic nature of marine ecosystems.
 
@@
 
-\begin{landscape}
+# CONCLUSION
 
-```{r comparision, out.width='40%', fig.show='hold', fig.cap="Comparative perfomance with scenarios s01, s1 and s7 to recruit, spawing biomass, biomass ratio (hopefully equal to fraction of unfished) and Virgin biomas probability"}
-par(mfrow=c(2,2))
-knitr::include_graphics(c('Figs/compare10_recruits_uncertainty.png',
-                          'Figs/compare2_spawnbio_uncertainty.png',
-                          'Figs/compare4_Bratio_uncertainty.png',
-                          'Figs/compare17_densities_SSB_Virgin.png'))
-```
+The integration of ecosystem variables, such as penguin predation and environmental conditions, significantly enhances the accuracy of krill population dynamics models by capturing complex and time-varying ecological processes.
+This approach provides a more realistic reflection of fluctuations in krill biomass and recruitment.
+The results emphasize that environmental conditions, particularly primary productivity, directly affect krill recruitment dynamics, highlighting the importance of considering climate variability in the sustainability of krill fisheries, especially in light of global climate change.
+Furthermore, predation by species like penguins plays a crucial role in the natural mortality of krill, and including this factor in the models improves estimates of natural mortality, offering a more accurate assessment of krill population health, which is vital for sustainable management.
+Ecosystem-informed models, while providing more biologically accurate projections, tend to produce more conservative and less optimistic estimates compared to traditional models.
+This helps mitigate the risk of overestimations that could lead to unsustainable management decisions.
+However, these models also introduce challenges such as increased sensitivity and retrospective bias, particularly in fishing mortality estimates, underscoring the need for a balance between ecological realism and model stability to ensure reliable management decisions.
+Overall, these findings reinforce that ecosystem-based approaches provide a stronger scientific foundation for fishery management, but they must be carefully managed to avoid potential instability in model outputs.
 
-\end{landscape}
+\newpage
+
+# ACKNOWLEDGMENTS
+
+This research was supported by the following funding sources: the INACH "Marine Protected Areas" Program (Grant No.
+2409052), the ANID/Millennium Science Initiative Program (Grant No.
+ICN2021_002), the CCAMLR Scholarship Scheme (2023-2024), and the Doctorate Scholarship from CENTRO-IDEAL at the Universidad Austral de Chile.
+The authors extend their gratitude to the Secretariat of the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) for providing access to the krill fishery data, which was instrumental to this study.
+
+# SUPPLEMENTARY MATERIAL
+
+-   Supplementary Material 1: analisis, formulas, and outputs of stock assessment of Antarctic krill in WAP can be found in this [link](https://mauromardones.github.io/SA_Krill/).
+
+-   Supplementary Material 2: Correlation analysis between krill recruitment and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_recruit_env/)
 

### 2025-04-11 index.Rmd
friday!

@@
 library(ggridges)
+library(ggrepel)
 ```
@@
 
+The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale (Figure \@ref(fig:mapa)).
+To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021; REF] (Fig 1 S1).
+While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale.
+This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics.
+This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.
+
+```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
+knitr::include_graphics('Figs/481.png')
+```
+
+In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets.
+Differences among strata have been identified from various perspectives of krill population dynamics, as well as from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).
+
+```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
+knitr::include_graphics('Figs/conceptual.jpeg')
+```
+
 ## Statistical Model (SS3)
@@
 These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.
+
+```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
+knitr::include_graphics('Figs/pathmod.png')
+```
 
@@
 ```
-Source of data inpit
+Source of data input
 
@@
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(legend.position = "top",
+    axis.text.x = element_text(angle = 90, hjust = 1),
+         axis.text.y = element_blank()) +
   labs(title = "", x = "", y = "Standard index", color = "Category")
@@
 
-```{r }
-par(mfrow=c(1,2), mar=c(5,4,2,1))
+```{r fig.height=8, fig.width=6}
+par(mfrow=c(2,1), mar=c(5,4,2,1))
 SSplotData(base.model1.1, 
@@
 
-```{r fig.height=3, fig.width=7}
+```{r fig.height=2, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1)) 
 SSplotTimeseries(base.model1.1,
-                 subplot = 1)
+                 subplot = 1,
+                 xlab = "")
 SSplotTimeseries(base.model1.1,
-                 subplot = 9)
+                 subplot = 9,
+                 xlab = "")
 SSplotTimeseries(base.model1.1,
-                 subplot = 7)
+                 subplot = 7,
+                 xlab = "")
 ```
@@
 
-```{r fig.height=3, fig.width=7}
+```{r fig.height=2, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.2,
-                 subplot = 1)
+                 subplot = 1,
+                 xlab = "")
 SSplotTimeseries(base.model1.2,
-                 subplot = 9)
+                 subplot = 9,
+                 xlab = "")
 SSplotTimeseries(base.model1.2,
-                 subplot = 7)
+                 subplot = 7,
+                 xlab = "")
 ```
@@
 
-```{r fig.height=3, fig.width=7}
+```{r fig.height=2, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.3,
-                 subplot = 1)
+                 subplot = 1,
+                 xlab = "")
 SSplotTimeseries(base.model1.3,
-                 subplot = 9)
+                 subplot = 9,
+                 xlab = "")
 SSplotTimeseries(base.model1.3,
-                 subplot = 7)
+                 subplot = 7,
+                 xlab = "")
 ```
@@
 
-```{r fig.height=3, fig.width=7}
+```{r fig.height=2, fig.width=7}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
 SSplotTimeseries(base.model1.4,
-                 subplot = 1)
+                 subplot = 1,
+                 xlab = "")
 SSplotTimeseries(base.model1.4,
-                 subplot = 9)
+                 subplot = 9,
+                 xlab = "")
 SSplotTimeseries(base.model1.4,
-                 subplot = 7)
+                 subplot = 7,
+                 xlab = "")
 ```
@@
 SSplotSelex(base.model1.1,
-            subplots = 1)
+            subplots = 1,
+            mainTitle = FALSE)
 mtext("s1.1", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.2,
-            subplots = 1)
+            subplots = 1,
+            mainTitle = FALSE)
 mtext("s1.2", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.3,
-            subplots = 1)
+            subplots = 1,
+            mainTitle = FALSE)
 mtext("s1.3", side=3, line=0.5, cex=1.2)
 SSplotSelex(base.model1.4,
-            subplots = 1)
+            subplots = 1,
+            mainTitle = FALSE)
 mtext("s1.4", side=3, line=0.5, cex=1.2)
 ```
-```{r}
-SSplotBiology(base.model1.1,
-            subplots = 24)
-```
-
-Natural Mortality
-
 ```{r}
@@
 SSplotBiology(base.model1.1,
+            subplots = 1)
+SSplotBiology(base.model1.2,
+            subplots = 1)
+SSplotBiology(base.model1.3,
+            subplots = 1)
+SSplotBiology(base.model1.4,
+            subplots = 1)
+```
+
+Natural Mortality
+
+```{r eval=FALSE}
+par(mfrow=c(4,2), mar=c(5,4,2,1))
+SSplotBiology(base.model1.1,
               subplots = 21)
-mtext("s1.1", side=3, line=0.5, cex=1.2)
 SSplotBiology(base.model1.2,
@@
 ```
-
 Total biomass
@@
   geom_smooth(method = "loess",
-              span=0.5, 
-              se=F)+
+              span = 0.5,
+              se = FALSE) +
   labs(title = "",
@@
   theme_few() +
-  scale_color_viridis_d(option="H",
-                        name="Scenarios")+
+  scale_color_manual(name = "",
+                     values = c("Without Ecosystem Variables" = "#ca0020",  
+                                "W/ Predator" = "#f4a582",  
+                                "W/ Environment" = "#bababa",  
+                                "With Predator and Ecosystem" = "#404040")) +
   theme(axis.text.x = element_text(angle = 90, 
                                    hjust = 1))
+
 ```
@@
 ```{r}
-df <- base.model1.4$lendbase
-
+
+create_heatmap <- function(df) {
   df1 <- df %>%
-  dplyr::filter(Pearson<5) %>% 
+    dplyr::filter(Pearson < 5) %>%
     dplyr::select(c(1, 6, 19, 16)) %>%
@@
       Fleet == 11 ~ "PREDATOR",
-    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
+      TRUE ~ NA_character_
     ))
 
-# Crea el heatmap
   ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
     geom_tile() +
-  scale_fill_gradient2(low = "blue", 
+    scale_fill_gradient2(low = "black", 
                          mid = "white", 
@@
                          midpoint = 0) +
-  facet_wrap(Fleet~., ncol=5) + 
+    facet_wrap(Fleet~., ncol=6) + 
     labs(title = "", 
@@
          fill = "Pearson") + 
-  theme_few()+
+    theme_few() + 
     theme(axis.text.x = element_text(angle = 90, 
                                      hjust = 1, 
-                                   size = 7))
-
-```
-hexagon
+                                     size = 7),
+          legend.position = "bottom")
+}
+
+# Crear heatmaps para los diferentes dataframes
+heatmap1 <- create_heatmap(base.model1.1$lendbase)
+heatmap2 <- create_heatmap(base.model1.2$lendbase)
+heatmap3 <- create_heatmap(base.model1.3$lendbase)
+heatmap4 <- create_heatmap(base.model1.4$lendbase)
+
+# Combina los plots en un solo gráfico con 8 columnas y 1 fila, compartiendo la leyenda
+ggarrange(heatmap1, heatmap2, heatmap3, heatmap4, 
+          ncol = 2, 
+          nrow = 2, 
+          common.legend = TRUE, 
+          legend = "bottom")
+```
+
+This figure shows the Pearson residuals of predicted length distributions for krill across four modeling scenarios, each incorporating different levels of ecosystem complexity.
+The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population.
+In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
+Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
+The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals.
+However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
+In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
+This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
+Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure.
+The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
+
+Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the **joint inclusion of predator and environmental variables** yielding the most robust representation of observed length compositions.
 
@@
                plotdir = Figs)
-```
-
-```{r}
-par(mfrow=c(3,3), mar=c(5,4,2,1))
+par(mfrow=c(2,5), mar=c(5,4,2,1)) 
 SSplotRunstest(base.model1.2,
@@
                plotdir = Figs)
-```
-
-```{r}
 par(mfrow=c(2,4), mar=c(5,4,2,1)) 
@@
                plotdir = Figs)
-```
-
-```{r}
-par(mfrow=c(3,3), mar=c(5,4,2,1))
+par(mfrow=c(2,5), mar=c(5,4,2,1)) 
 SSplotRunstest(base.model1.4,
@@
 
-```{r eval=FALSE}
+```{r}
+par(mfrow=c(2,5), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.1,
@@
                add=T,
-               print = TRUE)
+               plot = TRUE,
+               plotdir = Figs)
+par(mfrow=c(2,5), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.2,
                subplots = "cpue",
-               add=T)
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
+par(mfrow=c(2,5), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.3,
                subplots = "cpue",
-               add=T)
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
+par(mfrow=c(2,5), mar=c(5,4,2,1))
 SSplotRunstest(base.model1.4,
                subplots = "cpue",
-               add=T)
+               add=T,
+               plot = TRUE,
+               plotdir = Figs)
 ```
@@
 
-```{r}
-# Crear un data frame con los valores de RMSE
+```{r fig.height=3, fig.width=3}
 dfrmse <- data.frame(
-  Model1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
-  Model1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
-  Model1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
-  Model1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])
+  s1.1 = as.numeric(base.model1.1$index_variance_tuning_check$RMSE),
+  s1.2 = as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10]),
+  s1.3 = as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10]),
+  s1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])
 
@@
 
-t.test(dfrmse$Model1.1, dfrmse$Model1.2)  # Comparar Model1.1 vs Model1.2
-t.test(dfrmse$Model1.1, dfrmse$Model1.3)  # Comparar Model1.1 vs Model1.3
-t.test(dfrmse$Model1.2, dfrmse$Model1.3)
-t.test(dfrmse$Model1.3, dfrmse$Model1.4)
-# ANOVA para ver si hay diferencias globales
+t.test(dfrmse$s1.1, dfrmse$s1.2)  # Comparar s1.1 vs s1.2
+t.test(dfrmse$s1.1, dfrmse$s1.3)  # Comparar s1.1 vs s1.3
+t.test(dfrmse$s1.2, dfrmse$s1.3)
+t.test(dfrmse$s1.3, dfrmse$s1.4)
 anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
-summary(anova_rmse)
-
-# Prueba post-hoc si ANOVA es significativa
-#TukeyHSD(anova_rmse)
 
 dfrmse_long <- reshape2::melt(dfrmse)
-ggplot(dfrmse_long, aes(x = variable, y = value, fill = variable)) +
-  geom_boxplot(width=0.3) +
+rmse <-ggplot(dfrmse_long, aes(x = variable, 
+                        y = value, 
+                        fill= variable)) +
+  geom_boxplot(width=0.2,
+               alpha=0.8) +
   geom_jitter(width = 0.01, alpha = 0.3) +
@@
   scale_fill_manual(name= "Scenario",
-                    values = c("Model1.1" = "#ca0020", 
-                               "Model1.2" = "#f4a582", 
-                               "Model1.3" = "#bababa", 
-                               "Model1.4" = "#404040")) +
-  labs(title = "", x = "", y = "RMSE")
-```
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  labs(title = "", x = "", y = "RMSE")+
+  theme(legend.position = "none")
+rmse
+```
+
 This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill.
@@
 ```{r}
-# First dataset
-l <- base.model1.1$recruit$SpawnBio  
-m <- base.model1.1$recruit$pred_recr 
-yr <- base.model1.1$timeseries$Yr[1:35]
-
-# Second dataset
-leco <- base.model1.4$recruit$SpawnBio  
-meco <- base.model1.4$recruit$pred_recr 
-yreco <- base.model1.4$timeseries$Yr[1:35]
-
-# Beverton-Holt models
-bev_holt_model1 <- nls(m ~ (a * l) / (1 + b * l), 
+# Lista de modelos base
+models <- list(
+  s1.1 = base.model1.1,
+  s1.2 = base.model1.2,
+  s1.3 = base.model1.3,
+  s1.4 = base.model1.4
+)
+
+colors <- c("s1.1" = "#ca0020", "s1.2" = "#f4a582", "s1.3" = "#bababa", "s1.4" = "#404040")
+shapes <- c("s1.1" = 5, "s1.2" = 6, "s1.3" = 15, "s1.4" = 4)
+
+plot_data <- list()
+curve_data <- list()
+label_data <- list()
+params <- data.frame(Model = character(), a = numeric(), b = numeric(), stringsAsFactors = FALSE)
+for (group in names(models)) {
+  model <- models[[group]]
+  l <- model$recruit$SpawnBio
+  m <- model$recruit$pred_recr
+  yr <- model$timeseries$Yr[1:35]
+
+  # Filtrar valores válidos
+  valid <- which(is.finite(l) & is.finite(m) & l > 0 & m > 0)
+  l <- l[valid]
+  m <- m[valid]
+  yr <- yr[valid]
+
+  plot_data[[group]] <- data.frame(l = l, m = m, yr = yr, Group = group)
+  label_data[[group]] <- plot_data[[group]] %>% filter(yr %% 3 == 0)
+
+  fit <- NULL
+  tryCatch({
+    fit <- nls(m ~ (a * l) / (1 + b * l),
                start = list(a = max(m), b = 1 / max(l)),
@@
 
-bev_holt_model2 <- nls(meco ~ (a * leco) / (1 + b * leco), 
-                        start = list(a = max(meco), b = 1 / max(leco)), 
-                        data = data.frame(leco, meco))
-
-# Predictions
-l_pred1 <- seq(min(l), max(l), length.out = 100)
-m_pred1 <- predict(bev_holt_model1, newdata = data.frame(l = l_pred1))
-
-l_pred2 <- seq(min(leco), max(leco), length.out = 100)
-m_pred2 <- predict(bev_holt_model2, newdata = data.frame(leco = l_pred2))
-
-# Create data frames for plotting
-data1 <- data.frame(l = l, m = m, yr = yr, Group = "s1.1")
-data2 <- data.frame(l = leco, m = meco, yr = yreco, Group = "s1.4")
-
-# Plot
-ggplot() +
-  geom_point(data = data1, aes(x = l, y = m, shape = Group),
-             size = 1.5) +
-  geom_text(data = data1, aes(x = l, y = m, label = yr), 
-            hjust = 0.5, vjust = -1.5, size = 3) +  
-  geom_line(data = data.frame(l = l_pred1, m = m_pred1), 
-            aes(x = l, y = m), linewidth = 1,
-            col="blue") +
-  geom_point(data = data2, aes(x = l, y = m, shape = Group), 
-             size = 1.5) +
-  geom_text(data = data2, aes(x = l, y = m, label = yr), 
-            hjust = 0.5, vjust = 1.5, size = 3) +
-  geom_line(data = data.frame(l = l_pred2, m = m_pred2), 
-            aes(x = l, y = m), linewidth = 1,
-            col="blue") +
-  scale_shape_manual(values = c("s1.1" = 5, 
-                                "s1.4" = 4)) +
-  theme_few() +
-  ylim(0, max(m, meco)) +
-  labs(x = "SSB", y = "Recruit", 
-       title = "", shape = "Model") +
+    l_pred <- seq(min(l), max(l), length.out = 100)
+    m_pred <- predict(fit, newdata = data.frame(l = l_pred))
+    curve_data[[group]] <- data.frame(l = l_pred, m = m_pred, Group = group)
+
+    coefs <- coef(fit)
+    params <- rbind(params, data.frame(Model = group, a = coefs["a"], b = coefs["b"]))
+
+  }, error = function(e) {
+    message(paste("Model", group, ": Beverton-Holt fit failed ->", e$message))
+  })
+}
+
+# Unir data frames
+plot_data_all <- bind_rows(plot_data)
+curve_data_all <- bind_rows(curve_data)
+label_data_all <- bind_rows(label_data)
+
+# Graficar
+srr_sce <- ggplot() +
+  geom_point(data = plot_data_all, aes(x = l, y = m, shape = Group), size = 2) +
+  geom_line(data = curve_data_all, aes(x = l, y = m, color = Group), linewidth = 1.2) +
+  ggrepel::geom_text_repel(data = label_data_all, aes(x = l, y = m, label = yr, color = Group),
+                           size = 4, max.overlaps = 50) +
+  scale_shape_manual(values = shapes) +
+  scale_color_manual(values = colors) +
+  theme_bw() +
+  ylim(0, max(plot_data_all$m)) +
+  labs(x = "Spawning Stock Biomass (SSB)", y = "Predicted Recruitment", shape = "Model", color = "Model") +
+  theme(legend.position = "bottom")
+```
+
+### Productivity and Interannual Variability by Scenario
+
+Estimating the productivity of Antarctic krill is critical for understanding the species’ capacity to replenish its population in response to varying levels of spawning biomass.
+Productivity, defined as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
+Comparing productivity across scenarios—each representing different assumptions about environmental drivers, fishing mortality, or predator dynamics—enables a robust evaluation of how krill populations reflect this changes.
+
+For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):
+
+$$
+\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
+$$
+
+We also calculated the **interannual percentage change** in recruitment and SSB as:
+
+$$
+\text{Change in Recruitment}_{i,t} = \left( \frac{\text{Recruitment}_{i,t} - \text{Recruitment}_{i,t-1}}{\text{Recruitment}_{i,t-1}} \right) \times 100
+$$
+
+$$
+\text{Change in SSB}_{i,t} = \left( \frac{\text{SSB}_{i,t} - \text{SSB}_{i,t-1}}{\text{SSB}_{i,t-1}} \right) \times 100
+$$
+
+These metrics allow us to analyze both the productivity and the temporal dynamics of the population under each scenario \( i \).
+
+```{r}
+# Simular 4 dataframes por escenario (usa tus datos reales)
+escenarios <- list(
+  s1.1 = base.model1.1$SPAWN_RECR_CURVE,
+  s1.2 = base.model1.2$SPAWN_RECR_CURVE,
+  s1.3 = base.model1.3$SPAWN_RECR_CURVE,
+  s1.4 = base.model1.4$SPAWN_RECR_CURVE
+)
+
+# Crea una tabla con productividad y cambios para cada escenario
+resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
+  df %>%
+    mutate(
+      Scenario = nombre,
+      Productividad = Recruitment / SSB,
+      Cambio_Recruit = c(NA, diff(Recruitment)) / lag(Recruitment) * 100,
+      Cambio_SSB = c(NA, diff(SSB)) / lag(SSB) * 100
+    )
+})
+
+prod <- ggplot(resultados, aes(x = `SSB/SSB_virgin`, y = Productividad, 
+                       color = Scenario)) +
+  geom_point(size = 1.1) +
+  labs(title = "",
+       y = "Recruitment / SSB",
+       x = "SSB / SSB_virgin") +
+  scale_color_manual(name= "",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```
+
+These two panels provide insight into the stock-recruitment dynamics of Antarctic krill under four assessment scenarios, each incorporating different levels of ecosystem complexity.
+
+**Left panel – Predicted Recruitment vs Spawning Stock Biomass (SSB):**  
+This plot compares the predicted recruitment across years with corresponding spawning biomass values for the four model scenarios (s1.1 to s1.4).
+Scenario s1.1 (red) shows a relatively flat and optimistic Beverton-Holt relationship, with high recruitment predicted even at low levels of SSB.
+The other scenarios, particularly s1.3 (gray) and s1.4 (black), show much more constrained recruitment estimates and a more saturating relationship, where recruitment increases only slightly with increasing SSB.
+This suggests that incorporating environmental (s1.3) and predator (s1.4) effects leads to a more conservative and ecologically realistic recruitment dynamic.
+Year labels illustrate how certain years (e.g., 2004, 2016, 2022) deviate between scenarios, especially at low biomass levels.
+
+**Right panel – Recruitment per SSB vs Relative SSB (SSB / SSBo):**  
+This plot displays recruitment efficiency (Recruitment/SSB) as a function of relative spawning biomass (SSB / SSB₀).
+In all models, recruitment efficiency is highest at very low biomass levels and declines as SSB increases, consistent with compensatory dynamics.
+However, s1.3 and s1.4 (which include environmental effects) show significantly lower overall productivity across all SSB levels compared to s1.1 and s1.2.
+This indicates that the inclusion of ecosystem drivers reduces the per-capita productivity of the stock, especially under depleted conditions.
+
+**Interpretation:**  
+Together, these plots indicate that ecosystem-informed models (s1.3 and s1.4) produce more constrained and precautionary estimates of krill productivity.
+The more optimistic recruitment predicted by s1.1 (and to a lesser extent s1.2) may overestimate stock resilience, particularly in low SSB conditions.
+These findings emphasize the importance of accounting for environmental variability and predator dynamics in stock assessment models to avoid overoptimistic management advice and better reflect the biological limits of the krill population.
+
+```{r, fig.height=4, fig.width=6, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios.
+Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
+Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
+ggarrange(srr_sce,
+          prod,
+          ncol = 2)
+```
+
+### Compara Fishing Mortality.
+
+```{r eval=FALSE}
+escenarios <- list(
+  s1.1 = base.model1.1$exploitation[1:5],
+  s1.2 = base.model1.2$exploitation[1:5],
+  s1.3 = base.model1.3$exploitation[1:5],
+  s1.4 = base.model1.4$exploitation[1:5]
+)
+
+# Calcula el cambio porcentual entre años para F_std y annual_F
+# Calcula el cambio relativo entre años para F_std y annual_F
+# Calcula el cambio relativo entre años para F_std y annual_F
+resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
+  df %>%
+    mutate(
+      Scenario = nombre,
+      Relativo_Fstd = c(NA, diff(F_std) / head(F_std, -1)),  # Cambio relativo de F_std
+      Relativo_annual_F = c(NA, diff(annual_F) / head(annual_F, -1))  # Cambio relativo de annual_F
+    )
+})
+
+# Visualiza los resultados en un gráfico
+ggplot(resultados, aes(x = Yr, y = Relativo_Fstd, color = Scenario)) +
+  geom_line(size = 1.1) +
+  labs(title = "Cambio porcentual en F_std entre años por escenario",
+       x = "Año",
+       y = "Cambio porcentual en F_std") +
+  scale_color_manual(name= "Scenario",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  theme_minimal(base_size = 13) +
+  theme(legend.position = "bottom")
+
+# Gráfico para annual_F
+ggplot(resultados, aes(x = Yr, y = Relativo_annual_F, color = Scenario)) +
+  geom_line(size = 1.1) +
+  labs(title = "Cambio porcentual en annual_F entre años por escenario",
+       x = "Año",
+       y = "Cambio porcentual en annual_F") +
+  scale_color_manual(name= "Scenario",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  theme_minimal(base_size = 13) +
   theme(legend.position = "bottom")
@@
 ```{r}
-
 par(mfrow=c(2,2), mar=c(5,4,1,1)) 
-
 retro1.1 <- SSplotRetro(retroSummary1.1,
@@
 ```
-```{r}
-
+
+```{r}
 par(mfrow=c(2,2), mar=c(5,4,1,1)) 
-
 retro1.1 <- SSplotRetro(retroSummary1.1,
@@
 ```{r eval=FALSE}
-
 nanios<-base.model1.1$Kobe$Yr
@@
 kobe1
-
 ```
@@
 
-```{r}
+```{r, fig.height=4, fig.width=6}
 tablebias_combined <- tablebias_combined %>%
@@
 ```{r}
-
 like1 <- base.model1.1$likelihoods_used
@@
                    like4)
-
-```
-
-```{r}
-ggplot(totalike)+
+```
+
+```{r}
+ggplot(totalike %>% 
+         filter(values>1))+
   geom_bar( aes(y = reorder(Description, values), 
@@
           p3,
-          p4, ncol=2, nrow = 2,
+          p4, ncol=1, nrow = 4,
           common.legend = T)
@@
 This suggests that s1.4 provides the best overall model fit, while s1.1 (dark red) has the lowest likelihood, indicating a poorer fit.
+
 **Survey Component:** The survey likelihood follows a similar pattern, with s1.4 and s1.2 showing better agreement with the data compared to s1.1.
+
 **Recruitment, Parameter Priors, and Parameter Deviations:** These components contribute minimally to the total likelihood, indicating that they do not heavily influence model fit differences.
+
 **Length Composition:** There is some variation in this component, with s1.2 showing a larger contribution than s1.1 and s1.3.
 
-### **Overall Conclusion:**  
 Scenario s1.4 seems to fit the data best, followed closely by s1.2.
@@
 ```
+
 ### AKL

### 2025-04-14 MS2_Krill.Rmd
adding conceptual mapa to MJ

@@
 
-Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under four modeling scenarios from 1990 to 2020.
-The reference model (blue), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015.
+Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under most significant models in this analisys, s1.1 and s1.4.
+The reference model (red), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015.
 In contrast, the models that incorporate ecosystem components—predator data (green), environmental data (yellow), or both (red)—estimate consistently lower spawning output throughout the time series.
@@
 
-```{r compara,eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
-knitr::include_graphics('Figs/COM1compare2_spawnbio_uncertainty.png')
+```{r compara,eval=TRUE, out.width="60%", fig.cap=""}
+knitr::include_graphics('Figs/alloutput_s1_4.png')
 
@@
 
-```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated spawning output for Antarctic krill under four assessment scenarios"}
-knitr::include_graphics('Figs/COM1compare16_densities_SR_LN(R0).png')
+```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
+knitr::include_graphics('Figs/R0plot.png')
 

### 2025-04-14 index.Rmd
adding conceptual mapa to MJ

@@
 
-```{r mapa, out.width='40%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
+```{r mapa, out.width='50%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
 knitr::include_graphics('Figs/481.png')
@@
 
-In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets.
-Differences among strata have been identified from various perspectives of krill population dynamics, as well as from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).
+In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014].
+This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).
 
@@
 
-## Statistical Model (SS3)
+## Statistical Model (`SS3`)
 
@@
 ```
-Source of data input
-
-```{r fig.width=9, fig.height=4}
+Source of data input.
+
+Ibdex abundance in Figure \@ref(fig:index)
+
+```{r}
 dat1_std <- dat1$CPUE %>%
@@
   )
+```
+
+```{r index, fig.width=9, fig.height=4, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1.
+Each panel represents a distinct spatial or functional stratum, with trend lines indicating temporal variation from 1990 to 2020.
+Colors denote data source categories: green for fishery, orange for scientific surveys, and purple for predator-based indices.
+These patterns highlight spatial and temporal heterogeneity in krill dynamics across the subarea."}
 
@@
 
-```{r fig.width=10, fig.height=4, warning=FALSE}
+```{r warning=FALSE}
 dat_long <- dat1$lencomp %>% 
@@
   uncount(weights = Frecuencia)
-
+```
+
+Length compositions in Figure \@ref(fig:length)
+
+```{r length, fig.width=10, fig.height=4, warning=FALSE, fig.cap="Annual length-frequency distributions of Antarctic krill (Euphausia superba) across different data sources and spatial strata within Subarea 48.1 from 1991 to 2020.
+Each panel represents a distinct stratum for either fishery-dependent (green), fishery-independent survey (orange), or predator-based (purple) observations.
+Density ridgelines illustrate variation in krill size structure across years.
+The red vertical line marks a recruit references length (3.6 cm)."}
 p2 <- ggplot(dat_expanded, aes(x = Talla, y = as.factor(Yr), fill = FltSvy)) +
@@
 # Crear una columna de modelo en cada data frame
-outps1$Model <- "Ref Model: No Env-Predator"
+outps1$Model <- "s1.1 (Ref Model: No Env-Predator)"
 outps2$Model <- "s1.2 w/ Predator data"
 outps3$Model <- "s1.3 w/ Env data"
-outps4$Model <- "s1.4 w/ Env and Predator data"
+outps4$Model <- "s1.4 (Env and Predator data)"
 
@@
   pivot_longer(cols = c(Bio_smry, SpawnBio, Recruit_0), 
-               names_to = "Variable", values_to = "Value")
+               names_to = "Variable", values_to = "Value") %>% 
+  filter(Yr >= 1989, Yr <= 2020)
 
 # Gráfico de barras agrupadas por variable y modelo
-ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
+alloutput <- ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
   geom_bar(stat = "identity", position = "dodge") +
@@
   scale_fill_manual(name= "Scenario",
-                    values = c("Ref Model: No Env-Predator" = "#ca0020",
-
-                               "s1.4 w/ Env and Predator data" = "#404040")) +
+                    values = c("s1.1 (Ref Model: No Env-Predator)" = "#ca0020",
+
+                               "s1.4 (Env and Predator data)" = "#404040")) +
   labs(title = "",
        x = "", y = "") +
-  theme_bw() +
+  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1),
         legend.position = "bottom")
+alloutput
 ```
@@
 SSplotComparisons(summaryoutput1_4,
-                  legendlabels = c("Ref Model: No Env-Predator",
-                 "S1.1 w/ Predator data",
-                 "S1.1 w/ Env data",
-                 "S1.1 w/ Env and Predator data"),
+                  legendlabels = c("s1.1 (Ref Model)",
+                 "s1.2",
+                 "s1.3",
+                 "s1.4"),
                  filenameprefix = "COM1",
@@
 
+Antoher comparision in `R0` (Natural log of virgin recruitment level)
+
+```{r}
+r0_vals <- data.frame(
+  model = c("s1.1", "s1.2", "s1.3", "s1.4"),
+  value = c(base.model1.1$jitter_info[1,1],
+            base.model1.2$jitter_info[2,1],
+            base.model1.3$jitter_info[1,1],
+            base.model1.4$jitter_info[2,1]),
+  cv = c(base.model1.1$jitter_info[1,6],
+         base.model1.2$jitter_info[2,6],
+         base.model1.3$jitter_info[1,6],
+         base.model1.4$jitter_info[2,6])
+)
+
+x_vals <- seq(10, 30, length.out = 1000)
+
+dens_data <- r0_vals %>%
+  rowwise() %>%
+  mutate(sd =  cv,  
+         density = list(data.frame(
+           x = x_vals,
+           y = dnorm(x_vals, mean = value, sd = sd),
+           model = model
+         ))) %>%
+  select(density) %>%
+  unnest(cols = c(density))
+
+colores <- c("s1.1" = "#ca0020", 
+             "s1.2" = "#f4a582", 
+             "s1.3" = "#bababa", 
+             "s1.4" = "#404040")
+
+r0plot <- ggplot(dens_data, aes(x = x, y = y, color = model, fill = model)) +
+  geom_area(alpha = 0.5, position = "identity") +
+  geom_line(size = 1) +
+  scale_color_manual(values = colores) +
+  scale_fill_manual(values = colores) +
+  labs(title = "",
+       x = "Natural log of virgin recruitment level (R0)", y = "Density") +
+  theme_bw() +
+  xlim(16, 22)
+r0plot
+```
+
 ### Comparsion in sd long term time series forecasting

### 2025-04-14 MS2_Krill.Rmd
adding numbers section off

@@
 
+We comparethe time series of differente population variables for Antarctic krill under most significant models in this analysis; a reference model without environmental or predator data and an alternative model that incorporates both environmental and predator information.
+Total biomass estimates remain relatively stable across both scenarios until around 2010.
+From that point onward, both models show a marked increase in biomass, although the reference model predicts higher values than the ecosystem-informed model, particularly during the final assessment years (2018–2020).
+Recruitment at age 0 exhibits strong variability over time.
+This is coincidence with [REF].
+While both models show similar patterns, the reference model tends to estimate higher recruitment peaks, especially in recent years.
+Finally, spawning biomass (SpawnBio) follows trends similar to total biomass.
+Both models show a gradual increase after 2010, but the reference model again yields higher values in recent years.
+
 ```{r compara,eval=TRUE, out.width="60%", fig.cap=""}
@@
 
+Population variables: Overall, the inclusion of ecosystem and predator data (s1.4) results in more conservative estimates of biomass, recruitment, and spawning stock, suggesting that ecosystem-informed models may provide a more precautionary and ecologically realistic basis for krill stock assessment and management.
+
 ### Ecosystem Variables: Ecological Relevance

### 2025-04-14 index.Rmd
adding numbers section off

@@
 output:
-  html_document:
+  bookdown::html_document2:
     keep_md: true
@@
     code-tools: true
+    number_sections: false
 editor_options: 
@@
 
-```{r mapa, out.width='50%', fig.show='hold',fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
+```{r mapa, out.width='50%', fig.cap="Subarea 48.1 and management strata considered in the spatio-temporal analysis of intrinsic productivity of Krill (BS=Brainsfield Strait, EI= Elephant Island, Gerlache= Gerlache strait, JOIN= Joinville Island, SSWI= South West)"}
 knitr::include_graphics('Figs/481.png')
@@
 
-```{r fig.height=8, fig.width=6}
+```{r fig.height=10, fig.width=6}
 par(mfrow=c(2,1), mar=c(5,4,2,1))
@@
 
+### Code Repository
+
+The repository with code to replicate this analysis can be found in this [GitHub author link](https://github.com/MauroMardones/SA_Krill)
+
 \newpage

### 2025-04-15 MS2_Krill.Rmd
update scenarios

@@
 
-Multiple sources of data to inform the population dynamics model are included in this analysis.
-These data sources include data from the monitoring program of the krill fishery were used, as standar CPUE (kg/time hauls) and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program.
-This database has information about vessel, nationality, georeferenced, among others.
-Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including biomass index by strata, and biological characteristics of krill collected through net sampling.
-All this data was processed in @Dornam2021.
-Krill length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl.
-Mat.
-2).
+Multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation.
+These data include records from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been consistently collected on board fishing vessels by scientific observers through the CCAMLR Scheme of International Scientific Observation (SISO).
+This dataset contains detailed information including vessel identity, nationality, and georeferenced positions used in the exploratory analysis.
+In addition, scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling.
+Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were formatted according to model evaluation template requirements (see Supplementary Material 2).
+Environmental and predator-related data were also incorporated, such as satellite-derived chlorophyll-a concentrations from the Copernicus Marine Service and predator abundance and diet composition, particularly from penguin colonies reported in the [MAPPPD database](https://www.penguinmap.com/mapppd/sources/).
+A summary of all data sources, including temporal coverage and spatial scale, is provided in Table \@ref(tab:data-sources).
 
-```{r eval=TRUE}
+```{r data-sources, eval=TRUE}
 data <- data.frame(
-  Source = c("Krill Fishery Monitoring Program",
-                 "Krill Fishery Monitoring Program, collected by scientific observers under CCAMLR SISO program (vessel, nationality, georeferenced)",
-                 "Krill Fishery Monitoring Program",
-                 "AMLR Research Cruises Program", 
-                  "Satellite data",
-                 "Predator Abundance",
-                 "AMLR Research Cruises Program"),
-  Description = c("Catches",
-                  "CPUE (standarized by vessel, area and deep)", 
-                  "Krill Length compositions (grouped by 0.2 cm)",
-                  "Biomass index from scientific AMLR cruises",
-                  "Satellite-based chlorophyll-a concentration (monthly, interpolated) from the EU Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).",
-                  "Data on penguin colonies, sourced from 'MAPPPD'",
-                  "Predator Length composition krill diet"),
-    TimeSeries = c("1997-2020",
-                 "1997-2020",
-                 "2000-2020",
-                 "1991-2011 (based on the research cruise periods).",
-                 "Monthly data from 1997 to present (based on satellite data availability).",
-                 "Data from 2000s-present.",
-                 "1977-2020 (from 2000 used in this analysis)"),
-   Scale = c("by Strata",
-                           "by Strata",
-                           "by Strata",
-                           "by Strata",
-                           "48.1",
-                           "by colony",
-                           "by colony"))
-ft <- flextable(data) %>%  
-  set_caption("Comparison of Model Parameters") %>%  
+  Source = c(
+    "Krill Fishery Monitoring Program (CCAMLR SISO)",
+    "Krill Fishery Monitoring Program (CCAMLR SISO)",
+    "Krill Fishery Monitoring Program (CCAMLR SISO)",
+    "AMLR Scientific Research Cruises",
+    "Satellite Remote Sensing",
+    "Predator Monitoring (MAPPPD)",
+    "Predator Diet Composition (AMLR)"
+  ),
+  Description = c(
+    "Total reported catches",
+    "Standardized CPUE (by vessel, area, and depth)",
+    "Length-frequency distributions (0.2 cm bins)",
+    "Biomass indices by strata from net sampling",
+    "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
+    "Abundance and colony-specific trends of penguin populations",
+    "Length-frequency of krill in predator diets (e.g., penguins)"
+  ),
+  TimeSeries = c(
+    "1997–2020",
+    "1997–2020",
+    "2000–2020",
+    "1991–2011 (based on cruise periods)",
+    "1997–present",
+    "2000s–present",
+    "1977–2020 (subset from 2000 used in this analysis)"
+  ),
+  SpatialScale = c(
+    "By strata",
+    "By strata",
+    "By strata",
+    "By strata",
+    "Subarea 48.1",
+    "By colony",
+    "By colony"
+  )
+)
+
+ft_data <- flextable(data) %>%
+  set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
   autofit() %>%
   theme_vanilla() %>%
-  fontsize(size = 9, 
-           part = "all") %>%  
-  font(fontname = "Arial", 
-       part = "all") %>%  
+  fontsize(size = 9, part = "all") %>%
+  font(fontname = "Arial", part = "all") %>%
   set_table_properties(layout = "autofit", width = .85)
-ft
+
+ft_data
+
 ```
+
 Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
@@
 To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables previously recognized as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
-SST data were obtained from ERA5 monthly means (1991–present), SIC data from Nimbus-7 (1978–present), and Chla data from the Bio-Geo-Chemical L4 satellite product, downloaded from the E.U.
-Copernicus Marine Service (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148).
+Table \@ref(tab:env) summarizes the key characteristics of the environmental datasets used in the analysis, including sea ice, sea surface temperature (SST), and chlorophyll a.
+These variables were obtained from different satellite platforms and reanalysis products, namely Nimbus-7, ERA5, and Copernicus.
+The table details spatial and temporal resolution, as well as the temporal coverage of each dataset.
+
+```{r env, eval=TRUE}
+env_features <- data.frame(
+  Feature = c("Sea Ice", "SST", "Chlorophyll a"),
+  Platform = c("Nimbus-7", "ERA5", "E.U. Copernicus Marine Service"),
+  Spatial_Resolution = c("25 × 25 km", "0.25º × 0.25º", "0.25º × 0.25º"),
+  Temporal_Resolution = c("Daily", "Monthly", "Daily-Monthly"),
+  Time_Range = c("1978 - Present", "1991 - Present", "2000 - Present"),
+  stringsAsFactors = FALSE
+)
+
+ft_env <- flextable(env_features) |>
+  set_header_labels(
+    Feature = "Feature",
+    Platform = "Platform",
+    Spatial_Resolution = "Spatial Resolution",
+    Temporal_Resolution = "Temporal Resolution",
+    Time_Range = "Time Range"
+  ) |>
+  set_caption("Summary of environmental variables used in correlation analysis") |>
+  add_footer_lines("doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148") |>
+  color(part = "header", color = "black") |>
+  theme_booktabs() |>
+  fontsize(size = 10) |>
+  set_table_properties(layout = "autofit", width = .85)
+ft_env
+```
+
 All environmental data were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations.
@@
 
-As @methot2020stock establish in SS3 manual, natural mortality attributable to predation (M2) can be incorporated through different types of observational data, each informing distinct aspects of predator-prey dynamics.
-First, total mortality due to predation can be input as a "discard" component in the data structure.
-Since M2 represents a proportion of total mortality (Z), the quantity of individuals removed by predators can be estimated using the relative contribution of M2 to Z, i.e., `M2 / Z`.
-This allows the model to estimate the biomass consumed by predators, analogous to how catch is calculated for fishing fleets.
-Second, predator abundance can be input as an index of effort, treating M2 as a fishing mortality analog.
-This approach uses the same structure as a fishery-independent survey, where predator biomass serves as an index of predation pressure over time.
-In this framework, the predation rate is proportional to predator abundance scaled by a catchability coefficient \( Q \), following the equation:
+Following the SS3 framework described by @methot2020stock, natural mortality due to predation (*M2*) can be integrated into stock assessment models through multiple data types.
+In our krill assessment, we implement this by incorporating both predator biomass and diet composition from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*).
+Predator abundance is treated as an index of predation "effort," allowing time-varying estimation of predation mortality rates.
+This follows the equation:
 
 $$
-M2_t = Q * I_t
+M2_t = Q \cdot I_t
 $$
-where \( M2_t \) is the predation mortality in year \( t \), \( I_t \) is the predator abundance index in year \( t \), and \( Q \) is an estimated or fixed catchability parameter linking the predator index to predation pressure.
-Finally, information on the size or age composition of predated individuals, such as those derived from gut content analyses, can be used to inform the selectivity pattern of M2.
-These data are input analogously to length- or age-composition observations from fisheries, and allow the model to estimate the predation selectivity curve, capturing how predation probability varies with prey size or age.
 
-Regarding krill predators, we will incorporate both biomass and dietary data from the three main penguin colonies —Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*) —into the stock assessment model through two complementary pathways..
-First, the abundance of these predator colonies will be treated as an index of predation "effort" exerted on the krill population, allowing estimation of time-varying predation mortality rates.
-Second, empirical length-frequency distributions of krill derived from the diet composition of each penguin species will be used to parameterize the size-selective nature of M2, enabling the estimation of predator-specific selectivity functions.
-This dual approach facilitates a temporally and size-structured characterization of M2, consistent with ecosystem-based stock assessment principles and improving the ecological realism of krill population dynamics modeling.
-The total natural mortality (M) will thus explicitly include the predation component, expressed as \( M = M_1 + \sum M_2 \) [@methot2020stock], where \( M_1 \) represents background (non-predation) natural mortality, and \( \sum M_2 \) accounts for mortality attributed to each predator type.
-Consequently, total mortality is represented as \( Z = M_1 + M_2 + F \), incorporating both natural and fishing sources of mortality.
+where \( M2_t \) is the predation mortality in year *t*, \( I_t \) is the predator abundance index, and \( Q \) is a catchability coefficient.
+Additionally, length-frequency data of krill obtained from penguin diet samples are used to characterize the size-selectivity of predation.
+These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size.
+This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.
+Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions.
+Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality.
 
@@
 `s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers affecting krill recruitment and biomass.
-Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more comprehensive assessment of ecosystem interactions and their role in krill population dynamics (Table 1).
+Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more comprehensive assessment of ecosystem interactions and their role in krill population dynamics (Table \@ref(tab:scenario)).
 
-| Scenario | Description                                                |
-|:--------:|:-----------------------------------------------------------|
-|   s1.1   | Spatial data without environmental and predator components |
-|   s1.2   | "s1.1" with predator components                            |
-|   s1.3   | "s1.1" with environmental variable                         |
-|   s1.4   | "s1.1" w/ both, predator fleet and environmental variable  |
+```{r scenario, eval=TRUE}
+scenario_table <- data.frame(
+  Scenario = c("s1.1", "s1.2", "s1.3", "s1.4"),
+  Description = c(
+    "Spatial data without environmental and predator components",
+    "\"s1.1\" with predator components",
+    "\"s1.1\" with environmental variable",
+    "\"s1.1\" with both predator fleet and environmental variable"
+  ),
+  stringsAsFactors = FALSE
+)
+ft_scenarios <- flextable(scenario_table) |>
+  set_header_labels(
+    Scenario = "Scenario",
+    Description = "Description"
+  ) |>
+  set_caption("Scenarios used for modelling dynamics in krill") |>
+  theme_booktabs() |>
+   fontsize(size = 10) |>
+  set_table_properties(layout = "autofit", width = .85) %>% 
+  color(part = "header", color = "black") 
+ft_scenarios
+```
 
-: Scenarios to modelling dynamics in krill
-
-Regarding the performance and comparison between scenarios, @Carvalho2017 and @Carvalho2021b propose a structured approach for diagnosing stock assessment models to ensure their reliability and robustness.
-This diagnostic process was applied to the krill stock assessment and involved some key steps.
+Regarding the performance and comparison between scenarios, @Carvalho2017;  @Carvalho2021b; @Merino2022a and @Kell2021 propose good practice to made a structured approach for diagnosing stock assessment models to ensure their reliability and robustness.
+This diagnostic process was applied to the krill stock assessment and involved several key steps propoesed by previus research.
 First, convergence analysis was conducted to evaluate whether the estimation algorithm reached a stable solution.
@@
 Next, model fit to data was assessed by comparing observed and predicted values across key data sources, including catch, survey indices, and length composition.
-Residual plots were examined to detect potential patterns indicative of model misspecification.
-A retrospective analysis was performed by sequentially removing recent years of data (peeling) and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
-The model's forecasting capability was evaluated through one-step-ahead predictions and their comparison to observed values in recent years, helping identify potential bias or overfitting in recruitment or abundance estimates.
-Sensitivity analyses were conducted by varying ecosystem variables to assess their influence on model outputs.
+The p-values for the residual runs tests were computed for each of the four joint length compositions that are common in all models.
+The p-values were then combined into a single test statistic using Fisher's method @@Merino2022a:
+
+$$ 
+\chi^2_{2k} = -2 \sum \log(p_i)
+$$
+
+where \( p_i \) is the p-value for Length \( i \) and \( k \) are the degrees of freedom of the four p-values from the joint Length indices.
 Likelihood profiling was also carried out to quantify parameter uncertainty and identify the influence of priors.
-Finally, prediction skill was assessed using hindcasting techniques to evaluate the model’s ability to reproduce historical stock dynamics.Four estimation models were developed depending on how the ecosystem components was integrated where it was included as an additional covariate within the stock-recruitment (S-R) relationship.
-To evaluate statistical differences in model performance, we analyzed the Root Mean Squared Error (RMSE) values across four scenario configurations (`s1.1`, `s1.2`, `s1.3`, and `s1.4`).
-The RMSE values were extracted from the variance tuning checks of each scenario and compiled into a data frame for statistical analysis.
+Finally, prediction skill was assessed using hindcasting techniques to evaluate the model’s ability to reproduce historical stock dynamics.
+Retrospective analysis and hindcast cross-validation were performed by sequentially removing five years of data.
+regarding @HurtadoF2015, Mohn parametrer was calculated as the mean relative error, defined as:
+
+$$ 
+\rho_M = \frac{1}{h} \sum_{y=1}^{h} \left( \frac{X_{T - y} - \hat{X}_{T - y}}{\hat{X}_{T - y}} \right)
+$$
+
+where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which ρM was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset.
+Year \( y \) was nominally considered as season one, \( T \) refers to the terminal assessment data year, and \( h \) denotes the number of retrospective peels, removing recent years of data ("peeling") and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
+To evaluate statistical differences in model performance, we analyzed Root Mean Squared Error (RMSE) values across four scenario configurations (`s1.1`, `s1.2`, `s1.3`, and `s1.4`).
+The RMSE values were extracted from the variance tuning diagnostics of each scenario and compiled into a data frame for statistical analysis.
 To test for overall differences among scenarios, a one-way ANOVA [@REF] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
-If the ANOVA indicated significant differences, a post-hoc Tukey test would be performed to determine which scenario comparisons were driving the differences.
-By systematically applying these diagnostic steps, stock assessment models can be rigorously evaluated, ensuring they provide robust and unbiased estimates.
+AIC
+
+Table \@ref(tab:metrics) summarizes the diagnostic metrics used to evaluate model performance, including goodness-of-fit, prediction accuracy, information criteria, retrospective bias, and estimation stability.
+
+```{r metrics, eval=TRUE}
+df_model_metrics <- data.frame(
+  Metric = c(
+    "Pearson Residuals",
+    "Root Mean Square Error (RMSE)",
+    "Mean Absolute Scaled Error (MASE)",
+    "Akaike Information Criterion (AIC)",
+    "Rho Parameter (B and SSB)",
+    "R0 Deviation"
+  ),
+  Description = c(
+    "Standardized residuals calculated as the difference between observed and expected values, divided by the standard deviation of the expected values.
+Useful for assessing model fit across data points.",
+    "Measures the average magnitude of the residuals (prediction errors), giving higher weight to large errors. Lower values indicate better predictive accuracy.",
+    "A scale-independent measure of prediction accuracy. It compares forecast errors to the errors from a naive baseline model. Values <1 indicate better-than-baseline performance.",
+    "A relative measure of model quality that balances model fit and complexity. Lower AIC values indicate better models, penalizing overfitting.",
+    "The autoregressive correlation coefficient from retrospective analysis, used to assess systematic bias in biomass (B) and spawning stock biomass (SSB) estimates across years.",
+    "Measures the variability or uncertainty in the estimated unfished recruitment (R₀). High deviations may indicate poor estimation stability or model overparameterization."
+  ),
+  Reference = c(
+    NA,
+    "@Carvalho2021b",
+    "@Hyndman2006",
+    "@Akaike1974",
+    "@Mohn1999",
+    "@Methot2013"
+  ),
+  stringsAsFactors = FALSE
+)
+
+ft_stat <- flextable(df_model_metrics) |> 
+  set_header_labels(
+    Metric = "Metric",
+    Description = "Description",
+    Reference = "Reference"
+  ) |> 
+  set_caption("Summary of model diagnostic metrics used for krill stock assessment") |> 
+  autofit() |> 
+  fontsize(size = 10) |> 
+  theme_booktabs() |> 
+  color(part = "header", color = "black")|>
+  set_table_properties(layout = "autofit", width = .85)
+ft_stat
+```
 
@@
 
-Log-likelihood profiles were examined for each model scenario (s1.1 to s1.4) to assess how different components contributed to the overall model fit (Figure \@ref(fig:liketotal).
+Log-likelihood profiles were examined for each model scenario (s1.1 to s1.4) to assess how different components contributed to the overall model fit.
 Across all scenarios, the TOTAL log-likelihood is primarily influenced by the Survey, Length composition, and Recruitment components, with minor contributions from parameter priors.
@@
 These variations may reflect differences in model assumptions or input data between scenarios, which could affect how well each model fits particular data sources.
-
-```{r liketotal, eval=TRUE, fig.cap="Log-likelihood contributions for four model components across four krill population model scenarios.
-Bars show the contribution of each likelihood component to the total model fit"}
-knitr::include_graphics('Figs/likesel.png')
-```
-
 Table \@ref(tab:tablike) shows the log-likelihood component values for each scenario.
-It summarizes the contributions of different data sources—such as survey indices, length composition, recruitment, and parameter priors—to the total likelihood in each model configuration (s1.1 to s1.4).
-This comparison helps assess how each scenario fits the data and highlights potential differences in the influence of specific components across models.
 
@@
 ```{r rmseplot, eval=TRUE, fig.cap="Root Mean Square Error (RMSE) distributions for four krill population model scenarios"}
-knitr::include_graphics('Figs/unnamed-chunk-29-1.jpeg')
+knitr::include_graphics('Figs/diags2.png')
 ```
 
-FINAL TABLE Suggested by MJ
+(FINAL TABLE Suggested by MJ)
 
@@
 
-```{r popvar, eval=TRUE, fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
-Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)."}
-knitr::include_graphics('Figs/unnamed-chunk-65-1.jpeg')
-```
-
-Figure \@ref(fig:compara) shows the time series of estimated spawning output for Antarctic krill under most significant models in this analisys, s1.1 and s1.4.
-The reference model (red), which excludes ecosystem components, consistently estimates higher spawning output values over time, with a steep increase observed after 2015.
-In contrast, the models that incorporate ecosystem components—predator data (green), environmental data (yellow), or both (red)—estimate consistently lower spawning output throughout the time series.
-The red line, representing the model with both predator and environmental data, remains the most conservative, suggesting that ecosystem-informed models lead to more precautionary estimates of reproductive potential.
-The rigth panel displays the posterior density distributions of unfished spawning biomass (SSB~o~) for the same scenarios.
-The reference model estimates the highest SSBo values (mode ~ 3.2), while the ecosystem-informed models estimate considerably lower SSBo, with the most constrained distribution occurring under the full ecosystem model (red).
-The shift in the distributions indicates that including ecosystem drivers reduces estimates of unfished reproductive capacity, which has implications for reference point calculations and harvest strategies.
-
-We comparethe time series of differente population variables for Antarctic krill under most significant models in this analysis; a reference model without environmental or predator data and an alternative model that incorporates both environmental and predator information.
+We compare the time series of different population variables for Antarctic krill under most significant models in this analysis; a reference model without environmental or predator data and an alternative model that incorporates both environmental and predator information.
 Total biomass estimates remain relatively stable across both scenarios until around 2010.
@@
 
-```{r compara,eval=TRUE, out.width="60%", fig.cap=""}
+```{r compara,eval=TRUE, out.width="60%", fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
+Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)"}
 knitr::include_graphics('Figs/alloutput_s1_4.png')
-
 ```
 
-The different configurations to estimated posterior distribution of Natural log of virgin recruitment level (R0) suggesting that the choice of covariates has a substantial impact on the estimation of unfished recruitment (Figure \@ref(fig:compara2)).
-The model with both environmental and predator data yields the highest R0 values, while the model with only environmental data results in the lowest estimates.
+Biologically, ln(R0) represents the expected log-scale recruitment in the absence of fishing, providing a proxy for the reproductive potential or productivity of the population under pristine conditions.
+Figure \@ref(fig:compara2) displays posterior density distributions for the natural logarithm of virgin recruitment level (ln(R0)) across four model configurations.
+The narrow and well-separated peaks across models suggest that each model configuration yields distinctly different estimates of recruitment potential, likely reflecting differences in underlying assumptions such as natural mortality, selectivity, or the inclusion of predation effects.
+From a biological perspective, the higher ln(R0) values in `s1.3` and `s1.4` suggest that, under those model assumptions, the krill population could sustain higher levels of recruitment in unfished conditions, potentially reflecting stronger compensatory dynamics or reduced natural mortality (e.g., lower predation pressure).
+Conversely, the lower ln(R0) values in `s1.1` and `s1.2` imply a population with lower reproductive output, which might make it more vulnerable to overfishing or environmental variability.
 This highlights the sensitivity of the recruitment parameter to assumptions about ecological drivers in the stock-recruitment relationship.
@@
 Notably, the inclusion of ecosystem variables appears to reduce extreme recruitment events and leads to a more stable, density-dependent recruitment pattern aligned with theoretical expectations.
-
 Estimating the productivity of Antarctic krill is critical for understanding the species capacity to replenish its population in response to varying levels of spawning biomass.
@@
 This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill.
+
+### Model scenarios diagnostics summary
+
+The comparative evaluation of the four model scenarios (s1.1–s1.4) revealed key differences in model performance and complexity (Table \@ref(tab:diagnostics)).
+Scenario s1.1, the simplest model with no ecosystem covariates, exhibited the lowest maximum gradient component (2.4e−05), suggesting a well-converged solution.
+It also used the fewest estimated parameters (94) and achieved a relatively low RMSE (0.88), indicating a good fit to the data despite its simplicity.
+However, its negative log-likelihood (NLL = 1,291.04) was higher compared to more complex models.
+Model s1.2, which includes only environmental covariates, showed a substantial increase in the likelihood fit (NLL = 25,963.90), although it also required a larger number of parameters (152).
+Similarly, s1.3, which includes only predator effects, maintained a moderate parameter count (140) but performed slightly worse in terms of RMSE (0.93), indicating potential overfitting or reduced explanatory power relative to model complexity.
+The full model, s1.4, incorporating both environmental and predator effects, achieved a strong likelihood value (NLL = 25,188.80), albeit with the highest number of parameters (175).
+Its RMSE (0.89) was similar to the other scenarios, suggesting that the inclusion of additional covariates may have improved likelihood without substantial overfitting.
+The gradient component (5.18e−05) remained within acceptable bounds, supporting convergence.
+These results suggest a trade-off between model complexity and fit, where models incorporating ecological covariates (s1.2 and s1.4) improved the likelihood function but required more parameters, while the simpler model (s1.1) retained strong convergence properties with fewer degrees of freedom.
+
+```{r diagnostics, eval=TRUE, warning=FALSE, message=FALSE}
+diagnostics <- read_csv("model_diagnosis.csv")
+#write_csv(diagnostics, "model_diagnosis.csv")
+flextable(diagnostics) %>%
+  set_caption("diffenrent statistical componentes to evaluate perfomances and diagnostic of model scenarios") %>%
+  autofit() %>%
+  theme_booktabs() %>%
+  fontsize(size = 10, part = "all") %>%
+  set_table_properties(width = .85, layout = "autofit")
+```
 

### 2025-04-15 index.Rmd
update scenarios

@@
            subplots = 1,
-           pheight = 15)
+           ptsize = 0.5)
 
@@
            subplots = 1,
-           pheight = 15)
+           ptsize = 0.5)
 ```
@@
 
+another
+
+```{r}
+dfpearson_long <- data.frame(
+  Pearson = c(base.model1.1$lendbase$Pearson,
+              base.model1.2$lendbase$Pearson,
+              base.model1.3$lendbase$Pearson,
+              base.model1.4$lendbase$Pearson),
+  Scenario = factor(c(rep("s1.1", length(base.model1.1$lendbase$Pearson)),
+                      rep("s1.2", length(base.model1.2$lendbase$Pearson)),
+                      rep("s1.3", length(base.model1.3$lendbase$Pearson)),
+                      rep("s1.4", length(base.model1.4$lendbase$Pearson))))
+)
+
+pearson_plot <- ggplot(dfpearson_long, aes(x = Scenario, 
+                                            y = Pearson, 
+                                            fill = Scenario)) +
+  geom_boxplot(width = 0.2, alpha = 0.8,
+               outliers = FALSE) +
+  #geom_jitter(width = 0.01, alpha = 0.3) +
+  theme_minimal() +
+  scale_fill_manual(name = "Scenario",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  labs(title = "", x = "", y = "Residual") +
+  theme(legend.position = "none")
+
+```
+
 ### Comparision RMSE
@@
   geom_boxplot(width=0.2,
-               alpha=0.8) +
-  geom_jitter(width = 0.01, alpha = 0.3) +
+               alpha=0.8,
+               outliers = FALSE) +
+  #geom_jitter(width = 0.01, alpha = 0.3) +
   theme_minimal() +
@@
   theme(legend.position = "none")
-rmse
+```
+
+```{r}
+ggarrange(rmse,
+          pearson_plot,
+          ncol=2)
 ```
@@
 
+### Table Diagnosis
+
+```{r}
+# Construcción de tabla resumen
+diagnostics <- data.frame(
+  Model = c("s1.1", "s1.2", "s1.3", "s1.4"),
+  Gradient = c(
+    base.model1.1$maximum_gradient_component,
+    base.model1.2$maximum_gradient_component,
+    base.model1.3$maximum_gradient_component,
+    base.model1.4$maximum_gradient_component
+  ),
+  Likelihood = c(
+    base.model1.1$likelihoods_used[1,1],
+    base.model1.2$likelihoods_used[1,1],
+    base.model1.3$likelihoods_used[1,1],
+    base.model1.4$likelihoods_used[1,1]
+  ),
+  Parameters = c(
+    dim(base.model1.1$parameters)[1],
+    dim(base.model1.2$parameters)[1],
+    dim(base.model1.3$parameters)[1],
+    dim(base.model1.4$parameters)[1]
+  ),
+  RMSE = c(
+    mean(as.numeric(base.model1.1$index_variance_tuning_check$RMSE)),
+    mean(as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10])),
+    mean(as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10])),
+    mean(as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10]))
+  )
+)
+#write_csv(diagnostics, "model_diagnosis.csv")
+flextable(diagnostics) %>%
+  set_caption("Diagnostic summary of model scenarios") %>%
+  autofit() %>%
+  theme_booktabs() %>%
+  fontsize(size = 10, part = "all") %>%
+  set_table_properties(width = .85, layout = "autofit")
+
+```
+
 ## Outputs
@@
 
-# Gráfico de barras agrupadas por variable y modelo
-alloutput <- ggplot(outpsall_long, aes(x = factor(Yr), y = Value, fill = Model)) +
-  geom_bar(stat = "identity", position = "dodge") +
+# Gráfico de líneas por variable y modelo
+alloutput <- ggplot(outpsall_long, aes(x = Yr, y = Value, color = Model, group = Model)) +
+  geom_line(linewidth = 1) +
+  geom_point(size = 2) +
   facet_wrap(~Variable, 
-             ncol=1,
+             ncol = 1, 
              scales = "free_y") + 
-  scale_fill_manual(name= "Scenario",
+  scale_color_manual(name = "",
                      values = c("s1.1 (Ref Model: No Env-Predator)" = "#ca0020",
-
                                 "s1.4 (Env and Predator data)" = "#404040")) +
-  labs(title = "",
-       x = "", y = "") +
-  theme_minimal() +
+  labs(title = "", x = "", y = "") +
+  theme_bw() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1),
-        legend.position = "bottom")
+        legend.position = "none")
+
 alloutput
+
 ```
@@
 
+### Explotation Rate
+
+```{r}
+df1 <- base.model1.1$exploitation[, c(1, 4)] %>% mutate(model = "s1.1")
+df2 <- base.model1.2$exploitation[, c(1, 4)] %>% mutate(model = "s1.2")
+df3 <- base.model1.3$exploitation[, c(1, 4)] %>% mutate(model = "s1.3")
+df4 <- base.model1.4$exploitation[, c(1, 4)] %>% mutate(model = "s1.4")
+
+df_all <- bind_rows(df1, df2, df3, df4)
+colnames(df_all) <- c("year", "exploitation_rate", "model")
+
+colores <- c("s1.1" = "#ca0020", 
+             "s1.2" = "#f4a582", 
+             "s1.3" = "#bababa", 
+             "s1.4" = "#404040")
+ggplot(df_all, aes(x = year, y = exploitation_rate, color = model)) +
+  geom_line() +
+  geom_point() +
+  scale_color_manual(values = colores) +
+  labs(
+    title = "Exploitation Rate Over Time",
+    x = "Year",
+    y = "Exploitation Rate",
+    color = "Model"
+  ) +
+  theme_minimal()
+```
+
 ### Platoons analisis

### 2025-04-15 MS2_Krill.Rmd
add table summary

@@
 
-# ABSTRACT
+# Abstract
 
@@
 
-# INTRODUCTION
+# Introduction
 
@@
 
-# METHODOLOGY
+# Methodology
 
@@
 
-# RESULTS
+# Results
 
@@
 The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population (Figure \@ref(fig:pearson)).
-In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
+In scenario s1.1 (Without Ecosystem Variables), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
 Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
-The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals.
+s1.2 (With Predator) shows some improvement in model fit, with slightly more balanced residuals.
 However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
-In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
+In s1.3 (With Environment), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
 This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
-Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure.
-The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
+Finally, s1.4 (With Predator and Ecosystem) shows the most consistent reduction in Pearson residuals across time and size structure.
+The residual patterns also indicate potentially concerning model misfits, particularly among adult individuals in several survey fleets operating in the Bransfield Strait, as well as among the largest size classes in the predator fleet, which likely represents penguin consumption data.
+The residuals exhibit consistent over- or under-estimation in specific length bins over multiple years, suggesting that the current model structure may not fully capture key processes affecting the abundance or selectivity of adult individuals.
+In the case of surveys in Bransfield, this could reflect spatial mismatches between predicted and observed size distributions or biases in survey selectivity for larger size classes.
+This spatial heterogeneity aligns with previous findings identifying this area as a core juvenile krill habitat [@Mardones2025; @Veytia2021, Vetia et al., 2019, REF, REF].
+Similarly, the residuals associated with the predator fleet suggest that the model may inadequately represent size-selective predation or the temporal dynamics of predator consumption.
+These patterns warrant further investigation into the assumptions underlying length selectivity and mortality, particularly for the adult component of the population, as well as potential ecosystem-driven variability not currently integrated in the model.
+
+Overall, the distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
 
 ```{r pearson, eval=TRUE, fig.cap= "Pearson residuals of predicted versus observed length-frequency distributions of Antarctic krill across four modeling scenarios (s1.1 to s1.4), each incorporating increasing levels of ecosystem complexity.
-Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components"}
+Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components (Red = positive, Black = negative"}
 knitr::include_graphics('Figs/Pearson.png')
@@
 The comparative evaluation of the four model scenarios (s1.1–s1.4) revealed key differences in model performance and complexity (Table \@ref(tab:diagnostics)).
-Scenario s1.1, the simplest model with no ecosystem covariates, exhibited the lowest maximum gradient component (2.4e−05), suggesting a well-converged solution.
-It also used the fewest estimated parameters (94) and achieved a relatively low RMSE (0.88), indicating a good fit to the data despite its simplicity.
-However, its negative log-likelihood (NLL = 1,291.04) was higher compared to more complex models.
-Model s1.2, which includes only environmental covariates, showed a substantial increase in the likelihood fit (NLL = 25,963.90), although it also required a larger number of parameters (152).
-Similarly, s1.3, which includes only predator effects, maintained a moderate parameter count (140) but performed slightly worse in terms of RMSE (0.93), indicating potential overfitting or reduced explanatory power relative to model complexity.
-The full model, s1.4, incorporating both environmental and predator effects, achieved a strong likelihood value (NLL = 25,188.80), albeit with the highest number of parameters (175).
-Its RMSE (0.89) was similar to the other scenarios, suggesting that the inclusion of additional covariates may have improved likelihood without substantial overfitting.
-The gradient component (5.18e−05) remained within acceptable bounds, supporting convergence.
-These results suggest a trade-off between model complexity and fit, where models incorporating ecological covariates (s1.2 and s1.4) improved the likelihood function but required more parameters, while the simpler model (s1.1) retained strong convergence properties with fewer degrees of freedom.
+All four scenarios achieved convergence, as indicated by a maximum gradient component of zero, ensuring that optimization reached a stable solution.
+Notably, the number of estimated parameters varied across models, with the full ecosystem model (s1.4) estimating the same number of parameters as the predator-only model (s1.2), while the environment-only model (s1.3) and the base model (s1.1) used fewer parameters.
+The Akaike Information Criterion (AIC) reveals stark contrasts in model performance.
+The environment-inclusive models (s1.3 and s1.1) have substantially lower AIC values (3,234.9 and 2,668.1, respectively) compared to the predator-inclusive models (s1.2 and s1.4), which exceed 50,000.
+These results suggest that environmental covariates contribute more effectively to explaining krill population dynamics than predator effects alone.
+The total likelihood (−logL) further supports this, showing much better fit in models s1.1 and s1.3, while predator-inclusive models (s1.2 and s1.4) show considerably poorer likelihood values.
+This pattern is reflected in the survey and length composition likelihood components, where again the environment-only model (s1.3) shows superior performance compared to predator scenarios.
+Root Mean Square Error (RMSE) for both index and length composition highlights better predictive accuracy in models without predators.
+Model s1.1 and s1.3 achieve lower RMSE values for both indices and length composition, indicating better alignment with observed data.
+In contrast, RMSE values are higher in predator-inclusive models (s1.2 and s1.4), suggesting poorer fit.
+Similarly, the Mean Absolute Scaled Error (MASE) remains lower in s1.1 and s1.3 (0.427), compared to s1.2 and s1.4 (0.517), again highlighting improved predictive performance in models that exclude predator effects.
+Retrospective biases (rho values) provide insight into model consistency over time.
+Retrospective bias in spawning stock biomass (Retro_Rho_ssb) and forecast biomass (Forecast_Rho_ssb) is substantially reduced in models without predators, particularly s1.1 (−0.233 and −0.333).
+Models including predators (s1.2 and s1.4) show larger negative rho values, indicating stronger retrospective bias and less stability in biomass estimation.
+Fishing mortality bias (Rho_f) and its forecast also present serious concerns in predator-inclusive scenarios.
+Model s1.4, for instance, shows a high retrospective bias in fishing mortality (Rho_f = 6.912), compared to 1.585 in the base model (s1.1).
+This indicates over-sensitivity in mortality estimates when predator effects are included without sufficient constraint or supporting data.
 
@@
 flextable(diagnostics) %>%
-  set_caption("diffenrent statistical componentes to evaluate perfomances and diagnostic of model scenarios") %>%
+  colformat_num(digits = 3) |>  
+  set_caption("Statistical componentes to evaluate perfomances and diagnostic of model scenarios") %>%
   autofit() %>%
@@
 
-# DISCUSION
+# Discusion
 
@@
 
+Tabla summary
+
+The analysis suggests that incorporating environmental variables (s1.3 and s1.4) improves model fit and predictive accuracy, particularly when compared to predator-only models.
+The model that includes only environmental covariates (s1.3) strikes the best balance between fit quality, complexity, and retrospective performance, suggesting that environmental variability is a key driver of krill population dynamics in the Western Antarctic Peninsula.
+
+In contrast, the inclusion of predators alone (s1.2) leads to poorer model fit and higher retrospective bias, possibly due to data limitations or unaccounted complexity in predator-prey dynamics.
+While the full ecosystem model (s1.4) shows some improvements over the predator-only model, it still exhibits significant retrospective biases, highlighting the need for refined integration of ecosystem components and further data support, particularly regarding predator consumption rates and spatial foraging behavior.
+
+These findings support growing consensus that ecosystem-informed stock assessments, particularly those that integrate environmental covariates, provide more robust and ecologically realistic estimates for Antarctic krill and should inform precautionary management frameworks under CCAMLR.
+
 \newpage
 
-# CONCLUSION
+# Conclusion
 
@@
 
-# ACKNOWLEDGMENTS
+# Acknowledgments
 
@@
 
-# SUPPLEMENTARY MATERIAL
+# Supplementary Material
 
@@
 
+# Data Availability
+
+Data will be made available on request.
+
 \newpage
 
-# REFERENCES
+# References

### 2025-04-15 index.Rmd
add table summary

@@
 
-Heatmap
-
-```{r}
-
-create_heatmap <- function(df) {
-  df1 <- df %>%
+Pearson residual
+
+```{r}
+create_heatmap_df <- function(df, model_name) {
+  df %>%
     dplyr::filter(Pearson < 5) %>%
@@
       TRUE ~ NA_character_
-    ))
-
-  ggplot(df1, aes(x = Yr, y = Bin, fill = Pearson)) +
+    )) %>%
+    dplyr::mutate(Model = model_name)
+}
+
+# Unir los datasets en uno solo
+df_all <- bind_rows(
+  create_heatmap_df(base.model1.1$lendbase, "Model 1.1"),
+  create_heatmap_df(base.model1.2$lendbase, "Model 1.2"),
+  create_heatmap_df(base.model1.3$lendbase, "Model 1.3"),
+  create_heatmap_df(base.model1.4$lendbase, "Model 1.4")
+)
+
+# Plot combinado: facet_grid por Modelo y Fleet
+heatmapres <- ggplot(df_all, aes(x = Yr, y = Bin, fill = Pearson)) +
   geom_tile() +
-    scale_fill_gradient2(low = "black", 
+  scale_fill_gradient2(low = "red", 
                        mid = "white", 
-                         high = "red", 
+                       high = "blue",
                        midpoint = 0) +
-    facet_wrap(Fleet~., ncol=6) + 
-    labs(title = "", 
-         x = "", 
-         y = "Length (cm)", 
-         fill = "Pearson") + 
+  facet_grid(Fleet ~ Model) +
   theme_few() +
-    theme(axis.text.x = element_text(angle = 90, 
-                                     hjust = 1, 
-                                     size = 7),
-          legend.position = "bottom")
-}
-
-# Crear heatmaps para los diferentes dataframes
-heatmap1 <- create_heatmap(base.model1.1$lendbase)
-heatmap2 <- create_heatmap(base.model1.2$lendbase)
-heatmap3 <- create_heatmap(base.model1.3$lendbase)
-heatmap4 <- create_heatmap(base.model1.4$lendbase)
-
-# Combina los plots en un solo gráfico con 8 columnas y 1 fila, compartiendo la leyenda
-ggarrange(heatmap1, heatmap2, heatmap3, heatmap4, 
-          ncol = 2, 
-          nrow = 2, 
-          common.legend = TRUE, 
-          legend = "bottom")
+  theme(
+    axis.text.x = element_text(angle = 90, hjust = 1, size = 5),         
+    axis.text.y = element_text(size = 5),                               
+    strip.text = element_text(size = 8),                                 
+    legend.position = "none"
+  ) +
+  labs(x = "", y = "Length (cm)", fill = "Pearson")
+
+heatmapres
 ```
@@
 
-```{r eval=FALSE}
-# Assuming your data is in a data frame called 'lendbase'
-# Filter and select relevant columns
-data <- base.model1.4$lendbase %>%
-  dplyr::filter(Pearson<5) %>% 
-  dplyr::select(c(1, 6, 19, 16)) %>% 
-  dplyr::mutate(Fleet = dplyr::case_when(
-    Fleet == 1 ~ "FISHERYBS",
-    Fleet == 2 ~ "FISHERYEI",
-    Fleet == 3 ~ "FISHERYGS",
-    Fleet == 4 ~ "FISHERYJOIN",
-    Fleet == 5 ~ "FISHERYSSIW",
-    Fleet == 6 ~ "SURVEYBS",
-    Fleet == 7 ~ "SURVEYEI",
-    Fleet == 8 ~ "SURVEYGS",
-    Fleet == 9 ~ "SURVEYJOIN",
-    Fleet == 10 ~ "SURVEYSSIW",
-    Fleet == 11 ~ "PREDATOR",
-    TRUE ~ NA_character_  # Para manejar cualquier valor que no coincida
-  )) %>% 
-  dplyr::select(Yr, Bin, Pearson)
-
-# Create the hexagonal heatmap
-ggplot(data, aes(x = Yr, y = Bin)) +
-  geom_hex(bins=50) +  # Hexagonal bins
-  scale_fill_gradient2(low = "blue", 
-                       mid = "white", 
-                       high = "red", 
-                       midpoint = 0) +  # Color scale
-  theme_minimal() +  # Minimal theme for cleaner look
-  labs(title = "Hexagonal Heatmap of Pearson Residuals",
-       x = "Year",
-       y = "Bin",
-       fill = "Pearson Residual")  # Labels for plot and axes
-
-```
-
 ## Diagnosis Base Model
@@
 This framework follows the recommendations outlined by @Carvalho2021b, aiming to enhance transparency and reproducibility in model evaluation.
+
+### Convergence Criteria
+
+The convergence criterion used for model calibration is set to a final threshold of **0.0001** (or equivalently **1.0e-04**).
+This criterion defines the minimum acceptable difference between successive model iterations.
+Convergence is considered achieved when the absolute change in the objective function value or key parameters falls below this threshold.
+A smaller convergence value ensures that the model achieves a high degree of accuracy and stability in its final estimates, indicating that further iterations are unlikely to result in significant changes to the parameter estimates.
 
@@
 par(mfrow=c(2,2), mar=c(5,4,2,1))
-SSplotJABBAres(base.model1.1,
+rmse1l <- SSplotJABBAres(base.model1.1,
                subplots = "len",
                add=T)
-SSplotJABBAres(base.model1.2,
+rmse2l <- SSplotJABBAres(base.model1.2,
                subplots = "len",
                add=T)
-SSplotJABBAres(base.model1.3,
+rmse3l <- SSplotJABBAres(base.model1.3,
                subplots = "len",
                add=T)
-SSplotJABBAres(base.model1.4,
+rmse4l <- SSplotJABBAres(base.model1.4,
                subplots = "len",
@@
 par(mfrow=c(2,2), mar=c(5,4,2,1))
-SSplotJABBAres(base.model1.1,
+rmse1c <- SSplotJABBAres(base.model1.1,
                subplots = "cpue",
                add=T)
-SSplotJABBAres(base.model1.2,
+rmse2c <- SSplotJABBAres(base.model1.2,
                subplots = "cpue",
                add=T)
-SSplotJABBAres(base.model1.3,
+rmse3c <- SSplotJABBAres(base.model1.3,
                subplots = "cpue",
                add=T)
-SSplotJABBAres(base.model1.4,
+rmse4c <- SSplotJABBAres(base.model1.4,
                subplots = "cpue",
@@
 par(mfrow=c(2,4), mar=c(5,4,2,1))
-hci = SSplotHCxval(retroSummary1.1, 
+hci1 = SSplotHCxval(retroSummary1.1, 
                    add = T, 
@@
 par(mfrow=c(3,3), mar=c(5,4,2,1))
-hci = SSplotHCxval(retroSummary1.2, 
+hci2 = SSplotHCxval(retroSummary1.2, 
                    add = T, 
@@
 par(mfrow=c(2,4), mar=c(5,4,2,1))
-hci = SSplotHCxval(retroSummary1.3, 
+hci3 = SSplotHCxval(retroSummary1.3, 
                    add = T, 
@@
 par(mfrow=c(3,3), mar=c(5,4,2,1))
-hci = SSplotHCxval(retroSummary1.4, 
+hci4 = SSplotHCxval(retroSummary1.4, 
                    add = T, 
@@
 
-### `SSplotProfile()`
+### Likelihood Profile
 
@@
 ```
-
-### Convergence Criteria
-
-The convergence criterion used for model calibration is set to a final threshold of **0.0001** (or equivalently **1.0e-04**).
-This criterion defines the minimum acceptable difference between successive model iterations.
-Convergence is considered achieved when the absolute change in the objective function value or key parameters falls below this threshold.
-A smaller convergence value ensures that the model achieves a high degree of accuracy and stability in its final estimates, indicating that further iterations are unlikely to result in significant changes to the parameter estimates.
 
@@
 ```{r}
-# Construcción de tabla resumen
-diagnostics <- data.frame(
-  Model = c("s1.1", "s1.2", "s1.3", "s1.4"),
-  Gradient = c(
-    base.model1.1$maximum_gradient_component,
-    base.model1.2$maximum_gradient_component,
-    base.model1.3$maximum_gradient_component,
-    base.model1.4$maximum_gradient_component
-  ),
-  Likelihood = c(
-    base.model1.1$likelihoods_used[1,1],
-    base.model1.2$likelihoods_used[1,1],
-    base.model1.3$likelihoods_used[1,1],
-    base.model1.4$likelihoods_used[1,1]
-  ),
-  Parameters = c(
-    dim(base.model1.1$parameters)[1],
-    dim(base.model1.2$parameters)[1],
-    dim(base.model1.3$parameters)[1],
-    dim(base.model1.4$parameters)[1]
-  ),
-  RMSE = c(
-    mean(as.numeric(base.model1.1$index_variance_tuning_check$RMSE)),
-    mean(as.numeric(base.model1.2$index_variance_tuning_check$RMSE[1:10])),
-    mean(as.numeric(base.model1.3$index_variance_tuning_check$RMSE[1:10])),
-    mean(as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10]))
+# Modelo s1.1
+diag1.1 <- data.frame(
+  Scenario = "s1.1",
+  Convergency = base.model1.1$maximum_gradient_component,
+  AIC = as.numeric(2 * dim(base.model1.1$estimated_non_dev_parameters)[1] + 2 * base.model1.1$likelihoods_used[1, 1]),
+  Total_like = base.model1.1$likelihoods_used$values[rownames(base.model1.1$likelihoods_used) == "TOTAL"],
+  N_Params = as.numeric(2 * dim(base.model1.1$estimated_non_dev_parameters)[1]),
+  Survey_like = base.model1.1$likelihoods_used$values[rownames(base.model1.1$likelihoods_used) == "Survey"],
+  Length_comp_like = base.model1.1$likelihoods_used$values[rownames(base.model1.1$likelihoods_used) == "Length_comp"],
+  RMSE_index = rmse1c$RMSE.perc[rmse1c$indices == "Combined"],
+  RMSE_length = rmse1l$RMSE.perc[rmse1l$indices == "Combined"],
+  MASE = mean(hci1$MAE.base, na.rm = TRUE),
+  Retro_Rho_ssb = tablebias01[5, 3],
+  Forecast_Rho_ssb = tablebias01[5, 4],
+  Forecast_Rho_f = tablebias01a[5, 3],
+  Rho_f = tablebias01a[5, 4]
 )
+
+# Modelo s1.2
+diag1.2 <- data.frame(
+  Scenario = "s1.2",
+  Convergency = base.model1.2$maximum_gradient_component,
+  AIC = as.numeric(2 * dim(base.model1.2$estimated_non_dev_parameters)[1] + 2 * base.model1.2$likelihoods_used[1, 1]),
+  Total_like = base.model1.2$likelihoods_used$values[rownames(base.model1.2$likelihoods_used) == "TOTAL"],
+  N_Params = as.numeric(2 * dim(base.model1.2$estimated_non_dev_parameters)[1]),
+  Survey_like = base.model1.2$likelihoods_used$values[rownames(base.model1.2$likelihoods_used) == "Survey"],
+  Length_comp_like = base.model1.2$likelihoods_used$values[rownames(base.model1.2$likelihoods_used) == "Length_comp"],
+  RMSE_index = rmse2c$RMSE.perc[rmse2c$indices == "Combined"],
+  RMSE_length = rmse2l$RMSE.perc[rmse2l$indices == "Combined"],
+  MASE = mean(hci2$MAE.base, na.rm = TRUE),
+  Retro_Rho_ssb = tablebias2[5, 3],
+  Forecast_Rho_ssb = tablebias2[5, 4],
+  Forecast_Rho_f = tablebias2a[5, 3],
+  Rho_f = tablebias2a[5, 4]
 )
-#write_csv(diagnostics, "model_diagnosis.csv")
-flextable(diagnostics) %>%
-  set_caption("Diagnostic summary of model scenarios") %>%
-  autofit() %>%
-  theme_booktabs() %>%
-  fontsize(size = 10, part = "all") %>%
-  set_table_properties(width = .85, layout = "autofit")
-
+
+# Modelo s1.3
+diag1.3 <- data.frame(
+  Scenario = "s1.3",
+  Convergency = base.model1.3$maximum_gradient_component,
+  AIC = as.numeric(2 * dim(base.model1.3$estimated_non_dev_parameters)[1] + 2 * base.model1.3$likelihoods_used[1, 1]),
+  Total_like = base.model1.3$likelihoods_used$values[rownames(base.model1.3$likelihoods_used) == "TOTAL"],
+  N_Params = as.numeric(2 * dim(base.model1.3$estimated_non_dev_parameters)[1]),
+  Survey_like = base.model1.3$likelihoods_used$values[rownames(base.model1.3$likelihoods_used) == "Survey"],
+  Length_comp_like = base.model1.3$likelihoods_used$values[rownames(base.model1.3$likelihoods_used) == "Length_comp"],
+  RMSE_index = rmse3c$RMSE.perc[rmse3c$indices == "Combined"],
+  RMSE_length = rmse3l$RMSE.perc[rmse3l$indices == "Combined"],
+  MASE = mean(hci3$MAE.base, na.rm = TRUE),
+  Retro_Rho_ssb = tablebias3[5, 3],
+  Forecast_Rho_ssb = tablebias3[5, 4],
+  Forecast_Rho_f = tablebias3a[5, 3],
+  Rho_f = tablebias3a[5, 4]
+)
+
+# Modelo s1.4
+diag1.4 <- data.frame(
+  Scenario = "s1.4",
+  Convergency = base.model1.4$maximum_gradient_component,
+  AIC = as.numeric(2 * dim(base.model1.4$estimated_non_dev_parameters)[1] + 
+                     2 * base.model1.4$likelihoods_used[1, 1]),
+  Total_like = base.model1.4$likelihoods_used$values[rownames(base.model1.4$likelihoods_used) == "TOTAL"],
+  N_Params = as.numeric(2 * dim(base.model1.4$estimated_non_dev_parameters)[1]),
+  Survey_like = base.model1.4$likelihoods_used$values[rownames(base.model1.4$likelihoods_used) == "Survey"],
+  Length_comp_like = base.model1.4$likelihoods_used$values[rownames(base.model1.4$likelihoods_used) == "Length_comp"],
+  RMSE_index = rmse4c$RMSE.perc[rmse4c$indices == "Combined"],
+  RMSE_length = rmse4l$RMSE.perc[rmse4l$indices == "Combined"],
+  MASE = mean(hci4$MAE.base, na.rm = TRUE),
+  Retro_Rho_ssb = tablebias4[5, 3],
+  Forecast_Rho_ssb = tablebias4[5, 4],
+  Forecast_Rho_f = tablebias4a[5, 3],
+  Rho_f = tablebias4a[5, 4]
+)
+
+diag_all <- rbind(diag1.1,
+                  diag1.2, 
+                  diag1.3, 
+                  diag1.4)
+
+diag_tidy <- diag_all |>
+  pivot_longer(cols = -Scenario, names_to = "Description", values_to = "Value") |>
+  pivot_wider(names_from = Scenario, values_from = Value)
+
+diag_tidy[, 2:5] <- lapply(diag_tidy[, 2:5], function(x) as.numeric(format(round(x, 3), nsmall = 3)))
+
+#write_csv(diag_tidy, "model_diagnosis.csv")
+flextable(diag_tidy) |>
+  colformat_num(digits = 3) |>  
+  autofit() |>
+  theme_booktabs() |>
+  align(align = "center", part = "all") |>
+  bold(part = "header")
 ```

### 2025-05-06 MS2_Krill.Rmd
adding edits

@@
 ---
-title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated catch-at-length model perspective"
+title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
 author:
@@
 Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP).
-In this study, we implement an integrated Catch-at-Length model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
+In this study, we implement an integrated length-to-age model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
 We explored multiple scenarios that accounted for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species.
@@
 Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through post hoc analyses.
-Overlooking these drivers may result in less accurate conclusions and limit the effectiveness of management recommendations.
-Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s conservation and sustainable use objectives.
+Overlooking these drivers may result in less accurate conclusions, hence limiting the effectiveness of management recommendations.
+Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
 
-*Keywords: Antarctic krill, population dynamics, stock-recruitment, ecosystem variables, Stock Synthesis model, perfomance model, fisheries management, CCAMLR*
+*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, perfomance model, fisheries management, CCAMLR*
 
@@
 
-Antarctic krill (*Euphausia superba*) is a keystone species in the Antarctic ecosystem and one of the largest populations on the planet in terms of biomass (REF).
-Currently, its population is predominantly concentrated in the Western Antarctic Peninsula (WAP), a critical feeding ground for predators and the primary area for krill harvesting over the past four decades (REF).
-At the same time, the WAP is one of the most vulnerable regions to climate change (CC) and has already undergone rapid environmental shifts [@Turner2005; @Lima2013; @McBride2021; @Morley2020; @Atkinson2019a].
-These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective.
-Given its ecological significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
+Antarctic krill (*Euphausia superba*) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @Kawaguchi2024; @McBride2021].
+Currently, the population, and consecuently, the fishery is predominantly concentrated in the Western Antarctic Peninsula (WAP).
+This area is also a critical feeding ground for predators such as penguins, seals, and whales, and the primary area for krill harvesting over the past four decades [@Gallagher2023a; @Watters2013; @Hill2016].
+At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Lima2013; @Turner2005; @Carrasco2021; @Atkinson2019a; @Morley2020; @McBride2021].
+These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].
 
-Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has adopted a new management strategy (CCAMLR 2019) that follows an ecosystem-based approach.
-This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species.
-It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks.
+Given its ecosystem significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
+Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) that follows an ecosystem-based approach has been developing [@Cavanagh2016; @Constable2023; @Constable2011; @Watters2013].
+This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and time.
+It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Kawaguchi2024; @Warwick-Evans2022].
 
-As part of this strategy, CCAMLR has tested various modeling approaches to assess krill stock status and support fisheries management decisions [@Hill2016].
-While model-based stock assessments have been widely used (REF), many struggle to align with management objectives or accurately represent krill population dynamics due to their spatial variability and ecosystem complexity.
-Traditionally, stock assessments have relied on single-source data, either from fisheries catches [@Kinzey2015a; @Wang2021] or population monitoring surveys [@Green2023; @Watters2013].
-One widely applied approach is the General Yield Model (GRYM), which provides global catch limits based on predefined harvest thresholds [@Maschette2020].
-However, such approaches often neglect key ecological drivers.
-Considering the influence of habitat on population dynamics, the role of environmental variability in shaping krill productivity, and the contribution of predators as a source of natural mortality, stock assessment frameworks should explicitly incorporate these ecosystem components to more accurately represent population processes.
+To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Green2023; @Hill2019; @Dornam2021].
+However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, and the high costs associated with such studies have further motivated the search for alternative stock estimation methods.
+More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019].
+Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019a; @Green2023].
+Nowadays, a model-based approach used for generating management recommendations in CCAMLR is the Generalized Yield Model (GRYM) [@Maschette2020], which allows adjustments to the decision rules to predefined captures set by CCAMLR [@Constable2000; @CCAMLR2021a; @CCAMLR2010].
+However, these model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics and/or failure to account for the high spatial and temporal variability intrinsic to krill systems [@Meyer2020; @Trathan2021].
 
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
-However, integrating these complexities presents several challenges.
-Long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies.
-Additionally, some areas have sparse or incomplete datasets on krill dynamics.
-Furthermore, indirect drivers, such as environmental variability and predator abundance, introduce another layer of complexity.
-Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations.
+Increasing evidence shows that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Flores2012; @Morley2020; @Saba2014; @Perry2020; @Veytia2021; @Walsh2020; @Murphy2007; @Vorrath2020].
+Likewise, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Reisinger2022; @Riaz2023; @Baines2022; @Biuw2024; @Moosa2017].
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions [REF].
+Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Marshall2019; @Crone2019; @Richards2016].
+Therefore, any stock assessment frameworks related with new strategy of CCAMLR should explicitly incorporate these ecosystem components to more accurately represent krill population processes.
 
-To address this need, we propose a conceptual framework for implementing a statistical model that integrates multiple data sources to assess the population dynamics of Antarctic krill in the Western Antarctic Peninsula (WAP) over the past decades.
-This approach incorporates biological, fishery, survey, and ecosystem-related information to provide a more comprehensive understanding of krill stock fluctuations and productivity.
+Integrating ecological complexity into krill stock assessments presents multiple challenges.
+Long-term data consistency is often affected by changes in monitoring programs and sampling methodologies [REF], while some key regions of the Western Antarctic Peninsula (WAP) lack sufficient data to construct reliable krill population indices.
+Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [REF].
+Despite these challenges, substantial data exist on krill and their environment.
+Fishery monitoring programs provide detailed catch and effort data through onboard scientific observers.
+Ecosystem monitoring, like the "CCAMLR Ecosystem Monitoring Program" (CEMP), track predator populations and environmental variability at key sites.
+Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.
 
-Specifically, we employ an integrated catch-at-length implemented in Stock Synthesis model [@Methot2013] that allows us to evaluate the effects of environmental drivers on recruitment and productivity, as well as the impact of predator components into an integrated stock assessment model to Antarctic krill.
-The primary objective of this study was to assess the statistical performance of different management metrics, population variables and parameters based on alternative approaches to incorporating both environmental and predator data into a typical krill stock assessment model.
-By analyzing differences in estimated biomass, recruitment patterns, and overall krill productivity across these scenarios, we aim to quantify the extent to which ecosystem factors drive krill population trends.
-This comparative approach enables us to assess the relative importance of environmental conditions and predator dynamics in shaping krill stocks, providing critical insights for both fisheries management and conservation efforts.
+This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations.
+Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
+For statistical catch-at-age assessments, key frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013] among others.
+These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and populations dynamics [@Hoyle2020].
 
-By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions.
-This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
+This study aims to integrate different sources of available data on Antarctic krill (*Euphausia superba*) —including fishery operations (catches, effort distribution), biological characteristics (growth rates, reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (environmental forcing and predator-prey interactions)—into a unified stock assessment framework.
+Specifically, we employ an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting  biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
+The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices.
+Our results provide empirical evidence for the value of ecosystem-informed in model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in  model fit, predictive skill and structural uncertainty.
+
+This approach seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach [@CCAMLRWGSAM2024; @Hill2019; @Constable2023; @Zhao2023; @WGEMM2024; @Cavanagh2016; @Constable2011].
 Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
@@
 
-The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale.
-To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021] (Fig 1 S1).
-While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale.
-This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics.
-This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.
+The study area covers the Western Antarctic Peninsula (WAP), corresponding to FAO Subarea 48.1, which hosts a high of krill fishing activity.
+This is also the region where CCAMLR is actively developing a revised management approach at a finer spatial scale.
+To enhance the spatial resolution of our analysis of krill population dynamics, we adopted the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
+1, S1).
+It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological or population-specific criteria [@Zaldua2024a; @Constable2023; @Watters2013].
+However, their subdivision enables a finer-scale analysis of krill data.
+This increased spatial resolution allows for the detection of heterogeneity among various data sources—whether from fisheries or environmental  variables—thus providing a more detailed understanding of krill population dynamics.
 
-## Conceptual Model
+## Spatially Implicit Model
 
-In this approach we include spatial structure how an implicit approach where different areas are treated as separate fleets [REF].
-Differences among strata have been identified from various perspectives of krill population dynamics [REF], as well as from the influence of environmental variability within Subarea 48.1 [REF] (Fig 2 S1).
-This approach allows integrate spatial heterogeneity into the stock assessment model to ensure that spatially structured processes are adequately represented.
-Although strata are treated as separate fleets, recruitment and growth processes remain unified, ensuring that the overall krill stock is modeled as a single population while accounting for spatially structured impacts at the Subarea 48.1 level.
-Each stratum is assigned a unique fishery and survey history, reflecting differences in harvesting intensity and spatial exploitation patterns.
-Survey indices and length composition data are fleet-specific, allowing for variations in data collection methods and regional abundance estimates.
-Selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
-The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January regarding @REF.
-This approach represents a spatially implicit model, as it captures spatial variability in exploitation patterns without explicitly modeling movement or recruitment connectivity between areas.
-This allows for a more realistic assessment of krill stock productivity, in line with CCAMLR’s ecosystem-based management framework.
+To represent spatial heterogeneity in krill dynamics we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* concept [@Punt2019; @Nielsen2021; @Waterhouse2014].
+The areas-as-fleets approach aims treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations can exert different effects across space.
+In this approach, each strata within the Western Antarctic Peninsula is treated as a distinct fleet and allows a spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices.
+This is especially relevant for krill, given the spatial variability in both population processes and fishing effort, given the abundant evidence of spatial heterogeneity [@Perry2019; @Perry2020; @Reiss2020; @SantaCruz2022; @Veytia2021].
+By modeling each MU as a separate fleet, we allow the assessment to estimate region-specific parameters that reflect this variability.
+Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location based in @Dornam2021.
+Similarly, fishery-dependent data—such as catch and composition—are stratified by MU according to the reported fishing effort allowing for regional variation in both abundance and observational processes.
+Given the difference en the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
+The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January-February [@Perry2019; @McBride2021].
+While strata are modeled as separate fleets, the population dynamics—such as recruitment and growth—remain unified, representing the krill stock as a single, interconnected population.
+This disaggregated structure enables the model to account for differences in small scale spatial dynamics and providing a flexible yet informative way to include spatial processes in stock assessments.
+Areas-as-fleet approach captures regional variability in dynamic and observation while avoiding the complexity of modeling explicit movement.
+This approach provides a more realistic representation of krill productivity within FAO Subarea 48.1.
+
+## Length-to-Age Model Implementation
+
+The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
+SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
+The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources described previously and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
+All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
+
+Like many invertebrates, krill present challenges for age determination [@Punt2013].
+Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017].
+As a result, stock assessments have relied on length-to-age models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length.
+Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics.
+Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
+Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure.
+To date, all integrated krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021].
+Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003].
+An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes [@Lee2024].
+ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
+This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks.
+The ALK matrix is typically defined as:
+
+$$
+P(A = a \mid L = l) = \frac{N_{a,l}}{\sum_{a} N_{a,l}}
+$$
+where \( P(A = a \mid L = l) \) is the probability that an individual of length \( l \) belongs to age \( a \),  \( N_{a,l} \) is the number of individuals of age \( a \) observed at length \( l \) and the denominator \( \sum_{a} N_{a,l} \) ensures that the probabilities sum to 1 for each length class.
+This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:
+
+$$
+N_a = \sum_{l} P(A = a \mid L = l) \cdot N_l
+$$
+
+where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age.
+In integrated stock assessment models, such as those using statistical catch-at-age (SCAA) or integrated length-based frameworks, the ALK provides a critical link between observable length data and underlying age-based population dynamics.
+
+The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
+The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
+In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
+In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
+This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass.
+Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in S1 Fig 7.
+
+Growth was modeled following a von Bertalanffy growth function with lenght-at-age parameterization.
+The model assumed linear growth below a predefined settlement age of 1.5 years.
+The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary.
+Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
+Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
+Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
+Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
+The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
+The model used a direct parameterization without additional offsets between sexes.
+Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
+Regarding selectivity functions, it was modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
+The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
+The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
+The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata.
+Parameter estimation phases were assigned, with negative values indicating fixed parameters.
+The survey selectivity was also modeled using a length-based logistic function.
+Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
+The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
+These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (Fig 6, S1).
+The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
+assuming that the predator can capture a wider range of krill sizes.
+These parameters were estimated in later phases, allowing flexibility in their estimation.
+Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
+This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF].
+The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10.
+The parameter was freely estimated in phase 1.
+The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
+The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
+Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
+The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
+The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
+Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
+Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
+To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
+Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.
+All of these configurations can be reviewed in the repository section of S1.
 
@@
 
-Multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation.
-These data include records from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been consistently collected on board fishing vessels by scientific observers through the CCAMLR Scheme of International Scientific Observation (SISO).
-This dataset contains detailed information including vessel identity, nationality, and georeferenced positions used in the exploratory analysis.
-In addition, scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling.
-Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were formatted according to model evaluation template requirements (see Supplementary Material 2).
+A set of multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation.
+These data included register from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been systematically collected on board fishing vessels by scientific observers through the "Scheme of International Scientific Observation" (SISO) Program.
+Scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling.
+Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were handled according to stock assessment model templates requirements (Suppl.
+Mat.
+2).
 Environmental and predator-related data were also incorporated, such as satellite-derived chlorophyll-a concentrations from the Copernicus Marine Service and predator abundance and diet composition, particularly from penguin colonies reported in the [MAPPPD database](https://www.penguinmap.com/mapppd/sources/).
@@
 
-Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
-Source data and temporal scale used in this stock assessment is showed in Fig 4, S1.
+Life history parameters, like growth, weight-length relation, natural mortality and maturity were used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
+Source data and temporal scale used in this stock assessment are shown in Fig 4, S1.
 
 ```{r parainit1, eval=TRUE}
-dir1.1<-here("s1.1") 
-start1 <- SS_readstarter(file = file.path(dir1.1,
-                                          "starter.ss"),
-                              verbose = FALSE)
-dat1 <- SS_readdat(file = file.path(dir1.1, start1$datfile),
-                        verbose = FALSE)
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.1,
-                                    start1$ctlfil),
-                        verbose = FALSE,
-                        use_datlist = TRUE, 
-                   datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1.1, 
-                                                "forecast.ss"),
-                              verbose = FALSE)
-parbio<-ctl1$MG_parms[1:10,c(1:4,7)]
- row.names( parbio)<-c("Nat M",
-                       "Lmin", 
-                       "Lmax",
-                       "VonBert K",
-                       "CV young",
-                       "CV old", 
-                       "Wt a", 
-                       "Wt b",
-                       "L50%", 
-                       "Mat slope")
- SRpar<-ctl1$SR_parms[1:5,c(1:4,7)]
- Qpar<-ctl1$Q_parms[1:2,c(1:4,7)]
- Selpar<-ctl1$size_selex_parms[1:22,c(1:4,7)]
+dir1.1 <- here("s1.1")
 
- parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-parInit %>% 
-  kable(format = "latex",
-        booktabs = TRUE, 
-        caption = "Input parameters for the initial SS3 model of krill.
-Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
-If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-  kable_styling(latex_options = c("striped"), 
-                full_width = FALSE, 
-                font_size = 9) %>%
-  pack_rows(index = c("Natural Mortality" = 1,
-                      "Growth" = 5,
-                      "Length-Weight Relation" = 2,
-                      "Maturity" = 2,
-                      "Stock-Recruit Relation" = 5,
-                      "Catchability" = 2,
-                      "Selectivity" = 4))
+start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
+dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
+ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
+                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
+fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)
+
+# Extraer parámetros
+parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
+rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
+                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")
+
+SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
+Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
+Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]
+
+# Combinar parámetros
+parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))
+
+# Añadir nombres de fila como columna "Parameter"
+parInit$Parameter <- rownames(parInit)
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(INIT  = round(INIT, 2),
+         PRIOR = round(PRIOR, 2))
+
+# Crear vector de categorías (39 filas)
+categories <- c(rep("Natural Mortality", 1),
+                rep("Growth", 4),
+                rep("Length-Weight Relation", 2),
+                rep("Maturity", 2),
+                rep("Stock-Recruit Relation", 5),
+                rep("Catchability", 2),
+                rep("Selectivity", 21))  # OJO: corregido a 23
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(Category = categories) %>%
+  dplyr::relocate(Category, Parameter)
+
+# Crear flextable
+ft <- flextable(parInit)
+ft <- set_caption(ft, "Input parameters for the initial SS3 krill model")
+ft <- autofit(ft)
+ft <- theme_booktabs(ft)
+ft <- fontsize(ft, size = 9, part = "all")
+ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")
+
+# Mostrar flextable
+ft
+
 ```
-
-## Catch-to-Length Model Implementation
-
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
-SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources described previously and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
-
-Like many invertebrates, krill present challenges for age determination [@Punt2013].
-Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017].
-As a result, stock assessments have relied on catch-at-length models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length.
-Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics.
-Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
-Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure.
-To date, all integrated krill assessments have used catch-at-length frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021].
-Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003].
-An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes.
-ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
-This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks.
-The ALK matrix is typically defined as:
-
-$$
-P(A = a \mid L = l) = \frac{N_{a,l}}{\sum_{a} N_{a,l}}
-$$
-where \( P(A = a \mid L = l) \) is the probability that an individual of length \( l \) belongs to age \( a \),  \( N_{a,l} \) is the number of individuals of age \( a \) observed at length \( l \) and the denominator \( \sum_{a} N_{a,l} \) ensures that the probabilities sum to 1 for each length class.
-This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:
-
-$$
-N_a = \sum_{l} P(A = a \mid L = l) \cdot N_l
-$$
-
-where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age.
-In integrated stock assessment models, such as those using statistical catch-at-age (SCAA) or integrated length-based frameworks, the ALK provides a critical link between observable length data and underlying age-based population dynamics.
-
-The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
-The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
-In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
-In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
-This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass.
-Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in Figure \@ref(fig:AKL)).
-
-```{r AKL, eval=TRUE, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in Subarea 48.1"}
-knitr::include_graphics('Figs/AKL.png')
-```
-
-Growth was modeled following a von Bertalanffy growth function with size-at-age parameterization.
-The model assumed linear growth below a predefined settlement age of 1.5 years.
-The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary.
-Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
-Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
-Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
-Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
-The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
-The model used a direct parameterization without additional offsets between sexes.
-Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
-Regarding selectivity functions, it was modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
-The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
-The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
-The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata.
-Parameter estimation phases were assigned, with negative values indicating fixed parameters.
-The survey selectivity was also modeled using a length-based logistic function.
-Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
-The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
-These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (Fig 6, S1).
-The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
-assuming that the predator can capture a wider range of krill sizes.
-These parameters were estimated in later phases, allowing flexibility in their estimation.
-Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
-This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF].
-The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10.
-The parameter was freely estimated in phase 1.
-The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
-The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
-Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
-The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
-The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
-Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
-Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
-To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.
-All of these configurations can be reviewed in the repository section of S1.
 
@@
 To test for overall differences among scenarios, a one-way ANOVA [@REF] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
-AIC
+Another quantitative statistical method used to evaluate model performance was the Akaike Information Criterion (AIC)[REF], computed as:
+
+$$
+\text{AIC} = 2k + 2L
+$$
+
+where \( k \) is the number of estimated non-deviation parameters and \( L \) is the negative log-likelihood.
+This criterion balances model fit and complexity, facilitating comparisons across scenarios.
+Lower AIC values indicate more parsimonious models with better overall fit.
 
@@
 This index accounts for variations in colonies abundance providing a robust metric that reflects the overall status of the penguin population.
-The index appears to exhibit fluctuations over time, with a general decline from the early 1990s to the mid-2000s, followed by a period of relative stabilization and another decline toward the end of the time series.
-This index captures broad decadal declines in predator presence, particularly after 2015, potentially reflecting changes in prey availability, environmental conditions, or predator demography.
+The index appears to exhibit fluctuations over time, with a general decline from the early 1990s to the mid-2000s, followed by a period of relative stabilization and a strong decline toward the end of the time series.
+This decline has been extensively documented reported in the Antarctic Peninsula by studies such as @Waterhouse2014, @Dunn2019a and @Wethington2023, primarily associated with colonies of the Adélie penguin [@Strycker2020].
+This abundance series captures broad decadal declines in predator presence, particularly after 2015, potentially reflecting changes in prey availability, environmental conditions, or predator demography.
 
@@
 A red horizontal line marks the 3.6 cm as recruit size for reference purposes.
-Middle panel: Spatial foraging locations and relative intensity by species across years.
-Colored circles indicate locations and intensity of sampling by species, with circle size suggesting sampling effort or frequency.
+Middle panel: Spatial foraging colonies and relative intensity by species across years.
+Circles indicate locations and intensity of sampling by species, with circle size suggesting sampling effort or frequency.
 Bottom panel: Synoptic index of relative predator biomass, used as an ecosystem input in the stock assessment model."}
@@
 ```
+
+Natural mortality (M) of Antarctic krill varied with age, exhibiting a sharp increase from age-0 (M = 0.30) to age-1 (M = 0.62), followed by a gradual asymptotic rise reaching approximately M = 0.76 from age-4 onward (S1.
+Fig x).
+This age-specific pattern suggests heightened vulnerability during the early life stages, potentially associated with increased predation pressure and lower physiological resilience.
+From age-1 onward, the progressive stabilization in mortality rates may reflect reduced susceptibility to size-selective predation and improved swimming capabilities.
+Notably, adult krill (ages 3–7) still experience considerable natural mortality (M ~ 0.75), suggesting sustained predation pressure.
+This is consistent with the dietary reliance of apex predators such as *Pygoscelis* penguins, which feed extensively on sub-adult and adult krill during the breeding season.
+The overlap between peak penguin chick-rearing periods and krill availability reinforces the ecological significance of these age classes in sustaining predator populations.
+Additionally, the extended presence of krill in predator diets across age classes supports the inference of sustained natural mortality beyond early ontogeny.
+
+Would you like to incorporate other predator groups like seals or whales into the analysis?
 
@@
 
+By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions.
+This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
+Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
+
+Data consistency
+
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
+However, integrating these complexities presents several challenges.
+Long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies.
+Additionally, some areas have sparse or incomplete datasets on krill dynamics.
+Furthermore, indirect drivers, such as environmental variability and predator abundance, introduce another layer of complexity.
+Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations.
+
 The use of integrated models for assessing krill populations in the Southern Ocean has posed significant challenges for CCAMLR, particularly in light of the inherent uncertainties and limitations associated with data-poor regions.
@@
 
+@Hollowed2000 demostrate a successful integration of predation mortality into Gulf of Alaska walleye pollock assessments showing that the value of predator data (e.g., diet composition, predator abundance) for refining natural mortality estimates and understanding ecosystem dynamics improve the estimations.
+Explicit modeling of key predators (e.g., arrowtooth flounder, Steller sea lions) improved biomass estimates by addressing previously unquantified mortality sources, particularly for younger age classes.
+
+However, reliance on predator data introduces challenges: (1) interannual and age-specific variability in natural mortality (especially for vulnerable stages) requires high-resolution trophic data; (2) dome-shaped selectivity patterns in predators may complicate mortality rate estimation; and (3) gaps in spatiotemporal coverage of predator diets can bias assessments.
+
+To enhance robustness, future work should prioritize: (1) expanded monitoring of predator diets across seasons, (2) dynamic selectivity models to account for predator behavior, and (3) ecosystem-based frameworks that explicitly link krill dynamics to predator demand and environmental shifts.
+While predator data significantly reduces uncertainty in mortality estimates, its efficacy depends on systematic data collection and integration into holistic management strategies.
+
+Inclusion of predator in this analisis coincide with @Moustahfid2009, where they found that explicitly including predation mortality in stock assessment models significantly alters the outputs.
+When predation was not accounted for, the model underestimated both the magnitude and uncertainty of spawning-stock biomass (SSB) and recruitment levels.
+
+(reflection)
+
+> Although penguin predation on krill does not reach the magnitude required to decimate the planktonic crustacean population—as might be the case with baleen whales [@Baines2022; @Biuw2024]—it is nevertheless important that stock assessment models incorporate signals regarding the trends of this factor on population variables.
+Since the index is relative rather than absolute, it provides valuable information for modeling krill population dynamics.
+
+> The substantial variability between consumption estimates presents challenges for model integration.
+While some studies suggest predators consume approximately 20% of the krill standing stock [@Baines2022], others indicate consumption levels around 20% of the total krill catch [@Biuw2024].
+This discrepancy in baseline metrics complicates comparative analysis and requires careful reconciliation when incorporating these estimates into integrated models.
+
 ### Stock-Recruitment and Biomass Estimates
@@
 The ecosystem-informed model reduces the occurrence of extreme recruitment outliers and yields a more consistent relationship between SSB and recruitment.
-[REF]
-
+[REF].
 Models that omit ecosystem components tend to produce higher and more uncertain projections, potentially overestimating stock productivity.
@@
 
+Like @Crone2019, the study concluded that both methods for including environmental data resulted in relatively high-quality estimates for a diverse group of output variables.
+The best performing models were those that included a correct and bias-corrected environmental variable, indicating the importance of accurately incorporating environmental factors in stock assessments
+
+Regarding to evironmental impact in modelliling of main population variables in stock assessment process, @Haltuch2009 recommends that low-frequency environmental variability significantly affects fish population abundance through recruitment, and biomass reference points should be based on average recruitment and/or dynamic B0, especially if comprehensive catch and survey data are available, y a su vez, la calidad y cantidad de datos mejorar la exactitud de resultados.
+
 Tabla summary
@@
 These findings support growing consensus that ecosystem-informed stock assessments, particularly those that integrate environmental covariates, provide more robust and ecologically realistic estimates for Antarctic krill and should inform precautionary management frameworks under CCAMLR.
+
+### Limitations of the Paper
+
+The study on krill population dynamics presents several limitations that should be acknowledged for a comprehensive understanding of its findings. Here are the key limitations identified:
+
+- Integrating multiple data sources and accounting for temporal and spatial variability in stock assessments presents significant challenges.
+The study highlights that long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies, which can affect the robustness of the results .
+
+- While the study emphasizes the importance of incorporating ecosystem variables into stock assessments, it also acknowledges that the models used may not fully capture all relevant interactions within the ecosystem.
+This could lead to biased estimations of krill population dynamics and stock health .
+
+- The paper mentions that spatial heterogeneity was not fully addressed in the modeling approach (implicit approach).
+Future research is suggested to incorporate spatially explicit models, which were not utilized in this study.
+This limitation may affect the understanding of how krill populations are distributed and managed across different regions .
+
+-  The retrospective analysis conducted in the study indicates patterns of bias in the models.
+However, the findings are based on specific scenarios, and the generalizability of these results to other contexts or conditions may be limited .
+
+In summary, while the study provides valuable insights into krill population dynamics, these limitations highlight the need for further research and improved methodologies to enhance the accuracy and applicability of stock assessments in the context of ecosystem-based management.
 

### 2025-05-06 index.Rmd
adding edits

@@
 library(ggrepel)
+library(broom)
+library(lmtest)
+library(flextable)
 ```
@@
 knitr::include_graphics('Figs/pathmod.png')
+```
+
+### Age Key Length (ALK)
+
+In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process
+
+```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
+alk_matrix <- base.model1.1$ALK[,,1] # Ajusta las dimensiones según tu matriz
+
+# Convertir la matriz en un data.frame
+alk_df <- as.data.frame(as.table(alk_matrix))
+
+# Renombrar columnas para mayor claridad
+colnames(alk_df) <- c("Length", "TrueAge", "Value")
+
+# Asegúrate de que las columnas Length y TrueAge sean numéricas
+alk_df$Length <- as.numeric(as.character(alk_df$Length))
+alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
+
+# Asegúrate de que la columna Value sea numérica
+alk_df$Value <- as.numeric(alk_df$Value)
+
+# Crear el plot usando ggplot
+ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
+  geom_tile() +
+  scale_fill_gradient(low = "white", high = "blue") +
+  labs(x = "True Age",
+       y = "Length",
+       fill = "Value") +
+  theme_bw()+
+  theme(legend.position = "none")
 ```
@@
                         forecast=T)
-
 ```
@@
 
+### Otra tab
+
+```{r eval=FALSE}
+cpue_resid <- bind_rows(
+  base.model1.1$cpue %>% mutate(model = "s1.1"),
+  base.model1.2$cpue %>% mutate(model = "s1.2"),
+  base.model1.3$cpue %>% mutate(model = "s1.3"),
+  base.model1.4$cpue %>% mutate(model = "s1.4")
+)
+
+indres <- ggplot(cpue_resid, aes(x = Exp, y = Obs, color = model)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
+  facet_wrap(~Fleet_name, scales = "free") +
+  theme_minimal() +
+  labs(title = "Observed vs Expected (Index)",
+       x = "Expected",
+       y = "Observed")
+
+len_resid <- bind_rows(
+  base.model1.1$lendbase %>% mutate(model = "s1.1"),
+  base.model1.2$lendbase %>% mutate(model = "s1.2"),
+  base.model1.3$lendbase %>% mutate(model = "s1.3"),
+  base.model1.4$lendbase %>% mutate(model = "s1.4")
+)
+
+lenres <- ggplot(len_resid, aes(x = Exp, y = Obs, color = model)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
+  facet_wrap(~Fleet, scales = "free") +
+  theme_minimal() +
+  labs(title = "Observed vs Expected (Index)",
+       x = "Expected",
+       y = "Observed")
+
+ggarrange(indres,
+          lenres,
+          ncol=1,
+          common.legend = T)
+```
+```{r eval=FALSE}
+
+cpue_df <- cpue_resid %>%
+  transmute(
+    Yr = Yr,
+    Obs = Obs,
+    Exp = Exp,
+    model = model,
+    Fleet_name = as.character(Fleet_name),  # Asegura que es character
+    type = "Index"
+  )
+
+len_df <- len_resid %>%
+  transmute(
+    Yr = Yr,
+    Obs = Obs,
+    Exp = Exp,
+    model = model,
+    Fleet_name = as.character(Fleet),  # Lo convertimos para que combine bien
+    type = "Length"
+  )
+
+resid_all <- resid_all %>%
+  mutate(Fleet_name = as.character(Fleet_name),
+         Fleet_name = case_when(
+           Fleet_name == "1"  ~ "FISHERYBS",
+           Fleet_name == "2"  ~ "FISHERYEI",
+           Fleet_name == "3"  ~ "FISHERYGS",
+           Fleet_name == "4"  ~ "FISHERYJOIN",
+           Fleet_name == "5"  ~ "FISHERYSSIW",
+           Fleet_name == "6"  ~ "SURVEYBS",
+           Fleet_name == "7"  ~ "SURVEYEI",
+           Fleet_name == "8"  ~ "SURVEYGS",
+           Fleet_name == "9"  ~ "SURVEYJOIN",
+           Fleet_name == "10" ~ "SURVEYSSIW",
+           Fleet_name == "11" ~ "PREDATOR",
+           TRUE ~ Fleet_name  # mantiene los nombres ya correctos
+         ))
+
+ggplot(resid_all, aes(x = Exp, y = Obs, color = model)) +
+  geom_point(alpha = 0.6) +
+  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
+  facet_grid(type ~ Fleet_name, scales = "free") +
+  theme_minimal(base_size = 13) +
+  labs(title = "Observed vs Expected residuals",
+       x = "Expected", y = "Observed")
+
+```
+```{r}
+ggplot(resid_all, aes(x = Obs - Exp, fill = model)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(Fleet_name~type, scales = "free") +
+  labs(title = "Distribución de residuales", x = "Residual (Obs - Exp)", y = "Densidad") +
+  theme_minimal()
+
+```
+### Statistics
+
+```{r eval=FALSE, resumen_residuales, message=FALSE, warning=FALSE}
+resid_all <- resid_all %>%
+  mutate(residual = Obs - Exp)
+
+# Estadísticas básicas
+resid_stats <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    N = n(),
+    Mean = mean(residual, na.rm = TRUE),
+    SD = sd(residual, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Shapiro-Wilk (normalidad)
+shapiro_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    shapiro_p = ifelse(n() > 3, shapiro.test(residual)$p.value, NA),
+    .groups = "drop"
+  )
+
+# Ljung-Box (autocorrelación temporal)
+ljung_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    ljung_p = ifelse(n() > 10, Box.test(residual, lag = 10, type = "Ljung-Box")$p.value, NA),
+    .groups = "drop"
+  )
+
+# Breusch-Pagan (heterocedasticidad)
+bp_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    bp_p = ifelse(n() > 3, bptest(lm(residual ~ Yr, data = cur_data()))$p.value, NA),
+    .groups = "drop"
+  )
+
+# Unión de todo
+tabla_resumen <- resid_stats %>%
+  left_join(shapiro_p, by = c("type", "model")) %>%
+  left_join(ljung_p, by = c("type", "model")) %>%
+  left_join(bp_p, by = c("type", "model")) %>%
+  mutate(across(where(is.numeric), ~ round(.x, 3)))
+
+# Mostrar tabla
+flextable(tabla_resumen) %>%
+  set_caption("Tabla 1. Estadísticos y pruebas sobre residuales por tipo y modelo") %>%
+  autofit()
+
+```
+
 ## Outputs
@@
 
-### AKL
-
-In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process
-
-```{r AKL, out.width='60%',  fig.cap="Representation of ALK Matrix to krill in 48.1"}
-alk_matrix <- base.model1.1$ALK[,,1] # Ajusta las dimensiones según tu matriz
-
-# Convertir la matriz en un data.frame
-alk_df <- as.data.frame(as.table(alk_matrix))
-
-# Renombrar columnas para mayor claridad
-colnames(alk_df) <- c("Length", "TrueAge", "Value")
-
-# Asegúrate de que las columnas Length y TrueAge sean numéricas
-alk_df$Length <- as.numeric(as.character(alk_df$Length))
-alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))
-
-# Asegúrate de que la columna Value sea numérica
-alk_df$Value <- as.numeric(alk_df$Value)
-
-# Crear el plot usando ggplot
-ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
-  geom_tile() +
-  scale_fill_gradient(low = "white", high = "blue") +
-  labs(x = "True Age",
-       y = "Length",
-       fill = "Value") +
-  theme_bw()+
-  theme(legend.position = "none")
-```
-
 ### Code Repository

### 2025-05-12 MS2_Krill.Rmd
adding .bib refs

@@
 Given the difference en the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
-The model considers interannual fluctuations of growth dynamics, maturity and reproduction, with a spawning pulse per year occurring in January-February [@Perry2019; @McBride2021].
-While strata are modeled as separate fleets, the population dynamics—such as recruitment and growth—remain unified, representing the krill stock as a single, interconnected population.
-This disaggregated structure enables the model to account for differences in small scale spatial dynamics and providing a flexible yet informative way to include spatial processes in stock assessments.
-Areas-as-fleet approach captures regional variability in dynamic and observation while avoiding the complexity of modeling explicit movement.
-This approach provides a more realistic representation of krill productivity within FAO Subarea 48.1.
-
-## Length-to-Age Model Implementation
-
-The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21.
-SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
-The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources described previously and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population.
-All analysis were executed in R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
-
-Like many invertebrates, krill present challenges for age determination [@Punt2013].
-Historically, aging techniques have not been available for this species, although recent advancements have been made [@Kilada2017].
-As a result, stock assessments have relied on length-to-age models with age dynamics approximated through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length.
-Integrated models can effectively capture the age structure of such populations by transforming length observations into population-level dynamics.
-Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
-Length data are generally more abundant and cost-effective to collect [@Canales2021, @Chong2019], and, due to the correlation between length and age, they provide valuable information about population structure.
-To date, all integrated krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021].
-Stock assessment models often rely on Age-Length Keys (ALKs) to infer the age structure of a population from observed length-frequency data [@ICCA2003].
-An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes [@Lee2024].
-ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases.
-This approach allows the transformation of catch-at-length data into age-structured information, which is essential for age-based modeling frameworks.
+This disaggregated structure allows the model to reflect small-scale spatial dynamics while maintaining analytical simplicity.
+The areas-as-fleets approach captures regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019].
+The model accounts for interannual variability in growth dynamics, maturity, and reproduction, with a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021].
+Although strata are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single, interconnected population providing a more realistic representation of krill dynamics population within FAO Subarea 48.1.
+
+## Krill Length-to-Age Model Description
+
+Like many invertebrates, krill is a hard to age determination, and this presents challenge for modeling population dynamics, because most modern stock assessments that aim to estimate marine population variables rely on age-structured modeling approaches [@Punt2013].
+Although recent advancements have been made [@Kilada2017], aging techniques have historically not been available for krill.
+Therefore, to develop a stock assessment model for this type of species use length data to approximate age dynamics through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length [@Laslett2004a; @Lee2019; @Lee2024; @Rudd2021; @Zhang2022].
+Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].
+On the other hand, length data, which are cost-effective and readily available [@Canales2021; @Chong2019], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014].
+For made this transformation, the models often use an Age-Length Keys (ALKs).
+An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classed and this method assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024].
 The ALK matrix is typically defined as:
@@
 where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age.
-In integrated stock assessment models, such as those using statistical catch-at-age (SCAA) or integrated length-based frameworks, the ALK provides a critical link between observable length data and underlying age-based population dynamics.
-
-The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK.
-The majority of ALKs are based on growth parameters according to the most suitable formula for the species.
-In the case of krill, these parameters are derived from the von Bertalanffy growth curve.
-In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV.
-This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomass with model estimates of vulnerable biomass.
-Weight-at-length was assumed known and calculated using @Maschette2020 parameters in krill resulting in a AKL exposed in S1 Fig 7.
-
-Growth was modeled following a von Bertalanffy growth function with lenght-at-age parameterization.
+Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
+To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021; @Green2023].
+The resulting ALK is shown in S1 Fig 7.
+
+In the stock assessment catch-at-length model, growth is modeled using a von Bertalanffy growth function lenght-at-age parameterization with parameters L~inf~, k, and coefficient of variation (CV), as outlined in @Mardones2023.
+This growth relationship is integrated into the model alongside weight-at-length, which is assumed known and calculated using the parameters from @Maschette2020 for krill.
 The model assumed linear growth below a predefined settlement age of 1.5 years.
@@
 Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
-Maturity was modeled as an age-based logistic function, with the first age of maturity set at 3 years.
+Maturity was modeled as an age-based logistic function, with the first age of maturity set at 2-3 years.
 Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
@@
 The model used a direct parameterization without additional offsets between sexes.
-Under this assumption, male parameters with a value of 0 and a negative estimation phase were set equal to their corresponding female parameters.
-Regarding selectivity functions, it was modeled using a length-based logistic formulation for the fishery, survey, and predator fleets.
-The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95% width).
+Regarding selectivity functions, it was modeled using a logistic formulation for the fishery fleet.
+The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95%).
 The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
-The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata.
-Parameter estimation phases were assigned, with negative values indicating fixed parameters.
+The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata or "fleet".
 The survey selectivity was also modeled using a length-based logistic function.
@@
 The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
-These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (Fig 6, S1).
+These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (S1 Fig 6).
 The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
@@
 Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt formulation, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
-This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@REF].
-The unfished recruitment ($R_0$) was estimated with a log-transformed mean recruitment value of 23, with a prior of 10.
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019a, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
+This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019a; @Perry2020].
+The unfished recruitment ($R_0$) was estimated with a mean recruitment of 23 in log-scale, with a prior of 10.
 The parameter was freely estimated in phase 1.
@@
 To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-Cyclic patterns in recruitment were assumed each tree and five years [@REF], and advanced recruitment options were enabled for model flexibility.
-All of these configurations can be reviewed in the repository section of S1.
-
-## Data Sources
-
-A set of multiple sources of data were incorporated into the krill population dynamics model to improve inference and parameter estimation.
-These data included register from the krill fishery monitoring program, such as standardized CPUE (kg/time haul) and length compositions, which have been systematically collected on board fishing vessels by scientific observers through the "Scheme of International Scientific Observation" (SISO) Program.
-Scientific research cruises conducted under the AMLR program provided critical data such as krill biomass indices by strata and biological characteristics of krill obtained via net sampling.
-Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) were handled according to stock assessment model templates requirements (Suppl.
+Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019a; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility.
+Life history parameters—such as growth, weight-length relationship, natural mortality, and maturity—were used as priors to model the initial conditions of krill population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
+These parameters, along with initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1).
+
+```{r parainit1, eval=TRUE}
+dir1.1 <- here("s1.1")
+
+start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
+dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
+ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
+                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
+fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)
+
+# Extraer parámetros
+parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
+rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
+                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")
+
+SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
+Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
+Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]
+
+# Combinar parámetros
+parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))
+
+# Añadir nombres de fila como columna "Parameter"
+parInit$Parameter <- rownames(parInit)
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(INIT  = round(INIT, 2),
+         PRIOR = round(PRIOR, 2))
+
+# Crear vector de categorías (39 filas)
+categories <- c(rep("Natural Mortality", 1),
+                rep("Growth", 4),
+                rep("Length-Weight Relation", 2),
+                rep("Maturity", 2),
+                rep("Stock-Recruit Relation", 5),
+                rep("Catchability", 2),
+                rep("Selectivity", 21))  # OJO: corregido a 23
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(Category = categories) %>%
+  dplyr::relocate(Category, Parameter)
+
+# Crear flextable
+ft <- flextable(parInit)
+ft <- set_caption(ft, "Input parameters for the initial SS3 krill model")
+ft <- autofit(ft)
+ft <- theme_booktabs(ft)
+ft <- fontsize(ft, size = 9, part = "all")
+ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")
+
+# Mostrar flextable
+ft
+
+```
+
+The statistical model to krill was configured using Stock Synthesis v 3.30.21 (SS3 hereafter) [@Methot2013].
+SS3 is a structured stock assessment model, in the class of models called *"Integrated stock assessment model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
+The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
+Model conditioning is available in the code repository section of Suppl.
+Mat 1.
+All analysis about modeling was executed in R-CRAN [@R-base] and diagnoses and visualization of the outputs was made was through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
+
+## Environmental variable modelling
+
+To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables described previously as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
+Table \@ref(tab:env) summarizes the key characteristics of the environmental datasets used in the analysis, including sea ice, sea surface temperature (SST), and chlorophyll a.
+These variables were obtained from different satellite platforms and reanalysis products, namely Nimbus-7, ERA5, and Copernicus.
+The table details spatial and temporal resolution, as well as the temporal coverage of each dataset.
+
+```{r env, eval=TRUE}
+env_features <- data.frame(
+  Feature = c("Sea Ice", "SST", "Chlorophyll a"),
+  Platform = c("Nimbus-7", "ERA5", "E.U. Copernicus Marine Service"),
+  Spatial_Resolution = c("25 × 25 km", "0.25º × 0.25º", "0.25º × 0.25º"),
+  Temporal_Resolution = c("Daily", "Monthly", "Daily-Monthly"),
+  Time_Range = c("1978 - Present", "1991 - Present", "2000 - Present"),
+  stringsAsFactors = FALSE
+)
+
+ft_env <- flextable(env_features) |>
+  set_header_labels(
+    Feature = "Feature",
+    Platform = "Platform",
+    Spatial_Resolution = "Spatial Resolution",
+    Temporal_Resolution = "Temporal Resolution",
+    Time_Range = "Time Range"
+  ) |>
+  set_caption("Summary of environmental variables used in correlation analysis") |>
+  add_footer_lines("doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148") |>
+  color(part = "header", color = "black") |>
+  theme_booktabs() |>
+  fontsize(size = 10) |>
+  set_table_properties(layout = "autofit", width = .85)
+ft_env
+```
+
+All environmental data were processed at their original spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations.
+We modelled the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power.
+The initial model considered only the effect of strata (spatial), while subsequent models sequentially included Chla, SST, SIC and interaction terms.
+This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process.
+All models and their results can be found in the Supplementary Material 2.
+
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [REF] and SS3 addresses these approaches.
+According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment.
+This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;
+
+$$
+E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
+$$
+$R$ represents the variability of deviations, adding to the variance caused by environmental factors.
+$SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable.
+Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
+However, the model does not automatically adjust for this.
+Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment.
+This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability.
+By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics esttimates.
+
+## Predator modelling
+
+Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics.
+In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified, located along the northern arc of the Antarctic Peninsula during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)).
+Predator abundance is processed as an index of predation "effort" allowing time-varying estimation of predation mortality rates in time.
+This follows the equation:
+
+$$
+M2_t = Q \cdot I_t
+$$
+
+where \( M2_t \) is the predation mortality in year *t*, \( I_t \) is the predator abundance index, and \( Q \) is a catchability coefficient.
+Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation.
+These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size.
+Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions.
+Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality.
+This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.
+
+## Model Input Data: Fishery and Survey
+
+Along with data on predators and environmental conditions, a fundamental input to the model was derived from fishery-dependent data and scientific monitoring programs.
+Fishery-dependent data were obtained from the Krill Fishery Monitoring Program under the CCAMLR SISO framework, including total reported catches, standardized CPUE (adjusted by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm.
+Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul.
+This ensured consistency across years and strata and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020.
+Fishery-independent data originated from U.S.
+AMLR Program acoustic surveys included standarized biomass indices conducted between 1991 and 2011.
+Krill length compositions from AMLR surveys (provided by the Ecosystem Science Division of NOAA) was manipulate in the same structure that fishery-dependent data and handled according to stock assessment model templates requirements (Suppl.
 Mat.
 2).
-Environmental and predator-related data were also incorporated, such as satellite-derived chlorophyll-a concentrations from the Copernicus Marine Service and predator abundance and diet composition, particularly from penguin colonies reported in the [MAPPPD database](https://www.penguinmap.com/mapppd/sources/).
-A summary of all data sources, including temporal coverage and spatial scale, is provided in Table \@ref(tab:data-sources).
+All these data, together with the previously described ecosystem variables (environmental and predator-related), along with their temporal coverage and spatial scale, are summarized in Table \@ref(tab:data-sources) and illustrated in in S1, Fig.
+4.
 
@@
     "Krill Fishery Monitoring Program (CCAMLR SISO)",
-    "AMLR Scientific Research Cruises",
+    "AMLR Scientific Research Program",
     "Satellite Remote Sensing",
@@
   set_table_properties(layout = "autofit", width = .85)
-
 ft_data
-
-```
-
-Life history parameters, like growth, weight-length relation, natural mortality and maturity were used as priors to model initial condition in krill dynamics population was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit1).
-Source data and temporal scale used in this stock assessment are shown in Fig 4, S1.
-
-```{r parainit1, eval=TRUE}
-dir1.1 <- here("s1.1")
-
-start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
-dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
-ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
-                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
-fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)
-
-# Extraer parámetros
-parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
-rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
-                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")
-
-SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
-Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
-Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]
-
-# Combinar parámetros
-parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))
-
-# Añadir nombres de fila como columna "Parameter"
-parInit$Parameter <- rownames(parInit)
-
-# Añadir categorías y reordenar columnas
-parInit <- parInit %>%
-  mutate(INIT  = round(INIT, 2),
-         PRIOR = round(PRIOR, 2))
-
-# Crear vector de categorías (39 filas)
-categories <- c(rep("Natural Mortality", 1),
-                rep("Growth", 4),
-                rep("Length-Weight Relation", 2),
-                rep("Maturity", 2),
-                rep("Stock-Recruit Relation", 5),
-                rep("Catchability", 2),
-                rep("Selectivity", 21))  # OJO: corregido a 23
-
-# Añadir categorías y reordenar columnas
-parInit <- parInit %>%
-  mutate(Category = categories) %>%
-  dplyr::relocate(Category, Parameter)
-
-# Crear flextable
-ft <- flextable(parInit)
-ft <- set_caption(ft, "Input parameters for the initial SS3 krill model")
-ft <- autofit(ft)
-ft <- theme_booktabs(ft)
-ft <- fontsize(ft, size = 9, part = "all")
-ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")
-
-# Mostrar flextable
-ft
-
-```
-
-## Environmental variable modelling
-
-To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables previously recognized as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
-Table \@ref(tab:env) summarizes the key characteristics of the environmental datasets used in the analysis, including sea ice, sea surface temperature (SST), and chlorophyll a.
-These variables were obtained from different satellite platforms and reanalysis products, namely Nimbus-7, ERA5, and Copernicus.
-The table details spatial and temporal resolution, as well as the temporal coverage of each dataset.
-
-```{r env, eval=TRUE}
-env_features <- data.frame(
-  Feature = c("Sea Ice", "SST", "Chlorophyll a"),
-  Platform = c("Nimbus-7", "ERA5", "E.U. Copernicus Marine Service"),
-  Spatial_Resolution = c("25 × 25 km", "0.25º × 0.25º", "0.25º × 0.25º"),
-  Temporal_Resolution = c("Daily", "Monthly", "Daily-Monthly"),
-  Time_Range = c("1978 - Present", "1991 - Present", "2000 - Present"),
-  stringsAsFactors = FALSE
-)
-
-ft_env <- flextable(env_features) |>
-  set_header_labels(
-    Feature = "Feature",
-    Platform = "Platform",
-    Spatial_Resolution = "Spatial Resolution",
-    Temporal_Resolution = "Temporal Resolution",
-    Time_Range = "Time Range"
-  ) |>
-  set_caption("Summary of environmental variables used in correlation analysis") |>
-  add_footer_lines("doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148") |>
-  color(part = "header", color = "black") |>
-  theme_booktabs() |>
-  fontsize(size = 10) |>
-  set_table_properties(layout = "autofit", width = .85)
-ft_env
-```
-
-All environmental data were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations.
-We model the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power.
-The initial model considered only the effect of strata (spat), while subsequent models sequentially included Chla, SST, Sea Ice Index and interaction terms.
-This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process.
-All models and their results can be found in the Supplementary Material 2.
-
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [REF] and SS3 addresses these approaches.
-According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment.
-This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;
-
-$$
-E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-$$
-$R$ represents the variability of deviations, adding to the variance caused by environmental factors.
-$SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
-However, the model does not automatically adjust for this.
-Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment.
-This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability.
-By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics esttimates.
-
-## Predator modelling
-
-Following the SS3 framework described by @methot2020stock, natural mortality due to predation (*M2*) can be integrated into stock assessment models through multiple data types.
-In our krill assessment, we implement this by incorporating both predator biomass and diet composition from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*).
-Predator abundance is treated as an index of predation "effort," allowing time-varying estimation of predation mortality rates.
-This follows the equation:
-
-$$
-M2_t = Q \cdot I_t
-$$
-
-where \( M2_t \) is the predation mortality in year *t*, \( I_t \) is the predator abundance index, and \( Q \) is a catchability coefficient.
-Additionally, length-frequency data of krill obtained from penguin diet samples are used to characterize the size-selectivity of predation.
-These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size.
-This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.
-Total natural mortality (*M*) is thus partitioned as \(M = M_1 + \sum M_2 \), where \( M_1 \) is background mortality and \( \sum M_2 \) accounts for predator-specific contributions.
-Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality.
+```
 
@@
 
-To account for the uncertainty associated with ecosystem influence in krill dynamic popupation, we evaluated a set of model configurations that incorporate variations in predator "fleet" and environmental effects, and the relationship between spawning biomass and recruitment.
-Spatial heterogeneity implicity is considering in all scenarios.
+We evaluated a set of model configurations that incorporate variations in predator "fleet" and environmental effects, and the impacts in spawning biomass, recruitment among others.
+Spatial heterogeneity was considered in all scenarios.
 The methodology involves assessing the statistical performance of the model under these configurations.
-Specifically, `s1.1` serves as the reference model, incorporating only fishery and survey data while accounting for spatial heterogeneity.
+Specifically, `s1.1` serves as the reference model, incorporating only fishery and survey data while accounting for spatial heterogeneity (implicit spatial model).
 The subsequent scenarios build upon this configuration to evaluate the effects of additional ecological components.
 `s1.2` extends `s1.1` by incorporating predator influences, allowing the model to assess the impact of predation on krill population dynamics.
-`s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers affecting krill recruitment and biomass.
-Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more comprehensive assessment of ecosystem interactions and their role in krill population dynamics (Table \@ref(tab:scenario)).
+`s1.3` modifies `s1.1` by introducing an environmental variable, capturing potential environmental drivers.
+Finally, `s1.4` integrates both predator and environmental influences into the `s1.1` framework, providing a more realistic assessment of ecosystem interactions and their role in krill population dynamics (Table \@ref(tab:scenario)).
 
@@
 
-Regarding the performance and comparison between scenarios, @Carvalho2017;  @Carvalho2021b; @Merino2022a and @Kell2021 propose good practice to made a structured approach for diagnosing stock assessment models to ensure their reliability and robustness.
-This diagnostic process was applied to the krill stock assessment and involved several key steps propoesed by previus research.
+Regarding performance and scenario comparison, @Carvalho2017;  @Carvalho2021b; @Merino2022a and @Kell2021 propose a set of "good practices" for conducting a structured diagnosis to ensure the reliability and robustness of stock assessment analyses.
+This diagnostic process was applied to the krill stock assessment and involved several key steps proposed by previous research.
 First, convergence analysis was conducted to evaluate whether the estimation algorithm reached a stable solution.
@@
 Next, model fit to data was assessed by comparing observed and predicted values across key data sources, including catch, survey indices, and length composition.
-The p-values for the residual runs tests were computed for each of the four joint length compositions that are common in all models.
-The p-values were then combined into a single test statistic using Fisher's method @@Merino2022a:
+The *p-values* for the residual runs tests were computed for each of the four joint length compositions that are common in all models.
+The *p-values* were then combined into a single test statistic using Fisher's method @@Merino2022a:
 
@@
 
-where \( p_i \) is the p-value for Length \( i \) and \( k \) are the degrees of freedom of the four p-values from the joint Length indices.
+where \( p_i \) is the p-value for Length \( i \) and \( k \) are the degrees of freedom of the four *p-values* from the joint length compositions.
 Likelihood profiling was also carried out to quantify parameter uncertainty and identify the influence of priors.
-Finally, prediction skill was assessed using hindcasting techniques to evaluate the model’s ability to reproduce historical stock dynamics.
-Retrospective analysis and hindcast cross-validation were performed by sequentially removing five years of data.
-regarding @HurtadoF2015, Mohn parametrer was calculated as the mean relative error, defined as:
+Regarding model fit, Pearson residuals of predicted length distributions for krill were evaluated across four modeling scenarios, each incorporating different levels of ecosystem complexity.
+Pearson residuals are used to assess the goodness-of-fit by standardizing the differences between observed and predicted values, allowing for the detection of systematic patterns or biases in model predictions across size classes and sampling sources.
+Retrospective analysis and hindcast cross-validation were performed by sequentially removing five years of data regarding @HurtadoF2015.
+As part of the restrospective analysis, the Mohn parametrer was calculated as the mean relative error, defined as:
 
@@
 
-where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which ρM was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset.
+where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which \rho_M was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset.
 Year \( y \) was nominally considered as season one, \( T \) refers to the terminal assessment data year, and \( h \) denotes the number of retrospective peels, removing recent years of data ("peeling") and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
@@
 The RMSE values were extracted from the variance tuning diagnostics of each scenario and compiled into a data frame for statistical analysis.
-To test for overall differences among scenarios, a one-way ANOVA [@REF] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
-Another quantitative statistical method used to evaluate model performance was the Akaike Information Criterion (AIC)[REF], computed as:
-
-$$
-\text{AIC} = 2k + 2L
-$$
-
-where \( k \) is the number of estimated non-deviation parameters and \( L \) is the negative log-likelihood.
+To test for overall differences among scenarios, a one-way ANOVA [@Sthle1989] was conducted, treating RMSE values as the dependent variable and scenario configurations as the categorical factor.
+Another quantitative statistical method used to evaluate model performance was the Akaike Information Criterion (AIC) [@Bozdogan1987].
 This criterion balances model fit and complexity, facilitating comparisons across scenarios.
@@
     "@Hyndman2006",
-    "@Akaike1974",
-    "@Mohn1999",
+    "@Bozdogan1987",
+    "@HurtadoF2015",
     "@Methot2013"
@@
 Based on linear models performances and ecological relevance, Chla emerged as the most influential predictor, therefore, this variable was selected for incorporation into the stock assessment model as an environmental covariate (Table 2, S2).
-This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
-Figure \@ref(fig:mapachl) presents a time series of chlorophyll-a (Chl-a) concentration (mg/m-3) in the waters around the Antarctic Peninsula from 2000 to 2020.
+Figure \@ref(fig:mapachl) presents a time series of Chl-a concentration (mg/m-3) in the waters around the Antarctic Peninsula from 2000 to 2020.
 Each panel represents a different year, showing spatial variations in Chl-a levels, which indicate phytoplankton biomass.
-The highest concentrations are observed along the coastal regions, while offshore areas exhibit lower Chl-a levels.
-This visualization helps assess interannual variability in primary productivity, which is a key driver of the Antarctic krill population dynamics.
-
-```{r mapachl, eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in 48.1 area"}
+The highest concentrations were observed along the coastal regions, while offshore areas exhibit lower Chl-a levels.
+
+```{r mapachl, eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in Subarea 48.1"}
 knitr::include_graphics('Figs/chla_grilled.png')
@@
 
-In another hand, Figure \@ref(fig:chlindex) illustrates the interannual variability of a Chlorophyll Variability Index from the year 2000 to 2020.
-The figure shows notable fluctuations across years, highlighting periods of higher-than-average chlorophyll variability (e.g., 2003, 2016) as well as years with significantly lower values (e.g., 2011, 2020).
-These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
+The interannual variability of this environmental index from 2000 to 2020 is shown in Figure \@ref(fig:chlindex).
+The figure shown illustrates the interannual variability of a Chlorophyll Variability Index from the year 2000 to 2020.
+The figure shows fluctuations across years, highlighting periods of higher-than-average chlorophyll variability (e.g., 2003, 2016) as well as years with significantly lower values (e.g., 2011, 2020).
 
@@
 
-The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lukcas!, REF Vicky].
 Figure \@ref(fig:mapapre) presents a temporal and spatial overview of penguin variables from 2001 to 2020.
-The top panel in Figure \@ref(fig:mapapre) shows annual boxplots of krill length distributions derived from penguin diet samples, stratified by predator species.
-Although median krill lengths remain relatively consistent over time, there is notable interannual variability in the spread and distribution of sizes, which may reflect fluctuations in recruitment strength and availability of krill age classes.
-Foraging activity is concentrated in the Bransfield Strait and surrounding areas, aligning with regions of persistent krill aggregation.
-This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics (middle panel in Figure \@ref(fig:mapapre)).
-For stock assessment purposes, we integrated all penguin colony abundances into a synoptic index of relative predator biomass, which represents a comprehensive measure of population trends over time (bottom panel in Figure \@ref(fig:mapapre).
+The top panel in Figure \@ref(fig:mapapre) shows annual krill length distributions derived from penguin diet samples, displayed by predator species.
+Although median krill lengths remain relatively consistent over time, there is an interannual variability in the spread and distribution of sizes.
+Spatiotemporal distribution patterns of the three penguin species across the Antarctic Peninsula showed interannual variability from 2001 to 2020.
+Adélie penguins exhibited a relatively consistent presence across the northern sector of the Peninsula, with larger colony sizes concentrated toward the southwestern Bransfield Strait, particularly during the early 2000s.
+Chinstrap penguins were widespread and dominant in both spatial extent and colony size across most years, especially in the central and northeastern portions of the Peninsula.
+Gentoo penguins appeared in fewer and more localized sites but displayed a slight increase in spatial occurrence in the latter part of the time series (middle panel in Figure \@ref(fig:mapapre)).
+For stock assessment purposes, we integrated all penguin colony abundances into a synoptic index of relative predator biomass, which represents a comprehensive measure of population abundance trends over time (bottom panel in Figure \@ref(fig:mapapre).
 This index accounts for variations in colonies abundance providing a robust metric that reflects the overall status of the penguin population.
@@
 
-The krill population dynamics model was evaluated under four increasingly complex scenarios.
 Scenario s1.1 represented the baseline spatial implicit model, incorporating only survey and fishery data, and resulted in the estimation of 71 parameters.
@@
 All model scenarios successfully converged within acceptable tolerance thresholds, with final gradients below 1.00e-04, and positive-definite Hessian matrices confirming robust parameter estimation and local identifiability.
-
-Regarding model fit, Pearson residuals of predicted length distributions for krill were evaluated across four modeling scenarios, each incorporating different levels of ecosystem complexity.
-Pearson residuals are used to assess the goodness-of-fit by standardizing the differences between observed and predicted values, allowing for the detection of systematic patterns or biases in model predictions across size classes and sampling sources.
 
@@
 
+* **Big summary / importance statement for the science**
+
+  * Expón de manera destacada la importancia científica de tus hallazgos generales.
+
+* **Scientific findings as extension of previous work**
+
+  * Relaciona tus hallazgos con trabajos previos y explica cómo los extienden o complementan.
+
+* **Findings' scientific implications**
+
+  * Detalla las implicancias científicas específicas de los resultados obtenidos.
+
+* **How this engineering will facilitate future research**
+
+  * Explica cómo tu enfoque metodológico o técnico puede ayudar a futuras investigaciones.
+
+* **Forward-looking statements about the field as a whole**
+
+  * Incluye reflexiones sobre cómo evolucionará el campo en general a partir de tu trabajo.
+
+* **A summary / importance of the engineering aspect of the work**
+
+  * Resume la importancia de la parte técnica o de ingeniería de tu estudio.
+
+* **How it facilitates scientific research**
+
+  * Describe cómo tus métodos o resultados facilitan la investigación científica práctica.
+
+* **Limitations of the platform shown with questions that would be unanswered**
+
+  * Reconoce limitaciones del estudio y plantea preguntas que aún quedan por responder.
+
 By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions.
@@
 ### Predation and Natural Mortality
+
+> The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lukcas!, REF Vicky].
+
+> variability length; , which may reflect fluctuations in recruitment strength and availability of krill age classes.
+
+> . This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics
 
@@
 ### Ecosystem Variables: Ecological Relevance
+
+> regarding clorofila como driver; This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
+
+>  About index; These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
 

### 2025-05-16 MS2_Krill.Rmd
close work until..

@@
 In this study, we implement an integrated length-to-age model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
-We explored multiple scenarios that accounted for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species.
-Performance analyses reveal systematic differences across scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables.
+We explored four scenarios that accounted (or not) for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species.
+Performance analyses reveal systematic differences across those scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables.
 Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics.
-Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through post hoc analyses.
+Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses.
 Overlooking these drivers may result in less accurate conclusions, hence limiting the effectiveness of management recommendations.
-Integrated modeling approaches that draw on extensive biological, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
+Integrated modeling approaches that draw on extensive biological, ecologicl, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
+
+or
+
+Understanding the population dynamics of Antarctic krill (Euphausia superba) is essential for achieving ecosystem-based management objectives in the Western Antarctic Peninsula (WAP), particularly under the mandate of CCAMLR.
+In this study, we implement an integrated length-to-age assessment model designed to assimilate multiple data sources related to krill biology, fishery operations, and ecological interactions.
+A key feature of our approach is the explicit inclusion of ecosystem drivers—namely, environmental variability and predation pressure—as intrinsic components of the population dynamics.
+We evaluate four model scenarios, with and without ecosystem variables, and assess their performance in estimating recruitment, fishing mortality (F), total biomass, and spawning stock biomass (SSB).
+The comparison reveals systematic biases in models that omit ecological information, resulting in overestimated recruitment and biomass levels.
+In contrast, scenarios incorporating predator and environmental data yield more precautionary and biologically realistic outcomes, suggesting greater alignment with observed ecosystem variability.
+Despite ongoing challenges related to the availability, resolution, and integration of ecological data, we argue that environmental and trophic drivers should be embedded directly into the state-space structure of assessment models rather than treated as exogenous factors.
+Failing to do so may obscure key population processes and compromise the reliability of management advice.
+Our results underscore the potential of ecosystem-informed, integrated modeling frameworks to support robust, adaptive management strategies that align with CCAMLR’s conservation objectives, while also highlighting the continued need for methodological development and interdisciplinary data collection efforts in the Southern Ocean.
 
@@
 The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population (Figure \@ref(fig:pearson)).
-In scenario s1.1 (Without Ecosystem Variables), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
+In scenario s1.1, substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
 Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
-s1.2 (With Predator) shows some improvement in model fit, with slightly more balanced residuals.
+s1.2  shows some improvement in model fit, with slightly more balanced residuals.
 However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
-In s1.3 (With Environment), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
+In s1.3, there is a reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
 This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
-Finally, s1.4 (With Predator and Ecosystem) shows the most consistent reduction in Pearson residuals across time and size structure.
+Finally, s1.4 shows the most consistent reduction in Pearson residuals across time and size structure.
 The residual patterns also indicate potentially concerning model misfits, particularly among adult individuals in several survey fleets operating in the Bransfield Strait, as well as among the largest size classes in the predator fleet, which likely represents penguin consumption data.
@@
 ```{r pearson, eval=TRUE, fig.cap= "Pearson residuals of predicted versus observed length-frequency distributions of Antarctic krill across four modeling scenarios (s1.1 to s1.4), each incorporating increasing levels of ecosystem complexity.
-Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components (Red = positive, Black = negative"}
-knitr::include_graphics('Figs/Pearson.png')
+Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components (BS=Bransfield Strait, EI= Elephant Island, GS= Gerlache Strait, JOIN= Joinville Island, SSWI= South West)"}
+knitr::include_graphics('Figs/Pearson2.png')
 ```
@@
 
-Retrospective analyses were conducted to evaluate model performance across four scenarios incorporating different levels of ecosystem complexity (Figure \@ref(fig:rhoplot)).
-The retrospective pattern for fishing mortality (F) showed a marked increase in bias in scenario s1.4, which included both predator and environmental components, particularly in the terminal year.
+The retrospective pattern for fishing mortality (F) showed a marked increase in bias in scenario s1.2 and s1.4, which included predator components, particularly in the 2017 (Figure \@ref(fig:rhoplot)).
 In this scenario, the combined rho value exceeded 3.5, indicating a strong retrospective pattern.
@@
 In contrast, the retrospective pattern for spawning stock biomass (SSB) was consistently negative across all scenarios and years.
-However, models that included environmental components—s1.3 and s1.4—tended to show larger negative biases in SSB, with scenario s1.4 again exhibiting the most pronounced pattern.
+However, models that included environmental components—s1.3 and s1.4—tended to show larger negative biases in SSB, with scenarios s1.2 and s1.4 again exhibiting the most pronounced pattern.
 The combined rho values for SSB in s1.3 and s1.4 were close to –0.5 and –0.6, respectively, whereas s1.1 showed a comparatively milder retrospective trend.
@@
 
-Log-likelihood profiles were examined for each model scenario (s1.1 to s1.4) to assess how different components contributed to the overall model fit.
-Across all scenarios, the TOTAL log-likelihood is primarily influenced by the Survey, Length composition, and Recruitment components, with minor contributions from parameter priors.
+Across all scenarios, the total log-likelihood is primarily influenced by the Survey, Length composition, and Recruitment components, with minor contributions from parameter priors.
 The reference model (s1.1) displays the highest overall likelihood value, particularly due to a stronger fit to the Length composition and Survey data.
-However, this comes with an elevated penalty in the parm priors categories, suggesting potential overfitting or parameter instability.
+However, this comes with an elevated penalty in the parameters priors categories, suggesting potential overfitting or parameter instability.
 In contrast, the ecosystem-informed scenarios—s1.2 (with predator data), s1.3 (with environmental data), and s1.4 (with both)—show slightly lower total likelihoods.
@@
 Table \@ref(tab:tablike) shows the log-likelihood component values for each scenario.
+It summarizes the contributions of different data sources—such as survey indices, length composition, recruitment, and parameter priors—to the total likelihood in each model configuration (s1.1 to s1.4).
 
@@
 
-Another step to measure the model performance in krill stock assessment was evaluated using the Root Mean Square Error (RMSE) across four scenarios of increasing ecological complexity (Figure \@ref(fig:rmseplot)).
-Scenario Model1.1, which excludes both environmental and predator variables, showed the lowest median RMSE and the narrowest interquartile range, suggesting higher consistency in model fits to observed data.
-The inclusion of ecosystem variables in Model1.2 (predator only), Model1.3 (environment only), and Model1.4 (both) resulted in increases in RMSE dispersion and median values, although the differences were relatively moderate.
-Among the ecosystem-informed scenarios, Model1.3 and Model1.4 had comparable RMSE distributions, slightly exceeding those of the predator-only model (Model1.2).
+The Root Mean Square Error (RMSE) analysis across four scenarios of increasing ecological complexity revealed that Scenario s1.1 -which excludes both environmental and predator variables- showed the lowest median RMSE and the narrowest interquartile range, suggesting higher consistency in model fits to observed data.
+The inclusion of ecosystem variables in s1.2 (predator only), s1.3 (environment only), and s1.4 (both) resulted in increases in RMSE dispersion and median values, although the differences were relatively moderate.
+Among the ecosystem-informed scenarios, s1.3 and s1.4 had comparable RMSE distributions, slightly exceeding those of the predator-only model (s1.2) (Figure \@ref(fig:rmseplot)).
 The increased variability in RMSE under these scenarios suggests a reduction in overall fit quality when ecological complexity is introduced, likely reflecting increased model structural uncertainty or misalignment between ecological covariates and population dynamics.
@@
 ```{r rmseplot, eval=TRUE, fig.cap="Root Mean Square Error (RMSE) distributions for four krill population model scenarios"}
-knitr::include_graphics('Figs/diags2.png')
+knitr::include_graphics('Figs/unnamed-chunk-29-1.jpeg')
 ```
@@
 
-Model outputs under different ecosystem scenarios revealed substantial variation in krill population indicators, including total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio) (Figure \@ref(fig:popvar)).
+Model outputs under different ecosystem scenarios revealed substantial variation in krill population indicators, including total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio) (Figure \@ref(fig:compara)).
 The reference model (s1.1), which excluded environmental and predator components, consistently produced higher median estimates and wider uncertainty intervals across all population indicators.
@@
 Scenario s1.4, which incorporated both environmental and predator components, resulted in intermediate biomass and recruitment estimates, with moderate variance compared to the other scenarios.
-Recruitment (Recruit_0) was especially sensitive to the inclusion of ecosystem components.
+Recruitment was especially sensitive to the inclusion of ecosystem components.
 The highest median recruitment was observed under s1.2 (predator-only), while the lowest was seen in s1.3 (environment-only), highlighting the contrasting influences of these ecological factors.
 
-We compare the time series of different population variables for Antarctic krill under most significant models in this analysis; a reference model without environmental or predator data and an alternative model that incorporates both environmental and predator information.
+```{r compara,eval=TRUE, fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
+Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)"}
+knitr::include_graphics('Figs/unnamed-chunk-66-1.jpeg')
+```
+
+Also, we compare the time series of different population variables (Figure \@ref(fig:compara1)).
 Total biomass estimates remain relatively stable across both scenarios until around 2010.
 From that point onward, both models show a marked increase in biomass, although the reference model predicts higher values than the ecosystem-informed model, particularly during the final assessment years (2018–2020).
-Recruitment at age 0 exhibits strong variability over time.
-This is coincidence with [REF].
+Recruitment at age 0 exhibits strong variability over time in coincidence with [@Dornam and REF].
 While both models show similar patterns, the reference model tends to estimate higher recruitment peaks, especially in recent years.
@@
 Both models show a gradual increase after 2010, but the reference model again yields higher values in recent years.
-
-```{r compara,eval=TRUE, out.width="60%", fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
-Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)"}
-knitr::include_graphics('Figs/alloutput_s1_4.png')
+A *t-test*, assuming normality and homoscedasticity, revealed a statistically significant difference in spawning biomass estimates between models `s1.1` and `s1.4` (*t = 4.2682, df = 72, p < 0.001*), with a 95% confidence interval for the mean difference ranging from *1,387,011 to 3,818,074*.
+Mean spawning biomass estimates were *4,600,304 t* for `s1.1` and *1,997,762 t* for `s1.4`.
+A non-parametric *Wilcoxon rank sum test* further confirmed this difference (*W = 1256, p < 0.001*), reinforcing the robustness of the result despite minor warnings due to ties in the dataset.
+These findings demonstrate that incorporating predator and environmental covariates into the population model (scenario `s1.4`) leads to significantly lower estimates of spawning biomass.
+
+```{r compara1,eval=TRUE, fig.cap="Time series of population variables of krill in WAP: (A) total biomass (Bio_all), (B) summary biomass (Bio_smry), (C) recruitment (Recruit_0), and (D) spawning biomass (SpawnBio).
+Axes labels should clarify units (e.g., biomass in metric tons, recruitment in millions).
+Projections assume current management/environmental conditions."}
+knitr::include_graphics('Figs/alloutput_variable2.png')
 ```
@@
 The narrow and well-separated peaks across models suggest that each model configuration yields distinctly different estimates of recruitment potential, likely reflecting differences in underlying assumptions such as natural mortality, selectivity, or the inclusion of predation effects.
-From a biological perspective, the higher ln(R0) values in `s1.3` and `s1.4` suggest that, under those model assumptions, the krill population could sustain higher levels of recruitment in unfished conditions, potentially reflecting stronger compensatory dynamics or reduced natural mortality (e.g., lower predation pressure).
-Conversely, the lower ln(R0) values in `s1.1` and `s1.2` imply a population with lower reproductive output, which might make it more vulnerable to overfishing or environmental variability.
+From a biological perspective, the higher ln(R0) values in s1.3 and s1.4 suggest that, under those model assumptions, the krill population could sustain higher levels of recruitment in unfished conditions, potentially reflecting stronger compensatory dynamics or reduced natural mortality (e.g., lower predation pressure).
 This highlights the sensitivity of the recruitment parameter to assumptions about ecological drivers in the stock-recruitment relationship.
@@
 
-The stock–recruitment relationship was examined under two model scenarios: s1.1, a reference model excluding ecosystem components, and s1.4, which incorporates both environmental and predator data  (Figure \@ref(fig:proplot)).
-The plot illustrates the relationship between spawning stock biomass (SSB) and recruitment across years, overlaid with a fitted Beverton-Holt function derived from the model with ecosystem covariates.
+We compare the stock–recruitment relationship and productivity under all model scenarios with focus in differences between scenarios with and without ecosystems consideratios (s1.1 and s1.4).
+The left panel in Figure \@ref(fig:proplot) illustrates the relationship between spawning stock biomass (SSB) and recruitment across years, overlaid with a fitted Beverton-Holt function derived from the model with ecosystem covariates.
 Scenario s1.1 (diamonds) displays a wider spread in recruitment estimates across the full range of SSB values, with several years (e.g., 2017, 2014) producing exceptionally high recruitment despite moderate to low SSB.
@@
 Notably, the inclusion of ecosystem variables appears to reduce extreme recruitment events and leads to a more stable, density-dependent recruitment pattern aligned with theoretical expectations.
-Estimating the productivity of Antarctic krill is critical for understanding the species capacity to replenish its population in response to varying levels of spawning biomass.
-Here, we defined the Productivity as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
-For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):
-
-$$
-\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
-$$
-This plot illustrates the relationship between recruitment per unit of spawning stock biomass (Recruitment/SSB) and relative spawning stock biomass (SSB/SSBo) for four scenarios (s1.1 to s1.4)  (Figure \@ref(fig:proplot)).
+We defined the stock productivity as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
+The The right panel in Figure \@ref(fig:proplot) illustrates the relationship between recruitment per unit of spawning stock biomass (Recruitment/SSB) and relative spawning stock biomass (SSB/SSBo) for four scenarios (s1.1 to s1.4).
 As expected under compensatory recruitment dynamics, all scenarios show a declining trend in productivity as SSB increases, indicating higher per-capita recruitment at low stock levels and reduced productivity as the population approaches its unfished biomass.
 The scenarios differ markedly in the magnitude of predicted productivity.
-Scenario s1.1 and s1.3 show higher productivity levels throughout the SSB range when compared to s1.2 and, most notably, s1.4.
+Scenario s1.1 and s1.3 show higher productivity levels throughout the SSB range when compared to s1.2 and s1.4.
 The latter scenario presents the lowest productivity values, indicating either reduced recruitment compensation or differences in the stock-recruitment parameters.
 These contrasting trends highlight how scenario-specific assumptions or structural choices influence estimates of productivity.
-These patterns reflect a more precautionary and ecologically realistic recruitment dynamic, where environmental variability limits the ability of the stock to maintain high productivity under depleted conditions.
-Overall, the results emphasize the influence of ecosystem variables on shaping recruitment dynamics and demonstrate that incorporating environmental drivers into the assessment model leads to more conservative productivity estimates, particularly at low spawning biomass levels.
-
-```{r proplot, eval=TRUE, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios.
-Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
-Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
+
+```{r proplot, eval=TRUE, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios (s1.1 to s1.4).
+Left panel: Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
+Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo)"}
 knitr::include_graphics('Figs/produc_com.png')
@@
 ```{r acfplot, eval=TRUE, fig.cap="Autocorrelation function (ACF) plots of model residuals for four krill assessment scenarios Blue dashed lines indicate 95% confidence intervals"}
-knitr::include_graphics('Figs/ACF2.png')
-```
-
-A *t-test*, assuming normality and homoscedasticity, revealed a statistically significant difference in spawning biomass estimates between models `s1.1` and `s1.4` (*t = 4.2682, df = 72, p < 0.001*), with a 95% confidence interval for the mean difference ranging from *1,387,011 to 3,818,074*.
-Mean spawning biomass estimates were *4,600,304 t* for `s1.1` and *1,997,762 t* for `s1.4`.
-A non-parametric *Wilcoxon rank sum test* further confirmed this difference (*W = 1256, p < 0.001*), reinforcing the robustness of the result despite minor warnings due to ties in the dataset.
-These findings demonstrate that incorporating predator and environmental covariates into the population model (scenario `s1.4`) leads to significantly lower estimates of spawning biomass.
-This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill.
+knitr::include_graphics('Figs/ACF_Plot.png')
+```
 
@@
 
-# Discusion
-
-* **Big summary / importance statement for the science**
-
-  * Expón de manera destacada la importancia científica de tus hallazgos generales.
-
-* **Scientific findings as extension of previous work**
-
-  * Relaciona tus hallazgos con trabajos previos y explica cómo los extienden o complementan.
+# Discussion
+
+(small guide)
+
+* **Big summary / importance of this kind of integrated model in Krill**
+
+* **Scientific main findings as extension of previous work (comparisons)**
 
@@
 
-  * Detalla las implicancias científicas específicas de los resultados obtenidos.
-
-* **How this engineering will facilitate future research**
-
-  * Explica cómo tu enfoque metodológico o técnico puede ayudar a futuras investigaciones.
-
 * **Forward-looking statements about the field as a whole**
 
-  * Incluye reflexiones sobre cómo evolucionará el campo en general a partir de tu trabajo.
-
 * **A summary / importance of the engineering aspect of the work**
 
-  * Resume la importancia de la parte técnica o de ingeniería de tu estudio.
-
-* **How it facilitates scientific research**
-
-  * Describe cómo tus métodos o resultados facilitan la investigación científica práctica.
-
-* **Limitations of the platform shown with questions that would be unanswered**
-
-  * Reconoce limitaciones del estudio y plantea preguntas que aún quedan por responder.
-
-By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions.
-This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
-Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
-
-Data consistency
+* **Limitations of Integrated models shown with questions that would be unanswered**
+
+* **future work or recommendations to improve this**
+
+One critical issue concerns the dependence on direct biomass estimates, which are typically derived from expensive and logistically demanding scientific surveys—most notably acoustic campaigns.
+These surveys, while highly informative, are methodologically complex and costly to repeat with sufficient temporal resolution.
+Furthermore, differences in survey design, calibration, and analytical protocols between campaigns introduce variability and uncertainty into biomass estimates, complicating their interpretation and use in long-term management decisions.
+On the other hand, CCAMLR currently employs a management framework that uses an indirect model called General Yield Model (GRYM) [REF].
+This model approach assumes stable environmental conditions—a premise that increasingly fails to hold true in the context of ongoing climate change [REF].
+GRYM, while useful for certain data-poor applications, faces difficulties due to its heavy dependence on life history parameters, such as growth rates, harvest strategies, and natural mortality.
+Moreover, the assumption of stable environmental conditions in such models is problematic [REF], as climate variability significantly influences the productivity and survival.
+As a result, predictions based on GRYM and similar models become less reliable, particularly in the face of changing environmental conditions or ecological considerations.
+This highlights the need made a transition to a **more realistic** for a integrated model and ecosystem-informed approaches to stock assessment that consider the evolving nature of environmental and ecological drivers, such as changes in sea ice extent, primary productivity, and predator-prey interactions.
 
@@
 
-The use of integrated models for assessing krill populations in the Southern Ocean has posed significant challenges for CCAMLR, particularly in light of the inherent uncertainties and limitations associated with data-poor regions.
-One of the key challenges is the reliance on models like the General Yield Model (GRYM) [REF], which assumes stable environmental conditions—a premise that increasingly fails to hold true in the context of ongoing climate change [REF].
-GRYM, while useful for certain data-poor applications, faces difficulties due to its heavy dependence on life history parameters, such as growth rates, harvest strategies, and natural mortality.
-Moreover, the assumption of stable environmental conditions in such models is problematic [REF], as climate variability significantly influences krill productivity and survival.
-As a result, predictions based on GRYM and similar models become less reliable, particularly in the face of changing environmental conditions or ecological considerations.
-This highlights the need for more dynamic, ecosystem-informed approaches to stock assessment that consider the evolving nature of environmental and ecological drivers, such as changes in sea ice extent, primary productivity, and predator-prey interactions.
-
 The transition from traditional, model-based approaches like GRYM to more integrated, ecosystem-based models presents a significant opportunity to improve the scientific basis of krill stock assessments.
@@
 While integrated models offer a more holistic view of krill population dynamics by incorporating ecosystem variables, they also introduce new challenges related to the accuracy and reliability of model inputs.
-Thus, the development of integrated models requires continuous refinement to ensure that they remain relevant and robust in the face of environmental change.
-[REF]
-
+Thus, the development of integrated models requires continuous refinement to ensure that they remain relevant and robust in the face of change in various dimensions [REF].
+
+By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions in a realistic way.
+This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
 Management strategies should incorporate ecosystem-based thresholds informed by environmental and predation dynamics, as these factors are critical in maintaining krill availability for predators, which is essential for overall ecosystem balance.
-Additionally, the ongoing warming in the Antarctic Peninsula highlights the urgent need to integrate climate-driven changes into habitat suitability models, ensuring that management approaches remain adaptive to shifting environmental conditions.
-
-Although integrated models are not new to krill in the Southern Ocean, CCAMLR has for years applied methodologies that incorporate diverse data sources.
+Additionally, the ongoing warming in the WAP highlights the urgent need to integrate climate-driven changes into habitat suitability models, ensuring that management approaches remain adaptive to shifting environmental conditions.
+Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
+
+Although integrated models are not new to krill in the Southern Ocean, CCAMLR has for years applied methodologies that incorporate diverse data sources for several years.
 External reviews, such as @Thomson2016, have endorsed these approaches as improvements over traditional models.
@@
 
+Tabla summary
+
+The analysis suggests that incorporating environmental variables (s1.3 and s1.4) improves model fit and predictive accuracy, particularly when compared to predator-only models.
+The model that includes only environmental covariates (s1.3) strikes the best balance between fit quality, complexity, and retrospective performance, suggesting that environmental variability is a key driver of krill population dynamics in the Western Antarctic Peninsula.
+
+In contrast, the inclusion of predators alone (s1.2) leads to poorer model fit and higher retrospective bias, possibly due to data limitations or unaccounted complexity in predator-prey dynamics.
+While the full ecosystem model (s1.4) shows some improvements over the predator-only model, it still exhibits significant retrospective biases, highlighting the need for refined integration of ecosystem components and further data support, particularly regarding predator consumption rates and spatial foraging behavior.
+
+Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the joint inclusion of predator and environmental variables yielding the most robust representation of observed length compositions.
+The inclusion of environmental covariates (s1.3 and s1.4) markedly improves the temporal independence of model residuals, as indicated by lower autocorrelation beyond lag 1.
+This implies that environmental variability contributes significantly to explaining observed trends in krill population dynamics.
+Conversely, the predator-only model (s1.2) introduces more persistent autocorrelation, indicating that its effects may require more nuanced representation or additional process structuring.
+
+The analysis indicates that incorporating environmental and predator data into the stock assessment framework introduces greater variability in model fit, as measured by RMSE.
+While the baseline model (Model 1.1) demonstrates more stable performance, it potentially achieves this at the cost of biological realism.
+The modest trade-off in predictive accuracy associated with ecosystem-informed models is expected and highlights the need to balance goodness-of-fit with ecological plausibility when developing integrated assessment models.
+
+The inclusion of predator and environmental variables significantly alters model behavior in retrospective diagnostics, particularly increasing bias in terminal-year estimates of fishing mortality and deepening the underestimation of spawning stock biomass (SSB).
+These patterns suggest that while ecosystem-informed models improve the biological realism of the stock assessment, they also pose challenges in terms of stability and predictive consistency.
+Such outcomes underline the importance of further refining the structure and parameterization of ecosystem components when integrating them into assessment frameworks, to balance realism with diagnostic robustness.
+
+Likelihood composition reveals that models incorporating ecological data (s1.2–s1.4) yield slightly lower total likelihoods than the reference model (s1.1), reflecting trade-offs between model fit and biological realism.
+The reduced penalty terms and more evenly distributed likelihood contributions in the ecosystem scenarios suggest improved model stability and reduced overparameterization.
+These results reinforce the value of integrating ecosystem variables not only for enhancing ecological interpretation but also for promoting robust model structure and parameter behavior.
+
+> (pocos estudios incluyen ambas variables. problemas en interpretacion y cofounding correlation between parameters)
+
+### Ecosystem Variables: krill population dynamics relevance
+
+> regarding clorofila como driver; This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
+
+>  About index; These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
+
+Variations in krill size composition in penguin diets have been shown to reflect underlying changes in krill population dynamics, including recruitment pulses and cohort strength [Jeff REFs].
+Thus, penguin diet data serve as a proxy for detecting cyclical patterns in krill recruitment that may not be evident through direct surveys alone.
+[REF]
+
+The ecosystem variables used in assessment processes vary from their definition to how they are applied.
+Environmental variables influencing krill recruitment, as well as predators that introduce an additional component of natural mortality into the model, are among those considered in this study.
+[REF]
+
+Marshall et al.
+(2019) [REF] classify ecosystem variables based on their impact on population dynamics and how they are incorporated into stock assessment models.
+They suggest that integrating variables, especially recruitment, directly within the model results in a more dynamic and accurate representation of population dynamics, as it captures real-time interactions between ecological factors.
+In contrast, external incorporation may not fully capture ecosystem feedback loops, potentially leading to less reliable assessments.
+
+Our analysis, which integrates ecosystem variables directly within the model, aligns with the conclusions of Marshall et al.
+(2019) and provides further evidence that this approach offers superior accuracy in modeling krill population dynamics.
+By including both environmental drivers and predation pressures within the model, our study captures the complexity of krill dynamics more effectively, offering a more realistic representation of krill biomass, recruitment, and mortality processes.
+This integrated approach not only improves model precision but also enhances its predictive power, addressing key ecological factors that influence krill populations.
+
+This analysis represents a significant step forward in the incorporation of ecosystem-based considerations into krill stock assessments in the Southern Ocean.
+By adopting a more comprehensive, ecosystem-informed model, we contribute to the growing body of research supporting the integration of ecosystem variables in stock assessment frameworks [REF].
+This approach is consistent with the new and evolving management strategies of the Commission for the CCAMLR, which increasingly emphasizes ecosystem-based management practices.
+As climate change and ecosystem variability continue to impact krill populations, this analysis provides a foundation for precautionary management strategies that are better aligned with the dynamic nature of marine ecosystems.
+
+Like @Crone2019, the study concluded that both methods for including environmental data resulted in relatively high-quality estimates for a diverse group of output variables.
+The best performing models were those that included a correct and bias-corrected environmental variable, indicating the importance of accurately incorporating environmental factors in stock assessments
+
+Regarding to evironmental impact in modelliling of main population variables in stock assessment process, @Haltuch2009 recommends that low-frequency environmental variability significantly affects fish population abundance through recruitment, and biomass reference points should be based on average recruitment and/or dynamic B0, especially if comprehensive catch and survey data are available, y a su vez, la calidad y cantidad de datos mejorar la exactitud de resultados.
+
+Overall, these results highlight the benefit of including environmental signals to reduce residual autocorrelation and enhance model fit. This analysis coincides with... [REF]
+
 ### Predation and Natural Mortality
-
-> The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lukcas!, REF Vicky].
-
-> variability length; , which may reflect fluctuations in recruitment strength and availability of krill age classes.
-
-> . This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics
 
@@
 
-Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the joint inclusion of predator and environmental variables yielding the most robust representation of observed length compositions.
-The inclusion of environmental covariates (s1.3 and s1.4) markedly improves the temporal independence of model residuals, as indicated by lower autocorrelation beyond lag 1.
-This implies that environmental variability contributes significantly to explaining observed trends in krill population dynamics.
-Conversely, the predator-only model (s1.2) introduces more persistent autocorrelation, indicating that its effects may require more nuanced representation or additional process structuring.
-Overall, these results highlight the benefit of including environmental signals to reduce residual autocorrelation and enhance model fit.
-This analysis coincides with...
-[REF]
-
-The analysis indicates that incorporating environmental and predator data into the stock assessment framework introduces greater variability in model fit, as measured by RMSE.
-While the baseline model (Model 1.1) demonstrates more stable performance, it potentially achieves this at the cost of biological realism.
-The modest trade-off in predictive accuracy associated with ecosystem-informed models is expected and highlights the need to balance goodness-of-fit with ecological plausibility when developing integrated assessment models.
-
-The inclusion of predator and environmental variables significantly alters model behavior in retrospective diagnostics, particularly increasing bias in terminal-year estimates of fishing mortality and deepening the underestimation of spawning stock biomass (SSB).
-These patterns suggest that while ecosystem-informed models improve the biological realism of the stock assessment, they also pose challenges in terms of stability and predictive consistency.
-Such outcomes underline the importance of further refining the structure and parameterization of ecosystem components when integrating them into assessment frameworks, to balance realism with diagnostic robustness.
-
-Likelihood decomposition reveals that models incorporating ecological data (s1.2–s1.4) yield slightly lower total likelihoods than the reference model (s1.1), reflecting trade-offs between model fit and biological realism.
-The reduced penalty terms and more evenly distributed likelihood contributions in the ecosystem scenarios suggest improved model stability and reduced overparameterization.
-These results reinforce the value of integrating ecosystem variables not only for enhancing ecological interpretation but also for promoting robust model structure and parameter behavior.
+> The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lucas, REF Vicky].
+
+> This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics
+
+> variability length; , which may reflect fluctuations in recruitment strength and availability of krill age classes.
 
@@
 
-### Stock-Recruitment and Biomass Estimates
+(Talk about decrease in penguins colonies at 2020 an impact in modeling)
+
+### Productivity
 
@@
 
-### Productivity
+> This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill.
 
@@
 
-### Ecosystem Variables: Ecological Relevance
-
-> regarding clorofila como driver; This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
-
->  About index; These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
-
-Variations in krill size composition in penguin diets have been shown to reflect underlying changes in krill population dynamics, including recruitment pulses and cohort strength [Jeff REFs].
-Thus, penguin diet data serve as a proxy for detecting cyclical patterns in krill recruitment that may not be evident through direct surveys alone.
-[REF]
-
-The ecosystem variables used in assessment processes vary from their definition to how they are applied.
-Environmental variables influencing krill recruitment, as well as predators that introduce an additional component of natural mortality into the model, are among those considered in this study.
-[REF]
-
-Marshall et al.
-(2019) [REF] classify ecosystem variables based on their impact on population dynamics and how they are incorporated into stock assessment models.
-They suggest that integrating variables, especially recruitment, directly within the model results in a more dynamic and accurate representation of population dynamics, as it captures real-time interactions between ecological factors.
-In contrast, external incorporation may not fully capture ecosystem feedback loops, potentially leading to less reliable assessments.
-
-Our analysis, which integrates ecosystem variables directly within the model, aligns with the conclusions of Marshall et al.
-(2019) and provides further evidence that this approach offers superior accuracy in modeling krill population dynamics.
-By including both environmental drivers and predation pressures within the model, our study captures the complexity of krill dynamics more effectively, offering a more realistic representation of krill biomass, recruitment, and mortality processes.
-This integrated approach not only improves model precision but also enhances its predictive power, addressing key ecological factors that influence krill populations.
-
-This analysis represents a significant step forward in the incorporation of ecosystem-based considerations into krill stock assessments in the Southern Ocean.
-By adopting a more comprehensive, ecosystem-informed model, we contribute to the growing body of research supporting the integration of ecosystem variables in stock assessment frameworks [REF].
-This approach is consistent with the new and evolving management strategies of the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR), which increasingly emphasizes ecosystem-based management practices.
-As climate change and ecosystem variability continue to impact krill populations, this analysis provides a foundation for precautionary management strategies that are better aligned with the dynamic nature of marine ecosystems.
-
-Like @Crone2019, the study concluded that both methods for including environmental data resulted in relatively high-quality estimates for a diverse group of output variables.
-The best performing models were those that included a correct and bias-corrected environmental variable, indicating the importance of accurately incorporating environmental factors in stock assessments
-
-Regarding to evironmental impact in modelliling of main population variables in stock assessment process, @Haltuch2009 recommends that low-frequency environmental variability significantly affects fish population abundance through recruitment, and biomass reference points should be based on average recruitment and/or dynamic B0, especially if comprehensive catch and survey data are available, y a su vez, la calidad y cantidad de datos mejorar la exactitud de resultados.
-
-Tabla summary
-
-The analysis suggests that incorporating environmental variables (s1.3 and s1.4) improves model fit and predictive accuracy, particularly when compared to predator-only models.
-The model that includes only environmental covariates (s1.3) strikes the best balance between fit quality, complexity, and retrospective performance, suggesting that environmental variability is a key driver of krill population dynamics in the Western Antarctic Peninsula.
-
-In contrast, the inclusion of predators alone (s1.2) leads to poorer model fit and higher retrospective bias, possibly due to data limitations or unaccounted complexity in predator-prey dynamics.
-While the full ecosystem model (s1.4) shows some improvements over the predator-only model, it still exhibits significant retrospective biases, highlighting the need for refined integration of ecosystem components and further data support, particularly regarding predator consumption rates and spatial foraging behavior.
-
-These findings support growing consensus that ecosystem-informed stock assessments, particularly those that integrate environmental covariates, provide more robust and ecologically realistic estimates for Antarctic krill and should inform precautionary management frameworks under CCAMLR.
-
-### Limitations of the Paper
+CC coments : **"perhaps it is worth haveing another read, and combine some bits of discussion, there’s repetition of highlighting the important of the model and env variables .
+This is true but there’s  no need of saying it everytime you discuss a parameter"**
+
+> SRR: These patterns reflect a more precautionary and ecologically realistic recruitment dynamic, where environmental variability limits the ability of the stock to maintain high productivity under depleted conditions.
+Overall, the results emphasize the influence of ecosystem variables on shaping recruitment dynamics and demonstrate that incorporating environmental drivers into the assessment model leads to more conservative productivity estimates, particularly at low spawning biomass levels.
+
+### Model integrated limitations
 
@@
 However, these models also introduce challenges such as increased sensitivity and retrospective bias, particularly in fishing mortality estimates, underscoring the need for a balance between ecological realism and model stability to ensure reliable management decisions.
-Overall, these findings reinforce that ecosystem-based approaches provide a stronger scientific foundation for fishery management, but they must be carefully managed to avoid potential instability in model outputs.
+Overall, these findings reinforce that integrated model in an ecosystem-based approach provide a stronger scientific foundation for krill fishery management.
 
@@
 
--   Supplementary Material 1: analisis, formulas, and outputs of stock assessment of Antarctic krill in WAP can be found in this [link](https://mauromardones.github.io/SA_Krill/).
+-   Supplementary Material 1: analysis, formulas, and outputs of stock assessment of Antarctic krill in WAP can be found in this [link](https://mauromardones.github.io/SA_Krill/).
 

### 2025-05-16 index.Rmd
close work until..

@@
                        midpoint = 0) +
-  facet_grid(Fleet ~ Model) +
+  facet_grid(Model ~ Fleet) +
   theme_few() +
@@
 heatmapres
+
+circleplot_res <- ggplot(df_all, aes(x = Yr, y = Bin)) +
+  geom_point(aes(fill = Pearson, size = abs(Pearson)),
+             shape = 21, color = "grey30", stroke = 0.2, alpha = 0.25) +
+  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0) +
+  scale_alpha("none") +
+  scale_size_continuous(range = c(0.5, 2.5)) +
+  facet_grid(Model ~ Fleet) +
+  theme_few() +
+  theme(
+    axis.text.x = element_text(angle = 90, hjust = 1, size = 5),         
+    axis.text.y = element_text(size = 5),                               
+    strip.text = element_text(size = 8),
+    legend.position = "bottom"
+  ) +
+  labs(x = "", y = "Length (cm)", fill = "Pearson") +
+  guides(size = "none", alpha = "none")  # Esto oculta las leyendas de size y alpha
+
+circleplot_res
 ```
@@
   theme(legend.position = "bottom")
+
 ```
@@
 # Crear una columna de modelo en cada data frame
-outps1$Model <- "s1.1 (Ref Model: No Env-Predator)"
-outps2$Model <- "s1.2 w/ Predator data"
-outps3$Model <- "s1.3 w/ Env data"
-outps4$Model <- "s1.4 (Env and Predator data)"
+outps1$Model <- "s1.1"
+outps2$Model <- "s1.2"
+outps3$Model <- "s1.3"
+outps4$Model <- "s1.4"
 
 # Unir los data frames en uno solo
-outpsall <- rbind(outps1,  outps4)
+outpsall <- rbind(outps1,  
+                  outps2,
+                  outps3,
+                  outps4)
 
 outpsall_long <- outpsall %>%
-  pivot_longer(cols = c(Bio_smry, SpawnBio, Recruit_0), 
+  pivot_longer(cols = c(Bio_all, Bio_smry, SpawnBio, Recruit_0), 
                names_to = "Variable", values_to = "Value") %>% 
@@
 alloutput <- ggplot(outpsall_long, aes(x = Yr, y = Value, color = Model, group = Model)) +
-  geom_line(linewidth = 1) +
-  geom_point(size = 2) +
+  geom_point() +
+  #geom_smooth(method = "lm")+
   facet_wrap(~Variable, 
-             ncol = 1, 
+             ncol = 4, 
              scales = "free_y") + 
-  scale_color_manual(name = "",
-                     values = c("s1.1 (Ref Model: No Env-Predator)" = "#ca0020",
-                                "s1.4 (Env and Predator data)" = "#404040")) +
+  scale_color_manual(name = "Scenario",
+                     values = c("s1.1" = "#ca0020", 
+                                "s1.2" = "#f4a582", 
+                                "s1.3" = "#bababa", 
+                                "s1.4" = "#404040")) +
   labs(title = "", x = "", y = "") +
-  theme_bw() +
+  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1),
-        legend.position = "none")
+        legend.position = "bottom")
 
@@
 r0plot <- ggplot(dens_data, aes(x = x, y = y, color = model, fill = model)) +
-  geom_area(alpha = 0.5, position = "identity") +
-  geom_line(size = 1) +
-  scale_color_manual(values = colores) +
-  scale_fill_manual(values = colores) +
-  labs(title = "",
-       x = "Natural log of virgin recruitment level (R0)", y = "Density") +
-  theme_bw() +
-  xlim(16, 22)
+  geom_area(alpha = 0.3, position = "identity", size = 0) +  # More transparent areas without borders
+  geom_line(size = 1.2) +  # Thicker lines for better visibility
+  scale_color_manual(values = colores, name = NULL) +  # Remove legend title for color
+  scale_fill_manual(values = colores, name = NULL) +   # Remove legend title for fill
+  labs(x = "ln(R0)", 
+       y = "Density") +
+  theme_bw(base_size = 14) +
+  theme(
+    legend.position = "bottom",  # Legend at bottom
+    legend.spacing.x = unit(0.5, 'cm'),  # Space between legend items
+    panel.grid.major = element_line(color = "gray90"),
+    panel.grid.minor = element_blank(),
+    axis.title = element_text(size = 12),
+    legend.text = element_text(size = 11)
+  ) +
+  xlim(16, 22) +
+  guides(color = guide_legend(nrow = 1),  # Single row legend
+         fill = guide_legend(nrow = 1))
+
 r0plot
@@
 
-ggplot(err_long, aes(x = Model, y = Value, fill = Model)) +
+comsum <- ggplot(err_long, aes(x = Model, y = Value, fill = Model)) +
   geom_boxplot(width = 0.5, outlier.shape = NA,
@@
         legend.position = "none")
+ggarrange(
+          alloutput,
+          comsum,
+          ncol = 1)
 ```
@@
 p1 <- ggAcf(base.model1.1$recruit$pred_recr) +
-  ggtitle("Ref Model: No Env-Predator") +
+  ggtitle("s1.1") +
   theme_few()
 p2 <- ggAcf(base.model1.2$recruit$pred_recr) +
-  ggtitle("S1.1 w/ Predator data") +
+  ggtitle("s1.2") +
   theme_few()
 p3 <- ggAcf(base.model1.3$recruit$pred_recr) +
-  ggtitle("S1.1 w/ Env data") +
+  ggtitle("s1.3") +
   theme_few()
 p4 <- ggAcf(base.model1.4$recruit$pred_recr) +
-  ggtitle("S1.1 w/ Env and Predator data") +
+  ggtitle("s1.4") +
   theme_few()
 
-ggarrange(p1,
-          p2,
-          p3,
-          p4, ncol=1, nrow = 4,
-          common.legend = T)
+ggarrange(p1, p2,p3,p4, ncol=4, nrow = 1, common.legend = T)
 

### 2025-05-21 MS2_Krill.Rmd
last changes in html

@@
 
+@Boldt2025
+
 ### Ecosystem Variables: krill population dynamics relevance
@@
 ### Model integrated limitations
+
+(Discussion)
+
+The observed retrospective bias in SSB and F aligns with findings from previous studies on krill population dynamics, where recruitment variability and model structure influence stock assessment estimates (Hill et al., 2016; Watters et al., 2020).
+The negative bias in SSB suggests that models may overestimate past biomass, potentially due to unaccounted environmental variability or underestimation of natural mortality (Meyer et al., 2021).
+The positive bias in F indicates a tendency to underestimate past fishing mortality, which could be linked to the inclusion of predator-based indices such as penguin foraging data (Trathan et al., 2018).
+These results highlight the importance of incorporating ecosystem interactions into stock assessments to improve predictive accuracy and support sustainable management of Antarctic krill.
 

### 2025-05-21 index.Rmd
last changes in html

@@
 subtitle: "Assessing ecosystem effects on the stock productivity of Antarctic Krill (*Euphausia superba*): An Integrated modeling perspective"
-author: "Mardones, M; Jarvis Mason, E.T.; Pinones, A.;  Santa Cruz, F.; Cárdenas, C.A"
+author: "Mardones, M; Jarvis Mason, E.T.; Piñones, A.; Krüger, L.,Santa-Cruz, F.; Watters, G., Cárdenas, C.A; Methot, R."
 date:  "`r format(Sys.time(), '%d %B, %Y')`"
 bibliography: SA_krill.bib
-#csl: apa.csl
-csl: icesjournal.csl
+csl: apa.csl
+#csl: icesjournal.csl
 link-citations: yes
@@
       smooth_scroll: false
-    theme: cosmo
+    theme: simplex
     fontsize: 0.9em
@@
 
+## MODELING SETTING
+
 ```{r setup1, echo=FALSE}
@@
                       warning = FALSE,
-                      fig.height = 6,
-                      fig.width = 7,
                       fig.align = 'center',
@@
 
-```{r message=FALSE, eval=TRUE}
+### Libraries
+
+Libraries necessary to made this analysis;
+
+```{r message=FALSE, eval=TRUE, echo=TRUE}
 # install.packages("devtools")
@@
 library(kableExtra)
-library(doParallel) # facilita la ejecución paralela en R
+library(doParallel) 
 detectCores()
@@
 library(flextable)
-```
+library(car)
+```
+
+### Code Repository
+
+The repository with files templates by scenario to replicate this analysis can be found in this [GitHub author link](https://github.com/MauroMardones/SA_Krill/tree/main/scenarios)
 
 ```{r message=FALSE, warning=FALSE, echo=TRUE}
-# dir01 <- here("s01") # agreggate data (no spatial diferences)
-# dir1<-here("s1") # Data strata flishery
-# dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
-# dir3<-here("s3") # without S-R
-# dir4<-here("s4") # 
-# dir5<-here("s5") # 
-# dir6<-here("s6") # 
-# dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
-# dir8<-here("s8") # s1 platoons
-# dir9<-here("s9") # s1 w/ blocks
-dir1.1<-here("s1.1") # sin predador ni ambiental
-dir1.2 <- here("s1.2") # s1.1 C/ predador
-dir1.3 <- here("s1.3") # s1.1 solo env
-dir1.4 <- here("s1.4") # s1.2 predator and env
-Figs <- here("Figs")# S
+dir1.1<-here("s1.1") # no predador- environmental
+dir1.2 <- here("s1.2") # s1.1 w/ predador
+dir1.3 <- here("s1.3") # s1.1 w/ env
+dir1.4 <- here("s1.4") # s1.1 predator and env
+Figs <- here("Figs")
+```
+
+```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
+### Run Models
+directorios <- c("s1.1",
+                 "s1.2",
+                 "s1.3",
+                 "s1.4")
+
+for (dir in directorios) {
+  r4ss::run(
+    dir = dir,
+    exe = "ss_osx",
+    skipfinished = FALSE,
+    show_in_console = TRUE
+  )
+}
+```
+
+```{r eval=FALSE, message=F, include=FALSE}
+#OR with rby separate
+#shell(cmd="ss") # run SS to windows
+# or 
+r4ss::run(
+  dir = dir1.1,
+  exe = "ss_osx",
+  skipfinished = FALSE, 
+  show_in_console = TRUE 
+)
+# s1.2
+r4ss::run(
+  dir = dir1.2,
+  exe = "ss_osx",
+  skipfinished = FALSE, 
+  show_in_console = TRUE 
+)
+# s1.3
+r4ss::run(
+  dir = dir1.3,
+  exe = "ss_osx",
+  skipfinished = FALSE, 
+  show_in_console = TRUE 
+)
+# s1.4
+r4ss::run(
+  dir = dir1.4,
+  exe = "ss_osx",
+  skipfinished = FALSE, 
+  show_in_console = TRUE 
+)
+```
+
+```{r message=F, include=FALSE}
+#s1.1
+base.model1.1 <- SS_output(dir=dir1.1,
+                        covar=T,
+                        forecast=T)
+#s1.2
+base.model1.2 <- SS_output(dir=dir1.2,
+                        covar=T,
+                        forecast=T)
+#s1.3
+base.model1.3 <- SS_output(dir=dir1.3,
+                        covar=T,
+                        forecast=T)
+#s1.4
+base.model1.4 <- SS_output(dir=dir1.4,
+                        covar=T,
+                        forecast=T)
 ```
@@
 By incorporating these factors into the assessment, we analyze how krill population variables respond to ecological variability, providing insights into their resilience and potential management implications.
-
-Here, the **reference model** represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables.
-This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure.
-By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.
-
-The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale (Figure \@ref(fig:mapa)).
-To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021; REF] (Fig 1 S1).
-While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale.
-This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics.
-This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.
+This code forms part of the supplementary material for the scientific article titled **"Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective."**
+
+## Spatial dimension of stock assessment
+
+The study focuses on Subarea 48.1 in the Western Antarctic Peninsula, where most krill fishing occurs and CCAMLR is advancing spatially refined management.
+We use five CCAMLR-defined strata to increase spatial resolution, allowing detection of regional differences in krill dynamics (Figure \@ref(fig:mapa)).
+This structure, integrated into the stock assessment model, enhances its capacity to capture spatial variability in both biological and environmental data.
+In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014].
+This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1.
 
@@
 
-In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014].
-This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).
-
-```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
+Figure \@ref(fig:conceptual) illustrates a conceptual model with the spatial distribution and movement of Antarctic krill in the Western Antarctic Peninsula (WAP).
+Adult krill are concentrated and move primarily in the northern areas, while juvenile krill dominate the southern regions.
+Arrows indicate directional flows, suggesting ontogenetic or environmentally-driven migration patterns.
+In terms of assumptions, although there are spatial differences in population structure, all these areas are considered part of a single, closed population unit upon which the stock assessment is conducted.
+By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions.
+These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.
+
+```{r conceptual, out.width='50%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
 knitr::include_graphics('Figs/conceptual.jpeg')
@@
 SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-In this exercise, SS3 is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers.
-The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends.
+The spurce code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code).
+In this exercise, `SS3` is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers.
+The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends in krill.
 The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023].
-By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions.
-These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.
-
-```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
-knitr::include_graphics('Figs/pathmod.png')
-```
-
-### Age Key Length (ALK)
-
-In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process
+Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].
+
+On the other hand, length data, which are cost-effective and readily available [@Canales2021; @Chong2019], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014].
+For made this transformation, the models often use an Age-Length Keys (ALKs).
+An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classed and this method assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024].
+The ALK matrix is typically defined as:
+
+In a catch-at-length model like krill assessment the AKL matrix (Figure \@ref(fig:AKL)) is modeled trough parametrization process and have this shape;
 
@@
 
-## Parameters
+## Data
+
+### Parameters
 
@@
       format = "html",
-      position = "ht!",
       caption = "Input parameters for the initial SS3 model of krill.
@@
 If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-  kable_paper("hover", full_width = TRUE) %>%  # Expande el ancho
-  kable_styling(latex_options = c("striped", "condensed"),
-                full_width = TRUE,  # Expande la tabla al ancho completo
-                font_size = 12,
-                html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
+      kable_styling(latex_options = "scale_down",  # Expande la tabla al ancho completo
+                font_size = 8) %>%  # Aplica la fuente Times New Roman
   pack_rows(index = c("Natural Mortality" = 1,
@@
 ```
-Source of data input.
-
-Ibdex abundance in Figure \@ref(fig:index)
+
+### Index
+
+Abundance index in Figure \@ref(fig:index)
 
@@
     index_std = (obs - min(obs)) / (max(obs) - min(obs)),
-    index = recode(index,
+    index = dplyr::recode(index,
                    `1` = "FISHERYBS",
@@
 
-```{r index, fig.width=9, fig.height=4, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1.
+```{r index, fig.width=9, fig.height=6, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1.
 Each panel represents a distinct spatial or functional stratum, with trend lines indicating temporal variation from 1990 to 2020.
@@
 
+### Length comositions
+
 ```{r warning=FALSE}
@@
   mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
-  mutate(FltSvy = recode(FltSvy,
+  mutate(FltSvy = dplyr::recode(FltSvy,
                          `1` = "FISHERYBS",
@@
 
-## Scenarios
-
-In Table 1 we have ten scenarios to test different option in modeling
-about main consideration in assessment of krill population.
-
-| Scenario | Description                                                                          |
-|:------------:|:---------------------------------------------------------|
-|    s1.1  | Spatial data without environmental and predator components                           |
-|    s1.2  | "s1.1" with predator components                                                      |
-|    s1.3  | "s1.1" with environmental variable                                                  |
-|    s1.4  | "s1.1" w/ both, predator fleet and environmental variable                           |
-
-### Run Models
-
-```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
-directorios <- c("s1.1",
-                 "s1.2",
-                 "s1.3",
-                 "s1.4")
-
-for (dir in directorios) {
-  r4ss::run(
-    dir = dir,
-    exe = "ss_osx",
-    skipfinished = FALSE,
-    show_in_console = TRUE
-  )
-}
-```
-
-```{r eval=FALSE, message=F, include=FALSE}
-#OR with rby separate
-#shell(cmd="ss") # run SS to windows
-# or 
-r4ss::run(
-  dir = dir1.1,
-  exe = "ss_osx",
-  skipfinished = FALSE, 
-  show_in_console = TRUE 
-)
-
-# s1.2
-r4ss::run(
-  dir = dir1.2,
-  exe = "ss_osx",
-  skipfinished = FALSE, 
-  show_in_console = TRUE 
-)
-
-# s1.3
-r4ss::run(
-  dir = dir1.3,
-  exe = "ss_osx",
-  skipfinished = FALSE, 
-  show_in_console = TRUE 
-)
-
-# s1.4
-r4ss::run(
-  dir = dir1.4,
-  exe = "ss_osx",
-  skipfinished = FALSE, 
-  show_in_console = TRUE 
-)
-```
-
-```{r message=F, include=FALSE}
-#s1.1
-base.model1.1 <- SS_output(dir=dir1.1,
-                        covar=T,
-                        forecast=T)
-#s1.2
-base.model1.2 <- SS_output(dir=dir1.2,
-                        covar=T,
-                        forecast=T)
-#s1.3
-base.model1.3 <- SS_output(dir=dir1.3,
-                        covar=T,
-                        forecast=T)
-#s1.4
-base.model1.4 <- SS_output(dir=dir1.4,
-                        covar=T,
-                        forecast=T)
-```
-
-Data used en both (spatial and No spatial models) `s1.1` and `s1.4` to compare
-
-```{r fig.height=10, fig.width=6}
-par(mfrow=c(2,1), mar=c(5,4,2,1))
-SSplotData(base.model1.1, 
-           subplots = 1,
-           ptsize = 0.5)
-
+## Summary data sources to Modelling
+
+Finally, this information and all sources can be represented through the following flow diagram (Figure \@ref(fig:path)) of inputs, model, and outputs.
+
+```{r path, fig.cap="Framework path to stock assessment model in krill in WAP (Yellow boxes is not implemeted yet)."}
+knitr::include_graphics('Figs/pathmod.png')
+```
+
+Figure \@ref(fig:dataserie) show time series of differente componentes of data sources to this krill stock assessment.
+
+```{r dataserie, fig.width=9, fig.height=7}
 SSplotData(base.model1.4, 
@@
 
+## Scenarios
+
+Here, the **reference model** is `s1.1`, wich represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables.
+This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure.
+By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.
+
+Table \@ref(tab:scenarios) show scenarios used for modelling dynamics in krill
+
+```{r scenario, eval=TRUE}
+# Tu código de flextable aquí
+scenario_table <- data.frame(
+  Scenario = c("s1.1", "s1.2", "s1.3", "s1.4"),
+  Description = c(
+    "Spatial data without environmental and predator components",
+    "\"s1.1\" with predator components",
+    "\"s1.1\" with environmental variable",
+    "\"s1.1\" with both predator fleet and environmental variable"
+  ),
+  stringsAsFactors = FALSE
+)
+
+scenario_table %>%
+  kbl(caption = "Scenarios used for modelling dynamics in krill",
+      align = "l",  # Left alignment (or use "c" for center)
+      booktabs = TRUE,
+      col.names = c("Scenario", "Description")) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
+                full_width = FALSE,
+                position = "center",
+                font_size = 10,  # Equivalent to fontsize(size = 10)
+                latex_options = "HOLD_position") %>%
+  row_spec(0, bold = TRUE, color = "black") %>%  # Header formatting
+  column_spec(1, width = "2cm") %>%  # Adjust column widths
+  column_spec(2, width = "10cm")
+```
+
 # RESULTS
+
+## Population Variables by scenario
 
@@
 
-Main Variables poulation in `s1.1` scenario
-
-```{r fig.height=2, fig.width=7}
+Main Variables poulation in `s1.1` scenario (Figure \@ref(fig:scen1))
+
+```{r scen1, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.1"}
 par(mfrow=c(1,3), mar=c(5,4,1,1)) 
@@
 
-Main Variables poulation in `s1.2` scenario
-
-```{r fig.height=2, fig.width=7}
+Main Variables poulation in `s1.2` scenario (Figure \@ref(fig:scen2))
+
+```{r scen2, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.2"}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
@@
 
-Main Variables poulation in `s1.3` scenario
-
-```{r fig.height=2, fig.width=7}
+Main Variables poulation in `s1.3` scenario (Figure \@ref(fig:scen3))
+
+```{r scen3, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.3"}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
@@
 
-Main Variables poulation in `s1.4` scenario
-
-```{r fig.height=2, fig.width=7}
+Main Variables poulation in `s1.4` scenario (Figure \@ref(fig:scen4))
+
+```{r scen4, fig.height=2, fig.width=7 , fig.cap="Main variables scenario s1.4"}
 par(mfrow=c(1,3), mar=c(5,4,1,1))
@@
 ```
-Selectivity
-
-```{r}
+
+Selectivity estimated by scenario in Figure \@ref(fig:selectivity).
+
+```{r selectivity, fig.height=8, fig.width=8, fig.cap="Selectivity by fleet in each scenario"}
 par(mfrow=c(2,2), mar=c(5,4,1,1))
@@
 ```
-```{r}
-par(mfrow=c(2,2), mar=c(5,4,1,1))
-SSplotBiology(base.model1.1,
-            subplots = 1)
-SSplotBiology(base.model1.2,
-            subplots = 1)
-SSplotBiology(base.model1.3,
-            subplots = 1)
-SSplotBiology(base.model1.4,
-            subplots = 1)
-```
-
-Natural Mortality
-
-```{r eval=FALSE}
-par(mfrow=c(4,2), mar=c(5,4,2,1))
-SSplotBiology(base.model1.1,
-              subplots = 21)
-SSplotBiology(base.model1.2,
-            subplots = 21)
-mtext("s1.2", side=3, line=0.5, cex=1.2)
-SSplotBiology(base.model1.3,
-            subplots = 21)
-mtext("s1.3", side=3, line=0.5, cex=1.2)
-SSplotBiology(base.model1.4,
-            subplots = 21)
-mtext("s1.4", side=3, line=0.5, cex=1.2)
-```
-
-```{r}
+
+This Figure \@ref(fig:index2) shows standardized time series of input indices used in four different model scenarios (s1.1 to s1.4) for the stock assessment of Antarctic krill in Subarea 48.1.
+Each panel presents fishery-dependent (FISHERY) and fishery-independent (SURVEY) indices across five management strata: Bransfield Strait (BS), Elephant Island (EI), Gerlache Strait (GS), Joinville Island (JOIN), and South West (SW), from the mid-1990s to 2020.
+Scenario s1.4 additionally includes a predator index, highlighting the incorporation of ecosystem variables.
+The figure illustrates spatial and temporal variability in data availability and trends among sources.
+
+```{r index2, fig.height=8, fig.width=8, fig.cap= "Standardized indices of krill abundance used as input in four model scenarios (s1.1 to s1.4), representing fishery-dependent (FISHERY) and fishery-independent (SURVEY) data across five spatial strata: Bransfield Strait (BS), Elephant Island (EI), Gerlache Strait (GS), Joinville Island (JOIN), and South West (SW).
+Scenario s1.4 also incorporates a predator index (PREDATOR), reflecting the integration of ecosystem variables into the assessment framework"}
 par(mfrow=c(2,2), mar=c(5,4,2,1)) 
-
 SSplotIndices(base.model1.1, subplots = 9)
 mtext("s1.1", side=3, line=0.5, cex=1.2)
-
 SSplotIndices(base.model1.2, subplots = 9)
 mtext("s1.2", side=3, line=0.5, cex=1.2)
-
 SSplotIndices(base.model1.3, subplots = 9)
 mtext("s1.3", side=3, line=0.5, cex=1.2)
-
 SSplotIndices(base.model1.4, subplots = 9)
 mtext("s1.4", side=3, line=0.5, cex=1.2)
-
-```
+```
+
+### Outputs Variables `s1.1`
 
 ```{r}
-SSplotComps(base.model1.1,
-            subplots = 21)
-SSplotComps(base.model1.2,
-            subplots = 21)
-SSplotComps(base.model1.3,
-            subplots = 21)
-SSplotComps(base.model1.4,
-            subplots = 21)
-```
-Total biomass
+outps1 <- base.model1.1$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps1 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP `s1.1`") %>%
+ kable_styling(latex_options = c("scale_down",
+                                    "striped"),# Expande la tabla al ancho completo
+                font_size = 8)  #%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Outputs Variables `s1.2`
 
 ```{r}
-total1 <- as.data.frame(base.model1.1$timeseries[,c(2,5)]) %>%
-  mutate(Serie = "Without Ecosystem Variables")
-total2 <- as.data.frame(base.model1.2$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "W/ Predator")
-total3 <- as.data.frame(base.model1.3$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "W/ Environment")
-total4 <- as.data.frame(base.model1.4$timeseries[,c(2,5)]) %>% 
-  mutate(Serie = "With Predator and Ecosystem")
-
-totalbio <- rbind(total1,
-              total2,
-              total3,
-              total4)
-
-ggplot(totalbio, aes(x = Yr, y = Bio_all, color=Serie)) +
+outps2 <- base.model1.2$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps2 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.2`") %>%
+    kable_styling(latex_options = c("scale_down",
+                                    "striped"),# Expande la tabla al ancho completo
+                font_size = 8) #%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Outputs Variables `s1.3`
+
+```{r}
+outps3 <- base.model1.3$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps3 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.3`") %>%
+  kable_styling(latex_options = c("scale_down",
+                                    "striped"),# Expande la tabla al ancho completo
+                font_size = 8) #%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Outputs Variables `s1.4`
+
+```{r}
+outps4 <- base.model1.3$timeseries[1:37, 2:8]
+
+# addWorksheet(wb, "variable_s2")
+# writeData(wb, "variable_s2", outps1)
+# 
+# # Guardar el workbook
+# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
+# 
+# 
+# out_s2 <- read_excel("DataKrill.xlsx", 
+#     sheet = "variable_s2")
+outps4 %>%
+  kbl(booktabs = TRUE,
+      format = "html",
+    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.4`") %>%
+ kable_styling(latex_options = c("scale_down",
+                                    "striped"),# Expande la tabla al ancho completo
+                font_size = 8) #%>% 
+  #pack_rows(index = c("Estimation" = 1,
+   #                     "Prediction" = 45))
+```
+
+### Comparision between variables
+
+Main variables population by scenario (Figure \@ref(fig:popvar2))
+
+```{r popvar2, fig.height=4, fig.width=9, fig.cap="Time series of different populations variables"}
+# Unir los data frames en uno solo
+outpsall <- rbind(outps1,  
+                  outps2,
+                  outps3,
+                outps4)
+
+# Crear una columna de modelo en cada data frame
+outps1$Model <- "s1.1"
+outps2$Model <- "s1.2"
+outps3$Model <- "s1.3"
+outps4$Model <- "s1.4"
+
+# Unir los data frames en uno solo
+outpsall <- rbind(outps1,  
+                  outps2,
+                  outps3,
+                  outps4)
+
+outpsall_long <- outpsall %>%
+  pivot_longer(cols = c(Bio_all, Bio_smry, SpawnBio, Recruit_0), 
+               names_to = "Variable", values_to = "Value") %>% 
+  filter(Yr >= 1989, Yr <= 2020)
+
+# Gráfico de líneas por variable y modelo
+alloutput <- ggplot(outpsall_long, aes(x = Yr, y = Value, color = Model, group = Model)) +
   geom_point() +
-  geom_smooth(method = "loess",
-              span = 0.5,
-              se = FALSE) +
+  #geom_smooth(method = "lm")+
+  facet_wrap(~Variable, 
+             ncol = 4, 
+             scales = "free_y") + 
+  scale_color_manual(name = "Scenario",
+                     values = c("s1.1" = "#ca0020", 
+                                "s1.2" = "#f4a582", 
+                                "s1.3" = "#bababa", 
+                                "s1.4" = "#404040")) +
+  labs(title = "", x = "", y = "") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        legend.position = "bottom")
+
+alloutput
+
+```
+
+```{r eval=FALSE}
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist1_4 <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir1.1,
+                                  dir1.2,
+                                  dir1.3,
+                                  dir1.4),
+                       getcovar = F)
+
+#create summary of model runs from list above
+summaryoutput1_4 <- SSsummarize(biglist1_4)
+
+SSplotComparisons(summaryoutput1_4,
+                  legendlabels = c("s1.1 (Ref Model)",
+                 "s1.2",
+                 "s1.3",
+                 "s1.4"),
+                 filenameprefix = "COM1",
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"),
+                 png=TRUE,
+                 plotdir=Figs)
+```
+
+```{r eval=FALSE}
+#read in all model runs
+#note if cover=T you need a hessian; if covar=F you do not need a hessian
+biglist14 <- SSgetoutput(keyvec = NULL,
+                       dirvec = c(
+                                  dir1.1,
+                                  dir1.4),
+                       getcovar = F)
+
+#create summary of model runs from list above
+summaryoutput14 <- SSsummarize(biglist14)
+
+comp14 <- SSplotComparisons(summaryoutput14,
+                  legendlabels = c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Env and Predator data"),
+                  filenameprefix = "COM2",
+                 labels = c("Year", 
+                            "Spawning biomass (t)",
+                            "Relative spawning biomass", 
+                            "Age-0 recruits (1,000s)",
+                            "Recruitment deviations", 
+                            "Index", "Log index", 
+                            "1 - SPR", 
+                            "Density",
+                            "Management target", 
+                            "Minimum stock size threshold",
+                            "Spawning output",
+                            "Harvest rate"),
+                 png=TRUE,plotdir=Figs)
+```
+
+Another comparison in `R0` (Natural log of virgin recruitment level)  (Figure \@ref(fig:r0))
+
+```{r r0, fig.height=5, fig.width=8, warning=FALSE, message=FALSE, fig.cap="R0 probability predicted by scenario"}
+r0_vals <- data.frame(
+  model = c("s1.1", "s1.2", "s1.3", "s1.4"),
+  value = c(base.model1.1$jitter_info[1,1],
+            base.model1.2$jitter_info[2,1],
+            base.model1.3$jitter_info[1,1],
+            base.model1.4$jitter_info[2,1]),
+  cv = c(base.model1.1$jitter_info[1,6],
+         base.model1.2$jitter_info[2,6],
+         base.model1.3$jitter_info[1,6],
+         base.model1.4$jitter_info[2,6])
+)
+
+x_vals <- seq(10, 30, length.out = 1000)
+
+dens_data <- r0_vals %>%
+  rowwise() %>%
+  mutate(sd =  cv,  
+         density = list(data.frame(
+           x = x_vals,
+           y = dnorm(x_vals, mean = value, sd = sd),
+           model = model
+         ))) %>%
+  select(density) %>%
+  unnest(cols = c(density))
+
+colores <- c("s1.1" = "#ca0020", 
+             "s1.2" = "#f4a582", 
+             "s1.3" = "#bababa", 
+             "s1.4" = "#404040")
+
+r0plot <- ggplot(dens_data, aes(x = x, y = y, color = model, fill = model)) +
+  geom_area(alpha = 0.3, position = "identity", size = 0) +  # More transparent areas without borders
+  geom_line(size = 1.2) +  # Thicker lines for better visibility
+  scale_color_manual(values = colores, name = NULL) +  # Remove legend title for color
+  scale_fill_manual(values = colores, name = NULL) +   # Remove legend title for fill
+  labs(x = "ln(R0)", 
+       y = "Density") +
+  theme_bw(base_size = 14) +
+  theme(
+    legend.position = "bottom",  # Legend at bottom
+    legend.spacing.x = unit(0.5, 'cm'),  # Space between legend items
+    panel.grid.major = element_line(color = "gray90"),
+    panel.grid.minor = element_blank(),
+    axis.title = element_text(size = 12),
+    legend.text = element_text(size = 11)
+  ) +
+  xlim(16, 22) +
+  guides(color = guide_legend(nrow = 1),  # Single row legend
+         fill = guide_legend(nrow = 1))
+
+r0plot
+```
+Comparsion in sd long term time series forecasting Figure \@ref(fig:cumsum)
+
+```{r cumsum, fig.height=5, fig.width=8, fig.cap="Comparsion in sd long term time series forecasting"}
+errt <- data.frame(
+  Bio_all_1 = base.model1.1$timeseries$Bio_all,
+  Bio_all_2 = base.model1.2$timeseries$Bio_all,
+  Bio_all_3 = base.model1.3$timeseries$Bio_all,
+  Bio_all_4 = base.model1.4$timeseries$Bio_all,
+
+  Bio_smry_1 = base.model1.1$timeseries$Bio_smry,
+  Bio_smry_2 = base.model1.2$timeseries$Bio_smry,
+  Bio_smry_3 = base.model1.3$timeseries$Bio_smry,
+  Bio_smry_4 = base.model1.4$timeseries$Bio_smry,
+
+  SpawnBio_1 = base.model1.1$timeseries$SpawnBio,
+  SpawnBio_2 = base.model1.2$timeseries$SpawnBio,
+  SpawnBio_3 = base.model1.3$timeseries$SpawnBio,
+  SpawnBio_4 = base.model1.4$timeseries$SpawnBio,
+
+  Recruit_0_1 = base.model1.1$timeseries$Recruit_0,
+  Recruit_0_2 = base.model1.2$timeseries$Recruit_0,
+  Recruit_0_3 = base.model1.3$timeseries$Recruit_0,
+  Recruit_0_4 = base.model1.4$timeseries$Recruit_0
+)
+
+err_long <- errt %>%
+  pivot_longer(cols = everything(), 
+               names_to = "Variable_Model",
+               values_to = "Value")
+
+err_long <- err_long %>%
+  extract(Variable_Model, into = c("Variable", "Model"), 
+          regex = "(.+?)_(\\d+)$")
+
+err_long <- err_long %>%
+  mutate(Model = case_when(
+    Model == "1" ~ "Ref Model: No Env-Predator",
+    Model == "2" ~ "S1.1 w/ Predator data",
+    Model == "3" ~ "S1.1 w/ Env data",
+    Model == "4" ~ "S1.1 w/ Env and Predator data",
+    TRUE ~ Model
+  ))
+
+comsum <- ggplot(err_long, aes(x = Model, y = Value, fill = Model)) +
+  geom_boxplot(width = 0.5, outlier.shape = NA,
+               outliers = F) +
+  #geom_jitter(width = 0.2, alpha = 0.1) +
+  facet_wrap(~Variable, ncol=4, 
+             scales = "free_y") +
+  scale_fill_manual(name = "Scenario",
+                    values = c("#ca0020", "#f4a582", "#bababa", "#404040")) +
   labs(title = "",
        x = "",
-       y = "Total biomass (mt)") +
+       y = "") +
   theme_few() +
-  scale_color_manual(name = "",
-                     values = c("Without Ecosystem Variables" = "#ca0020",  
-                                "W/ Predator" = "#f4a582",  
-                                "W/ Environment" = "#bababa",  
-                                "With Predator and Ecosystem" = "#404040")) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1),
+        legend.position = "none")
+comsum
+```
+
+## Autocorrelation in Recruit
+
+To evaluate the temporal correlation structure of krill recruitment under different model configurations, we performed an Autocorrelation Function (ACF) analysis (Figure \@ref(fig:acf)).
+The ACF measures the correlation between observations at different time lags, helping to assess whether recruitment estimates exhibit persistence or randomness across time.
+
+The analysis was conducted on recruitment estimates derived from four model scenarios:
+
+A reference model without environmental or predator influences (Ref Model: No Env-Predator).
+A model incorporating predator data (S1.1 w/ Predator data).
+A model incorporating environmental data (S1.1 w/ Env data).
+A model incorporating both environmental and predator data (S1.1 w/ Env and Predator data).
+Each model's residuals were extracted, and the autocorrelation function (ACF) was computed for a time lag range of up to 15 years.
+The dashed blue lines in the plots represent the 95% confidence intervals, indicating the threshold beyond which correlation values are statistically significant.
+If autocorrelation values remain within this range, it suggests that the recruitment estimates behave as a random process with no significant dependence on past values.
+Conversely, autocorrelation values exceeding these bounds indicate recruitment persistence or cyclic patterns.
+
+```{r acf, fig.height=5, fig.width=8, fig.cap="Autocorrelation in Recruit by scenario"}
+recs1 <- base.model1.1$recruit$pred_recr
+recs2 <- base.model1.2$recruit$pred_recr
+recs3 <- base.model1.3$recruit$pred_recr
+recs4 <- base.model1.4$recruit$pred_recr
+
+p1 <- ggAcf(base.model1.1$recruit$pred_recr) +
+  ggtitle("s1.1") +
+  theme_few()
+p2 <- ggAcf(base.model1.2$recruit$pred_recr) +
+  ggtitle("s1.2") +
+  theme_few()
+p3 <- ggAcf(base.model1.3$recruit$pred_recr) +
+  ggtitle("s1.3") +
+  theme_few()
+p4 <- ggAcf(base.model1.4$recruit$pred_recr) +
+  ggtitle("s1.4") +
+  theme_few()
+
+ggarrange(p1, p2,p3,p4, ncol=4, nrow = 1, common.legend = T)
+
+```
+
+The ACF plots indicate that the reference model (without environmental or predator data) exhibits weak but noticeable positive autocorrelation at certain lags, suggesting some degree of recruitment dependence over time.
+However, this autocorrelation does not appear strong or systematic.
+
+The model incorporating only predator data shows a slight reduction in autocorrelation magnitude, suggesting that predator-driven recruitment variability may have captured part of the temporal structure in the data.
+
+The model with only environmental data exhibits a further reduction in autocorrelation, implying that environmental variability explains a larger portion of recruitment trends than predator effects alone.
+
+Finally, the model that includes both environmental and predator data presents the lowest autocorrelation values, with nearly all bars remaining within the confidence bounds.
+This suggests that incorporating both factors provides the most effective explanation of recruitment fluctuations, reducing unexplained temporal structure.
+
+Overall, these results indicate that recruitment variability is at least partially driven by environmental and predator influences, and models integrating these factors provide more robust and independent recruitment estimates, minimizing systematic dependencies over time.
+
+## Relationship Stock-Recruit
+
+\[
+R = \frac{R_0 \cdot S}{S_0 (1 - h) + S (5h - 1)}
+\]
+
+Where:  
+- \( R \) is the predicted recruitment.  
+- \( S \) is the spawning stock biomass.  
+- \( R_0 \) is the recruitment at unfished equilibrium.  
+- \( S_0 \) is the spawning biomass at unfished equilibrium.  
+- \( h \) is the steepness parameter (the proportion of \( R_0 \) produced when \( S = 20\% \cdot S_0 \)).
+
+The blue line represents the Beverton–Holt stock-recruitment relationship, commonly used in fisheries models to describe the compensatory response of recruitment to changes in spawning biomass.
+
+```{r}
+# Lista de modelos base
+models <- list(
+  s1.1 = base.model1.1,
+  s1.2 = base.model1.2,
+  s1.3 = base.model1.3,
+  s1.4 = base.model1.4
+)
+
+colors <- c("s1.1" = "#ca0020", "s1.2" = "#f4a582", "s1.3" = "#bababa", "s1.4" = "#404040")
+shapes <- c("s1.1" = 5, "s1.2" = 6, "s1.3" = 15, "s1.4" = 4)
+
+plot_data <- list()
+curve_data <- list()
+label_data <- list()
+params <- data.frame(Model = character(), a = numeric(), b = numeric(), stringsAsFactors = FALSE)
+for (group in names(models)) {
+  model <- models[[group]]
+  l <- model$recruit$SpawnBio
+  m <- model$recruit$pred_recr
+  yr <- model$timeseries$Yr[1:35]
+
+  # Filtrar valores válidos
+  valid <- which(is.finite(l) & is.finite(m) & l > 0 & m > 0)
+  l <- l[valid]
+  m <- m[valid]
+  yr <- yr[valid]
+
+  plot_data[[group]] <- data.frame(l = l, m = m, yr = yr, Group = group)
+  label_data[[group]] <- plot_data[[group]] %>% filter(yr %% 3 == 0)
+
+  fit <- NULL
+  tryCatch({
+    fit <- nls(m ~ (a * l) / (1 + b * l),
+               start = list(a = max(m), b = 1 / max(l)),
+               data = data.frame(l, m))
+
+    l_pred <- seq(min(l), max(l), length.out = 100)
+    m_pred <- predict(fit, newdata = data.frame(l = l_pred))
+    curve_data[[group]] <- data.frame(l = l_pred, m = m_pred, Group = group)
+
+    coefs <- coef(fit)
+    params <- rbind(params, data.frame(Model = group, a = coefs["a"], b = coefs["b"]))
+
+  }, error = function(e) {
+    message(paste("Model", group, ": Beverton-Holt fit failed ->", e$message))
+  })
+}
+
+# Unir data frames
+plot_data_all <- bind_rows(plot_data)
+curve_data_all <- bind_rows(curve_data)
+label_data_all <- bind_rows(label_data)
+
+# Graficar
+srr_sce <- ggplot() +
+  geom_point(data = plot_data_all, aes(x = l, y = m, shape = Group), size = 2) +
+  geom_line(data = curve_data_all, aes(x = l, y = m, color = Group), linewidth = 1.2) +
+  ggrepel::geom_text_repel(data = label_data_all, aes(x = l, y = m, label = yr, color = Group),
+                           size = 4, max.overlaps = 50) +
+  scale_shape_manual(values = shapes) +
+  scale_color_manual(values = colors) +
+  theme_bw() +
+  ylim(0, max(plot_data_all$m)) +
+  labs(x = "Spawning Stock Biomass (SSB)", y = "Predicted Recruitment", shape = "Model", color = "Model") +
+  theme(legend.position = "bottom")
+
+```
+
+## Productivity and Interannual Variability by Scenario
+
+Estimating the productivity of Antarctic krill is critical for understanding the species’ capacity to replenish its population in response to varying levels of spawning biomass.
+Productivity, defined as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
+Comparing productivity across scenarios—each representing different assumptions about environmental drivers, fishing mortality, or predator dynamics—enables a robust evaluation of how krill populations reflect this changes.
+
+For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):
+
+$$
+\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
+$$
+
+We also calculated the **interannual percentage change** in recruitment and SSB as:
+
+$$
+\text{Change in Recruitment}_{i,t} = \left( \frac{\text{Recruitment}_{i,t} - \text{Recruitment}_{i,t-1}}{\text{Recruitment}_{i,t-1}} \right) \times 100
+$$
+
+$$
+\text{Change in SSB}_{i,t} = \left( \frac{\text{SSB}_{i,t} - \text{SSB}_{i,t-1}}{\text{SSB}_{i,t-1}} \right) \times 100
+$$
+
+These metrics allow us to analyze both the productivity and the temporal dynamics of the population under each scenario \( i \).
+
+```{r}
+# Simular 4 dataframes por escenario (usa tus datos reales)
+escenarios <- list(
+  s1.1 = base.model1.1$SPAWN_RECR_CURVE,
+  s1.2 = base.model1.2$SPAWN_RECR_CURVE,
+  s1.3 = base.model1.3$SPAWN_RECR_CURVE,
+  s1.4 = base.model1.4$SPAWN_RECR_CURVE
+)
+
+# Crea una tabla con productividad y cambios para cada escenario
+resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
+  df %>%
+    mutate(
+      Scenario = nombre,
+      Productividad = Recruitment / SSB,
+      Cambio_Recruit = c(NA, diff(Recruitment)) / lag(Recruitment) * 100,
+      Cambio_SSB = c(NA, diff(SSB)) / lag(SSB) * 100
+    )
+})
+
+prod <- ggplot(resultados, aes(x = `SSB/SSB_virgin`, y = Productividad, 
+                       color = Scenario)) +
+  geom_point(size = 1.1) +
+  labs(title = "",
+       y = "Recruitment / SSB",
+       x = "SSB / SSB_virgin") +
+  scale_color_manual(name= "",
+                    values = c("s1.1" = "#ca0020", 
+                               "s1.2" = "#f4a582", 
+                               "s1.3" = "#bababa", 
+                               "s1.4" = "#404040")) +
+  theme_bw() +
+  theme(legend.position = "bottom")
+```
+
+These two panels in Figure \@ref(fig:recpro) provide insight into the stock-recruitment dynamics of Antarctic krill under four assessment scenarios, each incorporating different levels of ecosystem complexity.
+Left panel compares the predicted recruitment across years with corresponding spawning biomass values for the four model scenarios (s1.1 to s1.4).
+Scenario s1.1 (red) shows a relatively flat and optimistic Beverton-Holt relationship, with high recruitment predicted even at low levels of SSB.
+The other scenarios, particularly s1.3 (gray) and s1.4 (black), show much more constrained recruitment estimates and a more saturating relationship, where recruitment increases only slightly with increasing SSB.
+This suggests that incorporating environmental (s1.3) and predator (s1.4) effects leads to a more conservative and ecologically realistic recruitment dynamic.
+Year labels illustrate how certain years (e.g., 2004, 2016, 2022) deviate between scenarios, especially at low biomass levels.
+
+Right panel displays recruitment efficiency (Recruitment/SSB) as a function of relative spawning biomass (SSB / SSB₀).
+In all models, recruitment efficiency is highest at very low biomass levels and declines as SSB increases, consistent with compensatory dynamics.
+However, s1.3 and s1.4 (which include environmental effects) show significantly lower overall productivity across all SSB levels compared to s1.1 and s1.2.
+This indicates that the inclusion of ecosystem drivers reduces the per-capita productivity of the stock, especially under depleted conditions.
+
+Together, these plots indicate that ecosystem-informed models (s1.3 and s1.4) produce more constrained and precautionary estimates of krill productivity.
+The more optimistic recruitment predicted by s1.1 (and to a lesser extent s1.2) may overestimate stock resilience, particularly in low SSB conditions.
+These findings emphasize the importance of accounting for environmental variability and predator dynamics in stock assessment models to avoid overoptimistic management advice and better reflect the biological limits of the krill population.
+
+```{r recpro, fig.height=4, fig.width=6, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios.
+Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
+Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
+ggarrange(srr_sce,
+          prod,
+          ncol = 2)
+```
+
+## Recruit deviation
+
+```{r, warning=FALSE}
+dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
+  mutate(Serie = "s1.1")
+dev2 <- base.model1.2$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.2")
+dev3 <- base.model1.3$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.3")
+dev4 <- base.model1.4$recruit[8:35,c(1,7)] %>% 
+  mutate(Serie = "s1.4")
+
+data_list <- list(dev1, 
+                  dev2,
+                  dev3,
+                  dev4)
+titles <- c("Ref Model: No Env-Predator",
+                 "S1.1 w/ Predator data",
+                 "S1.1 w/ Env data",
+                 "S1.1 w/ Env and Predator data")
+# Lista para almacenar los gráficos
+plots <- list()
+
+for (i in 1:length(data_list)) {
+  p <- ggplot(data_list[[i]], aes(x = Yr, y = dev)) +
+    geom_point() +
+    stat_smooth(method = "loess", 
+                span = 0.15, 
+                se = T) +
+    geom_hline(yintercept = 0,
+               col="red",
+               linetype = "dotdash")+
+    ggtitle(titles[i]) +
     theme(axis.text.x = element_text(angle = 90,
-                                   hjust = 1))
-
-```
-
-Pearson residual
-
-```{r}
+                                   hjust = 1))+
+    ylim(-2,2)+
+    theme_few()
+  plots[[i]] <- p
+}
+
+# Organizar los gráficos con ggarrange
+combined_plot <- ggarrange(plotlist = plots, ncol = 2, nrow = 2)
+
+combined_plot
+```
+
+## Explotation Rate
+
+Explotation rato (havest rate) in Figure \@ref(fig:hrate)
+
+```{r hrate, warning=FALSE, message=FALSE, fig.cap="Harves rate by scenario in krill overtime"}
+df1 <- base.model1.1$exploitation[, c(1, 4)] %>% mutate(model = "s1.1")
+df2 <- base.model1.2$exploitation[, c(1, 4)] %>% mutate(model = "s1.2")
+df3 <- base.model1.3$exploitation[, c(1, 4)] %>% mutate(model = "s1.3")
+df4 <- base.model1.4$exploitation[, c(1, 4)] %>% mutate(model = "s1.4")
+
+df_all <- bind_rows(df1, df2, df3, df4)
+colnames(df_all) <- c("year", "exploitation_rate", "model")
+
+colores <- c("s1.1" = "#ca0020", 
+             "s1.2" = "#f4a582", 
+             "s1.3" = "#bababa", 
+             "s1.4" = "#404040")
+ggplot(df_all, aes(x = year, y = exploitation_rate, color = model)) +
+  geom_line() +
+  geom_point() +
+  scale_color_manual(values = colores) +
+  labs(
+    title = "Exploitation Rate Over Time",
+    x = "Year",
+    y = "Exploitation Rate",
+    color = "Model"
+  ) +
+  xlim(1997,2020)+
+  theme_minimal()
+```
+
+<!-- ### Platoons analisis -->
+
+<!-- ```{r} -->
+<!-- # Definir los valores del eje x -->
+<!-- x_values <- seq(0, 7, by=0.1) -->
+
+<!-- # Definir las medias y desviaciones estándar para los valores específicos -->
+<!-- means <- c( 2.9,  3.5, 4.1) -->
+<!-- std_devs <- c(0.65, 0.8, 0.65) -->
+
+<!-- # Crear un dataframe que contenga los valores de x, la media y las desviaciones estándar -->
+<!-- data <- data.frame( -->
+<!--   x = rep(x_values, each = length(means)), -->
+<!--   mean = rep(means, times = length(x_values)), -->
+<!--   std_dev = rep(std_devs, times = length(x_values)) -->
+<!-- ) -->
+
+<!-- # Calcular las curvas de las desviaciones estándar -->
+<!-- data <- data %>% -->
+<!--   mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2))) -->
+
+<!-- # Graficar con ggplot -->
+<!-- ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) + -->
+<!--   geom_line(size = 1) + -->
+<!--   scale_linetype_manual(values = means, -->
+<!--                         name ="Platoon") + -->
+<!--   labs(x = "", y = "growth increment") + -->
+<!--   theme_minimal()+ -->
+<!--   xlim(0,7) -->
+
+<!-- ``` -->
+
+## Diagnosis and robustness
+
+A rigorous model diagnosis is essential to ensure the reliability and robustness of stock assessment models. The key steps for a good practice in model diagnosis include:
+
+1. Convergence Check: The model must reach a final convergence criterion of 1.0e-04 to ensure numerical stability and reliable parameter estimation.
+
+2. Residual Analysis: Both visual inspection and statistical metrics are used to evaluate model residuals, helping to detect patterns of bias or misfit.
+
+3.
+Retrospective Analysis: The Mohn’s rho parameter is used to assess the consistency of model estimates when sequentially removing recent years of data, identifying potential overestimation or underestimation trends.
+
+4. Likelihood Profile Analysis: This approach examines how the likelihood function behaves across a range of parameter values, providing insight into parameter uncertainty and model sensitivity.
+
+This framework follows the recommendations outlined by @Carvalho2021b, aiming to enhance transparency and reproducibility in model evaluation.
+
+### Convergence Criteria
+
+The convergence criterion used for model calibration is set to a final threshold of **0.0001** (or equivalently **1.0e-04**).
+This criterion defines the minimum acceptable difference between successive model iterations.
+Convergence is considered achieved when the absolute change in the objective function value or key parameters falls below this threshold.
+A smaller convergence value ensures that the model achieves a high degree of accuracy and stability in its final estimates, indicating that further iterations are unlikely to result in significant changes to the parameter estimates.
+
+### Residuals
+
+Figure \@ref(fig:bubble1), \@ref(fig:bubble2), \@ref(fig:bubble3) and \@ref(fig:bubble4) displays bubble plots of krill length-frequency distributions across different spatial strata and years.
+Each panel corresponds to a fishery-dependent or independent data source (e.g., FISHERYBS, SURVEYBS), and shows the proportional abundance of krill in different length classes (y-axis) over time (x-axis).
+The size of each circle is proportional to the relative frequency or standardized number of individuals in a given length bin for a specific year, providing a visual summary of size structure dynamics across regions and time periods.
+
+```{r bubble, fig.width=8, fig.height=6, fig.cap="Pearson residuals, comparing across fleets (plot 1 of 2) Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) s1.1"}
+SSplotComps(base.model1.1,
+            subplots = 24,
+            maxrows = 9)
+```
+
+```{r bubble2,  fig.width=8, fig.height=6, fig.cap="Pearson residuals, comparing across fleets (plot 1 of 2) Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) s1.2"}
+SSplotComps(base.model1.2,
+            subplots = 24,
+            maxrows = 10)
+```
+
+```{r bubble3,  fig.width=8, fig.height=6, fig.cap="Pearson residuals, comparing across fleets (plot 1 of 2) Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) s1.3"}
+SSplotComps(base.model1.3,
+            subplots = 24,
+            maxrows = 9)
+```
+
+```{r bubble4,  fig.width=8, fig.height=6, fig.cap="Pearson residuals, comparing across fleets (plot 1 of 2) Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected) s1.4"}
+SSplotComps(base.model1.4,
+            subplots = 24,
+            maxrows = 10)
+```
+
+This Figure \@ref(fig:pearson) shows the Pearson residuals of predicted length distributions for krill across four modeling scenarios, each incorporating different levels of ecosystem complexity.
+The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population.
+In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
+Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
+The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals.
+However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
+In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
+This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
+Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure.
+The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
+
+Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the **joint inclusion of predator and environmental variables** yielding the most robust representation of observed length compositions.
+
+```{r pearson, fig.cap="Pearson residual by scenario and fleet"}
 create_heatmap_df <- function(df, model_name) {
@@
 
-# Plot combinado: facet_grid por Modelo y Fleet
-heatmapres <- ggplot(df_all, aes(x = Yr, y = Bin, fill = Pearson)) +
-  geom_tile() +
-  scale_fill_gradient2(low = "red", 
-                       mid = "white", 
-                       high = "blue",
-                       midpoint = 0) +
-  facet_grid(Model ~ Fleet) +
-  theme_few() +
-  theme(
-    axis.text.x = element_text(angle = 90, hjust = 1, size = 5),         
-    axis.text.y = element_text(size = 5),                               
-    strip.text = element_text(size = 8),                                 
-    legend.position = "none"
-  ) +
-  labs(x = "", y = "Length (cm)", fill = "Pearson")
-
-heatmapres
+# # Plot combinado: facet_grid por Modelo y Fleet
+# heatmapres <- ggplot(df_all, aes(x = Yr, y = Bin, fill = Pearson)) +
+#   geom_tile() +
+#   scale_fill_gradient2(low = "red", 
+#                        mid = "white", 
+#                        high = "blue",
+#                        midpoint = 0) +
+#   facet_grid(Model ~ Fleet) +
+#   theme_few() +
+#   theme(
+#     axis.text.x = element_text(angle = 90, hjust = 1, size = 5),         
+#     axis.text.y = element_text(size = 5),                               
+#     strip.text = element_text(size = 8),                                 
+#     legend.position = "none"
+#   ) +
+#   labs(x = "", y = "Length (cm)", fill = "Pearson")
+# 
+# heatmapres
 
@@
 
-This figure shows the Pearson residuals of predicted length distributions for krill across four modeling scenarios, each incorporating different levels of ecosystem complexity.
-The residuals are visualized by year and length bin, allowing an assessment of model fit over time and across the size structure of the population.
-In the "Without Ecosystem Variables" scenario (top left), substantial lack of fit is evident across multiple years and length classes, particularly in the late 1990s and early 2000s.
-Positive Pearson residuals indicate that the model tends to underestimate frequencies of certain length bins, especially among smaller and mid-sized individuals, suggesting poor representation of recruitment and growth dynamics in the absence of environmental and predator effects.
-The "With Predator" scenario (top right) shows some improvement in model fit, with slightly more balanced residuals.
-However, there are still consistent deviations in the mid-size ranges across multiple years, indicating that predation alone does not fully explain observed variability in length composition.
-In the "With Environment" scenario (bottom left), there is a noticeable reduction in residual magnitude across several years and size classes, implying that incorporating environmental covariates helps align the model more closely with observed size distributions.
-This is particularly apparent in the early 2000s, where the residuals are closer to zero across most bins.
-Finally, the "With Predator and Ecosystem" scenario (bottom right) shows the most consistent reduction in Pearson residuals across time and size structure.
-The distribution of residuals is more homogeneous and centered around zero, indicating that the combined inclusion of predator and environmental drivers leads to the best model fit among the four scenarios.
-
-Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the **joint inclusion of predator and environmental variables** yielding the most robust representation of observed length compositions.
-
-## Diagnosis Base Model
-
-A rigorous model diagnosis is essential to ensure the reliability and robustness of stock assessment models. The key steps for a good practice in model diagnosis include:
-
-1. Convergence Check: The model must reach a final convergence criterion of 1.0e-04 to ensure numerical stability and reliable parameter estimation.
-
-2. Residual Analysis: Both visual inspection and statistical metrics are used to evaluate model residuals, helping to detect patterns of bias or misfit.
-
-3.
-Retrospective Analysis: The Mohn’s rho parameter is used to assess the consistency of model estimates when sequentially removing recent years of data, identifying potential overestimation or underestimation trends.
-
-4. Likelihood Profile Analysis: This approach examines how the likelihood function behaves across a range of parameter values, providing insight into parameter uncertainty and model sensitivity.
-
-This framework follows the recommendations outlined by @Carvalho2021b, aiming to enhance transparency and reproducibility in model evaluation.
-
-### Convergence Criteria
-
-The convergence criterion used for model calibration is set to a final threshold of **0.0001** (or equivalently **1.0e-04**).
-This criterion defines the minimum acceptable difference between successive model iterations.
-Convergence is considered achieved when the absolute change in the objective function value or key parameters falls below this threshold.
-A smaller convergence value ensures that the model achieves a high degree of accuracy and stability in its final estimates, indicating that further iterations are unlikely to result in significant changes to the parameter estimates.
-
+```{r}
+cpue_resid <- bind_rows(
+  base.model1.1$cpue %>% mutate(model = "s1.1"),
+  base.model1.2$cpue %>% mutate(model = "s1.2"),
+  base.model1.3$cpue %>% mutate(model = "s1.3"),
+  base.model1.4$cpue %>% mutate(model = "s1.4")
+)
+
+len_resid <- bind_rows(
+  base.model1.1$lendbase %>% mutate(model = "s1.1"),
+  base.model1.2$lendbase %>% mutate(model = "s1.2"),
+  base.model1.3$lendbase %>% mutate(model = "s1.3"),
+  base.model1.4$lendbase %>% mutate(model = "s1.4")
+)
+
+cpue_df <- cpue_resid %>%
+  transmute(
+    Yr = Yr,
+    Obs = Obs,
+    Exp = Exp,
+    model = model,
+    Fleet_name = as.character(Fleet_name),  # Asegura que es character
+    type = "Index"
+  )
+
+len_df <- len_resid %>%
+  transmute(
+    Yr = Yr,
+    Obs = Obs,
+    Exp = Exp,
+    model = model,
+    Fleet_name = as.character(Fleet),  # Lo convertimos para que combine bien
+    type = "Length"
+  )
+
+resid_all <- rbind(cpue_df,
+                   len_df)
+
+resid_all <- resid_all %>%
+  mutate(Fleet_name = as.character(Fleet_name),
+         Fleet_name = case_when(
+           Fleet_name == "1"  ~ "FISHERYBS",
+           Fleet_name == "2"  ~ "FISHERYEI",
+           Fleet_name == "3"  ~ "FISHERYGS",
+           Fleet_name == "4"  ~ "FISHERYJOIN",
+           Fleet_name == "5"  ~ "FISHERYSSIW",
+           Fleet_name == "6"  ~ "SURVEYBS",
+           Fleet_name == "7"  ~ "SURVEYEI",
+           Fleet_name == "8"  ~ "SURVEYGS",
+           Fleet_name == "9"  ~ "SURVEYJOIN",
+           Fleet_name == "10" ~ "SURVEYSSIW",
+           Fleet_name == "11" ~ "PREDATOR",
+           TRUE ~ Fleet_name  # mantiene los nombres ya correctos
+         ))
+
+```
+
+Also, we can check density of residual distribution in Figure \@ref(fig:resldens)
+
+```{r resldens, fig.width=10, fig.height=6, fig.cap="Density of residual distribution by scenarios in krill model"}
+ggplot(resid_all, aes(x = Obs - Exp, color = model)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(Fleet_name~type, scales = "free") +
+  labs(title = "Residual Distribution", 
+       x = "Residual (Obs - Exp)", 
+       y = "Density") +
+  scale_color_manual(name = "",
+                     values = c("s1.1" = "#ca0020",  
+                                "s1.2" = "#f4a582",  
+                                "s1.3" = "#bababa",  
+                                "s1.4" = "#404040")) +
+  theme_minimal()
+```
 ### Residual consistency
@@
 This process is applied to both length composition data and abundance indices such as CPUE (Catch Per Unit Effort) and survey-derived estimates.
-
 For length composition data, residuals represent the difference between observed and model-predicted length distributions.
@@
 When mean residual values are close to zero, the model fit is considered unbiased.
-
 By integrating these residual analyses for both length and abundance indices, stock assessment models can be refined, improving their reliability and increasing confidence in the assessment results.
@@
 
-By analyzing residual patterns and RMSE values, the model can be refined to improve the accuracy of mean length predictions, ultimately enhancing the reliability of stock assessment outcomes and management recommendations.
-
-```{r}
+By analyzing residual patterns and RMSE values, the model can be refined to improve the accuracy of mean length predictions, ultimately enhancing the reliability of stock assessment outcomes and management recommendations (Figure \@ref(fig:rmse1)).
+
+```{r rmse1, fig.height=8, fig.width=8, fig.cap="Time series of RMSE of length compositions by scenario"}
 par(mfrow=c(2,2), mar=c(5,4,2,1))
@@
 ```
-
-```{r}
+Figure \@ref(fig:rmse2) show RMSE to index.
+
+```{r rmse2, fig.height=8, fig.width=8, fig.cap="Time series of RMSE of CPUE compositions by scenario"}
 par(mfrow=c(2,2), mar=c(5,4,2,1))
@@
 ```
-
-another
+### Comparision RMSE
 
@@
 
-### Comparision RMSE
-
-```{r fig.height=3, fig.width=3}
+```{r}
 dfrmse <- data.frame(
@@
   s1.4 = as.numeric(base.model1.4$index_variance_tuning_check$RMSE[1:10])
-
-)
-summary(dfrmse)
-
-t.test(dfrmse$s1.1, dfrmse$s1.2)  # Comparar s1.1 vs s1.2
-t.test(dfrmse$s1.1, dfrmse$s1.3)  # Comparar s1.1 vs s1.3
-t.test(dfrmse$s1.2, dfrmse$s1.3)
-t.test(dfrmse$s1.3, dfrmse$s1.4)
-anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
-
+  )
+#summary(dfrmse)
+# t.test(dfrmse$s1.1, dfrmse$s1.2)  # Comparar s1.1 vs s1.2
+# t.test(dfrmse$s1.1, dfrmse$s1.3)  # Comparar s1.1 vs s1.3
+# t.test(dfrmse$s1.2, dfrmse$s1.3)
+# t.test(dfrmse$s1.3, dfrmse$s1.4)
+# anova_rmse <- aov(as.numeric(unlist(dfrmse)) ~ rep(1:4, each=nrow(dfrmse)))
+```
+
+```{r}
 dfrmse_long <- reshape2::melt(dfrmse)
@@
                outliers = FALSE) +
-  #geom_jitter(width = 0.01, alpha = 0.3) +
+  geom_jitter(width = 0.01, alpha = 0.3) +
   theme_minimal() +
@@
 
-```{r}
+The Figure \@ref(fig:rmse3) presents evaluation results for krill models across different scenarios (s1.1 to s1.4).
+The RMSE (Root Mean Square Error) values range from 0.5 to 1.5, with lower values indicating better model accuracy.
+The Residuals vary between -1.0 and 1.0, showing deviations between predicted and observed values.
+Smaller residuals suggest better model fit.
+Scenario s1.4 appears to have the lowest RMSE (0.5), indicating higher precision, while residuals near zero in some scenarios reflect balanced predictions.
+
+```{r rmse3, fig.cap="RMSE and Residual values for krill model evaluation across scenarios (s1.1-s1.4), highlighting precision and prediction errors."}
 ggarrange(rmse,
@@
 
-This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4) used to evaluate recruitment estimates of Antarctic krill.
+This boxplot compares a summary of the Root Mean Square Error (RMSE) across four different models (s1.1, s1.2, s1.3, and s1.4).
 RMSE serves as an indicator of model accuracy, with lower values representing better predictive performance.
@@
 
-### Relationship Stock-Recruit
-
-\[
-R = \frac{aS}{1 + bS}
-\]
-
-Where:  
-- \( S \) is the spawning stock biomass.  
-- \( R \) is the predicted recruitment.  
-- \( a \) is the maximum recruitment capacity.  
-- \( b \) regulates density dependence (higher values of \( b \) result in a lower plateau).
-
-The blue line will show the asymptotic curve that describes the relationship between spawning stock biomass and recruitment.
-
-```{r}
-# Lista de modelos base
-models <- list(
-  s1.1 = base.model1.1,
-  s1.2 = base.model1.2,
-  s1.3 = base.model1.3,
-  s1.4 = base.model1.4
-)
-
-colors <- c("s1.1" = "#ca0020", "s1.2" = "#f4a582", "s1.3" = "#bababa", "s1.4" = "#404040")
-shapes <- c("s1.1" = 5, "s1.2" = 6, "s1.3" = 15, "s1.4" = 4)
-
-plot_data <- list()
-curve_data <- list()
-label_data <- list()
-params <- data.frame(Model = character(), a = numeric(), b = numeric(), stringsAsFactors = FALSE)
-for (group in names(models)) {
-  model <- models[[group]]
-  l <- model$recruit$SpawnBio
-  m <- model$recruit$pred_recr
-  yr <- model$timeseries$Yr[1:35]
-
-  # Filtrar valores válidos
-  valid <- which(is.finite(l) & is.finite(m) & l > 0 & m > 0)
-  l <- l[valid]
-  m <- m[valid]
-  yr <- yr[valid]
-
-  plot_data[[group]] <- data.frame(l = l, m = m, yr = yr, Group = group)
-  label_data[[group]] <- plot_data[[group]] %>% filter(yr %% 3 == 0)
-
-  fit <- NULL
-  tryCatch({
-    fit <- nls(m ~ (a * l) / (1 + b * l),
-               start = list(a = max(m), b = 1 / max(l)),
-               data = data.frame(l, m))
-
-    l_pred <- seq(min(l), max(l), length.out = 100)
-    m_pred <- predict(fit, newdata = data.frame(l = l_pred))
-    curve_data[[group]] <- data.frame(l = l_pred, m = m_pred, Group = group)
-
-    coefs <- coef(fit)
-    params <- rbind(params, data.frame(Model = group, a = coefs["a"], b = coefs["b"]))
-
-  }, error = function(e) {
-    message(paste("Model", group, ": Beverton-Holt fit failed ->", e$message))
-  })
-}
-
-# Unir data frames
-plot_data_all <- bind_rows(plot_data)
-curve_data_all <- bind_rows(curve_data)
-label_data_all <- bind_rows(label_data)
-
-# Graficar
-srr_sce <- ggplot() +
-  geom_point(data = plot_data_all, aes(x = l, y = m, shape = Group), size = 2) +
-  geom_line(data = curve_data_all, aes(x = l, y = m, color = Group), linewidth = 1.2) +
-  ggrepel::geom_text_repel(data = label_data_all, aes(x = l, y = m, label = yr, color = Group),
-                           size = 4, max.overlaps = 50) +
-  scale_shape_manual(values = shapes) +
-  scale_color_manual(values = colors) +
-  theme_bw() +
-  ylim(0, max(plot_data_all$m)) +
-  labs(x = "Spawning Stock Biomass (SSB)", y = "Predicted Recruitment", shape = "Model", color = "Model") +
-  theme(legend.position = "bottom")
-
-```
-
-### Productivity and Interannual Variability by Scenario
-
-Estimating the productivity of Antarctic krill is critical for understanding the species’ capacity to replenish its population in response to varying levels of spawning biomass.
-Productivity, defined as the ratio of recruitment to spawning stock biomass, provides a standardized measure of reproductive success and population resilience under different ecological and fishing pressures.
-Comparing productivity across scenarios—each representing different assumptions about environmental drivers, fishing mortality, or predator dynamics—enables a robust evaluation of how krill populations reflect this changes.
-
-For each scenario \( i \) and year \( t \), we computed the productivity as the ratio between recruitment and spawning stock biomass (SSB):
-
-$$
-\text{Productivity}_{i,t} = \frac{\text{Recruitment}_{i,t}}{\text{SSB}_{i,t}}
-$$
-
-We also calculated the **interannual percentage change** in recruitment and SSB as:
-
-$$
-\text{Change in Recruitment}_{i,t} = \left( \frac{\text{Recruitment}_{i,t} - \text{Recruitment}_{i,t-1}}{\text{Recruitment}_{i,t-1}} \right) \times 100
-$$
-
-$$
-\text{Change in SSB}_{i,t} = \left( \frac{\text{SSB}_{i,t} - \text{SSB}_{i,t-1}}{\text{SSB}_{i,t-1}} \right) \times 100
-$$
-
-These metrics allow us to analyze both the productivity and the temporal dynamics of the population under each scenario \( i \).
-
-```{r}
-# Simular 4 dataframes por escenario (usa tus datos reales)
-escenarios <- list(
-  s1.1 = base.model1.1$SPAWN_RECR_CURVE,
-  s1.2 = base.model1.2$SPAWN_RECR_CURVE,
-  s1.3 = base.model1.3$SPAWN_RECR_CURVE,
-  s1.4 = base.model1.4$SPAWN_RECR_CURVE
-)
-
-# Crea una tabla con productividad y cambios para cada escenario
-resultados <- purrr::imap_dfr(escenarios, function(df, nombre) {
-  df %>%
-    mutate(
-      Scenario = nombre,
-      Productividad = Recruitment / SSB,
-      Cambio_Recruit = c(NA, diff(Recruitment)) / lag(Recruitment) * 100,
-      Cambio_SSB = c(NA, diff(SSB)) / lag(SSB) * 100
-    )
-})
-
-prod <- ggplot(resultados, aes(x = `SSB/SSB_virgin`, y = Productividad, 
-                       color = Scenario)) +
-  geom_point(size = 1.1) +
-  labs(title = "",
-       y = "Recruitment / SSB",
-       x = "SSB / SSB_virgin") +
-  scale_color_manual(name= "",
-                    values = c("s1.1" = "#ca0020", 
-                               "s1.2" = "#f4a582", 
-                               "s1.3" = "#bababa", 
-                               "s1.4" = "#404040")) +
-  theme_bw() +
-  theme(legend.position = "bottom")
-```
-
-These two panels provide insight into the stock-recruitment dynamics of Antarctic krill under four assessment scenarios, each incorporating different levels of ecosystem complexity.
-
-**Left panel – Predicted Recruitment vs Spawning Stock Biomass (SSB):**  
-This plot compares the predicted recruitment across years with corresponding spawning biomass values for the four model scenarios (s1.1 to s1.4).
-Scenario s1.1 (red) shows a relatively flat and optimistic Beverton-Holt relationship, with high recruitment predicted even at low levels of SSB.
-The other scenarios, particularly s1.3 (gray) and s1.4 (black), show much more constrained recruitment estimates and a more saturating relationship, where recruitment increases only slightly with increasing SSB.
-This suggests that incorporating environmental (s1.3) and predator (s1.4) effects leads to a more conservative and ecologically realistic recruitment dynamic.
-Year labels illustrate how certain years (e.g., 2004, 2016, 2022) deviate between scenarios, especially at low biomass levels.
-
-**Right panel – Recruitment per SSB vs Relative SSB (SSB / SSBo):**  
-This plot displays recruitment efficiency (Recruitment/SSB) as a function of relative spawning biomass (SSB / SSB₀).
-In all models, recruitment efficiency is highest at very low biomass levels and declines as SSB increases, consistent with compensatory dynamics.
-However, s1.3 and s1.4 (which include environmental effects) show significantly lower overall productivity across all SSB levels compared to s1.1 and s1.2.
-This indicates that the inclusion of ecosystem drivers reduces the per-capita productivity of the stock, especially under depleted conditions.
-
-**Interpretation:**  
-Together, these plots indicate that ecosystem-informed models (s1.3 and s1.4) produce more constrained and precautionary estimates of krill productivity.
-The more optimistic recruitment predicted by s1.1 (and to a lesser extent s1.2) may overestimate stock resilience, particularly in low SSB conditions.
-These findings emphasize the importance of accounting for environmental variability and predator dynamics in stock assessment models to avoid overoptimistic management advice and better reflect the biological limits of the krill population.
-
-```{r, fig.height=4, fig.width=6, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios.
-Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
-Right panel: Recruitment per unit of SSB plotted against relative SSB (SSB/SSBo), illustrating differences in per-capita productivity across models ( s1.1 (no ecosystem variables), s1.2 (with predator effects), s1.3 (with environmental variables), and s1.4 (with both predator and environmental effects))"}
-ggarrange(srr_sce,
-          prod,
-          ncol = 2)
-```
-
+```{r eval=FALSE}
 ### Compara Fishing Mortality.
-
-```{r eval=FALSE}
 escenarios <- list(
@@
 These analyses assess the consistency and reliability of stock assessment models by systematically removing the most recent years of data and comparing the resulting estimates with the full dataset.
-
 In this study, we conducted a retrospective analysis to examine the sensitivity of our recruitment and spawning stock biomass (SSB) estimates to the inclusion or exclusion of recent data.
 By applying this approach to multiple models, we identified potential biases and evaluated the stability of the recruitment estimates over time.
-
 The retrospective patterns were assessed by calculating the relative error between the predictions of truncated datasets and the full dataset.
 These differences allowed us to detect trends in model performance, such as systematic overestimation or underestimation of key population parameters.
-Understanding these deviations is crucial for improving the robustness of the stock assessment models and ensuring more reliable projections for fisheries management.
 
@@
 
-```{r}
+Using `retro()` and `SSplotRetro()` functions, we obtain main results
+
+Retrospective analysis for spawning biomass (Figure \@ref(fig:retrossb))
+
+```{r retrossb, fig.cap="Retrospective analysis for spawning biomass by scenario in krill"}
 par(mfrow=c(2,2), mar=c(5,4,1,1)) 
@@
 
-```{r}
+Retrospective analysis for fishing mortality (Figure \@ref(fig:retrof))
+
+```{r retrof, fig.cap="Retrospective analysis for fishing mortality by scenario in krill"}
 par(mfrow=c(2,2), mar=c(5,4,1,1)) 
@@
             subplots = c("F"))
-```
-
-### Hindcast Cross-Validation and Prediction Skill
-
-The Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis is implemented using the model outputs generated by the `r4ss::SS_doRetro()` function.
-This diagnostic evaluates the predictive performance of the model by comparing hindcast predictions with observed data.
-
-To assess prediction skill, we employ the Mean Absolute Scaled Error (MASE) as a robust metric.
-MASE is calculated by scaling the mean absolute error of the model predictions relative to the mean absolute error of a naïve baseline prediction.
-Specifically, the MASE score is computed as follows:
-
-- A MASE score greater than 1 indicates that the model’s average forecasts are less accurate than a random walk model.
-- A MASE score equal to 1 suggests that the model’s accuracy is similar to that of a random walk.
-- A MASE score less than 1 indicates that the model performs better than a random walk.
-- A MASE score of 0.5, for example, indicates that the model’s forecasts are twice as accurate as the naïve baseline prediction, suggesting the model has predictive skill.
-
-This approach provides a rigorous evaluation of model forecasting capabilities and helps identify improvements for model calibration.
-
-```{r}
-par(mfrow=c(2,4), mar=c(5,4,2,1))
-hci1 = SSplotHCxval(retroSummary1.1, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-```
-
-```{r}
-par(mfrow=c(3,3), mar=c(5,4,2,1))
-hci2 = SSplotHCxval(retroSummary1.2, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-```
-
-```{r}
-par(mfrow=c(2,4), mar=c(5,4,2,1))
-hci3 = SSplotHCxval(retroSummary1.3, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7)
-```
-
-```{r}
-par(mfrow=c(3,3), mar=c(5,4,2,1))
-hci4 = SSplotHCxval(retroSummary1.4, 
-                   add = T, 
-                   verbose = F, 
-                   legendcex = 0.7) 
-```
-
-### Kobe (status)
-
-```{r eval=FALSE}
-mvln = SSdeltaMVLN(base.model1.1,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-mvln = SSdeltaMVLN(base.model1.2,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-mvln = SSdeltaMVLN(base.model1.3,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-mvln = SSdeltaMVLN(base.model1.4,
-                   Fref = "MSY", 
-                   plot = TRUE,
-                   addprj=TRUE,
-                   virgin = TRUE,
-                   mc=100)
-
-```
-another
-
-```{r eval=FALSE}
-nanios<-base.model1.1$Kobe$Yr
-SSBDep<-base.model1.1$Kobe$B.Bmsy
-y_BDPR<-base.model1.1$Kobe$F.Fmsy
-
-kobe1 <- ggplot(base.model1.1$Kobe)+
-  geom_rect(aes(xmin = 0, xmax = 0.5, ymin = 0, ymax = 5),
-            fill = "#E43338", alpha = 0.5) +
-  geom_rect(aes(xmin = 0.5, xmax = 0.75, ymin = 0, ymax = 5),
-            fill = "#F2ED23", alpha = 0.5) +
-  geom_rect(aes(xmin = 0.75, xmax = 1.25, ymin = 0, ymax = 5),
-            fill = "#ACC39A", alpha = 0.5) +
-  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 0, ymax = 1),
-            fill = "#608D68", alpha = 0.5) +
-  geom_rect(aes(xmin = 1.25, xmax = 9, ymin = 1, ymax = 5),
-            fill = "#808080", alpha = 0.5) +
-  geom_path(aes(x=base.model1.1$Kobe$B.Bmsy,y=base.model1.1$Kobe$F.Fmsy,
-                label=base.model1.1$Kobe$Yr))+
-  geom_point(aes(base.model1.1$Kobe$B.Bmsy, base.model1.1$Kobe$F.Fmsy),
-             lwd=2) +
-  geom_hline(yintercept = 1) +
-  geom_vline(xintercept = c(0.5, 0.75, 1.75, 1, 1.25), linetype=2)+
-  theme_few()+
-  labs(x = expression("BD/BD"[RMS]), y = expression("F/F"[RMS]))+
-  # geom_text(data = texto_coords, aes(x = x, y = y, label = etiqueta),
-  #            vjust = -0.5)+
-  geom_text(aes(x=base.model1.1$Kobe$B.Bmsy,y=base.model1.1$Kobe$F.Fmsy,label=base.model1.1$Kobe$Yr),
-            nudge_y = 0.1,size = 3,
-            check_overlap = TRUE)
-kobe1
 ```
@@
 ```
-Viz
-
-```{r, fig.height=4, fig.width=6}
+
+Figure \@ref(fig:bias) presents the retrospective bias (*Rho*) in Spawning Stock Biomass (SSB) and Fishing Mortality (F) across different peel years (2019–2016) and for the combined period under four model scenarios (s1.1 to s1.4).
+In the SSB panel, all models exhibit negative *Rho* values across the years, indicating a downward bias.
+The magnitude of the bias varies among models, with s1.3 and s1.4 showing the most pronounced deviations.
+In the F panel, *Rho* values are consistently positive, suggesting an upward bias.
+The combined *Rho* values for each metric confirm these trends, showing persistent differences among model scenarios.
+
+```{r bias, fig.cap= "Summary of retrospective analisis by scenario in F and SSB"}
 tablebias_combined <- tablebias_combined %>%
@@
 ```
-
-The plot presents the retrospective bias (*Rho*) in Spawning Stock Biomass (SSB) and Fishing Mortality (F) across different peel years (2019–2016) and for the combined period under four model scenarios (s1.1 to s1.4).
-In the SSB panel, all models exhibit negative *Rho* values across the years, indicating a downward bias.
-The magnitude of the bias varies among models, with s1.3 and s1.4 showing the most pronounced deviations.
-In the F panel, *Rho* values are consistently positive, suggesting an upward bias.
-The combined *Rho* values for each metric confirm these trends, showing persistent differences among model scenarios.
-
-(Discussion)
-
-The observed retrospective bias in SSB and F aligns with findings from previous studies on krill population dynamics, where recruitment variability and model structure influence stock assessment estimates (Hill et al., 2016; Watters et al., 2020).
-The negative bias in SSB suggests that models may overestimate past biomass, potentially due to unaccounted environmental variability or underestimation of natural mortality (Meyer et al., 2021).
-The positive bias in F indicates a tendency to underestimate past fishing mortality, which could be linked to the inclusion of predator-based indices such as penguin foraging data (Trathan et al., 2018).
-These results highlight the importance of incorporating ecosystem interactions into stock assessments to improve predictive accuracy and support sustainable management of Antarctic krill.
-
-```{r}
-kbl(tablebias_combined, booktabs = T, 
+See Table \@ref(tab:rho_parameters) for details.
+
+```{r rho_parameters}
+kbl(tablebias_combined, 
+    booktabs = TRUE, 
     format = "html", 
-    caption = "Rho parameter by model and quantity (SSB and F)") %>% 
+    caption = "Rho parameter by model and quantity (SSB and F)" ) %>%
   kable_styling(latex_options = "HOLD_position")
@@
 
-### Likelihood tables
-
-```{r}
-like1 <- base.model1.1$likelihoods_used
-like1$model <- rep("s1.1", nrow(like1))
-like1 <- rownames_to_column(like1, var = "Description")
-like2 <- base.model1.2$likelihoods_used
-like2$model <- rep("s1.2", nrow(like2))
-like2 <- rownames_to_column(like2, var = "Description")
-like3 <- base.model1.3$likelihoods_used
-like3$model <- rep("s1.3", nrow(like3))
-like3 <- rownames_to_column(like3, var = "Description")
-like4 <- base.model1.4$likelihoods_used
-like4$model <- rep("s1.4", nrow(like4))
-like4 <- rownames_to_column(like4, var = "Description")
-
-totalike <- rbind(like1,
-                   like2,
-                   like3,
-                   like4)
-```
-
-```{r}
-ggplot(totalike %>% 
-         filter(values>1))+
-  geom_bar( aes(y = reorder(Description, values), 
-                x =log(values)),
-            stat = "identity", position = "dodge") +
-  theme_few() +
-  facet_wrap(.~model, 
-             ncol=4)+
-  labs(y="",
-       x="Log Likelihood")+
-  theme(axis.text.x = element_text(angle = 90, 
-                                   hjust = 1))+
-  xlim(0,30)
-```
-
-### Likelihood Profile
+### Hindcast Cross-Validation and Prediction Skill
+
+The Hindcast Cross-Validation (HCxval) diagnostic in Stock Synthesis is implemented using the model outputs generated by the `r4ss::SS_doRetro()` and using `SSplotHCval()` function.
+This diagnostic evaluates the predictive performance of the model by comparing hindcast predictions with observed data.
+To assess prediction skill, we employ the Mean Absolute Scaled Error (MASE) as a robust metric.
+MASE is calculated by scaling the mean absolute error of the model predictions relative to the mean absolute error of a naïve baseline prediction.
+Specifically, the MASE score is computed as follows:
+
+- A MASE score greater than 1 indicates that the model’s average forecasts are less accurate than a random walk model.
+- A MASE score equal to 1 suggests that the model’s accuracy is similar to that of a random walk.
+- A MASE score less than 1 indicates that the model performs better than a random walk.
+- A MASE score of 0.5, for example, indicates that the model’s forecasts are twice as accurate as the naïve baseline prediction, suggesting the model has predictive skill.
+
+Hindcast validation for `s1.1` (Figure \@ref(fig:hcval1))
+
+```{r hcval1, fig.cap="Hindcast validation for s1.1 by fleet"}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
+hci1 = SSplotHCxval(retroSummary1.1, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+```
+
+Hindcast validation for `s1.2` (Figure \@ref(fig:hcval2))
+
+```{r hcval2, fig.cap="Hindcast validation for s1.2 by fleet"}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
+hci2 = SSplotHCxval(retroSummary1.2, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+```
+
+Hindcast validation for `s1.3` (Figure \@ref(fig:hcval3))
+
+```{r hcval3, fig.cap="Hindcast validation for s1.3 by fleet"}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
+hci3 = SSplotHCxval(retroSummary1.3, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7)
+```
+
+Hindcast validation for `s1.4` (Figure \@ref(fig:hcval4))
+
+```{r hcval4, fig.cap="Hindcast validation for s1.4 by fleet"}
+par(mfrow=c(3,3), mar=c(5,4,2,1))
+hci4 = SSplotHCxval(retroSummary1.4, 
+                   add = T, 
+                   verbose = F, 
+                   legendcex = 0.7) 
+```
+
+```{r eval=FALSE}
+### Kobe (status) dont run
+mvln = SSdeltaMVLN(base.model1.1,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+mvln = SSdeltaMVLN(base.model1.2,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+mvln = SSdeltaMVLN(base.model1.3,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+mvln = SSdeltaMVLN(base.model1.4,
+                   Fref = "MSY", 
+                   plot = TRUE,
+                   addprj=TRUE,
+                   virgin = TRUE,
+                   mc=100)
+
+```
+
+<!-- ### Likelihood Profile  -->
 
@@
 # 1. Identificar el directorio donde se encuentra el modelo base ----
-dirname.model.run <- here("s1.2")
+dirname.model.run <- here("s1.1")
 # 2. Crear un nuevo directorio para el "Perfil_Verosimilitud"  
-dirname.R0.profile <- here("s1.2",
+dirname.R0.profile <- here("s1.1",
                            "Perfil_Verosimilitud")
@@
                                  'Recruitment likelihood') 
-
-```
-
-### Likelihood Profile
+```
 
@@
 
-```{r eval=FALSE, echo = FALSE}
+```{r eval=FALSE}
 # Comparación de series de tiempo 
@@
 
-piner Plot
-
 ```{r eval = FALSE}
+#piner Plot
 #### R0_profile_plot_Length_like ----
@@
 
-### Table Diagnosis
+### Likelihood tables
+
+```{r eval=FALSE}
+like1 <- base.model1.1$likelihoods_used
+like1$model <- rep("s1.1", nrow(like1))
+like1 <- rownames_to_column(like1, var = "Description")
+like2 <- base.model1.2$likelihoods_used
+like2$model <- rep("s1.2", nrow(like2))
+like2 <- rownames_to_column(like2, var = "Description")
+like3 <- base.model1.3$likelihoods_used
+like3$model <- rep("s1.3", nrow(like3))
+like3 <- rownames_to_column(like3, var = "Description")
+like4 <- base.model1.4$likelihoods_used
+like4$model <- rep("s1.4", nrow(like4))
+like4 <- rownames_to_column(like4, var = "Description")
+
+totalike <- rbind(like1,
+                   like2,
+                   like3,
+                   like4)
+
+ggplot(totalike %>% 
+         filter(values>1))+
+  geom_bar( aes(y = reorder(Description, values), 
+                x =log(values)),
+            stat = "identity", position = "dodge") +
+  theme_few() +
+  facet_wrap(.~model, 
+             ncol=4)+
+  labs(y="",
+       x="Log Likelihood")+
+  theme(axis.text.x = element_text(angle = 90, 
+                                   hjust = 1))+
+  xlim(0,30)
+```
 
@@
 diag_tidy[, 2:5] <- lapply(diag_tidy[, 2:5], function(x) as.numeric(format(round(x, 3), nsmall = 3)))
-
-#write_csv(diag_tidy, "model_diagnosis.csv")
-flextable(diag_tidy) |>
-  colformat_num(digits = 3) |>  
-  autofit() |>
-  theme_booktabs() |>
-  align(align = "center", part = "all") |>
-  bold(part = "header")
-```
-
-### Otra tab
-
-```{r eval=FALSE}
-cpue_resid <- bind_rows(
-  base.model1.1$cpue %>% mutate(model = "s1.1"),
-  base.model1.2$cpue %>% mutate(model = "s1.2"),
-  base.model1.3$cpue %>% mutate(model = "s1.3"),
-  base.model1.4$cpue %>% mutate(model = "s1.4")
-)
-
-indres <- ggplot(cpue_resid, aes(x = Exp, y = Obs, color = model)) +
-  geom_point() +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
-  facet_wrap(~Fleet_name, scales = "free") +
-  theme_minimal() +
-  labs(title = "Observed vs Expected (Index)",
-       x = "Expected",
-       y = "Observed")
-
-len_resid <- bind_rows(
-  base.model1.1$lendbase %>% mutate(model = "s1.1"),
-  base.model1.2$lendbase %>% mutate(model = "s1.2"),
-  base.model1.3$lendbase %>% mutate(model = "s1.3"),
-  base.model1.4$lendbase %>% mutate(model = "s1.4")
-)
-
-lenres <- ggplot(len_resid, aes(x = Exp, y = Obs, color = model)) +
-  geom_point() +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
-  facet_wrap(~Fleet, scales = "free") +
-  theme_minimal() +
-  labs(title = "Observed vs Expected (Index)",
-       x = "Expected",
-       y = "Observed")
-
-ggarrange(indres,
-          lenres,
-          ncol=1,
-          common.legend = T)
-```
-```{r eval=FALSE}
-
-cpue_df <- cpue_resid %>%
-  transmute(
-    Yr = Yr,
-    Obs = Obs,
-    Exp = Exp,
-    model = model,
-    Fleet_name = as.character(Fleet_name),  # Asegura que es character
-    type = "Index"
-  )
-
-len_df <- len_resid %>%
-  transmute(
-    Yr = Yr,
-    Obs = Obs,
-    Exp = Exp,
-    model = model,
-    Fleet_name = as.character(Fleet),  # Lo convertimos para que combine bien
-    type = "Length"
-  )
-
-resid_all <- resid_all %>%
-  mutate(Fleet_name = as.character(Fleet_name),
-         Fleet_name = case_when(
-           Fleet_name == "1"  ~ "FISHERYBS",
-           Fleet_name == "2"  ~ "FISHERYEI",
-           Fleet_name == "3"  ~ "FISHERYGS",
-           Fleet_name == "4"  ~ "FISHERYJOIN",
-           Fleet_name == "5"  ~ "FISHERYSSIW",
-           Fleet_name == "6"  ~ "SURVEYBS",
-           Fleet_name == "7"  ~ "SURVEYEI",
-           Fleet_name == "8"  ~ "SURVEYGS",
-           Fleet_name == "9"  ~ "SURVEYJOIN",
-           Fleet_name == "10" ~ "SURVEYSSIW",
-           Fleet_name == "11" ~ "PREDATOR",
-           TRUE ~ Fleet_name  # mantiene los nombres ya correctos
-         ))
-
-ggplot(resid_all, aes(x = Exp, y = Obs, color = model)) +
-  geom_point(alpha = 0.6) +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray40") +
-  facet_grid(type ~ Fleet_name, scales = "free") +
-  theme_minimal(base_size = 13) +
-  labs(title = "Observed vs Expected residuals",
-       x = "Expected", y = "Observed")
-
-```
-```{r}
-ggplot(resid_all, aes(x = Obs - Exp, fill = model)) +
-  geom_density(alpha = 0.5) +
-  facet_wrap(Fleet_name~type, scales = "free") +
-  labs(title = "Distribución de residuales", x = "Residual (Obs - Exp)", y = "Densidad") +
-  theme_minimal()
-
-```
-### Statistics
-
-```{r eval=FALSE, resumen_residuales, message=FALSE, warning=FALSE}
-resid_all <- resid_all %>%
-  mutate(residual = Obs - Exp)
-
-# Estadísticas básicas
-resid_stats <- resid_all %>%
-  group_by(type, model) %>%
-  summarise(
-    N = n(),
-    Mean = mean(residual, na.rm = TRUE),
-    SD = sd(residual, na.rm = TRUE),
-    .groups = "drop"
-  )
-
-# Shapiro-Wilk (normalidad)
-shapiro_p <- resid_all %>%
-  group_by(type, model) %>%
-  summarise(
-    shapiro_p = ifelse(n() > 3, shapiro.test(residual)$p.value, NA),
-    .groups = "drop"
-  )
-
-# Ljung-Box (autocorrelación temporal)
-ljung_p <- resid_all %>%
-  group_by(type, model) %>%
-  summarise(
-    ljung_p = ifelse(n() > 10, Box.test(residual, lag = 10, type = "Ljung-Box")$p.value, NA),
-    .groups = "drop"
-  )
-
-# Breusch-Pagan (heterocedasticidad)
-bp_p <- resid_all %>%
-  group_by(type, model) %>%
-  summarise(
-    bp_p = ifelse(n() > 3, bptest(lm(residual ~ Yr, data = cur_data()))$p.value, NA),
-    .groups = "drop"
-  )
-
-# Unión de todo
-tabla_resumen <- resid_stats %>%
-  left_join(shapiro_p, by = c("type", "model")) %>%
-  left_join(ljung_p, by = c("type", "model")) %>%
-  left_join(bp_p, by = c("type", "model")) %>%
-  mutate(across(where(is.numeric), ~ round(.x, 3)))
-
-# Mostrar tabla
-flextable(tabla_resumen) %>%
-  set_caption("Tabla 1. Estadísticos y pruebas sobre residuales por tipo y modelo") %>%
-  autofit()
-
-```
-
-## Outputs
-
-### Outputs Model `s1.1`
-
-```{r}
-outps1 <- base.model1.1$timeseries[1:37, 2:8]
-
-# addWorksheet(wb, "variable_s2")
-# writeData(wb, "variable_s2", outps1)
-# 
-# # Guardar el workbook
-# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
-# 
-# 
-# out_s2 <- read_excel("DataKrill.xlsx", 
-#     sheet = "variable_s2")
-outps1 %>%
-  kbl(booktabs = TRUE,
-      format = "html",
-    caption = "Main variables outputs from stock asssessment krill in WAP `s1.1`") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
+```
+
+See Table \@ref(tab:likecom) for details.
+
+```{r likecom}
+diag_tidy %>%
+  kbl(digits = 3, 
+      caption = "Model Diagnosis Results",
+      align = "c",  # Centrar todas las columnas
+      booktabs = TRUE) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
                 full_width = FALSE,
-                font_size=9,
-                html_font = "arial")#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-### Outputs Model `s1.2`
-
-```{r}
-outps2 <- base.model1.2$timeseries[1:37, 2:8]
-
-# addWorksheet(wb, "variable_s2")
-# writeData(wb, "variable_s2", outps1)
-# 
-# # Guardar el workbook
-# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
-# 
-# 
-# out_s2 <- read_excel("DataKrill.xlsx", 
-#     sheet = "variable_s2")
-outps2 %>%
-  kbl(booktabs = TRUE,
-      format = "html",
-    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.2`") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = FALSE,
-                font_size=9,
-                html_font = "arial")#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-
-### Outputs Model `s1.3`
-
-```{r}
-outps3 <- base.model1.3$timeseries[1:37, 2:8]
-
-# addWorksheet(wb, "variable_s2")
-# writeData(wb, "variable_s2", outps1)
-# 
-# # Guardar el workbook
-# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
-# 
-# 
-# out_s2 <- read_excel("DataKrill.xlsx", 
-#     sheet = "variable_s2")
-outps3 %>%
-  kbl(booktabs = TRUE,
-      format = "html",
-    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.3`") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = FALSE,
-                font_size=9,
-                html_font = "arial")#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-
-### Outputs Model `s1.4`
-
-```{r}
-outps4 <- base.model1.3$timeseries[1:37, 2:8]
-
-# addWorksheet(wb, "variable_s2")
-# writeData(wb, "variable_s2", outps1)
-# 
-# # Guardar el workbook
-# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
-# 
-# 
-# out_s2 <- read_excel("DataKrill.xlsx", 
-#     sheet = "variable_s2")
-outps4 %>%
-  kbl(booktabs = TRUE,
-      format = "html",
-    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.4`") %>%
-  kable_paper("hover", 
-              full_width = TRUE)%>%
-  kable_styling(latex_options = c("striped",
-                                  "condensed"),
-                full_width = FALSE,
-                font_size=9,
-                html_font = "arial")#%>% 
-  #pack_rows(index = c("Estimation" = 1,
-   #                     "Prediction" = 45))
-```
-
-```{r}
-# Unir los data frames en uno solo
-outpsall <- rbind(outps1,  
-                  outps2,
-                  outps3,
-                outps4)
-```
-
-```{r}
-# Crear una columna de modelo en cada data frame
-outps1$Model <- "s1.1"
-outps2$Model <- "s1.2"
-outps3$Model <- "s1.3"
-outps4$Model <- "s1.4"
-
-# Unir los data frames en uno solo
-outpsall <- rbind(outps1,  
-                  outps2,
-                  outps3,
-                  outps4)
-
-outpsall_long <- outpsall %>%
-  pivot_longer(cols = c(Bio_all, Bio_smry, SpawnBio, Recruit_0), 
-               names_to = "Variable", values_to = "Value") %>% 
-  filter(Yr >= 1989, Yr <= 2020)
-
-# Gráfico de líneas por variable y modelo
-alloutput <- ggplot(outpsall_long, aes(x = Yr, y = Value, color = Model, group = Model)) +
-  geom_point() +
-  #geom_smooth(method = "lm")+
-  facet_wrap(~Variable, 
-             ncol = 4, 
-             scales = "free_y") + 
-  scale_color_manual(name = "Scenario",
-                     values = c("s1.1" = "#ca0020", 
-                                "s1.2" = "#f4a582", 
-                                "s1.3" = "#bababa", 
-                                "s1.4" = "#404040")) +
-  labs(title = "", x = "", y = "") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1),
-        legend.position = "bottom")
-
-alloutput
-
-```
-
-## Comparison
-
-### Comparison outputs betwwen scenarios
-
-```{r}
-
+                position = "center") %>%
+  row_spec(0, bold = TRUE)  # Encabezados en negrita
+```
+
+As shown in Table \@ref(tab:parameter-comparison), the models differ substantially in key parameter estimates and likelihood contributions.
+
+```{r parameter-comparison, warning=FALSE, message=FALSE}
 #read in all model runs
@@
 summaryoutputall <- SSsummarize(biglist)
-```
-
-```{r}
+
 legend.labels <- c('s1.1','s1.2','s1.3', 's1.4')
@@
 
-tablelike <- SStableComparisons(summaryoutput,
-                   likenames = c("TOTAL",
-                                 "Survey",
-                                 "Length_comp",
-                                 "Age_comp",
-                                 "priors",
-                                 "Size_at_age"), 
-                   names = c("Recr_Virgin",
-                             "R0",
-                             "SSB_Virg", 
-                             "Bratio_2020",
-                             "SPRratio_2020"),
+# Crear tabla con SStableComparisons
+tablelike <- SStableComparisons(
+  summaryoutput,
+  likenames = c("TOTAL", "Survey", "Length_comp", "Age_comp", "priors", "Size_at_age"), 
+  names = c("Recr_Virgin", "R0", "SSB_Virg", "Bratio_2020", "SPRratio_2020"),
   digits = NULL,
   modelnames = legend.labels,
-                   csv=TRUE,
+  csv = TRUE,
   csvdir = "/Users/mauriciomardones/DOCAS/SA_Krill",
-                   csvfile = "parameter_comparison_table.csv", verbose = TRUE,
-                   mcmc = FALSE)
-kbl(tablelike, booktabs = T,
-    format = "html",
-    caption = "")  %>% 
+  csvfile = "parameter_comparison_table.csv",
+  verbose = TRUE,
+  mcmc = FALSE
+)
+
+# Limpiar la tabla: eliminar columnas o filas con solo NAs y evitar notación científica
+table_clean <- tablelike %>%
+  mutate(across(where(is.numeric), ~ format(.x, scientific = FALSE, digits = 4))) %>%
+  select(where(~ !all(is.na(.))))  # elimina columnas que son completamente NA
+
+# Renderizar con kable
+kbl(table_clean, booktabs = TRUE, format = "html", 
+    caption = "Model parameter and likelihood comparison") %>%
   kable_styling(latex_options = "scale_down")
-```
-
-```{r eval=FALSE}
-#read in all model runs
-#note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist1_4 <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(
-                                  dir1.1,
-                                  dir1.2,
-                                  dir1.3,
-                                  dir1.4),
-                       getcovar = F)
-
-#create summary of model runs from list above
-summaryoutput1_4 <- SSsummarize(biglist1_4)
-
-SSplotComparisons(summaryoutput1_4,
-                  legendlabels = c("s1.1 (Ref Model)",
-                 "s1.2",
-                 "s1.3",
-                 "s1.4"),
-                 filenameprefix = "COM1",
-                 labels = c("Year", 
-                            "Spawning biomass (t)",
-                            "Relative spawning biomass", 
-                            "Age-0 recruits (1,000s)",
-                            "Recruitment deviations", 
-                            "Index", "Log index", 
-                            "1 - SPR", 
-                            "Density",
-                            "Management target", 
-                            "Minimum stock size threshold",
-                            "Spawning output",
-                            "Harvest rate"),
-                 png=TRUE,
-                 plotdir=Figs)
-```
-
-Comparison between select models `Ref Model: No Env-Predator` and `S1.1 w/ Env and Predator data`
-
-```{r eval=FALSE}
-#read in all model runs
-#note if cover=T you need a hessian; if covar=F you do not need a hessian
-biglist14 <- SSgetoutput(keyvec = NULL,
-                       dirvec = c(
-                                  dir1.1,
-                                  dir1.4),
-                       getcovar = F)
-
-#create summary of model runs from list above
-summaryoutput14 <- SSsummarize(biglist14)
-
-comp14 <- SSplotComparisons(summaryoutput14,
-                  legendlabels = c("Ref Model: No Env-Predator",
-                 "S1.1 w/ Env and Predator data"),
-                  filenameprefix = "COM2",
-                 labels = c("Year", 
-                            "Spawning biomass (t)",
-                            "Relative spawning biomass", 
-                            "Age-0 recruits (1,000s)",
-                            "Recruitment deviations", 
-                            "Index", "Log index", 
-                            "1 - SPR", 
-                            "Density",
-                            "Management target", 
-                            "Minimum stock size threshold",
-                            "Spawning output",
-                            "Harvest rate"),
-                 png=TRUE,plotdir=Figs)
-```
-
-Antoher comparision in `R0` (Natural log of virgin recruitment level)
-
-```{r}
-r0_vals <- data.frame(
-  model = c("s1.1", "s1.2", "s1.3", "s1.4"),
-  value = c(base.model1.1$jitter_info[1,1],
-            base.model1.2$jitter_info[2,1],
-            base.model1.3$jitter_info[1,1],
-            base.model1.4$jitter_info[2,1]),
-  cv = c(base.model1.1$jitter_info[1,6],
-         base.model1.2$jitter_info[2,6],
-         base.model1.3$jitter_info[1,6],
-         base.model1.4$jitter_info[2,6])
-)
-
-x_vals <- seq(10, 30, length.out = 1000)
-
-dens_data <- r0_vals %>%
-  rowwise() %>%
-  mutate(sd =  cv,  
-         density = list(data.frame(
-           x = x_vals,
-           y = dnorm(x_vals, mean = value, sd = sd),
-           model = model
-         ))) %>%
-  select(density) %>%
-  unnest(cols = c(density))
-
-colores <- c("s1.1" = "#ca0020", 
-             "s1.2" = "#f4a582", 
-             "s1.3" = "#bababa", 
-             "s1.4" = "#404040")
-
-r0plot <- ggplot(dens_data, aes(x = x, y = y, color = model, fill = model)) +
-  geom_area(alpha = 0.3, position = "identity", size = 0) +  # More transparent areas without borders
-  geom_line(size = 1.2) +  # Thicker lines for better visibility
-  scale_color_manual(values = colores, name = NULL) +  # Remove legend title for color
-  scale_fill_manual(values = colores, name = NULL) +   # Remove legend title for fill
-  labs(x = "ln(R0)", 
-       y = "Density") +
-  theme_bw(base_size = 14) +
-  theme(
-    legend.position = "bottom",  # Legend at bottom
-    legend.spacing.x = unit(0.5, 'cm'),  # Space between legend items
-    panel.grid.major = element_line(color = "gray90"),
-    panel.grid.minor = element_blank(),
-    axis.title = element_text(size = 12),
-    legend.text = element_text(size = 11)
-  ) +
-  xlim(16, 22) +
-  guides(color = guide_legend(nrow = 1),  # Single row legend
-         fill = guide_legend(nrow = 1))
-
-r0plot
-```
-
-### Comparsion in sd long term time series forecasting
-
-```{r fig.height=5, fig.width=8}
-
-errt <- data.frame(
-  Bio_all_1 = base.model1.1$timeseries$Bio_all,
-  Bio_all_2 = base.model1.2$timeseries$Bio_all,
-  Bio_all_3 = base.model1.3$timeseries$Bio_all,
-  Bio_all_4 = base.model1.4$timeseries$Bio_all,
-
-  Bio_smry_1 = base.model1.1$timeseries$Bio_smry,
-  Bio_smry_2 = base.model1.2$timeseries$Bio_smry,
-  Bio_smry_3 = base.model1.3$timeseries$Bio_smry,
-  Bio_smry_4 = base.model1.4$timeseries$Bio_smry,
-
-  SpawnBio_1 = base.model1.1$timeseries$SpawnBio,
-  SpawnBio_2 = base.model1.2$timeseries$SpawnBio,
-  SpawnBio_3 = base.model1.3$timeseries$SpawnBio,
-  SpawnBio_4 = base.model1.4$timeseries$SpawnBio,
-
-  Recruit_0_1 = base.model1.1$timeseries$Recruit_0,
-  Recruit_0_2 = base.model1.2$timeseries$Recruit_0,
-  Recruit_0_3 = base.model1.3$timeseries$Recruit_0,
-  Recruit_0_4 = base.model1.4$timeseries$Recruit_0
-)
-
-err_long <- errt %>%
-  pivot_longer(cols = everything(), 
-               names_to = "Variable_Model",
-               values_to = "Value")
-
-err_long <- err_long %>%
-  extract(Variable_Model, into = c("Variable", "Model"), 
-          regex = "(.+?)_(\\d+)$")
-
-err_long <- err_long %>%
-  mutate(Model = case_when(
-    Model == "1" ~ "Ref Model: No Env-Predator",
-    Model == "2" ~ "S1.1 w/ Predator data",
-    Model == "3" ~ "S1.1 w/ Env data",
-    Model == "4" ~ "S1.1 w/ Env and Predator data",
-    TRUE ~ Model
-  ))
-
-comsum <- ggplot(err_long, aes(x = Model, y = Value, fill = Model)) +
-  geom_boxplot(width = 0.5, outlier.shape = NA,
-               outliers = F) +
-  #geom_jitter(width = 0.2, alpha = 0.1) +
-  facet_wrap(~Variable, ncol=4, 
-             scales = "free_y") +
-  scale_fill_manual(name = "Scenario",
-                    values = c("#ca0020", "#f4a582", "#bababa", "#404040")) +
-  labs(title = "",
-       x = "",
-       y = "") +
-  theme_few() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1),
-        legend.position = "none")
-ggarrange(
-          alloutput,
-          comsum,
-          ncol = 1)
-```
-
-Statistical diferences
-
-```{r}
-# Extraer los datos
-spawn_bio_1 <- base.model1.1$timeseries$SpawnBio
-spawn_bio_4 <- base.model1.4$timeseries$SpawnBio
-
-# Prueba de normalidad (Shapiro-Wilk)
-shapiro.test(spawn_bio_1)
-shapiro.test(spawn_bio_4)
-
-# Prueba de igualdad de varianzas (Levene)
-library(car)
-leveneTest(c(spawn_bio_1, spawn_bio_4), 
-           group = rep(c("Model 1.1", "Model 1.4"), each = length(spawn_bio_1)))
-
-# Prueba t de Student (si los datos son normales)
-t.test(spawn_bio_1, spawn_bio_4, var.equal = TRUE)
-
-# Prueba de Wilcoxon (si los datos no son normales)
-wilcox.test(spawn_bio_1, spawn_bio_4)
-
-```
-
-**Two-Sample t-Test**
-
-- **t-value:** 4.2682  
-- **Degrees of Freedom (df):** 72  
-- **p-value:** 5.915e-05 (very small)  
-- **95% Confidence Interval for the Mean Difference:** [1,387,011; 3,818,074]  
-- **Sample Means:**  
-  - Model 1.1: **4,600,304**  
-  - Model 1.4: **1,997,762**
-
-Since the **p-value (5.915e-05) is much smaller than 0.05**, we reject the null hypothesis.
-This means there is **a statistically significant difference** between the spawning biomass predicted by Model 1.1 and Model 1.4.
-The confidence interval suggests that Model 1.1 tends to estimate a higher spawning biomass than Model 1.4, with an estimated difference between **1.39 and 3.82 million tons**.
-
-**Wilcoxon Rank Sum Test (Mann-Whitney U Test)**
-
-- **W-statistic:** 1256  
-- **p-value:** 6.698e-10 (extremely small)  
-- **Warning:** "Cannot compute exact p-value with ties" (indicates that some values are repeated)
-
-The **Wilcoxon test also rejects the null hypothesis** with an extremely small **p-value (6.698e-10)**.
-This confirms that the distribution of spawning biomass values is significantly different between the two models.
-Since the Wilcoxon test does not assume normality, it supports the conclusion from the **t-test**, reinforcing that
-
-Both tests strongly indicate that **Model 1.1 and Model 1.4 produce significantly different estimates of spawning biomass**.
-Specifically, **Model 1.1 predicts higher values**.
-This suggests that the factors or assumptions included in Model 1.4 result in **lower spawning biomass estimates**, which could have important implications for stock assessment and fisheries management.
-
-### Autocorrelation in Recruit
-
-To evaluate the temporal correlation structure of krill recruitment under different model configurations, we performed an Autocorrelation Function (ACF) analysis.
-The ACF measures the correlation between observations at different time lags, helping to assess whether recruitment estimates exhibit persistence or randomness across time.
-
-The analysis was conducted on recruitment estimates derived from four model scenarios:
-
-A reference model without environmental or predator influences (Ref Model: No Env-Predator).
-A model incorporating predator data (S1.1 w/ Predator data).
-A model incorporating environmental data (S1.1 w/ Env data).
-A model incorporating both environmental and predator data (S1.1 w/ Env and Predator data).
-Each model's residuals were extracted, and the autocorrelation function (ACF) was computed for a time lag range of up to 15 years.
-The dashed blue lines in the plots represent the 95% confidence intervals, indicating the threshold beyond which correlation values are statistically significant.
-If autocorrelation values remain within this range, it suggests that the recruitment estimates behave as a random process with no significant dependence on past values.
-Conversely, autocorrelation values exceeding these bounds indicate recruitment persistence or cyclic patterns.
-
-```{r}
-recs1 <- base.model1.1$recruit$pred_recr
-recs2 <- base.model1.2$recruit$pred_recr
-recs3 <- base.model1.3$recruit$pred_recr
-recs4 <- base.model1.4$recruit$pred_recr
-
-p1 <- ggAcf(base.model1.1$recruit$pred_recr) +
-  ggtitle("s1.1") +
-  theme_few()
-p2 <- ggAcf(base.model1.2$recruit$pred_recr) +
-  ggtitle("s1.2") +
-  theme_few()
-p3 <- ggAcf(base.model1.3$recruit$pred_recr) +
-  ggtitle("s1.3") +
-  theme_few()
-p4 <- ggAcf(base.model1.4$recruit$pred_recr) +
-  ggtitle("s1.4") +
-  theme_few()
-
-ggarrange(p1, p2,p3,p4, ncol=4, nrow = 1, common.legend = T)
-
-```
-
-The ACF plots indicate that the reference model (without environmental or predator data) exhibits weak but noticeable positive autocorrelation at certain lags, suggesting some degree of recruitment dependence over time.
-However, this autocorrelation does not appear strong or systematic.
-
-The model incorporating only predator data shows a slight reduction in autocorrelation magnitude, suggesting that predator-driven recruitment variability may have captured part of the temporal structure in the data.
-
-The model with only environmental data exhibits a further reduction in autocorrelation, implying that environmental variability explains a larger portion of recruitment trends than predator effects alone.
-
-Finally, the model that includes both environmental and predator data presents the lowest autocorrelation values, with nearly all bars remaining within the confidence bounds.
-This suggests that incorporating both factors provides the most effective explanation of recruitment fluctuations, reducing unexplained temporal structure.
-
-Overall, these results indicate that recruitment variability is at least partially driven by environmental and predator influences, and models integrating these factors provide more robust and independent recruitment estimates, minimizing systematic dependencies over time.
-
-### Recruit deviation
-
-```{r, warning=FALSE}
-dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
-  mutate(Serie = "s1.1")
-dev2 <- base.model1.2$recruit[8:35,c(1,7)] %>% 
-  mutate(Serie = "s1.2")
-dev3 <- base.model1.3$recruit[8:35,c(1,7)] %>% 
-  mutate(Serie = "s1.3")
-dev4 <- base.model1.4$recruit[8:35,c(1,7)] %>% 
-  mutate(Serie = "s1.4")
-
-data_list <- list(dev1, 
-                  dev2,
-                  dev3,
-                  dev4)
-titles <- c("Ref Model: No Env-Predator",
-                 "S1.1 w/ Predator data",
-                 "S1.1 w/ Env data",
-                 "S1.1 w/ Env and Predator data")
-# Lista para almacenar los gráficos
-plots <- list()
-
-for (i in 1:length(data_list)) {
-  p <- ggplot(data_list[[i]], aes(x = Yr, y = dev)) +
-    geom_point() +
-    stat_smooth(method = "loess", 
-                span = 0.15, 
-                se = T) +
-    geom_hline(yintercept = 0,
-               col="red",
-               linetype = "dotdash")+
-    ggtitle(titles[i]) +
-    theme(axis.text.x = element_text(angle = 90,
-                                   hjust = 1))+
-    ylim(-2,2)+
-    theme_few()
-  plots[[i]] <- p
-}
-
-# Organizar los gráficos con ggarrange
-combined_plot <- ggarrange(plotlist = plots, ncol = 2, nrow = 2)
-
-combined_plot
-```
-
-```{r}
+
+```
+
+Total likelihood in Figure \@ref(fig:likecompo2)
+
+```{r likecompo2, fig.width=7, fig.height=3, fig.cap="total likelihood composition by scenario"}
 df1 <- base.model1.1$likelihoods_used %>%
@@
 ```
+
 This bar plot presents the log-normal likelihood contributions of different model components in an Antarctic krill stock assessment under four scenarios (s1.1, s1.2, s1.3, and s1.4).
@@
 
-### Explotation Rate
+### Statistics analisys differences bewteen models
+
+To evaluate the residual behavior across model scenarios, we computed residuals as the difference between observed and expected values (`residual = Obs - Exp`).
+Basic statistics, including sample size (`n()`), mean (`mean()`), and standard deviation (`sd()`), were calculated for each combination of model and type.
+To test the normality of residuals, we applied the Shapiro-Wilk test (`shapiro.test()`) [@shapiro1965analysis], which is appropriate for small to moderate sample sizes.
+Temporal autocorrelation was assessed using the Ljung-Box test (`Box.test()` [@ljung1978measure] with `type = "Ljung-Box"` and `lag = 10`), evaluating the null hypothesis of independence across lags.
+To detect heteroscedasticity, we used the Breusch-Pagan test (`bptest()` from the `lmtest` package) [@breusch1979simple], fitting a linear model of residuals against year (`residual ~ Yr`) and testing for non-constant variance in the residuals.
+These diagnostics provide insight into the validity of model assumptions across different scenarios.
+
+```{r message=FALSE, warning=FALSE}
+resid_all <- resid_all %>%
+  mutate(residual = Obs - Exp)
+
+# Estadísticas básicas
+resid_stats <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    N = n(),
+    Mean = mean(residual, na.rm = TRUE),
+    SD = sd(residual, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Shapiro-Wilk (normalidad)
+shapiro_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    shapiro_p = ifelse(n() > 3, shapiro.test(residual)$p.value, NA),
+    .groups = "drop"
+  )
+
+# Ljung-Box (autocorrelación temporal)
+ljung_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    ljung_p = ifelse(n() > 10, Box.test(residual, lag = 10, type = "Ljung-Box")$p.value, NA),
+    .groups = "drop"
+  )
+
+# Breusch-Pagan (heterocedasticidad)
+bp_p <- resid_all %>%
+  group_by(type, model) %>%
+  summarise(
+    bp_p = ifelse(n() > 3, bptest(lm(residual ~ Yr, data = cur_data()))$p.value, NA),
+    .groups = "drop"
+  )
+
+```
+
+As shown in Table \@ref(tab:residual-summary), the residuals exhibit different statistical properties across model scenarios.
+
+```{r residual-summary, message=FALSE, warning=FALSE}
+tabla_resumen <- resid_stats %>%
+  left_join(shapiro_p, by = c("type", "model")) %>%
+  left_join(ljung_p, by = c("type", "model")) %>%
+  left_join(bp_p, by = c("type", "model")) %>%
+  mutate(across(where(is.numeric), ~ round(.x, 5)))
+
+tabla_resumen %>%
+  kbl(caption = "Summary statistics and residual diagnostic tests by type and model",
+      digits = 5,  # Redondeo a 5 decimales
+      align = "c",  # Centrado
+      booktabs = TRUE) %>%
+  kable_styling(bootstrap_options = c("striped", "hover"),
+                full_width = FALSE,
+                position = "center",
+                latex_options = "HOLD_position") %>%
+  row_spec(0, bold = TRUE) %>%  # Encabezados en negrita
+  column_spec(1, bold = TRUE)   # Primera columna en negrita (opcional)
+```
+
+The residual analysis shows distinct patterns between the Index and Length-based models across all four scenarios (`s1.1` to `s1.4`).
+For Index models, the mean residuals are notably large (ranging from approximately 105,610 to 143,409), with high standard deviations exceeding 1 million units, indicating substantial variability.
+The Shapiro-Wilk test consistently returned p-values of 0, strongly rejecting the null hypothesis of normality.
+The Ljung-Box test yielded very low p-values (≤ 0.00080), suggesting significant temporal autocorrelation in the residuals.
+In terms of heteroscedasticity, the Breusch-Pagan test showed moderate evidence of non-constant variance, with p-values ranging from 0.27150 (s1.1) to 0.01052 (s1.3), the latter indicating strong evidence of heteroscedasticity.
+
+In contrast, Length models present mean residuals near zero with low standard deviations (approximately 0.038–0.047), suggesting good overall model fit and low dispersion.
+However, despite the small residual magnitude, the Shapiro-Wilk and Ljung-Box tests consistently return p-values of 0, indicating violation of normality and presence of temporal autocorrelation.
+Furthermore, the Breusch-Pagan test returned extremely low p-values (≤ 0.00067), indicating strong evidence of heteroscedasticity across all Length-based scenarios.
+
+These results suggest that while Length models achieve better central tendencies (mean near zero and low variance), they still suffer from violations of key assumptions—most notably autocorrelation and heteroscedasticity.
+Index models, on the other hand, display large-scale variability and systematic residual patterns that may require model refinement or transformation.
+
+Another statistical analysis;
 
 ```{r}
-df1 <- base.model1.1$exploitation[, c(1, 4)] %>% mutate(model = "s1.1")
-df2 <- base.model1.2$exploitation[, c(1, 4)] %>% mutate(model = "s1.2")
-df3 <- base.model1.3$exploitation[, c(1, 4)] %>% mutate(model = "s1.3")
-df4 <- base.model1.4$exploitation[, c(1, 4)] %>% mutate(model = "s1.4")
-
-df_all <- bind_rows(df1, df2, df3, df4)
-colnames(df_all) <- c("year", "exploitation_rate", "model")
-
-colores <- c("s1.1" = "#ca0020", 
-             "s1.2" = "#f4a582", 
-             "s1.3" = "#bababa", 
-             "s1.4" = "#404040")
-ggplot(df_all, aes(x = year, y = exploitation_rate, color = model)) +
-  geom_line() +
-  geom_point() +
-  scale_color_manual(values = colores) +
-  labs(
-    title = "Exploitation Rate Over Time",
-    x = "Year",
-    y = "Exploitation Rate",
-    color = "Model"
-  ) +
-  theme_minimal()
-```
-
-### Platoons analisis
-
-```{r}
-# Definir los valores del eje x
-x_values <- seq(0, 7, by=0.1)
-
-# Definir las medias y desviaciones estándar para los valores específicos
-means <- c( 2.9,  3.5, 4.1)
-std_devs <- c(0.65, 0.8, 0.65)
-
-# Crear un dataframe que contenga los valores de x, la media y las desviaciones estándar
-data <- data.frame(
-  x = rep(x_values, each = length(means)),
-  mean = rep(means, times = length(x_values)),
-  std_dev = rep(std_devs, times = length(x_values))
-)
-
-# Calcular las curvas de las desviaciones estándar
-data <- data %>%
-  mutate(y = exp(-((x - mean)^2) / (2 * std_dev^2)))
-
-# Graficar con ggplot
-ggplot(data, aes(x = x, y = y, linetype = as.factor(mean))) +
-  geom_line(size = 1) +
-  scale_linetype_manual(values = means,
-                        name ="Platoon") +
-  labs(x = "", y = "growth increment") +
-  theme_minimal()+
-  xlim(0,7)
-
-```
-
-### Code Repository
-
-The repository with code to replicate this analysis can be found in this [GitHub author link](https://github.com/MauroMardones/SA_Krill)
-
-\newpage
+# Extraer los datos
+spawn_bio_1 <- base.model1.1$timeseries$SpawnBio
+spawn_bio_4 <- base.model1.4$timeseries$SpawnBio
+
+# Prueba de normalidad (Shapiro-Wilk)
+shapiro.test(spawn_bio_1)
+shapiro.test(spawn_bio_4)
+
+# Prueba de igualdad de varianzas (Levene)
+leveneTest(c(spawn_bio_1, spawn_bio_4), 
+           group = rep(c("Model 1.1", "Model 1.4"), each = length(spawn_bio_1)))
+
+# Prueba t de Student (si los datos son normales)
+t.test(spawn_bio_1, spawn_bio_4, var.equal = TRUE)
+
+# Prueba de Wilcoxon (si los datos no son normales)
+wilcox.test(spawn_bio_1, spawn_bio_4)
+
+```
+
+**Two-Sample t-Test**
+
+- **t-value:** 4.2682  
+- **Degrees of Freedom (df):** 72  
+- **p-value:** 5.915e-05 (very small)  
+- **95% Confidence Interval for the Mean Difference:** [1,387,011; 3,818,074]  
+- **Sample Means:**  
+  - Model 1.1: **4,600,304**  
+  - Model 1.4: **1,997,762**
+
+Since the **p-value (5.915e-05) is much smaller than 0.05**, we reject the null hypothesis.
+This means there is **a statistically significant difference** between the spawning biomass predicted by Model 1.1 and Model 1.4.
+The confidence interval suggests that Model 1.1 tends to estimate a higher spawning biomass than Model 1.4, with an estimated difference between **1.39 and 3.82 million tons**.
+
+**Wilcoxon Rank Sum Test (Mann-Whitney U Test)**
+
+- **W-statistic:** 1256  
+- **p-value:** 6.698e-10 (extremely small)  
+- **Warning:** "Cannot compute exact p-value with ties" (indicates that some values are repeated)
+
+The **Wilcoxon test also rejects the null hypothesis** with an extremely small **p-value (6.698e-10)**.
+This confirms that the distribution of spawning biomass values is significantly different between the two models.
+Since the Wilcoxon test does not assume normality, it supports the conclusion from the **t-test**, reinforcing that
+
+Both tests strongly indicate that **Model 1.1 and Model 1.4 produce significantly different estimates of spawning biomass**.
+Specifically, **Model 1.1 predicts higher values**.
+This suggests that the factors or assumptions included in Model 1.4 result in **lower spawning biomass estimates**, which could have important implications for stock assessment and fisheries management.
 

### 2025-06-01 MS2_Krill.Rmd
adding word format chgs

@@
   - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
-  - Erica Jarvis Mason^[Antarctic División NOAA, USA]
-  - Andrea Pinones ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
+  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
+  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
   - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
@@
   - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
-  - Richard Methot ^[NOAA, Fishery Science División, NorthWeast, Seattle, USA]
+  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
 output:
-  bookdown::word_document2:
+  bookdown::pdf_document2:
     reference_docx: MS2_Krill_template.docx.docx
@@
 
+\newpage
+\tableofcontents
+\newpage
+
 ```{r setup1, echo=FALSE}
@@
 
-Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem management objectives in the Western Antarctic Peninsula (WAP).
-In this study, we implement an integrated length-to-age model that not only allows the incorporation of diverse data sources related to krill population dynamics, but also enables the evaluation of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
-We explored four scenarios that accounted (or not) for ecosystem influences on krill population dynamics, yielding results that align with the expected behavior of a stock assessment framework tailored to this type of species.
-Performance analyses reveal systematic differences across those scenarios, highlighting increased bias in fishing mortality (F) and spawning stock biomass (SSB) in models that exclude ecosystem variables.
+Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula.
+In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
+We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species.
+Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables.
 Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics.
-Despite ongoing challenges in incorporating ecosystem variables into integrated model like this, we argue that these factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses.
-Overlooking these drivers may result in less accurate conclusions, hence limiting the effectiveness of management recommendations.
-Integrated modeling approaches that draw on extensive biological, ecologicl, fishery, and environmental datasets provide a valuable foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
-
-or
-
-Understanding the population dynamics of Antarctic krill (Euphausia superba) is essential for achieving ecosystem-based management objectives in the Western Antarctic Peninsula (WAP), particularly under the mandate of CCAMLR.
-In this study, we implement an integrated length-to-age assessment model designed to assimilate multiple data sources related to krill biology, fishery operations, and ecological interactions.
-A key feature of our approach is the explicit inclusion of ecosystem drivers—namely, environmental variability and predation pressure—as intrinsic components of the population dynamics.
-We evaluate four model scenarios, with and without ecosystem variables, and assess their performance in estimating recruitment, fishing mortality (F), total biomass, and spawning stock biomass (SSB).
-The comparison reveals systematic biases in models that omit ecological information, resulting in overestimated recruitment and biomass levels.
-In contrast, scenarios incorporating predator and environmental data yield more precautionary and biologically realistic outcomes, suggesting greater alignment with observed ecosystem variability.
-Despite ongoing challenges related to the availability, resolution, and integration of ecological data, we argue that environmental and trophic drivers should be embedded directly into the state-space structure of assessment models rather than treated as exogenous factors.
-Failing to do so may obscure key population processes and compromise the reliability of management advice.
-Our results underscore the potential of ecosystem-informed, integrated modeling frameworks to support robust, adaptive management strategies that align with CCAMLR’s conservation objectives, while also highlighting the continued need for methodological development and interdisciplinary data collection efforts in the Southern Ocean.
-
-*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, perfomance model, fisheries management, CCAMLR*
+Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses.
+Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations.
+Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
+Such approaches are particularly relevant for ecologically complex systems like the krill fishery in the Western Antarctic Peninsula.
+
+*Keywords: Antarctic krill, population dynamics, length-to-age model, stock-recruitment, ecosystem variables, performance model, fisheries management, CCAMLR*
 
@@
 
-Antarctic krill (*Euphausia superba*) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @Kawaguchi2024; @McBride2021].
-Currently, the population, and consecuently, the fishery is predominantly concentrated in the Western Antarctic Peninsula (WAP).
-This area is also a critical feeding ground for predators such as penguins, seals, and whales, and the primary area for krill harvesting over the past four decades [@Gallagher2023a; @Watters2013; @Hill2016].
-At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Lima2013; @Turner2005; @Carrasco2021; @Atkinson2019a; @Morley2020; @McBride2021].
+Antarctic krill (*Euphausia superba*, hereafter krill) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024].
+Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a].
+At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Turner2005; @Lima2013; @Atkinson2019a; @Morley2020; @Carrasco2021; @McBride2021].
 These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].
 
-Given its ecosystem significance and commercial value, effective management of krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
-Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) that follows an ecosystem-based approach has been developing [@Cavanagh2016; @Constable2023; @Constable2011; @Watters2013].
-This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and time.
-It includes three key components: (i) stock assessments to determine precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a risk assessment framework, now referred to as spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Kawaguchi2024; @Warwick-Evans2022].
-
-To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Green2023; @Hill2019; @Dornam2021].
-However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, and the high costs associated with such studies have further motivated the search for alternative stock estimation methods.
+Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
+Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023].
+This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales.
+It comprises three key components: (i) stock assessments to establish precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Warwick-Evans2022; @Kawaguchi2024].
+
+To provide scientific advice and recommendations on the status and standing stock of krill, CCAMLR has historically relied on dedicated scientific surveys [@Watters2013; @Hill2019; @Dornam2021; @Green2023].
+However, changes in survey design—both methodological and spatial—have raised concerns about the reliability of these estimates, while the high costs associated with such surveys have further motivated the search for alternative stock estimation methods.
 More recently, and as part of a new strategy, CCAMLR has explored a variety of model-based (indirect) approaches to assess krill stock status and support fisheries management decisions [@Hill2019].
-Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019a; @Green2023].
-Nowadays, a model-based approach used for generating management recommendations in CCAMLR is the Generalized Yield Model (GRYM) [@Maschette2020], which allows adjustments to the decision rules to predefined captures set by CCAMLR [@Constable2000; @CCAMLR2021a; @CCAMLR2010].
-However, these model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics and/or failure to account for the high spatial and temporal variability intrinsic to krill systems [@Meyer2020; @Trathan2021].
-
-Increasing evidence shows that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Flores2012; @Morley2020; @Saba2014; @Perry2020; @Veytia2021; @Walsh2020; @Murphy2007; @Vorrath2020].
-Likewise, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Reisinger2022; @Riaz2023; @Baines2022; @Biuw2024; @Moosa2017].
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions [REF].
-Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Marshall2019; @Crone2019; @Richards2016].
-Therefore, any stock assessment frameworks related with new strategy of CCAMLR should explicitly incorporate these ecosystem components to more accurately represent krill population processes.
+Some of these stock assessment models have used a single source of data—either fishery-dependent catches [@Wang2021] or survey-based monitoring programs [@Kinzey2011; @Kinzey2019; @Green2023].
+Nowadays, the Generalized Yield Model (GRYM) [@Maschette2020] is a model-based approach used by CCAMLR to generate management recommendations, allowing adjustments to the decision rules to predefined catch limits set by CCAMLR [@Constable2000; @CCAMLR2010; @CCAMLR2021a].
+However, model-based approaches exhibit limitations, including an effective integration of multiple data sources, incorporation of ecosystem-level variables into population dynamics, and accounting for the high spatial and temporal intrinsic variability of krill productivity [@Meyer2020; @Trathan2021].
+
+Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021].
+Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024].
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@REF].
+Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019].
+Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.
 
 Integrating ecological complexity into krill stock assessments presents multiple challenges.
-Long-term data consistency is often affected by changes in monitoring programs and sampling methodologies [REF], while some key regions of the Western Antarctic Peninsula (WAP) lack sufficient data to construct reliable krill population indices.
-Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [REF].
-Despite these challenges, substantial data exist on krill and their environment.
-Fishery monitoring programs provide detailed catch and effort data through onboard scientific observers.
-Ecosystem monitoring, like the "CCAMLR Ecosystem Monitoring Program" (CEMP), track predator populations and environmental variability at key sites.
+Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@REF], while some key regions of the WAP lack sufficient data to construct reliable krill population indices.
+Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@REF].
+Despite these challenges, substantial data exist on krill and associated environments.
+Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers.
+Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites.
 Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.
@@
 Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
-For statistical catch-at-age assessments, key frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013] among others.
-These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and populations dynamics [@Hoyle2020].
-
-This study aims to integrate different sources of available data on Antarctic krill (*Euphausia superba*) —including fishery operations (catches, effort distribution), biological characteristics (growth rates, reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (environmental forcing and predator-prey interactions)—into a unified stock assessment framework.
-Specifically, we employ an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting  biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
+For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others.
+These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].
+
+This study aims to integrate various data sources on krill (*Euphausia superba*), including fishery operations (e.g.
+catches and effort distribution), biological characteristics (e.g.
+growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g.
+environmental forcing and predator-prey interactions) into a unified stock assessment framework.
+Specifically, we use an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
+This approach strengthens the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR's ecosystem-based management approach [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024].
 The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices.
-Our results provide empirical evidence for the value of ecosystem-informed in model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in  model fit, predictive skill and structural uncertainty.
-
-This approach seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach [@CCAMLRWGSAM2024; @Hill2019; @Constable2023; @Zhao2023; @WGEMM2024; @Cavanagh2016; @Constable2011].
-Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
+Our results provide empirical evidence for the value of ecosystem-informed model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in model fit, predictive skill, and structural uncertainty.
 
@@
 
-The study area covers the Western Antarctic Peninsula (WAP), corresponding to FAO Subarea 48.1, which hosts a high of krill fishing activity.
-This is also the region where CCAMLR is actively developing a revised management approach at a finer spatial scale.
-To enhance the spatial resolution of our analysis of krill population dynamics, we adopted the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
+The study area covers the WAP, corresponding to FAO Subarea 48.1, one of the main regions for krill fishing activity.
+This area is also where CCAMLR is actively revising management approach at a finer spatial scale.
+To enhance the spatial resolution of our analysis of krill population dynamics, we used the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
 1, S1).
-It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological or population-specific criteria [@Zaldua2024a; @Constable2023; @Watters2013].
-However, their subdivision enables a finer-scale analysis of krill data.
-This increased spatial resolution allows for the detection of heterogeneity among various data sources—whether from fisheries or environmental  variables—thus providing a more detailed understanding of krill population dynamics.
+It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a].
+Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources—whether from fisheries or environmental variables—thus providing a more detailed understanding of krill population dynamics.
 
@@
 
-To represent spatial heterogeneity in krill dynamics we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* concept [@Punt2019; @Nielsen2021; @Waterhouse2014].
-The areas-as-fleets approach aims treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations can exert different effects across space.
-In this approach, each strata within the Western Antarctic Peninsula is treated as a distinct fleet and allows a spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices.
-This is especially relevant for krill, given the spatial variability in both population processes and fishing effort, given the abundant evidence of spatial heterogeneity [@Perry2019; @Perry2020; @Reiss2020; @SantaCruz2022; @Veytia2021].
-By modeling each MU as a separate fleet, we allow the assessment to estimate region-specific parameters that reflect this variability.
-Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location based in @Dornam2021.
-Similarly, fishery-dependent data—such as catch and composition—are stratified by MU according to the reported fishing effort allowing for regional variation in both abundance and observational processes.
-Given the difference en the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each strata to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
+To represent spatial heterogeneity in krill dynamics, we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* methodology [@Waterhouse2014; @Punt2019; @Nielsen2021].
+This approach treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations may have varying effects across space.
+Within this approach, each MU of the WAP is treated as a distinct fleet, allowing spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices.
+This is especially relevant for krill due to documented spatial variability and heterogeneity in both population processes and fishing effort [@Perry2019; @Perry2020; @Reiss2020; @Veytia2021; @SantaCruz2022].
+By modeling each MU as a separate fleet, the assessment can estimate region-specific parameters that reflect this variability.
+Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location as defined by [@Dornam2021].
+Similarly, fishery-dependent data—such as catch and size composition—are stratified by MU according to the reported fishing effort, allowing for regional variation in both abundance and observational processes.
+Given known difference in the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each MU to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
 This disaggregated structure allows the model to reflect small-scale spatial dynamics while maintaining analytical simplicity.
-The areas-as-fleets approach captures regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019].
-The model accounts for interannual variability in growth dynamics, maturity, and reproduction, with a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021].
-Although strata are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single, interconnected population providing a more realistic representation of krill dynamics population within FAO Subarea 48.1.
+The areas-as-fleets approach effectively incorporates regional variability in both population dynamics and observational processes without the added complexity of explicitly modeling movement [@Punt2019].
+The model accounts for interannual variability in growth, maturity, and reproduction, assuming a single annual spawning pulse occurring between January and February [@Perry2019; @McBride2021].
+Although MUs are treated as separate fleets, core population processes—such as recruitment and growth—are modeled collectively, representing the krill stock as a single interconnected population.
+This provides a more realistic representation of the krill population dynamics within Subarea 48.1.
 
@@
 
-Like many invertebrates, krill is a hard to age determination, and this presents challenge for modeling population dynamics, because most modern stock assessments that aim to estimate marine population variables rely on age-structured modeling approaches [@Punt2013].
-Although recent advancements have been made [@Kilada2017], aging techniques have historically not been available for krill.
-Therefore, to develop a stock assessment model for this type of species use length data to approximate age dynamics through transition matrices, which estimate the probability of individuals belonging to specific age classes based on their length [@Laslett2004a; @Lee2019; @Lee2024; @Rudd2021; @Zhang2022].
-Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].
-On the other hand, length data, which are cost-effective and readily available [@Canales2021; @Chong2019], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014].
-For made this transformation, the models often use an Age-Length Keys (ALKs).
-An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classed and this method assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024].
-The ALK matrix is typically defined as:
+Like many invertebrates, krill present challenges for age determination, which complicates modeling population dynamics.
+Most modern stock assessments rely on age-structured modeling approaches to estimate marine population parameters [@Punt2013].
+Although recent advancements have improved krill aging techniques [@Kilada2017], such methods have historically been unavailable for krill.
+Consequently, stock assessment for krill typically uses length data to approximate age structure through transition matrices.
+These matrices estimate the probability that individuals of a given length belong to specific age classes [@Laslett2004; @Lee2019; @Rudd2021; @Zhang2022; @Lee2024].
+Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
+To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015; @Kinzey2019; @Wang2021; @Green2023].
+Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Punt2013; @Lee2024].
+On the other hand, length data, which are cost-effective and readily available [@Chong2019; @Canales2021], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014].
+To perform this transformation, models often use Age-Length Keys (ALKs).
+An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classes.
+This method assumes both aged and measured individuals are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024].
+The ALK matrix is typically defined as follows:
 
@@
 
-where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age.
-Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
-To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015a; @Kinzey2019a; @Wang2021; @Green2023].
-The resulting ALK is shown in S1 Fig 7.
+where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age. The resulting ALK is shown in S1 Fig 7.
 
@@
 Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
-The model did not account for hermaphroditism, assuming that all individuals maintained a fixed sex throughout their lifespan.
 The model used a direct parameterization without additional offsets between sexes.
@@
 Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019a, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
-This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019a; @Perry2020].
+The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
+This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019; @Perry2020].
 The unfished recruitment ($R_0$) was estimated with a mean recruitment of 23 in log-scale, with a prior of 10.
@@
 To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019a; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility.
+Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility.
 Life history parameters—such as growth, weight-length relationship, natural mortality, and maturity—were used as priors to model the initial conditions of krill population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
@@
 ft <- theme_booktabs(ft)
+ft <- set_table_properties(ft, width = .5, layout = "autofit")
 ft <- fontsize(ft, size = 9, part = "all")
 ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")
-
-# Mostrar flextable
 ft
-
 ```
@@
 
+#to word
 ft_env <- flextable(env_features) |>
@@
   theme_booktabs() |>
-  fontsize(size = 10) |>
-  set_table_properties(layout = "autofit", width = .85)
-ft_env
+  fontsize(size = 10)
+
+ft_1 <- autofit(ft_env)
+ft_2 <- set_table_properties(ft_1, width = .5, layout = "autofit")
+ft_2
+
 ```
@@
 
-ft_data <- flextable(data) %>%
-  set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
-  autofit() %>%
-  theme_vanilla() %>%
-  fontsize(size = 9, part = "all") %>%
-  font(fontname = "Arial", part = "all") %>%
-  set_table_properties(layout = "autofit", width = .85)
-ft_data
+# # to word
+# ft_data <- flextable(data) %>%
+#   set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
+#   theme_vanilla() %>%
+#   fontsize(size = 9, part = "all") %>%
+#   font(fontname = "Arial", part = "all") %>%
+#   autofit() %>% 
+#   set_table_properties(width = .5, layout = "autofit")
+# 
+# ft_data
+
+# to PDF
+
+kbl(data, format = "latex", 
+    booktabs = TRUE, 
+    caption = "Table: Data sources used in the krill population dynamics modeling") %>%
+  kable_styling(latex_options = c("hold_position", "striped"), 
+                font_size = 9, 
+                full_width = FALSE) %>%
+  column_spec(1, width = "3cm") %>%
+  column_spec(2, width = "5cm") %>%
+  column_spec(3, width = "3cm") %>%
+  column_spec(4, width = "3cm")
+
 ```
@@
    fontsize(size = 10) |>
-  set_table_properties(layout = "autofit", width = .85) %>% 
+  autofit() %>% 
+  set_table_properties(width = 1, layout = "autofit") %>% 
   color(part = "header", color = "black") 
@@
 
-where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which \rho_M was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset.
+where \( X \) represents the quantities (such as spawning stock biomass (SSB) or fishing mortality (F)) for which \(\rho_M\) was calculated, and \( \hat{X} \) is the corresponding estimate from the reference model fitted to the full dataset.
 Year \( y \) was nominally considered as season one, \( T \) refers to the terminal assessment data year, and \( h \) denotes the number of retrospective peels, removing recent years of data ("peeling") and re-estimating the model, allowing for the calculation of retrospective bias (*Rho*) for key metrics such as Spawning Stock Biomass (SSB) and Fishing Mortality (F).
@@
     "The autoregressive correlation coefficient from retrospective analysis, used to assess systematic bias in biomass (B) and spawning stock biomass (SSB) estimates across years.",
-    "Measures the variability or uncertainty in the estimated unfished recruitment (R₀). High deviations may indicate poor estimation stability or model overparameterization."
+    "Measures the variability or uncertainty in the estimated unfished recruitment (R0). High deviations may indicate poor estimation stability or model overparameterization."
   ),
@@
 
-ft_stat <- flextable(df_model_metrics) |> 
-  set_header_labels(
-    Metric = "Metric",
-    Description = "Description",
-    Reference = "Reference"
-  ) |> 
-  set_caption("Summary of model diagnostic metrics used for krill stock assessment") |> 
-  autofit() |> 
-  fontsize(size = 10) |> 
-  theme_booktabs() |> 
-  color(part = "header", color = "black")|>
-  set_table_properties(layout = "autofit", width = .85)
-ft_stat
-```
+# ft_stat <- flextable(df_model_metrics) |> 
+#   set_header_labels(
+#     Metric = "Metric",
+#     Description = "Description",
+#     Reference = "Reference"
+#   ) |> 
+#   set_caption("Summary of model diagnostic metrics used for krill stock assessment") |> 
+#   #fontsize(size = 7) |> 
+#   theme_booktabs() |> 
+#   color(part = "header", color = "black")|>
+#   set_table_properties(layout = "autofit")
+# ft_stat
+
+# to pdf
+
+kbl(df_model_metrics, format = "latex", booktabs = TRUE,
+    caption = "Summary of model diagnostic metrics used for krill stock assessment") |>
+  kable_styling(latex_options = c("hold_position", "scale_down", "striped"), 
+                font_size = 9) |>
+  column_spec(1, width = "3.5cm", latex_valign = "m") |>
+  column_spec(2, width = "9cm", latex_valign = "m") |>
+  column_spec(3, width = "3.5cm", latex_valign = "m")
+
+```
+\pagebreak
 
@@
 
-```{r mapachl, eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in Subarea 48.1"}
+```{r mapachl, out.width="100%", eval=TRUE, fig.cap="Time series (2000-2020) to Chlorophyll in Subarea 48.1"}
 knitr::include_graphics('Figs/chla_grilled.png')
@@
 The interannual variability of this environmental index from 2000 to 2020 is shown in Figure \@ref(fig:chlindex).
-The figure shown illustrates the interannual variability of a Chlorophyll Variability Index from the year 2000 to 2020.
 The figure shows fluctuations across years, highlighting periods of higher-than-average chlorophyll variability (e.g., 2003, 2016) as well as years with significantly lower values (e.g., 2011, 2020).
@@
 
-### Predator comoponent as driver of krill dynamics
-
-Figure \@ref(fig:mapapre) presents a temporal and spatial overview of penguin variables from 2001 to 2020.
-The top panel in Figure \@ref(fig:mapapre) shows annual krill length distributions derived from penguin diet samples, displayed by predator species.
-Although median krill lengths remain relatively consistent over time, there is an interannual variability in the spread and distribution of sizes.
+### Predator component as driver of krill dynamics
+
+The top panel in Figure \@ref(fig:mapapre) shows annual krill length distributions derived from penguin diet samples, displayed by predator species, showing that although median krill lengths remain relatively consistent over time, there is an interannual variability in the spread and distribution of sizes.
 Spatiotemporal distribution patterns of the three penguin species across the Antarctic Peninsula showed interannual variability from 2001 to 2020.
@@
 Gentoo penguins appeared in fewer and more localized sites but displayed a slight increase in spatial occurrence in the latter part of the time series (middle panel in Figure \@ref(fig:mapapre)).
-For stock assessment purposes, we integrated all penguin colony abundances into a synoptic index of relative predator biomass, which represents a comprehensive measure of population abundance trends over time (bottom panel in Figure \@ref(fig:mapapre).
-This index accounts for variations in colonies abundance providing a robust metric that reflects the overall status of the penguin population.
+
+The synoptic index of relative predator biomass accounts for variations in colony abundance, providing a robust metric that reflects the overall status of the penguin population (bottom panel in Figure \@ref(fig:mapapre)).
 The index appears to exhibit fluctuations over time, with a general decline from the early 1990s to the mid-2000s, followed by a period of relative stabilization and a strong decline toward the end of the time series.
-This decline has been extensively documented reported in the Antarctic Peninsula by studies such as @Waterhouse2014, @Dunn2019a and @Wethington2023, primarily associated with colonies of the Adélie penguin [@Strycker2020].
+This decline has been extensively documented in the Antarctic Peninsula by studies such as [@Waterhouse2014; @Dunn2019; @Wethington2023], primarily associated with colonies of the Adélie penguin and Chinstrap penguin [@Strycker2020; @Talis2023].
 This abundance series captures broad decadal declines in predator presence, particularly after 2015, potentially reflecting changes in prey availability, environmental conditions, or predator demography.
 
-```{r mapapre, eval=TRUE, fig.cap= "Ecosystem indicators derived from penguin predator data in the Western Antarctic Peninsula.
+```{r mapapre, eval=TRUE, out.width="100%", fig.cap= "Ecosystem indicators derived from penguin predator data in the Western Antarctic Peninsula.
 Top panel: Annual krill length distributions from penguin diet samples by species: Adélie (red), chinstrap (orange), and gentoo (black).
@@
 
-Natural mortality (M) of Antarctic krill varied with age, exhibiting a sharp increase from age-0 (M = 0.30) to age-1 (M = 0.62), followed by a gradual asymptotic rise reaching approximately M = 0.76 from age-4 onward (S1.
+Natural mortality (M) of krill varied with age, exhibiting a sharp increase from age-0 (M = 0.30) to age-1 (M = 0.62), followed by a gradual asymptotic rise reaching approximately M = 0.76 from age-4 onward (S1.
 Fig x).
@@
 This is consistent with the dietary reliance of apex predators such as *Pygoscelis* penguins, which feed extensively on sub-adult and adult krill during the breeding season.
-The overlap between peak penguin chick-rearing periods and krill availability reinforces the ecological significance of these age classes in sustaining predator populations.
-Additionally, the extended presence of krill in predator diets across age classes supports the inference of sustained natural mortality beyond early ontogeny.
-
-Would you like to incorporate other predator groups like seals or whales into the analysis?
 
@@
 
-```{r pearson, eval=TRUE, fig.cap= "Pearson residuals of predicted versus observed length-frequency distributions of Antarctic krill across four modeling scenarios (s1.1 to s1.4), each incorporating increasing levels of ecosystem complexity.
+\pagebreak
+
+\begin{landscape}
+
+```{r pearson, out.width="100%",  eval=TRUE, fig.cap= "Pearson residuals of predicted versus observed length-frequency distributions of Antarctic krill across four modeling scenarios (s1.1 to s1.4), each incorporating increasing levels of ecosystem complexity.
 Residuals are shown by fleet, survey, and predator source, allowing visual assessment of model fit across time and sampling components (BS=Bransfield Strait, EI= Elephant Island, GS= Gerlache Strait, JOIN= Joinville Island, SSWI= South West)"}
@@
 ```
+
+\end{landscape}
+
+\pagebreak
 
@@
   #font(fontname = "Arial", part = "all") %>%
-  set_table_properties(layout = "autofit", width = .85)
+  set_table_properties(width = .5, layout = "autofit")
 
@@
 
-Model outputs under different ecosystem scenarios revealed substantial variation in krill population indicators, including total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio) (Figure \@ref(fig:compara)).
+Model outputs under different ecosystem scenarios revealed substantial variation in krill population indicators, including total biomass, summary biomass, recruitment and spawning biomass (Figure \@ref(fig:compara)).
 The reference model (s1.1), which excluded environmental and predator components, consistently produced higher median estimates and wider uncertainty intervals across all population indicators.
@@
 
-```{r compara,eval=TRUE, fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity.
-Metrics include total biomass (Bio_all), summary biomass (Bio_smry), recruitment (Recruit_0), and spawning biomass (SpawnBio)"}
+```{r compara, eval=TRUE,  out.width="100%", fig.cap="Comparisons of krill population variables under scenarios of increasing ecosystem complexity"}
 knitr::include_graphics('Figs/unnamed-chunk-66-1.jpeg')
@@
 From that point onward, both models show a marked increase in biomass, although the reference model predicts higher values than the ecosystem-informed model, particularly during the final assessment years (2018–2020).
-Recruitment at age 0 exhibits strong variability over time in coincidence with [@Dornam and REF].
+Recruitment at age 0 exhibits strong variability over time in coincidence with [@Dornam2021 and REF].
 While both models show similar patterns, the reference model tends to estimate higher recruitment peaks, especially in recent years.
@@
 
-```{r compara1,eval=TRUE, fig.cap="Time series of population variables of krill in WAP: (A) total biomass (Bio_all), (B) summary biomass (Bio_smry), (C) recruitment (Recruit_0), and (D) spawning biomass (SpawnBio).
-Axes labels should clarify units (e.g., biomass in metric tons, recruitment in millions).
-Projections assume current management/environmental conditions."}
+```{r compara1,eval=TRUE,  out.width="100%", fig.cap="Time series of population variables of krill in WAP: (A) total biomass, (B) summary biomass, (C) recruitment, and (D) spawning biomass"}
 knitr::include_graphics('Figs/alloutput_variable2.png')
@@
 
-```{r compara2, eval=TRUE, out.width="60%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
+```{r compara2, eval=TRUE, out.width="75%", fig.cap="Estimated R0 parameter for Antarctic krill under four assessment scenarios. Models including predator and environmental data"}
 knitr::include_graphics('Figs/R0plot.png')
@@
 
-We compare the stock–recruitment relationship and productivity under all model scenarios with focus in differences between scenarios with and without ecosystems consideratios (s1.1 and s1.4).
+We compare the stock–recruitment relationship and productivity under all model scenarios with focus on differences between scenarios with and without ecosystem considerations (s1.1 and s1.4).
 The left panel in Figure \@ref(fig:proplot) illustrates the relationship between spawning stock biomass (SSB) and recruitment across years, overlaid with a fitted Beverton-Holt function derived from the model with ecosystem covariates.
@@
 
-```{r proplot, eval=TRUE, fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios (s1.1 to s1.4).
+```{r proplot, eval=TRUE,  out.width="100%", fig.cap="Left panel: Stock–recruitment relationships for Antarctic krill under four assessment scenarios (s1.1 to s1.4).
 Left panel: Predicted recruitment as a function of spawning stock biomass (SSB), with fitted Beverton-Holt curves and annual labels for selected years.
@@
 
-```{r acfplot, eval=TRUE, fig.cap="Autocorrelation function (ACF) plots of model residuals for four krill assessment scenarios Blue dashed lines indicate 95% confidence intervals"}
+```{r acfplot, eval=TRUE,  out.width="100%", fig.cap="Autocorrelation function (ACF) plots of model residuals for four krill assessment scenarios Blue dashed lines indicate 95% confidence intervals"}
 knitr::include_graphics('Figs/ACF_Plot.png')
@@
 These results suggest that environmental covariates contribute more effectively to explaining krill population dynamics than predator effects alone.
-The total likelihood (−logL) further supports this, showing much better fit in models s1.1 and s1.3, while predator-inclusive models (s1.2 and s1.4) show considerably poorer likelihood values.
+The total likelihood (-logL) further supports this, showing much better fit in models s1.1 and s1.3, while predator-inclusive models (s1.2 and s1.4) show considerably poorer likelihood values.
 This pattern is reflected in the survey and length composition likelihood components, where again the environment-only model (s1.3) shows superior performance compared to predator scenarios.
@@
 Retrospective biases (rho values) provide insight into model consistency over time.
-Retrospective bias in spawning stock biomass (Retro_Rho_ssb) and forecast biomass (Forecast_Rho_ssb) is substantially reduced in models without predators, particularly s1.1 (−0.233 and −0.333).
+Retrospective bias in spawning stock biomass (Retro_Rho_ssb) and forecast biomass (Forecast_Rho_ssb) is substantially reduced in models without predators, particularly s1.1 (-0.233 and -0.333).
 Models including predators (s1.2 and s1.4) show larger negative rho values, indicating stronger retrospective bias and less stability in biomass estimation.
@@
   set_caption("Statistical componentes to evaluate perfomances and diagnostic of model scenarios") %>%
-  autofit() %>%
   theme_booktabs() %>%
   fontsize(size = 10, part = "all") %>%
-  set_table_properties(width = .85, layout = "autofit")
-```
-
-# Discussion
-
-(small guide)
-
-* **Big summary / importance of this kind of integrated model in Krill**
-
-* **Scientific main findings as extension of previous work (comparisons)**
-
-* **Findings' scientific implications**
-
-* **Forward-looking statements about the field as a whole**
-
-* **A summary / importance of the engineering aspect of the work**
-
-* **Limitations of Integrated models shown with questions that would be unanswered**
-
-* **future work or recommendations to improve this**
-
-One critical issue concerns the dependence on direct biomass estimates, which are typically derived from expensive and logistically demanding scientific surveys—most notably acoustic campaigns.
+  autofit() %>%
+  set_table_properties(width = .5, layout = "autofit")
+```
+
+<!-- # Discussion -->
+
+<!-- (small guide) -->
+
+<!-- * **Big summary / importance of this kind of integrated model in Krill** -->
+
+<!-- * **Scientific main findings as extension of previous work (comparisons)** -->
+
+<!-- * **Findings' scientific implications** -->
+
+<!-- * **Forward-looking statements about the field as a whole** -->
+
+<!-- * **A summary / importance of the engineering aspect of the work** -->
+
+<!-- * **Limitations of Integrated models shown with questions that would be unanswered** -->
+
+<!-- * **future work or recommendations to improve this** -->
+
+<!-- One critical issue concerns the dependence on direct biomass estimates, which are typically derived from expensive and logistically demanding scientific surveys—most notably acoustic campaigns.
 These surveys, while highly informative, are methodologically complex and costly to repeat with sufficient temporal resolution.
@@
 This highlights the need made a transition to a **more realistic** for a integrated model and ecosystem-informed approaches to stock assessment that consider the evolving nature of environmental and ecological drivers, such as changes in sea ice extent, primary productivity, and predator-prey interactions.
-
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
+-->
+
+<!-- There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and account for ecosystem interactions.
 However, integrating these complexities presents several challenges.
@@
 Despite these gaps, a substantial amount of data exists on krill populations and their ecosystem, making it crucial to identify knowledge gaps and integrate available information into comprehensive models that support management recommendations.
-
-The transition from traditional, model-based approaches like GRYM to more integrated, ecosystem-based models presents a significant opportunity to improve the scientific basis of krill stock assessments.
+-->
+
+<!-- The transition from traditional, model-based approaches like GRYM to more integrated, ecosystem-based models presents a significant opportunity to improve the scientific basis of krill stock assessments.
 However, this shift also requires addressing the trade-offs between model complexity and data availability [REF].
@@
 Thus, the development of integrated models requires continuous refinement to ensure that they remain relevant and robust in the face of change in various dimensions [REF].
-
-By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions in a realistic way.
+-->
+
+<!-- By integrating ecosystem drivers into the stock assessment process, this study aims to demonstrate how incorporating environmental and ecological factors enhances the accuracy of Antarctic krill population estimates and improves the ability to predict future stock trends under changing climate and ecological conditions in a realistic way.
 This perspective seeks to strengthen the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR’s ecosystem-based management approach.
@@
 Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean.
-
-Although integrated models are not new to krill in the Southern Ocean, CCAMLR has for years applied methodologies that incorporate diverse data sources for several years.
+-->
+
+<!-- Although integrated models are not new to krill in the Southern Ocean, CCAMLR has for years applied methodologies that incorporate diverse data sources for several years.
 External reviews, such as @Thomson2016, have endorsed these approaches as improvements over traditional models.
-Previous studies like @Kinzey2011, @Kinzey2015a, @Kinzey2019a, and @Wang2021 have advanced integrated frameworks, though challenges like model convergence remain.
-
-This analysis represents a modern approach to stock assessment of complex dynamics such as krill, integrating ecosystem components like environmental variables and ecological aspects such as predators.
-Integrated stock assessment models can support adaptive management, wherein outputs like status or quotas are adjusted based on ecosystem monitoring indices [REF]
-
-Spatial heterogeneity is a fundamental aspect of krill ecology, driven by factors including oceanographic conditions and predator foraging behavior.
+Previous studies like @Kinzey2011, @Kinzey2015, @Kinzey2019, and @Wang2021 have advanced integrated frameworks, though challenges like model convergence remain.
+-->
+
+<!-- This analysis represents a modern approach to stock assessment of complex dynamics such as krill, integrating ecosystem components like environmental variables and ecological aspects such as predators.
+Integrated stock assessment models can support adaptive management, wherein outputs like status or quotas are adjusted based on ecosystem monitoring indices [REF] -->
+
+<!-- Spatial heterogeneity is a fundamental aspect of krill ecology, driven by factors including oceanographic conditions and predator foraging behavior.
 Incorporating strata into our model captures complex spatial patterns of krill distribution, enabling identification of hotspots, productivity zones, and ecological corridors.
-[REF]
-
-These progressive increments in parameter numbers reflect the added ecological realism and mechanistic detail captured in each scenario, emphasizing the role of top-down and bottom-up processes in krill dynamics.
-
-Tabla summary
-
-The analysis suggests that incorporating environmental variables (s1.3 and s1.4) improves model fit and predictive accuracy, particularly when compared to predator-only models.
+[REF] -->
+
+<!-- These progressive increments in parameter numbers reflect the added ecological realism and mechanistic detail captured in each scenario, emphasizing the role of top-down and bottom-up processes in krill dynamics.
+-->
+
+<!-- Tabla summary -->
+
+<!-- The analysis suggests that incorporating environmental variables (s1.3 and s1.4) improves model fit and predictive accuracy, particularly when compared to predator-only models.
 The model that includes only environmental covariates (s1.3) strikes the best balance between fit quality, complexity, and retrospective performance, suggesting that environmental variability is a key driver of krill population dynamics in the Western Antarctic Peninsula.
-
-In contrast, the inclusion of predators alone (s1.2) leads to poorer model fit and higher retrospective bias, possibly due to data limitations or unaccounted complexity in predator-prey dynamics.
+-->
+
+<!-- In contrast, the inclusion of predators alone (s1.2) leads to poorer model fit and higher retrospective bias, possibly due to data limitations or unaccounted complexity in predator-prey dynamics.
 While the full ecosystem model (s1.4) shows some improvements over the predator-only model, it still exhibits significant retrospective biases, highlighting the need for refined integration of ecosystem components and further data support, particularly regarding predator consumption rates and spatial foraging behavior.
-
-Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the joint inclusion of predator and environmental variables yielding the most robust representation of observed length compositions.
+-->
+
+<!-- Overall, the results support the conclusion that ecosystem-informed models better capture the temporal and size-structured dynamics of krill, with the joint inclusion of predator and environmental variables yielding the most robust representation of observed length compositions.
 The inclusion of environmental covariates (s1.3 and s1.4) markedly improves the temporal independence of model residuals, as indicated by lower autocorrelation beyond lag 1.
@@
 Conversely, the predator-only model (s1.2) introduces more persistent autocorrelation, indicating that its effects may require more nuanced representation or additional process structuring.
-
-The analysis indicates that incorporating environmental and predator data into the stock assessment framework introduces greater variability in model fit, as measured by RMSE.
+-->
+
+<!-- The analysis indicates that incorporating environmental and predator data into the stock assessment framework introduces greater variability in model fit, as measured by RMSE.
 While the baseline model (Model 1.1) demonstrates more stable performance, it potentially achieves this at the cost of biological realism.
 The modest trade-off in predictive accuracy associated with ecosystem-informed models is expected and highlights the need to balance goodness-of-fit with ecological plausibility when developing integrated assessment models.
-
-The inclusion of predator and environmental variables significantly alters model behavior in retrospective diagnostics, particularly increasing bias in terminal-year estimates of fishing mortality and deepening the underestimation of spawning stock biomass (SSB).
+-->
+
+<!-- The inclusion of predator and environmental variables significantly alters model behavior in retrospective diagnostics, particularly increasing bias in terminal-year estimates of fishing mortality and deepening the underestimation of spawning stock biomass (SSB).
 These patterns suggest that while ecosystem-informed models improve the biological realism of the stock assessment, they also pose challenges in terms of stability and predictive consistency.
 Such outcomes underline the importance of further refining the structure and parameterization of ecosystem components when integrating them into assessment frameworks, to balance realism with diagnostic robustness.
-
-Likelihood composition reveals that models incorporating ecological data (s1.2–s1.4) yield slightly lower total likelihoods than the reference model (s1.1), reflecting trade-offs between model fit and biological realism.
+-->
+
+<!-- Likelihood composition reveals that models incorporating ecological data (s1.2–s1.4) yield slightly lower total likelihoods than the reference model (s1.1), reflecting trade-offs between model fit and biological realism.
 The reduced penalty terms and more evenly distributed likelihood contributions in the ecosystem scenarios suggest improved model stability and reduced overparameterization.
 These results reinforce the value of integrating ecosystem variables not only for enhancing ecological interpretation but also for promoting robust model structure and parameter behavior.
-
-> (pocos estudios incluyen ambas variables. problemas en interpretacion y cofounding correlation between parameters)
-
-@Boldt2025
-
-### Ecosystem Variables: krill population dynamics relevance
-
-> regarding clorofila como driver; This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
-
->  About index; These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
-
-Variations in krill size composition in penguin diets have been shown to reflect underlying changes in krill population dynamics, including recruitment pulses and cohort strength [Jeff REFs].
+-->
+
+<!-- > (pocos estudios incluyen ambas variables. problemas en interpretacion y cofounding correlation between parameters)  -->
+
+<!-- @Boldt2025 -->
+
+<!-- ### Ecosystem Variables: krill population dynamics relevance -->
+
+<!-- > regarding clorofila como driver; This is consistent with previous studies that have identified this variable as an important driver of variability in different biological components of the krill population and shaping krill habitat suitability and productivity [REF REF REF].
+-->
+
+<!-- >  About index; These patterns suggest substantial year-to-year environmental changes, which may have implications for marine ecosystem productivity and associated trophic dynamics and krill population impacts [REF].
+-->
+
+<!-- Variations in krill size composition in penguin diets have been shown to reflect underlying changes in krill population dynamics, including recruitment pulses and cohort strength [Jeff REFs].
 Thus, penguin diet data serve as a proxy for detecting cyclical patterns in krill recruitment that may not be evident through direct surveys alone.
-[REF]
-
-The ecosystem variables used in assessment processes vary from their definition to how they are applied.
+[REF] -->
+
+<!-- The ecosystem variables used in assessment processes vary from their definition to how they are applied.
 Environmental variables influencing krill recruitment, as well as predators that introduce an additional component of natural mortality into the model, are among those considered in this study.
-[REF]
-
-Marshall et al.
+[REF]  -->
+
+<!-- Marshall et al.
 (2019) [REF] classify ecosystem variables based on their impact on population dynamics and how they are incorporated into stock assessment models.
@@
 In contrast, external incorporation may not fully capture ecosystem feedback loops, potentially leading to less reliable assessments.
-
-Our analysis, which integrates ecosystem variables directly within the model, aligns with the conclusions of Marshall et al.
+-->
+
+<!-- Our analysis, which integrates ecosystem variables directly within the model, aligns with the conclusions of Marshall et al.
 (2019) and provides further evidence that this approach offers superior accuracy in modeling krill population dynamics.
@@
 This integrated approach not only improves model precision but also enhances its predictive power, addressing key ecological factors that influence krill populations.
-
-This analysis represents a significant step forward in the incorporation of ecosystem-based considerations into krill stock assessments in the Southern Ocean.
+-->
+
+<!-- This analysis represents a significant step forward in the incorporation of ecosystem-based considerations into krill stock assessments in the Southern Ocean.
 By adopting a more comprehensive, ecosystem-informed model, we contribute to the growing body of research supporting the integration of ecosystem variables in stock assessment frameworks [REF].
@@
 As climate change and ecosystem variability continue to impact krill populations, this analysis provides a foundation for precautionary management strategies that are better aligned with the dynamic nature of marine ecosystems.
-
-Like @Crone2019, the study concluded that both methods for including environmental data resulted in relatively high-quality estimates for a diverse group of output variables.
-The best performing models were those that included a correct and bias-corrected environmental variable, indicating the importance of accurately incorporating environmental factors in stock assessments
-
-Regarding to evironmental impact in modelliling of main population variables in stock assessment process, @Haltuch2009 recommends that low-frequency environmental variability significantly affects fish population abundance through recruitment, and biomass reference points should be based on average recruitment and/or dynamic B0, especially if comprehensive catch and survey data are available, y a su vez, la calidad y cantidad de datos mejorar la exactitud de resultados.
-
-Overall, these results highlight the benefit of including environmental signals to reduce residual autocorrelation and enhance model fit. This analysis coincides with... [REF]
-
-### Predation and Natural Mortality
-
-The study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
+-->
+
+<!-- Like @Crone2019, the study concluded that both methods for including environmental data resulted in relatively high-quality estimates for a diverse group of output variables.
+The best performing models were those that included a correct and bias-corrected environmental variable, indicating the importance of accurately incorporating environmental factors in stock assessments -->
+
+<!-- Regarding to evironmental impact in modelliling of main population variables in stock assessment process, @Haltuch2009 recommends that low-frequency environmental variability significantly affects fish population abundance through recruitment, and biomass reference points should be based on average recruitment and/or dynamic B0, especially if comprehensive catch and survey data are available, y a su vez, la calidad y cantidad de datos mejorar la exactitud de resultados.
+-->
+
+<!-- Overall, these results highlight the benefit of including environmental signals to reduce residual autocorrelation and enhance model fit. This analysis coincides with... [REF] -->
+
+<!-- ### Predation and Natural Mortality -->
+
+<!-- The study area encompasses the primary spawning, breeding, and wintering regions of Antarctic krill in the southwest Atlantic Ocean.
 Evidence indicates that penguin predation exerts strong top-down control on krill populations.
 Penguins, as krill specialists, have reproductive success tied closely to krill availability.
-[REF]
-
-Incorporating predation into stock models addresses the underestimation of uncertainty in natural mortality, which can otherwise skew biomass and productivity estimates by as much as 20% [@Hollowed2000].
-
-> The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lucas, REF Vicky].
-
-> This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics
-
-> variability length; , which may reflect fluctuations in recruitment strength and availability of krill age classes.
-
-@Hollowed2000 demostrate a successful integration of predation mortality into Gulf of Alaska walleye pollock assessments showing that the value of predator data (e.g., diet composition, predator abundance) for refining natural mortality estimates and understanding ecosystem dynamics improve the estimations.
+[REF] -->
+
+<!-- Incorporating predation into stock models addresses the underestimation of uncertainty in natural mortality, which can otherwise skew biomass and productivity estimates by as much as 20% [@Hollowed2000].
+-->
+
+<!-- > The three main penguin colonies Adélie, Chinstrap, and Gentoo are central place foragers with diet compositions that are strongly linked to krill abundance and demographic structure, particularly during the breeding season when foraging is spatially and temporally constrained [REF Lucas, REF Vicky].
+-->
+
+<!-- > This spatial overlap reinforces the ecological relevance of penguin-based indicators in reflecting krill dynamics  -->
+
+<!-- > variability length; , which may reflect fluctuations in recruitment strength and availability of krill age classes. -->
+
+<!-- @Hollowed2000 demostrate a successful integration of predation mortality into Gulf of Alaska walleye pollock assessments showing that the value of predator data (e.g., diet composition, predator abundance) for refining natural mortality estimates and understanding ecosystem dynamics improve the estimations.
 Explicit modeling of key predators (e.g., arrowtooth flounder, Steller sea lions) improved biomass estimates by addressing previously unquantified mortality sources, particularly for younger age classes.
-
-However, reliance on predator data introduces challenges: (1) interannual and age-specific variability in natural mortality (especially for vulnerable stages) requires high-resolution trophic data; (2) dome-shaped selectivity patterns in predators may complicate mortality rate estimation; and (3) gaps in spatiotemporal coverage of predator diets can bias assessments.
-
-To enhance robustness, future work should prioritize: (1) expanded monitoring of predator diets across seasons, (2) dynamic selectivity models to account for predator behavior, and (3) ecosystem-based frameworks that explicitly link krill dynamics to predator demand and environmental shifts.
+-->
+
+<!-- However, reliance on predator data introduces challenges: (1) interannual and age-specific variability in natural mortality (especially for vulnerable stages) requires high-resolution trophic data; (2) dome-shaped selectivity patterns in predators may complicate mortality rate estimation; and (3) gaps in spatiotemporal coverage of predator diets can bias assessments.
+-->
+
+<!-- To enhance robustness, future work should prioritize: (1) expanded monitoring of predator diets across seasons, (2) dynamic selectivity models to account for predator behavior, and (3) ecosystem-based frameworks that explicitly link krill dynamics to predator demand and environmental shifts.
 While predator data significantly reduces uncertainty in mortality estimates, its efficacy depends on systematic data collection and integration into holistic management strategies.
-
-Inclusion of predator in this analisis coincide with @Moustahfid2009, where they found that explicitly including predation mortality in stock assessment models significantly alters the outputs.
+-->
+
+<!-- Inclusion of predator in this analisis coincide with @Moustahfid2009, where they found that explicitly including predation mortality in stock assessment models significantly alters the outputs.
 When predation was not accounted for, the model underestimated both the magnitude and uncertainty of spawning-stock biomass (SSB) and recruitment levels.
-
-(reflection)
-
-> Although penguin predation on krill does not reach the magnitude required to decimate the planktonic crustacean population—as might be the case with baleen whales [@Baines2022; @Biuw2024]—it is nevertheless important that stock assessment models incorporate signals regarding the trends of this factor on population variables.
+-->
+
+<!-- (reflection)  -->
+
+<!-- > Although penguin predation on krill does not reach the magnitude required to decimate the planktonic crustacean population—as might be the case with baleen whales [@Baines2022; @Biuw2024]—it is nevertheless important that stock assessment models incorporate signals regarding the trends of this factor on population variables.
 Since the index is relative rather than absolute, it provides valuable information for modeling krill population dynamics.
-
-> The substantial variability between consumption estimates presents challenges for model integration.
+-->
+
+<!-- > The substantial variability between consumption estimates presents challenges for model integration.
 While some studies suggest predators consume approximately 20% of the krill standing stock [@Baines2022], others indicate consumption levels around 20% of the total krill catch [@Biuw2024].
 This discrepancy in baseline metrics complicates comparative analysis and requires careful reconciliation when incorporating these estimates into integrated models.
-
-(Talk about decrease in penguins colonies at 2020 an impact in modeling)
-
-### Productivity
-
-Integrating environmental and predator components into the population model (s1.4) substantially affects the shape and variability of the stock–recruitment relationship.
+-->
+
+<!-- (Talk about decrease in penguins colonies at 2020 an impact in modeling) -->
+
+<!-- The overlap between peak penguin chick-rearing periods and krill availability reinforces the ecological significance of these age classes in sustaining predator populations.
+Additionally, the extended presence of krill in predator diets across age classes supports the inference of sustained natural mortality beyond early ontogeny.
+-->
+
+<!-- ### Productivity -->
+
+<!-- Integrating environmental and predator components into the population model (s1.4) substantially affects the shape and variability of the stock–recruitment relationship.
 The ecosystem-informed model reduces the occurrence of extreme recruitment outliers and yields a more consistent relationship between SSB and recruitment.
@@
 The inclusion of these variables leads to more conservative and constrained outputs, which may better reflect the ecological dynamics governing krill populations.
-
-> This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill.
-
-The inclusion of environmental variables (s1.3 and s1.4) substantially lowers estimates of per-capita productivity, especially at low stock levels.
+-->
+
+<!-- > This highlights the critical role of ecosystem components in influencing recruitment and population structure, and their importance in comprehensive stock assessment frameworks for krill. -->
+
+<!-- The inclusion of environmental variables (s1.3 and s1.4) substantially lowers estimates of per-capita productivity, especially at low stock levels.
 These ecosystem-informed scenarios reflect more realistic ecological constraints and reduce the risk of overestimating krill productivity under unfavorable conditions.
 This highlights the importance of accounting for environmental drivers in stock assessment models to avoid overly optimistic projections of stock resilience.
-[REF]
-
-Population variables: Overall, the inclusion of ecosystem and predator data (s1.4) results in more conservative estimates of biomass, recruitment, and spawning stock, suggesting that ecosystem-informed models may provide a more precautionary and ecologically realistic basis for krill stock assessment and management.
-
-CC coments : **"perhaps it is worth haveing another read, and combine some bits of discussion, there’s repetition of highlighting the important of the model and env variables .
-This is true but there’s  no need of saying it everytime you discuss a parameter"**
-
-> SRR: These patterns reflect a more precautionary and ecologically realistic recruitment dynamic, where environmental variability limits the ability of the stock to maintain high productivity under depleted conditions.
+[REF] -->
+
+<!-- Population variables: Overall, the inclusion of ecosystem and predator data (s1.4) results in more conservative estimates of biomass, recruitment, and spawning stock, suggesting that ecosystem-informed models may provide a more precautionary and ecologically realistic basis for krill stock assessment and management.
+-->
+
+<!-- CC coments : **"perhaps it is worth haveing another read, and combine some bits of discussion, there’s repetition of highlighting the important of the model and env variables .
+This is true but there’s  no need of saying it everytime you discuss a parameter"** -->
+
+<!-- > SRR: These patterns reflect a more precautionary and ecologically realistic recruitment dynamic, where environmental variability limits the ability of the stock to maintain high productivity under depleted conditions.
 Overall, the results emphasize the influence of ecosystem variables on shaping recruitment dynamics and demonstrate that incorporating environmental drivers into the assessment model leads to more conservative productivity estimates, particularly at low spawning biomass levels.
-
-### Model integrated limitations
-
-(Discussion)
-
-The observed retrospective bias in SSB and F aligns with findings from previous studies on krill population dynamics, where recruitment variability and model structure influence stock assessment estimates (Hill et al., 2016; Watters et al., 2020).
+-->
+
+<!-- ### Model integrated limitations  -->
+
+<!-- > While the inclusion of ecosystem variables in stock assessment models has gained momentum in recent years, there remains a notable paucity of studies that integrate both predation and environmental forcing within a single modeling framework.
+This gap may stem from the inherent complexity of disentangling the individual and potentially interactive effects of these covariates, as they are often correlated in space and time.
+Such confounding can obscure causal relationships, making it difficult to attribute observed changes in population dynamics to specific ecological drivers.
+Furthermore, methodological limitations and the lack of standardized approaches for simultaneously incorporating multiple ecosystem components have constrained broader implementation.
+Nevertheless, advancing our understanding of how predation pressure and environmental variability jointly affect stock status is essential for developing robust, ecosystem-informed management strategies.
+Addressing this challenge requires both conceptual and analytical efforts to adequately represent these dynamics, evaluate their contributions, and quantify associated uncertainties.
+-->
+
+<!-- (Discussion) -->
+
+<!-- The observed retrospective bias in SSB and F aligns with findings from previous studies on krill population dynamics, where recruitment variability and model structure influence stock assessment estimates (Hill et al., 2016; Watters et al., 2020).
 The negative bias in SSB suggests that models may overestimate past biomass, potentially due to unaccounted environmental variability or underestimation of natural mortality (Meyer et al., 2021).
@@
 These results highlight the importance of incorporating ecosystem interactions into stock assessments to improve predictive accuracy and support sustainable management of Antarctic krill.
-
-The study on krill population dynamics presents several limitations that should be acknowledged for a comprehensive understanding of its findings. Here are the key limitations identified:
-
-- Integrating multiple data sources and accounting for temporal and spatial variability in stock assessments presents significant challenges.
+-->
+
+<!-- The study on krill population dynamics presents several limitations that should be acknowledged for a comprehensive understanding of its findings. Here are the key limitations identified: -->
+
+<!-- - Integrating multiple data sources and accounting for temporal and spatial variability in stock assessments presents significant challenges.
 The study highlights that long-term data consistency is often lacking due to variations in monitoring programs and sampling methodologies, which can affect the robustness of the results .
-
-- While the study emphasizes the importance of incorporating ecosystem variables into stock assessments, it also acknowledges that the models used may not fully capture all relevant interactions within the ecosystem.
+-->
+
+<!-- - While the study emphasizes the importance of incorporating ecosystem variables into stock assessments, it also acknowledges that the models used may not fully capture all relevant interactions within the ecosystem.
 This could lead to biased estimations of krill population dynamics and stock health .
-
-- The paper mentions that spatial heterogeneity was not fully addressed in the modeling approach (implicit approach).
+-->
+
+<!-- - The paper mentions that spatial heterogeneity was not fully addressed in the modeling approach (implicit approach).
 Future research is suggested to incorporate spatially explicit models, which were not utilized in this study.
 This limitation may affect the understanding of how krill populations are distributed and managed across different regions .
-
--  The retrospective analysis conducted in the study indicates patterns of bias in the models.
+-->
+
+<!-- -  The retrospective analysis conducted in the study indicates patterns of bias in the models.
 However, the findings are based on specific scenarios, and the generalizability of these results to other contexts or conditions may be limited .
-
-In summary, while the study provides valuable insights into krill population dynamics, these limitations highlight the need for further research and improved methodologies to enhance the accuracy and applicability of stock assessments in the context of ecosystem-based management.
-
-\newpage
+-->
+
+<!-- In summary, while the study provides valuable insights into krill population dynamics, these limitations highlight the need for further research and improved methodologies to enhance the accuracy and applicability of stock assessments in the context of ecosystem-based management.
+-->
+
+<!-- Problemas de desempeño... -->
+
+<!-- \newpage -->
+
+\pagebreak
 
@@
 However, these models also introduce challenges such as increased sensitivity and retrospective bias, particularly in fishing mortality estimates, underscoring the need for a balance between ecological realism and model stability to ensure reliable management decisions.
-Overall, these findings reinforce that integrated model in an ecosystem-based approach provide a stronger scientific foundation for krill fishery management.
+
+Overall, these findings reinforce that integrated models within an ecosystem-based approach provide a stronger scientific foundation for krill fishery management.
+However, this work represents a preliminary effort that must continue to be refined and strengthened through feedback and collaboration within the Antarctic scientific community.
+Disseminating and discussing these results will be essential for improving ecosystem-based modeling approaches and enhancing their utility in future management strategies.
+
+<!-- Ultimately, the insights gained from this integrated modeling effort contribute to CCAMLR’s ongoing strategy to balance the sustainability of the krill fishery with the conservation of the broader Antarctic marine ecosystem, supporting both the species that depend on krill as a primary food source and the long-term resilience of the Southern Ocean -->
 

### 2025-06-01 index.Rmd
adding word format chgs

@@
   mutate(Model = case_when(
-    Model == "1" ~ "Ref Model: No Env-Predator",
-    Model == "2" ~ "S1.1 w/ Predator data",
-    Model == "3" ~ "S1.1 w/ Env data",
-    Model == "4" ~ "S1.1 w/ Env and Predator data",
+    Model == "1" ~ "s1.1",
+    Model == "2" ~ "s1.2",
+    Model == "3" ~ "s1.3",
+    Model == "4" ~ "s1.4",
     TRUE ~ Model
@@
 
+```{r eval=FALSE, warning=FALSE}
 ## Recruit deviation
-
-```{r, warning=FALSE}
 dev1 <- base.model1.1$recruit[8:35,c(1,7)]%>% 
@@
   geom_line() +
-  geom_point() +
+  geom_point(size = 3, shape = 16) +
   scale_color_manual(values = colores) +
+  scale_x_continuous(
+    breaks = seq(1998, 2020, by = 1)  # etiquetas cada 2 años
+  ) +
   labs(
-    title = "Exploitation Rate Over Time",
-    x = "Year",
+    title = "",
+    x = "",
     y = "Exploitation Rate",
@@
   xlim(1997,2020)+
-  theme_minimal()
+  theme_minimal()+
+  theme(
+    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)  # texto vertical
+  )
 ```

### 2025-06-03 MS2_Krill.Rmd
adding scenarios

@@
 bibliography: SA_krill.bib
-#csl: apa.csl
-csl: icesjournal.csl
+csl: apa.csl
+#csl: icesjournal.csl
 link-citations: yes
@@
 Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024].
-There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@REF].
+There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019].
 Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019].
@@
 Integrating ecological complexity into krill stock assessments presents multiple challenges.
-Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@REF], while some key regions of the WAP lack sufficient data to construct reliable krill population indices.
-Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@REF].
+Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices.
+Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023].
 Despite these challenges, substantial data exist on krill and associated environments.
@@
 
-## Krill Length-to-Age Model Description
+## Length-to-Age Model Description
 
@@
 ft_2
-
 ```
@@
 
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [REF] and SS3 addresses these approaches.
+One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [@Marshall2019] and SS3 addresses these approaches.
 According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment.
@@
 In the case of surveys in Bransfield, this could reflect spatial mismatches between predicted and observed size distributions or biases in survey selectivity for larger size classes.
-This spatial heterogeneity aligns with previous findings identifying this area as a core juvenile krill habitat [@Mardones2025; @Veytia2021, Vetia et al., 2019, REF, REF].
+This spatial heterogeneity aligns with previous findings identifying this area as a core juvenile krill habitat [@Mardones2025; @Veytia2021, @Perry2020: @Perry2019].
 Similarly, the residuals associated with the predator fleet suggest that the model may inadequately represent size-selective predation or the temporal dynamics of predator consumption.
@@
 
-(FINAL TABLE Suggested by MJ)
+<!-- (FINAL TABLE Suggested by MJ) -->
 
@@
 
+<!-- Because M2 is a MGparm, it can be time-varying like any other MGparm. This is -->
+<!-- important because M2, as a component added to base M, will probably always need to be -->
+<!-- time-varying by blocks, random walk or linkage to external driver. (SS3 manual) -->
+
 <!-- @Hollowed2000 demostrate a successful integration of predation mortality into Gulf of Alaska walleye pollock assessments showing that the value of predator data (e.g., diet composition, predator abundance) for refining natural mortality estimates and understanding ecosystem dynamics improve the estimations.
@@
 
-# Data Availability
-
-Data will be made available on request.
+<!-- # Data Availability -->
+
+<!-- Data will be made available on request. -->
 

### 2025-06-10 MS2_Krill.Rmd
SSplotDataEnv()

@@
 author:
-  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl, Instituto Milenio Base]
+  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
   - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
-  - Andrea Pinones ^[Universidad Austral de Chile, Centro IDEAL, Instituto Milenio Base, Chile]
+  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
   - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
-  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Instituto Milenio Base]
-  - César Cárdenas^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
+  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
+  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
   - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
@@
     "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
-    "Abundance and colony-specific trends of penguin populations",
-    "Length-frequency of krill in predator diets (e.g., penguins)"
+    "Abundance by colony-specific trends of penguin populations",
+    "Length-frequency of krill in predator diets"
   ),

### 2025-06-10 index.Rmd
SSplotDataEnv()

@@
 SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013].
-The spurce code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code).
+The source code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code).
 In this exercise, `SS3` is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers.
@@
 ```{r dataserie, fig.width=9, fig.height=7}
-SSplotData(base.model1.4, 
+SSplotDataEnv(base.model1.4, 
            subplots = 1,
-           ptsize = 0.5)
+           ptsize = -2)
 ```
@@
 ```{r}
-outps4 <- base.model1.3$timeseries[1:37, 2:8]
+outps4 <- base.model1.4$timeseries[1:37, 2:8]
 
@@
 alloutput
+
+outpsall %>%
+  filter(Model %in% c("s1.3", "s1.4")) %>%
+  group_by(Model) %>%
+  summarise(
+    n = n(),
+    Bio_all_sum = sum(Bio_all, na.rm = TRUE),
+    Recruit_0_sum = sum(Recruit_0, na.rm = TRUE)
+  )
+
+```
+
+Difference between scenarios regarding to summary in populations variables in Table \@ref(tab:var_compar) for details.
+
+```{r var_compar}
+prom_hist <-outpsall %>%
+  group_by(Model) %>%
+  summarise(
+    Bio_all_hist = mean(Bio_all/1e6, na.rm = TRUE),
+    Recruit_0_hist = mean(Recruit_0/1e6, na.rm = TRUE)
+  )
+
+ult_5 <-outpsall %>%
+  group_by(Model) %>%
+  filter(Yr %in% sort(unique(Yr), decreasing = TRUE)[1:5]) %>%
+  summarise(
+    Bio_all_5yr = mean(Bio_all/1e6, na.rm = TRUE),
+    Recruit_0_5yr = mean(Recruit_0/1e6, na.rm = TRUE)
+  )
+
+resumen <- left_join(prom_hist, ult_5, by = "Model")
+
+base <- resumen %>% filter(Model == "s1.1")
+
+resumen <- resumen %>%
+  mutate(
+    Bio_all_hist_pct = 100 * (Bio_all_hist - base$Bio_all_hist) / base$Bio_all_hist,
+    Recruit_0_hist_pct = 100 * (Recruit_0_hist - base$Recruit_0_hist) / base$Recruit_0_hist,
+    Bio_all_5yr_pct = 100 * (Bio_all_5yr - base$Bio_all_5yr) / base$Bio_all_5yr,
+    Recruit_0_5yr_pct = 100 * (Recruit_0_5yr - base$Recruit_0_5yr) / base$Recruit_0_5yr
+  )
+
+resumen %>%
+  mutate(across(where(is.numeric), ~round(.x, 2))) %>%
+  kbl(caption = "Resumen por modelo: promedios históricos, últimos 5 años y cambio porcentual respecto a s1.1") %>%
+  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
 

### 2025-06-14 MS2_Krill.Rmd
adding last chgs

@@
   ),
-  TimeSeries = c(
+  TemporalScale = c(
     "1997–2020",

### 2025-08-29 MS2_Krill.Rmd
adding last chgs

@@
 title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
+subtitle: "Version 2"
 author:
@@
 output:
-  bookdown::pdf_document2:
+  bookdown::word_document2:
     reference_docx: MS2_Krill_template.docx.docx
@@
 Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables.
-Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics.
-Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses.
+Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary, and arguably more realistic, representation of krill population dynamics.
+Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressing them in peripheral analyses.
 Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations.
 Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
-Such approaches are particularly relevant for ecologically complex systems like the krill fishery in the Western Antarctic Peninsula.
+Such integrative approaches provide a stronger foundation for CCAMLR's objectives, particularly in managing the ecologically complex krill fishery of the Western Antarctic Peninsula.
 
@@
 
-Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021].
+Increasing evidence demonstrates that environmental conditions, such as sea surface temperature, sea ice extent, and phytoplankton availability, strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021].
 Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024].
 There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019].
-Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019].
+Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019; @Robertson2024].
 Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.
@@
 Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices.
-Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023].
+Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023].
 Despite these challenges, substantial data exist on krill and associated environments.
@@
 
-This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations.
-Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
+This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations, which have traditionally been assessed from a data-limited context using model-based approaches.
+A wide variety of integrated models allow the use of diverse data sources, such as spatially explicit approaches including the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
 For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others.
@@
 
-This study aims to integrate various data sources on krill (*Euphausia superba*), including fishery operations (e.g.
+This study aims to integrate various data sources on krill (Euphausia superba), including fishery operations (e.g.
 catches and effort distribution), biological characteristics (e.g.
@@
 environmental forcing and predator-prey interactions) into a unified stock assessment framework.
-Specifically, we use an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
-This approach strengthens the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR's ecosystem-based management approach [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024].
-The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices.
-Our results provide empirical evidence for the value of ecosystem-informed model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in model fit, predictive skill, and structural uncertainty.
+Specifically, we use an integrated length-to-age model, which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
+The primary objective is to evaluate the performance of alternative model configurations that represent different hypotheses about ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments.
+Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence for the value of ecosystem-informed models, showing that incorporating environmental and predator covariates has an impact on model fit, predictive skill, and structural uncertainty, and this has direct implications for sustainable management recommendations and precautionary krill harvesting strategies.
 
@@
 
+# Discussion
+
+This research represents an advance the understanding Antarctic krill (*Euphausia superba*) population dynamics by addressing two fundamental objectives.
+First, it consolidates and harmonizes multiple long-term datasets collected over decades, particularly from Subarea 48.1 in the Western Antarctic Peninsula, into a unified statistical framework, responding to a recurrent management priority about intregrated model in krill emphasized by the CCAMLR.
+Second, it integrates ecosystem variables, such as environmental drivers and predator impacts, thereby moving beyond the traditional single-stocks species approach and taking an initial step toward ecosystem-informed stock assessments.
+This approach not only enhances the biological realism of krill population modelling but also reveals the structural complexity and statistical challenges that must be addressed to ensure its effective application in management Antarctic fishery contexts.
+
+### Model Performance and Diagnostic Evaluation
+
+Increasing model complexity can destabilize internal matrices and processes, potentially leading to convergence issues [@Kinzey2015; @Kinzey2019; @Dong2025].
+While all models in this exercise successfully converged and produced Hessian matrices suitable for quantifying parameter uncertainty, the comprehensive diagnostic evaluation revealed important trade-offs between biological and ecological realism and statistical performance.
+While the ecosystem-informed models (particularly `s1.4`) showed improved biological realism through reduced Pearson residuals and better representation of length-frequency distributions, they also exhibited increased retrospective bias and reduced likelihood values compared to the baseline model (`s1.1`).
+The elevated retrospective bias in fishing mortality estimates for ecosystem scenarios, particularly the pronounced bias (rho = 6.912) in scenario `s1.4`, suggests potential overfitting or structural instability when both environmental and predator variables are simultaneously incorporated.
+The observed bias patterns may also reflect temporal mismatches between ecosystem variables and population responses, suggesting that lagged effects or non-linear relationships should be explored in future model development.
+
+The shift toward ecosystem-based fisheries management highlights the need for stock assessment models that explicitly incorporate species interactions, environmental drivers, and predator pressures, as addressed in this analysis.
+For Antarctic krill, integrating ecosystem variables—such as predation pressure and environmental variability—can improve the realism of population dynamics; however, it also increases model complexity.
+This, in turn, affects model fit and raises parameter uncertainty, revealing a trade-off between traditional single-species assessments and ecosystem-informed approaches, as described by @Collie2016 and @Robertson2024.
+This chapter embraced that challenge and confirmed the inherent difficulties, yet we argue that this approach is a step in the right direction.
+It offers valuable insights for the evolving management strategy that CCAMLR seeks to implement for the krill fishery.
+
+### Spatial Considerations and the Areas-as-Fleets Approach
+
+Building on the analyses from [Chapter 3](#chap3), recognizing and accounting for spatial heterogeneity in krill populations across the large area of Subarea 48.1 was essential, particularly for determining how to incorporate this dimension into the integrated model.
+To address this, we applied the areas-as-fleets approach, a methodology that captures spatial differences and incorporates them into global estimates of population variables.
+This framework allows for a more realistic representation of local variability in abundance, selectivity, and population processes [@Nielsen2021; @Waterhouse2014].
+However, some Management Units (MUs), which served as the structural basis for the areas-as-fleets approach, had limited data—for example, Joinville Island, with few years of survey and mfihery monitoring data.
+While these data limitations were not explicitly evaluated in this thesis, they could introduce complications in future integrated krill modeling efforts and should be carefully considered in subsequent analyses.
+
+### Environmental Drivers and Recruitment Dynamics
+
+The identification of chlorophyll-a concentration as the primary environmental driver affecting krill recruitment (see [Chapter 2](#chap2)) highlights the central role of primary productivity in shaping krill population dynamics [@Ainley2017; @Constable2023; @Saba2014].
+This finding is consistent with recent studies emphasizing that krill recruitment is primarily driven by environmental conditions rather than spawning stock biomass [@ichii2023; @Perry2020; @Veytia2021].
+The research identified chlorophyll-a concentration (Chla) as the most influential environmental predictor for krill recruitment.
+Models incorporating environmental variables (`s1.3` and `s1.4`) showed significantly lower Akaike Information Criterion (AIC) values, indicating better model performance and parsimony compared to models without environmental data.
+While these results highlight that integrating such environmental variables introduces complexity to model structure and performance, it also enhances model realism, a key challenge currently faced by stock assessment approaches.
+By explicitly integrating variables that are mechanistically linked to key population processes—such as recruitment—these models move beyond purely stochastic representations of stock productivity [@Haltuch2009; @Crone2019].
+Such integration allows for a more process informed understanding of population variability, improving both explanatory power and predictive capacity.
+
+### Predator-Prey Interactions and Natural Mortality
+
+The integration of penguin predation data into natural mortality estimates represents a novel approach in an integrated stock assessment, and even more so for krill.
+This size-selective predation pattern is biologically consistent with foraging theory and observed predator behavior, where no significant differences were identified amongkey predators like *Pygoscelis* penguins in the size ranges of krill in their diet [@Panasiuk2024].
+The declining trend in the synoptic predator biomass index, particularly after 2015, reflects documented population declines in Antarctic Peninsula penguin colonies [@Emmerson2022; @Kruger2019; @Salmeron2023; @Trivelpiece2011a].
+These declines, primarily affecting Adélie and Chinstrap penguin populations, have been attributed to climate change impacts, sea ice loss, and shifts in prey availability [@Kruger2019].
+The incorporation of this declining predation pressure into the stock assessment model provides a mechanism for capturing ecosystem mediated changes in natural mortality that would otherwise be overlooked in traditional single-species approaches.
+
+Regarding scenarios of modeling, the scenarios with explicit inclusion of predator components (`s1.2` and `s1.4`), showed increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB), they resulted in lower estimates of recruitment, total biomass, and spawning biomass.
+Incorporating predation mortality into stock assessment models is essential for accurately representing population dynamics in marine ecosystems and offering a more precautionary and arguably more realistic representation of krill dynamics.
+Efforts to incorporate predation as a modelling component in marine stock assessment processes are not new.
+@Hollowed2000 revealed higher total natural mortality rates than traditional life-history estimates and showed that ignoring predation uncertainty underestimated stock biomass uncertainty by up to 20%.
+Similarly, @Richards2016 introduced a predation pressure index for northern shrimp, combining predator biomass and diet frequency to scale natural mortality over time.
+This method improved model fit and reduced retrospective bias without requiring absolute predator consumption data, which are often unavailable.
+For Antarctic krill stock assessments, incorporating predation variability improves mortality estimates and supports ecosystem-based management.
+
+### Implications for Precautionary Management
+
+While there are concerns about the performance of models that consider ecosystem variables, their estimates of key population variables—including spawning biomass, total biomass, and recruitment—are generally lower than those produced by models without ecosystem variables (S1).
+The significant reduction in spawning biomass estimates between scenarios `s1.2` and `s1.4` (mean difference: 2.6 million tonnes) demonstrates that neglecting ecosystem variables may lead to overly optimistic stock assessments and potentially unsustainable harvest recommendations.
+This conservative tendency could have significant implications for fisheries management, as these more precautionary estimates align with CCAMLR's new strategic guidelines, which call for the integration of ecosystem variables into predictive models to support precautionary management approaches.
+The more conservative estimates produced by ecosystem-informed models provide a buffer against uncertainty in population status, which is particularly important given the high variability and unpredictability of Antarctic marine ecosystems [@Hill2016].
+However, the increased model complexity and associated statistical challenges must be balanced against the improved biological realism to ensure that management advice remains scientifically credible and operationally feasible.
+
+### Model Limitations and Future Directions
+
+Several limitations of the current modeling approach warrant discussion.
+First, the integration of predator effects relied on relatively coarse scale abundance indices that may not fully capture fine-scale predator-prey interactions or behavioral responses.
+This is primarily due to two reasons: the large spatial scale of Area 48.1, which covers more than 658,739 $km^{2}$, and the fact that CCAMLR's ecosystem observer deployment does not have observation points distributed throughout this entire area, which hampers comprehensive data collection across the region.
+In this sense, other predators such as whales [@Pallin2023; @Reisinger2022], which have a much greater impact on krill availability and abundance, were not included in this analysis.
+This exclusion was due to the high variability in methods and estimates used to assess whale abundances and consumption rates across time series, particularly in the Western Antarctic Peninsula (WAP), which complicates the integration of these data in a uniform framework.
+Studies suggest that whale consumption rates may even exceed around 20 times the total krill catch [@Biuw2024], highlighting the necessity of incorporating these key predators in future work to more accurately represent predation mortality.
+More detailed predator foraging data, including spatial overlap indices and prey switching behavior, could further enhance this analysis.
+Second, the environmental covariate framework, while effective in capturing chlorophyll driven recruitment variability, represents a simplified view of the complex environmental drivers affecting krill populations.
+Incorporating multivariate ecosystem variables must also be coupled with the development of models capable of handling such complexity; otherwise, contradictory signals may arise, potentially reducing model performance and compromising the reliability of outputs.
+Future studies should consider multivariate environmental relationships, threshold effects, and lagged responses that may better represent the mechanistic links between climate variability and krill population dynamics.
+Third, the current model assumes static relationships between ecosystem variables and parameters estimations.
+However, climate change may alter these relationships over time, requiring adaptive modeling approaches that can accommodate non-stationary ecosystem conditions.
+The development of time-varying parameter frameworks or regime specific models may be necessary to address this challenge.
+
+Despite these limitations, and considering that this is an academic-scientific work, we strongly believe that this represents the future direction for the development of integrated krill models.
+Such approaches provide a foundation to generate scientifically grounded recommendations to CCAMLR, employing the most advanced fisheries science methods currently used worldwide.
+
+\newpage
+
 <!-- One critical issue concerns the dependence on direct biomass estimates, which are typically derived from expensive and logistically demanding scientific surveys—most notably acoustic campaigns.
@@
 
+---
+
+This research demonstrates that integrating ecosystem variables into krill stock assessment models provides important advances in both biological realism and precautionary management advice.
+By harmonizing diverse long-term datasets and embedding environmental and predator covariates within an integrated framework, the work highlights both the potential and the challenges of moving beyond traditional single-species assessments.
+The findings show that ecosystem-informed models have a measurable impact on estimates of recruitment, mortality, and biomass, often producing more conservative outcomes that align with CCAMLR’s precautionary management principles.
+At the same time, the increased model complexity introduces statistical trade-offs and uncertainty that require careful evaluation and transparent communication.
+
+Looking forward, future efforts should refine ecosystem integration by improving predator-prey datasets, exploring multivariate and non-linear environmental relationships, and adopting adaptive modelling frameworks capable of capturing non-stationary dynamics under climate change.
+Despite the inherent challenges, this research supports the view that ecosystem-informed assessment is both necessary and feasible, offering a pathway for CCAMLR to strengthen its ecosystem-based management of the Antarctic krill fishery.
+
+---
+
 \newpage

### 2025-09-16 MS2_Krill.Rmd
adding last edits previous Results section

@@
 The study area covers the WAP, corresponding to FAO Subarea 48.1, one of the main regions for  krill fishing activity.
-This area is also where CCAMLR is actively revising management approach at a finer spatial scale.
-To enhance the spatial resolution of our analysis of krill population dynamics, we used the five Management Units (MUs) (strata) proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
+This area is also where CCAMLR is actively developing a revised management approach at a finer spatial scale.
+To enhance the spatial resolution of our analysis of krill population dynamics, five Management Units (MUs) proposed by CCAMLR (B) were used: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
 1, S1).
-It is worth noting that these strata were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a].
-Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources—whether from fisheries or environmental variables—thus providing a more detailed understanding of krill population dynamics.
+It is worth noting that these MU were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a].
+Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources (whether from fisheries or environmental variables) thus providing a more detailed understanding of krill population dynamics.
 
@@
 Survey indices and biological compositions (e.g., length-frequency data) are assigned to their corresponding MU based on sampling location as defined by [@Dornam2021].
-Similarly, fishery-dependent data—such as catch and size composition—are stratified by MU according to the reported fishing effort, allowing for regional variation in both abundance and observational processes.
+Similarly, fishery-dependent data (such as catch and size composition) are stratified by MU according to the reported fishing effort, allowing for regional variation in both abundance and observational processes.
 Given known difference in the population structure [@Kinzey2011; @Reiss2020; @Cutter2022], the selectivity is modeled separately for each MU to account for potential variations in size composition due to spatial heterogeneity or environmental conditions.
@@
 Given the scarcity—or absence—of direct age data in krill monitoring programs, length-based models offer a practical alternative to age-structured approaches.
-To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015; @Kinzey2019; @Wang2021; @Green2023].
+To date, all model-based krill assessments have used length-to-age frameworks [@Kinzey2015; @Kinzey2019; @Wang2021; @Green2023; @Zhang2022].
 Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Punt2013; @Lee2024].
@@
 $$
-where \( P(A = a \mid L = l) \) is the probability that an individual of length \( l \) belongs to age \( a \),  \( N_{a,l} \) is the number of individuals of age \( a \) observed at length \( l \) and the denominator \( \sum_{a} N_{a,l} \) ensures that the probabilities sum to 1 for each length class.
+where \(P(A = a \mid L = l)\) is the probability that an individual of length *l* belongs to age *a*, $N_{a,l}$ is the number of individuals of age *a* observed at length *l* and the denominator \(\sum_{a} N_{a,l}\) ensures that the probabilities sum to 1 for each length class.
 This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:
@@
 $$
-
-where \( N_l \) is the number of fish observed at length \( l \), and \( N_a \) is the estimated number at age. The resulting ALK is shown in S1 Fig 7.
-
-In the stock assessment catch-at-length model, growth is modeled using a von Bertalanffy growth function lenght-at-age parameterization with parameters L~inf~, k, and coefficient of variation (CV), as outlined in @Mardones2023.
-This growth relationship is integrated into the model alongside weight-at-length, which is assumed known and calculated using the parameters from @Maschette2020 for krill.
-The model assumed linear growth below a predefined settlement age of 1.5 years.
-The asymptotic length ($L_{\infty}$) was determined by the estimated length-at-age, indicating that the model used $L_{\infty}$ as an upper growth boundary.
-Exponential decay of growth beyond maximum age was disabled, ensuring that growth patterns followed the von Bertalanffy assumption without artificial constraints.
-Growth variability was not explicitly modeled, and the coefficient of variation (CV) was assumed to be a function of length-at-age.
-Maturity was modeled as an age-based logistic function, with the first age of maturity set at 2-3 years.
-Fecundity was estimated as a power function of body weight, where fecundity was computed as $Eggs = Wt \times (a + b \times Wt)$.
-The model used a direct parameterization without additional offsets between sexes.
-Regarding selectivity functions, it was modeled using a logistic formulation for the fishery fleet.
-The selectivity curve was defined by two key parameters: the size at 50% selectivity (inflection point) and the width of the selection range (95%).
-The fishery was characterized by multiple selectivity patterns, each defined by an inflection point and a selection range width.
-The estimated inflection points varied between 2.0 and 3.5 cm, while the selection range widths spanned between 2.0 and 4.0 cm depending of strata or "fleet".
-The survey selectivity was also modeled using a length-based logistic function.
-Inflection points ranged from 2.0 to 3.0 cm, with selection range widths consistently set at 1.0 cm.
-The phase assignment dictated whether parameters were freely estimated (positive phase values) or fixed (negative phase values).
-These differences in selectivity parameters are crucial for interpreting and modeling spatial heterogeneity among strata, as they reflect variability in size-dependent catchability across different regions and fleets (S1 Fig 6).
-The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
-assuming that the predator can capture a wider range of krill sizes.
-These parameters were estimated in later phases, allowing flexibility in their estimation.
-Across all selectivity functions, parameters with negative phase values were held fixed during the estimation process, ensuring model stability and preventing overparameterization in a initial step.
-The stock-recruitment (SR) relationship was modeled using a standard Beverton-Holt model curve like @Kinzey2019, with the steepness parameter fixed at 0.85 in phase -4 to ensure consistency with previous studies.
-This reflects a weak stock-recruitment relationship, as krill recruitment largely depends on environmental conditions rather than spawning biomass [@Dong2025; @Kinzey2019; @Perry2020].
-The unfished recruitment ($R_0$) was estimated with a mean recruitment of 23 in log-scale, with a prior of 10.
-The parameter was freely estimated in phase 1.
-The standard deviation of recruitment deviations ($SR\_sigmaR$) was set at 1.2 and fixed in phase -4 to constrain model variability.
-The regime shift parameter ($SR\_regime$) was set at 0 and fixed, implying no consideration of long-term recruitment regime shifts.
-Recruitment deviations were estimated to account for interannual variability in recruitment dynamics.
-The model used the deviation vector approach ($do\_recdev = 1$), where recruitment was modeled as a function of spawning stock biomass plus a deviation term.
-The main recruitment deviation period spanned from 1998 to 2020, with forecast deviations applied in subsequent years.
-Early recruitment deviations prior to 1998 were included, with an early deviation phase of -4 to ensure these values remained fixed.
-Bias adjustment was implemented, with key adjustment years defined as 1995.7 as the last early year with no bias adjustment, 2000.0 as the first year of full bias adjustment, 2015.7 as the last year of full bias adjustment, and 2022.3 as the first recent year with no bias adjustment.
-To prevent overparameterization, recruitment deviations were constrained between -5 and 5.
-Cyclic patterns in recruitment were assumed each tree and five years [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014], and advanced recruitment options were enabled for model flexibility.
-Life history parameters—such as growth, weight-length relationship, natural mortality, and maturity—were used as priors to model the initial conditions of krill population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
-These parameters, along with initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1).
+where $N_{l}$ is the number of fish observed at length *l*, and $N_{a}$ is the estimated number at age. The resulting ALK is shown in S1 Fig. 3.
+
+In the catch-at-length stock assessment model, growth was parameterized using a von Bertalanffy growth function (VBGF) for length-at-age.
+This formulation was integrated with the weight-at-length relationship, which was assumed known and derived from the parameters reported by @Maschette2020 for Antarctic krill.
+To account for early ontogeny, linear growth was assumed below a predefined settlement age of 1.5 years.
+The asymptotic length ($L_{\infty}$) was determined from the estimated length-at-age values, establishing $L_{\infty}$ as the upper boundary for growth.
+Exponential decay of growth beyond the maximum age was disabled, ensuring that growth trajectories followed the VBGF without artificial constraints.
+
+Maturity was modeled as an age-based logistic function, with the onset of maturity occurring between 2 and 3 years.
+Fishery selectivity was represented by a logistic function defined by the size at 50% selectivity ($L_{50}$) and the selection range ($L_{95}$).
+The estimated inflection points ranged from 2.0 to 3.5 cm, while the widths spanned from 2.0 to 4.0 cm across fleets or MU.
+Survey selectivity was also modeled using a logistic formulation, with inflection points between 2.0 and 3.0 cm and a fixed width of 1.0 cm.
+These differences in selectivity parameters are critical for interpreting spatial heterogeneity among MU, as they reflect variability in size-dependent catchability across fleets (S1 Fig 6).
+
+The stock–recruitment (SR) relationship was modeled using a Beverton–Holt function, following @Kinzey2019.
+The steepness parameter was fixed at 0.85 (phase -4) to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020].
+Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 (log-scale), with a prior of 10.
+Recruitment deviations were included to capture interannual variability using the deviation vector approach *recdev = 1*, whereby recruitment is modeled as a function of spawning stock biomass plus a deviation term.
+Cyclic recruitment patterns were also incorporated, assuming periodicity at three and five year intervals [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014].
+
+Finally, life-history parameters—including growth, weight–length relationship, natural mortality, and maturity—were specified as priors to define the initial population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
+These biological parameters, together with the initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1).
 
@@
 Mat 1.
-All analysis about modeling was executed in R-CRAN [@R-base] and diagnoses and visualization of the outputs was made was through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
+All analysis about modeling was executed in R-CRAN [@R-base] where visualization and diagnoses was made was through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
 
@@
 
-To identify the most influential environmental drivers of krill recruitment to be considered in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables described previously as critical in shaping krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chla) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
-Table \@ref(tab:env) summarizes the key characteristics of the environmental datasets used in the analysis, including sea ice, sea surface temperature (SST), and chlorophyll a.
-These variables were obtained from different satellite platforms and reanalysis products, namely Nimbus-7, ERA5, and Copernicus.
-The table details spatial and temporal resolution, as well as the temporal coverage of each dataset.
+To identify the most influential environmental drivers of krill recruitment to be use in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables known to shape krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
+Table \@ref(tab:env) summarizes the main characteristics of the environmental datasets used in this analysis.
+These variables were obtained from a combination of satellite and reanalysis products, specifically *Nimbus-7* [doi.org/10.5067/QOZIVYV3V9JP](https://doi.org/10.5067/QOZIVYV3V9JP), *ERA5* [doi.org/10.48670/moi-00021](https://doi.org/10.48670/moi-00021), and the *E.U.
+Copernicus Marine Service* [doi.org/10.48670/moi-00148](https://doi.org/10.48670/moi-00148).
+The table details their spatial and temporal resolutions, as well as the time ranges covered by each dataset.
+Post-processing techniques, including data cleaning and handling algorithms, manipulation of spatial data, rasterization and projection transformations were facilitated by the *sf* and *ncdf4* packages in R [@Pebezma2018; @Pierce2021].
 
@@
   set_caption("Summary of environmental variables used in correlation analysis") |>
-  add_footer_lines("doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148") |>
   color(part = "header", color = "black") |>
@@
 We modelled the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power.
-The initial model considered only the effect of strata (spatial), while subsequent models sequentially included Chla, SST, SIC and interaction terms.
+The initial model considered only the effect of MU (spatial), while subsequent models sequentially included Chla, SST, SIC and interaction terms.
 This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process.
@@
 $R$ represents the variability of deviations, adding to the variance caused by environmental factors.
-$SpBio$ represent spawning biomass, $envdata$ is vector of time series of environmental variable.
+$SpBio$ represent spawning biomass, $envdata$ is a vector of time series of environmental variables.
 Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
@@
 This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability.
-By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics esttimates.
+By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics estimates.
 
@@
 Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics.
-In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified, located along the northern arc of the Antarctic Peninsula during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)).
-Predator abundance is processed as an index of predation "effort" allowing time-varying estimation of predation mortality rates in time.
-This follows the equation:
+In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)).
+Predator abundance is processed as an index of predation "effort" follows the equation:
 
@@
 
-where \( M2_t \) is the predation mortality in year *t*, \( I_t \) is the predator abundance index, and \( Q \) is a catchability coefficient.
+where \(M2_t\) is the predation mortality in year *t*, \(I_t\) is the predator abundance index, and \( Q \) is a catchability coefficient.
 Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation.
+The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
+assuming that the predator can capture a wider range of krill sizes.
 These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size.
@@
 Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul.
-This ensured consistency across years and strata and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020.
+This ensured consistency across years and MU and fleets (fishery, survey and predator), facilitating reliable temporal and spatial comparisons of size structure, covering the period from 1997 to 2020.
 Fishery-independent data originated from U.S.
@@
     "Length-frequency distributions (0.2 cm bins)",
-    "Biomass indices by strata from net sampling",
+    "Biomass indices by MU from net sampling",
     "Monthly chlorophyll-a concentration from EU Copernicus Marine Service (interpolated; doi:10.48670/moi-00021, doi:10.48670/moi-00148)",
@@
   SpatialScale = c(
-    "By strata",
-    "By strata",
-    "By strata",
-    "By strata",
+    "By MU",
+    "By MU",
+    "By MU",
+    "By MU",
     "Subarea 48.1",
@@
 datos <- data.frame(
-  anio = 2001:2020,
+  anio = 2000:2020,
   variable = c(
-    0.684, 0.693, 0.737, 0.683, 0.716, 0.693, 0.670,
+    0.8, 0.684, 0.693, 0.737, 0.683, 0.716, 0.693, 0.670,
     0.708, 0.653, 0.699, 0.642, 0.647, 0.672, 0.658, 0.694,
@@
 
-min_val <- min(datos$variable)
-max_val <- max(datos$variable)
-datos$indice_ambiental <- 2 * (datos$variable - min_val) / (max_val - min_val) - 1
-
-ggplot(datos, aes(x = anio, y = indice_ambiental)) +
-  geom_col(aes(fill = indice_ambiental > 0), width = 0.5) +
-  scale_fill_manual(values = c("black", "red")) +
-  #geom_line(color = "grey20", size = 1) +
-  geom_hline(yintercept = 0, linetype = "dashed") +
-  labs(title = "",
-       x = "", y = "Env Index") +
-  theme_bw() +
-  theme(legend.position = "none",
-        axis.text.x = element_text(angle = 90, hjust = 2))+
-  scale_x_continuous(breaks = seq(from = 2000, to = 2020, by = 1))
+# estandarización tipo z-score
+datos$indice_ambiental <- scale(datos$variable)[,1]
+
+ggplot(datos, aes(x=anio, y=indice_ambiental))+
+  geom_col(aes(fill=indice_ambiental>0), width=0.5)+
+  scale_fill_manual(values=c("black","red"))+
+  geom_hline(yintercept=0, linetype="dashed")+
+  labs(x="", y="Env Index (z-score)")+
+  theme_bw()+
+  theme(legend.position="none",
+        axis.text.x=element_text(angle=90,hjust=1))+
+  scale_x_continuous(breaks=seq(2000,2020,1))
+
 ```
@@
 
+> suben biomasas como en @Zhang2022
+
 ### Model Limitations and Future Directions
@@
 
--   Supplementary Material 2: Correlation analysis between krill recruitment and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_recruit_env/)
+-   Supplementary Material 2: Mathematical Formulas in SS3 for Krill Stock Assessment (PDF)
+
+-   Supplementary Material 3: Correlation analysis between krill recruitment and environmental variables can be found in this [link](https://mauromardones.github.io/Krill_recruit_env/)
 

### 2025-09-16 Supp_Mat_2.Rmd
adding last edits previous Results section

---
title: "Supplementary Material 2"
subtitle: "Mathematical Formulas in SS3 for Krill Stock Assessment"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: false
geometry: margin=2cm
fontsize: 11pt
header-includes:
  - \usepackage{amsmath}
  - \usepackage{amssymb}
  - \usepackage{bm}
classoption: fleqn
---

\pagebreak

### State Model

- Population Dynamics

\begin{equation}
N_{a,l,s,t+1} = N_{a-1,l,s,t} \cdot e^{-(M_{a-1,l,s,t} + F_{a-1,l,s,t})} \cdot T_{a-1 \rightarrow a,l,s,t} + R_{a,l,s,t}
\end{equation}

- Total Mortality

\begin{equation}
Z_{a,l,s,t} = M_{a,l,s,t} + F_{a,l,s,t}
\end{equation}

- Spawning Stock Biomass

\begin{equation}
SSB_t = \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot W_{a,l,s,t} \cdot \text{Mat}_{a,l,s,t} \cdot \text{Fec}_{a,l,s,t}
\end{equation}

- Total Biomass

\begin{equation}
B_t = \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot W_{a,l,s,t}
\end{equation}

### Sub-models

- Selectivity

\begin{equation}
S_{a,l,f} = \frac{1}{1 + e^{-\ln(19) \cdot \frac{(L_l - L_{50,f})}{(L_{95,f} - L_{50,f})}}}
\end{equation}

\begin{equation}
S_{a,f,t} = \sum_l S_{l,f,t} \cdot P(L_l|a,t)
\end{equation}

- Stock-Recruitment

\begin{equation}
R_t = \frac{\alpha \cdot SSB_{t-\tau}}{\beta + SSB_{t-\tau}} \cdot e^{\delta_t - \frac{\sigma_R^2}{2}}
\end{equation}

- Environmental Effects on Recruitment

\begin{equation}
E(\text{Recruitment}) = f(\text{SSB}) \times \exp(B \times \text{env}) \times \exp\left(-0.5 \times \pi_R^2\right)
\end{equation}

- Maturity

\begin{equation}
\text{Mat}_{l} = \frac{1}{1 + e^{-\ln(19) \cdot \frac{(L_l - L_{50})}{(L_{95} - L_{50})}}}
\end{equation}

\begin{equation}
\text{Mat}_{a,t} = \sum_l \text{Mat}_l \cdot P(L_l|a,t)
\end{equation}

- Growth (Von Bertalanffy)

\begin{equation}
L_{a,s} = L_{\infty,s} \cdot (1 - e^{-K_s \cdot (a - t_{0,s})})
\end{equation}

- Weight-Length Relationship

\begin{equation}
W_{l,s} = w_s \cdot L_l^{b_s}
\end{equation}

- Length-at-Age Variability

\begin{equation}
P(L_l|a,s,t) = \frac{1}{\sqrt{2\pi}\sigma_{L|a,s}} \exp\left(-\frac{(L_l - \bar{L}_{a,s,t})^2}{2\sigma_{L|a,s}^2}\right)
\end{equation}

- Length-to-Age Transition Matrix

\begin{equation}
T_{a,l \rightarrow a+1,l',s,t} = P(L_{l'}|a+1,s,t+1) \cdot \text{Growth}_{l \rightarrow l'}
\end{equation}

- Fishing Mortality

\begin{equation}
F_{a,l,s,f,t} = F'_{f,t} \cdot S_{a,l,s,f,t}
\end{equation}

\begin{equation}
F_{a,l,s,t} = \sum_f F_{a,l,s,f,t}
\end{equation}

- Predator Mortality

\begin{equation}
M_{a,t} = M1_{a,t} + M2_{a,t}
\end{equation}

\begin{equation}
M2_{a,t} = \sum_{p=1}^{n_{\text{pred}}} \frac{C_{p,t} \cdot S_{p,a}}{B_{\text{avail},t}}
\end{equation}

### Catch Equations

\begin{equation}
C_{a,l,s,f,t} = \frac{F_{a,l,s,f,t}}{Z_{a,l,s,t}} \cdot N_{a,l,s,t} \cdot (1 - e^{-Z_{a,l,s,t}}) \cdot W_{a,l,s,t}
\end{equation}

\begin{equation}
C_{f,t} = \sum_a \sum_l \sum_s C_{a,l,s,f,t}
\end{equation}

### Observation Model

\begin{equation}
\hat{I}_{q,t} = Q_q \cdot \sum_a \sum_l \sum_s N_{a,l,s,t} \cdot S_{a,l,s,q,t} \cdot W_{a,l,s,t}
\end{equation}

\begin{equation}
\hat{P}_{a,f,t} = \frac{C_{a,f,t}}{\sum_a C_{a,f,t}}
\end{equation}

\begin{equation}
\hat{P}_{l,f,t} = \frac{\sum_a C_{a,l,f,t}}{\sum_a \sum_l C_{a,l,f,t}}
\end{equation}

### Likelihood Components

\begin{equation}
L_{\text{survey}} = -\sum_q \sum_t \frac{1}{2} \left[ \frac{(\ln(I_{q,t}) - \ln(\hat{I}_{q,t}))^2}{\sigma_{I,q}^2} + \ln(2\pi\sigma_{I,q}^2) \right]
\end{equation}

\begin{equation}
L_{\text{age}} = -\sum_f \sum_t N_{eff,f,t} \sum_a P_{a,f,t} \ln(\hat{P}_{a,f,t})
\end{equation}

\begin{equation}
L_{\text{length}} = -\sum_f \sum_t N_{eff,f,t} \sum_l P_{l,f,t} \ln(\hat{P}_{l,f,t})
\end{equation}

\begin{equation}
L_{\text{rec}} = -\frac{1}{2} \sum_t \left[ \frac{\delta_t^2}{\sigma_R^2} + \ln(2\pi\sigma_R^2) \right]
\end{equation}

### Parameter Estimation

\begin{equation}
-\ln(L_{\text{total}}) = -\sum_i w_i \ln(L_i)
\end{equation}

\begin{equation}
\text{Penalty} = \lambda \sum_j \left( \frac{\theta_j - \theta_{j,\text{prior}}}{\sigma_{\theta,j}} \right)^2
\end{equation}

\pagebreak

### Glossary of Terms and Parameters

| **Symbol** | **Description** | **Units** |
|------------|-----------------|-----------|
| $N_{a,l,s,t}$ | Number of individuals at age $a$, length $l$, sex $s$, time $t$ | Thousands |
| $M_{a,l,s,t}$ | Natural mortality rate | yr$^{-1}$ |
| $F_{a,l,s,t}$ | Fishing mortality rate | yr$^{-1}$ |
| $Z_{a,l,s,t}$ | Total mortality rate | yr$^{-1}$ |
| $R_t$ | Recruitment at time $t$ | Thousands |
| $SSB_t$ | Spawning stock biomass at time $t$ | Metric tons |
| $B_t$ | Total biomass at time $t$ | Metric tons |
| $W_{a,l,s,t}$ | Weight-at-age-length-sex | Grams |
| $\text{Mat}_{a,l,s,t}$ | Proportion mature at age-length-sex | Proportion |
| $\text{Fec}_{a,l,s,t}$ | Fecundity-at-age-length-sex | Eggs/individual |
| $S_{a,l,f}$ | Selectivity for fleet $f$ at age $a$, length $l$ | Proportion |
| $L_{\infty,s}$ | Asymptotic length for sex $s$ | mm |
| $K_s$ | Growth rate parameter for sex $s$ | yr$^{-1}$ |
| $t_{0,s}$ | Theoretical age at length zero for sex $s$ | Years |
| $w_s$ | Weight-length scaling parameter for sex $s$ | g/mm$^b$ |
| $b_s$ | Weight-length allometric exponent for sex $s$ | Dimensionless |
| $L_{50}$ | Length at 50% maturity | mm |
| $L_{95}$ | Length at 95% maturity | mm |
| $L_{50,f}$ | Length at 50% selectivity for fleet $f$ | mm |
| $L_{95,f}$ | Length at 95% selectivity for fleet $f$ | mm |
| $\alpha$ | Stock-recruitment steepness parameter | Various |
| $\beta$ | Stock-recruitment capacity parameter | Various |
| $\sigma_R$ | Recruitment variability | Dimensionless |
| $\delta_t$ | Recruitment deviation in year $t$ | Log-scale |
| $Q_q$ | Catchability coefficient for survey $q$ | Various |
| $C_{a,l,s,f,t}$ | Catch at age-length-sex by fleet $f$ at time $t$ | Metric tons |
| $I_{q,t}$ | Observed survey index for survey $q$ at time $t$ | Various |
| $\hat{I}_{q,t}$ | Expected survey index for survey $q$ at time $t$ | Various |
| $P_{a,f,t}$ | Observed age composition for fleet $f$ at time $t$ | Proportion |
| $\hat{P}_{a,f,t}$ | Expected age composition for fleet $f$ at time $t$ | Proportion |
| $P_{l,f,t}$ | Observed length composition for fleet $f$ at time $t$ | Proportion |
| $\hat{P}_{l,f,t}$ | Expected length composition for fleet $f$ at time $t$ | Proportion |
| $N_{eff,f,t}$ | Effective sample size for fleet $f$ at time $t$ | Dimensionless |
| $\sigma_{I,q}$ | Standard deviation of survey index $q$ | Log-scale |
| $M1_{a,t}$ | Base natural mortality at age $a$, time $t$ | yr$^{-1}$ |
| $M2_{a,t}$ | Predation mortality at age $a$, time $t$ | yr$^{-1}$ |
| $C_{p,t}$ | Consumption by predator $p$ at time $t$ | Metric tons |
| $S_{p,a}$ | Age-specific selectivity by predator $p$ | Proportion |
| $B_{\text{avail},t}$ | Available biomass for predation at time $t$ | Metric tons |
| $\beta_i$ | Environmental coefficient for covariate $i$ | Various |
| $X_{i,t}$ | Environmental covariate $i$ at time $t$ | Various |
| $\lambda$ | Penalty weight | Dimensionless |
| $w_i$ | Likelihood component weight | Dimensionless |
| $\theta_j$ | Estimated parameter $j$ | Various |
| $\theta_{j,\text{prior}}$ | Prior mean for parameter $j$ | Various |
| $\sigma_{\theta,j}$ | Prior standard deviation for parameter $j$ | Various |
| $\tau$ | Recruitment lag (spawning to recruitment age) | Years |
| $F'_{f,t}$ | Fully-selected fishing mortality for fleet $f$ | yr$^{-1}$ |
| $R_0$ | Unfished recruitment level | Thousands |
| $h$ | Steepness parameter in stock-recruitment | Proportion |

\pagebreak

### References

Methot, R.D. and Wetzel, C.R. (2013). Stock Synthesis: A biological and statistical framework for fish stock assessment and fishery management. *Fisheries Research*, 142: 86-99.

NMFS-OST (2025). Stock Synthesis User Manual Version 3.30.23.2. Available at: https://nmfs-ost.github.io/ss3-doc/SS330_User_Manual_release.html

### 2025-09-16 index.Rmd
adding last edits previous Results section

@@
 # install.packages("r4ss")
-library(r4ss)
-library(here)
-#remotes::install_github("PIFSCstockassessments/ss3diags")
-library(ss3diags)
-library(kableExtra)
+pkgs<-c("r4ss","ss3diags","doParallel",
+        "tibble","tidyr","tidyverse",
+        "readxl","openxlsx","broom",
+        "forecast","mixR","lmtest",
+        "car","ggpubr","ggthemes",
+        "ggridges","ggrepel",
+        "kableExtra","flextable","here")
+
+instalar<-pkgs[!pkgs%in%installed.packages()]
+if(length(instalar)>0)install.packages(instalar)
+
+invisible(lapply(pkgs,library,character.only=TRUE))
+
 library(doParallel)
-detectCores()
 registerDoParallel(8)
-library(ggpubr)
-library(tibble)
-library(openxlsx)
-library(ggthemes)
-library(forecast)
-library(tidyr)
-library(mixR)
-library(readxl)
-library(tidyverse)
-library(ggridges)
-library(ggrepel)
-library(broom)
-library(lmtest)
-library(flextable)
-library(car)
 ```
@@
 
-### Length comositions
+### Length compositions
 

### 2025-12-16 MS2_Krill.Rmd
fix abstract

@@
 
-Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula.
-In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status.
-We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species.
-Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables.
-Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary, and arguably more realistic, representation of krill population dynamics.
-Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressing them in peripheral analyses.
-Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations.
-Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use.
-Such integrative approaches provide a stronger foundation for CCAMLR's objectives, particularly in managing the ecologically complex krill fishery of the Western Antarctic Peninsula.
+Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is fundamental for ecosystem-based fisheries management in the Western Antarctic Peninsula, where environmental variability and predator–prey interactions are known to strongly influence population productivity.
+However, most stock assessment frameworks applied to krill and other exploited marine resources are based on integrated models in which ecosystem processes are not explicitly embedded within the core population dynamics.
+In particular, environmental variability and predation pressure are seldom incorporated directly into the state equations governing recruitment, natural mortality, and biomass dynamics in integrated assessment models.
+Instead, these processes are typically explored through external sensitivity analyses or post hoc diagnostics, thereby limiting their influence on estimated stock status and management advice.
+In this study, an integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics.
+Within a Stock Synthesis–based framework, alternative model configurations are evaluated in which environmental covariates and predator indices are either explicitly linked to key state processes (recruitment and natural mortality) or excluded from the structural equations.
+This design allows a direct comparison of their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality.
+Systematic and coherent differences are observed among model configurations.
+When environmental variability and predator effects are embedded within the state equations, lower estimates of recruitment, total biomass, and spawning biomass are obtained, together with altered estimates of fishing mortality, relative to models in which ecosystem drivers are ignored.
+In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, less stable Monod-type parameter estimates, and more structured residual patterns.
+This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional ecological processes are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves.
+Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecological realism.
+Despite the challenges associated with integrating ecosystem information into fully integrated assessment models, it is demonstrated that environmental and predator effects can be consistently incorporated within the core population dynamics of krill assessments.
+The direct embedding of these drivers in the state equations provides a more precautionary and ecologically realistic representation of krill population dynamics and strengthens the scientific basis for management advice consistent with CCAMLR’s objectives for conservation and sustainable use in the Western Antarctic Peninsula.
 
@@
 The steepness parameter was fixed at 0.85 (phase -4) to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020].
-Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 (log-scale), with a prior of 10.
+Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 mm.
+(log-scale), with a prior of 10.
 Recruitment deviations were included to capture interannual variability using the deviation vector approach *recdev = 1*, whereby recruitment is modeled as a function of spawning stock biomass plus a deviation term.
@@
 Following @methot2020stock, natural mortality due to predation (*M2*) was incorporate through different types of observational data types, with each informing distinct aspects of predator-prey dynamics.
-In our krill assessment, we incorporated predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)).
+In our krill assessment, we incorporated annual predator abundance from the three main penguin species in the Antarctic Peninsula: Adélie (*Pygoscelis adeliae*), Chinstrap (*Pygoscelis antarcticus*), and Gentoo (*Pygoscelis papua*)  from 399 colonies identified during the last two decades (2000–2020) obtained from the Mapping Application for Penguin Populations and Projected Dynamics ([MAPPPD](https://www.penguinmap.com/mapppd/sources/)).
 Predator abundance is processed as an index of predation "effort" follows the equation:
@@
 Additionally, length-frequency data of krill obtained from penguin diet samples from AMLR Program are used to characterize the size-selectivity of predation.
-The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.2 cm and a corresponding selection range width of 0.2 cm.
-assuming that the predator can capture a wider range of krill sizes.
+The predator fleet followed a similar selectivity function, with a size-at-inflection point of 0.5 cm and a corresponding selection range width of 3.5 cm assuming that the predator can capture a wider range of krill sizes.
 These data inform predator-specific selectivity functions for *M2*, enabling the model to represent how predation varies with krill size.
@@
 Total mortality is expressed as \( Z = M_1 + M_2 + F \) combining natural and fishing-induced mortality.
-This dual approach—using predator abundance and diet composition—supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.
+This dual approach (using predator abundance and diet composition) supports a temporally and size-structured representation of predation mortality, aligning with ecosystem-based assessment principles.
 
@@
 Along with data on predators and environmental conditions, a fundamental input to the model was derived from fishery-dependent data and scientific monitoring programs.
-Fishery-dependent data were obtained from the Krill Fishery Monitoring Program under the CCAMLR SISO framework, including total reported catches, standardized CPUE (adjusted by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm.
+Fishery-dependent data were obtained from the CCAMLR Krill Fishery Monitoring Program and SISO Scheme, including total catches in weight, effort, (agruped by vessel, area, and depth), and krill length-frequency distributions data were standardized using a fixed bin structure of 28 length intervals, each representing the lower edge of 0.2 cm width bins from 1.2 to 6.6 cm.
 Length data were assigned to partition code 0 (combined discard and retained) and sex code 0 (combined sexes), with a fixed sample size of 75 individuals per haul.
@@
     booktabs = TRUE, 
-    caption = "Table: Data sources used in the krill population dynamics modeling") %>%
+    caption = "Data sources used in the krill population dynamics modeling") %>%
   kable_styling(latex_options = c("hold_position", "striped"), 

### 2025-12-16 index.Rmd
fix abstract

@@
 circleplot_res
+
+# 1. Calcular pendiente y p-value por Fleet y Modelo
+lm_stats <- df_all %>%
+  group_by(Model, Fleet) %>%
+  do({
+    mod <- lm(Pearson ~ Yr, data = .)
+    out <- tidy(mod)
+    slope <- out$estimate[out$term == "Yr"]
+    pval <- out$p.value[out$term == "Yr"]
+    data.frame(slope = round(slope, 3),
+               pval = round(pval, 3))
+  }) %>%
+  ungroup()
+
+# 2. Plot con nubes de puntos y línea de tendencia
+ggplot(df_all, aes(x = Yr, y = Pearson)) +
+  geom_point(alpha = 0.5, size = 0.7, color = "grey30") +
+  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 0.6) +
+  facet_grid(Model ~ Fleet) +
+  theme_bw(base_size = 9) +
+  labs(x = "Year", y = "Pearson residuals") +
+  theme(strip.text = element_text(size = 7)) +
+  # 3. Añadir pendiente y p-value dentro de cada panel
+  geom_text(data = lm_stats,
+            aes(x = min(df_all$Yr) + 2,
+                y = max(df_all$Pearson, na.rm = TRUE) - 0.3,
+                label = paste0("slope=", slope, "\np=", pval)),
+            inherit.aes = FALSE,
+            size = 2.5, hjust = 0)
+
 ```
@@
 ```
+
+```{r}
+# Lista de nombres de modelos
+model_names <- paste0("base.model1.", 1:4)
+
+# Bucle para calcular Pearson residual
+for (mn in model_names) {
+  # Obtener el objeto del modelo
+  mod <- get(mn)
+
+  # Calcular residual de Pearson
+  mod$cpue$PearsonRes <- with(mod$cpue, (Obs - Exp) / SE_input)
+
+  # Guardar de vuelta en el entorno global
+  assign(mn, mod)
+}
+
+# Revisar el resultado, ejemplo para el primero
+head(base.model1.1$cpue[, c("Obs", "Exp", "SE_input", "PearsonRes")])
+
+library(ggplot2)
+library(dplyr)
+
+# Lista de modelos
+model_names <- paste0("base.model1.", 1:4)
+
+# Crear un data.frame combinado con todos los modelos
+cpue_all <- bind_rows(
+  lapply(model_names, function(mn) {
+    mod <- get(mn)
+    df <- mod$cpue
+    df$Model <- mn  # añadir columna con el nombre del modelo
+    return(df)
+  })
+)
+
+# Plot de residuos de Pearson por año, por flota, facetado por modelo
+ggplot(cpue_all, aes(x = Yr, y = PearsonRes)) +
+  geom_point(alpha = 0.6) +
+  facet_grid(Model ~ Fleet_name) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  theme_minimal() +
+  labs(
+    title = "Residuos de Pearson por modelo y flota",
+    x = "Año",
+    y = "Residual de Pearson"
+  )+
+  ylim( -10,10)
+```
+
 ### Residual consistency

### 2025-12-18 MS2_Krill.Rmd
improve intro

@@
 ---
-title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
+title: "Integrating Environmental Forcing and Predator Effects into Antarctic Krill (*Euphausia superba*) Population Dynamics Using a Length-to-Age Stock Assessment Framework"
 subtitle: "Version 2"
@@
   bookdown::word_document2:
-    reference_docx: MS2_Krill_template.docx.docx
+    reference_docx: MS2_Krill_template.docx
     number_sections: false
@@
 Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is fundamental for ecosystem-based fisheries management in the Western Antarctic Peninsula, where environmental variability and predator–prey interactions are known to strongly influence population productivity.
-However, most stock assessment frameworks applied to krill and other exploited marine resources are based on integrated models in which ecosystem processes are not explicitly embedded within the core population dynamics.
-In particular, environmental variability and predation pressure are seldom incorporated directly into the state equations governing recruitment, natural mortality, and biomass dynamics in integrated assessment models.
-Instead, these processes are typically explored through external sensitivity analyses or post hoc diagnostics, thereby limiting their influence on estimated stock status and management advice.
-In this study, an integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics.
+However, most stock assessment frameworks applied to krill and other exploited marine resources are based on model-based approaches in which ecosystem drivers are not explicitly embedded within the core population dynamics.
+These drivers are typically explored through external sensitivity analyses or post-hoc diagnostics, thereby limiting their influence on estimated stock status and management advice.
+In this study, a spatial implicit integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics.
 Within a Stock Synthesis–based framework, alternative model configurations are evaluated in which environmental covariates and predator indices are either explicitly linked to key state processes (recruitment and natural mortality) or excluded from the structural equations.
-This design allows a direct comparison of their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality.
+This design allows assess their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality.
 Systematic and coherent differences are observed among model configurations.
 When environmental variability and predator effects are embedded within the state equations, lower estimates of recruitment, total biomass, and spawning biomass are obtained, together with altered estimates of fishing mortality, relative to models in which ecosystem drivers are ignored.
-In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, less stable Monod-type parameter estimates, and more structured residual patterns.
-This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional ecological processes are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves.
-Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecological realism.
+In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, retrospective paterns and more structured residual patterns.
+This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional components are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves.
+Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecosystem realism.
 Despite the challenges associated with integrating ecosystem information into fully integrated assessment models, it is demonstrated that environmental and predator effects can be consistently incorporated within the core population dynamics of krill assessments.
@@
 
-Antarctic krill (*Euphausia superba*, hereafter krill) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024].
+Antarctic krill (*Euphausia superba*, hereafter krill) is the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024].
 Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a].
@@
 Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries.
-Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023].
+Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023; @Zhao2023; @Kawaguchi2024].
 This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales.
@@
 Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices.
-Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023].
-Despite these challenges, substantial data exist on krill and associated environments.
-Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers.
+Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions, which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses in krill dynamic, and the interactions of multiple stressors [@Constable2023].
+Despite these challenges, substantial data exist on krill and their associated environments.
+Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers, notably through CCAMLR’s Scheme of International Scientific Observation (SISO), which ensures standardized and independent biological sampling of the krill fishery across the Convention Area.
+Additionally, Multiple biomass assessment surveys based on hydroacoustic cruises are also conducted by CCAMLR, primarily led by countries such as the United States and China, whose programs represent the longest-running efforts and the broadest spatial coverage of krill hydroacoustic assessments in the Southern Ocean [@Hill2019; @Dornam2021; @Green2023].
 Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites.
 Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.
-
-This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations, which have traditionally been assessed from a data-limited context using model-based approaches.
-A wide variety of integrated models allow the use of diverse data sources, such as spatially explicit approaches including the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
+This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations.
+
+CCAMLR has traditionally relied on assessments conducted within a data-limited context using relatively simple model-based approaches, such as GRYM-type models, which do not explicitly incorporate environmental or ecological variables in the representation of population dynamics.
+In contrast, a wide range of integrated assessment frameworks now allow the incorporation of diverse data sources and increasing levels of structural complexity.
+These include spatially explicit approaches such as the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011].
 For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others.
-These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].
-
-This study aims to integrate various data sources on krill (Euphausia superba), including fishery operations (e.g.
-catches and effort distribution), biological characteristics (e.g.
-growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g.
-environmental forcing and predator-prey interactions) into a unified stock assessment framework.
-Specifically, we use an integrated length-to-age model, which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration.
-The primary objective is to evaluate the performance of alternative model configurations that represent different hypotheses about ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments.
-Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence for the value of ecosystem-informed models, showing that incorporating environmental and predator covariates has an impact on model fit, predictive skill, and structural uncertainty, and this has direct implications for sustainable management recommendations and precautionary krill harvesting strategies.
+Notably, Stock Synthesis can also be applied to species that are difficult to age, such as Antarctic krill, through integrated length-to-age transformation models that explicitly link observed length compositions to underlying age-structured population dynamics.
+These advanced tools provide robust and flexible methodologies to address data limitations while accommodating complex biological processes and ecosystem interactions, thereby enabling a more realistic representation of population dynamics and improved scientific support for ecosystem-based fisheries management [@Hoyle2020].
+
+This study has two main objectives.
+The first is to integrate multiple data sources on Antarctic krill (*Euphausia superba*), including fishery operations (e.g., catches and effort distribution), biological characteristics (e.g., growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g., environmental forcing and predator–prey interactions), into a unified stock assessment framework.
+Specifically, we apply an integrated length-to-age model that enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and overall productivity, as well as the potential influence of predator populations through the configuration of natural mortality.
+The second objective is to evaluate the performance of alternative model configurations representing different hypotheses regarding ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments.
+Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence of the value of ecosystem-informed assessment models, demonstrating that the incorporation of environmental and predator covariates influences model fit, predictive skill, and structural uncertainty, with direct implications for precautionary and sustainable management advice for the Antarctic krill fishery.
 
@@
 
-One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [@Marshall2019] and SS3 addresses these approaches.
-According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment.
-This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;
-
-$$
-E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
-$$
-$R$ represents the variability of deviations, adding to the variance caused by environmental factors.
-$SpBio$ represent spawning biomass, $envdata$ is a vector of time series of environmental variables.
-Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced.
-However, the model does not automatically adjust for this.
-Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment.
-This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability.
-By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics estimates.
+To identify environmental drivers suitable for explicit inclusion in the stock assessment model, we evaluated the relationship between a krill recruitment index and three environmental variables known to influence krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
+These variables represent thermal, habitat, and trophic controls acting on early life stages of Antarctic krill.
+
+Environmental data were obtained from a combination of satellite observations and reanalysis products (*Nimbus-7*, *ERA5*, and the EU Copernicus Marine Service).
+A summary of the spatial and temporal characteristics of each dataset is provided in Table @ref(tab:env).
+All variables were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], including data cleaning, spatial manipulation, rasterization, and coordinate transformation.
+
+Prior to integration into the stock assessment model, environmental variables were statistically screened using a set of linear regression models with recruitment as the response variable.
+The baseline model accounted for spatial structure through management unit (MU), while subsequent models progressively incorporated Chl-a, SST, SIC, and selected interaction terms.
+This stepwise approach was used to identify the environmental variable exhibiting the strongest and most consistent statistical relationship with recruitment.
+Results of these screening analyses are provided in Supplementary Material 2.
+
+### Environmental link formulation in Stock Synthesis
+
+Stock Synthesis version 3.30.21 allows environmental covariates to be incorporated directly into the population dynamics through the *env-link* option in the control file, whereby external time series modify biological processes in a deterministic and multiplicative manner [@methot2020stock].
+
+In this study, the selected environmental variable was linked to recruitment via an env-link acting on the recruitment deviations. Expected recruitment in year (t) was modeled as:
+
+$$
+E(R_t) = f(SSB_t) \times \exp(\beta_E E_t) \times \exp\left(-\tfrac{1}{2}\sigma_R^2\right),
+$$
+
+where ( f(SSB_t) ) represents the stock–recruitment relationship, assumed to follow a Beverton–Holt formulation; ( E_t ) corresponds to the standardized environmental time series provided in the Stock Synthesis data file; ( \beta_E ) denotes the environment–recruitment link coefficient estimated within the control file; and ( \sigma_R^2 ) is the variance associated with recruitment deviations.
+
+On the log scale, this formulation is equivalent to:
+
+$$
+\log(R_t) = \log(f(SSB_t)) + \beta_E E_t + \eta_t,
+\quad \eta_t \sim \mathcal{N}(0,\sigma_R^2).
+$$
+
+The environmental time series (E_t) was standardized prior to model fitting, such that (\beta_E) represents the change in log-recruitment associated with a one-standard-deviation change in the environmental variable.
+
+The env-link coefficient (\beta_E) quantifies the sensitivity of recruitment to environmental variability:
+
+$$
+\beta_E = \frac{\partial \log(R_t)}{\partial E_t}.
+$$
+
+Positive values of (\beta_E) indicate environmental conditions that enhance recruitment, while negative values imply adverse effects.
+Because the env-link operates multiplicatively, environmental effects scale recruitment exponentially rather than additively.
+
+By incorporating environmental forcing through the env-link, part of the interannual variability in recruitment that would otherwise be attributed to unexplained recruitment deviations is instead explained by the environmental covariate.
+Consequently, improved model fit is expected to manifest as a reduction in the estimated recruitment deviation variance (\sigma_R^2), although this reduction is not imposed a priori and arises through likelihood optimization.
+
+The inclusion of the selected environmental variable via the env-link framework provides a biologically interpretable and statistically coherent mechanism to integrate ecosystem drivers into the core population dynamics of the krill stock assessment.
 

### 2025-12-19 MS2_Krill.Rmd
edit env link

@@
 The second objective is to evaluate the performance of alternative model configurations representing different hypotheses regarding ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments.
-Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence of the value of ecosystem-informed assessment models, demonstrating that the incorporation of environmental and predator covariates influences model fit, predictive skill, and structural uncertainty, with direct implications for precautionary and sustainable management advice for the Antarctic krill fishery.
+Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence of the value of ecosystem-informed assessment models, demonstrating that the incorporation of environmental and predator covariates influences population variables, model fit, predictive skill, and structural uncertainty, with implications for precautionary and sustainable management advice for the Antarctic krill fishery.
 
@@
 
-The study area covers the WAP, corresponding to FAO Subarea 48.1, one of the main regions for  krill fishing activity.
+The study area covers the WAP, corresponding to FAO Statistical Subarea 48.1, one of the main regions for  krill fishing activity.
 This area is also where CCAMLR is actively developing a revised management approach at a finer spatial scale.
-To enhance the spatial resolution of our analysis of krill population dynamics, five Management Units (MUs) proposed by CCAMLR (B) were used: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (Fig.
-1, S1).
-It is worth noting that these MU were originally developed for management purposes, without explicitly considering biological, ecological or population-specific criteria [@Watters2013; @Constable2023; @Zaldua2024a].
-Nevertheless, their subdivision enables a finer-scale analysis of krill data, increasing spatial resolution and enabling detection of heterogeneity among various data sources (whether from fisheries or environmental variables) thus providing a more detailed understanding of krill population dynamics.
-
-## Spatially Implicit Model
-
-To represent spatial heterogeneity in krill dynamics, we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* methodology [@Waterhouse2014; @Punt2019; @Nielsen2021].
+To enhance the spatial resolution of our analysis of krill population dynamics, five Management Units (MUs) or stratas proposed by CCAMLR were used: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest sector [@Dornam2021] (S1 Fig.
+1).
+Although these management units were not originally designed to incorporate krill biological or population-specific characteristics [@Watters2013; @Constable2023; @Zaldua2024a], their subdivision allows for finer-scale examination of krill data, enhancing spatial resolution and facilitating the identification of heterogeneity among different data sources, including fishery-dependent and environmental information, thereby supporting a more detailed characterization of krill population dynamics.
+
+## Spatial implicit model
+
+To represent spatial heterogeneity in krill dynamics, we implemented a spatially implicit modeling approach based on the *"areas-as-fleets"* [@Waterhouse2014; @Punt2019; @Nielsen2021].
 This approach treats distinct geographical regions as independent fleets, under the assumption that fishing and survey operations may have varying effects across space.
 Within this approach, each MU of the WAP is treated as a distinct fleet, allowing spatial differentiation in key parameters such as selectivity, catchability (*q*) and abundance indices.
-This is especially relevant for krill due to documented spatial variability and heterogeneity in both population processes and fishing effort [@Perry2019; @Perry2020; @Reiss2020; @Veytia2021; @SantaCruz2022].
+This is especially relevant for krill due to documented spatial variability and heterogeneity in both population processes and fishing effort [@Perry2019; @Perry2020; @Reiss2020; @Veytia2021; @SantaCruz2022; @Mardones2026].
 By modeling each MU as a separate fleet, the assessment can estimate region-specific parameters that reflect this variability.
@@
 
-## Length-to-Age Model Description
+## Length-to-Age Model description
 
@@
 This method assumes both aged and measured individuals are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024].
-The ALK matrix is typically defined as follows:
-
-$$
-P(A = a \mid L = l) = \frac{N_{a,l}}{\sum_{a} N_{a,l}}
-$$
-where \(P(A = a \mid L = l)\) is the probability that an individual of length *l* belongs to age *a*, $N_{a,l}$ is the number of individuals of age *a* observed at length *l* and the denominator \(\sum_{a} N_{a,l}\) ensures that the probabilities sum to 1 for each length class.
-This matrix is then used in the model to allocate observed length compositions into age classes, typically through multiplication with a length-frequency vector:
-
-$$
-N_a = \sum_{l} P(A = a \mid L = l) \cdot N_l
-$$
-where $N_{l}$ is the number of fish observed at length *l*, and $N_{a}$ is the estimated number at age. The resulting ALK is shown in S1 Fig. 3.
-
+Using this statistical framework, we constructed the age–length key (ALK) for Antarctic krill, which was subsequently applied within the assessment model (S1 Fig.
+3).
 In the catch-at-length stock assessment model, growth was parameterized using a von Bertalanffy growth function (VBGF) for length-at-age.
@@
 To account for early ontogeny, linear growth was assumed below a predefined settlement age of 1.5 years.
-The asymptotic length ($L_{\infty}$) was determined from the estimated length-at-age values, establishing $L_{\infty}$ as the upper boundary for growth.
-Exponential decay of growth beyond the maximum age was disabled, ensuring that growth trajectories followed the VBGF without artificial constraints.
-
-Maturity was modeled as an age-based logistic function, with the onset of maturity occurring between 2 and 3 years.
+Maturity was modeled as an age-based logistic function, with the onset of maturity occurring between 2 and 3 years [@Thanassekos2014].
 Fishery selectivity was represented by a logistic function defined by the size at 50% selectivity ($L_{50}$) and the selection range ($L_{95}$).
 The estimated inflection points ranged from 2.0 to 3.5 cm, while the widths spanned from 2.0 to 4.0 cm across fleets or MU.
-Survey selectivity was also modeled using a logistic formulation, with inflection points between 2.0 and 3.0 cm and a fixed width of 1.0 cm.
-These differences in selectivity parameters are critical for interpreting spatial heterogeneity among MU, as they reflect variability in size-dependent catchability across fleets (S1 Fig 6).
-
+Survey selectivity was also modeled using a logistic formulation, with inflection points between 2.0 and 3.0 cm.
+These differences in selectivity parameters are critical for interpreting spatial heterogeneity among MU, as they reflect variability in size-dependent catchability across fleets (S1 Fig 5).
 The stock–recruitment (SR) relationship was modeled using a Beverton–Holt function, following @Kinzey2019.
-The steepness parameter was fixed at 0.85 (phase -4) to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020].
+The steepness parameter was fixed at 0.85 to maintain consistency with previous studies, reflecting the weak relationship between spawning biomass and recruitment in krill populations, where recruitment is largely driven by environmental variability [@Dong2025; @Kinzey2019; @Perry2020].
 Unfished recruitment ($R_0$) was estimated assuming a mean recruitment of 23 mm.
@@
 Cyclic recruitment patterns were also incorporated, assuming periodicity at three and five year intervals [@Perry2019; @McBride2021; @Kinzey2019; @Thanassekos2014].
-
-Finally, life-history parameters—including growth, weight–length relationship, natural mortality, and maturity—were specified as priors to define the initial population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
-These biological parameters, together with the initial selectivity assumptions, are summarized in Table \@ref(tab:parainit1).
-
-```{r parainit1, eval=TRUE}
-dir1.1 <- here("s1.1")
-
-start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
-dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
-ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
-                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
-fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)
-
-# Extraer parámetros
-parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
-rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
-                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")
-
-SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
-Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
-Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]
-
-# Combinar parámetros
-parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))
-
-# Añadir nombres de fila como columna "Parameter"
-parInit$Parameter <- rownames(parInit)
-
-# Añadir categorías y reordenar columnas
-parInit <- parInit %>%
-  mutate(INIT  = round(INIT, 2),
-         PRIOR = round(PRIOR, 2))
-
-# Crear vector de categorías (39 filas)
-categories <- c(rep("Natural Mortality", 1),
-                rep("Growth", 4),
-                rep("Length-Weight Relation", 2),
-                rep("Maturity", 2),
-                rep("Stock-Recruit Relation", 5),
-                rep("Catchability", 2),
-                rep("Selectivity", 21))  # OJO: corregido a 23
-
-# Añadir categorías y reordenar columnas
-parInit <- parInit %>%
-  mutate(Category = categories) %>%
-  dplyr::relocate(Category, Parameter)
-
-# Crear flextable
-ft <- flextable(parInit)
-ft <- set_caption(ft, "Input parameters for the initial SS3 krill model")
-ft <- autofit(ft)
-ft <- theme_booktabs(ft)
-ft <- set_table_properties(ft, width = .5, layout = "autofit")
-ft <- fontsize(ft, size = 9, part = "all")
-ft <- add_header_lines(ft, values = "Each parameter includes limits (LO-HI), initial value (INIT), and phase. Negative phase means the parameter is fixed.")
-ft
-```
+All these life-history parameters and assumptions were specified as priors to define the initial population dynamics, based on data from @Smith2023a, @Maschette2020, and @Kinzey2011.
+These biological parameters, together with the initial selectivity assumptions, are summarized in Table 1 S1.
 
 The statistical model to krill was configured using Stock Synthesis v 3.30.21 (SS3 hereafter) [@Methot2013].
-SS3 is a structured stock assessment model, in the class of models called *"Integrated stock assessment model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
+SS3 is a structured stock assessment model, in the class of models called "Integrated stock assessment model" and has a set of sub-model that simulates growth, maturity, fecundity, recruitment, and mortality processes, and observation, with expected values for different types of data.
 The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters.
-Model conditioning is available in the code repository section of Suppl.
-Mat 1.
+Model conditioning is available in the code repository section of Supporting Information 1.
 All analysis about modeling was executed in R-CRAN [@R-base] where visualization and diagnoses was made was through *r4ss* [@Taylor2019] and *ss3diags* packages [@Winker2023].
 
-## Environmental variable modelling
-
-To identify the most influential environmental drivers of krill recruitment to be use in stock evaluation processes, we examined the relationship between a recruitment index and three key environmental variables known to shape krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
-Table \@ref(tab:env) summarizes the main characteristics of the environmental datasets used in this analysis.
-These variables were obtained from a combination of satellite and reanalysis products, specifically *Nimbus-7* [doi.org/10.5067/QOZIVYV3V9JP](https://doi.org/10.5067/QOZIVYV3V9JP), *ERA5* [doi.org/10.48670/moi-00021](https://doi.org/10.48670/moi-00021), and the *E.U.
-Copernicus Marine Service* [doi.org/10.48670/moi-00148](https://doi.org/10.48670/moi-00148).
-The table details their spatial and temporal resolutions, as well as the time ranges covered by each dataset.
-Post-processing techniques, including data cleaning and handling algorithms, manipulation of spatial data, rasterization and projection transformations were facilitated by the *sf* and *ncdf4* packages in R [@Pebezma2018; @Pierce2021].
-
-```{r env, eval=TRUE}
-env_features <- data.frame(
-  Feature = c("Sea Ice", "SST", "Chlorophyll a"),
-  Platform = c("Nimbus-7", "ERA5", "E.U. Copernicus Marine Service"),
-  Spatial_Resolution = c("25 × 25 km", "0.25º × 0.25º", "0.25º × 0.25º"),
-  Temporal_Resolution = c("Daily", "Monthly", "Daily-Monthly"),
-  Time_Range = c("1978 - Present", "1991 - Present", "2000 - Present"),
-  stringsAsFactors = FALSE
-)
-
-#to word
-ft_env <- flextable(env_features) |>
-  set_header_labels(
-    Feature = "Feature",
-    Platform = "Platform",
-    Spatial_Resolution = "Spatial Resolution",
-    Temporal_Resolution = "Temporal Resolution",
-    Time_Range = "Time Range"
-  ) |>
-  set_caption("Summary of environmental variables used in correlation analysis") |>
-  color(part = "header", color = "black") |>
-  theme_booktabs() |>
-  fontsize(size = 10)
-
-ft_1 <- autofit(ft_env)
-ft_2 <- set_table_properties(ft_1, width = .5, layout = "autofit")
-ft_2
-```
-
-All environmental data were processed at their original spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], incorporating spatial manipulation, rasterization, and coordinate transformations.
-We modelled the relation the environmental influence on krill recruitment with a serie of linear models, progressively incorporating the environmental variables to identify the one with the greatest explanatory power.
-The initial model considered only the effect of MU (spatial), while subsequent models sequentially included Chla, SST, SIC and interaction terms.
-This modeling framework allowed us to determine the relative contribution of each factor and their interactions on the recruitment process.
-All models and their results can be found in the Supplementary Material 2.
-
-To identify environmental drivers suitable for explicit inclusion in the stock assessment model, we evaluated the relationship between a krill recruitment index and three environmental variables known to influence krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
-These variables represent thermal, habitat, and trophic controls acting on early life stages of Antarctic krill.
-
-Environmental data were obtained from a combination of satellite observations and reanalysis products (*Nimbus-7*, *ERA5*, and the EU Copernicus Marine Service).
-A summary of the spatial and temporal characteristics of each dataset is provided in Table @ref(tab:env).
-All variables were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], including data cleaning, spatial manipulation, rasterization, and coordinate transformation.
-
-Prior to integration into the stock assessment model, environmental variables were statistically screened using a set of linear regression models with recruitment as the response variable.
-The baseline model accounted for spatial structure through management unit (MU), while subsequent models progressively incorporated Chl-a, SST, SIC, and selected interaction terms.
-This stepwise approach was used to identify the environmental variable exhibiting the strongest and most consistent statistical relationship with recruitment.
-Results of these screening analyses are provided in Supplementary Material 2.
-
-### Environmental link formulation in Stock Synthesis
-
-Stock Synthesis version 3.30.21 allows environmental covariates to be incorporated directly into the population dynamics through the *env-link* option in the control file, whereby external time series modify biological processes in a deterministic and multiplicative manner [@methot2020stock].
-
-In this study, the selected environmental variable was linked to recruitment via an env-link acting on the recruitment deviations. Expected recruitment in year (t) was modeled as:
+## Environmental link formulation
+
+Within the Stock Synthesis (SS3) framework, environmental covariates can be incorporated into the estimation of recruitment through link functions that modify the annual value of key model parameters as a function of an external environmental index.
+Following the approach described by @Caserta2025 and @methot2020stock, recruitment deviations are adjusted relative to a baseline level according to the magnitude of the environmental index in each year, allowing either a multiplicative or additive response structure.
+In this formulation, the strength and direction of the environmental influence are governed by an estimable link parameter, while the environmental covariate is standardized to ensure comparability across years.
+For years in which environmental information is unavailable, the covariate is assumed to take its mean value, effectively removing its influence on recruitment.
+This approach enables the explicit propagation of environmental variability into recruitment dynamics, allowing recruitment estimates to respond coherently to interannual environmental forcing within the integrated assessment model.
+Under a log-linear environmental link, the estimated coefficient ( \beta_E )can be interpreted as the partial derivative of log-recruitment with respect to the environmental index.
+On the log scale, this formulation is equivalent to:
 
 $$
-E(R_t) = f(SSB_t) \times \exp(\beta_E E_t) \times \exp\left(-\tfrac{1}{2}\sigma_R^2\right),
-$$
-
-where ( f(SSB_t) ) represents the stock–recruitment relationship, assumed to follow a Beverton–Holt formulation; ( E_t ) corresponds to the standardized environmental time series provided in the Stock Synthesis data file; ( \beta_E ) denotes the environment–recruitment link coefficient estimated within the control file; and ( \sigma_R^2 ) is the variance associated with recruitment deviations.
-
-On the log scale, this formulation is equivalent to:
-
-$$
-\log(R_t) = \log(f(SSB_t)) + \beta_E E_t + \eta_t,
+\log(R_t) = \log(R_{base}) + \beta_E E_t + \eta_t,
 \quad \eta_t \sim \mathcal{N}(0,\sigma_R^2).
@@
 $$
-
 Positive values of (\beta_E) indicate environmental conditions that enhance recruitment, while negative values imply adverse effects.
 Because the env-link operates multiplicatively, environmental effects scale recruitment exponentially rather than additively.
-
-By incorporating environmental forcing through the env-link, part of the interannual variability in recruitment that would otherwise be attributed to unexplained recruitment deviations is instead explained by the environmental covariate.
+By incorporating environmental forcing through the env-link, part of the interannual variability in krill recruitment that would otherwise be attributed to unexplained recruitment deviations is instead explained by the environmental covariate.
 Consequently, improved model fit is expected to manifest as a reduction in the estimated recruitment deviation variance (\sigma_R^2), although this reduction is not imposed a priori and arises through likelihood optimization.
-
 The inclusion of the selected environmental variable via the env-link framework provides a biologically interpretable and statistically coherent mechanism to integrate ecosystem drivers into the core population dynamics of the krill stock assessment.
+
+To identify the most influential environmental drivers of krill recruitment for use in stock assessment processes, we examined the relationship between a recruitment index and three key environmental variables known to shape krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020].
+Table @ref(tab:env) summarizes the main characteristics of the environmental datasets used in this analysis.
+These variables were obtained from a combination of satellite and reanalysis products, including *Nimbus-7* [doi.org/10.5067/QOZIVYV3V9JP], *ERA5* [doi.org/10.48670/moi-00021], and the *EU Copernicus Marine Service* [doi.org/10.48670/moi-00148].
+The influence of environmental variability on krill recruitment was evaluated using a series of linear models, progressively incorporating environmental covariates to identify those with the greatest explanatory power.
+The initial model included only spatial structure through management units (MUs), while subsequent models sequentially added Chl-a, SST, SIC, and interaction terms.
+Results from correlation analyses and linear models (Supporting Information 2) consistently indicated that chlorophyll-a exhibited the strongest relationship with recruitment, outperforming SST and SIC across model configurations.
+Based on these results, chlorophyll-a was selected as the environmental covariate incorporated into the SS3.
 
@@
 
-## Model Input Data: Fishery and Survey
+## Data sources
 
@@
 
-# # to word
-# ft_data <- flextable(data) %>%
-#   set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
-#   theme_vanilla() %>%
-#   fontsize(size = 9, part = "all") %>%
-#   font(fontname = "Arial", part = "all") %>%
-#   autofit() %>% 
-#   set_table_properties(width = .5, layout = "autofit")
-# 
-# ft_data
-
-# to PDF
-
-kbl(data, format = "latex", 
-    booktabs = TRUE, 
-    caption = "Data sources used in the krill population dynamics modeling") %>%
-  kable_styling(latex_options = c("hold_position", "striped"), 
-                font_size = 9, 
-                full_width = FALSE) %>%
-  column_spec(1, width = "3cm") %>%
-  column_spec(2, width = "5cm") %>%
-  column_spec(3, width = "3cm") %>%
-  column_spec(4, width = "3cm")
-
+# to word
+ft_data <- flextable(data) %>%
+  set_caption("Table: Data sources used in the krill population dynamics modeling") %>%
+  theme_vanilla() %>%
+  fontsize(size = 9, part = "all") %>%
+  font(fontname = "Arial", part = "all") %>%
+  autofit() %>%
+  set_table_properties(width = .5, layout = "autofit")
+
+ft_data
 ```

### 2025-12-19 index.Rmd
edit env link

@@
 output:
-  bookdown::html_document2:
+  bookdown::pdf_document2:
     keep_md: true
@@
 
-The following table summarizes the key parameters to conditioning the reference model, including biological, growth, and population dynamics factors.
-
-```{r}
-start1 <- SS_readstarter(file = file.path(dir1.4,
-                                          "starter.ss"),
-                              verbose = FALSE)
-dat1 <- SS_readdat(file = file.path(dir1.4, start1$datfile),
-                        verbose = FALSE)
-ctl1 <-  r4ss::SS_readctl(file = file.path(dir1.4,
-                                    start1$ctlfil),
-                        verbose = FALSE,
-                        use_datlist = TRUE, 
-                   datlist = dat1)
-fore1 <- r4ss::SS_readforecast(file = file.path(dir1.4, 
-                                                "forecast.ss"),
-                              verbose = FALSE)
-# can also read in wtatage.ss for an empirical wt at age model using
-# r4ss::SS_readwtatage()
-```
-
-```{r}
-parbio<-ctl1$MG_parms[1:10,c(1:3,7)]
- row.names( parbio)<-c("Nat M",
-                       "Lmin", 
-                       "Lmax",
-                       "VonBert K",
-                       "CV young",
-                       "CV old", 
-                       "Wt a", 
-                       "Wt b",
-                       "L50%", 
-                       "Mat slope")
-
- SRpar<-ctl1$SR_parms[1:5,c(1:3,7)]
- Qpar<-ctl1$Q_parms[1:2,c(1:3,7)]
- Selpar<-ctl1$size_selex_parms[1:22,c(1:3,7)]
- parInit<-as.data.frame(rbind(parbio,SRpar,Qpar,Selpar))
-
-wb <- createWorkbook()
-
-addWorksheet(wb, "parameters")
-writeData(wb, "parameters", parInit)
-
-# Guardar el workbook
-#saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
-
-```
-
-```{r}
-parInit %>%
-  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
-  kbl(booktabs = TRUE,
-      format = "html",
-      caption = "Input parameters for the initial SS3 model of krill.
+The following table summarizes the key parameters to conditioning the reference model, including biological, growth, and population dynamics factors (\@ref(tab:parainit1)).
+
+```{r parainit1, eval=TRUE}
+dir1.1 <- here("s1.1")
+
+start1 <- SS_readstarter(file = file.path(dir1.1, "starter.ss"), verbose = FALSE)
+dat1   <- SS_readdat(file = file.path(dir1.1, start1$datfile), verbose = FALSE)
+ctl1   <- SS_readctl(file = file.path(dir1.1, start1$ctlfil),
+                     verbose = FALSE, use_datlist = TRUE, datlist = dat1)
+fore1  <- SS_readforecast(file = file.path(dir1.1, "forecast.ss"), verbose = FALSE)
+
+# Extraer parámetros
+parbio <- ctl1$MG_parms[1:10, c(1:4,7)]
+rownames(parbio) <- c("Nat M", "Lmin", "Lmax", "VonBert K",
+                      "CV young", "CV old", "Wt a", "Wt b", "L50%", "Mat slope")
+
+SRpar  <- ctl1$SR_parms[1:5, c(1:4,7)]
+Qpar   <- ctl1$Q_parms[1:2, c(1:4,7)]
+Selpar <- ctl1$size_selex_parms[1:20, c(1:4,7)]
+
+# Combinar parámetros
+parInit <- as.data.frame(rbind(parbio, SRpar, Qpar, Selpar))
+
+# Añadir nombres de fila como columna "Parameter"
+parInit$Parameter <- rownames(parInit)
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(INIT  = round(INIT, 2),
+         PRIOR = round(PRIOR, 2))
+
+# Crear vector de categorías (39 filas)
+categories <- c(rep("Natural Mortality", 1),
+                rep("Growth", 4),
+                rep("Length-Weight Relation", 2),
+                rep("Maturity", 2),
+                rep("Stock-Recruit Relation", 5),
+                rep("Catchability", 2),
+                rep("Selectivity", 21))  # OJO: corregido a 23
+
+# Añadir categorías y reordenar columnas
+parInit <- parInit %>%
+  mutate(Category = categories) %>%
+  dplyr::relocate(Category, Parameter)
+
+# Crear flextable
+ft <- flextable(parInit)
+ft <- set_caption(ft, "Input parameters for the initial SS3 model of krill.
 Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT).
-If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
-      kable_styling(latex_options = "scale_down",  # Expande la tabla al ancho completo
-                font_size = 8) %>%  # Aplica la fuente Times New Roman
-  pack_rows(index = c("Natural Mortality" = 1,
-                      "Growth" = 5,
-                      "relationship Length-Weigth" = 2,
-                      "Maturity" = 2,
-                      "S-R relation" = 5,
-                      "Catchability" = 2,
-                      "Selectivity" = 4))
-
-```
-
+If the phase (PHASE) for the parameter is negative, the parameter is fixed as input")
+ft <- autofit(ft)
+ft <- theme_booktabs(ft)
+ft <- set_table_properties(ft, width = .5, layout = "autofit")
+ft <- fontsize(ft, size = 9, part = "all")
+ft
+```
 ### Index
@@
 
-```{r dataserie, fig.width=9, fig.height=7}
+```{r dataserie, fig.width=9, fig.height=7, fig.cap="Data series used in krill modelling in 48.1 Subarea"}
 SSplotDataEnv(base.model1.4, 
